### 2021-01-14 Mest.tex
add files

\documentclass[11pt]{article} % use larger type; default would be 10pt

\usepackage[utf8]{inputenc} % set input encoding (not needed with XeLaTeX)
\usepackage{amsmath,amssymb}
\usepackage{bm}
\usepackage{natbib}

\newcommand{\EE}{\mathbb{E}}
\newcommand{\pp}{e(\bm{X})}
\newcommand{\ppi}{e(\bm{X_i})}
\newcommand{\hpp}{\hat{e}(\bm{X})}
\newcommand{\hppi}{\hat{e}(\bm{X_i})}

\newcommand\independent{\protect\mathpalette{\protect\independenT}{\perp}}
\def\independenT#1#2{\mathrel{\rlap{$#1#2$}\mkern2mu{#1#2}}}

\title{MOM PS Estimator under Strong Monotonicity}

\begin{document}
\maketitle

Say members of the control group of an RCT have no access to the treatment, but among the treatment group there are compliers and non-compliers---i.e. one-sided non-compliance.
(Of course, this could be, e.g., compliance in two different ways, etc.)
Then there are two principal strata $S$, never-takers and compliers. 
Say the exclusion restriction doesn't hold.

\citet{feller2017principal} and \citet{dingLu} and others discuss estimating principal effects in one-sided noncompliance (which they call ``strong monotonicity'') under the assumption of principal ignorability, viz. $Y_C\independent S |\bm{X}$.
This has two problems, to my mind:
\begin{enumerate}
\item Sometimes principal ignorability is not plausible
\item If you assume the principal ignorability for $Y_T$ (``strong'' principal ignorability) then, as \citet{feller2017principal} points out, the principal effects conditional on covariates are equal to the ITT conditional on covariates---the stratum plays no role. But sometimes the role of the strata is what we're trying to estimate!
\end{enumerate}

The approach here takes a different tack, based on a different assumption which is maybe even less plausible---but which, I think, can be relaxed.

\section{Setup}
Estimating $\EE[Y_T|S]$ is straightforward, so for the remainder 
just consider the ``control'' group, $Z=0$, with $n$ (control) subjects and drop the $C$ subscript. 
The control group is a mixture of compliers and never-takers.

For subject $i$, $i=1,\dots,n$, we have:
\begin{itemize}
\item $Y_i$ observed outcome
\item $S_i \in \{0,1\}$ unobserved stratum
\item $\ppi=Pr(S_i=1|\bm{X}_i)$ %take as given (like, estimated from
  % treatment group)
\item $\hppi$ is an estimate of $\ppi$ from a separate sample
  (treatment group?) with $\EE[\hppi|\ppi]=\ppi$
\end{itemize}

Let $\mu_0=\EE[Y|S=0]$ and $\mu_1=\EE[Y|S=1]$.
The goal is to estimate $\mu_0$ and $\mu_1$.

\textbf{Assumption:}\\
\begin{equation}\label{eq:assumption}
\EE[Y_i|\bm{X}_i,S_i]=\EE[Y_i|S_i]=\mu_0\text{ or }\mu_1
\end{equation}
i.e. $Y$ is mean-independent of $\bm{X}$ conditional on $S$. 
This is problematic---why should $S$ contain all information about $Y$?

Note, it is kinda related to ``principal ignorability`` of \citet{feller2017principal} and \citet{dingLu}, $Y\independent S |\bm{X}_i$ but, obviously, different too.

\section{M-Estimator}
As a preliminary, note that
\begin{align*}
  \EE[S|\hpp]&=\EE\left\{\EE[S|\pp,\hpp]|\hpp\right\}\\
             &=\EE\left\{\EE[S|\pp]|\hpp\right\}\\
             &=\EE[\pp|\hpp]=\hpp
\end{align*}

Then, note
\begin{align*}
  \EE[Y|\hpp]&=\EE\left\{\EE[Y|\hpp,S]|\hpp\right\}\\
             &=\EE\left\{\EE[Y|S]|\hpp\right\}\\
             &=\EE[\mu_1S+\mu_0(1-S)|\hpp]\\
             &=\mu_1\hpp+\mu_0(1-\hpp)
\end{align*}

Then we have
\begin{equation*}
  \EE[Y]=\EE\EE[Y|\hpp]=\mu_1\EE\hpp+\mu_0(1-\EE\hpp)
\end{equation*}

Next we have

\begin{align*}
  \EE[Y\hpp]&=\EE\left\{\EE[Y\hpp|\hpp]\right\}\\
            &=\EE\left\{\hpp\EE[Y|\hpp]\right\}\\
            &=\EE\left\{\hpp\left[\mu_1\hpp+\mu_0(1-\hpp)\right]\right\}\\
            &=\mu_1\EE[\hpp^2]+\mu_0\left(\EE[\hpp]-\EE[\hpp^2]\right)
\end{align*}

That gives us four parameters:
\begin{align*}
  \theta_1&=\EE[\hpp]\\
  \theta_2&=\EE[\hpp^2]\\
  \theta_3&=\mu_0\\
  \theta_4&=\mu_1
\end{align*}

and four estimating equations:
\begin{align*}
  \sum_i \hppi-\theta_1&=0\\
  \sum_i \hppi^2-\theta_2&=0\\
  \sum_i Y_i-\theta_4\theta_1-\theta_3(1-\theta_1)&=0\\
  \sum_i Y_i\hppi-\theta_4\theta_2-\theta_3(\theta_1-\theta_2)&=0
\end{align*}

\section{Relaxing \eqref{eq:assumption} with regression}
There's no particular reason to assume \eqref{eq:assumption} and it seems like it would typically be pretty implausible.

But say you believe the model
\begin{equation}\label{eq:regression}
Y_i=\bm{x}_i'\bm{\beta}+\gamma S_i+\epsilon_i
\end{equation}
 Then instead of \eqref{eq:assumption} we could assume something like
\begin{equation*}
\EE[Y_i-\bm{x}_i'\bm{\beta}|\bm{X}_i,S_i]=\EE[Y_i-\bm{x}_i'\bm{\beta}|S_i]
\end{equation*}

in other words, no interaction between $\bm{X}$ and $S$ in \eqref{eq:assumption}. 
Then, if you had an estimate for $\beta$, you could just substitute $Y_i-\bm{x}_i'\bm{\beta}$ for $Y_i$ in the estimates. 
Alternatively, you could use a stacked estimating equation approach and estimate it all together.

\bibliographystyle{unsrtnat}
\bibliography{MOM}

\end{document}

### 2021-01-14 MOM.r

@@ -1,0 +1,170 @@
+library(geex)
+library(pbapply)
+### try it out
+
+### under simple assumption
+n <- 1000
+
+mu0 <- 0
+mu1 <- 0.2
+
+ps <- runif(n)
+
+S <- rbinom(n,1,ps)
+
+Y <- ifelse(S==1,mu1,mu0)+rnorm(n)
+
+data <- data.frame(ps,Y)
+
+### parameter vector
+## mu0, mu1
+
+estFun1 <- function(data){
+    function(theta){
+        c(theta[2]*data$ps+theta[1]*(1-data$ps)-data$Y,
+          theta[2]*data$ps^2+theta[1]*(data$ps-data$ps^2)-data$Y*data$ps)
+    }
+}
+
+bbb <- create_basis(estFun,data)
+mycontrol <- new('geex_control', .root = setup_root_control(start = c(1, 1)))
+bbb@.control <- mycontrol
+
+estimate_GFUN_roots(bbb)
+
+mu0hat <- mean(ps)*mean(Y*ps)/((mean(Y)-1)*mean(ps^2)+mean(ps)^2)
+
+results <- m_estimate(estFun,data=data,    root_control = setup_root_control(start = c(1,1)))
+
+
+sim <- function(){
+    mu0 <- 0
+    mu1 <- 0.2
+
+    ps <- runif(n)
+
+    S <- rbinom(n,1,ps)
+
+    Y <- ifelse(S==1,mu1,mu0)+rnorm(n)
+
+    data <- data.frame(ps,Y)
+
+    results <- m_estimate(estFun,data=data,    root_control = setup_root_control(start = c(1,1)))
+
+    c(coef(results), sqrt(diag(vcov(results))))
+}
+
+
+res <- replicate(1000,sim())
+
+
+#### bigger
+
+n <- 200
+
+x1 <- rnorm(2*n)
+x2 <- rnorm(2*n)
+x3 <- rnorm(2*n)
+
+psTrue <- plogis(x1+x2+x3)
+
+S <- rbinom(2*n,1,psTrue)
+
+Z <- rep(c(1,0),n)
+
+mu10 <- 0
+mu00 <- 0
+mu11 <- .5
+mu01 <- 0.2
+
+Y <- 0.5*(x1+x2+x3)+ifelse(Z==1,ifelse(S==1,mu11,mu10),ifelse(S==1,mu01,mu00))+rnorm(2*n,0,0.01)
+
[…]

### 2021-01-14 mom2step.r

@@ -1,0 +1,53 @@
+twoStep <- function(data,justEst=TRUE){
+    mod1 <- lm(Y~x1+x2+S,data=data,subset=Z==1)
+    lmod <- glm(S~x1+x2,data=data,subset=Z==1,family=binomial)
+
+    dat1 <- data.frame(
+        r=data$Y[data$Z==0]-#coef(mod1)['x1']*data$x1[data$Z==0]-coef(mod1)['x2']*data$x2[data$Z==0],
+                                        predict(mod1,subset(data,Z==0)),
+        ps=predict(lmod,subset(data,Z==0),type='response')
+    )
+
+    estFun2 <- function(data){
+        function(theta){
+            c(
+                theta[1]-data$ps,
+                theta[2]-data$ps^2,
+                theta[4]*theta[1]+theta[3]*(1-theta[1])-data$r,
+                theta[4]*theta[2]+theta[3]*(theta[1]-theta[2])-data$r*data$ps
+            )
+        }
+    }
+
+    if(justEst){
+        bbb <- create_basis(estFun2,dat1)
+        mycontrol <- new('geex_control', .root = setup_root_control(start = rep(.5,4)))
+        bbb@.control <- mycontrol
+        theta <- estimate_GFUN_roots(bbb)$root
+        theta <- c(theta,
+                   -theta[3],#+mean(data$Y[data$Z==1&data$S==0]),
+                   coef(mod1)['S']-theta[4]
+                   )
+        names(theta) <- c('Epi','Epi2','mu0','mu1','eff0','eff1')
+
+        theta <- c(theta,attr(data,"trueEff"))
+
+        return(theta)
+    }
+
+    m_estimate(estFun2,data=dat1,root_control=setup_root_control(start = rep(.5,4)))
+}
+
+
+
+
+
+## #clusterExport(cl, c('twoStep','makeDat','sim2Step'))
+
+## res2000 <- pbreplicate(500,sim2Step(2000),cl=cl)
+## res500 <- pbreplicate(500,sim2Step(500),cl=cl)
+## resNoSep  <- pbreplicate(500,sim2Step(2000,mu00=0,mu01=0,mu10=0.5,mu11=0.5),cl=cl)
+## res0  <- pbreplicate(1000,sim2Step(500,mu00=0,mu01=0,mu10=0,mu11=0),cl=cl)
[…]

### 2021-01-14 momStan.r

@@ -1,0 +1,31 @@
+
+
+rstan_options("auto_write" = TRUE)
+
+
+mle <- function(data){
+    sdat <- with(data, list(
+                           nctl=sum(1-Z),
+                           ntrt=sum(Z),
+                           x1t=x1[Z==1],
+                           x1c=x1[Z==0],
+                           x2t=x2[Z==1],
+                           x2c=x2[Z==0],
+                           Ytrt=Y[Z==1],
+                           Yctl=Y[Z==0],
+                           St=S[Z==1]
+                       )
+                 )
+                                        #fit1 <-
+    fit <- optimizing(stanMod,data=sdat)
+#    fit2 <- optimizing(stanMod,data=sdat)
+
+ #   fit <- if(fit1$value>fit2$value) fit1 else fit2
+
+    c(eff0=fit$par['mu10']-fit$par['mu00'],
+      eff1=fit$par['mu11']-fit$par['mu01'])
+}
+
+stanMod <- stan_model('ps.stan',auto_write=TRUE)
+
+#fit <- optimizing(stanMod,data=sdat)

### 2021-01-14 simTot.r

@@ -1,0 +1,167 @@
+library(tidyverse)
+library(rstan)
+library(geex)
+library(parallel)
+library(pbapply)
+
+rstan_options("auto_write" = TRUE)
+
+
+mle <- function(data){
+    sdat <- with(data, list(
+                           nctl=sum(1-Z),
+                           ntrt=sum(Z),
+                           x1t=x1[Z==1],
+                           x1c=x1[Z==0],
+                           x2t=x2[Z==1],
+                           x2c=x2[Z==0],
+                           Ytrt=Y[Z==1],
+                           Yctl=Y[Z==0],
+                           St=S[Z==1]
+                       )
+                 )
+                                        #fit1 <-
+    fit <- optimizing(stanMod,data=sdat)
+#    fit2 <- optimizing(stanMod,data=sdat)
+
+ #   fit <- if(fit1$value>fit2$value) fit1 else fit2
+
+    c(eff0=fit$par['mu10']-fit$par['mu00'],
+      eff1=fit$par['mu11']-fit$par['mu01'])
+}
+
+stanMod <- stan_model('ps.stan',auto_write=TRUE)
+
+
+twoStep <- function(data,justEst=TRUE){
+    mod1 <- lm(Y~x1+x2+S,data=data,subset=Z==1)
+    lmod <- glm(S~x1+x2,data=data,subset=Z==1,family=binomial)
+
+    dat1 <- data.frame(
+        r=data$Y[data$Z==0]-#coef(mod1)['x1']*data$x1[data$Z==0]-coef(mod1)['x2']*data$x2[data$Z==0],
+                                        predict(mod1,subset(data,Z==0)),
+        ps=predict(lmod,subset(data,Z==0),type='response')
+    )
+
+    estFun2 <- function(data){
+        function(theta){
+            c(
+                theta[1]-data$ps,
+                theta[2]-data$ps^2,
+                theta[4]*theta[1]+theta[3]*(1-theta[1])-data$r,
+                theta[4]*theta[2]+theta[3]*(theta[1]-theta[2])-data$r*data$ps
+            )
+        }
+    }
+
+    if(justEst){
+        bbb <- create_basis(estFun2,dat1)
[…]

### 2021-04-30 simTot.r
try to incorporate X:Z interactions

@@ -94,14 +94,75 @@ twoStep <- function(data,justEst=TRUE){
 }
 
 
-oneStep <- function(dat){
+oneStep <- function(dat,int=FALSE){
+    estFun <- function(data,int){
+
+        function(theta){
+            ## ols z=1
+            xb1 <- with(data,(cbind(1,S,x1,x2)%*%theta[1:4])[,1])
+
+            ## xb for z=0
+            xb2 <- if(int){
+                       with(data,(cbind(x1,x2)%*%theta[12:13])[,1]) ##  (b differs btw trt grps)
+                   } else{
+                       with(data,(cbind(x1,x2)%*%theta[3:4])[,1]) ## (b is same in trt groups)
+                   }
+
+            ## logit z=1
+            xb3 <- with(data,(cbind(1,x1,x2)%*%theta[5:7])[,1])
+            ps <- plogis(xb3)
+
+            out <- c(
+### regression for treatment group
+                ifelse(data$Z==1,data$Y-xb1,0),
+                ifelse(data$Z==1,data$S*(data$Y-xb1),0),
+                ifelse(data$Z==1,data$x1*(data$Y-xb1),0),
+                ifelse(data$Z==1,data$x2*(data$Y-xb1),0),
+### logistic regression
+                ifelse(data$Z==1,data$S-ps,0),
+                ifelse(data$Z==1,data$x1*(data$S-ps),0),
+                ifelse(data$Z==1,data$x2*(data$S-ps),0),
+### mixture model in control group
+                ifelse(data$Z==0,theta[9]*ps+theta[8]*(1-ps)-data$Y+xb2,0),
+                ifelse(data$Z==0,theta[9]*ps^2+theta[8]*(ps-ps^2)-(data$Y-xb2)*ps,0),
+                theta[10]-(theta[1]-theta[8]),
+                theta[11]-(theta[1]+theta[2]-theta[9])
+            )
+
+            if(int)
+                out <- c(out,
+                         ifelse(data$Z==0,data$x1*(data$Y-xb2-theta[9]*ps-theta[8]*(1-ps)),0),
+                         ifelse(data$Z==0,data$x2*(data$Y-xb2-theta[9]*ps-theta[8]*(1-ps)),0)
+                         )
+            out
+        }
+    }
+
[…]

### 2021-04-30 Mest.tex
took out estimated principal scores from writeup--math was wrong

@@
 \maketitle
+
+[NOTE: I updated this taking **out** references to \emph{estimated}
+principal scores, cuz I realized the math was wrong. I'm following the
+propensity score literature and ignoring estimation error, for the moment]\\
 
@@
 \item $S_i \in \{0,1\}$ unobserved stratum
-\item $\ppi=Pr(S_i=1|\bm{X}_i)$ %take as given (like, estimated from
+\item $\ppi=Pr(S_i=1|\bm{X}_i)$ take as given (like, estimated from
   % treatment group)
-\item $\hppi$ is an estimate of $\ppi$ from a separate sample
-  (treatment group?) with $\EE[\hppi|\ppi]=\ppi$
+%\item $\hppi$ is an estimate of $\ppi$ from a separate sample
+%  (treatment group?) with $\EE[\hppi|\ppi]=\ppi$
 \end{itemize}
@@
 As a preliminary, note that
+% \begin{align*}
+%   \EE[S|\hpp]&=\EE\left\{\EE[S|\pp,\hpp]|\hpp\right\}\\
+%              &=\EE\left\{\EE[S|\pp]|\hpp\right\}\\
+%              &=\EE[\pp|\hpp]=\hpp
+% \end{align*}
+
+%Then, note
 \begin{align*}
-  \EE[S|\hpp]&=\EE\left\{\EE[S|\pp,\hpp]|\hpp\right\}\\
-             &=\EE\left\{\EE[S|\pp]|\hpp\right\}\\
-             &=\EE[\pp|\hpp]=\hpp
-\end{align*}
-
-Then, note
-\begin{align*}
-  \EE[Y|\hpp]&=\EE\left\{\EE[Y|\hpp,S]|\hpp\right\}\\
-             &=\EE\left\{\EE[Y|S]|\hpp\right\}\\
-             &=\EE[\mu_1S+\mu_0(1-S)|\hpp]\\
-             &=\mu_1\hpp+\mu_0(1-\hpp)
+  \EE[Y|\pp]&=\EE\left\{\EE[Y|\pp,S]|\pp\right\}\\
+             &=\EE\left\{\EE[Y|S]|\pp\right\}\text{ by \eqref{eq:assumption}}\\
+             &=\EE[\mu_1S+\mu_0(1-S)|\pp]\\
+             &=\mu_1\pp+\mu_0(1-\pp)
 \end{align*}
@@
 \begin{equation*}
-  \EE[Y]=\EE\EE[Y|\hpp]=\mu_1\EE\hpp+\mu_0(1-\EE\hpp)
+  \EE[Y]=\EE\EE[Y|\pp]=\mu_1\EE\pp+\mu_0(1-\EE\pp)
 \end{equation*}
@@
 \begin{align*}
-  \EE[Y\hpp]&=\EE\left\{\EE[Y\hpp|\hpp]\right\}\\
-            &=\EE\left\{\hpp\EE[Y|\hpp]\right\}\\
-            &=\EE\left\{\hpp\left[\mu_1\hpp+\mu_0(1-\hpp)\right]\right\}\\
-            &=\mu_1\EE[\hpp^2]+\mu_0\left(\EE[\hpp]-\EE[\hpp^2]\right)
+  \EE[Y\pp]&=\EE\left\{\EE[Y\pp|\pp]\right\}\\
+            &=\EE\left\{\pp\EE[Y|\pp]\right\}\\
+            &=\EE\left\{\pp\left[\mu_1\pp+\mu_0(1-\pp)\right]\right\}\\
+            &=\mu_1\EE[\pp^2]+\mu_0\left(\EE[\pp]-\EE[\pp^2]\right)
 \end{align*}
@@
 \begin{align*}
-  \theta_1&=\EE[\hpp]\\
-  \theta_2&=\EE[\hpp^2]\\
+  \theta_1&=\EE[\pp]\\
+  \theta_2&=\EE[\pp^2]\\
   \theta_3&=\mu_0\\
@@
 \begin{align*}
-  \sum_i \hppi-\theta_1&=0\\
-  \sum_i \hppi^2-\theta_2&=0\\
+  \sum_i \ppi-\theta_1&=0\\
+  \sum_i \ppi^2-\theta_2&=0\\
   \sum_i Y_i-\theta_4\theta_1-\theta_3(1-\theta_1)&=0\\
-  \sum_i Y_i\hppi-\theta_4\theta_2-\theta_3(\theta_1-\theta_2)&=0
+  \sum_i Y_i\ppi-\theta_4\theta_2-\theta_3(\theta_1-\theta_2)&=0
 \end{align*}
@@
 \begin{equation}\label{eq:regression}
-Y_i=\bm{x}_i'\bm{\beta}+\gamma S_i+\epsilon_i
+Y_i=\mu_1S_i+\mu_0(1-S_i)+\bm{x}_i'\bm{\beta}+\epsilon_i
 \end{equation}
@@
 \end{equation*}
+
+Under model \eqref{eq:regression}, we have new estimating equations.
+
+\begin{align*}
+  \EE[Y]&=\EE\left[\mu_1\pp+\mu_0(1-\pp)+\bm{X}\bm{\beta}\right]\\
+  \EE[Y\pp]&=\EE\left[\pp\left(\mu_1\pp+\mu_0(1-\pp)+\bm{X}\bm{\beta}\right)\right]\\
+  \EE[\bm{X}Y]&\EE\left[\bm{X}\left(\mu_1\pp+\mu_0(1-\pp)+\bm{X}\bm{\beta}\right)\right]\\
+\end{align*}
 

### 2021-07-22 writeUps/MOM.tex
reorganize

--- 2021-01-14 Mest.tex
@@
 \item $S_i \in \{0,1\}$ unobserved stratum
-\item $\ppi=Pr(S_i=1|\bm{X}_i)$ %take as given (like, estimated from
-  % treatment group)
-\item $\hppi$ is an estimate of $\ppi$ from a separate sample
-  (treatment group?) with $\EE[\hppi|\ppi]=\ppi$
+\item $\ppi=Pr(S_i=1|\bm{X}_i)$ take as given (like, estimated from treatment group)
 \end{itemize}
@@
 
-\section{M-Estimator}
-As a preliminary, note that
+\section{MOM Estimator}
+
+First, relating $\bar{Y}$ to $\mu_0$ and $\mu_1$.
+
 \begin{align*}
-  \EE[S|\hpp]&=\EE\left\{\EE[S|\pp,\hpp]|\hpp\right\}\\
-             &=\EE\left\{\EE[S|\pp]|\hpp\right\}\\
-             &=\EE[\pp|\hpp]=\hpp
+  \EE[Y_i|\bm{X}_i]&=\EE\left\{\EE[Y_i|\bm{X}_i,S_i]|\bm{X}_i\right\}\\
+  &=\EE\left\{\EE[Y_i|S_i]|\bm{X}_i\right\}\\
+  =\mu_1\ppi+\mu_0(1-\ppi)
+\end{equation*}
+Hence the first estimating equation is
+\begin{equation}\label{eq:est1}
+\bar{Y}=\hat{\mu}_1\overline{\pp}+\hat{\mu}_0(1-\overline{\pp})
+\end{equation}
+
+Second, 
+\begin{align*}
+\EE[Y_i\ppi]&=\EE\EE[Y_i\ppi|\ppi]\\
+&=\EE\left\{ \ppi \EE\EE[Y_i|\ppi,S_i]\right\}\\
+&=\EE\left\{ \ppi \EE\EE[Y_i|S_i]\right\}\text{ by \eqref{eq:assumption}}\\
+&=\EE\left\{\mu_1\ppi^2+\mu_0(1-\ppi)\ppi\right\}
 \end{align*}
 
-Then, note
-\begin{align*}
-  \EE[Y|\hpp]&=\EE\left\{\EE[Y|\hpp,S]|\hpp\right\}\\
-             &=\EE\left\{\EE[Y|S]|\hpp\right\}\\
-             &=\EE[\mu_1S+\mu_0(1-S)|\hpp]\\
-             &=\mu_1\hpp+\mu_0(1-\hpp)
-\end{align*}
-
-Then we have
+Then the 2nd estimating equation is
 \begin{equation*}
-  \EE[Y]=\EE\EE[Y|\hpp]=\mu_1\EE\hpp+\mu_0(1-\EE\hpp)
+\overline{Y\pp}=\hat{\mu}_1\overline{\pp^2}+\hat{\mu}_0(\overline{\pp}-\overline{\pp^2})
 \end{equation*}
 
-Next we have
-
-\begin{align*}
-  \EE[Y\hpp]&=\EE\left\{\EE[Y\hpp|\hpp]\right\}\\
-            &=\EE\left\{\hpp\EE[Y|\hpp]\right\}\\
-            &=\EE\left\{\hpp\left[\mu_1\hpp+\mu_0(1-\hpp)\right]\right\}\\
-            &=\mu_1\EE[\hpp^2]+\mu_0\left(\EE[\hpp]-\EE[\hpp^2]\right)
-\end{align*}
-
-That gives us four parameters:
-\begin{align*}
-  \theta_1&=\EE[\hpp]\\
-  \theta_2&=\EE[\hpp^2]\\
-  \theta_3&=\mu_0\\
-  \theta_4&=\mu_1
-\end{align*}
-
-and four estimating equations:
-\begin{align*}
-  \sum_i \hppi-\theta_1&=0\\
-  \sum_i \hppi^2-\theta_2&=0\\
-  \sum_i Y_i-\theta_4\theta_1-\theta_3(1-\theta_1)&=0\\
-  \sum_i Y_i\hppi-\theta_4\theta_2-\theta_3(\theta_1-\theta_2)&=0
-\end{align*}
+Solving gives the estimates:
+\begin{equation*}
+\hat{\mu}_0=\frac{\overline{\pp}\;\overline{Y\pp}}{(\bar{Y}-1)\overline{\pp^2}+\overline{\pp}^2}
+\end{equation*}
+and
+\begin{equation*}
+\hat{\mu}_1=\frac{\bar{Y}-(1-\overline{\pp})\hat{\mu}_0}{\overline{\pp}}
+\end{equation*}
 

### 2021-07-22 writeUps/Mest.tex
reorganize

--- 2021-04-30 Mest.tex
@@
 \usepackage{natbib}
+\usepackage{hyperref}
+\usepackage{graphicx}
 
@@
 \maketitle
-
-[NOTE: I updated this taking **out** references to \emph{estimated}
-principal scores, cuz I realized the math was wrong. I'm following the
-propensity score literature and ignoring estimation error, for the moment]\\
 
@@
 \item $S_i \in \{0,1\}$ unobserved stratum
-\item $\ppi=Pr(S_i=1|\bm{X}_i)$ take as given (like, estimated from
-  % treatment group)
+\item $\ppi=Pr(S_i=1|\bm{X}_i)$ take as given  
 %\item $\hppi$ is an estimate of $\ppi$ from a separate sample
@@
 Note, it is kinda related to ``principal ignorability`` of \citet{feller2017principal} and \citet{dingLu}, $Y\independent S |\bm{X}_i$ but, obviously, different too.
+
+The principal scores will typically
+  be estimated using data from the treatment group. Lemmas 1 \& 2 of
+  \citealt{fellerEtAl2016} imply that
+  $Pr(S=1|\bm{X},Z)=Pr(S=1|\bm{X})$, so that's cool.
+The math below ignores estimation error in principal scores, treating $\pp$ as
+known. However, as long as principal scores are estimated
+consistently, the asymptotics work out by Slutsky's theorem.
 
@@
 
+\section{Simulation}
+\subsection{Design}
+
+\subsubsection{Data Generation}
+
+Three indpendent standard normal covariates, $x_k$ $k=1,\dots,3$.
+Principal scores are set as:
+\begin{equation*}
+  \ppi=logit^{-1}\left[\beta(x_{1i}+x_{2i}+x_{3i})\right]
+\end{equation*}
+Principal stratum $S_i\in \{0,1\}$ generated as:
+\begin{equation*}
+S_i\sim Bern(\ppi)
+\end{equation*}
+where $\beta$ is a manipuated factor.
+When $\beta$ was higher, $S$ was more easily predicted by covariates
+$x_k$.
+
+Outcomes are generated as
+\begin{equation*}
+  Y_i=0.5(x_{1i}+x_{2i}+x_{3_i})+\mu_{01}S_i+\tau_iZ_i+\epsilon_i
+\end{equation*}
+where $Z_i$ is treatment assignment $Z_i=i\%2$, so that half of
+subjects are in the treatment group, $\tau_i$ is the treatment effect,
+\begin{equation*}
+  \tau_i=\begin{cases}
+    \mu_{10}-0&\text{ if }S_i=0\\
+    \mu_{11}-\mu_{01}&\text{ if }S_i=1
+  \end{cases}
+\end{equation*}
+and $\epsilon_i\sim \mathcal{N}(0,0.2)$ or $\epsilon_i$ has a Gumbel
+distribution with location parameter 0 and scale parameter 0.16
+(corresponding to a standard deviation of approximately 0.2).
+Errors $\epsilon$ and covariates $x_k$ are centered, so each has
+sample mean exactly 0.
+
+\subsubsection{Manipulated Factors}
+We manipulated four factors in the simulation:
+\begin{itemize}
+\item Sample size $n\in\{100,500,1000\}$
+\item $\mu_{01}\in\{0,0.3\}$
+\item $\mu_{10}\in\{0,0.3\}$
+\item $\beta \in \{0,0.2,0.5,1\}$
+\end{itemize}
+$\mu_{11}=0.3$ in all cases.
+
+When $\mu_{01}=0$ there is only one mixture component in the control
+group---i.e. no separation. When $\mu_{01}=\mu_{01}=0$ the treatment effect is
+0.3 for $S=1$ and 0 for $S=0$.  When $\mu_{01}=\mu_{10}=0.3$, the
+treatment effect is 0.3 for $S=0$ and 0 for $S=1$. When $\mu_{01}=0.3$
+and $\mu_{10}=0$, the treatment effect is 0 for both strata, and
+when $\mu_{01}=0$ and $\mu_{10}=0.3$ the effect is 0.3 for both
+strata.
+
+\subsubsection{Analysis Models}
+Analysis models had access to treatment assignment $Z$, outcomes $Y$,
+principal stratum $S$ for members of the treatment group, and
+covariates.
+Even though three covariates, $x_1$, $x_2$ and $x_3$ were included as
+predictors of principal stratum $S$ and outcome $Y$ in the data
+generating model, both analysis approaches only had access to the
+first two covariates, $x_1$ and $x_2$.
+Hence, principal ignorability was violated---$S$ and $Y$ were
+``confounted'' by unobserved variable $x_3$.
+
+Simulated data were analyzed first with a ``stacked equations''/M-Estimation/GEE
+approach, using the \texttt{geex} package in \texttt{R}
+\citep{geex,rcite}, and then with a more conventional maximum
+likelihood approach via \texttt{rstan} \citep{rstan}.
+
+Both methods assumed (correctly) that there were no interactions
+between covariates $x_1$ and $x_2$ and treatment assignment (after
+accounting for $S$).
+
+In the GEE approach included a set of nine estimating equations.
+First, a GEE logistic regression fit to the treatment group to
+estimate principal scores
+$\ppi=logit^{-1}(\beta_0+\beta_1x_{1i}+\beta_2x_{2i})$ (three parameters):
+\begin{align*}
+  \sum_iZ_i\left[S_i-\ppi\right]&=0\\
+  \sum_iZ_ix_{1i}\left[S_i-\ppi\right]&=0\\
+  \sum_iZ_ix_{2i}\left[S_i-\ppi\right]&=0
+\end{align*}
+Next, four equations to estimate the outcome regression in the
+treatment group, including terms for $S$:
+\begin{align*}
+  \sum_i
+  Z_i\left[Y_i-(\gamma_1x_{1i}+\gamma_2x_{2i}+S_i\mu_{11}+(1-S_i)\mu_{10})\right]&=0\\
+  \sum_i
+  Z_ix_{1i}\left[Y_i-(\gamma_1x_{1i}+\gamma_2x_{2i}+S_i\mu_{11}+(1-S_i)\mu_{10})\right]&=0\\
+  \sum_i
+  Z_ix_{2i}\left[Y_i-(\gamma_1x_{1i}+\gamma_2x_{2i}+S_i\mu_{11}+(1-S_i)\mu_{10})\right]&=0\\
+  \sum_i
+  Z_iS_i\left[Y_i-(\gamma_1x_{1i}+\gamma_2x_{2i}+S_i\mu_{11}+(1-S_i)\mu_{10})\right]&=0
+\end{align*}
+
+Finally, two equations to estimate the mixture model in the control
+group, for whom $S$ is unobserved:
+\begin{align*}
+  \sum_i
+  (1-Z_i)\left[Y_i-(\gamma_1x_{1i}+\gamma_2x_{2i}+\ppi\mu_{01}+(1-\ppi)\mu_{00})\right]&=0\\
+  \sum_i
+  (1-Z_i)\left[Y_i-(\gamma_1x_{1i}+\gamma_2x_{2i}+\ppi^2\mu_{01}+(\ppi-\ppi^2)\mu_{00})\right]&=0
+\end{align*}                                                                            
+The average treatment effect for $S=0$ was $\mu_{10}-\mu_{00}$ and for
+$S=1$ was $\mu_{11}-\mu_{01}$.
+
+Next, we estimated effects with MLE, using the built-in optimizer of
+\texttt{rstan}.
+
+In this model we regressed $S$ and $Y$ for the treatment group on an
+intercept, $x_1$, and $x_2$ using a logistic and normal linear
+specification, respectively.
+Then, we specified a two-component mixture model for control
+outcomes.
+
+Code for both models and for the data generating process can be found
+at \url{https://github.com/adamSales/psGee}.
+
+\subsection{Results}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=0.95\textwidth]{biasBetaN.jpg}
+  \caption{Bias as a function of $\beta$, $n$, and analysis model,
+    pooling over $\mu_{10}$ and $\mu_{01}$.}
+  \label{fig:bias}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=0.95\textwidth]{rmseBetaN.jpg}
+  \caption{Root mean sequared error (RMSE) as a function of $\beta$, $n$, and analysis model,
+    pooling over $\mu_{10}$ and $\mu_{01}$.}
+  \label{fig:rmse}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=0.95\textwidth]{coverBetaN.jpg}
+  \caption{Coverage of nominal 95\% confidence intervals as a function of $\beta$, $n$, and analysis model,
+    pooling over $\mu_{10}$ and $\mu_{01}$.}
+  \label{fig:coverage}
+\end{figure}
+
 \bibliographystyle{unsrtnat}

### 2021-07-22 code/fastSimInt.r

@@ -1,0 +1,39 @@
+
+
+sim1 <- function(n){
+
+    x1 <- rnorm(2*n)
+    x2 <- rnorm(2*n)
+    x3 <- rnorm(2*n)
+
+    x1 <- x1-mean(x1)
+    x2 <- x2-mean(x2)
+    x3 <- x3-mean(x3)
+
+    psTrue <- plogis(x1+x2+x3)
+
+    S <- rbinom(2*n,1,psTrue)
+
+    Z <- rep(c(1,0),n)
+
+
+
+    psDat <- data.frame(x1,x2,S,Z)
+    psMod <- glm(S~x1+x2,data=psDat,subset=Z==1,family=binomial)
+    ps <- predict(psMod,subset(psDat,Z==0),type='response')
+
+    x1 <- x1[Z==0]
+    x2 <- x2[Z==0]
+
+    Y <- 0.5*(x1+x2+x3[Z==0])+.2*S[Z==0]+rnorm(n,0,.2)
+
+    a <- rbind(
+        c(sum(ps), sum(1-ps), sum(x1), sum(x2)),
+        c(sum(ps^2), sum(ps*(1-ps)), sum(ps*x1), sum(ps*x2)),
+        c(sum(x1*ps), sum(x1*(1-ps)), sum(x1*x1), sum(x1*x2)),
+        c(sum(x2*ps), sum(x2*(1-ps)), sum(x2*x1), sum(x2*x2)))
+
+    b <- c(sum(Y),sum(ps*Y),sum(x1*Y),sum(x2*Y))
+
+    solve(a,b)
+}

### 2021-07-22 code/readSim.r

@@ -1,0 +1,395 @@
+library(tidyverse)
+library(rstan)
+library(geex)
+library(parallel)
+library(pbapply)
+
+bias <- function(sss){
+
+    sss1 <- sss[,1:6]
+    sss2 <- sss[,7:11]
+
+    colMeans(cbind(sss1[,3:6]-sss1[,c(1,2,1,2)],sss2))
+}
+
+rmse <- function(sss){
+
+    sss1 <- sss[,1:6]
+    sss2 <- sss[,7:11]
+
+    c(
+        sqrt(colMeans((sss1[,3:6]-sss1[,c(1,2,1,2)])^2)),
+        colMeans(sss2)
+    )
+}
+
+
+summ <- function(sss){
+    sss1 <- sss[,1:6]
+    sss2 <- sss[,7:11]
+    rbind(colMeans(sss1),
+          apply(sss1,2,sd),
+          sqrt(row1Means((sss-sss[rep(1:2,3),])^2))
+          )
+    }
+
+
+plotRes <- function(rrr,fac,S){
+    www <- (max(rrr[fac,])-min(rrr[fac,]))*0.02
+    plot(rrr[fac,]-www,rrr[paste0('mom',S),],ylim=range(c(rrr[paste0('mom',S),],rrr[paste0('mle',S),])),xlim=c(min(rrr[fac,])-2*www,max(rrr[fac,])+2*www),pch=16,xlab=fac)
+    points(rrr[fac,]+www,rrr[paste0('mle',S),],col='red',pch=16)
+    legend('topright',legend=c('MOM','MLE'),col=c('black','red'),pch=16)
+    abline(h=0,lty=2)
+}
+
+
+getSumms <- function(res,S,EST){
+    #cases <- results%>%select(starts_with('mu'),n,b1,gumb)%>%distinct()
+    true <- res[[paste0('true',S)]]
+    est <- res[[paste0(EST,S)]]
+    se <- res[[paste0(EST,S,'se')]]
+
+    cbind(
+        res[1,which(sapply(res,n_distinct)==1)],
+        data.frame(
+            EST=EST,
+            S=S,
+            getEst=mean(is.finite(est)),
+            getSE=mean(is.finite(se)),
+            bias=mean(est-true),
+            rmse=sqrt(mean((est-true)^2)),
+            biasV=mean(se^2)-var(est-true),
+            avgSE=mean(se),
[…]

### 2021-07-22 code/sim.r

@@ -1,0 +1,100 @@
+library(tidyverse)
+library(rstan)
+library(geex)
+library(parallel)
+library(pbapply)
+
+
+source('mom2step.r')
+source('momStan.r')
+
+
+
+makeDat <- function(n,mu00=0,mu01=0.2,mu10=0,mu11=0.5,b1=1,gumb=FALSE){
+
+    x1 <- rnorm(2*n)
+    x2 <- rnorm(2*n)
+    x3 <- rnorm(2*n)
+
+    x1 <- x1-mean(x1)
+    x2 <- x2-mean(x2)
+    x3 <- x3-mean(x3)
+
+    psTrue <- plogis(b1*(x1+x2+x3))
+
+    S <- rbinom(2*n,1,psTrue)
+
+    Z <- rep(c(1,0),n)
+
+    error <- if(gumb) evd::rgumbel(2*n,loc=0,scale=0.16) else rnorm(2*n,0,0.2)
+    error <- error-mean(error)
+
+    Yc <- 0.5*(x1+x2+x3)+mu01*S
+    Y <- Yc+(mu11-mu01)*S*Z+error
+
+
+    dat <- data.frame(Y,Z,S=ifelse(Z==1,S,0),x1,x2,Strue=S)
+    attr(dat,'trueEffs') <- c(
+        S0=mean(Y[Z==1&S==0])-mean(Y[Z==0&S==0]),
+        S1=mean(Y[Z==1&S==1])-mean(Y[Z==0&S==1])
+    )
+    dat
+
+}
+
+sim2Step <- function(n,...){
+    dat <- makeDat(n,...)
+    twoStep(dat)
+}
+
+simStan2step <- function(n,...){
+    dat <- makeDat(n,...)
+    two <- twoStep(dat)
+    MLE <- mle(dat)
+
+    out <- setNames(c(two[c('S0','S1','eff0','eff1')],MLE),
+             c('true0','true1','mom0','mom1','mle0','mle1')
+             )
+    #if(any(abs(out)>2)) return(list(out,dat))
+    out
+}
+
+
+summ <- function(sss){
+    rbind(rowMeans(sss),
+          apply(sss,1,sd),
+          sqrt(rowMeans((sss-sss[rep(1:2,3),])^2))
+          )
+    }
+
+
+fullsim <- function(nsim,
+                    ns=c(100,500,1000),
+                    mu01=c(0,.3),#sepTs=c(TRUE,FALSE),
+                    mu10=c(0,.3),#sepCs=c(TRUE,FALSE),
[…]

### 2021-07-22 code/simTot.r

@@ -1,0 +1,336 @@
+#library(dplyr)
+#library(rstan)
+print('here')
+library(geex)
+library(parallel)
+#library(pbapply)
+
+load('stanMod.RData')
+
+print('here')
+
+
+#rstan_options("auto_write" = TRUE)
+#pboptions(type='none')
+
+mle <- function(data,se=TRUE){
+    sdat <- with(data, list(
+                           nctl=sum(1-Z),
+                           ntrt=sum(Z),
+                           x1t=x1[Z==1],
+                           x1c=x1[Z==0],
+                           x2t=x2[Z==1],
+                           x2c=x2[Z==0],
+                           Ytrt=Y[Z==1],
+                           Yctl=Y[Z==0],
+                           St=S[Z==1]
+                       )
+                 )
+                                        #fit1 <-
+    fit <- rstan::optimizing(stanMod,data=sdat,hessian=se)
+#    fit2 <- optimizing(stanMod,data=sdat)
+
+ #   fit <- if(fit1$value>fit2$value) fit1 else fit2
+
+    if(se){
+        invHess <- -solve(fit$hessian)
+
+        return(
+            c(eff0=fit$par['mu10']-fit$par['mu00'],
+              se0=sqrt(invHess['mu10','mu10']+invHess['mu00','mu00']-2*invHess['mu10','mu00']),
+              eff1=fit$par['mu11']-fit$par['mu01'],
+              se1=sqrt(invHess['mu11','mu11']+invHess['mu01','mu01']-2*invHess['mu11','mu01'])
+              )
+        )
+    }
+
+    c(eff0=fit$par['mu10']-fit$par['mu00'],
+      eff1=fit$par['mu11']-fit$par['mu01']
+      )
+}
+
+#stanMod <- rstan::stan_model('ps.stan')#,auto_write=TRUE)
+
+
+twoStep <- function(data,justEst=TRUE){
+    mod1 <- lm(Y~x1+x2+S,data=data,subset=Z==1)
+    lmod <- glm(S~x1+x2,data=data,subset=Z==1,family=binomial)
+
+    dat1 <- data.frame(
+        r=data$Y[data$Z==0]-#coef(mod1)['x1']*data$x1[data$Z==0]-coef(mod1)['x2']*data$x2[data$Z==0],
[…]

### 2021-07-22 simTot.r
use ctl to estimate beta in simulation?

@@ -1,336 +1,0 @@
-#library(dplyr)
-#library(rstan)
-print('here')
-library(geex)
-library(parallel)
-#library(pbapply)
-
-load('stanMod.RData')
-
-print('here')
-
-
-#rstan_options("auto_write" = TRUE)
-#pboptions(type='none')
-
-mle <- function(data,se=TRUE){
-    sdat <- with(data, list(
-                           nctl=sum(1-Z),
-                           ntrt=sum(Z),
-                           x1t=x1[Z==1],
-                           x1c=x1[Z==0],
-                           x2t=x2[Z==1],
-                           x2c=x2[Z==0],
-                           Ytrt=Y[Z==1],
-                           Yctl=Y[Z==0],
-                           St=S[Z==1]
-                       )
-                 )
-                                        #fit1 <-
-    fit <- rstan::optimizing(stanMod,data=sdat,hessian=se)
-#    fit2 <- optimizing(stanMod,data=sdat)
-
- #   fit <- if(fit1$value>fit2$value) fit1 else fit2
-
-    if(se){
-        invHess <- -solve(fit$hessian)
-
-        return(
-            c(eff0=fit$par['mu10']-fit$par['mu00'],
-              se0=sqrt(invHess['mu10','mu10']+invHess['mu00','mu00']-2*invHess['mu10','mu00']),
-              eff1=fit$par['mu11']-fit$par['mu01'],
-              se1=sqrt(invHess['mu11','mu11']+invHess['mu01','mu01']-2*invHess['mu11','mu01'])
-              )
-        )
-    }
-
-    c(eff0=fit$par['mu10']-fit$par['mu00'],
-      eff1=fit$par['mu11']-fit$par['mu01']
-      )
-}
-
-#stanMod <- rstan::stan_model('ps.stan')#,auto_write=TRUE)
-
-
-twoStep <- function(data,justEst=TRUE){
-    mod1 <- lm(Y~x1+x2+S,data=data,subset=Z==1)
-    lmod <- glm(S~x1+x2,data=data,subset=Z==1,family=binomial)
-
-    dat1 <- data.frame(
-        r=data$Y[data$Z==0]-#coef(mod1)['x1']*data$x1[data$Z==0]-coef(mod1)['x2']*data$x2[data$Z==0],
[…]

### 2021-07-22 code/ct.r
estimate for CT

@@ -1,0 +1,70 @@
+library(tidyverse)
+library(rstan)
+library(geex)
+
+#load('../data/problemLevelUsageData/probLevelData.RData')
+load('../data/RANDstudyData/HSdata.RData')
+hs <- dat
+#print(load('../data/RANDstudyData/HSdata.RData'))
+#ms <- dat
+
+hs$xirt <- hs$xirt/ifelse(hs$year==1,sd(hs$xirt[hs$year==1]),sd(hs$xirt[hs$year==2]))
+
+load('../data/sectionLevelUsageData/advanceDataWInfo.RData')
+
+numProbs <- x%>%group_by(field_id,study.year,gradeLevel)%>%summarize(n=n(),nprob=n_distinct(Prob1))
+
+hist(numProbs$n,100)
+hist(numProbs$nprob,100)
+
+numProbs%>%ggplot(aes(nprob,y=..density..))+geom_histogram(bins=100)+facet_grid(gradeLevel~study.year)
+
+numProbs%>%
+    filter(study.year<3,n<1000)%>%
+    group_by(gradeLevel,study.year)%>%
+    mutate(N=n())%>%
+    group_by(gradeLevel,study.year,n)%>%
+    summarize(dens=n()/N)%>%
+    ggplot(aes(n,y=dens))+geom_point()+facet_grid(gradeLevel~study.year)
+
+numSec <- x%>%group_by(field_id,study.year,gradeLevel)%>%summarize(nsec=n_distinct(unit,section))
+
+hist(numSec$nsec,100)
+
+numSec%>%ggplot(aes(nsec,y=..density..))+geom_histogram(bins=100)+facet_grid(gradeLevel~study.year)
+
+numSec%>%
+    filter(study.year<3)%>%
+    group_by(gradeLevel,study.year)%>%
+    mutate(N=n())%>%
+    group_by(gradeLevel,study.year,nsec)%>%
+    summarize(dens=n()/N)%>%
+    ggplot(aes(nsec,y=dens))+geom_point()+facet_grid(gradeLevel~study.year)
+
+totalTime <-
+    x%>%filter(total_t1>0)%>%group_by(field_id,study.year,gradeLevel)%>%summarize(time=sum(total_t1,na.rm=TRUE))
+
+hist(log(totalTime$time),100)
+
+
+numCP <- advance%>%group_by(field_id,year)%>%summarize(ncp=sum(status=='changed placement',na.rm=TRUE))
[…]

### 2021-11-08 writeUps/regression.tex
putting some things down in writing

\documentclass[11pt]{article} % use larger type; default would be 10pt

\usepackage[utf8]{inputenc} % set input encoding (not needed with XeLaTeX)
\usepackage{amsmath,amssymb}
\usepackage{bm}
\usepackage{natbib}
\usepackage{hyperref}
\usepackage{graphicx}

\newcommand{\EE}{\mathbb{E}}
\newcommand{\pp}{e(\bm{X})}
\newcommand{\ppi}{e(\bm{X_i})}
\newcommand{\hpp}{\hat{e}(\bm{X})}
\newcommand{\hppi}{\hat{e}(\bm{X_i})}

\newcommand\independent{\protect\mathpalette{\protect\independenT}{\perp}}
\def\independenT#1#2{\mathrel{\rlap{$#1#2$}\mkern2mu{#1#2}}}

\title{MOM Estimator: Regression Form}

\begin{document}
\maketitle

Let $p_i=Pr(S_i=1|\bm{x}_i)$.
Then the original estimating equations were:
\begin{align*}
  Z_i=0:&\\
  &Y_i-\mu_{00}(1-p_i)+\mu_{01}p_i=Y_i-\mu_{00}-(\mu_{01}-\mu_{00})p_i\\
  p_i&Y_i-\mu_{00}p_i-(\mu_{01}-\mu_{00})p_i^2\\
  Z_i=1:&\\
        &Y_i-\mu_{10}-S_i(\mu_{11}-\mu_{10})\\
  S_i&Y_i-S_i\mu_{10}-S_i(\mu_{11}-\mu_{10})=S_iY_i-S_i\mu_{11}\\
\end{align*}

Let
\begin{equation}
  r_i=
  \begin{cases}
    p_i&\text{ if }Z_i=0\\
    S_i&\text{ if }Z_i=1
  \end{cases}
\end{equation}

Note that since $r_i$ depends on $Z_i$, we can replace both $p$ and $S$ in the estimating equations with $r$, without changing them.

\begin{align*}
  Z_i=0:&\\
  &Y_i-\mu_{00}(1-r_i)+\mu_{01}r_i=Y_i-\mu_{00}-(\mu_{01}-\mu_{00})r_i\\
  r_i&Y_i-\mu_{00}r_i-(\mu_{01}-\mu_{00})r_i^2\\
  Z_i=1:&\\
        &Y_i-\mu_{10}-r_i(\mu_{11}-\mu_{10})\\
  r_i&Y_i-r_i\mu_{10}-r_i(\mu_{11}-\mu_{10})=r_iY_i-r_i\mu_{11}\\
\end{align*}

Then we can combine these into four equations that look more like OLS. The first:

\begin{align*}
  &Y_i-(1-Z_i)(\mu_{00}-(\mu_{01}-\mu_{00})r_i)-Z_i(\mu_{10}-r_i(\mu_{11}-\mu_{10}))\\
  =& Y_i-\mu_{00}-(\mu_{01}-\mu_{00})r_i-Z_i(\mu_{10}-\mu_{00})-r_iZ_i
\end{align*}
and the 2nd:
\begin{align*}
  &r_iY_i-(1-Z_i)\left[\mu_{00}r_i-(\mu_{01}-\mu_{00})r_i^2\right]-Z_ir_i\mu_{11}\\
  =&r_iY_i-\mu_{00}r_i-(\mu_{01}-\mu_{00})r_i^2-Z_ir_i\left[\mu_{11}-\mu_{00}-r_i(\mu_{01}-\mu_{00})\right]\\
\end{align*}

The 3rd and 4th are the same as above:
\begin{align*}
  Z_i&Y_i-Z_i\mu_{10}-Z_ir_i(\mu_{11}-\mu_{10})\text{ and }\\
  Z_ir_i&Y_i-Z_ir_i\mu_{11}\\
\end{align*}

These are the estimating equations for the OLS model:
\begin{equation}\label{eq:regression0}
  Y_i=\beta_0+\beta_1r_i+\beta_2Z_i+\beta_3Z_ir_i+\epsilon_i
\end{equation}
with
\begin{align*}
  \beta_0&=\mu_{00}\\
  \beta_1&=\mu_{01}-\mu_{00}\\
  \beta_2&=\mu_{10}-\mu_{00}\\
  \beta_3&=\mu_{11}-\mu_{10}-(\mu_{01}-\mu_{00})
\end{align*}

\section{Missing $S$ in Treatment Group}
When $S$ is missing at random (conditional on $\bm{x}$) for some observations in the treatment group,
and $\mathbb{E}[Y|Z=1,S,p]=\mathbb{E}[Y|Z=1,S]$,
a similar mixed model will work for estimating $\mu_{10}$ and $\mu_{11}$.
\begin{align*}
  Z_i=1\text{ and }S_i\text{ Missing}:&\\
  &Y_i-\mu_{10}(1-p_i)+\mu_{11}p_i=Y_i-\mu_{10}-(\mu_{11}-\mu_{10})p_i\\
  p_i&Y_i-\mu_{10}p_i-(\mu_{11}-\mu_{10})p_i^2\\
  Z_i=1\text{ and }S_i\text{ Observed}:&\\
        &Y_i-\mu_{10}-S_i(\mu_{11}-\mu_{10})\\
  S_i&Y_i-S_i\mu_{10}-S_i(\mu_{11}-\mu_{10})=S_iY_i-S_i\mu_{11}\\
    Z_i=0:&\\
  &Y_i-\mu_{00}(1-p_i)+\mu_{01}p_i=Y_i-\mu_{00}-(\mu_{01}-\mu_{00})p_i\\
  p_i&Y_i-\mu_{00}p_i-(\mu_{01}-\mu_{00})p_i^2\\
\end{align*}

This works because
\begin{equation*}
  \mathbb{E}[Y|p]=\mathbb{E}\left[\mathbb{E}[Y|p,S]|p\right]=\mathbb{E}\left[\mathbb{E}[Y|S]|p\right]=\mu_{10}(1-p)+\mu_{11}p
\end{equation*}

Then re-define

\begin{equation}
  r_i=
  \begin{cases}
    p_i&\text{ if }Z_i=0\text{ or }Z_i=1\text{ and }S_i\text{ missing}\\
    S_i&\text{ if }Z_i=1\text{ and }S_i\text{ observed}
  \end{cases}
\end{equation}
Then the rest of the logic holds, and the regression (\ref{eq:regression0}) can be used to estimate $\bm{\mu}$.

\section{Sandwich Standard Errors}

So we're estimating principal effects in two stages.
First, fit the model
\begin{equation}\label{eq:psMod}
  Pr(S_{Ti}=1|\bm{\tilde{x}}_i)=f(\bm{\alpha}'\bm{\tilde{x}}_i)
\end{equation}
for some inverse link function $f(\cdot)$, where $\bm{\tilde{x}}_i=[1,\bm{x}_i]$, using (observed) values from the treatment group, and estimating $\hat{\alpha}$.
Then let
\begin{equation}\label{eq:ps}
  \hat{p}_i=f(\bm{\hat{\alpha}}'\bm{\tilde{x}}_i)
\end{equation}
for all subjects in the experiment.

Finally, fit model
\begin{equation}\label{eq:regression}
  Y_i=\beta_0+\beta_1r_i+\beta_2Z_i+\beta_3Z_ir_i+\bm{\gamma}^t\bm{x}_i+\epsilon_i
\end{equation}
to estimate $\bm{\beta}$ and hence principal effects.
[Note that \eqref{eq:regression} is just \eqref{eq:regression0} with $\bm{\gamma}^t\bm{x}_i$ added on the end].
Let $\bm{\beta}=[\beta_0,\beta_1,\beta_2,\beta_3,\bm{\gamma}^t]^t$.

The standard error matrix for $\bm{\hat{\beta}}$ can be estimated as:
\begin{equation}\label{eq:sandwich}
  \widehat{var}(\bm{\hat{\beta}})=A^{-1}BA^{-t}
\end{equation}
Where
\begin{equation}\label{eq:Amat}
  A=\sum_i\frac{\partial}{\partial [\alpha,\beta]^t} \Psi(Y_i;\alpha,\beta,\bm{x}_i)
\end{equation}
and
\begin{equation}\label{eq:Bmat}
  B=\sum_i \Psi(Y_i;\alpha,\beta,\bm{x}_i)\Psi(Y_i;\alpha,\beta,\bm{x}_i)^t
\end{equation}
where $\Psi(\cdot)$ is the estimating equations.

Following \citet[][p. 373]{carroletal06}, we %separate the estimating equations $\Psi$ into $\phi$ and $\psi$, estimating equations for models \eqref{
can decompose the matrices into diagonal elements $A_{1,1}$ and $A_{2,2}$ and $B_{1,1}$ and $B_{2,2}$ that pertain to the parameter sets $\alpha$ and $\beta$ and the estimating equations for models \eqref{eq:psMod} and  \eqref{eq:regression}, respectively, and $A_{21}$, and $B_{12}=B_{21}^t$, which capture to the dependence of model \eqref{eq:regression} on the parameters $\alpha$ from \eqref{psMod} and the covariance between the estimating equations of the two models. The sub-matrix $A_{12}=0$, since \eqref{eq:psMod} does not depend on $\beta$.

The diagonal matrices $A_{1,1}$ and $A_{2,2}$ and $B_{1,1}$ and $B_{2,2}$ are all the typical ``bread'' and ``meat'' matrices from M-estimation of generalized linear models and OLS. In practice, we use the estimates from the \texttt{sandwich} package in \texttt{R}, adjusted in two ways: first, the function \texttt{bread} actually gives $A^{-1}$, not $A$; second, we must pay careful attention to sample sizes, since the sample size for \eqref{eq:psMod} includes only treated observations (with observed $S$) and \eqref{eq:regression} contains all observations.

\subsection{$A_{21}$ Matrix}
The estimating equations for the regression \eqref{eq:regression} are
\begin{equation}\label{eq:eeOLS}
  \psi(Y_i,\bm{x}_i,\beta,\alpha)=X_iY_i-X_iX_i^t\beta
\end{equation}
Where $X_i=[1,r_i,Z_i,r_iZ_i,\bm{x}_i^t]^t$.
These depend on $\alpha$ when $r_i=p_i$, i.e. when $Z_i=0$ or $S_i$ is missing.
Then note that, following \eqref{eq:ps}, and letting $\eta_i=\alpha^t\bm{\tilde{x}}_i$
\begin{equation}\label{eq:derivP}
  \frac{\partial p_i}{\partial \alpha^t}=f'(\eta_i)\frac{\partial \eta_i}{\partial \alpha^t}=f'(\eta_i)\bm{\tilde{x}}_i^t
\end{equation}
and that
\begin{equation}
  \frac{\partial p_i^2}{\partial \alpha^t}=2p\frac{\partial p_i}{\partial \alpha^t}=2f(\eta)f'(\eta_i)\bm{\tilde{x}}_i^t
\end{equation}

Then if $r_i=p_i$,
\begin{align*}
  \frac{\partial}{\partial \alpha^t}&\psi(Y_i,\bm{x}_i,\beta,\alpha)=\\
  \left[\begin{array}{c}
          -\beta_1\frac{\partial p_i}{\partial \alpha^t}\\
          (Y_i-\beta_0-\beta_2Z_i-\bm{\gamma}^t\bm{x}_{i})\partial p/\partial\alpha^t-(\beta_1+\beta_3Z_i)\partial p^2/\partial \alpha^t\\
          -\beta_1\frac{\partial p_i}{\partial \alpha^t}Z_i\\
          Z_i\left[(Y_i-\beta_0-\beta_2Z_i-\bm{\gamma}^t\bm{x}_{i})\partial p/\partial\alpha^t-(\beta_1+\beta_3Z_i)\partial p^2/\partial \alpha^t\right]\\
          -\beta_1\frac{\partial p_i}{\partial \alpha^t}\bm{x}_i
        \end{array}\right]\\
  =& \left[\begin{array}{c}
          -\beta_1f'(\eta_i)\bm{\tilde{x}}_i^t\\
          \left[Y_i-X_i^t\beta-p_i(\beta_1+\beta_3Z_i)\right]f'(\eta_i)\bm{\tilde{x}}_i\\
          -Z_i\beta_1f'(\eta_i)\bm{\tilde{x}}_i^t\\
          Z_i\left[Y_i-X_i^t\beta-p_i(\beta_1+\beta_3)\right]f'(\eta_i)\bm{\tilde{x}}_i\\
          -\beta_1f'(\eta_i)\bm{x}_i\bm{x}_i^t
           \end{array}\right]
\end{align*}
Note that unless there is missing $S$ in the treatment group, $Z_i=0$ whenever $r_i=p_i$, so the 3rd and 4th lines of the matrix are all 0.

If $r_i=S_i$, $\frac{\partial}{\partial \alpha^t}&\psi(Y_i,\bm{x}_i,\beta,\alpha)=0$.

\end{document}

### 2021-11-08 code/regression.r
try dividing by sample sizes

@@ -52,16 +52,16 @@ sandwichMats <- function(psMod,outMod,data,clust=NULL,int=any(grepl(":x",names(c
     efOut <- estfun(outMod)
 
     out <- list(
-        a11inv = bread(psMod)/sum(data$Z),
-        a22inv = bread(outMod)/nrow(data),
-        a21 = A21(psMod,outMod,data),
-        b11=(if(is.null(clust)) meat(psMod) else meatCL(psMod,cluster=clust[data$Z==1]))*nrow(model.frame(psMod)),
-        b22 = (if(is.null(clust)) meat(outMod) else meatCL(outMod,cluster=clust))*nrow(data)#,adjust=TRUE)
+        a11inv = bread(psMod),#/sum(data$Z),
+        a22inv = bread(outMod),#/nrow(data),
+        a21 = A21(psMod,outMod,data)/nrow(data),
+        b11=(if(is.null(clust)) meat(psMod) else meatCL(psMod,cluster=clust[data$Z==1])),#*nrow(model.frame(psMod)),
+        b22 = (if(is.null(clust)) meat(outMod) else meatCL(outMod,cluster=clust))#*nrow(data)#,adjust=TRUE)
     )
     out <- within(out,b12 <-
         if(int){
             matrix(0,nrow(b11),ncol(b22))
-        } else if(is.null(clust)) crossprod(efPS,efOut)
+        } else if(is.null(clust)) crossprod(efPS,efOut)/nrow(data)
         else crossprod(apply(efPS,2L,rowsum,clust),
                        apply(efOut,2L,rowsum,clust))#/sum(data$Z==0)
         )

### 2021-11-09 writeUps/regression.tex
more writing up

@@
   \frac{\partial}{\partial \alpha^t}&\psi(Y_i,\bm{x}_i,\beta,\alpha)=\\
-  \left[\begin{array}{c}
+  &\left[\begin{array}{c}
           -\beta_1\frac{\partial p_i}{\partial \alpha^t}\\
@@
 
-If $r_i=S_i$, $\frac{\partial}{\partial \alpha^t}&\psi(Y_i,\bm{x}_i,\beta,\alpha)=0$.
+If $r_i=S_i$, $\frac{\partial}{\partial \alpha^t}\text{ \& }\psi(Y_i,\bm{x}_i,\beta,\alpha)=0$.
+
+\section{Sample Sizes}
+
+I am a bit confused over this one.
+In Sefanski and Boos, the sums that define $A$ and $B$ are divided by $n$, the sample size.
+In Carrol et al., the sums are not divided by anything, but the variance formula has an $n^{-1}$ in it, $A^{-1}BA^{-t}/n$.
+The dilemma is that the sample size for \eqref{eq:psMod} is $\approx n/2$ and the sample size for \eqref{eq:regression} is $n$.
+
+My intuition is based on the fact that $A$ is supposed to estimate $\mathbb{E}[-d\Psi/d\theta]$ and $B$ is supposed to estimate $\mathbb{E}[\Psi\Psi^T]$.
+Hence, $A_{11}$ and $B_{11}$ should be divided by the sample size for \eqref{eq:psMod}, and $A_{22}$ and $B_{22}$ should be divided by $n$.
+What about $A_{21}$ and $B_{21}$?
+Well, the covariance between the estimating equations of the two models should account for the fact that roughly half of the estimating equations for the outcome model's sample is uncorrelated with the estimating equations for the principal score model, and the other half is not dependent on its parameters---i.e., there are zeroes in the sum.
+Dividing by $n$ treats those zeros as informative, which seems correct to me.
+
+On the other hand, it turns out that if you don't divide anything by anything---i.e. there's no $n$ or $n/2$ in any denominators---you reproduce the standard errors from \texttt{geex} almost exactly.
+
+In any event, it doesn't seem to matter much in practice, and standard errors computed either way, dividing by sample sizes or not, are all almost unbiased, in the scenarios I've simulated.
+So rn I'm not sure.
 

### 2021-11-09 code/regression.r

@@ -96,7 +96,7 @@ vcvPS <- function(psMod,outMod,data,clust=NULL,int=any(grepl(":x",names(coef(out
 
     n <- nrow(model.frame(outMod))
 
-    vcvFull <- solve(A)%*%B%*%t(solve(A))
+    vcvFull <- solve(A)%*%B%*%t(solve(A))/nrow(data)
     main <- a22inv%*%b22%*%t(a22inv)
     vcv <- vcvFull[-(1:nrow(b11)),-(1:nrow(b11))]
     dimnames(vcv) <- dimnames(main)

### 2021-11-09 code/simFuncs.r

@@ -374,7 +374,7 @@ fullsimJustM <- function(nsim,
              1:nrow(cases),
              function(i){
                  cat(round(i/nrow(cases)*100))
-                 replicate(nsim,effs(do.call("makeData",cases[i,])))
+                 replicate(nsim,effs(do.call("makeDat",cases[i,])))
              }
          )
          )

### 2022-06-24 writeUps/paper.tex
update

\documentclass[11pt]{article} % use larger type; default would be 10pt

\usepackage[utf8]{inputenc} % set input encoding (not needed with XeLaTeX)
\usepackage{amsmath,amssymb}
\usepackage{bm}
\usepackage{natbib}
\usepackage{hyperref}
\usepackage{graphicx}
\usepackage{mathtools}
\usepackage{fullpage}
\usepackage{makecell}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{../../../generalCode/numdef}

\input{notation.tex}

\title{M-Estimation of Principal Effects using Principal Scores}

\begin{document}
\maketitle

\section{Introduction}

In a recent educational field experiment, middle school students completing their math schoolwork on a computer program were randomized between two conditions. In one condition, students were instructed to work on a fixed number of math problems, and later informed of their scores. In a second condition, students received immediate, informative error messages after entering a wrong answer, and were required to enter a correct answer in order to proceed. Students in the second treatment arm also had access to a series of hints for each problem, the last of which contained the correct answer. Students who ``bottomed out'' on a problem, i.e. requested all available hints on a problem, and were told the correct answer, did not receive credit for that problem. Students remained in their randomly-assigned condition over the course of a school year, and were assessed with a 10-item posttest administered within the computer program, as well with state standardized tests.

The literature on hints in online learning is mixed \citep{stuff}, in particular when it comes to bottoming out. On the one hand, hints that include the correct answer are essentially worked examples, which can be beneficial to learning \citep{worked example stuff}. On the other hand, they allow students to ``game the system,'' by simply requesting enough hints to see the right answer, and use the software without ever working to solve any problems. However, in this experiment, the ability to bottom out was only one of several differnces between the treatment conditions. One way to disentangle the role of bottoming out in any treatment effect would be to estimate average treatment effects separately for subjects who would (if assigned to treatment) bottom out, and for subjects who would not. If hints (that do not supply the correct answer) and error messages are helpful, but the ability to bottom out is harmful to learning, then we might expect to estimate a positive effect of being assigned to the feedback condition for students who do not bottom out, but a negative effect for students who do. If hints are helpful for learning only if they can be read as worked examples, then we might expect no effect of condition on students who do not bottom out, but a positive effect for students who do. (These hypothetical results may be masked by other treatment effect heterogeneity between students who would or wouldn't bottom out, but the results could still be informative.)

\section{Background}
In a randomized experiment with two conditions, let $Z_i=0$ if subject $i$, $i=1,\dots,n$ is assigned to the treatment condition, and let $Z_i=0$ otherwise.
If interest is in the effect of $Z$ on an outcome $Y$, under the ``stable unit treatment value assumption'' \citep{sutva} of no interference between units and no hidden versions of the treatment, define potential outcomes \citep{neyman,rubin} $\yti$ as the value of $Y$ that $i$ would exhibit if $Z_i=1$, and let $\yci$ be the value if $Z_i=0$. Then the observed value of $Y$ satisfies $Y_i=Z_i\yti+(1-Z_i)\yci$, and the treatment effect of $Z$ on $Y$ for stubject $i$ may be defined as $\tau_i=\yti-\yci$. The average treatment effect (ATE) is $\EE[\tau]$. %

Let $S_i\in\{0,1\}$ be a measurement on subject $i$ taken subsequent to treatment assignment, so that $S$ may be affected by $Z$.
Then $S$ itself has potential values $\sti$ and $\sci$, which $i$ would exibit if $Z_i=1$ or $Z_i=0$, respectively.
Assume ``strong monotonicity'' \citep[c.f.][]{dingLu} (also referred to as ``one-way noncompliance'' \citealt{imbensRubin}):
\begin{ass}[Strong Monotonicity]\label{ass:SM}
\begin{equation*}%\label{eq:strongMonotonicity}
  \sci=0\text{ for all }i
\end{equation*}
\end{ass}
Like potential outcomes $\yti$ and $\yci$, $\sci$ and $\sti$ are defined for all subjects, though $\sci$ is only observed when $Z_i=0$ and $\sti$ is only observed when $Z_i=1$. Under strong monotonicity, since $S$ is binary, every subject belongs to one of two ``principal strata,'' $\sti=1$ or $\sti=0$.
The average treatment effect in each principal stratum, $\eff1=\EE[\tau|\st=1]$ or $\eff1=\EE[\tau|\st=0]$, is called a principal effect. %
The challange in estimating principal effects is due to the fact that strata membership $\sti$ is only observed when $Z_i=1$, in which case $\sti=S_i$, the observed value of $S$.

\subsection{Estimating Principal Effects}
In estimating principal effects for a binary $S$, it will be useful to define four conditional means:
\begin{equation}\label{eq:mus}
  \begin{split}
    \muc0&=\EE[Y_C|\st=0]\\
    \muc1&=\EE[Y_C|\st=1]\\
    \mut0&=\EE[Y_T|\st=0]\\
    \mut1&=\EE[Y_T|\st=1]
  \end{split}
\end{equation}
Then, since
\begin{equation*}
  \begin{split}
    \eff0&=\mut0-\muc0\\
    \eff1&=\mut1-\muc1
  \end{split}
\end{equation*}
estimating principal effects is equivalent to estimating the four conditional means.

We will focus on estimation in completely randomized experiments, that is, we will assume
\begin{ass}[Randomization]\label{ass:rand}
\begin{equation*}%\label{eq:randomization}
  \yci,\yti,\sti, \bm{x}_i \independent Z_i
\end{equation*}
\end{ass}
where $\bm{x}_i$ is a vector of pre-treatment covariates.

Under Assumptions \ref{ass:sm} and \ref{ass:rand}, conditional means $\mut0$ and $\mut1$ are non-parametrically identified.
In particular, the means of observed outcomes for subjects with $Z=1$ and $S=0$ or $S=1$ are unbiased for $\mut0$ and $\mut1$, respectively, since when $Z_i=1$, $Y_i=\yti$ and $S_i=\sti$.
In contrast, $\muc0$ and $\muc1$ are not fully identified without further assumptions because $\sti$ and $\yci$ are never observed simultaneously (though they may be nonparametrically bounded; see, e.g., \citealt{bounding}).
When $S$ is a measure of compliance, and $\tau_i=0$ whenever $\sti=0$ (the ``exclusion restriction''), $\muc1$---and hence $\eff1$---may be estimated with instrumental variables methods \citep{angrist1996identification}.
In this case, $\eff1$ is often called the ``local average treatment effect'' or the average effect of the treatment on the treated.
We will briefly review two approaches to estimating $\muc0$ and $\muc1$ that do not assume the exclusion restriction---finite mixture modeling and weighting---with some comments on the role for covariates in-between.

\subsubsection{Finite Mixture Modeling}
The finite mixture modeling approch to estimating $\muc0$ and $\muc1$ \citep{imbens1997bayesian} assumes that, conditional on $\st$, $Y_C$ is normally distributed.
Then the probability density of $Y_C$ may be written as
\begin{equation}\label{eq:mixtureModel}
  f_{Y_C}(y)=Pr(\st=0)\phi\left(\frac{y-\muc0}{\sigma_C^0}\right)+Pr(\st=1)\phi\left(\frac{y-\muc1}{\sigma_C^1}\right)
\end{equation}
where $\phi(\cdot)$ is the standard normal density function, and $\sigma_C^0>0$ and $\sigma_C^1>0$ are standard deviations, which are sometimes assumed equal.
The probabilities $Pr(\st=0)$ and $Pr(\st=1)$ may be estimated first using data from the treatment group, since Assumption \ref{ass:rand} implies that $Pr(\st=1|Z=1)=Pr(\st=1)$.
Given these estimates and \eqref{eq:mixtureModel}, parameters $\muc0$ and $\muc1$ may be estimated using maximum likelihood, method of moments, or Bayesian techniques.

This approach has a number of setbacks.
First, the conditional normality assumption is untestable and limiting.
Unfortunately, even when it holds, mixture model estimates can be unreliable and, in some cases, severely biased \citep{griffin2008application,feller2017principal}.

\subsection{Principal Scores}
The principal score \citep[e.g.][]{jo2009use} is defined as
\begin{equation}\label{eq:pscore}
  \ppi=Pr(\sti=1|\bxsi)
\end{equation}
where $\bxsi\subseteq\bm{x}_i$, a not-necessarily-proper subset of covariates used to model $S$.
Note that under randomization, $Pr(\sti=1|\bxsi,Z_i)=Pr(\sti=1|\bxsi)$, so a model for principal scores can be estimated using data from the treatment group and extrapolated to the control group.
Assume a model for principal scores:
\begin{equation}\label{eq:prinScoreMod}
  \ppip{\bm{\alpha}}=f(\bxs;\bm{\alpha})
\end{equation}
with parameter vector $\bm{\alpha}$. An analyst may estimate $\bm{\alpha}$ by fitting model \eqref{eq:prinScoreMod} using observed $\bm{S}$ and $\bxs$ for subjects with $Z=1$, and then compute estimated principal scores $\ppip{\bm{\hat{\alpha}}}$ for subjects with $Z_i=1$.

Going forward, we will occasionally suppress dependence on $\bm{\alpha}$, and write $\ppip{\bm{\alpha}}=\ppi$.

In order for principal scores to be useful, they need to vary with $\bm{x}$. In particular, we will assume the following:
\begin{ass}[Variable Principal Scores]\label{ass:vps}
  \begin{equation*}
    Var_{\bx} (\pp)>0
  \end{equation*}
\end{ass}

Principal scores can potentially improve inference in a finite mixture model, by re-writing \eqref{eq:mixtureModel} as
\begin{equation}\label{eq:mixtureModelPS}
  f_{Y_C}(y)=(1-\ppip{\bm{\alpha}})\phi\left(\frac{y-\muc0}{\sigma_C^0}\right)+\ppip{\bm{\alpha}}\phi\left(\frac{y-\muc1}{\sigma_C^1}\right)
\end{equation}
Ideally, a statistician will estimate parameters $\muc0$, $\muc1$, $\mut0$, $\mut1$, $\sigma^0_C$, $\sigma^1_C$, and $\bm{\alpha}$ simultaneously, using maximum likelihood or Bayesian techniques.

Alternatively, assume ``principal ignorability'' \citep{jo2009use,dingLu}:
\begin{ass}[Principal Ignorability]\label{ass:PI}
\begin{equation}\label{eq:pi}
  Y_{C}\independent \st |\bxs
\end{equation}
\end{ass}
or a somewhat weaker version, $\EE[Y_C|\st,\bxs]=\EE[Y_C|\bxs]$ \citep{feller2017principal}.
That is, the principal stratum is unrelated to control potential outcomes conditional on covariates.
Principal ignorability \eqref{eq:pi} is reminiscent of ignorability assumptions typical in observational studies \citep[e.g.][]{rosenbaum2002observational}.
Both require a rich set of covariates and background knowledge of the determinants of $S$; in particular, an unobserved covariate $u$ that is correlated with both $\st$ and $Y_C$ would invalidate \eqref{eq:pi}.

Under principal ignorability, $\muc0$ and $\muc1$ can be estimated via weighting:
\begin{equation}\label{eq:psw}
  \hat{\muc0}_w=\frac{\sum_{i: Z_i=0} Y_i(1-\pp)}{\sum_{i:Z_i=0}1-\pp}\text{ and } \hat{\muc1}_w=\frac{\sum_{i: Z_i=0} Y_i(\pp)}{\sum_{i:Z_i=0}\pp}
\end{equation}

Principal score weighting does not require any distributional assumptions and (after estimating principal scores) very easy to implement.
On the other hand, the principal ignorability assumption is strong and restrictive.

\citet{feller2017principal} recommends the case-resampling bootstrap to estimate the sampling variance of $\hat{\muc0}_w$, $\hat{\muc1}_w$, and their associated principal effect estimators.

\section{M-Estimator}

The approach we introduce here incorporates principal scores into an M-Estimator for the mixture model \eqref{eq:mixtureModel}.
Like principal score weighting, our M-estimator does not require distributional assumptions, but instead requires a conditional independence assumption.
We will begin by describing a stronger-than-necessary conditional independence assumption in order to build intuition; in \S~\ref{sec:regression} we will present a considerably weaker alternative.

\subsection{Building Intuition: Covariate Ignorability}\label{sec:ci}

We begin by introducing an assumption that we term ``Covariate Ignorability,'' or CI.
\begin{ass}[Covariate Ignorability]\label{ass:ci}
\begin{equation}\label{eq:assumption}
\EE[\yci|\bxsi,\sti]=\EE[\yci|\sti]=\muc0\text{ or }\muc1
\end{equation}
\end{ass}
i.e. $Y_C$ is mean-independent of $\bxs$ conditional on $\st$.
Under CI, covariates are not informative of the mean of $Y_C$ within principal strata.
As stated above, CI will rarely be plausible.
However, in some circumstances researchers can identify a subset of observed covariates that satisfy CI, and use those covariates in estimation.

It turns out that under strong monotonicity, randomization, and CI, and given a set of principal scores, the two principal effects $\eff0$ and $\eff1$ can be estimted by a simple OLS regression.
To see how, we'll %first
derive the estimating equations for $\muc0$, $\muc1$, $\mut0$, and $\mut1$, and show that they are equivalent to estimating equations for a model fit by OLS.% assuming that principal scores are known exactly, and then generalize to the case in which they are estimated using data from the treatment group.
The details of the calculations and the proofs are in the appendix.

The argument stems from a set of expressions for expectations, summarized in the following lemma:
\begin{lemma}\label{lemma:expectations}
  Under Assumptions \ref{ass:sm}, \ref{ass:rand}, \ref{ass:vps}, and \ref{ass:ci}, %if principal scores $\ppi$, for $i:Z_i=0$ are known exactly, then:
  \begin{equation*}
    \begin{split}
      \EE[Y|Z=0]=&\muc0+\EE\pp(\muc1-\muc0)\\
      \EE[Y\pp|Z=0]&=\EE[\pp]\muc0+\EE[\pp^2](\muc1-\muc0)\\
      \EE[Y|Z=1]&=\mut0+\EE[\st](\mut1-\mut0)\\
      \EE[YS|Z=1]&=\EE[\st]\mut0+\EE[\st^2](\mut1-\mut0)
    \end{split}
  \end{equation*}
\end{lemma}
Since $\st=\st^2$, the expression for $\EE[\st Y_T]$ in Lemma \ref{lemma:expectation} can be simplified to $\EE[\st]\mut1$; however, we retain the unsimplified version, since its similarity with the analogous expression when $Z_i=0$ simplifies the reasoning below.

Lemma \ref{lemma:expectation} in turn implies a set of estimating equations $\sum_{i=1}^n\tilde{\Psi}_i=\bm{0}$, where
\begin{equation}\label{eq:estEq0}
\tilde{\Psi}_i=\begin{pmatrix}
    (1-Z_i)\left[Y_i-\muc0-\ppi(\muc1-\muc0)\right]\\
    (1-Z_i)\left[\ppi Y_i-\ppi\muc0-\ppi^2(\muc1-\muc0)\right]\\
    Z_i\left[Y_i-\mut0-\sti (\mut1-\mut0)\right]\\
    Z_i\left[\sti Y_i -\sti\mut0-\sti^2(\mut1-\mut0)\right]
  \end{pmatrix}
\end{equation}
Note that Lemma \ref{lemma:expectation} implies that $\EE[\Psi_i]=\bm{0}$.

Now, the estimating equations $\sum_i\tilde{\Psi}_i=\bm{0}$ are algebraically equivalent to the estimating equations for a particular OLS fit, after some transformations, the most important of which follows.
Let
\begin{equation}\label{eq:ri}
\ri=Z_iS_i+(1-Z_i)\ppi=\begin{cases}
\ppi &Z_i=0\\
\sti &Z_i=1
\end{cases}
\end{equation}
Note that, in general, $(1-Z_i)\ppi=(1-Z_i)\ri$ and $Z_i\sti=Z_i\ri$, allowing $\ri$ to take the place of $\ppi$ and $S_i$ througout $\tilde{\Psi}_i$.

\begin{prop}\label{prop:reg1}
  Under Assumptions \ref{ass:sm}, \ref{ass:randomization}, \ref{ass:vps}, and \ref{ass:ci}, if %principal scores are known exactly,
  $\ri$ is defined as in \eqref{eq:ri} and the model
\begin{equation}\label{eq:regression0}
  \EE[Y_i|Z_i,\ri]=\beta_0+\beta_1\ri+\beta_2Z_i+\beta_3Z_i\ri
\end{equation}
is fit with OLS, yielding coefficient estimates $\hat{\beta}_k$, $k=0,\dots,3$, then
\begin{equation}
  \begin{split}
    \hat{\eff0}&\equiv \hat{\beta}_2\\
    \hat{\eff1}&\equiv \hat{\beta}_3-\hat{\beta}_2
  \end{split}
\end{equation}
are M-estimators, with $\hat{\eff0}\rightarrow_p\eff0$ and $\hat{\eff1}\rightarrow_p\eff1$ as $n\rightarrow\infty$
\end{prop}

Thus, given principal scores, Assumption \ref{ass:ci} enables an analyst to estimate principal effects with a simple regression.
If principal scores were known exactly, the M-estimation structure of $\hat{\eff0}$ and $\hat{\eff1}$ implies that their sampling variance could be estimated by the standard sandwich estimator for OLS \citep{sefanskiBoosArticle}.
This will rarely be the case.
However, if the principal scores are modeled as \eqref{eq:prinScoreMod}, and parameters $\bm{\alpha}$ are estimated by M-estimation, with estimating equations $\sum_{i: Z_i=1}\Omega_i=\bm{0}$, then the estimating equations for the principal score model and principal effect estimation can be stacked \citep[c.f.][]{boosStefanskiBook}, as
\begin{equation}\label{eq:stacked}
  \sum_{i=1}^n \begin{pmatrix}
    Z_i\Omega_i(\bxsi,S_i;\bm{\alpha})\\
    \Psi_i\left(Y_i,Z_i,\ri(\bm{\alpha}); \bm{\beta}\right)\end{pmatrix}=\begin{pmatrix} \bm{0}\\\bm{0}\end{pmatrix}
\end{equation}
where $\ri\left(\bm{\alpha}\right)=S_iZ_i+\ppip{\bm{\alpha}}$ emphasizes $\ppi$'s dependence on $\bm{\alpha}$.
In practice, the principal score model \eqref{eq:prinScoreMod} could be fit first, yielding estimates $\bm{\hat{\alpha}}$, and model \eqref{eq:regression0} could be fit next, using $\ri=\ri\left(\bm{\hat{\alpha}}\right)$.
Together, the vector of estimates $[\bm{\hat{\alpha}},\bm{\hat{\beta}}]$ represents a zero of the stacked estimating equations \eqref{eq:stacked}.

Then, the standard error matrix for $[\bm{\hat{\alpha}},\bm{\hat{\beta}}$ can be estimated as \citep[][ch. 7]{boosStefanskiBook}:
\begin{equation}\label{eq:sandwich}
  \widehat{var}(\bm{\hat{\beta}})=A^{-1}BA'
\end{equation}
Where
\begin{equation}\label{eq:Amat}
  A=\sum_i\frac{\partial}{\partial [\alpha,\beta]'} [Z_i\Omega_i,\Psi_i]'
\end{equation}
and
\begin{equation}\label{eq:Bmat}
  B=\sum_i [Z_i\Omega_i, \Psi_i]'[Z_i\Omega_i, \Psi_i]
\end{equation}
where $[Z_i \Omega_i, \Psi_i]$ is the stacked vector of estimating equations evaluated at $Z_i$, $\bxsi$, $S_i$, and $Y_i$.
The appendix gives details of this calculation, giving a consistent variance estimator.

To summarize, under strong monotonicity, randomization, and covariate ignorability \eqref{eq:assumption}, principal effects can be estimated with the following procedure:
\begin{enumerate}
\item Estimate parameters $\bm{\alpha}$ as $\bm{\hat{\alpha}}$ by fitting model \eqref{eq:prinScoreMod} using data from subjects with $Z=1$.
\item Estimate principal scores for subjects with $Z_i=0$ as $\ppip{\bm{\hat{\alpha}}}$
\item Set $\ri=Z_iS_i+(1-Z_i)\ppip{\bm{\hat{\alpha}}}$
\item Using OLS, fit model \eqref{eq:regression0}
\item Estimate $\hat{\tau^0}=\hat{\beta}_2$ and $\hat{\tau^1}=\hat{\beta}_3-\hat{\beta}_2$
\end{enumerate}

\subsection{Relaxing Covariate Ignorability}
In most applications, identifying a set of covariates $\bxs$ satisfying CI will be difficult or impossible.
Fortunately, CI may be relaxed or obviated by regression.
Let $\bxy\in \mathbb{R}^k$, be a vector of covariates (possibly the same as $\bxs$); $\bxy$ can contain transformations and interactions of baseline covariates, allowing for a wide range of models.
Then note that $\st$ is, essentially, a covariate, so by randomization, (Assumption \ref{ass:rand}), $\EE[\bxy|\st]=\EE[\bxy|\st,Z=0]=\EE[\bxy|\st,Z=1]$.
Then information about the stratum-specific means of $\bxy$ can be extrapolated from the treatment to the control group.
This leads to the following proposition:
\begin{prop}\label{prop:interactions}
  Say
  \begin{equation}\label{eq:interaction}
    \begin{split}
      Y_{Ci}=\beta_0^C+\bm{\beta_1}^{C'}\bxy_i+\beta_2^C\sti+\bm{\beta_3^C}'\bxy_i\sti+\epsilon_{Ci}\\
      Y_{Ti}=\beta_0^T+\bm{\beta_1}^{T'}\bxy_i+\beta_2^T\sti+\bm{\beta_3^T}'\bxy_i\sti+\epsilon_{Ti}\\
    \end{split}
  %  \end{equation}
%      +\beta_3Z_i+\beta_4\sti Z_i+\bm{\beta_5}'\bxy_i\sti+
 %   \bm{\beta_6}'\bxy Z_i+\bm{\beta_7}'\bxy_i Z_i\sti+
  \end{equation}
  with $\EE[\epsilon_i]=0$ and $\{Z_i,\sti,\bxy_i\}\independent \epsilon_i$.
  Then say the following model is fit with OLS
  \begin{equation}\label{eq:fullInteractionModel}
    Y_i=\beta_0+\bm{\beta_1}'\bxy_i+\beta_2\ri+\beta_3Z_i+\beta_4\ri Z_i+\bm{\beta_5}'\bxy_i\ri+
    \bm{\beta_6}'\bxy Z_i+\bm{\beta_7}'\bxy_i Z_i\ri+\epsilon_i
  \end{equation}
  with $\ri$ as defined in \eqref{eq:ri},
  yielding coefficient estimates $\hat{\beta}_k$, $k=0,\dots,3$.
  Then
   \begin{equation}\label{eq:prinEffEst}
  \begin{split}
    \hat{\eff0}_{int}&\equiv \hat{\beta}_3+\bm{\beta}_6'\overline{\bxy}_{Z=1,S=0}\\
    \hat{\eff1}_{int}&\equiv \hat{\beta}_3+(\bm{\hat{\beta}_6}+\bm{\hat{\beta}_7})'\overline{\bxy}_{Z=1,S=1}
  \end{split}
\end{equation}
   are M-estimators, with $\hat{\eff0}_{int}\rightarrow_p\eff0$ and $\hat{\eff1}_{int}\rightarrow_p\eff1$ as $n\rightarrow\infty$,
   where $\overline{\bxy}_{Z=1,S=s}$ is a vector of the $k$ column means of the rows of $\bxy$ corresponding to subjects with $Z=1$ and $S=s$.
\end{prop}

Proposition \ref{prop:interactions} would seem to nearly solve the problem of estimating principal effects, or at least reducing it to the difficulty of other commen problems in causal modeling.
Unfortunately, simulation studies presented in the appendix suggest that, although consistent estimates for principal effects may be available, these estimates are far too inefficient for practical use; standard errors will be prohibitively high in all but the largest sample sizes.
For instance, if there is one standard normal covariate $x$, and $\pp_i=\left[1+exp\{-x_i/2\right]^{-1}$, then $Var\left(\hat{\muc0}_{int}\right)/Var(Y_C)\approx 8,500/n$.
In contrast, an analogous simulation in which Assumption \ref{ass:ci} holds, and the estimation routine of Proposition \ref{prop:reg1} was used, $Var\left(\hat{\muc0}_{int}\right)/Var(Y_C)\approx 15/n$.

Fortunately, simpler models can be fit much more efficiently. In particular, if interactions between covariates and treatment assignment are omitted, then data from the treatment group---in which $\st$ is fully observed---can be used to estimate the coefficients on covariates.
Such a model implies a fairly strong assumption, that the only interactions between covariates and treatment assignment are due to differences between principal effects.
However, there are two mitigating points.
First, since there is no requirement that all relevant covariates be included in the model, a researcher can exclude covariates from a model if he or she thinks their relationship with $Y$ may vary substantially between treatment groups, even in the same principal stratum.
Second, simulation results in the next section suggest that some models excluding important interactions may still yield reliable inferences.

As in the simpler case discussed in \S~\ref{sec:ci}, since the estimators in \eqref{eq:prinEffEst} are M-Estimators, their sampling variance can be estimated with a consistent sandwich estimator.
Uncertainty in principal score estimation can be incorporated by stacking the estimating equations for the principal score model \eqref{eq:prinScoreMod} and estimating the combined covariance matrix \eqref{eq:sandwich}..

To summarize, incorporating covariates into an outcome regression can substantially relax covariate ignorability, but estimation is only practical if  if the regression omits interactions between covariates and $Z$.
Standard errors can be estimated via stacked estimating equations.

\section{A Simulation Study}

We conducted a simulation study to compare the performance of our regression estimator compared to a Bayesian mixture model and to a principal score weighting estimator, in both favorable and unfavorable circumstances.
The study was designed to answer three sets of overarching questions.
First, how does the performance of the M-estimator vary as sample size and the extent to which $\bx$ predicts $\st$, and the principal effects themselves vary, and when important interactions were omitted from the outcome model.
Second, how does the M-estimator could compete with the mixture model? Is it competitive in circumstances favorable to both techniques? Does it avoid the pitfalls of mixture modeling when the parametric assumptions of the normal mixture model fail?
Third, are there ways or circumstances when the principal score weighting estimator outperforms the M-estimator even though principal ignorability does not hold?

\subsection{Simultation Design}
The simulation study was conducted in \texttt{R} \citep{rcite} and Stan \citep{rstan}, and full replication code is available at \url{www.github.com/[Redacted]}.
\subsubsection{Data Generation}

Across simulation runs, there were $n=500$ subjects in each treatment arm.

Given three indpendent standard normal covariates, $x_k$ $k=1,\dots,3$,
principal scores were simulated as
\begin{equation*}
  \ppi=logit^{-1}\left[\alpha(x_{1i}+x_{2i}+x_{3i})\right]
\end{equation*}
where $\alpha$ is a manipuated factor.
When $\alpha$ was higher, $\st$ was more easily predicted by covariates; when $\alpha=0$, Assumption \ref{ass:vps}, was violated, i.e. $Var(\ppi)=0$.

Then principal stratum $\sti\in \{0,1\}$ was simulated as:
\begin{equation*}
S_i\sim Bern(\ppi)
\end{equation*}

In most simulation runs, there were no interactions between
Control potential outcomes were generated as
\begin{equation}\label{eq:yc-sim}
  \yci=\beta_1(x_{1i}+x_{2i})+[\muc1+\beta_2(x_{1i}+x_{2i})]\sti +x_{3i}/\sqrt{6}+\epsilon_i
\end{equation}
where $\EE[\epsilon_i]=0$ $Var(\epsilon_i)=1/2$, and the distribution of $\epsilon_i$ and $\muc1$ were manipulated factors in the study.
In most simulations, there was no interaction between $\st$ and $x_1$ and $x_2$; in those cases, $\beta_1=1/\sqrt{6}$, and $\beta_2=0$.
In simulation runs including an interaction with $\st$, $\beta_1=3/(4\sqrt{6})$ and $\beta_2=5/(4\sqrt{6})$.

Finally, treatment potential outcomes were generated as
\begin{equation*}
  \yti=\yci+(0.3-\muc1)\sti+\beta_3(x_{1i}+x_{2i})
\end{equation*}
and observed outcomes were $Y_i=Z_i\yti+(1-Z_i)\yci$, with even subjects assigned to treatment and odd subjects assigned to control.
In most simulation runs, there were no interactions between treatment assignemnt and covariates, so $\beta_3=0$; when such interactions were included, $\beta_3=1/(2\sqrt{6})$.

Since the covariates were all centered, $\muc0=\mut0=\eff0=0$ throughout the study.
Similarly, $\mut1=0.3$, while $\muc1$, and hence $\eff1$, varied between runs.

Note that when there were no interactions included in the model, since $Var(\beta_1(x_1+x2+x_3))=1/2$ and $Var(\epsilon_i)=1/2$, covariates explained half of the variance of $Y$ within treatment group and principal stratum. However, as we shall see, $x_3$ was ``unobserved,'' so observed covariates explained 1/3 of that variance.

\subsubsection{Manipulated Factors}
We manipulated five factors in the simulation:
\begin{itemize}
%\item Sample size $n\in\{500,1000\}$
\item $\muc1\in\{0,0.3\}$
\item $\alpha \in \{0,0.2,0.5,1\}$
\item The distribution of $\epsilon_i$
\item The presence of interactions between $\st$ and covariates
\item The presence of interactions between $Z$ and covariates
\end{itemize}

When $\muc1=0$ there is only one mixture component in the control
group, which can be problematic for mixture modeling \citep{griffin2008application}. Conversely, when $\muc1=0.3$, $\eff0=\eff1=0$, so there is no treatment effect for any subjects.

As mentioned above, $\alpha$ controls the variance of the principal scores, i.e. the extent to which $\st$ is predictible as a function of covariates. When $\alpha=0$, the covariates are unrelated to $\st$ and principal scores do not vary between participants, violating \eqref{eq:vps}. We expect that the more predictive the covariates, the better the performance of all three estimators.

$\epsilon$ is the regression error in \eqref{eq:yc-sim}. Across simulation runs, $E[\epsilon]=0$ and $Var(\epsilon)=1/2$. However, the shape of $\epsilon$'s distribution varied between the following three possibilities:
\begin{itemize}
\item[Normal] $\epsilon~\mathcal{N}(0,1/2)$, as assumed by the mixture model estimator
\item[Uniform] $\epsilon~\mathcal{U}(-\sqrt{6}/2,\sqrt{6}/2)$
\item[Mixture] $f_\epsilon(x)=\frac{1}{4}\phi\left((x+1/3)\sqrt{6}\right)+\frac{1}{4}\phi\left((x-1)\sqrt{6})\right)$
\end{itemize}
The mixture distribution for $\epsilon$ was designed to be particularly difficult for the mixture model estimator.

The last two factors---the presence of interactions between covariates and $\st$ and $Z$, respectively---test the performance of a model that does not include interactions when that modeling assumption is false.

\subsubsection{Analysis Models}\label{sec:simMods}

Using each simulated dataset, we estimatinged $\eff0$ and $\eff1$ with three methods: a mixture model \eqref{eq:mixtureModelPS} including principal scores, the regression M-estimator proposed here, and third, the principal score weighting estimator \eqref{eq:psw}.

The methods all used the same ``observed'' dataset, consisting of outcomes $Y$, treatment assignements $Z$, two covariates $x_1$ and $x_2$, and $S$ observed only when $Z=1$.
The third covariate used in the data generating model, $x_3$, was ``unobserved''--since $x_3$ was correlated with both $\st$ and $Y$, Principal Ignorability (Assumption \ref{ass:pi}) was violated.

All three methods used the same principal score model, fit using data from the treatment group in which $S=\st$ was observed:
\begin{equation}\label{eq:psModsim}
  \fpp{\bm{\alpha}}{x_1,x_2}=logit^{-1}(\alpha_0+\alpha_1x_1+\alpha_2x_2)
\end{equation}

The mixture estimator then fit two outcome models.
For the treatment group, a normal regression:
\begin{equation}\label{eq:mix1-sim}
  f_{Y|Z=1}(y;\mut0,\mut1,\bm{\beta},\sigma_T)=\phi\left(\frac{y-[\mut0+\beta_1x_1+\beta_2x_2+(\mut1-\mut0)S_i]}{\sigma_T}\right)
\end{equation}
and for the control group, mixture model:
\begin{equation}\label{eq:mix-sim}
  f_{Y|Z=0}(y_i)=fpp{\bm{\alpha}}{x_{1i},x_{2i}}\phi\left(\frac{y-\muc1-\beta_1x_1-\beta_2x_2}{\sigma_C}\right)+(1-fpp{\bm{\alpha}}{x_{1i},x_{2i}})\phi\left(\frac{y-\muc0-\beta_1x_1-\beta_2x_2}{\sigma_C}\right)
\end{equation}
Note that the coefficients $\beta_1$ and $\beta_2$ are the same across treatment groups and principal strata, while the residual variance differs across treatment groups, but not across principal strata.
The models \eqref{eq:psModsim}, \eqref{eq:mix1-sim}, and \eqref{eq:mix-sim} were fit simultaneously using Bayesian Markov Chain Monte Carlo with Stan \citep{rstan}. We interpreted the posterior mean as a point estimate and the posterior standard deviation as a standard error.

In contrast, for regression M-estimation and principal score weighting the principal score model \eqref{eq:psModsim} was fit by itself, to data from the treatment group, using standard maximum likelihood methods.
For principal score weighting we extracted the estimated probabilities from \eqref{eq:psModsim}---the principal scores---plugged them in to \eqref{eq:psw} to estimate $\muc0$ and $\muc1$, and estimated $\mut0$ and $\mut1$ with the sample means of subjects with $Z=1$ and $S=0$ or $S=1$, respectively.

For regression M-estimation, we used the estimated principal scores in the control group, along with the observed $S$ in the treatment group, to construct $R$, and then fit the model
\begin{equation}\label{eq:simReg}
  Y_i=\muc0+(\muc1-\muc0) \ri+(\mut0-\muc0)Z_i+(\mut1-\muc1-\mut0+\muc0)Z_iR_i+\beta_1x_{1i}+\beta_2x_{2i}+\epsilon_i
\end{equation}
with OLS.
To estimate standard errors, we used the \texttt{sandwich} package \citep{sandwich} to extract estimating equations from \eqref{eq:psModsim} and \eqref{eq:simReg} and computed the sandwich covariance matrix following the procedure detailed in the appendix.

\subsection{Results}

\begin{figure}
  \centring
  \includegraphics{../simFigs/normalOutcomesNoInteractionPosB1.jpg}
  \caption{Violin and scatterplots of 500 simulation estimation errors for estimators described in Section~\ref{sec:simMods} when the distribution of residuals in \eqref{eq:yc-sim} was normal, there were no interactions between covariates and $S$ or $Z$, and $\alpha=0.2,0.5$. Annotations at the upper and lower edges of the plot indicate the number of extreme outliers excluded from the plotting area.}
  \label{fig:simNorm}
\end{figure}

\begin{figure}
  \centring
  \includegraphics{../simFigs/unifOutcomesNoInteractionPosB1.jpg}
  \caption{Violin and scatterplots of 500 simulation estimation errors for estimators described in Section~\ref{sec:simMods} when the distribution of residuals in \eqref{eq:yc-sim} was uniform, there were no interactions between covariates and $S$ or $Z$, and $\alpha=0.2,0.5$. Annotations at the upper and lower edges of the plot indicate the number of extreme outliers excluded from the plotting area.}
  \label{fig:simUnif}
\end{figure}

\begin{figure}
  \centring
  \includegraphics{../simFigs/InteractionPosB1.jpg}
  \caption{Violin and scatterplots of 500 simulation estimation errors for estimators described in Section~\ref{sec:simMods} when the distribution of residuals in \eqref{eq:yc-sim} was normal, there were possibly interactions between covariates and $S$ or $Z$, $\muc1=0.3$, and $\alpha=0.5$. Annotations at the upper and lower edges of the plot indicate the number of extreme outliers excluded from the plotting area.}
  \label{fig:simUnif}
\end{figure}

  \begin{table}
    \centering
  \input{rmseTab}
  \caption{RMSE }
\end{table}

\begin{table}
  \input{coverageTab}
  \caption{Coverage}
\end{table}

Analysis models had access to treatment assignment $Z$, outcomes $Y$,
principal stratum $S$ for members of the treatment group, and
covariates.
Even though three covariates, $x_1$, $x_2$ and $x_3$ were included as
predictors of principal stratum $S$ and outcome $Y$ in the data
generating model, both analysis approaches only had access to the
first two covariates, $x_1$ and $x_2$.
Hence, principal ignorability was violated---$S$ and $Y$ were
``confounted'' by unobserved variable $x_3$.

Simulated data were analyzed first with a ``stacked equations''/M-Estimation/GEE
approach, using the \texttt{geex} package in \texttt{R}
\citep{geex,rcite}, and then with a more conventional maximum
likelihood approach via \texttt{rstan} \citep{rstan}.

Both methods assumed (correctly) that there were no interactions
between covariates $x_1$ and $x_2$ and treatment assignment (after
accounting for $S$).

In the GEE approach included a set of nine estimating equations.
First, a GEE logistic regression fit to the treatment group to
estimate principal scores
$\ppi=logit^{-1}(\beta_0+\beta_1x_{1i}+\beta_2x_{2i})$ (three parameters):
\begin{align*}
  \sum_iZ_i\left[S_i-\ppi\right]&=0\\
  \sum_iZ_ix_{1i}\left[S_i-\ppi\right]&=0\\
  \sum_iZ_ix_{2i}\left[S_i-\ppi\right]&=0
\end{align*}
Next, four equations to estimate the outcome regression in the
treatment group, including terms for $S$:
\begin{align*}
  \sum_i
  Z_i\left[Y_i-(\gamma_1x_{1i}+\gamma_2x_{2i}+S_i\mu_{11}+(1-S_i)\mu_{10})\right]&=0\\
  \sum_i
  Z_ix_{1i}\left[Y_i-(\gamma_1x_{1i}+\gamma_2x_{2i}+S_i\mu_{11}+(1-S_i)\mu_{10})\right]&=0\\
  \sum_i
  Z_ix_{2i}\left[Y_i-(\gamma_1x_{1i}+\gamma_2x_{2i}+S_i\mu_{11}+(1-S_i)\mu_{10})\right]&=0\\
  \sum_i
  Z_iS_i\left[Y_i-(\gamma_1x_{1i}+\gamma_2x_{2i}+S_i\mu_{11}+(1-S_i)\mu_{10})\right]&=0
\end{align*}

Finally, two equations to estimate the mixture model in the control
group, for whom $S$ is unobserved:
\begin{align*}
  \sum_i
  (1-Z_i)\left[Y_i-(\gamma_1x_{1i}+\gamma_2x_{2i}+\ppi\mu_{01}+(1-\ppi)\mu_{00})\right]&=0\\
  \sum_i
  (1-Z_i)\left[Y_i-(\gamma_1x_{1i}+\gamma_2x_{2i}+\ppi^2\mu_{01}+(\ppi-\ppi^2)\mu_{00})\right]&=0
\end{align*}
The average treatment effect for $S=0$ was $\mu_{10}-\mu_{00}$ and for
$S=1$ was $\mu_{11}-\mu_{01}$.

Next, we estimated effects with MLE, using the built-in optimizer of
\texttt{rstan}.

In this model we regressed $S$ and $Y$ for the treatment group on an
intercept, $x_1$, and $x_2$ using a logistic and normal linear
specification, respectively.
Then, we specified a two-component mixture model for control
outcomes.

\section{Data Analysis}

\section{Discussion}

\appendix
\section{Proofs and Calculations}
\subsection{Proof for Lemma \ref{lemma:expectation}}
As a preliminary, note that
% \begin{align*}
%   \EE[\st|\hpp]&=\EE\left\{\EE[\st|\pp,\hpp]|\hpp\right\}\\
%              &=\EE\left\{\EE[\st|\pp]|\hpp\right\}\\
%              &=\EE[\pp|\hpp]=\hpp
% \end{align*}

%Then, note
\begin{align*}
  \EE[Y_C|\pp]&=\EE\left\{\EE[Y_C|\pp,\st]|\pp\right\}\\
             &=\EE\left\{\EE[Y_C|\st]|\pp\right\}\tag*{by \eqref{eq:assumption}}\\
             &=\EE[\muc1\st+\muc0(1-\st)|\pp]\\
             &=\muc1\pp+\muc0(1-\pp)
\end{align*}

Then we have
\begin{equation*}
  \begin{split}
    \EE[Y_C]=&\EE\EE[Y_C|\pp]=\muc1\EE\pp+\muc0(1-\EE\pp)\\
    =&\muc0+\EE\pp(\muc1-\muc0)
    \end{split}
\end{equation*}

Next we have

\begin{align*}
  \EE[Y_C\pp]&=\EE\left\{\EE[Y_C\pp|\pp]\right\}\\
            &=\EE\left\{\pp\EE[Y_C|\pp]\right\}\\
            &=\EE\left\{\pp\left[\muc1\pp+\muc0(1-\pp)\right]\right\}\\
            &=\EE[\pp]\muc0+\EE[\pp^2](\muc1-\muc0)
\end{align*}

In the treatment group, $\st$ is observed, so
\begin{align*}
    \EE[Y_T]=&\mut0+\EE[\st](\mut1-\mut0)\tag*{and}\\
    \EE[\st Y_T]=&\EE[\st]\mut0+\EE[\st^2](\mut1-\mut0)
\end{align*}

Due to Assumption \ref{ass:rand} (randomization), $\EE[Y|Z=0]=\EE[Y_C]$, $\EE[Y|Z=1]=\EE[Y_T]$, $\EE[Y\pp|Z=0]=\EE[Y_C\pp]$ and $\EE[YS|Z=1]=\EE[Y_T\st]$, completing the proof.

\subsection{Proof for Proposition \ref{prop:reg0}}
Carrying out those replacements in \eqref{eq:estEq0}, and rearranging\footnote{Technically, replacing $\tilde{\Psi}_i$ with $\Psi_i=\begin{psmallmatrix} 1 & 0&1&0\\ 0&1&0&1\\0&0&1&0\\0&0&0&1\end{psmallmatrix}\tilde{\Psi}_i$.} gives an equivalent set of estimating equations:
\begin{equation}\label{eq:estEq1}
\sum_{i=1}^{n_C}\Psi_i=\sum_{i=1}^{n_C}\begin{pmatrix}
    Y_i-\muc0-\ri(\muc1-\muc0)-Z_i(\mut0-\muc0)-Z_i\ri(\mut1-\mut0-\muc1+\muc0)\\
    \ri Y_i-\ri\muc0-\ri^2(\muc1-\muc0)-Z_i\ri(\mut0-\muc0)-Z_i\ri^2(\mut0-\mut1-\muc0+\muc1)\\
    Z_iY_i-Z_i\mut0-Z_i\ri (\mut1-\mut0)\\
    Z_i\ri Y_i -Z_i\ri\mut0-Z_i\ri^2(\mut1-\mut0)
  \end{pmatrix}=\begin{pmatrix} 0\\0\\0\\0\end{pmatrix}
\end{equation}

\subsection{Proof for Proposition \ref{prop:interaction}}
First, note that $\EE[S]=\EE\EE[S|\bx]=\EE[\pp]$ and
$$\EE[\bx S]=\EE[\bx S|Z=1]=\EE[\bx S|Z=0]=\EE[\bx \EE[\bx\EE[S|\bx]|Z=0]]=\EE[\bx\pp|Z=0]=\EE[\bx\pp]$$
implying that
\begin{equation*}
  \begin{split}
    \EE[Y|\bx,Z=0]&=\beta_0+\bm{\beta_1}\bxy+\beta_2\pp+\bm{\beta_5}'\bxy\pp\tag*{and}\\
    \EE[Y|\bx,S,Z=1]&=\beta_0+\beta_3+(\bm{\beta_1}+\bm{\beta_6})\bxy+(\beta_3+\beta_4)S+(\bm{\beta_5}+\bm{\beta_7})'\bxy S
  \end{split}
\end{equation*}
Therefore,
\begin{align*}
  \EE[Y]&=\EE[Y|Z=0]+\EE[Z]\left\{\EE[Y|Z=1]-\EE[Y|Z=0]\right\}\\
  &=\beta_0+\bm{\beta_1}\EE[\bxy]+\beta_2\EE[\pp]+\bm{\beta_5}'\EE[\bxy\pp]+\EE[Z]\left\{\beta_3+\bm{\beta_4}\EE[S]+\bm{\beta_6}'\EE[\bxy]+\bm{\beta_7}'\EE[\bxy S]\right\}\\
  &=\beta_0+\bm{\beta_1}\EE[\bxy]+\beta_2\EE[R]+\bm{\beta_5}'\EE[\bxyR]+\EE[Z]\left\{\beta_3+\bm{\beta_4}\EE[R]+\bm{\beta_6}'\EE[\bxy]+\bm{\beta_7}'\EE[\bxy R]\right\}\\
\end{align*}
Analogous reasoning leads to expressions for $\EE[RY]$, $\EE[\bxy Y]$, $\EE[ZY]$,  $\EE[ZRY]$, $\EE[\bxy YZ]$, and $\EE[\bxy RZY]$.
These, in turn, give rise to estimating equations
\begin{equation*}

  \displaystyle\sum_i \begin{pmatrix}
    Y_i-\beta_0-\bm{\beta_1}\bxy_i-\beta_2R_i-\bm{\beta_5}'\bxy_iR_i-Z_i\left\{\beta_3-\bm{\beta_4}R_i-\bm{\beta_6}'\bxy_i-\bm{\beta_7}'\bxy_i R_i\right\}\\
    R_iY_i-\beta_0R_i-\bm{\beta_1}\bxy_iR_i-\beta_2R_i^2-\bm{\beta_5}'\bxy_iR_i^2-Z_i\left\{\beta_3R_i-\bm{\beta_4}R_i^2-\bm{\beta_6}'\bxy_iR_i-\bm{\beta_7}'\bxy_i R_i^2\right\}\\
    \bxy_i Y_i-\beta_0\bxy_i-\bm{\beta_1}\bxy_i\bxy_i'-\beta_2R_i\bxy_i-\bm{\beta_5}'\bxy_iR_i-Z_i\left\{\beta_3\bxy_i-\bm{\beta_4}R_i\bxy_i-\bm{\beta_6}'\bxy_i\bxy_i'-\bm{\beta_7}'\bxy_i\bxy_i' R_i\right\}\\
    Z_iY_i-Z_i\left\{\beta_0-\bm{\beta_1}\bxy_i-\beta_2R_i-\bm{\beta_5}'\bxy_iR_i-\beta_3-\bm{\beta_4}R_i-\bm{\beta_6}'\bxy_i-\bm{\beta_7}'\bxy_i R_i\right\}\\
    Z_iR_iY_i-Z_iR_i\left\{\beta_0-\bm{\beta_1}\bxy_i-\beta_2R_i-\bm{\beta_5}'\bxy_iR_i-\beta_3-\bm{\beta_4}R_i-\bm{\beta_6}'\bxy_i-\bm{\beta_7}'\bxy_i R_i\right\}\\
    Z_iY_i\bxy_i-Z_i\left\{\beta_0\bxy_i-\bm{\beta_1}\bxy_i\bxy_i'-\beta_2R_i\bxy_i-\bm{\beta_5}'\bxy_i\bxy_i'R_i-\beta_3\bxy_i-\bm{\beta_4}R_i\bxy_i-\bm{\beta_6}'\bxy_i\bxy_i'-\bm{\beta_7}'\bxy_i\bxy_i' R_i\right\}\\
    R_iZ_iY_i\bxy_i-Z_iR_i\left\{\beta_0\bxy_i-\bm{\beta_1}\bxy_i\bxy_i'-\beta_2R_i\bxy_i-\bm{\beta_5}'\bxy_i\bxy_i'R_i-\beta_3\bxy_i-\bm{\beta_4}R_i\bxy_i-\bm{\beta_6}'\bxy_i\bxy_i'-\bm{\beta_7}'\bxy_i\bxy_i' R_i\right\}
  \end{pmatrix}=\begin{pmatrix} 0\\0\\0\\0\\0\\0\\0\end{pmatrix}
\end{equation*}
These are the estimating equations for the regression model \eqref{eq:fullInteractionModel}, fit by OLS. Now note that
\begin{equation*}
  \EE[Y_T-Y_C|\st=0]=\beta_0^T-\beta_0^C+(\bm{\beta_1}^T'-\bm{\beta_1}^C')\EE[\xby|S=0]=\beta_3+\beta_6'E[\xby|S=0]
\end{equation*}
and
\begin{equation*}
  \EE[Y_T-Y_C|\st=1]=\beta_0^T+\beta_2^T-(\beta_0^C+\beta_2^C)+(\bm{\beta_1}^T'+\bm{\beta_3^T}'-\bm{\beta_1}^C'-\bm{\beta_3^C}')\EE[\xby|S=0]=\beta_3+\beta_4+(\beta_6'+\beta_7')E[\xby|S=1]
\end{equation*}
the result follows

\end{document}

### 2022-06-24 code/fh2t/analysis_stanModels.R

@@ -517,8 +517,33 @@ stopifnot(max(sdatClass$classT)==max(sdatClass$classC))
 stopifnot(max(sdatClass$classT)==n_distinct(sdatClass$classT))
 stopifnot(max(sdatClass$classT)==n_distinct(sdatClass$classC))
 
+
 checkmodRandClass=stan('code/fh2t/psModRandClass.stan',data=sdatClass)
 save(checkmodRandClass,file='checkmodRandClass.RData')
+
+
+checkdathalf1=subset(checkdat,classN<67)
+stanDatHalf <-  list(nc= length(checkdathalf1[checkdathalf1$Z == 0,]$student_number), #
+                 nt= length(checkdathalf1[checkdathalf1$Z == 1,]$student_number), #
+                 ncovU= ncol(XctlU), # √
+                 ncovY= ncol(XctlY), # √
+                 YctlY=  (checkdathalf1[checkdathalf1$Z == 0, ]$Y), ### IS THE OUTCOME Y OR post.total_math_score?
+                 #    YctlU= , this doesn't exist, right?
+                 YtrtY= (checkdathalf1[checkdathalf1$Z == 1, ]$Y ), ### IS THE OUTCOME Y OR post.total_math_score?
+                 YtrtU=(ifelse(checkdathalf1[checkdathalf1$Z == 1, ]$anyBottom == "TRUE", 1, 0)),
+                 XctlU=as.matrix(XctlU)[checkdat$classN[checkdat$Z==0]<67,], # √
+                 XctlY=as.matrix(XctlY)[checkdat$classN[checkdat$Z==0]<67,], # √
+                 XtrtU=as.matrix(XtrtU)[checkdat$classN[checkdat$Z==1]<67,], # √
+                 XtrtY=as.matrix(XtrtY)[checkdat$classN[checkdat$Z==1]<67,], # √
+                 classT=checkdathalf1$classN[checkdathalf1$Z==1],
+                 classC=checkdathalf1$classN[checkdathalf1$Z==0],
+                 nclass=66,
+                 bottomOuter=(ifelse(checkdathalf1[checkdathalf1$Z == 1, ]$anyBottom == "TRUE", 1, 0))
+)
+
+modHalf1=stan('code/fh2t/psModRandClass.stan',data=stanDatHalf)
+save(modHalf1,file='modHalf1.RData')
+
 
                                         # saveRDS(checkmod2, "checkmod2.rds")
 # mod <- readRDS("model.rds")

### 2022-06-24 code/regression.r

@@ -39,9 +39,9 @@ datNames <- function(data,trt='Z',out='Y',use='S',block=NULL){
   data$Y <- as.numeric(data[[out]])
   data$Z <- as.numeric(data[[trt]])
   data$S <- as.numeric(data[[use]])
-  
+
   data$S[data$Z==0] <- NA
-  
+
   if(!is.null(block)){
     data$block <- as.factor(data[[block]])
     data <- data%>%
@@ -60,9 +60,9 @@ datNames <- function(data,trt='Z',out='Y',use='S',block=NULL){
 
 ### computes/returns regression models
 pointEst <- function(data,covFormU=~x1+x2,covFormY=covFormU,int=FALSE,psMod=NULL){
-  
+
   data$S[data$Z==0] <- NA
-  
+
     if(is.null(psMod)) psMod <- glm(
       update(covFormU,S~.),
       data=data,family=binomial,
@@ -75,17 +75,24 @@ pointEst <- function(data,covFormU=~x1+x2,covFormY=covFormU,int=FALSE,psMod=NULL
     if('Sp'%in%names(data)) warning('replacing Sp')
     data <- within(data,Sp <- ifelse(Z==1&!is.na(S),S,ps))
 
-    outForm <- 
+    outForm <-
       update(covFormY,
         if(int) Y~Z*(Sp+.) else Y~Z*Sp+.
         )
-    
+
   if('block'%in%names(data)) if(!is.null(data$block))
       outForm <- update(outForm,.~.+block)
-    
+
     outMod <- lm(outForm,data=data)
 
     list(psMod=psMod,outMod=outMod)
+}
+
+intEst0 <- function(dat){
+  Attach(pointEst(dat,int=TRUE))
+
+  coef(outMod)['Z']+
+    crossprod(coef(outMod)[c('Z:x1','Z:x2')],colMeans(subset(dat,Z==1&S==0,select=c('x1','x2'))))
 }
 
 ### estimating equations for stage 1 model (PS)
@@ -172,7 +179,7 @@ est <- function(data,covFormU=~x1+x2,covFormY=covFormU,psMod=NULL,clust=NULL,int
                 trt='Z',out='Y',use='S',block=NULL){
 
   data <- datNames(data,trt=trt,out=out,use=use,block=block)
-  
+
     Attach(pointEst(data=data,covFormU=covFormU,covFormY=covFormY,int=int,psMod=psMod))
 
[…]

### 2022-06-24 code/simulation/displaySim.r

@@ -6,34 +6,6 @@ source('code/simulation/readSimFuncs.r')
 #### after simulation results have been loaded and pre-processed...
 
 
-print(load('simResults/fullResults.RData'))
-print(load('simResults/pswResults.RData'))
-
-if(!is.data.frame(results)) results <- reduce(results,bind_rows)
-
-facsDF<-distinct(results,n,mu01,mu10,mu11,b1,errDist,intS,intZ,run)
-stopifnot(all.equal(facsDF$run,1:length(pswResults)))
-### transform pswResults
-pswResults <-
-  map(seq(length(pswResults)),
-      function(i){
-        res<-pswResults[[i]]
-        facs<-attr(res,'facs')
-        if(!all(as.data.frame(facsDF[i,names(facs)])==facs)){
-          print(i)
-          stop()
-        }
-        res<-cbind(res,facs)
-        res$run<-i
-        res$effDiff=res$eff1-res$eff0
-        res=pivot_longer(res,starts_with('eff'),names_to='eff',names_prefix='eff',values_to='est')
-        samp<-unique(na.omit(results$samp[results$run==i]))
-        res$samp <- if(length(samp)==nrow(res)) samp else NA
-        res
-      })%>%
-  reduce(bind_rows)
-
-pswResults$estimator<-'psw'
 
 ### rhats
 rhats <- results%>%
@@ -65,13 +37,6 @@ results%>%
   geom_smooth()+
   geom_hline(aes(yintercept=mu11-mu01))+
   facet_grid(estimator~mu01)+ylim(-1,1)
-
-
-
-results<-bind_rows(results,pswResults)%>%
-  group_by(run,eff)%>%
-  mutate(pop=na.omit(pop)[1],samp=ifelse(is.na(samp),pop,samp))%>%
-  ungroup()
 
 
 pd <- results %>%
@@ -518,7 +483,7 @@ rate%>%
 dev.off()
 
 
-
+### n=100 results
 pd100 <- res100 %>%
   group_by(b1)%>%
   mutate(AUCf=mean(auc,na.rm=TRUE))%>%
@@ -538,3 +503,73 @@ pd100 <- res100 %>%
     interactionS=ifelse(intS,"S interaction","No\nS interaction")
   ) %>%
[…]

### 2022-06-24 code/simulation/interactionSim.r

@@ -1,0 +1,127 @@
+sss0=function(n){
+  x=rnorm(n)
+  p=plogis(.5*x)
+  S=rbinom(n,1,p)
+  Y=0.3+0.3*S+rnorm(n)
+
+  vy=var(Y)
+
+  y0=mean(Y[S==0])
+
+  beta=cov(p,Y)/var(p)
+  p0=mean(p)
+  c(mu0=y0,mu0hat=mean(Y)-beta*p0,vy=vy)
+}
+
+sss=function(n){
+  x=rnorm(n)
+  ps=plogis(.5*x)
+  S=rbinom(n,1,ps)
+  Y=.3-0.5*x+0.3*S+rnorm(n)
+  vy=var(Y)
+#  mod=lm(Y~x+ps)
+  y0=mean(Y[S==0])
+  x0=mean(x[S==0])
+  rm(S);gc()
+                                        #  c(coef(mod),coef(mod)[1]+coef(mod)[2]*mean(x[S==0]),mean(Y[S==0]))
+  vx=var(x)
+  xy=cov(x,Y)
+  xp=cov(x,ps)
+  xb=mean(x)
+  rm(x); gc()
+  py=cov(ps,Y)
+  Yb=mean(Y)
+  rm(Y); gc()
+  vp=var(ps)
+  pb=mean(ps)
+
+  m1=(py-xp*xy/vx)/(vp-xp^2/vx)
+  b=(xy*vp-py*xp)/(vx*vp-xp^2)
+  m0=Yb-b*xb-m1*pb
+
+
+  c(mu0=y0,mu0hat=m0+x0*b,m1,b,m0,vy=vy)
+
+}
+
+
+sssInt=function(n){
+  x=rnorm(n)
+  p=plogis(.5*x)
+  S=rbinom(n,1,p)
+  y=.3-0.5*x+0.3*S+.3*S*x+rnorm(n)
+  vy=var(y)
+#  mod=lm(y~x+ps)
+  y0=mean(y[S==0])
+  x0=mean(x[S==0])
+  rm(S);gc()
+                                        #  c(coef(mod),coef(mod)[1]+coef(mod)[2]*mean(x[S==0]),mean(y[S==0]))
+  px=mean(p*x)
+  p2x=mean(p^2*x)
+  px2=mean(p*x^2)
+  p2x2=mean(p^2*x^2)
+  x2=mean(x^2)
+  xy=mean(x*y)
[…]

### 2022-06-24 code/simulation/readSim.r

@@ -1,461 +1,13 @@
 library(tidyverse)
-library(rstan)
-library(geex)
 library(parallel)
-library(pbapply)
 
-bias <- function(sss){
+source('code/simulation/readSimFuncs.r')
 
-    sss1 <- sss[,1:6]
-    sss2 <- sss[,7:11]
+#### read, process results from main simulation
+print(load('simResults/pswResults.RData'))
 
-    colMeans(cbind(sss1[,3:6]-sss1[,c(1,2,1,2)],sss2))
-}
+results=loadRes(pswResults=pswResults)
 
-rmse <- function(sss){
+resultsN100=loadRes(ext2='n100')
 
-    sss1 <- sss[,1:6]
-    sss2 <- sss[,7:11]
-
-    c(
-        sqrt(colMeans((sss1[,3:6]-sss1[,c(1,2,1,2)])^2)),
-        colMeans(sss2)
-    )
-}
-
-
-summ <- function(sss){
-    sss1 <- sss[,1:6]
-    sss2 <- sss[,7:11]
-    rbind(colMeans(sss1),
-          apply(sss1,2,sd),
-          sqrt(row1Means((sss-sss[rep(1:2,3),])^2))
-          )
-    }
-
-
-plotRes <- function(rrr,fac,S){
-    www <- (max(rrr[fac,])-min(rrr[fac,]))*0.02
-    plot(rrr[fac,]-www,rrr[paste0('mom',S),],ylim=range(c(rrr[paste0('mom',S),],rrr[paste0('mle',S),])),xlim=c(min(rrr[fac,])-2*www,max(rrr[fac,])+2*www),pch=16,xlab=fac)
-    points(rrr[fac,]+www,rrr[paste0('mle',S),],col='red',pch=16)
-    legend('topright',legend=c('MOM','MLE'),col=c('black','red'),pch=16)
-    abline(h=0,lty=2)
-}
-
-
-getSumms <- function(res,S,EST){
-    #cases <- results%>%select(starts_with('mu'),n,b1,gumb)%>%distinct()
-    true <- res[[paste0('true',S)]]
-    est <- res[[paste0(EST,S)]]
-    se <- res[[paste0(EST,S,'se')]]
-
-    cbind(
-        res[1,which(sapply(res,n_distinct)==1)],
-        data.frame(
-            EST=EST,
-            S=S,
-            getEst=mean(is.finite(est)),
-            getSE=mean(is.finite(se)),
[…]

### 2022-06-24 code/simulation/readSimFuncs.r

@@ -1,9 +1,3 @@
-
-
-library(tidyverse)
-library(rstan)
-library(parallel)
-
 bias <- function(sss){
 
     sss1 <- sss[,1:6]
@@ -39,27 +33,39 @@ muEff <- function(facs,eff)
          ifelse(eff==1,mu11-mu01,mu11-mu01-mu10))
          )
 
+psw1Proc=function(psw){
+  psw['effDiff']=psw['eff1']-psw['eff0']
+  cbind(
+    estimates=psw[c('eff0','eff1','effDiff')],
+    SE=NA,
+    psw=1
+  )
+}
 
 bayesProc1 <- function(eff, res1) {
   if (rownames(res1$mest)[3] == 'diff')
     rownames(res1$mest)[3] <- 'effDiff'
+
+  res1$psw <- if('psw'%in% names(res1)) psw1Proc(res1$psw) else NULL
+
+
   with(res1,
-       map_dfr(
-         list(bayes, mest),
-         ~
-           tibble(
-             eff = eff,
-             pop = muEff(facs, eff),
-             samp = ifelse(eff == 'Diff', true['S1'] -true['S0'], true[paste0('S', eff)]),
-             estimator = ifelse('mean' %in% colnames(.), 'bayes', 'mest'),
-             est = .[paste0('eff', eff), ifelse('mean' %in% colnames(.), 'mean', 'estimates')],
-             se = .[paste0('eff', eff), ifelse('sd' %in%colnames(.), 'sd', 'SE')],
-             CInormL = est - 2 *se,
-             CInormU = est + 2 *se,
-             CIpercL = ifelse('2.5%' %in% colnames(.), .[paste0('eff', eff), '2.5%'], CInormL),
-             CIpercU = ifelse('97.5%' %in% colnames(.), .[paste0('eff', eff), '97.5%'], CInormU),
-             rhat = bayes[paste0('eff', eff), 'Rhat'],
-             auc = attr(mest, 'auc')
+   map_dfr(
+    if('psw'%in%names(res1)) list(bayes,mest,psw) else list(bayes,mest),
+    ~
+      tibble(
+       eff = eff,
+       pop = muEff(facs, eff),
+       samp = ifelse(eff == 'Diff', true['S1'] -true['S0'], true[paste0('S', eff)]),
+       estimator=ifelse('mean'%in%colnames(.),'bayes',ifelse('psw'%in%colnames(.),'psw','mest')),
[…]

### 2022-06-24 code/simulation/runSimBayes.r

@@ -35,3 +35,24 @@ fullsim(500,ns=c(500,1000),mu01=c(0,.3),mu10=0,b1s=c(0,0.2,0.5),
 	ext='',cl=NULL,start=1)
 
 system('drive push --no-clobber -quiet')
+
+
+pswResults=sapply(seq(max(as.numeric(gsub('dat|\\.RData','',list.files('simData/'))),na.rm=TRUE)),
+                  function(i){
+                    print(i)
+                    load(paste0('simData/dat',i,'.RData'))
+                    facs=attributes(datasets[[1]])$facs
+                    out=as.data.frame(t(vapply(datasets,psw,psw(datasets[[1]]))))
+                    attr(out,'facs')=facs
+                    out
+                  },simplify=FALSE)
+save(pswResults,file='simResults/pswResults.RData')
+
+
+################################
+### some further investigations
+fullsim(500,ns=100,mu01=.3,mu10=0,errDist=c('norm','unif'),b1s=c(0,.2,.5),ext='n100')
+
+fullsim(500,ns=c(200,300,400),errDist='norm',b1s=.5,intS=FALSE,intZ=FALSE,mu01=.3,mu10=0,ext='ns')
+
+fullsim(500,ns=500,errDist='norm',b1s=c(.3,.4,seq(.6,1,.1)),intS=FALSE,intZ=FALSE,mu01=.3,mu10=0,ext='b1s')

### 2022-06-24 code/simulation/simFuncs.r

@@ -14,7 +14,7 @@ bayes <- function(data,...){
                        )
                  )
                                         #fit1 <-
-    output <- capture.output(fit <- stan('code/ps.stan',data=sdat,...))
+    output <- capture.output(fit <- stan('code/ps.stan',data=sdat))#,...))
 
     summary(fit, par=c('eff0','eff1','effDiff'))$summary
 }

### 2022-06-27 code/simulation/displaySim.r
one boxplot fig in paper

@@ -1,10 +1,15 @@
 library(tidyverse)
 library(broom)
 library(kableExtra)
+library(gridExtra)
 
 source('code/simulation/readSimFuncs.r')
-#### after simulation results have been loaded and pre-processed...
 
+#### after simulation results have been loaded and pre-processed...
+## source('code/simulation/readSim.r')
+load('simResults/fullResults.RData')
+load('simResults/resultsNs.RData')
+load('simResults/resultsB1s.RData')
 
 
 ### rhats
@@ -24,6 +29,22 @@ lm(Rhat1.1~I(n/500)+mu01+errDist+b1+intS+intZ,data=rhats)
 results%>%ggplot(aes(as.factor(b1),auc))+geom_jitter(alpha=0.2)+geom_violin(draw_quantiles = 0.5)
 
 results%>%ggplot(aes(as.factor(n),auc))+geom_jitter(alpha=0.2)+geom_violin(draw_quantiles = 0.5)+facet_wrap(~b1,nrow=1)
+
+
+resultsB1s%>%
+  group_by(b1)%>%
+  mutate(meanAUC=mean(auc))%>%
+  ggplot(aes(as.factor(b1),auc))+geom_boxplot()+geom_point(aes(y=meanAUC))+geom_smooth(se=FALSE)
+                                        #  scale_y_continuous('Avg. AUC',seq(.5,1,.1))
+resultsB1s%>%
+  group_by(b1)%>%
+  summarize(meanAUC=mean(auc))
+
+resultsB1s%>%
+  group_by(b1)%>%
+  summarize(meanAUC=mean(auc))%>%ggplot(aes(b1,meanAUC))+geom_point()+geom_smooth(method='lm')
+
+
 
 #### how do estimates vary with auc (simple case)?
 results%>%
@@ -50,12 +71,13 @@ pd <- results %>%
     AUCff=paste0('AUC=',round(AUCf,1)),
     N=paste0('n=',n),
     M1=factor(mu01,levels=c("0","0.3"),
-              labels=c(bquote(mu[c]^1-mu[c]^0==0),bquote(mu[c]^1-mu[c]^0==0.3))),
+              ##labels=c(bquote(mu[c]^1-mu[c]^0==0),bquote(mu[c]^1-mu[c]^0==0.3))),
+              labels=c(bquote(mu[c]^1==0),bquote(mu[c]^1==0.3))),
     PE=paste('Stratum',eff),
[…]

### 2022-06-28 writeUps/paper.tex
cleaning up displaySim.r pt 1

@@
     \begin{split}
-      Y_{Ci}=\beta_0^C+\bm{\beta_1}^{C'}\bxy_i+\beta_2^C\sti+\bm{\beta_3^C}'\bxy_i\sti+\epsilon_{Ci}\\
-      Y_{Ti}=\beta_0^T+\bm{\beta_1}^{T'}\bxy_i+\beta_2^T\sti+\bm{\beta_3^T}'\bxy_i\sti+\epsilon_{Ti}\\
+      Y_{Ci}=\beta_0^C+\bm{\beta_1}^{C'}\bxy_i+\beta_2^C\sti+\bm{\beta_3^{C'}}\bxy_i\sti+\epsilon_{Ci}\\
+      Y_{Ti}=\beta_0^T+\bm{\beta_1}^{T'}\bxy_i+\beta_2^T\sti+\bm{\beta_3^{T'}}\bxy_i\sti+\epsilon_{Ti}\\
     \end{split}
@@
 \subsubsection{Manipulated Factors}
-We manipulated five factors in the simulation:
+We manipulated six factors in the simulation:
 \begin{itemize}
-%\item Sample size $n\in\{500,1000\}$
+\item Sample size per treatment group $n\in\{100,200,\dots,1000\}$
+\item $\alpha \in \{0,0.2,0.3,\dots,1\}$
 \item $\muc1\in\{0,0.3\}$
-\item $\alpha \in \{0,0.2,0.5,1\}$
 \item The distribution of $\epsilon_i$
@@
 \end{itemize}
+The six factors are not completely crossed---in particular, we let $n$ and $\alpha$ each vary across all the levels listed while holding the other five factors fixed.
+
+The guarentees of M-estimation are all assymptotic, so we examined the behavior of our M-estimator at a range of sample sizes to determine its finite-sample properties.
 
@@
 We expect that the more predictive the covariates, the better the performance of all three estimators.
+To aid in the interpretation of the $\alpha$ parameter, the online appendix plots $\alpha$ against the area under the receiver operating characteristic curve for each fitted model (AUC), a common measure of classification accuracy \citep{bradley1997use}---an AUC of 0.5 implies that the model is no better than random guessing, and AUC of 1 implieds perfect prediction.
+For $0.5\le \alpha \le 1$, the average AUC varies from roughly 0.53 to 0.77; for $\alpha=0.5$, the average AUC was roughly 0.675.
 
@@
 Across simulation runs, $E[\epsilon]=0$ and $Var(\epsilon)=1/2$.
-However, the shape of $\epsilon$'s distribution varied between the following three possibilities:
+However, the shape of $\epsilon$'s distribution varied between the following two possibilities:
 \begin{itemize}
-\item[Normal] $\epsilon~\mathcal{N}(0,1/2)$, as assumed by the mixture model estimator
-\item[Uniform] $\epsilon~\mathcal{U}(-\sqrt{6}/2,\sqrt{6}/2)$
-\item[Mixture] $f_\epsilon(x)=\frac{1}{4}\phi\left((x+1/3)\sqrt{6}\right)+\frac{1}{4}\phi\left((x-1)\sqrt{6})\right)$
+\item[Normal] $\epsilon\sim\mathcal{N}(0,1/2)$, as assumed by the mixture model estimator
+\item[Uniform] $\epsilon\sim\mathcal{U}(-\sqrt{6}/2,\sqrt{6}/2)$
+%\item[Mixture] $f_\epsilon(x)=\frac{1}{4}\phi\left((x+1/3)\sqrt{6}\right)+\frac{1}{4}\phi\left((x-1)\sqrt{6})\right)$
 \end{itemize}
-The mixture distribution for $\epsilon$ was designed to be particularly difficult for the mixture model estimator.
+%The mixture distribution for $\epsilon$ was designed to be particularly difficult for the mixture model estimator.
 
@@
 \begin{equation}\label{eq:psModsim}
-  \fpp{\bm{\alpha}}{x_1,x_2}=logit^{-1}(\alpha_0+\alpha_1x_1+\alpha_2x_2)
+  \ppf{\bm{\alpha}}{x_1,x_2}=logit^{-1}(\alpha_0+\alpha_1x_1+\alpha_2x_2)
 \end{equation}
@@
 \subsection{Results}
+In this section we present the subset of simulation results that we think is the most telling and interesting; a fuller set of results is available in the online appendix.
+
+\subsubsection{Varying $n$ and $\alpha$}
+When $\alpha=0$, Assumption \ref{ass:vps} is violated, and M-estimates are highly erratic; however, as Table \ref{tab:coverage} shows, when other assumptions hold inference remains valid.
+For the sake of clarity, we will focus here on cases where $\alpha>0$; results for $\alpha=0$ are in Tables \ref{tab:rmse} and \ref{tab:coverage} and in the online appendix.
+
+Figures \ref{fig:n} and \ref{fig:alpha} show the empirical bias and standard error (SE) of the M-estimator and mixture estimator of $\eff1$ as sample size and $\alpha$ vary.
+Across $n$ and $\alpha$, the M-estimator has lower bias and higher SE than the mixture estimator.
+The M-estimator is positively biased for low $n$ with bias converging to close to 0 (up to simulation error) for $n\ge 300$, and appears to be roughly unbiased (up to simulation error) for $0.2\le \alpha 1$, with $n$ fixed at 500.
+The SE of the M-estimator decreases rapidly for $100\le n \le 500$ and $0.2\le \alpha \le 0.5$ and more gradually for higher values of $n$ and $\alpha$.
+
+For the remainder of the simulation, we focus on the moderate cases of $n=500$ and $\alpha=0.2$ or 0.5.
 
 \begin{figure}
-  \centring
-  \includegraphics{../simFigs/normalOutcomesNoInteractionPosB1.jpg}
-  \caption{Violin and scatterplots of 500 simulation estimation errors for estimators described in Section~\ref{sec:simMods} when the distribution of residuals in \eqref{eq:yc-sim} was normal, there were no interactions between covariates and $S$ or $Z$, and $\alpha=0.2,0.5$.
-Annotations at the upper and lower edges of the plot indicate the number of extreme outliers excluded from the plotting area.}
-  \label{fig:simNorm}
+  \centering
+  \includegraphics{../simFigs/biasSEbyN.jpg}
+  \caption{Empirical bias and standard error of M- and Mixture estimates of $\muc1$ as sample size per treatment group $n$ varies from 0 to 1.
+There were 500 replications for each $n$.
+Other factors were held fixed at the noted values.}
+  \label{fig:n}
 \end{figure}
@@
 \begin{figure}
-  \centring
-  \includegraphics{../simFigs/unifOutcomesNoInteractionPosB1.jpg}
-  \caption{Violin and scatterplots of 500 simulation estimation errors for estimators described in Section~\ref{sec:simMods} when the distribution of residuals in \eqref{eq:yc-sim} was uniform, there were no interactions between covariates and $S$ or $Z$, and $\alpha=0.2,0.5$.
-Annotations at the upper and lower edges of the plot indicate the number of extreme outliers excluded from the plotting area.}
-  \label{fig:simUnif}
+  \centering
+  \includegraphics{../simFigs/biasSEbyB1.jpg}
+  \caption{Empirical bias and standard error of M- and Mixture estimates of $\muc1$ as $\alpha$ varies from 0.1 to 1.
+There were 500 replications for each $n$.
+Other factors were held fixed at the noted values.}
+  \label{fig:alpha}
 \end{figure}
 
+\subsubsection{Other Factors}
+
 \begin{figure}
-  \centring
-  \includegraphics{../simFigs/InteractionPosB1.jpg}
-  \caption{Violin and scatterplots of 500 simulation estimation errors for estimators described in Section~\ref{sec:simMods} when the distribution of residuals in \eqref{eq:yc-sim} was normal, there were possibly interactions between covariates and $S$ or $Z$, $\muc1=0.3$, and $\alpha=0.5$.
+  \centering
+  \includegraphics{../simFigs/boxplots.pdf}
+  \caption{Violin and scatterplots of 500 simulation estimation errors for estimators described in Section~\ref{sec:simMods} under varying conditions.
+For all plots, $n=500$; there were 500 replications for each set of conditions.
 Annotations at the upper and lower edges of the plot indicate the number of extreme outliers excluded from the plotting area.}
-  \label{fig:simUnif}
+  \label{fig:boxplots}
 \end{figure}
+
+Figure \ref{fig:boxplots} shows violin plots, layered on top of jitters scatterplots, showing estimation error for the M-estimator, the mixture estimator, and the PSW estimator of $\eff1$ under various sets of conditions, and Table \ref{tab:rmse} shows root mean squared error (RMSE) estimates under a somewhat wider set of circumstances when $\muc1=0.3$.
+The leftmost panel of Figure \ref{fig:boxplots} and the top line of Table \ref{tab:rmse} show results under ``standard'' conditions---normal residuals and no interactions in the data generating model between covariates and either $S$ or $Z$.
+In this case, the M-estimator is roughly unbiased, but less precise than the other two methods---when $\alpha=0.2$ it is much noisier, and when $\alpha=0.5$ it is only slightly less precise.
+The mixture estimator's bias depends on $\muc1$---when $\muc1=0$, the bias is slight, and when $\muc1=0.3$ the bias is more substantial.
+This is somewhat surprising, because when $\muc1=0$, there is no separation (``dispersion'') between the principal strata in the control group, and previous literature \citep{griffin2008application} has identified this scenario as a particular challenge for mixture modeling.
+The PSW estimator is also biased---due, presumeably, to the violation of Principal Ignorability, Assumption \ref{ass:pi}--inherent in the data generating model.
+On the other hand, the PSW estimator has easily the lowest sampling variance of the three estimators.
+In fact, Table \ref{tab:rmse} shows that despite its bias, the PSW estimator has lower RMSE than the M-estimator when $\alpha=0$ or 0.2.
+The middle panel of Figure \ref{fig:boxplots} and the fifth line of Table \ref{tab:rmse} shows a similar set of results for when the residuals in \eqref{eq:yc-sim} are uniformly distributed.
+In this case, the M-estimator and PSW estimator behave in a similiar fashion to the normal case, but the mixture estimator shows a bimodal pattern--sometimes over-estimating $\eff0$ and sometimes underestimating, but rarely estimationg $\eff1$ accurately.
+
+The rightmost panel of Figure \ref{fig:boxplots} returns to the case of normal residuals---and only shows results when $\muc1=0.3$ and $\alpha=0.5$---but introduces interactions in the data generating model between covariates and either $\st$, $Z$, or both.
+That is, the coefficients on the covariates in the outcome model vary between principal strata and/or between treatment groups.
+Corresponding RMSE results, along with analogous results for uniform residuals and other values for $\alpha$, are in Table \ref{tab:rmse}.
+Presence of interactions could undermine the M-estimate by inducing a relationship between principal scores and potential outcomes, even after adjusting the outcomes for covariate effects.
+Fortunately, the presence of interactions in the data generating model in this scenario appear to make little difference, though RMSEs were slightly higher when interactions between $Z$ and covariates are present.
+Table \ref{tab:rmse} shows that when $\alpha=.2$, M-estimation RMSE can be substantially higher when $Z$ interactions are present than when they are not; plots in the online appendix suggest that this is due to an increase in sampling variance and kurtosis rather than bias.
+
+Table \ref{tab:coverate} shows estimates of the empirical coverage of nominal 95\% confidence intervals for the M-estimator and mixture estimator of $\eff1$, to evaluate their respective standard error estimates and the associated statistical inference.
+When $\alpha=0.5$, M-estimator confidence intervals consistently acheive their nominal levels\footnote{With 500 replications and true coverage probabilities of 0.95, the standard error for emperical coverage rates is approximately 1 percentage point.}, whereas confidence intervals from mixture estimators undercover in the presence of interactions between $Z$ and $X$, or when residuals are non-normal.
+When $\alpha=0$ or 0.2, M-estimation confidence intervals undercover in the presence of interactions between $X$ and $Z$.
+Across all conditions in the table, the M-estimation confidence intervals have higher empirical coverage than their mixture-model counterparts.
+
+%% \begin{figure}
+%%   \centring
+%%   \includegraphics{../simFigs/normalOutcomesNoInteractionPosB1.jpg}
+%%   \caption{Violin and scatterplots of 500 simulation estimation errors for estimators described in Section~\ref{sec:simMods} when the distribution of residuals in \eqref{eq:yc-sim} was normal, there were no interactions between covariates and $S$ or $Z$, and $\alpha=0.2,0.5$.
+Annotations at the upper and lower edges of the plot indicate the number of extreme outliers excluded from the plotting area.}
+%%   \label{fig:simNorm}
+%% \end{figure}
+
+%% \begin{figure}
+%%   \centring
+%%   \includegraphics{../simFigs/unifOutcomesNoInteractionPosB1.jpg}
+%%   \caption{Violin and scatterplots of 500 simulation estimation errors for estimators described in Section~\ref{sec:simMods} when the distribution of residuals in \eqref{eq:yc-sim} was uniform, there were no interactions between covariates and $S$ or $Z$, and $\alpha=0.2,0.5$.
+Annotations at the upper and lower edges of the plot indicate the number of extreme outliers excluded from the plotting area.}
+%%   \label{fig:simUnif}
+%% \end{figure}
+
+%% \begin{figure}
+%%   \centring
+%%   \includegraphics{../simFigs/InteractionPosB1.jpg}
+%%   \caption{Violin and scatterplots of 500 simulation estimation errors for estimators described in Section~\ref{sec:simMods} when the distribution of residuals in \eqref{eq:yc-sim} was normal, there were possibly interactions between covariates and $S$ or $Z$, $\muc1=0.3$, and $\alpha=0.5$.
+Annotations at the upper and lower edges of the plot indicate the number of extreme outliers excluded from the plotting area.}
+%%   \label{fig:simUnif}
+%% \end{figure}
 
@@
   \caption{Coverage}
+  \label{tab:coverage}
 \end{table}
@@
   \begin{split}
-    \EE[Y|\bx,Z=0]&=\beta_0+\bm{\beta_1}\bxy+\beta_2\pp+\bm{\beta_5}'\bxy\pp\tag*{and}\\
+    \EE[Y|\bx,Z=0]&=\beta_0+\bm{\beta_1}\bxy+\beta_2\pp+\bm{\beta_5}'\bxy\pp\\
     \EE[Y|\bx,S,Z=1]&=\beta_0+\beta_3+(\bm{\beta_1}+\bm{\beta_6})\bxy+(\beta_3+\beta_4)S+(\bm{\beta_5}+\bm{\beta_7})'\bxy S
@@
   &=\beta_0+\bm{\beta_1}\EE[\bxy]+\beta_2\EE[\pp]+\bm{\beta_5}'\EE[\bxy\pp]+\EE[Z]\left\{\beta_3+\bm{\beta_4}\EE[S]+\bm{\beta_6}'\EE[\bxy]+\bm{\beta_7}'\EE[\bxy S]\right\}\\
-  &=\beta_0+\bm{\beta_1}\EE[\bxy]+\beta_2\EE[R]+\bm{\beta_5}'\EE[\bxyR]+\EE[Z]\left\{\beta_3+\bm{\beta_4}\EE[R]+\bm{\beta_6}'\EE[\bxy]+\bm{\beta_7}'\EE[\bxy R]\right\}\\
+  &=\beta_0+\bm{\beta_1}\EE[\bxy]+\beta_2\EE[R]+\bm{\beta_5}'\EE[\bxy R]+\EE[Z]\left\{\beta_3+\bm{\beta_4}\EE[R]+\bm{\beta_6}'\EE[\bxy]+\bm{\beta_7}'\EE[\bxy R]\right\}\\
 \end{align*}
@@
 \begin{equation*}
-
   \displaystyle\sum_i \begin{pmatrix}
     Y_i-\beta_0-\bm{\beta_1}\bxy_i-\beta_2R_i-\bm{\beta_5}'\bxy_iR_i-Z_i\left\{\beta_3-\bm{\beta_4}R_i-\bm{\beta_6}'\bxy_i-\bm{\beta_7}'\bxy_i R_i\right\}\\
-    R_iY_i-\beta_0R_i-\bm{\beta_1}\bxy_iR_i-\beta_2R_i^2-\bm{\beta_5}'\bxy_iR_i^2-Z_i\left\{\beta_3R_i-\bm{\beta_4}R_i^2-\bm{\beta_6}'\bxy_iR_i-\bm{\beta_7}'\bxy_i R_i^2\right\}\\
-    \bxy_i Y_i-\beta_0\bxy_i-\bm{\beta_1}\bxy_i\bxy_i'-\beta_2R_i\bxy_i-\bm{\beta_5}'\bxy_iR_i-Z_i\left\{\beta_3\bxy_i-\bm{\beta_4}R_i\bxy_i-\bm{\beta_6}'\bxy_i\bxy_i'-\bm{\beta_7}'\bxy_i\bxy_i' R_i\right\}\\
+    R_iY_i-\beta_0R_i-\bm{\beta_1}\bxy_i R_i-\beta_2R_i^2-\bm{\beta_5}'\bxy_iR_i^2-Z_i\left\{\beta_3R_i-\bm{\beta_4}R_i^2-\bm{\beta_6}'\bxy_iR_i-\bm{\beta_7}'\bxy_i R_i^2\right\}\\
+    \bxy_i Y_i-\beta_0\bxy_i-\bm{\beta_1}\bxy_i\bx^{Y'}_i-\beta_2R_i\bxy_i-\bm{\beta_5}'\bxy_iR_i-Z_i\left\{\beta_3\bxy_i-\bm{\beta_4}R_i\bxy_i-\bm{\beta_6}'\bxy_i\bx^{Y'}_i-\bm{\beta_7}'\bxy_i\bx^{Y'}_i R_i\right\}\\
     Z_iY_i-Z_i\left\{\beta_0-\bm{\beta_1}\bxy_i-\beta_2R_i-\bm{\beta_5}'\bxy_iR_i-\beta_3-\bm{\beta_4}R_i-\bm{\beta_6}'\bxy_i-\bm{\beta_7}'\bxy_i R_i\right\}\\
     Z_iR_iY_i-Z_iR_i\left\{\beta_0-\bm{\beta_1}\bxy_i-\beta_2R_i-\bm{\beta_5}'\bxy_iR_i-\beta_3-\bm{\beta_4}R_i-\bm{\beta_6}'\bxy_i-\bm{\beta_7}'\bxy_i R_i\right\}\\
-    Z_iY_i\bxy_i-Z_i\left\{\beta_0\bxy_i-\bm{\beta_1}\bxy_i\bxy_i'-\beta_2R_i\bxy_i-\bm{\beta_5}'\bxy_i\bxy_i'R_i-\beta_3\bxy_i-\bm{\beta_4}R_i\bxy_i-\bm{\beta_6}'\bxy_i\bxy_i'-\bm{\beta_7}'\bxy_i\bxy_i' R_i\right\}\\
-    R_iZ_iY_i\bxy_i-Z_iR_i\left\{\beta_0\bxy_i-\bm{\beta_1}\bxy_i\bxy_i'-\beta_2R_i\bxy_i-\bm{\beta_5}'\bxy_i\bxy_i'R_i-\beta_3\bxy_i-\bm{\beta_4}R_i\bxy_i-\bm{\beta_6}'\bxy_i\bxy_i'-\bm{\beta_7}'\bxy_i\bxy_i' R_i\right\}
+    Z_iY_i\bxy_i-Z_i\left\{\beta_0\bxy_i-\bm{\beta_1}\bxy_i\bx^{Y'}_i-\beta_2R_i\bxy_i-\bm{\beta_5}'\bxy_i\bx^{Y'}_iR_i-\beta_3\bxy_i-\bm{\beta_4}R_i\bxy_i-\bm{\beta_6}'\bxy_i\bx^{Y'}_i-\bm{\beta_7}'\bxy_i\bx^{Y'}_i R_i\right\}\\
+    R_iZ_iY_i\bxy_i-Z_iR_i\left\{\beta_0\bxy_i-\bm{\beta_1}\bxy_i\bx^{Y'}_i-\beta_2R_i\bxy_i-\bm{\beta_5}'\bxy_i\bx^{Y'}_iR_i-\beta_3\bxy_i-\bm{\beta_4}R_i\bxy_i-\bm{\beta_6}'\bxy_i\bx^{Y'}_i-\bm{\beta_7}'\bxy_i\bx^{Y'}_i R_i\right\}
   \end{pmatrix}=\begin{pmatrix} 0\\0\\0\\0\\0\\0\\0\end{pmatrix}
@@
 \begin{equation*}
-  \EE[Y_T-Y_C|\st=0]=\beta_0^T-\beta_0^C+(\bm{\beta_1}^T'-\bm{\beta_1}^C')\EE[\xby|S=0]=\beta_3+\beta_6'E[\xby|S=0]
+  \EE[Y_T-Y_C|\st=0]=\beta_0^T-\beta_0^C+(\bm{\beta_1}^{T'}-\bm{\beta_1}^{C'})\EE[\bxy|S=0]=\beta_3+\beta_6'E[\bxy|S=0]
 \end{equation*}
@@
 \begin{equation*}
-  \EE[Y_T-Y_C|\st=1]=\beta_0^T+\beta_2^T-(\beta_0^C+\beta_2^C)+(\bm{\beta_1}^T'+\bm{\beta_3^T}'-\bm{\beta_1}^C'-\bm{\beta_3^C}')\EE[\xby|S=0]=\beta_3+\beta_4+(\beta_6'+\beta_7')E[\xby|S=1]
+  \EE[Y_T-Y_C|\st=1]=\beta_0^T+\beta_2^T-(\beta_0^C+\beta_2^C)+(\bm{\beta_1}^{T'}+\bm{\beta_3^{Ti}}-\bm{\beta_1}^{C'}-\bm{\beta_3^{C'}})\EE[\bxy|S=0]=\beta_3+\beta_4+(\beta_6'+\beta_7')E[\bxy|S=1]
 \end{equation*}

### 2022-06-28 code/simulation/displaySim.r

@@ -12,53 +12,182 @@ load('simResults/resultsNs.RData')
 load('simResults/resultsB1s.RData')
 
 
-### rhats
-rhats <- results%>%
-  filter(eff=='1',estimator=='bayes')%>%
-  group_by(n,mu01,errDist,b1,intS,intZ)%>%
-    summarize(Rhat1.1=mean(rhat<1.1,na.rm=TRUE),Rhat1.01=mean(rhat<1.01,na.rm=TRUE))%>%arrange(Rhat1.1)
-
-hist(rhats$Rhat1.1)
-hist(rhats$Rhat1.01)
-
-lm(Rhat1.1~I(n/500)+mu01+errDist+b1+intS+intZ,data=rhats)
-
-##### check some stuff out wrt
-
-### how does AUC very w b1 and n?
-results%>%ggplot(aes(as.factor(b1),auc))+geom_jitter(alpha=0.2)+geom_violin(draw_quantiles = 0.5)
-
-results%>%ggplot(aes(as.factor(n),auc))+geom_jitter(alpha=0.2)+geom_violin(draw_quantiles = 0.5)+facet_wrap(~b1,nrow=1)
-
-
-resultsB1s%>%
+###############################################################################
+### results across ns
+###############################################################################
+ pdns <- resultsNs %>%
   group_by(b1)%>%
-  mutate(meanAUC=mean(auc))%>%
-  ggplot(aes(as.factor(b1),auc))+geom_boxplot()+geom_point(aes(y=meanAUC))+geom_smooth(se=FALSE)
-                                        #  scale_y_continuous('Avg. AUC',seq(.5,1,.1))
-resultsB1s%>%
-  group_by(b1)%>%
-  summarize(meanAUC=mean(auc))
-
-resultsB1s%>%
-  group_by(b1)%>%
-  summarize(meanAUC=mean(auc))%>%ggplot(aes(b1,meanAUC))+geom_point()+geom_smooth(method='lm')
-
-
-
-#### how do estimates vary with auc (simple case)?
-results%>%
-  filter(n==500,errDist=='norm',!intS,!intZ,eff=='1')%>%
-  group_by(b1)%>%
-  mutate(AUC=round(mean(auc,na.rm=TRUE),1))%>%
+  mutate(AUCf=mean(auc,na.rm=TRUE))%>%
   ungroup()%>%
-  ggplot(aes(auc,est))+
[…]

### 2022-06-28 code/simulation/readSimFuncs.r
tweaking bp function

@@ -109,7 +109,7 @@ loadRes <- function(ext1='',ext2='',pswResults){
     reduce(results,bind_rows)
 }
 
-bp <- function(pd,subset,facet,title=deparse(substitute(subset)),ylim,Labeller="label_value"){
+bp <- function(pd,subset,facet,title=deparse(substitute(subset)),ylim,Labeller="label_value",labSize=5){
 
 
   r <- if (missing(subset))
@@ -126,20 +126,6 @@ bp <- function(pd,subset,facet,title=deparse(substitute(subset)),ylim,Labeller="
 
   if(missing(ylim)) ylim = quantile(pd$errP, c(0.01, 0.99))
 
-  if(any(pd$errP<ylim[1]|pd$errP>ylim[2])){
-    outliers=TRUE
-    grp=c(as.character(unlist(as.list(facet)[-1])),'estimator')
-    outDat=pd%>%
-      group_by(across(!!grp))%>%
-      summarize(
-        nBig=sum(errP>ylim[2]),
-        nSmall=sum(errP<ylim[1]))%>%
-      pivot_longer(c(nBig,nSmall),names_to="bs",values_to="out")%>%
-      mutate(
-        y=ifelse(bs=='nBig',ylim[2],ylim[1]),
-        lab=ifelse(out>0,paste0('+',out),''))%>%
-      ungroup()
-  } else outliers=FALSE
 
   p <- ggplot(pd,
          aes(
@@ -157,11 +143,33 @@ bp <- function(pd,subset,facet,title=deparse(substitute(subset)),ylim,Labeller="
     coord_cartesian(ylim =ylim)+
     facet_grid(facet , scales = "free",
                labeller = Labeller)+
-    labs(title = title,
-         x = NULL, y = 'Estimation Error') + theme(legend.pos = 'none')
-  if(outliers) p=p+geom_label(data=filter(outDat,out>0),
-                              inherit.aes=FALSE,
-                              mapping=aes(estimator,y,label=lab))
+    ggtitle(title)+
+                                        #labs(title = title,
+     #    x = NULL, y = 'Estimation Error') +
+    theme(legend.pos = 'none')
+
+  ## if(any(pd$errP<ylim[1]|pd$errP>ylim[2])){
+  ##   outliers=TRUE
+  ##   grp=c(as.character(unlist(as.list(facet)[-1])),'estimator')
[…]

### 2022-06-28 code/simulation/readSim.r
making simulation results files

@@ -7,7 +7,37 @@ source('code/simulation/readSimFuncs.r')
 print(load('simResults/pswResults.RData'))
 
 results=loadRes(pswResults=pswResults)
+save(results,file='simResults/fullResults.RData')
 
 resultsN100=loadRes(ext2='n100')
+save(resultsN100,file='simResults/resultsN100.RData')
+
 
 resultsNs=loadRes(ext2='ns')
+
+resultsB1s=loadRes(ext2='b1s')
+
+### merge results by n
+load('simResults/casesns.RData')
+casesNs=cases
+load('simResults/casesn100.RData')
+casesN100=cases
+load('simResults/cases.RData')
+
+resultsNs=bind_rows(
+  resultsNs,
+  filter(resultsN100,run%in%which(apply(casesN100[,-1],1,function(x) all(x==casesNs[1,-1])))),
+  filter(results,run%in%which(apply(cases[,-1],1,function(x) all(x==casesNs[1,-1])))))
+
+save(resultsNs,file='simResults/resultsNs.RData')
+
+### merge results by B1s
+casesTot=cases
+load('simResults/casesb1s.RData')
+casesB1=cases
+
+resultsB1s=bind_rows(
+  resultsB1s,
+  filter(results,run%in%(left_join(casesB1[1,],casesTot%>%mutate(run=1:n())%>%select(-b1))%>%pull(run)))
+  )
+save(resultsB1s,file='simResults/resultsB1s')

### 2022-08-03 code/fh2t/analysisGEE.r
considering alternatives

@@ -1,13 +1,15 @@
 library(tidyverse)
 library(arm)
+library(randomForest)
 source('code/regression.r')
 select <- dplyr::select
 
+load('data/psdatBAU.RData')
 load('data/psdat.RData')
 
 mod4imp <- glm(
   anyBottom ~ pretestIMP+ Scale.Score5IMP+ MALEIMP+ raceIMP+ virtualIMP+ EIPIMP+
-  IEPIMP+ ESOLIMP+ GIFTEDIMP+ pre.avg_time_on_tasksIMP+
+  IEPIMP+ ESOLIMP+ GIFTEDIMP+ log(pre.avg_time_on_tasksIMP)+
   pre_MA_total_scoreIMP+ pre_negative_reaction_scoreIMP+ pre_numerical_confindence_scoreIMP,
   data=subset(dat3,Z=='ASSISTments'),family=binomial)
 
@@ -18,20 +20,63 @@ binnedplot(mod4imp$fitted.values,resid(mod4imp,type='response'))
 
 
 
-estimate0 <- est(psdat,covFormU = formula(mod4imp)[-2])
+estimate0 <- est(psdatBAU,covFormU = formula(mod4imp)[-2])
                 # covFormY=
                 #   update(formula(mod4imp)[-2],
                 #          .~.-virtualIMP),block = "class")
 plot(estimate0$outMod,which=1)
 effsFromFit(estimate0)
 
-estimate1 <- est(psdat,covFormU = formula(mod4imp)[-2],
+estimate1 <- est(psdatBAU,covFormU = formula(mod4imp)[-2],
                 covFormY=
                   update(formula(mod4imp)[-2],
                          .~.-virtualIMP),block = "class")
 plot(estimate1$outMod,which=1)
 effsFromFit(estimate1)
 
+
+### robustness checks
+library(lspline)
+### 3-df linear splines on pretest
[…]

### 2022-08-03 code/fh2t/analysis_stanModels.R

@@ -14,218 +14,89 @@ options(mc.cores = 4)
 gc()
 
 load('data/psdat.RData')
-write.csv(psdat, "data/psdat.csv")
-psdat<- read.csv("data/psdat.csv")
-colnames(psdat)
-table(psdat$Z,
-      psdat$rdm_condition)
+## write.csv(psdat, "data/psdat.csv")
+## psdat<- read.csv("data/psdat.csv")
+## colnames(psdat)
+## table(psdat$Z,
+##       psdat$rdm_condition)
 
 # U --> Model predicting Bottom Outers
 # Y --> Model predicting Bottom Outcomes
 
-### remove classes that don't have representation across both scripts
-classes_by_z <- data.frame(table(psdat$class, psdat$Z)) %>%
-  spread(Var2, Freq) %>%
-  rename(class = Var1) %>%
-  filter(
-    `0` > 0 &
-    `1` > 0) %>%
-  dplyr::select(-`1`,
-         -`0`)
 
-psdat <- psdat %>%
-  inner_join(
-    classes_by_z
-  )
+xtabs(~class+Z,psdat)%>%as.data.frame()%>%summarize(sum(Freq==0))
+length(unique(psdat$class))
+
+psdat=psdat%>%
+    group_by(class)%>%
+    mutate(nz=n_distinct(Z))%>%
+    ungroup()%>%
+    filter(nz>1)
 
 length(unique(psdat$class))
 
 table(psdat$raceIMP, psdat$Z)
 
-### Data Prep ####
-###### Dummy Codes ####
-#  dummy codes for race
-race_dummy <- fastDummies::dummy_cols(psdat$raceIMP,
-                                      remove_first_dummy = T) %>%
-  distinct() %>%
-  rename(raceIMP = ".data")
-names(dummy)[-1]<-sub(".data*", "", names(dummy)[-1])
+formU <-    ~ pretestIMP+ Scale.Score5IMP+ MALEIMP+ raceIMP+ virtualIMP+ EIPIMP+
+  IEPIMP+ ESOLIMP+ GIFTEDIMP+ pre.avg_time_on_tasksIMP+
+  pre_MA_total_scoreIMP+ pre_negative_reaction_scoreIMP+ pre_numerical_confindence_scoreIMP
 
-#  dummy codes for class
-class_dummy<-fastDummies::dummy_cols(psdat$class,
[…]

### 2022-08-03 code/fh2t/data.r

@@ -6,7 +6,7 @@ select <- dplyr::select
 
 Scale <- function(x) (x-mean(x,na.rm=TRUE))/sd(x,na.rm=TRUE)
 
-dat <- read_csv('../../fh2t/data/data.csv')
+dat <- read_csv('data/data.csv')
 dat$hints <- dat%>%select(contains("hint_count"))%>%rowSums(na.rm=TRUE)
 dat$anyHint <- dat$hints>0
 
@@ -14,8 +14,8 @@ dat$bottom <- dat%>%select(contains("bottom"))%>%rowSums(na.rm=TRUE)
 dat$anyBottom <- dat$bottom>0
 
 
-full <- read_csv('../../fh2t/data/Assessment_merged_2022_01_19_N=4,343 - Sheet1.csv')
-older <- read_csv('../../fh2t/data/Assessment_merged_2021_07_16_state_assessment_N=4321 - Sheet1.csv')
+full <- read_csv('data/Assessment_merged_2022_01_19_N=4,343 - Sheet1.csv')
+older <- read_csv('data/Assessment_merged_2021_07_16_state_assessment_N=4321 - Sheet1.csv')
 
 full <- older%>%
   select(student_number,pre.avg_time_on_tasks,pre_MA_total_score,pre_negative_reaction_score,pre_numerical_confindence_score)%>%
@@ -47,7 +47,7 @@ dat0 <- dat0%>%
   )%>%
   filter(!is.na(Z))
 
-teachDrop <- read_csv('../../fh2t/data/IES_school_teacher_ID_list_opt_out - teacher (1).csv')
+teachDrop <- read_csv('data/IES_school_teacher_ID_list_opt_out - teacher (1).csv')
 
 dat0$schoolSupp <- dat0$initial_school_id
 dat0$schoolSupp[dat0$initial_teacher_id%in%teachDrop$teacher_id[teachDrop$note=='S03']] <- 'S03'
@@ -176,4 +176,13 @@ psdat <- dat3%>%
          S=as.numeric(anyBottom))
 
 save(psdat,dat3,file="data/psdat.RData")
+##
+
+psdatBAU <- dat3%>%
+  filter(Z%in%c('ASSISTments','BAU'))%>%
+  mutate(Z=ifelse(Z=='ASSISTments',1,0),
+         Y=ScaleScore7,
+         S=as.numeric(anyBottom))
+
+save(psdatBAU,dat3,file="data/psdatBAU.RData")
 ##

### 2022-08-03 code/fh2t/overallEffects.r

@@ -1,0 +1,29 @@
+library(tidyverse)
+library(arm)
+library(sandwich)
+library(loop.estimator)
+source('code/regression.r')
+select <- dplyr::select
+
+load('data/psdat.RData')
+
+### posttest
+dat3$z=ifelse(dat3$Z=='ASSISTments',1,0)
+### contrast assistments with each of the other conditions
+postMods=
+    lapply(c('BAU','Dragon','FH2T')%>%setNames(.,.),function(alt)
+           lm(  post.total_math_score ~ z + class+pretestIMP+ Scale.Score5IMP+ MALEIMP+ raceIMP+  EIPIMP+
+  IEPIMP+ ESOLIMP+ GIFTEDIMP+ pre.avg_time_on_tasksIMP+
+  pre_MA_total_scoreIMP+ pre_negative_reaction_scoreIMP+ pre_numerical_confindence_scoreIMP,
+  data=dat3,subset=Z=='ASSISTments'|Z==alt))
+
+postMods$pooledTrt=update(postMods$BAU,subset=Z!='BAU')
+
+vcvs=lapply(postMods,vcovHC,type='HC1')
+
+effs=sapply(postMods,function(x) coef(x)['z'])
+ses=sqrt(sapply(vcvs,function(x) x['z','z']))
+Ts=effs/ses
+pvals=2*pnorm(-abs(Ts))
+
+cbind(effs,ses,Ts,pvals)

### 2022-08-03 writeUps/paper.tex
fixing git stuff

@@
 \section{Introduction}
-
-In a recent educational field experiment, middle school students completing their math schoolwork on a computer program were randomized between two conditions.
-In one condition, students were instructed to work on a fixed number of math problems, and later informed of their scores.
-In a second condition, students received immediate, informative error messages after entering a wrong answer, and were required to enter a correct answer in order to proceed.
-Students in the second treatment arm also had access to a series of hints for each problem, the last of which contained the correct answer.
-Students who ``bottomed out'' on a problem, i.e.
-requested all available hints on a problem, and were told the correct answer, did not receive credit for that problem.
-Students remained in their randomly-assigned condition over the course of a school year, and were assessed with a 10-item posttest administered within the computer program, as well with state standardized tests.
-
-The literature on hints in online learning is mixed \citep{stuff}, in particular when it comes to bottoming out.
-On the one hand, hints that include the correct answer are essentially worked examples, which can be beneficial to learning \citep{worked example stuff}.
-On the other hand, they allow students to ``game the system,'' by simply requesting enough hints to see the right answer, and use the software without ever working to solve any problems.
-However, in this experiment, the ability to bottom out was only one of several differnces between the treatment conditions.
-One way to disentangle the role of bottoming out in any treatment effect would be to estimate average treatment effects separately for subjects who would (if assigned to treatment) bottom out, and for subjects who would not.
-If hints (that do not supply the correct answer) and error messages are helpful, but the ability to bottom out is harmful to learning, then we might expect to estimate a positive effect of being assigned to the feedback condition for students who do not bottom out, but a negative effect for students who do.
-If hints are helpful for learning only if they can be read as worked examples, then we might expect no effect of condition on students who do not bottom out, but a positive effect for students who do.
-(These hypothetical results may be masked by other treatment effect heterogeneity between students who would or wouldn't bottom out, but the results could still be informative.)
 
@@
 
-Analysis models had access to treatment assignment $Z$, outcomes $Y$,
-principal stratum $S$ for members of the treatment group, and
-covariates.
-Even though three covariates, $x_1$, $x_2$ and $x_3$ were included as
-predictors of principal stratum $S$ and outcome $Y$ in the data
-generating model, both analysis approaches only had access to the
-first two covariates, $x_1$ and $x_2$.
-Hence, principal ignorability was violated---$S$ and $Y$ were
-``confounted'' by unobserved variable $x_3$.
+\section{Application: Estimating Treatment Effects for Implementation Subgroups}
 
-Simulated data were analyzed first with a ``stacked equations''/M-Estimation/GEE
-approach, using the \texttt{geex} package in \texttt{R}
-\citep{geex,rcite}, and then with a more conventional maximum
-likelihood approach via \texttt{rstan} \citep{rstan}.
+In a recent educational field experiment, middle school students completing their math schoolwork on a computer program were randomized between two conditions.
+In one condition (which we arbitrarily label ``treatment''), students were instructed to work on a fixed number of math problems, and later informed of their scores.
+In a second condition, students received immediate, informative error messages after entering a wrong answer, and were required to enter a correct answer in order to proceed.
+Students in the second treatment arm (labeled ``control'') also had access to a series of hints for each problem, the last of which contained the correct answer.
+Students who ``bottomed out'' on a problem, i.e.
+requested all available hints on a problem, and were told the correct answer, did not receive credit for that problem.
+Students remained in their randomly-assigned condition over the course of a school year, and were assessed with a 10-item posttest administered within the computer program, as well with state standardized tests.
 
-Both methods assumed (correctly) that there were no interactions
-between covariates $x_1$ and $x_2$ and treatment assignment (after
-accounting for $S$).
+The literature on hints in online learning is mixed \citep{stuff}, in particular when it comes to bottoming out.
+On the one hand, hints that include the correct answer are essentially worked examples, which can be beneficial to learning \citep{worked example stuff}.
+On the other hand, they allow students to ``game the system,'' by simply requesting enough hints to see the right answer, and use the software without ever working to solve any problems.
+However, in this experiment, the ability to bottom out was only one of several differnces between the treatment conditions.
+One way to disentangle the role of bottoming out in any treatment effect would be to estimate average treatment effects separately for subjects who would (if assigned to treatment) bottom out, and for subjects who would not.
+If hints (that do not supply the correct answer) and error messages are helpful, but the ability to bottom out is harmful to learning, then we might expect to estimate a positive effect of being assigned to the feedback condition for students who do not bottom out, but a negative effect for students who do.
+If hints are helpful for learning only if they can be read as worked examples, then we might expect no effect of condition on students who do not bottom out, but a positive effect for students who do.
+(These hypothetical results may be masked by other treatment effect heterogeneity between students who would or wouldn't bottom out, but the results could still be informative.)
 
-In the GEE approach included a set of nine estimating equations.
-First, a GEE logistic regression fit to the treatment group to
-estimate principal scores
-$\ppi=logit^{-1}(\beta_0+\beta_1x_{1i}+\beta_2x_{2i})$ (three parameters):
-\begin{align*}
-  \sum_iZ_i\left[S_i-\ppi\right]&=0\\
-  \sum_iZ_ix_{1i}\left[S_i-\ppi\right]&=0\\
-  \sum_iZ_ix_{2i}\left[S_i-\ppi\right]&=0
-\end{align*}
-Next, four equations to estimate the outcome regression in the
-treatment group, including terms for $S$:
-\begin{align*}
-  \sum_i
-  Z_i\left[Y_i-(\gamma_1x_{1i}+\gamma_2x_{2i}+S_i\mu_{11}+(1-S_i)\mu_{10})\right]&=0\\
-  \sum_i
-  Z_ix_{1i}\left[Y_i-(\gamma_1x_{1i}+\gamma_2x_{2i}+S_i\mu_{11}+(1-S_i)\mu_{10})\right]&=0\\
-  \sum_i
-  Z_ix_{2i}\left[Y_i-(\gamma_1x_{1i}+\gamma_2x_{2i}+S_i\mu_{11}+(1-S_i)\mu_{10})\right]&=0\\
-  \sum_i
-  Z_iS_i\left[Y_i-(\gamma_1x_{1i}+\gamma_2x_{2i}+S_i\mu_{11}+(1-S_i)\mu_{10})\right]&=0
-\end{align*}
+For the purpose of demonstrating GEEPERS, we gathered dichotomized data on which students in the treatment condition requested at least one bottom-out hint.
+That is, $\st=1$ for students who, if assigned to the treatment condition, would request at least one bottom-out hint, and $\st=0$ for students who would not.
+The outcome of interest $Y$ in our example is students' total scores on a ten-item posttest completed within the online tutoring system; $Y$ is an integer ranging from 0 to 10.
+The goal of our analysis will be to determine the principal effects $\tau_1=\EE[\yt-\yc|\st=0]$ and $\tau_0=\EE[\yt-\yc|\st=0]$, the average effects of assignment to the treatment condition for bottom-outers and non-bottom-outers.
+We also gathered data on baseline covariates: total scores on a 10-item pretest, fifth-grade stanardized test scores, indicator variables for if students have an early intervention program (EIP)
 
-Finally, two equations to estimate the mixture model in the control
-group, for whom $S$ is unobserved:
-\begin{align*}
-  \sum_i
-  (1-Z_i)\left[Y_i-(\gamma_1x_{1i}+\gamma_2x_{2i}+\ppi\mu_{01}+(1-\ppi)\mu_{00})\right]&=0\\
-  \sum_i
-  (1-Z_i)\left[Y_i-(\gamma_1x_{1i}+\gamma_2x_{2i}+\ppi^2\mu_{01}+(\ppi-\ppi^2)\mu_{00})\right]&=0
-\end{align*}
-The average treatment effect for $S=0$ was $\mu_{10}-\mu_{00}$ and for
-$S=1$ was $\mu_{11}-\mu_{01}$.
+or individualized educational plan (IEP) in place, if students are learning English as a second language (ESOL)students'
 
-Next, we estimated effects with MLE, using the built-in optimizer of
-\texttt{rstan}.
+empress
+Mperes
 
-In this model we regressed $S$ and $Y$ for the treatment group on an
-intercept, $x_1$, and $x_2$ using a logistic and normal linear
-specification, respectively.
-Then, we specified a two-component mixture model for control
-outcomes.
+M-estimation of Principal
 
-\section{Data Analysis}
+peseer per
+
+M estimation regression principal effects
+
+M
 
@@
     Z_i\ri Y_i -Z_i\ri\mut0-Z_i\ri^2(\mut1-\mut0)
+
 \end{pmatrix}=\begin{pmatrix} 0\\0\\0\\0\end{pmatrix}

### 2022-08-03 code/fh2t/analysis_stanModels.R

@@ -192,11 +192,12 @@ traceplot(
 #
 # write.csv(checkdat, "FakeDataForCheck.csv")
 checkdat <- read.csv("data/FakeDataForCheck.csv")
-md.pattern(checkdat)
+#md.pattern(checkdat)
 ### Data Prep ####
 
 XU=makeCovMat(formU,data=checkdat)
 XY=makeCovMat(formY,data=checkdat)
+
 
 XctlU=XU[checkdat$Z==0,]
 XtrtU=XU[checkdat$Z==1,]
@@ -207,6 +208,159 @@ XtrtY=XY[checkdat$Z==1,]
 
 stanDat <-  list(nc= sum(1-checkdat$Z),#length(checkdat[checkdat$Z == 0,]$student_number), #
                  nt= sum(checkdat$Z),#length(psdat[psdat$Z == 1,]$student_number), #
+
+#### Covariates matrixes #######
+# Control Cov for U model
+XctlU <- checkdat %>%
+  dplyr::filter(Z == 0) %>%
+  dplyr::select(
+    pretestIMP,
+    Scale.Score5IMP,
+    MALEIMP,
+    raceIMP,
+    virtualIMP,
+    EIPIMP,
+    IEPIMP,
+    ESOLIMP,
+    GIFTEDIMP,
+    pre.avg_time_on_tasksIMP,
+    pre_MA_total_scoreIMP,
+    pre_negative_reaction_scoreIMP,
+    pre_numerical_confindence_scoreIMP
+  ) %>%
+  mutate(
+    pretestIMP = scale(pretestIMP),
+    Scale.Score5IMP = scale(Scale.Score5IMP),
+    pre.avg_time_on_tasksIMP = scale(pre.avg_time_on_tasksIMP),
+    pre_MA_total_scoreIMP = scale(pre_MA_total_scoreIMP),
+    pre_negative_reaction_scoreIMP = scale(pre_negative_reaction_scoreIMP),
+    pre_numerical_confindence_scoreIMP = scale(pre_numerical_confindence_scoreIMP)
+  ) %>%
+  left_join(
+    race_dummy,
+    by = "raceIMP"
+  ) %>%
+  dplyr::select(-raceIMP)
+colnames(XctlU)
+##md.pattern(XctlU)
+
+# Control Cov for Y model
+XctlY <- checkdat %>%
+  filter(Z == 0) %>%
+  dplyr::select(
+    pretestIMP,
+    Scale.Score5IMP,
+    MALEIMP,
[…]

### 2022-08-05 code/fh2t/tables.r
tables.r and gitignore

@@ -1,0 +1,93 @@
+library(tidyverse)
+library(arm)
+library(randomForest)
+library(tableone)
+library(kableExtra)
+library(xtable)
+source('code/regression.r')
+select <- dplyr::select
+
+options(knitr.kable.NA = '')
+
+load('data/psdatBAU.RData')
+load('data/psdat.RData')
+load('data/imputation.RData')
+
+tab1dat <-
+  dat3%>%
+  filter(Z%in%c('ASSISTments','BAU'))%>%
+  select(
+    pretest , Scale.Score5 , MALE , race ,
+    virtual , EIP , IEP , ESOL , GIFTED , pre.avg_time_on_tasks ,
+    pre_MA_total_score , pre_negative_reaction_score ,
+    pre_numerical_confindence_score,anyBottom,post.total_math_score,Z)%>%
+  mutate(pre.avg_time_on_tasks=log(pre.avg_time_on_tasks),
+         MALE=ifelse(MALE==1,'Male','Female'),
+         virtual=ifelse(virtual==1,'Remote','In-Person'),
+         anyBottom=anyBottom*1)
+
+
+oob=as.data.frame(rbind(imp$OOBerror))%>%  select(
+    `Has EIP`=EIP,
+    `Has IEP`=IEP,
+    ESOL,
+    `Gifted`=GIFTED,
+        Pretest=pretest,
+    `Grade 5 Stand. Test`=Scale.Score5,
+    Gender=MALE,
+    `Race/Ethnicity`=race,
+    `log(Pretest ToT)`=pre.avg_time_on_tasks,
+    `Math Anxiety`=pre_MA_total_score,
+    `Math Neg. React.`=pre_negative_reaction_score,
+    `Numerical Conf.`=pre_numerical_confindence_score)%>%t()
+
+
+
+tab1bin <- tab1dat%>%
+  select(
+    `Has EIP (n (%))`=EIP,
+    `Has IEP (n (%))`=IEP,
+    `ESOL (n (%))`=ESOL,
+    `Gifted (n (%))`=GIFTED,
+    `Bottom Out (n (%))`=anyBottom,Z)%>%
+  mutate(across(.fns=as.factor))%>%
+  CreateTableOne(vars=setdiff(names(.),'Z'),data=.,strata="Z")%>%
+  print(missing=TRUE,test=FALSE,dropEqual=TRUE,explain=FALSE)%>%
[…]

### 2022-08-05 code/fh2t/analysisGEE.r
cleaning up gee analysis A LITTLE

@@ -7,6 +7,10 @@ select <- dplyr::select
 load('data/psdatBAU.RData')
 load('data/psdat.RData')
 
+if(!exists('alt')) alt <- 'bau'
+
+dat <- if(alt=='bau') psdatBAU else psdat
+
 mod4imp <- glm(
   anyBottom ~ pretestIMP+ Scale.Score5IMP+ MALEIMP+ raceIMP+ virtualIMP+ EIPIMP+
   IEPIMP+ ESOLIMP+ GIFTEDIMP+ log(pre.avg_time_on_tasksIMP)+
@@ -20,14 +24,14 @@ binnedplot(mod4imp$fitted.values,resid(mod4imp,type='response'))
 
 
 
-estimate0 <- est(psdatBAU,covFormU = formula(mod4imp)[-2])
+estimate0 <- est(dat,covFormU = formula(mod4imp)[-2])
                 # covFormY=
                 #   update(formula(mod4imp)[-2],
                 #          .~.-virtualIMP),block = "class")
 plot(estimate0$outMod,which=1)
 effsFromFit(estimate0)
 
-estimate1 <- est(psdatBAU,covFormU = formula(mod4imp)[-2],
+estimate1 <- est(dat,covFormU = formula(mod4imp)[-2],
                 covFormY=
                   update(formula(mod4imp)[-2],
                          .~.-virtualIMP),block = "class")
@@ -40,8 +44,8 @@ probit=glm(
   S ~ pretestIMP+ Scale.Score5IMP+ MALEIMP+ raceIMP+ virtualIMP+ EIPIMP+
   IEPIMP+ ESOLIMP+ GIFTEDIMP+ log(pre.avg_time_on_tasksIMP)+
   pre_MA_total_scoreIMP+ pre_negative_reaction_scoreIMP+ pre_numerical_confindence_scoreIMP,
-  data=subset(psdatBAU,Z==1),family=binomial('probit'))
-estimate1.1 <- est(psdatBAU, psMod=probit,
+  data=subset(dat,Z==1),family=binomial('probit'))
+estimate1.1 <- est(dat, psMod=probit,
                 covFormY=
                   update(formula(mod4imp)[-2],
                          .~.-virtualIMP),block = "class")
@@ -52,8 +56,8 @@ bayes=bayesglm(
   S ~ pretestIMP+ Scale.Score5IMP+ MALEIMP+ raceIMP+ virtualIMP+ EIPIMP+
[…]

### 2022-08-05 writeUps/paper.tex
updates

@@
 \section{Introduction}
+
+We refer to the estimators presented here as ``GEEPERs'': generalized estimating equations for principal effects using regressions.
 
@@
 The goal of our analysis will be to determine the principal effects $\tau_1=\EE[\yt-\yc|\st=0]$ and $\tau_0=\EE[\yt-\yc|\st=0]$, the average effects of assignment to the treatment condition for bottom-outers and non-bottom-outers.
-We also gathered data on baseline covariates: total scores on a 10-item pretest, fifth-grade stanardized test scores, indicator variables for if students have an early intervention program (EIP)
+We also gathered data on baseline covariates: total scores on a 10-item pretest, fifth-grade stanardized test scores, race/ethnicity (white, Asian, Hispanic/Latino, or Other), indicator variables for if students have an early intervention program (EIP) or individualized educational plan (IEP) in place, if students are learning English as a second language (ESOL), students' average time on task for the pretest (logged), total scores from baseline evaluations of students' math anxiety, negative reaction to math, and numerical confidence, and an indicator for weather students began the school year in remote or in-person instruction.
+Students with missing pretest scores were dropped from the analysis; missing data in other covariates was imputed with a Random Forest algorithm \citep{missForest}.
+Table \ref{tab:tab1} describes each covariate--as well as the sucess of the randomForest imputation algorithm.
 
-or individualized educational plan (IEP) in place, if students are learning English as a second language (ESOL)students'
+\begin{table}
+  \centering
+  \input{table1}
+  \caption{Descriptive statistics on baseline covariates, $S$, and $Y$ for the analysis data.
+Values are mean (SD) unless otherwise noted.
+Imputation error is the ``out of box'' error estimate from the Random Forest imputation algorithm---root mean squared error divided by the observed standard deviation for continuous variables and misclassification proportion for categorical variables.}
+  \label{tab:tab1}
+\end{table}
 
-empress
-Mperes
-
-M-estimation of Principal
-
-peseer per
-
-M estimation regression principal effects
-
-M
+The data analysis we present here is intended as a demonstration of GEEPERS, rather than for its substantive conclusions, and our description omits discussion of some important methodological considerations, including attrition.
 

### 2022-08-05 code/fh2t/data.r

@@ -172,7 +172,7 @@ binnedplot(mod4imp$fitted.values,resid(mod4imp,type='response'))
 psdat <- dat3%>%
   filter(Z%in%c('ASSISTments','FH2T'))%>%
   mutate(Z=ifelse(Z=='ASSISTments',1,0),
-         Y=ScaleScore7,
+         Y=post.total_math_score,
          S=as.numeric(anyBottom))
 
 save(psdat,dat3,file="data/psdat.RData")
@@ -181,7 +181,7 @@ save(psdat,dat3,file="data/psdat.RData")
 psdatBAU <- dat3%>%
   filter(Z%in%c('ASSISTments','BAU'))%>%
   mutate(Z=ifelse(Z=='ASSISTments',1,0),
-         Y=ScaleScore7,
+         Y=post.total_math_score,
          S=as.numeric(anyBottom))
 
 save(psdatBAU,dat3,file="data/psdatBAU.RData")

### 2022-08-05 code/fh2t/tables.r

@@ -73,7 +73,7 @@ tab1oth<- tab1dat%>%
   CreateTableOne(vars=setdiff(names(.),'Z'),data=.,strata="Z")%>%
   print(missing=TRUE,test=FALSE,showAllLevels=TRUE,explain=FALSE)%>%
   cbind(.,
-        `Imputation Error`=round(oob[pmatch(gsub("( \\(%\\))|( \\(mean \\(SD\\)\\))","",rownames(.)),rownames(oob)),],2))
+        `Imp. Err.`=round(oob[pmatch(gsub("( \\(%\\))|( \\(mean \\(SD\\)\\))","",rownames(.)),rownames(oob)),],2))
 
 
 
@@ -85,7 +85,7 @@ tab1=rbind(
 
 colnames(tab1)[4]='Miss. %'
 
-print(xtable(tab1),hline.after=17,sanitize.text.function=function(x) {x})
+#print(xtable(tab1),hline.after=17,sanitize.text.function=function(x) {x})
 
 sink('writeUps/table1.tex')
 kbl(tab1,format='latex',booktabs=TRUE)%>% #c("p{1.5in}",rep("l",5)))%>%

### 2022-08-05 code/fh2t/data.r
new data

@@ -1,190 +1,43 @@
 library(tidyverse)
 library(arm)
 library(missForest)
-source('code/regression.r')
 select <- dplyr::select
 
 Scale <- function(x) (x-mean(x,na.rm=TRUE))/sd(x,na.rm=TRUE)
 
-dat <- read_csv('data/data.csv')
-dat$hints <- dat%>%select(contains("hint_count"))%>%rowSums(na.rm=TRUE)
-dat$anyHint <- dat$hints>0
+actDat0=read_csv('data/logs_for_adam.csv')
 
-dat$bottom <- dat%>%select(contains("bottom"))%>%rowSums(na.rm=TRUE)
-dat$anyBottom <- dat$bottom>0
+dat=read_csv('data/DATA20220202_3591.csv',na=c('','NA','#NULL!'))
 
+probMeta=read_csv('data/ASSISTments-Table 1.csv')
+probMeta=select(probMeta,!starts_with('...'))
+names(probMeta)=gsub(' ','',names(probMeta))
+probMeta$Hints[probMeta$Note=='3']=3
 
-full <- read_csv('data/Assessment_merged_2022_01_19_N=4,343 - Sheet1.csv')
-older <- read_csv('data/Assessment_merged_2021_07_16_state_assessment_N=4321 - Sheet1.csv')
+bottomOut=
+  probMeta%>%
+  filter(Hints!='n/a',CorrectAnswer!='n/a')%>%
+  mutate(Hints=as.numeric(Hints),
+         Subproblem=ifelse(Subproblem%in%c('n/a','N/A'),'A',Subproblem))%>%
+  select(ProblemID,ProblemSet,ProblemNumber,ProblemOrder,Hints,Subproblem)%>%
+  inner_join(mutate(actDat0,Subproblem=LETTERS[problem_part]),
+             by=c("ProblemID"="problem_id","Subproblem"))%>%
+  select(-'...1')%>%
+  group_by(StuID)%>%
+  summarize(nbo=sum(action=='answer_hint',na.rm=TRUE),nprob=n_distinct(ProblemID))
 
-full <- older%>%
-  select(student_number,pre.avg_time_on_tasks,pre_MA_total_score,pre_negative_reaction_score,pre_numerical_confindence_score)%>%
-  right_join(full)
+dat <- left_join(dat,bottomOut)%>%
+  filter(!is.na(post.total_math_score))
 
[…]

### 2022-09-08 code/fh2t/analysisGEE.r
update analysis new data

@@ -4,40 +4,243 @@ library(randomForest)
 source('code/regression.r')
 select <- dplyr::select
 
-load('data/psdatBAU.RData')
 load('data/psdat.RData')
 
-if(!exists('alt')) alt <- 'bau'
 
-dat <- if(alt=='bau') psdatBAU else psdat
+load('results/psModGlm7.RData')
 
-mod4imp <- glm(
-  anyBottom ~ pretestIMP+ Scale.Score5IMP+ MALEIMP+ raceIMP+ virtualIMP+ EIPIMP+
-  IEPIMP+ ESOLIMP+ GIFTEDIMP+ log(pre.avg_time_on_tasksIMP)+
-  pre_MA_total_scoreIMP+ pre_negative_reaction_scoreIMP+ pre_numerical_confindence_scoreIMP,
-  data=subset(dat3,Z=='ASSISTments'),family=binomial)
+psdat <- filter(psdat,SchIDPre!=7) ### only one student from school 7, in FH2T
 
-summary(mod4imp)
-AUCmod(mod4imp)
+psdat$Z <- ifelse(psdat$trt=='ASSISTments',1,0)
 
-binnedplot(mod4imp$fitted.values,resid(mod4imp,type='response'))
+alts <- unique(psdat$trt[psdat$Z==0])
+
+### do a little "reloop"
+## remnantMods <- map(setNames(alts,alts),
+##                    ~psdat%>%
+##                    filter(!trt%in%c('ASSISTments',.x))%>%
+##                    select(-StuID,-TeaIDPre,-ClaIDPre,-S,-virtual,-Z,-trt)%>%
+##                    randomForest(Y~.,data=.))
+
+
+getDat <- function(alt){
+  dat <- psdat%>%
+    filter(trt%in%c('ASSISTments',alt))
+ # dat$pred <- predict(remnantMods[[alt]],dat)
+
+  droplevels(dat)
+}
+
+### overall ATEs
+## ateMods <- sapply(setNames(alts,alts),
+##                   function(alt)
+##                     estimatr::lm_robust(
+##                                 Y~Z+pred+
+##                                   Scale.Score5+ESOL+IEP+pre.total_time_on_tasks+pre_MSE_total_score+
+##                             fullYear5+pre_PS_tasks_total_score+raceEth+Gender+GIFTED+
+##                             splines::ns(pre.total_math_score,3)+pre.total_math_scoreNA,
[…]

### 2022-09-08 code/fh2t/data.r

@@ -12,6 +12,31 @@ kirk <- read_csv('data/data_uptated_8_18.csv')
 dat$nbo <- kirk$num_problem_parts_used_bottom_out_hint[match(dat$StuID,kirk$StuID)]
 dat <- filter(dat,!is.na(post.total_math_score))
 
+## dat <- kirk%>%
+##   select(StuID,starts_with("num_")&contains("_parts_"))%>%
+##   right_join(dat)%>%
+##   rename(nbo=num_problem_parts_used_bottom_out_hint)%>%
+##   mutate(pbo1=nbo/num_problem_parts_started,
+##          pbo2=nbo/num_problem_parts_attempted,
+##          pbo3=nbo/num_graded_problem_parts_started,
+##          pbo4=nbo/num_graded_problem_parts_attempted)%>%
+##   mutate(across(starts_with("pbo"),~ifelse(is.finite(.),.,0)))%>%
+##   filter(!is.na(post.total_math_score))
+
+
+## dat%>%filter(rdm_condition=='ASSISTments')%>%
+##   select(starts_with("pbo"))%>%
+##   pivot_longer(everything(),names_to="pbo",values_to="prop")%>%
+##   ggplot(aes(log(prop+.01)))+geom_histogram()+facet_wrap(~pbo,ncol=1)
+
+## dat%>%filter(rdm_condition=='ASSISTments')%>%
+##   select(starts_with("pbo"))%>%
+##   map(quantile)
+
+## dat%>%filter(rdm_condition=='ASSISTments')%>%
+##   select(starts_with("pbo"))%>%
+##   map(~c(mean(.),mean(.<0.1)))
+
 
 cat(sum(is.na(dat$nbo)&dat$rdm_condition=='ASSISTments'),' students with NA for # bottom out hints; setting to 0\n')
 dat$nbo[is.na(dat$nbo)&dat$rdm_condition=='ASSISTments'] <- 0
@@ -19,7 +44,8 @@ dat$nbo[is.na(dat$nbo)&dat$rdm_condition=='ASSISTments'] <- 0
 plot(table(dat$nbo[dat$rdm_condition=='ASSISTments']),xlim=c(0,30))
 med <- median(dat$nbo[dat$rdm_condition=='ASSISTments'],na.rm=TRUE)
 
-dat$S <- dat$nbo>med
+dat$S <- #dat$pbo4>0.1#
+  dat$nbo>med
 
 psdat <- dat%>%
[…]

### 2022-09-08 writeUps/paper.tex
more writing

@@
 %%   \caption{Violin and scatterplots of 500 simulation estimation errors for estimators described in Section~\ref{sec:simMods} when the distribution of residuals in \eqref{eq:yc-sim} was normal, there were possibly interactions between covariates and $S$ or $Z$, $\muc1=0.3$, and $\alpha=0.5$.
-Annotations at the upper and lower edges of the plot indicate the number of extreme outliers excluded from the plotting area.}
+Annotations at the upper and lower edges of the plot indicate the number of extreme outliers excluded from the plotting area.}ana
 %%   \label{fig:simUnif}
@@
 
-For the purpose of demonstrating GEEPERS, we gathered dichotomized data on which students in the treatment condition requested at least one bottom-out hint.
-That is, $\st=1$ for students who, if assigned to the treatment condition, would request at least one bottom-out hint, and $\st=0$ for students who would not.
+For the purpose of demonstrating GEEPERS, we gathered dichotomized data on %which students in the treatment condition requested at least one bottom-out hint.
+whether students in the treatment condition requested more than the median number of bottom-out hints (11).
+That is, $\st=1$ for students who, if assigned to the treatment condition, would request at more than 11 bottom-out hints (``bottom-outers''), and $\st=0$ for students who would request 11 or fewer (``non-bottom-outers'').
+%least one bottom-out hint, and $\st=0$ for students who would not.
 The outcome of interest $Y$ in our example is students' total scores on a ten-item posttest completed within the online tutoring system; $Y$ is an integer ranging from 0 to 10.
@@
 
-The data analysis we present here is intended as a demonstration of GEEPERS, rather than for its substantive conclusions, and our description omits discussion of some important methodological considerations, including attrition.
+The data analysis we present here is intended as a demonstration of GEEPERS, rather than for its substantive conclusions, and our description omits discussion of some important methodological considerations, including attrition and post-selection inference.
+
+\subsection{Principal Score Model}
+
+We considered several models for the principal score, and full results (which are largely similar to each other) can be found in the online appendix.
+Here we present results from our model of choice, a logistic regression of observed $S$ on baseline variables chosen with a combination of expert judgement and backwards selection based on the AIC.
+These included school fixed effects and a set of student-level demographic and educational data: fifth-grade mathematics standardized test scores, %Gender, Race/Ethnicity,
+indicator variables for special education status (IEP), whether the student was taking English as a second language (ESOL), whether the student was classified as gifted, and whether students attended a full year of school in 5th grade.
+Finally, we also included data collected from students at the beginning of the study: the natural logrithm of the  time spent on task during the pretest, scores on baseline measurements of mathematics anxiety and self-efficacy and perceputal sensitivity, along with a missingness indicator for these measurements.
+Binned residual plots \citep{gelman2006data} from an initial model fit indicated a curviliniar relationship between percentual sensitivity measurements (modeled as linear) and model errors, so in the final model we included perceptual sensitivity as a quadratic polynomial.
+
+The model was fairly successful in distinguishing bottom-outers from non-bottom-outers---the AUC, evaluated using out-of-sample predictions in a 10-fold cross-validation, was roughly 0.795.
+That is, in roughly 80\% of bottom-outer/non-bottom-outer pairs of subjects, the bottom-outer will have a higher predicted probability from the model.
+Table \ref{tab:results} show the model's coefficient estimates.
+Noteably, conditional on the other covariates, students with higher fifth-grade standardized test scores were less likely to request many bottom out hints, as were students who spent little time on task during the pretest, while IEP and ESOL students, and students without baseline measurement data, were relatively unlikely to be bottom-outers.
+The relationship between perceptual sensitivity and bottom out requests was not monotonic---students with average levels were more likely than those with more extreme high or low levels to be bottom-outers; bottom-outers tended to have lower math self-efficacy, but the data are also consistent with a small positive relationship, and there appears to be little to no relationship between math anxiety and bottom-outer status at all.
+
+\subsection{Outcome Model and
 

### 2022-09-08 code/regression.r
first try at X:S interactions

@@ -19,7 +19,18 @@ Attach <- function(.list,.names=names(.list),env=environment()){
 
 }
 
-AUC = function(probs, true_Y){
+auc <- function(x,y)
+  if(length(unique(y))==2 & length(y)==length(x)){
+    wilcox.test(x~y)$statistic/(sum(y)*(sum(1-y)))
+  }else
+    wilcox.test(x,y)$statistic/(legnth(x)*length(y))
+
+aucMod <- function(mod)
+  auc(mod$linear,1-mod$y)
+
+
+
+AUC1 = function(probs, true_Y){
     probsSort = sort(probs, decreasing = TRUE, index.return = TRUE)
     val = unlist(probsSort$x)
     idx = unlist(probsSort$ix)
@@ -32,7 +43,7 @@ AUC = function(probs, true_Y){
     return(auc)
 }
 AUCmod <- function(mod){
-    AUC(mod$linear,mod$y)
+    auc(mod$linear,1-mod$y)
 }
 
 datNames <- function(data,trt='Z',out='Y',use='S',block=NULL){
@@ -59,7 +70,7 @@ datNames <- function(data,trt='Z',out='Y',use='S',block=NULL){
 }
 
 ### computes/returns regression models
-pointEst <- function(data,covFormU=~x1+x2,covFormY=covFormU,int=FALSE,psMod=NULL){
+pointEst <- function(data,covFormU=~x1+x2,covFormY=covFormU,intS=NULL,psMod=NULL){
 
   data$S[data$Z==0] <- NA
 
@@ -67,7 +78,10 @@ pointEst <- function(data,covFormU=~x1+x2,covFormY=covFormU,int=FALSE,psMod=NULL
       update(covFormU,S~.),
       data=data,family=binomial,
       subset=!is.na(S))
-    else covForm <- formula(psMod)[c(1,3)]
+    else covFormU <- formula(psMod)[c(1,3)]
+
+  if(is.null(covFormY)) covFormY <- covFormU
+
 
     attr(psMod,'auc') <- AUCmod(psMod)
 
@@ -75,10 +89,12 @@ pointEst <- function(data,covFormU=~x1+x2,covFormY=covFormU,int=FALSE,psMod=NULL
     if('Sp'%in%names(data)) warning('replacing Sp')
[…]

### 2022-09-12 code/fh2t/analysisGEE.r
get interactions right

@@ -159,6 +159,8 @@ estimatesInt1 <- lapply(setNames(alts,alts),
                             pre.total_math_score+pre.total_math_scoreNA+ClaIDPre,
                           intS=c("Scale.Score5","pre.total_math_score")))
 
+ sapply(estimatesInt1,effsFromFit,simplify=F)
+
 ### test standard errors w bootstrap
 bsSchool <- function(dat){
   datS <- split(dat,dat$SchIDPre)
@@ -170,15 +172,19 @@ bsInt1 <- lapply(setNames(alts,alts),
                  function(alt){
                    print(alt)
                    datAlt <- getDat(alt)
-                   replicate(5000, effsFromFit(
-                                     est(bsSchool(datAlt),
-                                         covFormU=formula(psMod7)[-2],
-                                         covFormY=~Scale.Score5+ESOL+IEP+pre.total_time_on_tasks+
-                                           pre_MSE_total_score+
-                                           fullYear5+pre_PS_tasks_total_score+raceEth+Gender+GIFTED+
-                                           pre.total_math_score+pre.total_math_scoreNA+ClaIDPre,
-                                         intS=c("Scale.Score5","pre.total_math_score"))))})
+                   replicate(5000, {
+                     bsDat <- datAlt[sample(1:nrow(datAlt),nrow(datAlt),replace=TRUE),]
+                     if(!any(with(bsDat,table(SchIDPre,Z))==0))
+                       effsFromFit(
+                         est(bsDat,#bsSchool(datAlt),
+                             covFormU=formula(psMod7)[-2],
+                             covFormY=~Scale.Score5+ESOL+IEP+pre.total_time_on_tasks+
+                               pre_MSE_total_score+
+                               fullYear5+pre_PS_tasks_total_score+raceEth+Gender+GIFTED+
+                               pre.total_math_score+pre.total_math_scoreNA+ClaIDPre,
+                             intS=c("Scale.Score5","pre.total_math_score")))})})
 
+bsInt1a <- lapply(bsInt1,function(x) do.call("abind",list(x,along=0)))
 save(bsInt1,file='results/bsInt1.RData')
 
 par(mfrow=c(3,3))
[…]

### 2022-09-12 code/regression.r

@@ -4,6 +4,8 @@ library(sandwich)
 ### want them to print every time!
 print.vcv <- function(x,...) print(x[1:NROW(x),1:NCOL(x)],...)
 
+print.geepers <- function(x,...) print(effsFromFit(x),...)
+
 ### convenience function for functions that return lists
 ### why cant R be more like python in this regard??
 Attach <- function(.list,.names=names(.list),env=environment()){
@@ -117,7 +119,7 @@ EFps <- function(psMod,data){
   efPS <- matrix(0,nrow(data),ncol(efPStmp))
   if(is.null(rownames(data))){
     efPS[as.numeric(rownames(efPStmp)),] <- efPStmp
-  } else efPS[match(rownames(data),rownames(efPStmp)),] <- efPStmp
+  } else efPS[match(rownames(efPStmp),rownames(data)),] <- efPStmp
     efPS
 }
 
@@ -131,7 +133,7 @@ sandwichMats <- function(psMod,outMod,data,clust=NULL,int=any(grepl(":x",names(c
     out <- list(
         a11inv = bread(psMod),#/sum(data$Z),
         a22inv = bread(outMod),#/nrow(data),
-        a21 = A21(psMod,outMod,data)/nrow(data),
+        a21 = A21(psMod,outMod,data),#/nrow(data),
         b11=(if(is.null(clust)) meat(psMod) else meatCL(psMod,cluster=clust[data$Z==1])),#*nrow(model.frame(psMod)),
         b22 = (if(is.null(clust)) meat(outMod) else meatCL(outMod,cluster=clust))#*nrow(data)#,adjust=TRUE)
     )
@@ -204,7 +206,9 @@ est <- function(data,covFormU=~x1+x2,covFormY=NULL,psMod=NULL,clust=NULL,intS=NU
 
   vcv <- vcvPS(psMod,outMod,data=data,clust=clust)
 
-  list(outMod=outMod,psMod=psMod,vcv=vcv)
+  out <- list(outMod=outMod,psMod=psMod,vcv=vcv)
+  class(out) <- c('geepers',class(out))
+  out
 }
 
 ### estimates effects of interest, starting from est() output
[…]

### 2022-09-14 writeUps/paper.tex
get interactions right
results figs, etc.

@@
 
-\title{M-Estimation of Principal Effects using Principal Scores}
+\title{GEEPERS: Principal Stratification using Principal Scores and Stacked Estimating Equations}
 
@@
 \section{Introduction}
+
+Randomized experiments easily admit subgroup analysis---estimation of treatment effects for a subset of the study population---when those subgroups are defined at baseline, and not themselves a function of treatment assignment.
+A simple, popular approach to estimating subgroup effects is to use ordinary least squares (OLS) regression, regressing the outcome on a treatment indicator interacted with an indicator for subgroup membership.
+However, when subgroup membership is a function of a post-treatment variable, then it may itself be affected by the treatment.
+Contrats between treatment and control conditions within those subgroups may not have a causal interpretation at all---and when they do, the same techniques that work for subgroups defined at baseline (such as OLS with interactions) will often produce inconsistant estimates.
+Principal stratification \citep{frangakis} is a broad framework for defining subgroup effects for post-treatment subgroups, based on the notition of \emph{potential} subgroups called principal strata, that is, groups of subjects defined based on the values a post-treatment variable would potentially take if the subject is assigned to one treatment condition or another.
+Since principal strata membership is based on subjects' responses to both factual and counter-factual treatment assignemtns, it is, in general unobserved.
+Nevertheless, researchers have devised estimators of principal effects, as average effects within principal strata are called, to address a broad range of applied statistical problems, including non-compliance in randomized experiments, the evaluation of surrogate outcomes \citep{}, the investigation of causal mechanisms \citep{lidsayPage}, and sample attrition \citep{zhouRubin, pengDing2011}.
+
+Because principal strata membership is generally unobserved, principal effects are not always identified, and when they are, estimation can still be difficult.
+Typically, principal effects are identified and estimated based on strong structural assumptions \citep[e.g.][]{air} and/or parametric modeling \citep[e.g.][]{imbensRubin98}.
+Parametric models for principal effects are often hard to specify and hard or impossible to test.
+Even when they are well-specified they can yield estimates that are severely biased and misleading \citep{griffin,feller}.
+Another route to identification and estimation of principal effects leads through auxilliary data, including baseline covariates \citep{} or secondary outcomes \citep{mealli}.
+Of particular importance is the ``principal score,'' \citep{jo,ding,feller}, or the probability that a subject belongs to a particular principal stratum, as a function of pre-treatment covariates.
+Under a strong, untestable assumption called ``principal ignorability,'' which is akin to the no-confounding assumption underlying observational studies, principal scores can be tranformed into weights, which in turn form an unbiased estimate of principal effects.
+
+In a separate, if related, line of research, \citet{ding2011}, \citet{jiangDing2021}, and others have shown that, under certain circumstances, researchers can use baseline covariates to identify and estimate principal effects without relying on principal ignorability.
+This paper expands on those identification results by presenting a new set of principal effect estimators, and associated standard error estimates, based on a generalized estimating equations (or M-estimation or method-of-moments) \citep{stefanskiBoos}.
+It turns out that, under certain general conditions, principal effects can be estimated with a simple, two-step procedure: first, estimate principal scores with logistic regression or another M-estimator; then, impute unknown principal subroup membership indicators with principal scores and estimate principal effects using OLS with interactions, just as in the pre-treatment subgroup case.
 
@@
 
-\section{M-Estimator}
+\section{GEEPERS}
 
@@
 
-\subsection{Relaxing Covariate Ignorability}
+\subsection{Relaxing Covariate Ignorability with Outcome Modeling}
 In most applications, identifying a set of covariates $\bxs$ satisfying CI will be difficult or impossible.
@@
 
-In a recent educational field experiment, middle school students completing their math schoolwork on a computer program were randomized between two conditions.
-In one condition (which we arbitrarily label ``treatment''), students were instructed to work on a fixed number of math problems, and later informed of their scores.
-In a second condition, students received immediate, informative error messages after entering a wrong answer, and were required to enter a correct answer in order to proceed.
-Students in the second treatment arm (labeled ``control'') also had access to a series of hints for each problem, the last of which contained the correct answer.
-Students who ``bottomed out'' on a problem, i.e.
-requested all available hints on a problem, and were told the correct answer, did not receive credit for that problem.
-Students remained in their randomly-assigned condition over the course of a school year, and were assessed with a 10-item posttest administered within the computer program, as well with state standardized tests.
+In a recent educational field experiment, middle school students completing their math schoolwork on a computer program were randomized between four conditions.
+In all four conditions, students worked on their assigned math program in class on 12 different occassions---a prior assessmenent, a mid-test, a post-test, and nine learning sessions---spaced throughout the school year.
+In three of the conditions, during the learning sessions, students worked through a series of math problems, while the forth condition was a computer game designed to teach math concepts.
+In this illustration, we will consider three different treatment effects, comparing one of the four conditions (which we arbitrarily label ``treatment''), with each of the three others (Business as Usual, or ``BAU'', From-Here-To-There!, or ``FH2T,'' and DragonBox), separately.
+During learning sessions, students in the treatment condition received immediate, informative error messages after entering a wrong answer, and were required to enter a correct answer in order to proceed.
+They also had access to a series of hints for each problem, the last of which (called the ``bottom out'' hint) contained the correct answer.
+Students in the other three conditions did not have access bottom out hints; our illustration will focus on that software feature.
 
-The literature on hints in online learning is mixed \citep{stuff}, in particular when it comes to bottoming out.
+The literature on hints in online learning is mixed \citep{stuff}, in particular when it comes to requesting a bottom-out hint (called ``bottoming out'').
 On the one hand, hints that include the correct answer are essentially worked examples, which can be beneficial to learning \citep{worked example stuff}.
@@
 The goal of our analysis will be to determine the principal effects $\tau_1=\EE[\yt-\yc|\st=0]$ and $\tau_0=\EE[\yt-\yc|\st=0]$, the average effects of assignment to the treatment condition for bottom-outers and non-bottom-outers.
-We also gathered data on baseline covariates: total scores on a 10-item pretest, fifth-grade stanardized test scores, race/ethnicity (white, Asian, Hispanic/Latino, or Other), indicator variables for if students have an early intervention program (EIP) or individualized educational plan (IEP) in place, if students are learning English as a second language (ESOL), students' average time on task for the pretest (logged), total scores from baseline evaluations of students' math anxiety, negative reaction to math, and numerical confidence, and an indicator for weather students began the school year in remote or in-person instruction.
+
+We also had access to data on a number of baseline covariates: total scores on a 10-item pretest, fifth-grade stanardized test scores, race/ethnicity (white, Asian, Hispanic/Latino, or Other), indicator variables for if students have an early intervention program (EIP) or individualized educational plan (IEP) in place, if students are learning English as a second language (ESOL), students' average time on task for the pretest (logged), total scores from baseline evaluations of students' math anxiety, negative reaction to math, and numerical confidence, and an indicator for weather students began the school year in remote or in-person instruction.
 Students with missing pretest scores were dropped from the analysis; missing data in other covariates was imputed with a Random Forest algorithm \citep{missForest}.
@@
 The data analysis we present here is intended as a demonstration of GEEPERS, rather than for its substantive conclusions, and our description omits discussion of some important methodological considerations, including attrition and post-selection inference.
+For a more complete discussion of the experimental design, the conditions being compared, and the impact analysis, please see \citet{impactPaper}.
 
 \subsection{Principal Score Model}
+
+\begin{figure}
+  \centering
+  \includegraphics{../figure/psModCoef.jpg}
+  \caption{Principal Score Model Coefficients}
+  \label{fig:psMod}
+\end{figure}
 
@@
 Binned residual plots \citep{gelman2006data} from an initial model fit indicated a curviliniar relationship between percentual sensitivity measurements (modeled as linear) and model errors, so in the final model we included perceptual sensitivity as a quadratic polynomial.
+The model was fit to data from students in the treatment group, for whom bottom-out status ($S$) was observed; hence, the same model fit was used for all three treatment contrasts.
 
@@
 
-\subsection{Outcome Model and
+\subsection{Outcome Regression and Principal Effects}
+\begin{figure}
+  \centering
+  \includegraphics{../figure/prinEffs.pdf}
+  \caption{Principal Effects}
+  \label{fig:effects}
+\end{figure}
+
+Figure \ref{fig:effects} plots the estimated principal effects for bottom-outers and non-bottom-outers, using two different models---alongside estimates of the overall average treatment effect, labeled ``ATE.'' In both models, effects were adjusted for all of the variables that were in the principal score model, in addition to students' pretest scores; also, the models substituted class fixed effects for school fixed effects.\footnote{Why use a corser set of fixed effects for the principal score model than the outcome regression?
+There were two reasons: first, the principal score model was fit to a smaller sample (only students in the ``Instant'' condition), so some classrooms only included one student, which can lead to poor fit.
+Second, the asymptotics of logistic regression are more restricted than OLS, and so logistic regression tends to perform poorly when the number of parameters increases with the sample size \citep{agresti}.}
+The two sets of esimated differed in that the first assumed that coefficients on covariates were equal across principal strata, whereas the second set of results allowed the coefficients on pretest and 5th-grade standardized test scores---the most important predictors of outcomes and principal strata membership, respectively---to vary between strata.
+
+The results suggest little difference, if any, between the effects in the two principal strata.
+Estimated effects compared to BAU or DragonBox were slightly lower for bottom-outers than for non-bottom-outers, and estimated effects compared to FH2T were higher for non-bottom-outers.
+However, none of these differences is statistically significant, so opposite patterns, or no difference betweeen principal effects at all, would also be consistant with the data.
+In fact, 95\% confidence intervals for all principal effects include both positive and negative principal effects.
+There is no discernable difference between estimated principal effects when coefficients on pretest and grade-5 standardized tests were or were not allowed to vary between strata, though standard errors from the models including an interaction were somewhat higher.
+
+\subsection{Mixture Model and Principal Score Weighting}
+\begin{figure}
+  \centering
+  \includegraphics{../figure/compareMethods.pdf}
+  \caption{Principal Effects}
+  \label{fig:compare}
+\end{figure}
 

### 2022-09-14 code/fh2t/analysisGEE.r

@@ -1,6 +1,7 @@
 library(tidyverse)
 library(arm)
 library(randomForest)
+library(estimatr)
 source('code/regression.r')
 select <- dplyr::select
 
@@ -14,13 +15,9 @@ psdat <- filter(psdat,SchIDPre!=7) ### only one student from school 7, in FH2T
 psdat$Z <- ifelse(psdat$trt=='ASSISTments',1,0)
 
 alts <- unique(psdat$trt[psdat$Z==0])
+alts <- setNames(alts,alts)
 
-### do a little "reloop"
-## remnantMods <- map(setNames(alts,alts),
-##                    ~psdat%>%
-##                    filter(!trt%in%c('ASSISTments',.x))%>%
-##                    select(-StuID,-TeaIDPre,-ClaIDPre,-S,-virtual,-Z,-trt)%>%
-##                    randomForest(Y~.,data=.))
+
 
 
 getDat <- function(alt){
@@ -31,31 +28,36 @@ getDat <- function(alt){
   droplevels(dat)
 }
 
-### overall ATEs
-## ateMods <- sapply(setNames(alts,alts),
-##                   function(alt)
-##                     estimatr::lm_robust(
-##                                 Y~Z+pred+
-##                                   Scale.Score5+ESOL+IEP+pre.total_time_on_tasks+pre_MSE_total_score+
-##                             fullYear5+pre_PS_tasks_total_score+raceEth+Gender+GIFTED+
-##                             splines::ns(pre.total_math_score,3)+pre.total_math_scoreNA,
-##                             data=getDat(alt),fixed_effects=~ClaIDPre),
-##                   simplify=FALSE)
 
-
-ateModsNoRem <- sapply(setNames(alts,alts),
+ates <- sapply(setNames(alts,alts),
                   function(alt)
                     estimatr::lm_robust(
                                 Y~Z+Scale.Score5+ESOL+IEP+pre.total_time_on_tasks+pre_MSE_total_score+
                             fullYear5+pre_PS_tasks_total_score+raceEth+Gender+GIFTED+
-                            splines::ns(pre.total_math_score,3)+pre.total_math_scoreNA,
+                            pre.total_math_score+pre.total_math_scoreNA+ClaIDPre,
                             data=getDat(alt),fixed_effects=~ClaIDPre),
[…]

### 2022-09-14 code/fh2t/analysisPSW.r

@@ -1,0 +1,73 @@
+library(tidyverse)
+library(arm)
+library(randomForest)
+library(estimatr)
+source('code/regression.r')
+select <- dplyr::select
+
+load('data/psdat.RData')
+
+
+load('results/psModGlm7.RData')
+
+psdat <- filter(psdat,SchIDPre!=7) ### only one student from school 7, in FH2T
+
+psdat$Z <- ifelse(psdat$trt=='ASSISTments',1,0)
+
+alts <- unique(psdat$trt[psdat$Z==0])
+alts <- setNames(alts,alts)
+
+bsSchool <- function(dat){
+  datS <- split(dat,dat$SchIDPre)
+  do.call("rbind",
+          lapply(datS,function(x) x[sample(1:nrow(x),nrow(x),replace=TRUE),]))
+}
+
+
+#### PSW
+psw1 <- function(dat,psMod){
+
+  dat0=subset(dat,Z==0)
+  dat1=subset(dat,Z==1)
+  dat0$ps=predict(psMod,dat0,type='response')
+
+
+  muc0=with(dat0,sum(Y*(1-ps))/sum(1-ps))
+  muc1=with(dat0,sum(Y*ps)/sum(ps))
+
+  mut0=with(dat1,mean(Y[S==0]))
+  mut1=with(dat1,mean(Y[S==1]))
+
+  c(eff0=mut0-muc0,
+    eff1=mut1-muc1,
+    diff=mut1-muc1-mut0+muc0
+    )
+}
+
+print.psw <- function(x, ...) print(x$coef,...)
+
+
+psw <- function(alt,psMod,B=5000,verbose=TRUE){
+  dat <- getDat(alt)
+
+  est=psw1(dat,psMod)
+
+  if(verbose) step <- if(B>10) round(B/10) else 1
+  bs <- matrix(nrow=B,ncol=3)
+  for(i in 1:B){
+    if(verbose) if(i%%step==0) cat(round(i/B*100),' ')
+    datStar <- bsSchool(dat)
+    psModStar <- update(psMod,data=subset(datStar,Z==1))
+    bs[i,] <- psw1(datStar,psModStar)
+  }
+  if(verbose) cat('\n')
+  out <-
+    list(
+      coef=cbind(est=est,se=apply(bs,2,sd)),
+      bs=bs)
+  class(out) <- 'psw'
+  out
+}
+
+pswResults <- lapply(setNames(alts,alts),psw,psMod=psMod7)
+save(pswResults,file='results/psw.RData')

### 2022-09-14 code/fh2t/tables.r

@@ -5,6 +5,7 @@ library(tableone)
 library(kableExtra)
 library(xtable)
 source('code/regression.r')
+library(huxtable)
 select <- dplyr::select
 
 options(knitr.kable.NA = '')

### 2022-09-15 writeUps/paper.tex
GEEPERs

@@
 
-\title{GEEPERS: Principal Stratification using Principal Scores and Stacked Estimating Equations}
+\title{GEEPERs: Principal Stratification using Principal Scores and Stacked Estimating Equations}
 
@@
 
-In a separate, if related, line of research, \citet{ding2011}, \citet{jiangDing2021}, and others have shown that, under certain circumstances, researchers can use baseline covariates to identify and estimate principal effects without relying on principal ignorability.
+In a separate, if related, line of research, \citet{jo2002}, \citet{ding2011}, \citet{jiangDing2021}, and others have shown that, under certain circumstances, researchers can use baseline covariates to identify and estimate principal effects without relying on principal ignorability.
 This paper expands on those identification results by presenting a new set of principal effect estimators, and associated standard error estimates, based on a generalized estimating equations (or M-estimation or method-of-moments) \citep{stefanskiBoos}.
 It turns out that, under certain general conditions, principal effects can be estimated with a simple, two-step procedure: first, estimate principal scores with logistic regression or another M-estimator; then, impute unknown principal subroup membership indicators with principal scores and estimate principal effects using OLS with interactions, just as in the pre-treatment subgroup case.
+We will refer to the estimators presented here as ``GEEPERs'': generalized estimating equations for principal effects using regressions.
+GEEPERs estimates rely on correct specification of regression functions, but unlike common fully-parametric methods, they make no assumptions about the distribution of regression errors.
+Unlike principal score weighting, GEEPERs does not assume principal ignorability, and admits to sandwich-style standard error estimates, which we provide in the appendix.
 
-We refer to the estimators presented here as ``GEEPERs'': generalized estimating equations for principal effects using regressions.
+This paper focuses on a special case of principal stratification---one-way, binary non-compliance---that is, there are only two principal strata, and principal stratum membership is observable in one treatment arm, but not in the other.
+This scenario would occur if, say, subjects in the control arm of a study had no access to the treatment, and not all subjects assigned to the treatment arm actually received the full treatment.
+While such a setup may represent the simpliest and easiest principal stratification problem, we believe it is a common scenario.
+Moreover, we believe that the methods developed here can extend to more complex scenarios, which we leave to future work.
+
+The following section gives a formal introduction to principal stratification and briefly describes two common estimation techniques, parametric mixture modeling and principal score weighting.
+Next, section \ref{sec:geepers} develops GEEPERs estimation, and gives conditions for its consistency.
+Section \ref{sec:simulation} describes a simulation study examining the operating characteristics of GEEPERs estimates and comparing them to estimates from parametric mixture modeling and principal score weighting.
+Under ideal circumstances, estimates from parametric mixture models and weighting estimators are somewhat more efficient than GEEPERs, but when assumptions are violated GEEPERs easily out-performs the other two methods.
+Section \ref{sec:fh2t} illustrates GEEPERs in analyzing data from a recent large-scale field trial of educational software, and Section \ref{sec:discussion} concludes.
 
@@
 
-\section{GEEPERS}
+\section{GEEPERs}\label{sec:geepers}
 
@@
 
-We begin by introducing an assumption that we term ``Covariate Ignorability,'' or CI.
+We begin by introducing an assumption that we term ``Covariate Ignorability,'' or CI. It is a slighty weaker version of \citet{jiangDing2021}'s ``auxilliary indpendence'' assumption.
 \begin{ass}[Covariate Ignorability]\label{ass:ci}
@@
 Thus, given principal scores, Assumption \ref{ass:ci} enables an analyst to estimate principal effects with a simple regression.
+Proposition \ref{prop:reg1} builds on Theorem 2 of \citet{jiangDing2021} which establishes identification of $\eff0$ and $\eff1$ by representing the principal effect estimators as a familiar regression using estimated principal scores.
+
 If principal scores were known exactly, the M-estimation structure of $\hat{\eff0}$ and $\hat{\eff1}$ implies that their sampling variance could be estimated by the standard sandwich estimator for OLS \citep{sefanskiBoosArticle}.
@@
   \end{equation}
-  with $\EE[\epsilon_i]=0$ and $\{Z_i,\sti,\bxy_i\}\independent \epsilon_i$.
-  Then say the following model is fit with OLS
+  with $\EE[\epsilon_i]=0$, $\{Z_i,\sti,\bxy_i\}\independent \epsilon_i$.
+  Say $\ppi$ linearly independent of $\bxy$, and
+  that the following model is fit with OLS
   \begin{equation}\label{eq:fullInteractionModel}
@@
 In particular, if interactions between covariates and treatment assignment are omitted, then data from the treatment group---in which $\st$ is fully observed---can be used to estimate the coefficients on covariates.
-Such a model implies a fairly strong assumption, that the only interactions between covariates and treatment assignment are due to differences between principal effects.
+Such a model relies on what \citet{bo2002} calls ``Additivity of Treatment Assignment Effect,'' that $\bm{\beta_1}^{C}=\bm{\beta_1}^{T}$.
+In other words, the only interactions between covariates and treatment assignment are due to differences between principal effects---a fairly strong assumpton.
 However, there are two mitigating points.
@@
 As in the simpler case discussed in \S~\ref{sec:ci}, since the estimators in \eqref{eq:prinEffEst} are M-Estimators, their sampling variance can be estimated with a consistent sandwich estimator.
-Uncertainty in principal score estimation can be incorporated by stacking the estimating equations for the principal score model \eqref{eq:prinScoreMod} and estimating the combined covariance matrix \eqref{eq:sandwich}..
+Uncertainty in principal score estimation can be incorporated by stacking the estimating equations for the principal score model \eqref{eq:prinScoreMod} and estimating the combined covariance matrix \eqref{eq:sandwich}, formulas for which are given in the appendix.
 
@@
 
-We conducted a simulation study to compare the performance of our regression estimator compared to a Bayesian mixture model and to a principal score weighting estimator, in both favorable and unfavorable circumstances.
+We conducted a simulation study to compare the performance of GEEPERs compared to a Bayesian mixture model and to a principal score weighting estimator, in both favorable and unfavorable circumstances.
 The study was designed to answer three sets of overarching questions.
-First, how does the performance of the M-estimator vary as sample size and the extent to which $\bx$ predicts $\st$, and the principal effects themselves vary, and when important interactions were omitted from the outcome model.
-Second, how does the M-estimator could compete with the mixture model?
+First, how does the performance of the GEEPERs estimator vary under different conditions, including sample size, the extent to which $\bx$ predicts $\st$, the principal effects themselves vary, and when important interactions were omitted from the outcome model.
+Second, how does GEEPERs compare with mixture modeling?
 Is it competitive in circumstances favorable to both techniques?
 Does it avoid the pitfalls of mixture modeling when the parametric assumptions of the normal mixture model fail?
-Third, are there ways or circumstances when the principal score weighting estimator outperforms the M-estimator even though principal ignorability does not hold?
+Third, are there ways or circumstances when the principal score weighting estimator outperforms GEEPERs even though principal ignorability does not hold?
 
@@
 
-Across simulation runs, there were $n=500$ subjects in each treatment arm.
-
-Given three indpendent standard normal covariates, $x_k$ $k=1,\dots,3$,
-principal scores were simulated as
+In each run of the simulation, we simulated three indpendent standard normal covariates, $x_k$ $k=1,\dots,3$.
+However, only the first two covariates were ``observed,'' i.e.
+were included in the analysis model.
+Given the covariates, the (true) principal scores were defined as
 \begin{equation*}
@@
 
-In most simulation runs, there were no interactions between
-Control potential outcomes were generated as
-\begin{equation}\label{eq:yc-sim}
-  \yci=\beta_1(x_{1i}+x_{2i})+[\muc1+\beta_2(x_{1i}+x_{2i})]\sti +x_{3i}/\sqrt{6}+\epsilon_i
-\end{equation}
-where $\EE[\epsilon_i]=0$ $Var(\epsilon_i)=1/2$, and the distribution of $\epsilon_i$ and $\muc1$ were manipulated factors in the study.
-In most simulations, there was no interaction between $\st$ and $x_1$ and $x_2$; in those cases, $\beta_1=1/\sqrt{6}$, and $\beta_2=0$.
-In simulation runs including an interaction with $\st$, $\beta_1=3/(4\sqrt{6})$ and $\beta_2=5/(4\sqrt{6})$.
-
-Finally, treatment potential outcomes were generated as
-\begin{equation*}
-  \yti=\yci+(0.3-\muc1)\sti+\beta_3(x_{1i}+x_{2i})
-\end{equation*}
-and observed outcomes were $Y_i=Z_i\yti+(1-Z_i)\yci$, with even subjects assigned to treatment and odd subjects assigned to control.
-In most simulation runs, there were no interactions between treatment assignemnt and covariates, so $\beta_3=0$; when such interactions were included, $\beta_3=1/(2\sqrt{6})$.
-
-Since the covariates were all centered, $\muc0=\mut0=\eff0=0$ throughout the study.
-Similarly, $\mut1=0.3$, while $\muc1$, and hence $\eff1$, varied between runs.
-
-Note that when there were no interactions included in the model, since $Var(\beta_1(x_1+x2+x_3))=1/2$ and $Var(\epsilon_i)=1/2$, covariates explained half of the variance of $Y$ within treatment group and principal stratum.
-However, as we shall see, $x_3$ was ``unobserved,'' so observed covariates explained 1/3 of that variance.
+Finally, potential outcomes were generated as in \eqref{eq:interaction}, with $\bxy=[\bm{x_1},\bm{x_2},\bm{x_3}]$,
+and with $\EE[\epsilon_i]=0$ and $Var(\epsilon_i)=1/2$.
+Across simulation runs, intercepts $\beta_0^C=\beta_0^T=0$ and the coefficient on the unobserved covariate $x_3$, $\beta_{1,3}^C=\beta_{1,3}^T= 1/\sqrt{6}$.
+The distribution of $\epsilon_i$ and other coefficients $\bm{\beta}$ were manipulated factors in the study.
 
@@
 The last two factors---the presence of interactions between covariates and $\st$ and $Z$, respectively---test the performance of a model that does not include interactions when that modeling assumption is false.
+
+In most simulations, there was no interaction between $\st$ and $x_1$ and $x_2$; in those cases, $\beta_1=1/\sqrt{6}$, and $\beta_2=0$.
+In simulation runs including an interaction with $\st$, $\beta_1=3/(4\sqrt{6})$ and $\beta_2=5/(4\sqrt{6})$.
+
+Finally, treatment potential outcomes were generated as
+\begin{equation*}
+  \yti=\yci+(0.3-\muc1)\sti+\beta_3(x_{1i}+x_{2i})
+\end{equation*}
+and observed outcomes were $Y_i=Z_i\yti+(1-Z_i)\yci$, with even subjects assigned to treatment and odd subjects assigned to control.
+In most simulation runs, there were no interactions between treatment assignemnt and covariates, so $\beta_3=0$; when such interactions were included, $\beta_3=1/(2\sqrt{6})$.
+
+Since the covariates were all centered, $\muc0=\mut0=\eff0=0$ throughout the study.
+Similarly, $\mut1=0.3$, while $\muc1$, and hence $\eff1$, varied between runs.
+
+Note that when there were no interactions included in the model, since $Var(\beta_1(x_1+x2+x_3))=1/2$ and $Var(\epsilon_i)=1/2$, covariates explained half of the variance of $Y$ within treatment group and principal stratum.
+However, as we shall see, $x_3$ was ``unobserved,'' so observed covariates explained 1/3 of that variance.
 
@@
 
-For the purpose of demonstrating GEEPERS, we gathered dichotomized data on %which students in the treatment condition requested at least one bottom-out hint.
+For the purpose of demonstrating GEEPERs, we gathered dichotomized data on %which students in the treatment condition requested at least one bottom-out hint.
 whether students in the treatment condition requested more than the median number of bottom-out hints (11).
@@
 
-The data analysis we present here is intended as a demonstration of GEEPERS, rather than for its substantive conclusions, and our description omits discussion of some important methodological considerations, including attrition and post-selection inference.
+The data analysis we present here is intended as a demonstration of GEEPERs, rather than for its substantive conclusions, and our description omits discussion of some important methodological considerations, including attrition and post-selection inference.
 For a more complete discussion of the experimental design, the conditions being compared, and the impact analysis, please see \citet{impactPaper}.
 
-\subsection{Principal Score Model}
+\subsection{Principal Score Model}\label{sec:psMod}
 
@@
 
-\subsection{Outcome Regression and Principal Effects}
+\subsection{Outcome Regression and Principal Effects}\label{sec:outMod}
 \begin{figure}
@@
 \end{figure}
+
+Figure \ref{fig:compare} shows principal effect estimates from GEEPERs (without the $S$-interactions) compared to analogous estimates using Bayesian mixture modeling and PSW.
+All three sets of estimates relied on the same principal score model described above, in Section \ref{sec:psMod}, though in the mixture model it was fit simultaneously with effect estimation using a Markov Chain Monte Carlo algorithm \citep{rstan}.
+The Bayesian mixture model assumed that outcomes were normally distributed, conditional on covariates (the same as in Section \ref{sec:outMod}) and principal stratum, with a residual standard deviation that varied between principal strata.
+This modeling assumption is necessarily false, since post-test scores were integers between 0 and 10.
+All model parameters with support in $\mathbb{R}$ were given standard normal priors, while standard deviation parameters were given half-standard normal priors.
+The PSW estimators did not include any covariate outcome modeling. Standard errors for the PSW estimates were estimated using the bootstrap, with resamping done within schools.
+For all three sets of estimators, approximate 95\% confidence intervals were estimated by subtracting twice the standard error from the point estimate (or twice the posterior standard deviation from the posterior mean, in the Bayesian mixture model).
+
+In this case, all three estimation strategies largely agreed.
+This may suggest that the estimators' various assumptions---additivity of treatment assignment effect for GEEPERs and the mixture model, normal residuals for the mixture model, and principal ignorability for PSW---were all innocuous and may have held approximately.
+Alternatively, violations of the different assumptions may have led to similar errors.
 

### 2022-09-15 code/simulation/displaySim.r

@@ -29,7 +29,7 @@ load('simResults/resultsB1s.RData')
               labels=c(bquote(mu[c]^1-mu[c]^0==0),bquote(mu[c]^1-mu[c]^0==0.3))),
     PE=paste('Stratum',eff),
     dist=paste(c(mix='Mixture',unif='Unform',norm='Normal')[errDist],'Errors'),
-    estimator=c(bayes='Mixture',mest='M-Est',psw='PSW')[estimator],
+    estimator=c(bayes='Mixture',mest='GEEPERs',psw='PSW')[estimator],
     interactionZ=ifelse(intZ,"Z interaction","No\nZ interaction"),
     interactionS=ifelse(intS,"S interaction","No\nS interaction"),
     N=paste0("n=",n)
@@ -122,7 +122,7 @@ ggsave('simFigs/byN.jpg',width=6,height=3)
               labels=c(bquote(mu[c]^1-mu[c]^0==0),bquote(mu[c]^1-mu[c]^0==0.3))),
     PE=paste('Stratum',eff),
     dist=paste(c(mix='Mixture',unif='Unform',norm='Normal')[errDist],'Errors'),
-    estimator=c(bayes='Mixture',mest='M-Est',psw='PSW')[estimator],
+    estimator=c(bayes='Mixture',mest='GEEPERs',psw='PSW')[estimator],
     interactionZ=ifelse(intZ,"Z interaction","No\nZ interaction"),
     interactionS=ifelse(intS,"S interaction","No\nS interaction"),
     N=paste0("n=",n)
@@ -204,7 +204,7 @@ pd <- results %>%
               labels=c(bquote(mu[c]^1==0),bquote(mu[c]^1==0.3))),
     PE=paste('Stratum',eff),
     dist=paste(c(mix='Mixture',unif='Unform',norm='Normal')[errDist],'Errors'),
-    estimator=c(bayes='Mixture',mest='M-Est',psw='PSW')[estimator],
+    estimator=c(bayes='Mixture',mest='GEEPERs',psw='PSW')[estimator],
     interactionZ=ifelse(intZ,"Z\ninteraction","No Z\ninteraction"),
     interactionS=ifelse(intS,"S\ninteraction","No S\ninteraction")
   ) %>%
@@ -417,7 +417,7 @@ bias%>%
[…]

### 2022-09-30 writeUps/paper.tex
update paper

@@
 \usepackage{multirow}
-\usepackage{../../../generalCode/numdef}
+\usepackage{numdef}
 
@@
 However, when subgroup membership is a function of a post-treatment variable, then it may itself be affected by the treatment.
-Contrats between treatment and control conditions within those subgroups may not have a causal interpretation at all---and when they do, the same techniques that work for subgroups defined at baseline (such as OLS with interactions) will often produce inconsistant estimates.
-Principal stratification \citep{frangakis} is a broad framework for defining subgroup effects for post-treatment subgroups, based on the notition of \emph{potential} subgroups called principal strata, that is, groups of subjects defined based on the values a post-treatment variable would potentially take if the subject is assigned to one treatment condition or another.
-Since principal strata membership is based on subjects' responses to both factual and counter-factual treatment assignemtns, it is, in general unobserved.
-Nevertheless, researchers have devised estimators of principal effects, as average effects within principal strata are called, to address a broad range of applied statistical problems, including non-compliance in randomized experiments, the evaluation of surrogate outcomes \citep{}, the investigation of causal mechanisms \citep{lidsayPage}, and sample attrition \citep{zhouRubin, pengDing2011}.
+Contrasts between treatment and control conditions within those subgroups may not have a causal interpretation at all---and when they do, the same techniques that work for subgroups defined at baseline (such as OLS with interactions) will often produce inconsistent estimates.
+Principal stratification \citep{frangakis} is a broad framework for defining subgroup effects for post-treatment subgroups, based on the notion of \emph{potential} subgroups called principal strata, that is, groups of subjects defined based on the values a post-treatment variable would potentially take if the subject is assigned to one treatment condition or another.
+Since principal strata membership is based on subjects' responses to both factual and counter-factual treatment assignments, it is, in general unobserved.
+Nevertheless, researchers have devised estimators of principal effects, as average effects within principal strata are called, to address a broad range of applied statistical problems, including non-compliance in randomized experiments, the evaluation of surrogate outcomes \citep{li2010bayesian}, the investigation of causal mechanisms \citep{lidsayPage}, and sample attrition \citep{zhangRubin, ding2011}.
 
 Because principal strata membership is generally unobserved, principal effects are not always identified, and when they are, estimation can still be difficult.
-Typically, principal effects are identified and estimated based on strong structural assumptions \citep[e.g.][]{air} and/or parametric modeling \citep[e.g.][]{imbensRubin98}.
+Typically, principal effects are identified and estimated based on strong structural assumptions \citep[e.g.][]{air} and/or parametric modeling \citep[e.g.][]{imbens1997bayesian}.
 Parametric models for principal effects are often hard to specify and hard or impossible to test.
-Even when they are well-specified they can yield estimates that are severely biased and misleading \citep{griffin,feller}.
-Another route to identification and estimation of principal effects leads through auxilliary data, including baseline covariates \citep{} or secondary outcomes \citep{mealli}.
-Of particular importance is the ``principal score,'' \citep{jo,ding,feller}, or the probability that a subject belongs to a particular principal stratum, as a function of pre-treatment covariates.
-Under a strong, untestable assumption called ``principal ignorability,'' which is akin to the no-confounding assumption underlying observational studies, principal scores can be tranformed into weights, which in turn form an unbiased estimate of principal effects.
+Even when they are well-specified they can yield estimates that are severely biased and misleading \citep{griffin2008application,feller2016principal}.
+Another route to identification and estimation of principal effects leads through auxiliary data, including baseline covariates or secondary outcomes \citep{mattei2013exploiting}.
+Of particular importance is the ``principal score,'' \citep{jo,ding2017principal,feller2017principal}, or the probability that a subject belongs to a particular principal stratum, as a function of pre-treatment covariates.
+Under a strong, untestable assumption called ``principal ignorability,'' which is akin to the no-confounding assumption underlying observational studies, principal scores can be transformed into weights, which in turn form an unbiased estimate of principal effects.
 
@@
 This scenario would occur if, say, subjects in the control arm of a study had no access to the treatment, and not all subjects assigned to the treatment arm actually received the full treatment.
-While such a setup may represent the simpliest and easiest principal stratification problem, we believe it is a common scenario.
+While such a setup may represent the simplest and easiest principal stratification problem, we believe it is a common scenario.
 Moreover, we believe that the methods developed here can extend to more complex scenarios, which we leave to future work.
@@
 In a randomized experiment with two conditions, let $Z_i=0$ if subject $i$, $i=1,\dots,n$ is assigned to the treatment condition, and let $Z_i=0$ otherwise.
-If interest is in the effect of $Z$ on an outcome $Y$, under the ``stable unit treatment value assumption'' \citep{sutva} of no interference between units and no hidden versions of the treatment, define potential outcomes \citep{neyman,rubin} $\yti$ as the value of $Y$ that $i$ would exhibit if $Z_i=1$, and let $\yci$ be the value if $Z_i=0$.
-Then the observed value of $Y$ satisfies $Y_i=Z_i\yti+(1-Z_i)\yci$, and the treatment effect of $Z$ on $Y$ for stubject $i$ may be defined as $\tau_i=\yti-\yci$.
+If interest is in the effect of $Z$ on an outcome $Y$, under the ``stable unit treatment value assumption'' \citep{air} of no interference between units and no hidden versions of the treatment, define potential outcomes \citep{splawa1990application,air} $\yti$ as the value of $Y$ that $i$ would exhibit if $Z_i=1$, and let $\yci$ be the value if $Z_i=0$.
+Then the observed value of $Y$ satisfies $Y_i=Z_i\yti+(1-Z_i)\yci$, and the treatment effect of $Z$ on $Y$ for subject $i$ may be defined as $\tau_i=\yti-\yci$.
 The average treatment effect (ATE) is $\EE[\tau]$.
@@
 Then $S$ itself has potential values $\sti$ and $\sci$, which $i$ would exibit if $Z_i=1$ or $Z_i=0$, respectively.
-Assume ``strong monotonicity'' \citep[c.f.][]{dingLu} (also referred to as ``one-way noncompliance'' \citealt{imbensRubin}):
-\begin{ass}[Strong Monotonicity]\label{ass:SM}
+Assume ``strong monotonicity'' \citep[c.f.][]{dingLu} (also referred to as ``one-way noncompliance'' \citealt{}):
+\begin{ass}[Strong Monotonicity]\label{ass:sm}
 \begin{equation*}%\label{eq:strongMonotonicity}
@@
 The average treatment effect in each principal stratum, $\eff1=\EE[\tau|\st=1]$ or $\eff1=\EE[\tau|\st=0]$, is called a principal effect. %
-The challange in estimating principal effects is due to the fact that strata membership $\sti$ is only observed when $Z_i=1$, in which case $\sti=S_i$, the observed value of $S$.
+The challenge in estimating principal effects is due to the fact that strata membership $\sti$ is only observed when $Z_i=1$, in which case $\sti=S_i$, the observed value of $S$.
 
@@
 In contrast, $\muc0$ and $\muc1$ are not fully identified without further assumptions because $\sti$ and $\yci$ are never observed simultaneously (though they may be nonparametrically bounded; see, e.g., \citealt{bounding}).
-When $S$ is a measure of compliance, and $\tau_i=0$ whenever $\sti=0$ (the ``exclusion restriction''), $\muc1$---and hence $\eff1$---may be estimated with instrumental variables methods \citep{angrist1996identification}.
+When $S$ is a measure of compliance, and $\tau_i=0$ whenever $\sti=0$ (the ``exclusion restriction''), $\muc1$---and hence $\eff1$---may be estimated with instrumental variables methods \citep{air}.
 In this case, $\eff1$ is often called the ``local average treatment effect'' or the average effect of the treatment on the treated.
@@
 
-In order for principal scores to be useful, they need to vary with $\bm{x}$. In particular, we will assume the following:
+In order for principal scores to be useful, they need to vary with $\bm{x}$. Otherwise, $\bm{x}$ is uninformative about $\st$. In particular, we will assume the following:
 \begin{ass}[Variable Principal Scores]\label{ass:vps}
-  \begin{equation*}
-    Var_{\bx} (\pp)>0
-  \end{equation*}
+ $\pp$ takes at least 3 distinct values
 \end{ass}
+In some estimators we will discuss below, two distinct values suffice. For a broader discussion, see \citet{ding2011,jiangDing2021}.
 
@@
 That is, the principal stratum is unrelated to control potential outcomes conditional on covariates.
-Principal ignorability \eqref{eq:pi} is reminiscent of ignorability assumptions typical in observational studies \citep[e.g.][]{rosenbaum2002observational}.
+Principal ignorability \eqref{eq:pi} is reminiscent of ignorability assumptions typical in observational studies \citep[e.g.][]{rosenbaum2010design}.
 Both require a rich set of covariates and background knowledge of the determinants of $S$; in particular, an unobserved covariate $u$ that is correlated with both $\st$ and $Y_C$ would invalidate \eqref{eq:pi}.
@@
 
-We begin by introducing an assumption that we term ``Covariate Ignorability,'' or CI. It is a slighty weaker version of \citet{jiangDing2021}'s ``auxilliary indpendence'' assumption.
+We begin by introducing an assumption that we term ``Covariate Ignorability,'' or CI. It is a slighty weaker version of \citet{jiangDing2021}'s ``auxiliary independence'' assumption.
 \begin{ass}[Covariate Ignorability]\label{ass:ci}
@@
 
-It turns out that under strong monotonicity, randomization, and CI, and given a set of principal scores, the two principal effects $\eff0$ and $\eff1$ can be estimted by a simple OLS regression.
+It turns out that under strong monotonicity, randomization, and CI, and given a set of principal scores, the two principal effects $\eff0$ and $\eff1$ can be estimated by a simple OLS regression.
 To see how, we'll %first
@@
 The argument stems from a set of expressions for expectations, summarized in the following lemma:
-\begin{lemma}\label{lemma:expectations}
+\begin{lemma}\label{lemma:expectation}
   Under Assumptions \ref{ass:sm}, \ref{ass:rand}, \ref{ass:vps}, and \ref{ass:ci}, %if principal scores $\ppi$, for $i:Z_i=0$ are known exactly, then:
@@
 \end{equation}
-Note that, in general, $(1-Z_i)\ppi=(1-Z_i)\ri$ and $Z_i\sti=Z_i\ri$, allowing $\ri$ to take the place of $\ppi$ and $S_i$ througout $\tilde{\Psi}_i$.
+Note that, in general, $(1-Z_i)\ppi=(1-Z_i)\ri$ and $Z_i\sti=Z_i\ri$, allowing $\ri$ to take the place of $\ppi$ and $S_i$ throughout $\tilde{\Psi}_i$.
 
 \begin{prop}\label{prop:reg1}
-  Under Assumptions \ref{ass:sm}, \ref{ass:randomization}, \ref{ass:vps}, and \ref{ass:ci}, if %principal scores are known exactly,
+  Under Assumptions \ref{ass:sm}, \ref{ass:rand}, \ref{ass:vps}, and \ref{ass:ci}, if %principal scores are known exactly,
   $\ri$ is defined as in \eqref{eq:ri} and the model
@@
 
-If principal scores were known exactly, the M-estimation structure of $\hat{\eff0}$ and $\hat{\eff1}$ implies that their sampling variance could be estimated by the standard sandwich estimator for OLS \citep{sefanskiBoosArticle}.
+If principal scores were known exactly, the M-estimation structure of $\hat{\eff0}$ and $\hat{\eff1}$ implies that their sampling variance could be estimated by the standard sandwich estimator for OLS \citep{stefanskiBoos}.
 This will rarely be the case.
@@
 \end{equation}
-where $\ri\left(\bm{\alpha}\right)=S_iZ_i+\ppip{\bm{\alpha}}$ emphasizes $\ppi$'s dependence on $\bm{\alpha}$.
+where $\ri\left(\bm{\alpha}\right)=S_iZ_i+\ppip{\bm{\alpha}}(1-Z_i)$ emphasizes $\ppi$'s dependence on $\bm{\alpha}$.
 In practice, the principal score model \eqref{eq:prinScoreMod} could be fit first, yielding estimates $\bm{\hat{\alpha}}$, and model \eqref{eq:regression0} could be fit next, using $\ri=\ri\left(\bm{\hat{\alpha}}\right)$.
@@
 
-\subsection{Relaxing Covariate Ignorability with Outcome Modeling}
+\subsection{Relaxing Covariate Ignorability with Outcome Modeling}\label{sec:regression}
 In most applications, identifying a set of covariates $\bxs$ satisfying CI will be difficult or impossible.
-Fortunately, CI may be relaxed or obviated by regression.
+Fortunately, CI may be relaxed or obviated by regression (see \citealt[][\S 3.4]{jiangDing2021} for an analogous identification result).
 Let $\bxy\in \mathbb{R}^k$, be a vector of covariates (possibly the same as $\bxs$); $\bxy$ can contain transformations and interactions of baseline covariates, allowing for a wide range of models.
@@
 \begin{prop}\label{prop:interactions}
-  Say
+  Say, for $i=1,\dots,n$, principal scores $\ppi$ are generated as \eqref{eq:pscore}, with parameters $\bm{\alpha}$ identified and estimable with M-estimation, and outcomes are generated as
   \begin{equation}\label{eq:interaction}
-    \begin{split}
-      Y_{Ci}=\beta_0^C+\bm{\beta_1}^{C'}\bxy_i+\beta_2^C\sti+\bm{\beta_3^{C'}}\bxy_i\sti+\epsilon_{Ci}\\
-      Y_{Ti}=\beta_0^T+\bm{\beta_1}^{T'}\bxy_i+\beta_2^T\sti+\bm{\beta_3^{T'}}\bxy_i\sti+\epsilon_{Ti}\\
-    \end{split}
-  %  \end{equation}
-%      +\beta_3Z_i+\beta_4\sti Z_i+\bm{\beta_5}'\bxy_i\sti+
- %   \bm{\beta_6}'\bxy Z_i+\bm{\beta_7}'\bxy_i Z_i\sti+
+    Y_i=\beta_0+\beta_1\sti+\beta_2 Z_i+\beta_3Z_i\sti+\bm{\gamma_1}'\bxy_i+\bm{\gamma_2}'\bxy_i\sti+
+    \bm{\gamma_3}'\bxy Z_i+\bm{\gamma_4}'\bxy_i Z_i\sti+\epsilon_i
   \end{equation}
-  with $\EE[\epsilon_i]=0$, $\{Z_i,\sti,\bxy_i\}\independent \epsilon_i$.
-  Say $\ppi$ linearly independent of $\bxy$, and
-  that the following model is fit with OLS
-  \begin{equation}\label{eq:fullInteractionModel}
-    Y_i=\beta_0+\bm{\beta_1}'\bxy_i+\beta_2\ri+\beta_3Z_i+\beta_4\ri Z_i+\bm{\beta_5}'\bxy_i\ri+
-    \bm{\beta_6}'\bxy Z_i+\bm{\beta_7}'\bxy_i Z_i\ri+\epsilon_i
-  \end{equation}
-  with $\ri$ as defined in \eqref{eq:ri},
-  yielding coefficient estimates $\hat{\beta}_k$, $k=0,\dots,3$.
-  Then
+  with $\EE[\epsilon_i]=0$, $\{Z_i,\sti,\bxy_i\}\independent \epsilon_i$, and $\epsilon_i\independent\epsilon_j$ when $i\ne j$.
+
+  Then, under Assumptions \ref{ass:sm}, \ref{ass:rand}, and \ref{ass:vps}, if $\ppi$ is linearly independent of $\bxy$, a researcher may follow the following procedure to estimate principal effects:
+  \begin{enumerate}
+  \item Estimate principal scores by fitting model \eqref{eq:pscore} to data from the treatment group
+  \item Replace $\sti$ with $\ri$ (as defined in \ref{eq:ri}) in model \eqref{eq:interaction} and fit with OLS
+  \item Estimate principal effects as:
    \begin{equation}\label{eq:prinEffEst}
@@
    \end{equation}
-   are M-estimators, with $\hat{\eff0}_{int}\rightarrow_p\eff0$ and $\hat{\eff1}_{int}\rightarrow_p\eff1$ as $n\rightarrow\infty$,
-   where $\overline{\bxy}_{Z=1,S=s}$ is a vector of the $k$ column means of the rows of $\bxy$ corresponding to subjects with $Z=1$ and $S=s$.
+   where $\overline{\bxy}_{Z=1,S=0}$ and $\overline{\bxy}_{Z=1,S=0}$ are the vector of covariate sample means for the subsets of subjects with $Z=1$ and $S=0$ or $S=1$, respectively.
+  \end{enumerate}
+  Then $\hat{\eff0}_{G}$ and $\hat{\eff1}_{int}$ are M-estimators, with $\hat{\eff0}_{int}\rightarrow_p\eff0$ and $\hat{\eff1}_{G}\rightarrow_p\eff1$ as $n\rightarrow\infty$.
+  Under suitible regularity conditions, they are jointly asymptotically normal, with a variance of the form \eqref{eq:sandwich}.
 \end{prop}
 
-Proposition \ref{prop:interactions} would seem to nearly solve the problem of estimating principal effects, or at least reducing it to the difficulty of other commen problems in causal modeling.
+In short, researchers can estimate principal effects under strong monotonicity by simply imputing missing $\st$ values with estimated principal scores, and fitting an OLS regression with appropriate interaction terms---just as if $\st$ were completely observed.
+
+Proposition \ref{prop:interactions} would seem to nearly solve the problem of estimating principal effects under strong monotonicity, or at least seem to reduce it to the difficulty of other common problems in causal modeling.
 Unfortunately, simulation studies presented in the appendix suggest that, although consistent estimates for principal effects may be available, these estimates are far too inefficient for practical use; standard errors will be prohibitively high in all but the largest sample sizes.
@@
 In particular, if interactions between covariates and treatment assignment are omitted, then data from the treatment group---in which $\st$ is fully observed---can be used to estimate the coefficients on covariates.
-Such a model relies on what \citet{bo2002} calls ``Additivity of Treatment Assignment Effect,'' that $\bm{\beta_1}^{C}=\bm{\beta_1}^{T}$.
+Such a model relies on what \citet{jo2002} calls ``Additivity of Treatment Assignment Effect,'' that $\bm{\gamma_3}=\bm{\gamma_4}=0$.
 In other words, the only interactions between covariates and treatment assignment are due to differences between principal effects---a fairly strong assumpton.
@@
 To summarize, incorporating covariates into an outcome regression can substantially relax covariate ignorability, but estimation is only practical if  if the regression omits interactions between covariates and $Z$.
-Standard errors can be estimated via stacked estimating equations.
 
-\section{A Simulation Study}
+\section{A Simulation Study}\label{sec:simulation}
 
@@
 
-\subsection{Simultation Design}
+\subsection{Simulation Design}
 The simulation study was conducted in \texttt{R} \citep{rcite} and Stan \citep{rstan}, and full replication code is available at \url{www.github.com/[Redacted]}.
@@
 \begin{equation*}
-  \ppi=logit^{-1}\left[\alpha(x_{1i}+x_{2i}+x_{3i})\right]
+  \ppi=logit^{-1}\left[\alpha(x_{1i}-x_{2i}+x_{3i})\right]
 \end{equation*}
-where $\alpha$ is a manipuated factor.
+where $\alpha$ is a manipulated factor.
 When $\alpha$ was higher, $\st$ was more easily predicted by covariates; when $\alpha=0$, Assumption \ref{ass:vps}, was violated, i.e. $Var(\ppi)=0$.
+Then principal stratum $\sti\in \{0,1\}$ was simulated as $S_i\sim Bern(\ppi)$
 
-Then principal stratum $\sti\in \{0,1\}$ was simulated as:
-\begin{equation*}
-S_i\sim Bern(\ppi)
-\end{equation*}
+Finally, potential and observed outcomes were generated as in \eqref{eq:interaction}, with $\bxy=[\bm{x_1},\bm{x_2},\bm{x_3}]$,
+and with $\EE[\epsilon_i]=0$ and $Var(\epsilon_i)=1/2$.
+Specifically,
+\begin{equation}\label{eq:y-sim}
+  \begin{split}
+  \yci&=\beta_1\sti+(\gamma_1+\gamma_2\sti)(x_{1i}+x_{2i})+\frac{1}{\sqrt{6}}x_{3i}+\epsilon_i\\
+  \yti&=0.3\sti+(\gamma_1+\gamma_2\sti+\gamma_3Z_i)(x_{1i}+x_{2i})+\frac{1}{\sqrt{6}}x_{3i}+\epsilon_i\\
+  Y_i&=Z_i\yti+(1-Z_i)\yci
+  \end{split}
+\end{equation}
+Hence, $\beta_0=\beta_2=0$, and $\beta_3=0.3-\beta_1$.
+When $\gamma_3=0$ (i.e.
+no interaction between $Z$ and covariates), there is no average treatment effect in the $\st=0$ stratum, the coefficient for the unobserved covariate $x_3$ is constant across treatment groups and principal strata, and we do not consider three-way interactions between $x_1$ or $x_2$, $\st$, and $Z$.
 
-Finally, potential outcomes were generated as in \eqref{eq:interaction}, with $\bxy=[\bm{x_1},\bm{x_2},\bm{x_3}]$,
-and with $\EE[\epsilon_i]=0$ and $Var(\epsilon_i)=1/2$.
-Across simulation runs, intercepts $\beta_0^C=\beta_0^T=0$ and the coefficient on the unobserved covariate $x_3$, $\beta_{1,3}^C=\beta_{1,3}^T= 1/\sqrt{6}$.
-The distribution of $\epsilon_i$ and other coefficients $\bm{\beta}$ were manipulated factors in the study.
+The distribution of $\epsilon_i$, $\beta_1$, and coefficients $\gamma_1$, $\gamma_2$, and $\gamma_3$ were manipulated factors in the study.
 
@@
 We manipulated six factors in the simulation:
-\begin{itemize}
-\item Sample size per treatment group $n\in\{100,200,\dots,1000\}$
-\item $\alpha \in \{0,0.2,0.3,\dots,1\}$
-\item $\muc1\in\{0,0.3\}$
-\item The distribution of $\epsilon_i$
-\item The presence of interactions between $\st$ and covariates
-\item The presence of interactions between $Z$ and covariates
-\end{itemize}
+
+\begin{table}
+  \centering
+  \begin{tabular}{cc}
+    \hline
+    Factor &Levels\\
+    \hline
+Sample size per treatment group $n$ &$100,200,\dots,1000$\\
+$\alpha$ &$0,0.2,0.3,\dots,1$\\
+The distribution of $\epsilon_i$& Normal, Uniform\\
+$\beta_1$ & 0, 0.3\\
+Interactions between $\st$ and covariates & No, Yes\\
+Interactions between $Z$ and covariates& No, Yes\\
+\hline
+  \end{tabular}
+  \caption{Manipulated factors in simulation study}
+  \label{tab:factor}
+  \end{table}
 The six factors are not completely crossed---in particular, we let $n$ and $\alpha$ each vary across all the levels listed while holding the other five factors fixed.
 
-The guarentees of M-estimation are all assymptotic, so we examined the behavior of our M-estimator at a range of sample sizes to determine its finite-sample properties.
-
-When $\muc1=0$ there is only one mixture component in the control
-group, which can be problematic for mixture modeling \citep{griffin2008application}. Conversely, when $\muc1=0.3$, $\eff0=\eff1=0$, so there is no treatment effect for any subjects.
+The guarentees of M-estimation are all asymptotic, so we examined the behavior of our M-estimator at a range of sample sizes to determine its finite-sample properties.
 
 As mentioned above, $\alpha$ controls the variance of the principal scores, i.e.
-the extent to which $\st$ is predictible as a function of covariates.
-When $\alpha=0$, the covariates are unrelated to $\st$ and principal scores do not vary between participants, violating \eqref{eq:vps}.
+the extent to which $\st$ is predictable as a function of covariates.
+When $\alpha=0$, the covariates are unrelated to $\st$ and principal scores do not vary between participants, violating \eqref{ass:vps}.
 We expect that the more predictive the covariates, the better the performance of all three estimators.
-To aid in the interpretation of the $\alpha$ parameter, the online appendix plots $\alpha$ against the area under the receiver operating characteristic curve for each fitted model (AUC), a common measure of classification accuracy \citep{bradley1997use}---an AUC of 0.5 implies that the model is no better than random guessing, and AUC of 1 implieds perfect prediction.
+To aid in the interpretation of the $\alpha$ parameter, the online appendix plots $\alpha$ against the area under the receiver operating characteristic curve for each fitted model (AUC), a common measure of classification accuracy \citep{bradley1997use}---an AUC of 0.5 implies that the model is no better than random guessing, and AUC of 1 implies perfect prediction.
 For $0.5\le \alpha \le 1$, the average AUC varies from roughly 0.53 to 0.77; for $\alpha=0.5$, the average AUC was roughly 0.675.
 
-$\epsilon$ is the regression error in \eqref{eq:yc-sim}.
+$\epsilon$ is the regression error in \eqref{eq:y-sim}.
 Across simulation runs, $E[\epsilon]=0$ and $Var(\epsilon)=1/2$.
@@
 
+When $\beta_1=0$ there is only one mixture component in the control
+group, which can be problematic for mixture modeling \citep{griffin2008application,feller2016principal}.
+Conversely, when $\beta_1=0.3$, unless there is an interaction between $Z$ and covariates (i.e.
+$\gamma_3\ne 0$), $\eff0=\eff1=0$, so there is no average treatment effect for either principal stratum.
+
 The last two factors---the presence of interactions between covariates and $\st$ and $Z$, respectively---test the performance of a model that does not include interactions when that modeling assumption is false.
-
-In most simulations, there was no interaction between $\st$ and $x_1$ and $x_2$; in those cases, $\beta_1=1/\sqrt{6}$, and $\beta_2=0$.
-In simulation runs including an interaction with $\st$, $\beta_1=3/(4\sqrt{6})$ and $\beta_2=5/(4\sqrt{6})$.
-
-Finally, treatment potential outcomes were generated as
-\begin{equation*}
-  \yti=\yci+(0.3-\muc1)\sti+\beta_3(x_{1i}+x_{2i})
-\end{equation*}
-and observed outcomes were $Y_i=Z_i\yti+(1-Z_i)\yci$, with even subjects assigned to treatment and odd subjects assigned to control.
-In most simulation runs, there were no interactions between treatment assignemnt and covariates, so $\beta_3=0$; when such interactions were included, $\beta_3=1/(2\sqrt{6})$.
-
-Since the covariates were all centered, $\muc0=\mut0=\eff0=0$ throughout the study.
-Similarly, $\mut1=0.3$, while $\muc1$, and hence $\eff1$, varied between runs.
-
-Note that when there were no interactions included in the model, since $Var(\beta_1(x_1+x2+x_3))=1/2$ and $Var(\epsilon_i)=1/2$, covariates explained half of the variance of $Y$ within treatment group and principal stratum.
-However, as we shall see, $x_3$ was ``unobserved,'' so observed covariates explained 1/3 of that variance.
+When neither interaction was present, $\gamma_2=\gamma_3=0$, then $\gamma_1=1/\sqrt{6}$.
+Then, since  $Var(\gamma_1(x_1+x2+x_3))=1/2$ and $Var(\epsilon_i)=1/2$, covariates explained half of the variance of $Y$ within treatment group and principal stratum, and observed covariates $x_1$ and $x_2$ explained 1/3 of that variance.
+When there was an interaction between covariates and $\sti$, then $\gamma_1=3/(4\sqrt{6})$ and $\gamma_2=1/(2\sqrt{6})$, so that the slopes for $x_1$ and $x_2$ varied by half their magnitude between principal strata, and on average they were equal to $1/\sqrt{6}$, as in the no-interaction case.
+Interactions between covariates and $Z$ were controlled by $\gamma_3$: when there were interactions, $\gamma_3=1/(2sqrt{6})$, otherwise $\gamma_3=0$.
 
@@
 
-Using each simulated dataset, we estimatinged $\eff0$ and $\eff1$ with three methods: a mixture model \eqref{eq:mixtureModelPS} including principal scores, the regression M-estimator proposed here, and third, the principal score weighting estimator \eqref{eq:psw}.
+Using each simulated dataset, we estimated $\eff0$ and $\eff1$ with three methods: a mixture model \eqref{eq:mixtureModelPS} including principal scores, GEEPERs, and third, the principal score weighting estimator \eqref{eq:psw}.
 
-The methods all used the same ``observed'' dataset, consisting of outcomes $Y$, treatment assignements $Z$, two covariates $x_1$ and $x_2$, and $S$ observed only when $Z=1$.
-The third covariate used in the data generating model, $x_3$, was ``unobserved''--since $x_3$ was correlated with both $\st$ and $Y$, Principal Ignorability (Assumption \ref{ass:pi}) was violated.
+The methods all used the same ``observed'' dataset, consisting of outcomes $Y$, treatment assignments $Z$, two covariates $x_1$ and $x_2$, and $S$ observed only when $Z=1$.
+The third covariate used in the data generating model, $x_3$, was ``unobserved''--since $x_3$ was correlated with both $\st$ and $Y$, Principal Ignorability (Assumption \ref{ass:PI}) was violated.
 
@@
 \begin{equation}\label{eq:mix1-sim}
-  f_{Y|Z=1}(y;\mut0,\mut1,\bm{\beta},\sigma_T)=\phi\left(\frac{y-[\mut0+\beta_1x_1+\beta_2x_2+(\mut1-\mut0)S_i]}{\sigma_T}\right)
+  f_{Y|Z=1}(y_i)=\phi\left(\frac{y_i-[\beta_0+\beta_1S_i+\beta_2 +\bm{\gamma}_1^T\bm{x}_{i,1:2}]}{\sigma_T}\right)
 \end{equation}
@@
 \begin{equation}\label{eq:mix-sim}
-  f_{Y|Z=0}(y_i)=fpp{\bm{\alpha}}{x_{1i},x_{2i}}\phi\left(\frac{y-\muc1-\beta_1x_1-\beta_2x_2}{\sigma_C}\right)+(1-fpp{\bm{\alpha}}{x_{1i},x_{2i}})\phi\left(\frac{y-\muc0-\beta_1x_1-\beta_2x_2}{\sigma_C}\right)
+  f_{Y|Z=0}(y_i)=\ppf{\bm{\alpha}}{\bm{x}_{i,1:2}}\phi\left(\frac{y_i-[\beta_0+\beta_1+\bm{\gamma}_1^T\bm{x}_{i,1:2}]}{\sigma_C}\right)}+(1-\ppf{\bm{\alpha}}{\bm{x}_{i,1:2}})\phi\left(\frac{y_i-[\beta_0+\bm{\gamma}_1^T\bm{x}_{i,1:2}]}{\sigma_C}\right)}
 \end{equation}
-Note that the coefficients $\beta_1$ and $\beta_2$ are the same across treatment groups and principal strata, while the residual variance differs across treatment groups, but not across principal strata.
+Note that the coefficients $\bm{\gamma}_1$ are the same across treatment groups and principal strata, while the residual variance differs across treatment groups, but not across principal strata.
 The models \eqref{eq:psModsim}, \eqref{eq:mix1-sim}, and \eqref{eq:mix-sim} were fit simultaneously using Bayesian Markov Chain Monte Carlo with Stan \citep{rstan}.
@@
 
-In contrast, for regression M-estimation and principal score weighting the principal score model \eqref{eq:psModsim} was fit by itself, to data from the treatment group, using standard maximum likelihood methods.
-For principal score weighting we extracted the estimated probabilities from \eqref{eq:psModsim}---the principal scores---plugged them in to \eqref{eq:psw} to estimate $\muc0$ and $\muc1$, and estimated $\mut0$ and $\mut1$ with the sample means of subjects with $Z=1$ and $S=0$ or $S=1$, respectively.
-
-For regression M-estimation, we used the estimated principal scores in the control group, along with the observed $S$ in the treatment group, to construct $R$, and then fit the model
+In contrast, for GEEPERs and principal score weighting the principal score model \eqref{eq:psModsim} was fit by itself, to data from the treatment group, using standard maximum likelihood methods.
+For GEEPERs, we used the estimated principal scores in the control group, along with the observed $S$ in the treatment group, to construct $R$, and then fit the model
 \begin{equation}\label{eq:simReg}
-  Y_i=\muc0+(\muc1-\muc0) \ri+(\mut0-\muc0)Z_i+(\mut1-\muc1-\mut0+\muc0)Z_iR_i+\beta_1x_{1i}+\beta_2x_{2i}+\epsilon_i
+  Y_i=\beta0+\beta_1 \ri+\beta_2Z_i+\beta_3Z_iR_i+\bm{\gamma}_1^T\bm{x}_{i,1:2}+\epsilon_i
 \end{equation}
@@
 To estimate standard errors, we used the \texttt{sandwich} package \citep{sandwich} to extract estimating equations from \eqref{eq:psModsim} and \eqref{eq:simReg} and computed the sandwich covariance matrix following the procedure detailed in the appendix.
+
+For the Bayesian mixture model and GEEPERs we estimated treatment effects as $\hat{\eff0}=\hat{\beta}_2$ and $\hat{\eff1}=\hat{\beta}_2+\hat{\beta}_3$.
+
+For principal score weighting we extracted the estimated probabilities from \eqref{eq:psModsim}---the principal scores---plugged them in to \eqref{eq:psw} to estimate $\muc0$ and $\muc1$, and estimated $\mut0$ and $\mut1$ with the sample means of subjects with $Z=1$ and $S=0$ or $S=1$, respectively.
+Since the PSW estimate did not use an outcome model, we estimated $\hat{\eff0}=\hat{\mut0}-\hat{\muc0}$ and $\hat{\eff1}=\hat{\mut1}-\hat{\muc1}$.
 
@@
 Figures \ref{fig:n} and \ref{fig:alpha} show the empirical bias and standard error (SE) of the M-estimator and mixture estimator of $\eff1$ as sample size and $\alpha$ vary.
-Across $n$ and $\alpha$, the M-estimator has lower bias and higher SE than the mixture estimator.
-The M-estimator is positively biased for low $n$ with bias converging to close to 0 (up to simulation error) for $n\ge 300$, and appears to be roughly unbiased (up to simulation error) for $0.2\le \alpha 1$, with $n$ fixed at 500.
-The SE of the M-estimator decreases rapidly for $100\le n \le 500$ and $0.2\le \alpha \le 0.5$ and more gradually for higher values of $n$ and $\alpha$.
+Across $n$ and $\alpha$, GEEPERs has lower bias and higher SE than the mixture estimator.
+GEEPERs is positively biased for low $n$ with bias converging to close to 0 (up to simulation error) for $n\ge 300$, and appears to be roughly unbiased (up to simulation error) for $\alpha \ge 0.2$, with $n$ fixed at 500.
+The SE of the GEEPERs decreases rapidly for $100\le n \le 500$ and $0.2\le \alpha \le 0.5$ and more gradually for higher values of $n$ and $\alpha$.
 
@@
 
-Figure \ref{fig:boxplots} shows violin plots, layered on top of jitters scatterplots, showing estimation error for the M-estimator, the mixture estimator, and the PSW estimator of $\eff1$ under various sets of conditions, and Table \ref{tab:rmse} shows root mean squared error (RMSE) estimates under a somewhat wider set of circumstances when $\muc1=0.3$.
-The leftmost panel of Figure \ref{fig:boxplots} and the top line of Table \ref{tab:rmse} show results under ``standard'' conditions---normal residuals and no interactions in the data generating model between covariates and either $S$ or $Z$.
-In this case, the M-estimator is roughly unbiased, but less precise than the other two methods---when $\alpha=0.2$ it is much noisier, and when $\alpha=0.5$ it is only slightly less precise.
+Figure \ref{fig:boxplots} shows violin plots, layered on top of jittered scatterplots, showing estimation error for GEEPERs, the mixture estimator, and the PSW estimator of $\eff1$ under various sets of conditions, and Table \ref{tab:rmse} shows root mean squared error (RMSE) estimates under a somewhat wider set of circumstances when $\muc1=0.3$.
+The leftmost panel of Figure \ref{fig:boxplots} and the top line of Tables \ref{tab:rmse} and \ref{tab:coverage} show results under ``standard'' conditions---normal residuals and no interactions in the data generating model between covariates and either $S$ or $Z$.
+In this case, GEEPERs is roughly unbiased, but less precise than the other two methods---when $\alpha=0.2$ it is much noisier, and when $\alpha=0.5$ it is only slightly less precise.
 The mixture estimator's bias depends on $\muc1$---when $\muc1=0$, the bias is slight, and when $\muc1=0.3$ the bias is more substantial.
-This is somewhat surprising, because when $\muc1=0$, there is no separation (``dispersion'') between the principal strata in the control group, and previous literature \citep{griffin2008application} has identified this scenario as a particular challenge for mixture modeling.
-The PSW estimator is also biased---due, presumeably, to the violation of Principal Ignorability, Assumption \ref{ass:pi}--inherent in the data generating model.
+This is somewhat surprising, because when $\muc1=0$, there is no separation between the principal strata in the control group, and previous literature \citep{griffin2008application} has identified this scenario as a particular challenge for mixture modeling.
+The PSW estimator is also biased---due, presumably, to the violation of Principal Ignorability, Assumption \ref{ass:pi}--inherent in the data generating model.
 On the other hand, the PSW estimator has easily the lowest sampling variance of the three estimators.
-In fact, Table \ref{tab:rmse} shows that despite its bias, the PSW estimator has lower RMSE than the M-estimator when $\alpha=0$ or 0.2.
-The middle panel of Figure \ref{fig:boxplots} and the fifth line of Table \ref{tab:rmse} shows a similar set of results for when the residuals in \eqref{eq:yc-sim} are uniformly distributed.
-In this case, the M-estimator and PSW estimator behave in a similiar fashion to the normal case, but the mixture estimator shows a bimodal pattern--sometimes over-estimating $\eff0$ and sometimes underestimating, but rarely estimationg $\eff1$ accurately.
+In fact, Table \ref{tab:rmse} shows that despite its bias, the PSW estimator has lower RMSE than GEEPERs when $\alpha=0$ or 0.2.
+The middle panel of Figure \ref{fig:boxplots} and the fifth line of Table \ref{tab:rmse} shows a similar set of results for when the residuals in \eqref{eq:y-sim} are uniformly distributed.
+In this case, GEEPERs and PSW estimator behave in a similiar fashion to the normal case, but the mixture estimator shows a bimodal pattern--sometimes over-estimating $\eff1$ and sometimes underestimating, but rarely estimating $\eff1$ accurately.
 
@@
 Corresponding RMSE results, along with analogous results for uniform residuals and other values for $\alpha$, are in Table \ref{tab:rmse}.
-Presence of interactions could undermine the M-estimate by inducing a relationship between principal scores and potential outcomes, even after adjusting the outcomes for covariate effects.
+Presence of interactions could undermine the GEEPERs by inducing a relationship between principal scores and potential outcomes, even after adjusting the outcomes for covariate effects.
 Fortunately, the presence of interactions in the data generating model in this scenario appear to make little difference, though RMSEs were slightly higher when interactions between $Z$ and covariates are present.
-Table \ref{tab:rmse} shows that when $\alpha=.2$, M-estimation RMSE can be substantially higher when $Z$ interactions are present than when they are not; plots in the online appendix suggest that this is due to an increase in sampling variance and kurtosis rather than bias.
+Table \ref{tab:rmse} shows that when $\alpha=.2$, GEEPERs RMSE can be substantially higher when $Z$ interactions are present than when they are not; plots in the online appendix suggest that this is due to an increase in sampling variance and kurtosis rather than bias.
 
-Table \ref{tab:coverate} shows estimates of the empirical coverage of nominal 95\% confidence intervals for the M-estimator and mixture estimator of $\eff1$, to evaluate their respective standard error estimates and the associated statistical inference.
-When $\alpha=0.5$, M-estimator confidence intervals consistently acheive their nominal levels\footnote{With 500 replications and true coverage probabilities of 0.95, the standard error for emperical coverage rates is approximately 1 percentage point.}, whereas confidence intervals from mixture estimators undercover in the presence of interactions between $Z$ and $X$, or when residuals are non-normal.
-When $\alpha=0$ or 0.2, M-estimation confidence intervals undercover in the presence of interactions between $X$ and $Z$.
-Across all conditions in the table, the M-estimation confidence intervals have higher empirical coverage than their mixture-model counterparts.
+Table \ref{tab:coverate} shows estimates of the empirical coverage of nominal 95\% confidence intervals for GEEPERs and the mixture estimator of $\eff1$, to evaluate their respective standard error estimates and the associated statistical inference.
+When $\alpha=0.5$, GEEPERs confidence intervals consistently achieve their nominal levels\footnote{With 500 replications and true coverage probabilities of 0.95, the standard error for empirical coverage rates is approximately 1 percentage point.}, whereas confidence intervals from mixture estimators undercover in the presence of interactions between $Z$ and $X$, or when residuals are non-normal.
+When $\alpha=0$ or 0.2, GEEPERs confidence intervals undercover in the presence of interactions between $X$ and $Z$.
+Across all conditions in the table, the GEEPERs confidence intervals have higher empirical coverage than their mixture-model counterparts.
 
@@
   \input{rmseTab}
-  \caption{RMSE }
+  \caption{Empirical root-mean-squared error (RMSE) estimates for GEEPERs, Bayesian mixture modeling, and PSW under varying conditions.}
+  \label{tab:rmse}
 \end{table}
@@
   \input{coverageTab}
-  \caption{Coverage}
+  \caption{Emprical coverage of nominal 95\% confidence intervals for GEEPERs and Baysian mixture modeling undery varying conditions.}
   \label{tab:coverage}
@@
 
-\section{Application: Estimating Treatment Effects for Implementation Subgroups}
+\section{Application: Estimating Treatment Effects for Implementation Subgroups}\label{sec:fh2t}
 
 In a recent educational field experiment, middle school students completing their math schoolwork on a computer program were randomized between four conditions.
-In all four conditions, students worked on their assigned math program in class on 12 different occassions---a prior assessmenent, a mid-test, a post-test, and nine learning sessions---spaced throughout the school year.
+In all four conditions, students worked on their assigned math program in class on 12 different occasions---a prior assessment, a mid-test, a post-test, and nine learning sessions---spaced throughout the school year.
 In three of the conditions, during the learning sessions, students worked through a series of math problems, while the forth condition was a computer game designed to teach math concepts.
@@
 The literature on hints in online learning is mixed \citep{stuff}, in particular when it comes to requesting a bottom-out hint (called ``bottoming out'').
-On the one hand, hints that include the correct answer are essentially worked examples, which can be beneficial to learning \citep{worked example stuff}.
+On the one hand, hints that include the correct answer are essentially worked examples, which can be beneficial to learning \citep{workedexamplestuff}.
 On the other hand, they allow students to ``game the system,'' by simply requesting enough hints to see the right answer, and use the software without ever working to solve any problems.
@@
 Students with missing pretest scores were dropped from the analysis; missing data in other covariates was imputed with a Random Forest algorithm \citep{missForest}.
-Table \ref{tab:tab1} describes each covariate--as well as the sucess of the randomForest imputation algorithm.
-
-\begin{table}
-  \centering
-  \input{table1}
-  \caption{Descriptive statistics on baseline covariates, $S$, and $Y$ for the analysis data.
+%Table \ref{tab:tab1} describes each covariate--as well as the sucess of the randomForest imputation algorithm.
+%% \begin{comment}
+%% \begin{table}
+%%   \centering
+%%   \input{table1}
+%%   \caption{Descriptive statistics on baseline covariates, $S$, and $Y$ for the analysis data.
 Values are mean (SD) unless otherwise noted.
 Imputation error is the ``out of box'' error estimate from the Random Forest imputation algorithm---root mean squared error divided by the observed standard deviation for continuous variables and misclassification proportion for categorical variables.}
-  \label{tab:tab1}
-\end{table}
+%%   \label{tab:tab1}
+%% \end{table}
+%% \end{comment}
+%% u
 
@@
 Finally, we also included data collected from students at the beginning of the study: the natural logrithm of the  time spent on task during the pretest, scores on baseline measurements of mathematics anxiety and self-efficacy and perceputal sensitivity, along with a missingness indicator for these measurements.
-Binned residual plots \citep{gelman2006data} from an initial model fit indicated a curviliniar relationship between percentual sensitivity measurements (modeled as linear) and model errors, so in the final model we included perceptual sensitivity as a quadratic polynomial.
+Binned residual plots \citep{gelman2006data} from an initial model fit indicated a curviliniar relationship between perceptual sensitivity measurements (modeled as linear) and model errors, so in the final model we included perceptual sensitivity as a quadratic polynomial.
 The model was fit to data from students in the treatment group, for whom bottom-out status ($S$) was observed; hence, the same model fit was used for all three treatment contrasts.
@@
 Second, the asymptotics of logistic regression are more restricted than OLS, and so logistic regression tends to perform poorly when the number of parameters increases with the sample size \citep{agresti}.}
-The two sets of esimated differed in that the first assumed that coefficients on covariates were equal across principal strata, whereas the second set of results allowed the coefficients on pretest and 5th-grade standardized test scores---the most important predictors of outcomes and principal strata membership, respectively---to vary between strata.
+The two sets of estimated differed in that the first assumed that coefficients on covariates were equal across principal strata, whereas the second set of results allowed the coefficients on pretest and 5th-grade standardized test scores---the most important predictors of outcomes and principal strata membership, respectively---to vary between strata.
 
@@
 Estimated effects compared to BAU or DragonBox were slightly lower for bottom-outers than for non-bottom-outers, and estimated effects compared to FH2T were higher for non-bottom-outers.
-However, none of these differences is statistically significant, so opposite patterns, or no difference betweeen principal effects at all, would also be consistant with the data.
+However, none of these differences is statistically significant, so opposite patterns, or no difference between principal effects at all, would also be consistent with the data.
 In fact, 95\% confidence intervals for all principal effects include both positive and negative principal effects.
-There is no discernable difference between estimated principal effects when coefficients on pretest and grade-5 standardized tests were or were not allowed to vary between strata, though standard errors from the models including an interaction were somewhat higher.
+There is no discernible difference between estimated principal effects when coefficients on pretest and grade-5 standardized tests were or were not allowed to vary between strata, though standard errors from the models including an interaction were somewhat higher.
 
@@
 
-\section{Discussion}
+\section{Discussion}\label{sec:discussion}
+GEEPERs is a straightforward approach to principal effect estimation under strong monotonicity, built on widely-used regression models, which is more robust---though less precise---than alternative approaches under a wide array of scenarios.
 
+There is good reason to hope that extensions to GEEPERs may be straightforward.
+First, GEEPERs may be extended to situations in which there is two-way noncompliance, and only a weaker form of monotonicity (Assumption \ref{ass:sm}) holds.
+For instance, in truncation by death problems \citep{e.g.][]{zhangRubin,ding2011}, $S_i=1$ if participant $i$ survives (or, more generally, if the outcome $Y_i$ is measured) and 0 otherwise, and interest is is in the principal effect for the ``always survive'' principal stratum in which $\sti=S_{ci}=1$.
+If, say, $\sti\ge S_{Ci}$, then every subject in the control condition with $S_i=1$ is in the always survive stratum, while those subjects in the treatment condition with $S_i=1$ are a mixture of the always survive and $\sti=1;\;S_{Ci}=0$ stratum.
+The scenario of truncation by death with (weak) monotonicity is broadly similar to the strong monotonicity, one-way noncompliance situation that this paper discussed.
+On the other hand, when no monotonicity assumption holds, both $\st$ and $S_C$ will have to be imputed for every subject; further research is necessary to determine the appropriate way to do so.
+
+Another direction of extension involved the principal score model \eqref{eq:pscore}.
+The performance of GEEPERs in the simulation study of Section \ref{sec:simulation} depended heavily on the factor $\alpha$, which controlled the extent to which covariates could predict $S$.
+That suggests that when covariates are high-dimensional, GEEPERs' performance in applications could be optimized with a high-dimensional semi- or non-parametric model.
+If so, several further questions emerge.
+First, can the parameter $\alpha$ be extended to a more general parameter measuring the prediction accuracy of a non-parametric model.
+Second, if the principal score model cannot be formulated as the solution to a set of estimating equations, how should the standard error be computed?
+Lastly, can over-fit principal score models cause bias or other estimation problems, and if so, are there ways to protect against overfitting?
+
+GEEPERs is a flexible, easily-implementable method for principal effect estimation for one-way noncompliance, with predictive covariates; extensions to a broader set of circumstances could be a boon to causal modeling.
+
+\bibliographystyle{unsrtnat}
+\bibliography{MOM}
 \appendix
+
+[NOTE: I NEED TO MAKE SURE THE NOTATION HERE MATCHES ABOVE]
 \section{Proofs and Calculations}
@@
 \end{equation*}
-the result follows
+the result follows.
+
+\subsection{Sandwich Matrix Calculations}
+
+So we're estimating principal effects in two stages.
+First, fit the model
+\begin{equation}\label{eq:psMod}
+  Pr(S_{Ti}=1|\bm{\tilde{w}}_i)=f(\bm{\alpha}'\bm{\tilde{w}}_i)
+\end{equation}
+for some inverse link function $f(\cdot)$, where $\bm{\tilde{x}}_i=[1,\bm{w}_i]$, using (observed) values from the treatment group, and estimating $\hat{\alpha}$.
+Then let
+\begin{equation}\label{eq:ps}
+  \hat{p}_i=f(\bm{\hat{\alpha}}'\bm{\tilde{w}}_i)
+\end{equation}
+for all subjects in the experiment.
+
+Finally, fit model
+\begin{equation}\label{eq:regression}
+  Y_i=\beta_0+\beta_1r_i+\beta_2Z_i+\beta_3Z_ir_i+\bm{\gamma}^t\bm{x}_i+\epsilon_i
+\end{equation}
+to estimate $\bm{\beta}$ and hence principal effects, where $\bm{x}_i$ is a set of covariates predictive of $Y$ within principal strata.
+[Note that \eqref{eq:regression} is just \eqref{eq:regression0} with $\bm{\gamma}^t\bm{x}_i$ added on the end].
+$\bm{x}_i$ should, generally, contain $\bm{w}_i$; the two may also be equivalent.
+Let $\bm{\beta}=[\beta_0,\beta_1,\beta_2,\beta_3,\bm{\gamma}^t]^t$.
+
+The standard error matrix for $\bm{\hat{\beta}}$ can be estimated as:
+\begin{equation}\label{eq:sandwich}
+  \widehat{var}(\bm{\hat{\beta}})=A^{-1}BA^{-t}
+\end{equation}
+Where
+\begin{equation}\label{eq:Amat}
+  A=\sum_i\frac{\partial}{\partial [\alpha,\beta]^t} \Psi(Y_i;\alpha,\beta,\bm{x}_i,\bm{w}_i)
+\end{equation}
+and
+\begin{equation}\label{eq:Bmat}
+  B=\sum_i \Psi(Y_i;\alpha,\beta,\bm{x}_i)\Psi(Y_i;\alpha,\beta,\bm{x}_i,\bm{w}_i)^t
+\end{equation}
+where $\Psi(\cdot)$ is the estimating equations.
+
+Following \citet[][p. 373]{carroletal06}, we %separate the estimating equations $\Psi$ into $\phi$ and $\psi$, estimating equations for models \eqref{
+can decompose the matrices into diagonal elements $A_{1,1}$ and $A_{2,2}$ and $B_{1,1}$ and $B_{2,2}$ that pertain to the parameter sets $\alpha$ and $\beta$ and the estimating equations for models \eqref{eq:psMod} and  \eqref{eq:regression}, respectively, and $A_{21}$, and $B_{12}=B_{21}^t$, which capture to the dependence of model \eqref{eq:regression} on the parameters $\alpha$ from \eqref{psMod} and the covariance between the estimating equations of the two models.
+The sub-matrix $A_{12}=0$, since \eqref{eq:psMod} does not depend on $\beta$.
+
+The diagonal matrices $A_{1,1}$ and $A_{2,2}$ and $B_{1,1}$ and $B_{2,2}$ are all the typical ``bread'' and ``meat'' matrices from M-estimation of generalized linear models and OLS.
+In practice, we use the estimates from the \texttt{sandwich} package in \texttt{R}, adjusted in two ways: first, the function \texttt{bread} actually gives $A^{-1}$, not $A$; second, we must pay careful attention to sample sizes, since the sample size for \eqref{eq:psMod} includes only treated observations (with observed $S$) and \eqref{eq:regression} contains all observations.
+
+\subsection{$A_{21}$ Matrix}
+The estimating equations for the regression \eqref{eq:regression} are
+\begin{equation}\label{eq:eeOLS}
+  \psi(Y_i,\bm{x}_i,\bm{w}_i,\beta,\alpha)=X_iY_i-X_iX_i^t\beta
+\end{equation}
+Where $X_i=[1,r_i,Z_i,r_iZ_i,\bm{x}_i^t]^t$.
+In other words,
+\begin{align*}
+  \psi(Y_i,&\bm{x}_i,\bm{w}_i,\beta,\alpha)=\\
+  &\left\{
+  \begin{array}{l}
+    Y_i-\left(\beta_0+\beta_1r_i+\beta_2Z_i+\beta_3Z_ir_i+\bm{\gamma}'\bm{x}_i\right)\\
+    r_iY_i-r_i\left(\beta_0+\beta_2Z_i+\bm{\gamma}'\bm{x}_i\right)-r_i^2\left(\beta_1+\beta_3Z_i\right)\\
+    Z_iY_i-Z_i\left(\beta_0+\beta_1r_i+\beta_2Z_i+\beta_3Z_ir_i+\bm{\gamma}'\bm{x}_i\right)\\
+    Z_ir_iY_i-Z_ir_i\left(\beta_0+\beta_2Z_i+\bm{\gamma}'\bm{x}_i\right)-Z_ir_i^2\left(\beta_1+\beta_3Z_i\right)\\
+    \bm{x}_iY_i-\bm{x}_i\left(\beta_0+\beta_1r_i+\beta_2Z_i+\beta_3Z_ir_i+\bm{\gamma}'\bm{x}_i\right)
+  \end{array}
+  \right\}
+\end{align*}
+(noting that $Z^2=Z$).
+These depend on $\alpha$ when $r_i=p_i$, i.e. when $Z_i=0$ or $S_i$ is missing.
+Then note that, following \eqref{eq:ps}, and letting $\eta_i=\alpha^t\bm{\tilde{w}}_i$
+\begin{equation}\label{eq:derivP}
+  \frac{\partial p_i}{\partial \alpha^t}=f'(\eta_i)\frac{\partial \eta_i}{\partial \alpha^t}=f'(\eta_i)\bm{\tilde{w}}_i^t
+\end{equation}
+and that
+\begin{equation}
+  \frac{\partial p_i^2}{\partial \alpha^t}=2p_i\frac{\partial p_i}{\partial \alpha^t}=2f(\eta)f'(\eta_i)\bm{\tilde{w}}_i^t=2p_if'(\eta_i)\bm{\tilde{w}}_i^t
+\end{equation}
+
+Then if $r_i=p_i$,
+\begin{align*}
+  \frac{\partial}{\partial \alpha^t}&\psi(Y_i,\bm{x}_i,\beta,\alpha)=\\
+  &\left[\begin{array}{c}
+          -(\beta_1+\beta_3Z_i)\frac{\partial p_i}{\partial \alpha^t}\\
+          (Y_i-\beta_0-\beta_2Z_i-\bm{\gamma}^t\bm{x}_{i})\partial p/\partial\alpha^t-(\beta_1+\beta_3Z_i)\partial p^2/\partial \alpha^t\\
+          -(\beta_1+\beta_3Z_i)\frac{\partial p_i}{\partial \alpha^t}Z_i\\
+          Z_i\left[(Y_i-\beta_0-\beta_2Z_i-\bm{\gamma}^t\bm{x}_{i})\partial p/\partial\alpha^t-(\beta_1+\beta_3Z_i)\partial p^2/\partial \alpha^t\right]\\
+          -(\beta_1+\beta_3Z_i)\frac{\partial p_i}{\partial \alpha^t}\bm{x}_i
+        \end{array}\right]\\
+  =& \left[\begin{array}{c}
+          -(\beta_1+\beta_3Z_i)f'(\eta_i)\bm{\tilde{w}}_i^t\\
+          \left[Y_i-\beta_0-\beta_2Z_i-\bm{x}_i^t\bm{\gamma}-2p_i(\beta_1+\beta_3Z_i)\right]f'(\eta_i)\bm{\tilde{w}}_i^t\\
+          -Z_i\beta_1f'(\eta_i)\bm{\tilde{w}}_i^t\\
+          Z_i\left[Y_i-\beta_0-\beta_2Z_i-\bm{x}_i^t\bm{\gamma}-2p_i(\beta_1+\beta_3)\right]f'(\eta_i)\bm{\tilde{w}}_i^t\\
+          -(\beta_1+\beta_3Z_i)f'(\eta_i)\bm{x}_i\bm{\tilde{w}}_i^t
+    \end{array}\right]
+\end{align*}
+
+When $S$ is observed for all members of the treatment group, $Z_i=0$ whenever $r_i=p_i$, so the latter expression reduces to
+\begin{align*}
+  \frac{\partial}{\partial \alpha^t}&\psi(Y_i,\bm{x}_i,\beta,\alpha)=\\
+  & \left[\begin{array}{c}
+          -(\beta_1)f'(\eta_i)\bm{\tilde{w}}_i^t\\
+          \left[Y_i-X_i^t\beta-2p_i(\beta_1)\right]f'(\eta_i)\bm{\tilde{w}}_i^t\\
+          0\\
+          0\\
+          -\beta_1f'(\eta_i)\bm{x}_i\bm{\tilde{w}}_i^t
+    \end{array}\right]
+\end{align*}
+
+If $r_i=S_i$, $\frac{\partial}{\partial \alpha^t}\psi(Y_i,\bm{x}_i,\beta,\alpha)=0$.
 

### 2022-09-30 writeUps/rmseTab.tex
update paper

\begin{tabular}[t]{lllrrrrrrrrr}
\toprule
\multicolumn{3}{c}{ } & \multicolumn{3}{c}{$\alpha=0$} & \multicolumn{3}{c}{$\alpha=0.2$} & \multicolumn{3}{c}{$\alpha=0.5$} \\
\cmidrule(l{3pt}r{3pt}){4-6} \cmidrule(l{3pt}r{3pt}){7-9} \cmidrule(l{3pt}r{3pt}){10-12}
\makecell[l]{Residual\\Dist.} & \makecell[c]{X:Z\\Int.?} & \makecell[r]{X:S\\Int.?} & GEEPERs & Mix. & PSW & GEEPERs & Mix. & PSW & GEEPERs & Mix. & PSW\\
\midrule
 & No & No & 2.07 & 0.19 & 0.17 & 0.55 & 0.20 & 0.20 & 0.19 & 0.16 & 0.24\\

 & Yes & No & 9.34 & 0.34 & 0.16 & 1.06 & 0.33 & 0.21 & 0.23 & 0.18 & 0.23\\

 & No & Yes & 2.56 & 0.19 & 0.16 & 0.51 & 0.22 & 0.21 & 0.19 & 0.15 & 0.24\\

\multirow{-4}{*}{\raggedright\arraybackslash Normal} & Yes & Yes & 7.16 & 0.34 & 0.17 & 1.81 & 0.33 & 0.20 & 0.23 & 0.20 & 0.24\\
\cmidrule{1-12}
 & No & No & 4.19 & 0.56 & 0.16 & 0.60 & 0.50 & 0.20 & 0.18 & 0.31 & 0.24\\

 & Yes & No & 6.99 & 0.60 & 0.17 & 0.91 & 0.57 & 0.20 & 0.22 & 0.35 & 0.24\\

 & No & Yes & 1.83 & 0.52 & 0.17 & 0.45 & 0.46 & 0.20 & 0.19 & 0.31 & 0.23\\

\multirow{-4}{*}{\raggedright\arraybackslash Uniform} & Yes & Yes & 7.55 & 0.60 & 0.18 & 1.28 & 0.56 & 0.20 & 0.24 & 0.35 & 0.24\\
\bottomrule
\end{tabular}

### 2022-09-30 code/fh2t/psMods.r

@@ -1,0 +1,223 @@
+library(arm)
+
+load('data/psdat.RData')
+
+psdat <- droplevels(filter(psdat,SchIDPre!=7))
+
+auc <- function(x,y)
+  if(length(unique(y))==2 & length(y)==length(x)){
+    wilcox.test(x~y)$statistic/(sum(y)*(sum(1-y)))
+  }else
+    wilcox.test(x,y)$statistic/(legnth(x)*length(y))
+
+aucMod <- function(mod)
+  auc(mod$linear,1-mod$y)
+
+
+psMod1 <- glm(
+  S~.-StuID-TeaIDPre-ClaIDPre-S-Y-virtual,
+  data=psdat%>%filter(trt=='ASSISTments')%>%select(-trt),
+  family=binomial)
+
+aucMod(psMod1)
+boxplot(fitted(psMod1)~psMod1$y)
+
+
+
+
+summary(psMod2 <- step(psMod1))
+aucMod(psMod2)
+
+psMod3 <- update(psMod2,.~.-SchIDPre+TeaIDPre)
+aucMod(psMod3)
+
+### what about lumping teachers with fewer than 10 kids, within school?
+psdat <- psdat%>%
+  group_by(SchIDPre)%>%
+  mutate(maxTch=levels(TeaIDPre)[which.max(table(TeaIDPre))])%>%
+  group_by(TeaIDPre)%>%
+  mutate(nTrt=sum(trt=='ASSISTments'))%>%
+  ungroup()%>%
+  mutate(tid2=factor(ifelse(nTrt>9,as.character(TeaIDPre),maxTch)))
+
+psMod4 <- update(psMod3,.~.-TeaIDPre+tid2)
+aucMod(psMod4)
+
+### but is it just overfit?
+### 10-fold cv
+cv <- function(mod,folds=10,seed=613){
+  mf=mod$data[names(mod$y),]
+  set.seed(seed)
+  folds <- seq(nrow(mf))%>%sample()%>%cut(10,labels=FALSE)
+  cvPred <- numeric(nrow(mf))
+  for(ff in 1:10)
+    cvPred[folds==ff] <- predict(update(mod,data=mf[folds!=ff,]),mf[folds==ff,])
+  AUC <-auc(cvPred,1-mod$y)
+  print(AUC)
+  invisible(list(cvPred=cvPred,folds=folds,auc=AUC))
+}
+
+cv(psMod2)
+cv(psMod4)
+
[…]

### 2022-12-05 writeUps/paper.tex
update writeup

@@
 \begin{prop}\label{prop:reg1}
-  Under Assumptions \ref{ass:sm}, \ref{ass:rand}, \ref{ass:vps}, and \ref{ass:ci}, if %principal scores are known exactly,
+  Under Assumptions \ref{ass:sm}, \ref{ass:rand}, \ref{ass:vps}, \ref{ass:ci}, and suitible regularity conditions \citep[e.g.][p. 327]{boosStefanskiBook}, if %principal scores are known exactly,
   $\ri$ is defined as in \eqref{eq:ri} and the model
@@
 \begin{equation}\label{eq:mix-sim}
-  f_{Y|Z=0}(y_i)=\ppf{\bm{\alpha}}{\bm{x}_{i,1:2}}\phi\left(\frac{y_i-[\beta_0+\beta_1+\bm{\gamma}_1^T\bm{x}_{i,1:2}]}{\sigma_C}\right)}+(1-\ppf{\bm{\alpha}}{\bm{x}_{i,1:2}})\phi\left(\frac{y_i-[\beta_0+\bm{\gamma}_1^T\bm{x}_{i,1:2}]}{\sigma_C}\right)}
+  f_{Y|Z=0}(y_i)=\ppf{\bm{\alpha}}{\bm{x}_{i,1:2}}\phi\left(\frac{y_i-[\beta_0+\beta_1+\bm{\gamma}_1^T\bm{x}_{i,1:2}]}{\sigma_C}\right)+(1-\ppf{\bm{\alpha}}{\bm{x}_{i,1:2}})\phi\left(\frac{y_i-[\beta_0+\bm{\gamma}_1^T\bm{x}_{i,1:2}]}{\sigma_C}\right)
 \end{equation}
@@
 First, GEEPERs may be extended to situations in which there is two-way noncompliance, and only a weaker form of monotonicity (Assumption \ref{ass:sm}) holds.
-For instance, in truncation by death problems \citep{e.g.][]{zhangRubin,ding2011}, $S_i=1$ if participant $i$ survives (or, more generally, if the outcome $Y_i$ is measured) and 0 otherwise, and interest is is in the principal effect for the ``always survive'' principal stratum in which $\sti=S_{ci}=1$.
+For instance, in truncation by death problems \citep[e.g.][]{zhangRubin,ding2011}, $S_i=1$ if participant $i$ survives (or, more generally, if the outcome $Y_i$ is measured) and 0 otherwise, and interest is is in the principal effect for the ``always survive'' principal stratum in which $\sti=S_{ci}=1$.
 If, say, $\sti\ge S_{Ci}$, then every subject in the control condition with $S_i=1$ is in the always survive stratum, while those subjects in the treatment condition with $S_i=1$ are a mixture of the always survive and $\sti=1;\;S_{Ci}=0$ stratum.
@@
 
-\subsection{Proof for Proposition \ref{prop:reg0}}
-Carrying out those replacements in \eqref{eq:estEq0}, and rearranging\footnote{Technically, replacing $\tilde{\Psi}_i$ with $\Psi_i=\begin{psmallmatrix} 1 & 0&1&0\\ 0&1&0&1\\0&0&1&0\\0&0&0&1\end{psmallmatrix}\tilde{\Psi}_i$.} gives an equivalent set of estimating equations:
+\subsection{Proof for Proposition \ref{prop:reg1}}
+replacing $\sti$ and $\pp$ in \eqref{eq:estEq0} with $\ri$, as in \eqref{eq:ri}, and replacing $\tilde{\Psi}_i$ with $\Psi_i=\begin{psmallmatrix} 1 & 0&1&0\\ 0&1&0&1\\0&0&1&0\\0&0&0&1\end{psmallmatrix}\tilde{\Psi}_i$ gives an equivalent set of estimating equations:
 \begin{equation}\label{eq:estEq1}
@@
 \end{equation}
+These are equivalent to the estimating equations for OLS model \eqref{eq:regression0} with $\beta_0=\muc0$, $\beta_1=\muc1-\muc0$, $\beta_2=\mut0-\muc0$, and $\beta_3=\mut1-\mut0-\muc1-\muc0$.
+Therefore, under standard OLS regularity conditions the estimated parameter vector $\bm{\hat{\beta}}$ is consistent, completing the proof.
 
-\subsection{Proof for Proposition \ref{prop:interaction}}
+\subsection{Proof for Proposition \ref{prop:interactions}}
 First, note that $\EE[S]=\EE\EE[S|\bx]=\EE[\pp]$ and
-$$\EE[\bx S]=\EE[\bx S|Z=1]=\EE[\bx S|Z=0]=\EE[\bx \EE[\bx\EE[S|\bx]|Z=0]]=\EE[\bx\pp|Z=0]=\EE[\bx\pp]$$
+\begin{align*}
+  \EE[\bx S]&=\EE[\bx S|Z=1]=\EE[\bx S|Z=0] \mbox{ due to \eqref{ass:rand}}\\
+  &=\EE[\bx\EE[S|\bx]|Z=0]=\EE[\bx\pp|Z=0]=\EE[\bx\pp]
+\end{align*}
+
 implying that

### 2022-12-05 writeUps/paper.tex
Updates from Overleaf

--- 2022-09-30 writeUps/paper.tex
@@
 \input{notation.tex}
+
+%%% TO DO
+%% Add detail to application section
+%% fill in missing citations
+%% Appendix
+%%   clean up/fix notation for math
+%%   add in full simulation results
+%%   small interaction simulation
+%%   look for what else I promised and either put it in or delete promise
+%% proofread
 
@@
 This paper expands on those identification results by presenting a new set of principal effect estimators, and associated standard error estimates, based on a generalized estimating equations (or M-estimation or method-of-moments) \citep{stefanskiBoos}.
-It turns out that, under certain general conditions, principal effects can be estimated with a simple, two-step procedure: first, estimate principal scores with logistic regression or another M-estimator; then, impute unknown principal subroup membership indicators with principal scores and estimate principal effects using OLS with interactions, just as in the pre-treatment subgroup case.
+It turns out that, under certain general conditions, principal effects can be estimated with a simple, two-step procedure: first, estimate principal scores with logistic regression or another M-estimator; then, impute unknown principal subgroup membership indicators with principal scores and estimate principal effects using OLS with interactions, just as in the pre-treatment subgroup case.
 We will refer to the estimators presented here as ``GEEPERs'': generalized estimating equations for principal effects using regressions.
@@
 \subsubsection{Manipulated Factors}
-We manipulated six factors in the simulation:
+We manipulated six factors in the simulation, listed in Table \ref{tab:factor}.
+The six factors are not completely crossed---in particular, we let $n$ and $\alpha$ each vary across all the levels listed while holding the other five factors fixed.
 
@@
     \hline
-Sample size per treatment group $n$ &$100,200,\dots,1000$\\
+Sample size per condition $n$ &$100,200,\dots,1000$\\
 $\alpha$ &$0,0.2,0.3,\dots,1$\\
@@
   \end{table}
-The six factors are not completely crossed---in particular, we let $n$ and $\alpha$ each vary across all the levels listed while holding the other five factors fixed.
 
-The guarentees of M-estimation are all asymptotic, so we examined the behavior of our M-estimator at a range of sample sizes to determine its finite-sample properties.
+\textbf{\emph{Sample Size per condition $\bm{n}$:}} The guarantees of M-estimation are all asymptotic, so we examined the behavior of our M-estimator at a range of sample sizes to determine its finite-sample properties.
 
-As mentioned above, $\alpha$ controls the variance of the principal scores, i.e.
+$\bm{\alpha}$\textbf{:} As mentioned above, $\alpha$ controls the variance of the principal scores, i.e.
 the extent to which $\st$ is predictable as a function of covariates.
@@
 
-$\epsilon$ is the regression error in \eqref{eq:y-sim}.
+\textbf{\emph{The distribution of $\bm{\epsilon_i}$:}} $\epsilon$ is the regression error in \eqref{eq:y-sim}.
 Across simulation runs, $E[\epsilon]=0$ and $Var(\epsilon)=1/2$.
 However, the shape of $\epsilon$'s distribution varied between the following two possibilities:
-\begin{itemize}
-\item[Normal] $\epsilon\sim\mathcal{N}(0,1/2)$, as assumed by the mixture model estimator
-\item[Uniform] $\epsilon\sim\mathcal{U}(-\sqrt{6}/2,\sqrt{6}/2)$
+Normal $\epsilon\sim\mathcal{N}(0,1/2)$, as assumed by the mixture model estimator, or Uniform $\epsilon\sim\mathcal{U}(-\sqrt{6}/2,\sqrt{6}/2)$.
 %\item[Mixture] $f_\epsilon(x)=\frac{1}{4}\phi\left((x+1/3)\sqrt{6}\right)+\frac{1}{4}\phi\left((x-1)\sqrt{6})\right)$
-\end{itemize}
 %The mixture distribution for $\epsilon$ was designed to be particularly difficult for the mixture model estimator.
 
-When $\beta_1=0$ there is only one mixture component in the control
+$\bm{\beta_1}$\textbf{:} When $\beta_1=0$ there is only one mixture component in the control
 group, which can be problematic for mixture modeling \citep{griffin2008application,feller2016principal}.
@@
 
-The last two factors---the presence of interactions between covariates and $\st$ and $Z$, respectively---test the performance of a model that does not include interactions when that modeling assumption is false.
+\textbf{\emph{Interactions between $\st$ or $Z$ and covariates:}} The last two factors---the presence of interactions between covariates and $\st$ and $Z$, respectively---test the performance of a model that does not include interactions when that modeling assumption is false.
 When neither interaction was present, $\gamma_2=\gamma_3=0$, then $\gamma_1=1/\sqrt{6}$.
@@
 
-Table \ref{tab:coverate} shows estimates of the empirical coverage of nominal 95\% confidence intervals for GEEPERs and the mixture estimator of $\eff1$, to evaluate their respective standard error estimates and the associated statistical inference.
+Table \ref{tab:coverage} shows estimates of the empirical coverage of nominal 95\% confidence intervals for GEEPERs and the mixture estimator of $\eff1$, to evaluate their respective standard error estimates and the associated statistical inference.
 When $\alpha=0.5$, GEEPERs confidence intervals consistently achieve their nominal levels\footnote{With 500 replications and true coverage probabilities of 0.95, the standard error for empirical coverage rates is approximately 1 percentage point.}, whereas confidence intervals from mixture estimators undercover in the presence of interactions between $Z$ and $X$, or when residuals are non-normal.
@@
 \begin{table}
+\centering
   \input{coverageTab}
-  \caption{Emprical coverage of nominal 95\% confidence intervals for GEEPERs and Baysian mixture modeling undery varying conditions.}
+  \caption{Empirical coverage of nominal 95\% confidence intervals for GEEPERs and Bayesian mixture modeling under varying conditions.}
   \label{tab:coverage}
@@
 
-We also had access to data on a number of baseline covariates: total scores on a 10-item pretest, fifth-grade stanardized test scores, race/ethnicity (white, Asian, Hispanic/Latino, or Other), indicator variables for if students have an early intervention program (EIP) or individualized educational plan (IEP) in place, if students are learning English as a second language (ESOL), students' average time on task for the pretest (logged), total scores from baseline evaluations of students' math anxiety, negative reaction to math, and numerical confidence, and an indicator for weather students began the school year in remote or in-person instruction.
+We also had access to data on a number of baseline covariates: total scores on a 10-item pretest, fifth-grade standardized test scores, race/ethnicity (white, Asian, Hispanic/Latino, or Other), indicator variables for if students have an early intervention program (EIP) or individualized educational plan (IEP) in place, if students are learning English as a second language (ESOL), students' average time on task for the pretest (logged), total scores from baseline evaluations of students' math anxiety, negative reaction to math, and numerical confidence, and an indicator for weather students began the school year in remote or in-person instruction.
 Students with missing pretest scores were dropped from the analysis; missing data in other covariates was imputed with a Random Forest algorithm \citep{missForest}.
@@
 indicator variables for special education status (IEP), whether the student was taking English as a second language (ESOL), whether the student was classified as gifted, and whether students attended a full year of school in 5th grade.
-Finally, we also included data collected from students at the beginning of the study: the natural logrithm of the  time spent on task during the pretest, scores on baseline measurements of mathematics anxiety and self-efficacy and perceputal sensitivity, along with a missingness indicator for these measurements.
-Binned residual plots \citep{gelman2006data} from an initial model fit indicated a curviliniar relationship between perceptual sensitivity measurements (modeled as linear) and model errors, so in the final model we included perceptual sensitivity as a quadratic polynomial.
+Finally, we also included data collected from students at the beginning of the study: the natural logarithm of the  time spent on task during the pretest, scores on baseline measurements of mathematics anxiety and self-efficacy and perceptual sensitivity, along with a missingness indicator for these measurements.
+Binned residual plots \citep{gelman2006data} from an initial model fit indicated a curvilinear relationship between perceptual sensitivity measurements (modeled as linear) and model errors, so in the final model we included perceptual sensitivity as a quadratic polynomial.
 The model was fit to data from students in the treatment group, for whom bottom-out status ($S$) was observed; hence, the same model fit was used for all three treatment contrasts.
@@
 Table \ref{tab:results} show the model's coefficient estimates.
-Noteably, conditional on the other covariates, students with higher fifth-grade standardized test scores were less likely to request many bottom out hints, as were students who spent little time on task during the pretest, while IEP and ESOL students, and students without baseline measurement data, were relatively unlikely to be bottom-outers.
+Notably, conditional on the other covariates, students with higher fifth-grade standardized test scores were less likely to request many bottom out hints, as were students who spent little time on task during the pretest, while IEP and ESOL students, and students without baseline measurement data, were relatively unlikely to be bottom-outers.
 The relationship between perceptual sensitivity and bottom out requests was not monotonic---students with average levels were more likely than those with more extreme high or low levels to be bottom-outers; bottom-outers tended to have lower math self-efficacy, but the data are also consistent with a small positive relationship, and there appears to be little to no relationship between math anxiety and bottom-outer status at all.
@@
 
-Figure \ref{fig:effects} plots the estimated principal effects for bottom-outers and non-bottom-outers, using two different models---alongside estimates of the overall average treatment effect, labeled ``ATE.'' In both models, effects were adjusted for all of the variables that were in the principal score model, in addition to students' pretest scores; also, the models substituted class fixed effects for school fixed effects.\footnote{Why use a corser set of fixed effects for the principal score model than the outcome regression?
+Figure \ref{fig:effects} plots the estimated principal effects for bottom-outers and non-bottom-outers, using two different models---alongside estimates of the overall average treatment effect, labeled ``ATE.'' In both models, effects were adjusted for all of the variables that were in the principal score model, in addition to students' pretest scores; also, the models substituted class fixed effects for school fixed effects.\footnote{Why use a coarser set of fixed effects for the principal score model than the outcome regression?
 There were two reasons: first, the principal score model was fit to a smaller sample (only students in the ``Instant'' condition), so some classrooms only included one student, which can lead to poor fit.
@@
 All model parameters with support in $\mathbb{R}$ were given standard normal priors, while standard deviation parameters were given half-standard normal priors.
-The PSW estimators did not include any covariate outcome modeling. Standard errors for the PSW estimates were estimated using the bootstrap, with resamping done within schools.
+The PSW estimators did not include any covariate outcome modeling. Standard errors for the PSW estimates were estimated using the bootstrap, with resampling done within schools.
 For all three sets of estimators, approximate 95\% confidence intervals were estimated by subtracting twice the standard error from the point estimate (or twice the posterior standard deviation from the posterior mean, in the Bayesian mixture model).
@@
 
-\bibliographystyle{unsrtnat}
+\bibliographystyle{plainnat}
 \bibliography{MOM}
@@
 
-\subsection{Proof for Proposition \ref{prop:reg0}}
+\subsection{Proof for Proposition \ref{prop:reg1}}
 Carrying out those replacements in \eqref{eq:estEq0}, and rearranging\footnote{Technically, replacing $\tilde{\Psi}_i$ with $\Psi_i=\begin{psmallmatrix} 1 & 0&1&0\\ 0&1&0&1\\0&0&1&0\\0&0&0&1\end{psmallmatrix}\tilde{\Psi}_i$.} gives an equivalent set of estimating equations:

### 2022-12-05 writeUps/paper.tex

@@
 \subsection{Proof for Proposition \ref{prop:reg1}}
-Carrying out those replacements in \eqref{eq:estEq0}, and rearranging\footnote{Technically, replacing $\tilde{\Psi}_i$ with $\Psi_i=\begin{psmallmatrix} 1 & 0&1&0\\ 0&1&0&1\\0&0&1&0\\0&0&0&1\end{psmallmatrix}\tilde{\Psi}_i$.} gives an equivalent set of estimating equations:
+
+replacing $\sti$ and $\pp$ in \eqref{eq:estEq0} with $\ri$, as in \eqref{eq:ri}, and replacing $\tilde{\Psi}_i$ with $\Psi_i=\begin{psmallmatrix} 1 & 0&1&0\\ 0&1&0&1\\0&0&1&0\\0&0&0&1\end{psmallmatrix}\tilde{\Psi}_i$ gives an equivalent set of estimating equations:
 \begin{equation}\label{eq:estEq1}

### 2022-12-08 writeUps/paper.tex
filled in missing citations/refs

@@
 Then $S$ itself has potential values $\sti$ and $\sci$, which $i$ would exibit if $Z_i=1$ or $Z_i=0$, respectively.
-Assume ``strong monotonicity'' \citep[c.f.][]{dingLu} (also referred to as ``one-way noncompliance'' \citealt{}):
+Assume ``strong monotonicity'' \citep[c.f.][]{dingLu} (also referred to as ``one-way noncompliance'' \citealt{wing2017can}):
 \begin{ass}[Strong Monotonicity]\label{ass:sm}
@@
 Where
-\begin{equation}\label{eq:Amat}
+\begin{equation*}
   A=\sum_i\frac{\partial}{\partial [\alpha,\beta]'} [Z_i\Omega_i,\Psi_i]'
-\end{equation}
+\end{equation*}
 and
-\begin{equation}\label{eq:Bmat}
+\begin{equation*}%\label{eq:Bmat}
   B=\sum_i [Z_i\Omega_i, \Psi_i]'[Z_i\Omega_i, \Psi_i]
-\end{equation}
+\end{equation*}
 where $[Z_i \Omega_i, \Psi_i]$ is the stacked vector of estimating equations evaluated at $Z_i$, $\bxsi$, $S_i$, and $Y_i$.
@@
   \begin{split}
-    \hat{\eff0}_{int}&\equiv \hat{\beta}_3+\bm{\beta}_6'\overline{\bxy}_{Z=1,S=0}\\
-    \hat{\eff1}_{int}&\equiv \hat{\beta}_3+(\bm{\hat{\beta}_6}+\bm{\hat{\beta}_7})'\overline{\bxy}_{Z=1,S=1}
+    \hat{\eff0}_{int}&\equiv \hat{\beta}_2+\bm{\hat{\gamma}_3}'\overline{\bxy}_{Z=1,S=0}\\
+    \hat{\eff1}_{int}&\equiv \hat{\beta}_2+\hat{\beta}_3+(\bm{\hat{\gamma}_3}+\bm{\hat{\gamma}_4})'\overline{\bxy}_{Z=1,S=1}
   \end{split}
@@
 This is somewhat surprising, because when $\muc1=0$, there is no separation between the principal strata in the control group, and previous literature \citep{griffin2008application} has identified this scenario as a particular challenge for mixture modeling.
-The PSW estimator is also biased---due, presumably, to the violation of Principal Ignorability, Assumption \ref{ass:pi}--inherent in the data generating model.
+The PSW estimator is also biased---due, presumably, to the violation of Principal Ignorability, Assumption \ref{ass:PI}--inherent in the data generating model.
 On the other hand, the PSW estimator has easily the lowest sampling variance of the three estimators.
@@
 In all four conditions, students worked on their assigned math program in class on 12 different occasions---a prior assessment, a mid-test, a post-test, and nine learning sessions---spaced throughout the school year.
-In three of the conditions, during the learning sessions, students worked through a series of math problems, while the forth condition was a computer game designed to teach math concepts.
-In this illustration, we will consider three different treatment effects, comparing one of the four conditions (which we arbitrarily label ``treatment''), with each of the three others (Business as Usual, or ``BAU'', From-Here-To-There!, or ``FH2T,'' and DragonBox), separately.
-During learning sessions, students in the treatment condition received immediate, informative error messages after entering a wrong answer, and were required to enter a correct answer in order to proceed.
-They also had access to a series of hints for each problem, the last of which (called the ``bottom out'' hint) contained the correct answer.
-Students in the other three conditions did not have access bottom out hints; our illustration will focus on that software feature.
+The four conditions ranged from a game---DragonBox---to the active control condition in which students simply used the computer to work on a sequence of math problems. 
+In the DragonBox game, students learned to solve algebraic equations by isolating a box containing a dragon; as students progressed in the game, pictures of monsters were replaced by mathematical symbols.
+Another condition, the From Here to There (FH2T) program, used visual cues such as color and spacing to teach algebraic concepts, and allowed students to solve equations by manipulating expressions via dragging and dropping.
+In the ASSISTments condition, students used computers to work through a series of math problems taken from the district's textbook.
+While working, students had access to a series hints for each problem, culminating in a ``bottom out'' hint that contained the answer.
+Students in the ASSISTments condition were given immediate feedback on their answers and had to enter the correct answer in order to proceed to the next problem.
+In the active control condition, labeled ``business as usual'' (BAU), the computer work mimicked traditional pencil-and-paper mathematics coursework---students worked through a series of math problems without access to hints and received feedback only after completing the assignment.
+In this illustration, we will compare the ASSISTments condition (which we arbitrarily label ``treatment'')---the only condition featuring bottom-out hints---with each of the three others separately.
+Our goal is to estimate different treatment effects for groups of students who would, if assigned to the ASSISTments condition, request few or many bottom out hints.
 
-The literature on hints in online learning is mixed \citep{stuff}, in particular when it comes to requesting a bottom-out hint (called ``bottoming out'').
-On the one hand, hints that include the correct answer are essentially worked examples, which can be beneficial to learning \citep{workedexamplestuff}.
-On the other hand, they allow students to ``game the system,'' by simply requesting enough hints to see the right answer, and use the software without ever working to solve any problems.
-However, in this experiment, the ability to bottom out was only one of several differnces between the treatment conditions.
+The literature on hints in online learning is mixed \citep[see, e.g.,][]{aleven2016help,goldin2012learner,sales2021student} in particular when it comes to requesting a bottom-out hint (called ``bottoming out'').
+On the one hand, hints that include the correct answer are essentially worked examples, which can be beneficial to learning \citep[e.g.]{sweller1985use}.
+On the other hand, they allow students to ``game the system,'' by simply requesting enough hints to see the right answer, and use the software without ever working to solve any problems \citep[e.g.][]{guo2008trying}.
+However, in this experiment, the ability to bottom out was only one of several differences between the treatment conditions.
 One way to disentangle the role of bottoming out in any treatment effect would be to estimate average treatment effects separately for subjects who would (if assigned to treatment) bottom out, and for subjects who would not.
@@
 That is, in roughly 80\% of bottom-outer/non-bottom-outer pairs of subjects, the bottom-outer will have a higher predicted probability from the model.
-Table \ref{tab:results} show the model's coefficient estimates.
+Table \ref{fig:psMod} show the model's coefficient estimates.
 Notably, conditional on the other covariates, students with higher fifth-grade standardized test scores were less likely to request many bottom out hints, as were students who spent little time on task during the pretest, while IEP and ESOL students, and students without baseline measurement data, were relatively unlikely to be bottom-outers.
@@
 
-replacing $\sti$ and $\pp$ in \eqref{eq:estEq0} with $\ri$, as in \eqref{eq:ri}, and replacing $\tilde{\Psi}_i$ with $\Psi_i=\begin{psmallmatrix} 1 & 0&1&0\\ 0&1&0&1\\0&0&1&0\\0&0&0&1\end{psmallmatrix}\tilde{\Psi}_i$ gives an equivalent set of estimating equations:
+Replacing $\sti$ and $\pp$ in \eqref{eq:estEq0} with $\ri$, as in \eqref{eq:ri}, and replacing $\tilde{\Psi}_i$ with $\Psi_i=\begin{psmallmatrix} 1 & 0&1&0\\ 0&1&0&1\\0&0&1&0\\0&0&0&1\end{psmallmatrix}\tilde{\Psi}_i$ gives an equivalent set of estimating equations $\sum_{i=1}^{n_C}\Psi_i=\bm{0}$ or
 \begin{equation}\label{eq:estEq1}
-\sum_{i=1}^{n_C}\Psi_i=\sum_{i=1}^{n_C}\begin{pmatrix}
+\sum_{i=1}^{n_C}\begin{pmatrix}
     Y_i-\muc0-\ri(\muc1-\muc0)-Z_i(\mut0-\muc0)-Z_i\ri(\mut1-\mut0-\muc1+\muc0)\\
@@
 \subsection{Proof for Proposition \ref{prop:interactions}}
+First of all, by \eqref{eq:interaction}, 
+\begin{equation*}
+  \EE[Y_T-Y_C|\st=0]=\EE[Y|Z=1,\st=0]-\EE[Y|Z=1,\st=0]=\beta_2+\bm{\gamma_3}'\EE[\bxy|\st=0]
+\end{equation*}
+and
+\begin{equation*}
+  \EE[Y_T-Y_C|\st=1]=\EE[Y|Z=1,\st=1]-\EE[Y|Z=1,\st=1]=\beta_2+\beta_3+(\bm{\gamma_3}'+\bm{\gamma_4}')\EE[\bxy|\st=1]
+\end{equation*}
+Furthermore, $\overline{\bxy}_{Z=1,S=0}\rightarrow \EE[\bxy|\st=0]$ and $\overline{\bxy}_{Z=1,S=1}\rightarrow \EE[\bxy|\st=1]$ as $n\rightarrow \infty$.
+
+We will show that the estimated coefficients from model \eqref{eq:interaction}, but with $R$ replacing $\st$, fit with OLS, are consistent for $\bm{\beta}$ and $\bm{\gamma}$ from \eqref{eq:interaction}.
+
 First, note that $\EE[S]=\EE\EE[S|\bx]=\EE[\pp]$ and
 \begin{align*}
-  \EE[\bx S]&=\EE[\bx S|Z=1]=\EE[\bx S|Z=0] \mbox{ due to \eqref{ass:rand}}\\
+  \EE[\bx S]&=\EE[\bx S|Z=1]=\EE[\bx S|Z=0] \mbox{ (due to randomization)}\\
   &=\EE[\bx\EE[S|\bx]|Z=0]=\EE[\bx\pp|Z=0]=\EE[\bx\pp]
 \end{align*}
-
-implying that
+implying that, according to \eqref{eq:interaction},
 \begin{equation*}
   \begin{split}
-    \EE[Y|\bx,Z=0]&=\beta_0+\bm{\beta_1}\bxy+\beta_2\pp+\bm{\beta_5}'\bxy\pp\\
-    \EE[Y|\bx,S,Z=1]&=\beta_0+\beta_3+(\bm{\beta_1}+\bm{\beta_6})\bxy+(\beta_3+\beta_4)S+(\bm{\beta_5}+\bm{\beta_7})'\bxy S
+    \EE[Y|\bx,Z=0]&=\beta_0+\beta_1\pp+\bm{\gamma_1}'\bm{X^Y}+\bm{\gamma_2}'\bm{X^Y}\pp\\
+                  &=\beta_0+\beta_1R+\bm{\gamma_1}'\bm{X^Y}+\bm{\gamma_2}'\bm{X^Y}R\\
+    \EE[Y|\bx,S,Z=1]&=\beta_0+\beta_2+(\beta_1+\beta_3)S+(\bm{\gamma_1}+\bm{\gamma_3})\bxy+(\bm{\gamma_2}+\bm{\gamma_4})\bxy S\\
+    &=\beta_0+\beta_2+(\beta_1+\beta_3)R+(\bm{\gamma_1}+\bm{\gamma_3})\bxy+(\bm{\gamma_2}+\bm{\gamma_4})\bxy R
   \end{split}
@@
   \EE[Y]&=\EE[Y|Z=0]+\EE[Z]\left\{\EE[Y|Z=1]-\EE[Y|Z=0]\right\}\\
-  &=\beta_0+\bm{\beta_1}\EE[\bxy]+\beta_2\EE[\pp]+\bm{\beta_5}'\EE[\bxy\pp]+\EE[Z]\left\{\beta_3+\bm{\beta_4}\EE[S]+\bm{\beta_6}'\EE[\bxy]+\bm{\beta_7}'\EE[\bxy S]\right\}\\
-  &=\beta_0+\bm{\beta_1}\EE[\bxy]+\beta_2\EE[R]+\bm{\beta_5}'\EE[\bxy R]+\EE[Z]\left\{\beta_3+\bm{\beta_4}\EE[R]+\bm{\beta_6}'\EE[\bxy]+\bm{\beta_7}'\EE[\bxy R]\right\}\\
+  &=\beta_0+\beta_1\EE[R]+\beta_2\EE[Z]+\beta_3\EE[ZR]+\bm{\gamma_1}'\EE[\bxy]+\bm{\gamma_2}'\EE[\bxy R]+\bm{\gamma_3}'\EE[\bxy Z]+\bm{\gamma_4}\EE[\bxy RZ]
 \end{align*}
+
 Analogous reasoning leads to expressions for $\EE[RY]$, $\EE[\bxy Y]$, $\EE[ZY]$,  $\EE[ZRY]$, $\EE[\bxy YZ]$, and $\EE[\bxy RZY]$.
@@
   \displaystyle\sum_i \begin{pmatrix}
-    Y_i-\beta_0-\bm{\beta_1}\bxy_i-\beta_2R_i-\bm{\beta_5}'\bxy_iR_i-Z_i\left\{\beta_3-\bm{\beta_4}R_i-\bm{\beta_6}'\bxy_i-\bm{\beta_7}'\bxy_i R_i\right\}\\
-    R_iY_i-\beta_0R_i-\bm{\beta_1}\bxy_i R_i-\beta_2R_i^2-\bm{\beta_5}'\bxy_iR_i^2-Z_i\left\{\beta_3R_i-\bm{\beta_4}R_i^2-\bm{\beta_6}'\bxy_iR_i-\bm{\beta_7}'\bxy_i R_i^2\right\}\\
-    \bxy_i Y_i-\beta_0\bxy_i-\bm{\beta_1}\bxy_i\bx^{Y'}_i-\beta_2R_i\bxy_i-\bm{\beta_5}'\bxy_iR_i-Z_i\left\{\beta_3\bxy_i-\bm{\beta_4}R_i\bxy_i-\bm{\beta_6}'\bxy_i\bx^{Y'}_i-\bm{\beta_7}'\bxy_i\bx^{Y'}_i R_i\right\}\\
-    Z_iY_i-Z_i\left\{\beta_0-\bm{\beta_1}\bxy_i-\beta_2R_i-\bm{\beta_5}'\bxy_iR_i-\beta_3-\bm{\beta_4}R_i-\bm{\beta_6}'\bxy_i-\bm{\beta_7}'\bxy_i R_i\right\}\\
-    Z_iR_iY_i-Z_iR_i\left\{\beta_0-\bm{\beta_1}\bxy_i-\beta_2R_i-\bm{\beta_5}'\bxy_iR_i-\beta_3-\bm{\beta_4}R_i-\bm{\beta_6}'\bxy_i-\bm{\beta_7}'\bxy_i R_i\right\}\\
-    Z_iY_i\bxy_i-Z_i\left\{\beta_0\bxy_i-\bm{\beta_1}\bxy_i\bx^{Y'}_i-\beta_2R_i\bxy_i-\bm{\beta_5}'\bxy_i\bx^{Y'}_iR_i-\beta_3\bxy_i-\bm{\beta_4}R_i\bxy_i-\bm{\beta_6}'\bxy_i\bx^{Y'}_i-\bm{\beta_7}'\bxy_i\bx^{Y'}_i R_i\right\}\\
-    R_iZ_iY_i\bxy_i-Z_iR_i\left\{\beta_0\bxy_i-\bm{\beta_1}\bxy_i\bx^{Y'}_i-\beta_2R_i\bxy_i-\bm{\beta_5}'\bxy_i\bx^{Y'}_iR_i-\beta_3\bxy_i-\bm{\beta_4}R_i\bxy_i-\bm{\beta_6}'\bxy_i\bx^{Y'}_i-\bm{\beta_7}'\bxy_i\bx^{Y'}_i R_i\right\}
-  \end{pmatrix}=\begin{pmatrix} 0\\0\\0\\0\\0\\0\\0\end{pmatrix}
+  Y_i\\
+  \phantom{}\\
+  Y_i\ri\\
+    \phantom{}\\
+  Y_iZ_i\\
+  \phantom{}\\
+  Y_iZ_i\ri\\
+  \phantom{}\\
+  Y_i\bxy_i\\
+  \phantom{}\\
+  Y_i\bxy_i\ri\\
+  \phantom{}\\
+  Y_i\bxy_iZ_i\\
+  \phantom{}\\
+  Y_i\bxy_iZ_i\ri\\
+  \phantom{}\end{pmatrix} - \begin{pmatrix}
+    \beta_0+\beta_1R_i+\beta_2Z_i+\beta_3Z_iR_i\phantom{-----}\\
+    \hfill+\left\{\bm{\gamma_1}'+R_i\bm{\gamma_2}'+Z_i\bm{\gamma_3}'+R_iZ_i\bm{\gamma_4}'\right\}\bxy\\
+    \beta_0\ri+\beta_1R_i^2+\beta_2Z_i\ri+\beta_3Z_iR_i^2\phantom{-----}\\
+    \hfill+\ri\left\{\bm{\gamma_1}'+R_i\bm{\gamma_2}'+Z_i\bm{\gamma_3}'+R_iZ_i\bm{\gamma_4}'\right\}\bxy\\
+    (\beta_0+\beta_2)Z_i+(\beta_1+\beta_3)R_iZ_i\phantom{-----}\\
+    \hfill+Z_i\left\{\bm{\gamma_1}'+R_i\bm{\gamma_2}'+Z_i\bm{\gamma_3}'+R_iZ_i\bm{\gamma_4}'\right\}\bxy\\
+    \beta_0Z_i\ri+(\beta_1+\beta_3)Z_iR_i^2+\beta_2Z_i\ri\phantom{-----}\\
+    \hfill+Z_iR_i\left\{\bm{\gamma_1}'+\bm{\gamma_3}'+R_i(\bm{\gamma_2}'+\bm{\gamma_4}')\right\}\bxy \hfill\\
+    \beta_0\bxyt+\beta_1R_i\bxyt+\beta_2Z_i\bxyt+\beta_3Z_iR_i\bxyt\phantom{---}\\
+    \hfill {}+\left\{\bm{\gamma_1}'+R_i\bm{\gamma_2}'+Z_i\bm{\gamma_3}'+R_iZ_i\bm{\gamma_4}'\right\}\bxy\bxyt\\
+    \beta_0\ri\bxyt+\beta_1R_i^2\bxyt+\beta_2Z_i\ri\bxyt+\beta_3Z_iR_i^2\bxyt\phantom{---}\\
+    \hfill {}+\ri\left\{\bm{\gamma_1}'+R_i\bm{\gamma_2}'+Z_i\bm{\gamma_3}'+R_iZ_i\bm{\gamma_4}'\right\}\bxy\bxyt\\
+        (\beta_0+\beta_2)Z_i\bxyt+(\beta_1+\beta_3)R_iZ_i\bxyt\phantom{-----}\\
+    \hfill {}+Z_i\left\{\bm{\gamma_1}'+R_i\bm{\gamma_2}'+Z_i\bm{\gamma_3}'+R_iZ_i\bm{\gamma_4}'\right\}\bxy\bxyt\\
+\beta_0\ri Z_i\bxyt+\beta_1R_i^2Z_i\bxyt+\beta_2\ri Z_i\bxyt+\beta_3Z_iR_i^2\bxyt\phantom{---}\\
+    \hfill {}+\ri Z_i\left\{\bm{\gamma_1}'+R_i\bm{\gamma_2}'+Z_i\bm{\gamma_3}'+R_iZ_i\bm{\gamma_4}'\right\}\bxy\bxyt\\
+  \end{pmatrix}=\bm{0}
 \end{equation*}
-These are the estimating equations for the regression model \eqref{eq:fullInteractionModel}, fit by OLS. Now note that
-\begin{equation*}
-  \EE[Y_T-Y_C|\st=0]=\beta_0^T-\beta_0^C+(\bm{\beta_1}^{T'}-\bm{\beta_1}^{C'})\EE[\bxy|S=0]=\beta_3+\beta_6'E[\bxy|S=0]
-\end{equation*}
-and
-\begin{equation*}
-  \EE[Y_T-Y_C|\st=1]=\beta_0^T+\beta_2^T-(\beta_0^C+\beta_2^C)+(\bm{\beta_1}^{T'}+\bm{\beta_3^{Ti}}-\bm{\beta_1}^{C'}-\bm{\beta_3^{C'}})\EE[\bxy|S=0]=\beta_3+\beta_4+(\beta_6'+\beta_7')E[\bxy|S=1]
-\end{equation*}
-the result follows.
+These are the estimating equations for the regression model \eqref{eq:interaction}, with $R$ replacing $\st$, fit by OLS. 
+Under standard OLS regularity conditions, the estimated coefficients of that model are consistent, completing the proof.
 
@@
 
-So we're estimating principal effects in two stages.
+Here we will derive the sandwich variance-covariance matrix for the GEEPERs estimate without interactions between $\bx$ and either $Z$ or $\st$---i.e.
+with $\bm{\gamma_2}=\bm{\gamma_3}=\bm{\gamma_4}=0$ in the notation of \eqref{eq:interaction}---and estimating principal scores using a generalized linear model.
+
+We propose estimating principal effects in two stages.
 First, fit the model
 \begin{equation}\label{eq:psMod}
-  Pr(S_{Ti}=1|\bm{\tilde{w}}_i)=f(\bm{\alpha}'\bm{\tilde{w}}_i)
+  \ppi=Pr(\sti=1|\bxsi)=f(\bm{\alpha}'\bxsit)
 \end{equation}
-for some inverse link function $f(\cdot)$, where $\bm{\tilde{x}}_i=[1,\bm{w}_i]$, using (observed) values from the treatment group, and estimating $\hat{\alpha}$.
+for some inverse link function $f(\cdot)$, where $\bxsit=[1,\bxsi]$, using (observed) values from the treatment group, and estimating $\hat{\alpha}$.
 Then let
 \begin{equation}\label{eq:ps}
-  \hat{p}_i=f(\bm{\hat{\alpha}}'\bm{\tilde{w}}_i)
+  \hat{p}_i=f(\bm{\hat{\alpha}}'\bxsit)
 \end{equation}
@@
 \begin{equation}\label{eq:regression}
-  Y_i=\beta_0+\beta_1r_i+\beta_2Z_i+\beta_3Z_ir_i+\bm{\gamma}^t\bm{x}_i+\epsilon_i
+  Y_i=\beta_0+\beta_1r_i+\beta_2Z_i+\beta_3Z_i\ri+\bm{\gamma}'\bxy_i+\epsilon_i
 \end{equation}
 to estimate $\bm{\beta}$ and hence principal effects, where $\bm{x}_i$ is a set of covariates predictive of $Y$ within principal strata.
-[Note that \eqref{eq:regression} is just \eqref{eq:regression0} with $\bm{\gamma}^t\bm{x}_i$ added on the end].
-$\bm{x}_i$ should, generally, contain $\bm{w}_i$; the two may also be equivalent.
-Let $\bm{\beta}=[\beta_0,\beta_1,\beta_2,\beta_3,\bm{\gamma}^t]^t$.
+Let $\bm{\beta}=[\beta_0,\beta_1,\beta_2,\beta_3,\bm{\gamma}']'$.
 
-The standard error matrix for $\bm{\hat{\beta}}$ can be estimated as:
-\begin{equation}\label{eq:sandwich}
-  \widehat{var}(\bm{\hat{\beta}})=A^{-1}BA^{-t}
-\end{equation}
-Where
-\begin{equation}\label{eq:Amat}
-  A=\sum_i\frac{\partial}{\partial [\alpha,\beta]^t} \Psi(Y_i;\alpha,\beta,\bm{x}_i,\bm{w}_i)
-\end{equation}
+Following \eqref{eq:stacked}, let $\blam(Z_i,S_i,\bxsi,\bxy_i,Y_i;\bm{\alpha},\bm{\beta})=\begin{pmatrix} Z_i\Omega(\bxsi,S_i;\bm{\alpha})\\ \Psi(\bxy_i,\bxsi,Y_i,Z_i,\ri(\bm{\alpha});\bm{\beta})\end{pmatrix}$, the stacked estimating equations of \eqref{eq:psMod} and \eqref{eq:regression}.
+Going forward, for the sake of brevety, we will write $\blam_i(\bm{\alpha},\bm{\beta})=\blam(Z_i,S_i,\bxsi,\bxy_i,Y_i;\bm{\hat{\alpha}},\bm{\hat{\beta}})$, where dependence on the data for $i$ is captured in the subscript $i$, with similiar meanings for $\Psi_i(\bm{\alpha},\bm{\beta})$ and $\blam_i(\bm{\alpha},\bm{\beta})$.
+%Then let $\blamh_i=\blam_i(Z_i,S_i,\bxsi,\bxy_i,Y_i;\bm{\hat{\alpha}},\bm{\hat{\beta}})$ be the estimating equations evaluated at the estimated parameters, and let 
+%\begin{equation*}
+%\dblam_i=\frac{\partial}{\partial [\alpha,\beta]'} \blam_i\Bigr|_{\substack{\bm{\alpha}=\bm{\hat{\alpha}}\\\bm{\beta}=\bm{\hat{\beta}}}}
+%\end{equation*}
+%the derivative matrix of $\blam$ evaluated at the estimated parameters.
+
+The variance-covariance matrix for $\bm{\hat{\alpha}}$ and $\bm{\hat{\beta}}$ can be estimated as:
+\begin{equation*}
+  \widehat{var}\left([\bm{\hat{\alpha}}',\bm{\hat{\beta}}']'\right)=A^{-1}BA^{-t}
+\end{equation*}
+where
+\begin{equation*}%\label{eq:Amat}
+  A=\sum_i \frac{\partial}{\partial [\bm{\alpha},\bm{\beta}]'} \blam_i\Bigr|_{\substack{\bm{\alpha}=\bm{\hat{\alpha}}\\\bm{\beta}=\bm{\hat{\beta}}}}
+\end{equation*}
 and
-\begin{equation}\label{eq:Bmat}
-  B=\sum_i \Psi(Y_i;\alpha,\beta,\bm{x}_i)\Psi(Y_i;\alpha,\beta,\bm{x}_i,\bm{w}_i)^t
-\end{equation}
-where $\Psi(\cdot)$ is the estimating equations.
+\begin{equation*}%\label{eq:Bmat}
+  B=\sum_i \blam_i(\bm{\hat{\alpha}},\bm{\hat{\beta}})\blam_i(\bm{\hat{\alpha}},\bm{\hat{\beta}})'
+\end{equation*}
 
-Following \citet[][p. 373]{carroletal06}, we %separate the estimating equations $\Psi$ into $\phi$ and $\psi$, estimating equations for models \eqref{
-can decompose the matrices into diagonal elements $A_{1,1}$ and $A_{2,2}$ and $B_{1,1}$ and $B_{2,2}$ that pertain to the parameter sets $\alpha$ and $\beta$ and the estimating equations for models \eqref{eq:psMod} and  \eqref{eq:regression}, respectively, and $A_{21}$, and $B_{12}=B_{21}^t$, which capture to the dependence of model \eqref{eq:regression} on the parameters $\alpha$ from \eqref{psMod} and the covariance between the estimating equations of the two models.
-The sub-matrix $A_{12}=0$, since \eqref{eq:psMod} does not depend on $\beta$.
+Following \citet[][p. 373]{carroll2006measurement}, we %separate the estimating equations $\Psi$ into $\phi$ and $\psi$, estimating equations for models \eqref{
+can decompose the matrices into diagonal elements 
+\begin{equation*}
+    \begin{split}
+        A_{1,1}&=\sum_i \partial \Omega_i/\partial \bm{\alpha}|_{\bm{\alpha}=\bm{\hat{\alpha}}}\\
+        A_{2,2}&=\sum_i\partial\Psi_i/\partial \bm{\beta}|_{\bm{\beta}=\bm{\hat{\beta}}}\\
+        B_{1,1}&=\sum_i\Omega_i(\bm{\hat\alpha})\Omega_i(\bm{\hat\alpha})'\\
+        B_{2,2}&=\sum_i \Psi_i(\bm{\hat\alpha},\bm{\hat\beta})\Psi_i(\bm{\hat\alpha},\bm{\hat\beta})'
+    \end{split}
+\end{equation*}
+ that pertain to the parameter sets $\bm{\alpha}$ and $\bm{\beta}$ and the estimating equations for models \eqref{eq:psMod} and  \eqref{eq:regression}, respectively, and 
+ \begin{equation*}
+     \begin{split}
+         A_{21}&=\sum_i\partial\Psi_i/\partial \bm{\alpha}|_{\bm{\alpha}=\bm{\hat{\alpha}}}\\
+         B_{12}=B_{21}'&=\sum_i \Omega_i(\bm{\hat{\alpha}})\Psi_i(\bm{\hat{\alpha}},\bm{\hat{\beta}})'
+     \end{split}
+ \end{equation*}
+ which capture to the dependence of model \eqref{eq:regression} on the parameters $\bm{\alpha}$ from \eqref{eq:psMod} and the covariance between the estimating equations of the two models. 
+The sub-matrix $A_{12}=\sum_i \partial \Omega_i/\partial \bm{\beta}=0$, since \eqref{eq:psMod} does not depend on $\bm{\beta}$.
 
 The diagonal matrices $A_{1,1}$ and $A_{2,2}$ and $B_{1,1}$ and $B_{2,2}$ are all the typical ``bread'' and ``meat'' matrices from M-estimation of generalized linear models and OLS. 
-In practice, we use the estimates from the \texttt{sandwich} package in \texttt{R}, adjusted in two ways: first, the function \texttt{bread} actually gives $A^{-1}$, not $A$; second, we must pay careful attention to sample sizes, since the sample size for \eqref{eq:psMod} includes only treated observations (with observed $S$) and \eqref{eq:regression} contains all observations.
+%In practice, we use the estimates from the \texttt{sandwich} package in \texttt{R}, adjusted in two ways: first, the function \texttt{bread} actually gives $A^{-1}$, not $A$; second, we must pay careful attention to sample sizes, since the sample size for \eqref{eq:psMod} includes only treated observations (with observed $S$) and \eqref{eq:regression} contains all observations.
+Calculation of the matrices $B_{12}$ and $B_{21}$ is straightforward after vectors $\Omega_i(\bm{\hat{\alpha}})$ and $\Psi_i(\bm{\hat{\alpha}},\bm{\hat{\beta}})$ have been calculated.
+Some specialized calculation is necessary for matrix $A_{21}$.
 
@@
 \begin{equation}\label{eq:eeOLS}
-  \psi(Y_i,\bm{x}_i,\bm{w}_i,\beta,\alpha)=X_iY_i-X_iX_i^t\beta
+  \psi(Y_i,\bxy_i,\bxsi,\bm{\beta},\alpha)=X_iY_i-X_iX_i'\bm{\beta}
 \end{equation}
-Where $X_i=[1,r_i,Z_i,r_iZ_i,\bm{x}_i^t]^t$.
+Where $X_i=[1,r_i,Z_i,r_iZ_i,\bxyt_i]'$.
 In other words,
 \begin{align*}
-  \psi(Y_i,&\bm{x}_i,\bm{w}_i,\beta,\alpha)=\\
+  \psi(Y_i,&\bxy_i,\bxsi,\bm{\beta},\bm{\alpha})=\\
   &\left\{
   \begin{array}{l}
-    Y_i-\left(\beta_0+\beta_1r_i+\beta_2Z_i+\beta_3Z_ir_i+\bm{\gamma}'\bm{x}_i\right)\\
-    r_iY_i-r_i\left(\beta_0+\beta_2Z_i+\bm{\gamma}'\bm{x}_i\right)-r_i^2\left(\beta_1+\beta_3Z_i\right)\\
-    Z_iY_i-Z_i\left(\beta_0+\beta_1r_i+\beta_2Z_i+\beta_3Z_ir_i+\bm{\gamma}'\bm{x}_i\right)\\
-    Z_ir_iY_i-Z_ir_i\left(\beta_0+\beta_2Z_i+\bm{\gamma}'\bm{x}_i\right)-Z_ir_i^2\left(\beta_1+\beta_3Z_i\right)\\
-    \bm{x}_iY_i-\bm{x}_i\left(\beta_0+\beta_1r_i+\beta_2Z_i+\beta_3Z_ir_i+\bm{\gamma}'\bm{x}_i\right)
+    Y_i-\left(\beta_0+\beta_1r_i+\beta_2Z_i+\beta_3Z_ir_i+\bm{\gamma}'\bxy_i\right)\\
+    r_iY_i-r_i\left(\beta_0+\beta_2Z_i+\bm{\gamma}'\bxy_i\right)-r_i^2\left(\beta_1+\beta_3Z_i\right)\\
+    Z_iY_i-Z_i\left(\beta_0+\beta_1r_i+\beta_2Z_i+\beta_3Z_ir_i+\bm{\gamma}'\bxy_i\right)\\
+    Z_ir_iY_i-Z_ir_i\left(\beta_0+\beta_2Z_i+\bm{\gamma}'\bxy_i\right)-Z_ir_i^2\left(\beta_1+\beta_3Z_i\right)\\
+    \bxy_iY_i-\bxy_i\left(\beta_0+\beta_1r_i+\beta_2Z_i+\beta_3Z_ir_i+\bm{\gamma}'\bxy_i\right)
   \end{array}
@@
 (noting that $Z^2=Z$).
-These depend on $\alpha$ when $r_i=p_i$, i.e. when $Z_i=0$ or $S_i$ is missing.
-Then note that, following \eqref{eq:ps}, and letting $\eta_i=\alpha^t\bm{\tilde{w}}_i$
+These depend on $\bm{\alpha}$ when $r_i=p_i$, i.e. when $Z_i=0$.% or $S_i$ is missing.
+Then note that, following \eqref{eq:ps}, and letting $\eta_i=\bm{\alpha}'\bxsit$
 \begin{equation}\label{eq:derivP}
-  \frac{\partial p_i}{\partial \alpha^t}=f'(\eta_i)\frac{\partial \eta_i}{\partial \alpha^t}=f'(\eta_i)\bm{\tilde{w}}_i^t
+  \frac{\partial p_i}{\partial \bm{\alpha}'}=f'(\eta_i)\frac{\partial \eta_i}{\partial \bm{\alpha}'}=f'(\eta_i)\bxsitp
 \end{equation}
@@
 \begin{equation}
-  \frac{\partial p_i^2}{\partial \alpha^t}=2p_i\frac{\partial p_i}{\partial \alpha^t}=2f(\eta)f'(\eta_i)\bm{\tilde{w}}_i^t=2p_if'(\eta_i)\bm{\tilde{w}}_i^t
+  \frac{\partial p_i^2}{\partial \bm{\alpha}'}=2p_i\frac{\partial p_i}{\partial \bm{\alpha}'}=2f(\eta)f'(\eta_i)\bxsitp=2p_if'(\eta_i)\bxsitp
 \end{equation}
@@
 Then if $r_i=p_i$,
+% \begin{align*}
+%   \frac{\partial}{\partial \bm{\alpha}'}&\psi(Y_i,\bxy_i,\bm{\beta},\bm{\alpha})=\\
+%   &\left[\begin{array}{c}
+%           -(\beta_1+\beta_3Z_i)\frac{\partial p_i}{\partial \bm{\alpha}'}\\
+%           (Y_i-\beta_0-\beta_2Z_i-\bm{\gamma}'\bm{x}_{i})\partial p/\partial\bm{\alpha}'-(\beta_1+\beta_3Z_i)\partial p^2/\partial \bm{\alpha}'\\
+%           -(\beta_1+\beta_3Z_i)\frac{\partial p_i}{\partial \bm{\alpha}'}Z_i\\
+%           Z_i\left[(Y_i-\beta_0-\beta_2Z_i-\bm{\gamma}'\bm{x}_{i})\partial p/\partial\bm{\alpha}'-(\beta_1+\beta_3Z_i)\partial p^2/\partial \bm{\alpha}'\right]\\
+%           -(\beta_1+\beta_3Z_i)\frac{\partial p_i}{\partial \bm{\alpha}'}\bxy_i
+%         \end{array}\right]\\
+%   =& \left[\begin{array}{c}
+%           -(\beta_1+\beta_3Z_i)f'(\eta_i)\bxsitp\\
+%           \left[Y_i-\beta_0-\beta_2Z_i-\bxy_i'\bm{\gamma}-2p_i(\beta_1+\beta_3Z_i)\right]f'(\eta_i)\bxsitp\\
+%           -Z_i\beta_1f'(\eta_i)\bxsitp\\
+%           Z_i\left[Y_i-\beta_0-\beta_2Z_i-\bxy_i'\bm{\gamma}-2p_i(\beta_1+\beta_3)\right]f'(\eta_i)\bxsitp\\
+%           -(\beta_1+\beta_3Z_i)f'(\eta_i)\bxy_i\bxsitp
+%     \end{array}\right]
+% \end{align*}
+
+% When $S$ is observed for all members of the treatment group, $Z_i=0$ whenever $r_i=p_i$, so the latter expression reduces to
 \begin{align*}
-  \frac{\partial}{\partial \alpha^t}&\psi(Y_i,\bm{x}_i,\beta,\alpha)=\\
-  &\left[\begin{array}{c}
-          -(\beta_1+\beta_3Z_i)\frac{\partial p_i}{\partial \alpha^t}\\
-          (Y_i-\beta_0-\beta_2Z_i-\bm{\gamma}^t\bm{x}_{i})\partial p/\partial\alpha^t-(\beta_1+\beta_3Z_i)\partial p^2/\partial \alpha^t\\
-          -(\beta_1+\beta_3Z_i)\frac{\partial p_i}{\partial \alpha^t}Z_i\\
-          Z_i\left[(Y_i-\beta_0-\beta_2Z_i-\bm{\gamma}^t\bm{x}_{i})\partial p/\partial\alpha^t-(\beta_1+\beta_3Z_i)\partial p^2/\partial \alpha^t\right]\\
-          -(\beta_1+\beta_3Z_i)\frac{\partial p_i}{\partial \alpha^t}\bm{x}_i
-        \end{array}\right]\\
-  =& \left[\begin{array}{c}
-          -(\beta_1+\beta_3Z_i)f'(\eta_i)\bm{\tilde{w}}_i^t\\
-          \left[Y_i-\beta_0-\beta_2Z_i-\bm{x}_i^t\bm{\gamma}-2p_i(\beta_1+\beta_3Z_i)\right]f'(\eta_i)\bm{\tilde{w}}_i^t\\
-          -Z_i\beta_1f'(\eta_i)\bm{\tilde{w}}_i^t\\
-          Z_i\left[Y_i-\beta_0-\beta_2Z_i-\bm{x}_i^t\bm{\gamma}-2p_i(\beta_1+\beta_3)\right]f'(\eta_i)\bm{\tilde{w}}_i^t\\
-          -(\beta_1+\beta_3Z_i)f'(\eta_i)\bm{x}_i\bm{\tilde{w}}_i^t
+  \frac{\partial}{\partial \bm{\alpha}'}&\psi(Y_i,\bxy_i,\bm{\beta},\bm{\alpha})=\\
+  & \left[\begin{array}{c}
+          -(\beta_1)f'(\eta_i)\bxsitp\\
+          \left[Y_i-X_i'\bm{\beta}-2p_i(\beta_1)\right]f'(\eta_i)\bxsitp\\
+          0\\
+          0\\
+          -\beta_1f'(\eta_i)\bxy_i\bxsitp
     \end{array}\right]
@@
 
-When $S$ is observed for all members of the treatment group, $Z_i=0$ whenever $r_i=p_i$, so the latter expression reduces to
-\begin{align*}
-  \frac{\partial}{\partial \alpha^t}&\psi(Y_i,\bm{x}_i,\beta,\alpha)=\\
-  & \left[\begin{array}{c}
-          -(\beta_1)f'(\eta_i)\bm{\tilde{w}}_i^t\\
-          \left[Y_i-X_i^t\beta-2p_i(\beta_1)\right]f'(\eta_i)\bm{\tilde{w}}_i^t\\
-          0\\
-          0\\
-          -\beta_1f'(\eta_i)\bm{x}_i\bm{\tilde{w}}_i^t
-    \end{array}\right]
-\end{align*}
-
-If $r_i=S_i$, $\frac{\partial}{\partial \alpha^t}\psi(Y_i,\bm{x}_i,\beta,\alpha)=0$.
+If $r_i=S_i$, $\frac{\partial}{\partial \bm{\alpha}'}\psi(Y_i,\bxy_i,\bm{\beta},\bm{\alpha})=0$.
 

### 2022-12-09 code/simulation/displaySim.r
combine figs

@@ -2,6 +2,8 @@ library(tidyverse)
 library(broom)
 library(kableExtra)
 library(gridExtra)
+library(ggpubr)
+
 
 source('code/simulation/readSimFuncs.r')
 
@@ -56,7 +58,7 @@ seN <- pdns%>%
 
 full_join(seN,biasN)%>%filter(eff==1,estimator!="PSW")%>%summarize(across(c(se,bias),max),coef=round(se/bias))
 
-bind_rows(
+plotByN=bind_rows(
   seN%>%mutate(se=se/3,meas="SE")%>%rename(what=se),
   biasN%>%mutate(meas="Bias")%>%rename(what=bias))%>%
   filter(eff==1,estimator!="PSW")%>%
@@ -65,9 +67,10 @@ bind_rows(
   scale_y_continuous(name="Bias",sec.axis=sec_axis(trans=~.*3,name='Standard Error'))+
   scale_shape_manual(values=c(0,16))+
   annotate('text',600,.125,label=list(bquote(atop("Normal Resid., No Interactions, ",alpha==0.5~", "~mu[C]^1-mu[C]^0==0.3))),parse=TRUE)+
-  scale_x_continuous(name="Sample Size Per Group",breaks=c(seq(100,500,200),1000))+
-  ggtitle("Bias and Standard Error by n")+theme(legend.title=element_blank())
-ggsave("simFigs/biasSEbyN.jpg",width=6,height=3)
+  scale_x_continuous(name="Sample Size Per Group (n)",breaks=c(seq(100,500,200),1000))+
+  #ggtitle("Bias and Standard Error by n")+
+  theme(legend.title=element_blank())
+ggsave("simFigs/biasSEbyN.jpg",plot=plotByN,width=6,height=3)
 
 
 ##### table for appendix
@@ -150,7 +153,7 @@ seB1 <- pdb1%>%
 
 full_join(seB1,biasB1)%>%filter(eff==1,estimator!="PSW",b1>0)%>%summarize(across(c(se,bias),max),coef=round(se/bias))
 
-bind_rows(
+plotByAlpha=bind_rows(
   seB1%>%mutate(se=se/5,meas="SE")%>%rename(what=se),
   biasB1%>%mutate(meas="Bias")%>%rename(what=bias))%>%
   filter(b1>0,eff==1,estimator!="PSW")%>%
@@ -158,9 +161,17 @@ bind_rows(
[…]

### 2022-12-09 writeUps/paper.tex

@@
-\documentclass[11pt]{article} % use larger type; default would be 10pt
+\documentclass{statsoc} % use larger type; default would be 10pt
 
@@
 
-\title{GEEPERs: Principal Stratification using Principal Scores and Stacked Estimating Equations}
+\title[GEEPERS]{GEEPERs: Principal Stratification using Principal Scores and Stacked Estimating Equations}
+
+\author[Sales and Vanacore]{Adam C Sales}
+
+\author{Kirk Vanacore}
+\address{Worcester Polytechnic Institute,
+Worcester, MA,
+USA.}
+\email{asales@wpi.edu}
 
@@
 \maketitle
+
+\begin{abstract}
+Principal stratification is a framework for making sense of causal effects conditioned on variables that themselves may have been affected by treatment.
+Most principal stratification estimators rely on strong structural or modeling assumptions, and many require advanced statistical training to fit and to check.
+In this paper, we introduce a new M-estimation principal effect estimator for one-way noncompliance based on a binary indicator.
+Estimates may be computed using conventional regressions (though the standard errors require a specialized sandwich formula) and do not rely on distributional assumptions.
+We illustrate the new technique in an analysis of student log data from a recent educational technology field experiment.
+\end{abstract}
+\keywords{Causal Inference; Principal Stratification; Educational Technology}
 
@@
 \begin{table}
+    \caption{\label{tab:factor} Manipulated factors in simulation study}
   \centering
-  \begin{tabular}{cc}
+\begin{tabular}{*{2}{c}}
   \hline
@@
   \end{tabular}
-  \caption{Manipulated factors in simulation study}
-  \label{tab:factor}
   \end{table}
@@
 \begin{equation}\label{eq:mix-sim}
-  f_{Y|Z=0}(y_i)=\ppf{\bm{\alpha}}{\bm{x}_{i,1:2}}\phi\left(\frac{y_i-[\beta_0+\beta_1+\bm{\gamma}_1^T\bm{x}_{i,1:2}]}{\sigma_C}\right)+(1-\ppf{\bm{\alpha}}{\bm{x}_{i,1:2}})\phi\left(\frac{y_i-[\beta_0+\bm{\gamma}_1^T\bm{x}_{i,1:2}]}{\sigma_C}\right)
+\begin{split}
+  f_{Y|Z=0}(y_i)&=\ppf{\bm{\alpha}}{\bm{x}_{i,1:2}}\phi\left(\frac{y_i-[\beta_0+\beta_1+\bm{\gamma}_1^T\bm{x}_{i,1:2}]}{\sigma_C}\right)\\
+  &+(1-\ppf{\bm{\alpha}}{\bm{x}_{i,1:2}})\phi\left(\frac{y_i-[\beta_0+\bm{\gamma}_1^T\bm{x}_{i,1:2}]}{\sigma_C}\right)
+\end{split}
 \end{equation}
@@
   \centering
-  \includegraphics{../simFigs/biasSEbyN.jpg}
-  \caption{Empirical bias and standard error of M- and Mixture estimates of $\muc1$ as sample size per treatment group $n$ varies from 0 to 1.
+  \includegraphics[width=\textwidth]{../simFigs/biasSEbyB1N.jpg}
+  \caption{Empirical bias and standard error of GEEPERs and Mixture estimates of $\muc1$ as sample size per treatment group $n$ or $\alpha$ varies.
+There were 500 replications for each $n$ or $\alpha$.
+Other factors were held fixed at the noted values.}
+  \label{fig:alphan}
+\end{figure}
+
+% \begin{figure}
+%   \centering
+%   \includegraphics{../simFigs/biasSEbyB1.jpg}
+%   \caption{Empirical bias and standard error of M- and Mixture estimates of $\muc1$ as $\alpha$ varies from 0.1 to 1.
 There were 500 replications for each $n$.
 Other factors were held fixed at the noted values.}
-  \label{fig:n}
-\end{figure}
-
-\begin{figure}
-  \centering
-  \includegraphics{../simFigs/biasSEbyB1.jpg}
-  \caption{Empirical bias and standard error of M- and Mixture estimates of $\muc1$ as $\alpha$ varies from 0.1 to 1.
-There were 500 replications for each $n$.
-Other factors were held fixed at the noted values.}
-  \label{fig:alpha}
-\end{figure}
+%   \label{fig:alpha}
+% \end{figure}
 
@@
   \centering
-  \includegraphics{../simFigs/boxplots.pdf}
+  \includegraphics[width=\textwidth]{../simFigs/boxplots.pdf}
   \caption{Violin and scatterplots of 500 simulation estimation errors for estimators described in Section~\ref{sec:simMods} under varying conditions.
@@
 
-  \begin{table}
-    \centering
-  \input{rmseTab}
-  \caption{Empirical root-mean-squared error (RMSE) estimates for GEEPERs, Bayesian mixture modeling, and PSW under varying conditions.}
-  \label{tab:rmse}
-\end{table}
+%   \begin{table}
+%     \centering
+%   \input{rmseTab}
+%   \caption{Empirical root-mean-squared error (RMSE) estimates for GEEPERs, Bayesian mixture modeling, and PSW under varying conditions.}
+%   \label{tab:rmse}
+% \end{table}
 
 \begin{table}
+\caption{Empirical coverage of nominal 95\% confidence intervals for GEEPERs and Bayesian mixture modeling under varying conditions.}
+  \label{tab:coverage}
 \centering
   \input{coverageTab}
-  \caption{Empirical coverage of nominal 95\% confidence intervals for GEEPERs and Bayesian mixture modeling under varying conditions.}
-  \label{tab:coverage}
 \end{table}

### 2022-12-13 code/simulation/displaySim.r
standardized y scale in violin plots

@@ -257,7 +257,7 @@ int<-bp(pd,
    subset=b1 > 0& errDist=='norm'& n==500&PE=='Stratum 1'&b1==.5&mu01==0.3,
    title="S&Z Interactions",
    facet = interactionZ~interactionS,
-   ylim=NULL
+   ylim=c(-1,1)
    )+
   labs(subtitle=bquote(~alpha==0.5~","~mu[C]^1==0.3~"Norm. Resid"),x=NULL,y=NULL)+
   theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1),

### 2022-12-13 code/simulation/readSimFuncs.r

@@ -141,7 +141,7 @@ bp <- function(pd,subset,facet,title=deparse(substitute(subset)),ylim,Labeller="
       draw_quantiles = 0.5) +
     geom_hline(yintercept = 0) +
     coord_cartesian(ylim =ylim)+
-    facet_grid(facet , scales = "free",
+    facet_grid(facet , #scales = "free",
                labeller = Labeller)+
     ggtitle(title)+
                                         #labs(title = title,

### 2022-12-13 code/simulation/displaySim.r
fixed annotation in alpha-n fig

@@ -66,7 +66,7 @@ plotByN=bind_rows(
   geom_point()+geom_line()+geom_hline(yintercept=0)+#,linetype="dotted",size=2)+
   scale_y_continuous(name="Bias",sec.axis=sec_axis(trans=~.*3,name='Standard Error'))+
   scale_shape_manual(values=c(0,16))+
-  annotate('text',600,.125,label=list(bquote(atop("Normal Resid., No Interactions, ",alpha==0.5~", "~mu[C]^1-mu[C]^0==0.3))),parse=TRUE)+
+  annotate('text',600,.11,label=list(bquote(atop("Normal Resid., No Interactions, ",alpha==0.5~", "~mu[C]^1-mu[C]^0==0.3))),parse=TRUE)+
   scale_x_continuous(name="Sample Size Per Group (n)",breaks=c(seq(100,500,200),1000))+
   #ggtitle("Bias and Standard Error by n")+
   theme(legend.title=element_blank())
@@ -161,7 +161,7 @@ plotByAlpha=bind_rows(
   geom_point()+geom_line()+geom_hline(yintercept=0)+#,linetype="dashed")+
   scale_y_continuous(name="Bias",sec.axis=sec_axis(trans=~.*5,name='Standard Error'))+
   scale_shape_manual(values=c(0,16))+
-  annotate('text',.7,.1,label=list(bquote(atop("Normal Resid., No Interactions, ",n==500~", "~mu[C]^1-mu[C]^0==0.3))),parse=TRUE)+
+  annotate('text',.7,.08,label=list(bquote(atop("Normal Resid., No Interactions, ",n==500~", "~mu[C]^1-mu[C]^0==0.3))),parse=TRUE)+
   xlab(bquote(alpha))#+ggtitle(bquote("Bias and Standard Error by "~alpha))
 ggsave("simFigs/biasSEbyB1.jpg",plot=plotByAlpha,width=6,height=3)
 

### 2022-12-13 writeUps/paper.tex

@@
 \usepackage{natbib}
-\usepackage{hyperref}
+\usepackage[hidelinks]{hyperref}
 \usepackage{graphicx}
@@
 Randomized experiments easily admit subgroup analysis---estimation of treatment effects for a subset of the study population---when those subgroups are defined at baseline, and not themselves a function of treatment assignment.
-A simple, popular approach to estimating subgroup effects is to use ordinary least squares (OLS) regression, regressing the outcome on a treatment indicator interacted with an indicator for subgroup membership.
+%A simple, popular approach to estimating subgroup effects is to use ordinary least squares (OLS) regression, regressing the outcome on a treatment indicator interacted with an indicator for subgroup membership.
 However, when subgroup membership is a function of a post-treatment variable, then it may itself be affected by the treatment.
-Contrasts between treatment and control conditions within those subgroups may not have a causal interpretation at all---and when they do, the same techniques that work for subgroups defined at baseline (such as OLS with interactions) will often produce inconsistent estimates.
-Principal stratification \citep{frangakis} is a broad framework for defining subgroup effects for post-treatment subgroups, based on the notion of \emph{potential} subgroups called principal strata, that is, groups of subjects defined based on the values a post-treatment variable would potentially take if the subject is assigned to one treatment condition or another.
-Since principal strata membership is based on subjects' responses to both factual and counter-factual treatment assignments, it is, in general unobserved.
-Nevertheless, researchers have devised estimators of principal effects, as average effects within principal strata are called, to address a broad range of applied statistical problems, including non-compliance in randomized experiments, the evaluation of surrogate outcomes \citep{li2010bayesian}, the investigation of causal mechanisms \citep{lidsayPage}, and sample attrition \citep{zhangRubin, ding2011}.
+Contrasts between treatment and control conditions within those subgroups may not have a causal interpretation at all---and when they do, the same techniques that work for subgroups defined at baseline %(such as OLS with interactions)
+will often produce inconsistent estimates.
+
+Principal stratification \citep{frangakis} is a broad framework for defining subgroup effects for post-treatment subgroups.
+It relies on the notion of \emph{potential} subgroups called principal strata: groups of subjects defined based on the values a post-treatment variable would potentially take if the subject were assigned to one treatment condition or another.
+Since principal strata membership is based on subjects' responses to both factual and counter-factual treatment assignments, it is in general unobserved.
+Nonetheless, researchers have devised estimators of principal effects, as average effects within principal strata are called, to address a broad range of applied statistical problems, including non-compliance in randomized experiments \citep{air}, the evaluation of surrogate outcomes \citep{li2010bayesian}, the investigation of causal mechanisms \citep{lidsayPage}, and sample attrition \citep{zhangRubin, ding2011}.
 
@@
 In a separate, if related, line of research, \citet{jo2002}, \citet{ding2011}, \citet{jiangDing2021}, and others have shown that, under certain circumstances, researchers can use baseline covariates to identify and estimate principal effects without relying on principal ignorability.
-This paper expands on those identification results by presenting a new set of principal effect estimators, and associated standard error estimates, based on a generalized estimating equations (or M-estimation or method-of-moments) \citep{stefanskiBoos}.
-It turns out that, under certain general conditions, principal effects can be estimated with a simple, two-step procedure: first, estimate principal scores with logistic regression or another M-estimator; then, impute unknown principal subgroup membership indicators with principal scores and estimate principal effects using OLS with interactions, just as in the pre-treatment subgroup case.
+This paper expands on those identification results by presenting a new set of principal effect estimators, and associated standard error estimates, based on a general estimating equations (or M-estimation or method-of-moments) \citep{stefanskiBoos}.
+It turns out that, under certain general conditions, principal effects can be estimated with a simple, two-step procedure: first, estimate principal scores with logistic regression or another M-estimator; then, impute unknown principal subgroup membership indicators with principal scores and estimate principal effects using ordinary least squares (OLS) regression with interactions.
+%, just as in the pre-treatment subgroup case.
 We will refer to the estimators presented here as ``GEEPERs'': generalized estimating equations for principal effects using regressions.
-GEEPERs estimates rely on correct specification of regression functions, but unlike common fully-parametric methods, they make no assumptions about the distribution of regression errors.
+GEEPERs estimates rely on correct specification of regression functions, but unlike common fully-parametric methods, they make no assumptions about the shape of the distribution of regression errors.
 Unlike principal score weighting, GEEPERs does not assume principal ignorability, and admits to sandwich-style standard error estimates, which we provide in the appendix.
@@
 In a randomized experiment with two conditions, let $Z_i=0$ if subject $i$, $i=1,\dots,n$ is assigned to the treatment condition, and let $Z_i=0$ otherwise.
-If interest is in the effect of $Z$ on an outcome $Y$, under the ``stable unit treatment value assumption'' \citep{air} of no interference between units and no hidden versions of the treatment, define potential outcomes \citep{splawa1990application,air} $\yti$ as the value of $Y$ that $i$ would exhibit if $Z_i=1$, and let $\yci$ be the value if $Z_i=0$.
+If interest is in the effect of $Z$ on an outcome $Y$, under the ``stable unit treatment value assumption'' \citep{air} of no interference between units and no hidden versions of the treatment, define potential outcomes \citep{splawa1990application} $\yti$ as the value of $Y$ that $i$ would exhibit if $Z_i=1$, and let $\yci$ be the value if $Z_i=0$.
 Then the observed value of $Y$ satisfies $Y_i=Z_i\yti+(1-Z_i)\yci$, and the treatment effect of $Z$ on $Y$ for subject $i$ may be defined as $\tau_i=\yti-\yci$.
@@
 \subsection{Estimating Principal Effects}
-In estimating principal effects for a binary $S$, it will be useful to define four conditional means:
+In estimating principal effects for a binary $S$, it will be useful to define four conditional means. For $s\in\{0,1\}$, let 
 \begin{equation}\label{eq:mus}
-  \begin{split}
-    \muc0&=\EE[Y_C|\st=0]\\
-    \muc1&=\EE[Y_C|\st=1]\\
-    \mut0&=\EE[Y_T|\st=0]\\
-    \mut1&=\EE[Y_T|\st=1]
-  \end{split}
+\mucs=\EE[Y_C|\st=s] \mbox{ and }\muts=\EE[Y_T|\st=s]%\mbox{ for }s=0,1
+  % \begin{split}
+  %   \muc0&=\EE[Y_C|\st=0]\\
+  %   \muc1&=\EE[Y_C|\st=1]\\
+  %   \mut0&=\EE[Y_T|\st=0]\\
+  %   \mut1&=\EE[Y_T|\st=1]
+  % \end{split}
 \end{equation}
@@
 \begin{equation*}
-  \begin{split}
-    \eff0&=\mut0-\muc0\\
-    \eff1&=\mut1-\muc1
-  \end{split}
-\end{equation*}
-estimating principal effects is equivalent to estimating the four conditional means.
+  \effs=\muts-\mucs
+  % \begin{split}
+  %   \eff0&=\mut0-\muc0\\
+  %   \eff1&=\mut1-\muc1
+  % \end{split}
+\end{equation*}
+estimating principal effects requires estimating the four conditional means.
 
@@
 In contrast, $\muc0$ and $\muc1$ are not fully identified without further assumptions because $\sti$ and $\yci$ are never observed simultaneously (though they may be nonparametrically bounded; see, e.g., \citealt{bounding}).
-When $S$ is a measure of compliance, and $\tau_i=0$ whenever $\sti=0$ (the ``exclusion restriction''), $\muc1$---and hence $\eff1$---may be estimated with instrumental variables methods \citep{air}.
-In this case, $\eff1$ is often called the ``local average treatment effect'' or the average effect of the treatment on the treated.
-We will briefly review two approaches to estimating $\muc0$ and $\muc1$ that do not assume the exclusion restriction---finite mixture modeling and weighting---with some comments on the role for covariates in-between.
-
-\subsubsection{Finite Mixture Modeling}
-The finite mixture modeling approch to estimating $\muc0$ and $\muc1$ \citep{imbens1997bayesian} assumes that, conditional on $\st$, $Y_C$ is normally distributed.
+% When $S$ is a measure of compliance, and $\tau_i=0$ whenever $\sti=0$ (the ``exclusion restriction''), $\muc1$---and hence $\eff1$---may be estimated with instrumental variables methods \citep{air}.
+% In this case, $\eff1$ is often called the ``local average treatment effect'' or the average effect of the treatment on the treated.
+We will briefly review two approaches to estimating $\muc0$ and $\muc1$% that do not assume the exclusion restriction
+---normal mixture modeling and weighting---with some comments in-between on the role for covariates.
+
+\subsubsection{Normal Mixture Modeling}
+The normal mixture modeling approach to estimating $\muc0$ and $\muc1$ \citep{imbens1997bayesian} assumes that, conditional on $\st$, $Y_C$ is normally distributed.
 Then the probability density of $Y_C$ may be written as
@@
 \end{equation}
-where $\bxsi\subseteq\bm{x}_i$, a not-necessarily-proper subset of covariates used to model $S$.
+where $\bxsi\subseteq\bm{x}_i$, the set of covariates used to model $S$.
 Note that under randomization, $Pr(\sti=1|\bxsi,Z_i)=Pr(\sti=1|\bxsi)$, so a model for principal scores can be estimated using data from the treatment group and extrapolated to the control group.
@@
 \begin{equation}\label{eq:prinScoreMod}
-  \ppip{\bm{\alpha}}=f(\bxs;\bm{\alpha})
+  \ppip{\bm{\alpha}}=f(\bxsi;\bm{\alpha})
 \end{equation}
@@
 \end{ass}
-In some estimators we will discuss below, two distinct values suffice. For a broader discussion, see \citet{ding2011,jiangDing2021}.
+%In some estimators we will discuss below, two distinct values suffice. 
+For a broader discussion, see \citet{ding2011,jiangDing2021}.
 
@@
 \end{equation}
-Ideally, a statistician will estimate parameters $\muc0$, $\muc1$, $\mut0$, $\mut1$, $\sigma^0_C$, $\sigma^1_C$, and $\bm{\alpha}$ simultaneously, using maximum likelihood or Bayesian techniques.
-
-Alternatively, assume ``principal ignorability'' \citep{jo2009use,dingLu}:
+%Ideally, a statistician will estimate parameters $\muc0$, $\muc1$, $\mut0$, $\mut1$, $\sigma^0_C$, $\sigma^1_C$, and $\bm{\alpha}$ simultaneously, using maximum likelihood or Bayesian techniques.
+
+Alternatively, an analyst may assume ``principal ignorability'' \citep{jo2009use,dingLu}:
 \begin{ass}[Principal Ignorability]\label{ass:PI}
@@
 Principal ignorability \eqref{eq:pi} is reminiscent of ignorability assumptions typical in observational studies \citep[e.g.][]{rosenbaum2010design}.
-Both require a rich set of covariates and background knowledge of the determinants of $S$; in particular, an unobserved covariate $u$ that is correlated with both $\st$ and $Y_C$ would invalidate \eqref{eq:pi}.
+% Both require a rich set of covariates and background knowledge of the determinants of $S$; 
+In particular, an unobserved covariate $u$ that is correlated with both $\st$ and $Y_C$ would invalidate \eqref{eq:pi}.
 
@@
 
-We begin by introducing an assumption that we term ``Covariate Ignorability,'' or CI. It is a slighty weaker version of \citet{jiangDing2021}'s ``auxiliary independence'' assumption.
+We begin by introducing an assumption that we term ``Covariate Ignorability,'' or CI. It is a slightly weaker version of \citet{jiangDing2021}'s ``auxiliary independence'' assumption.
 \begin{ass}[Covariate Ignorability]\label{ass:ci}
@@
 To see how, we'll %first
-derive the estimating equations for $\muc0$, $\muc1$, $\mut0$, and $\mut1$, and show that they are equivalent to estimating equations for a model fit by OLS.% assuming that principal scores are known exactly, and then generalize to the case in which they are estimated using data from the treatment group.
+derive the estimating equations for $\muc0$, $\muc1$, $\mut0$, and $\mut1$, and show that they are equivalent to estimating equations for a model fit by OLS.
+% assuming that principal scores are known exactly, and then generalize to the case in which they are estimated using data from the treatment group.
 The details of the calculations and the proofs are in the appendix.
@@
     \begin{split}
-      \EE[Y|Z=0]=&\muc0+\EE\pp(\muc1-\muc0)\\
+      \EE[Y|Z=0]&=\muc0+\EE\pp(\muc1-\muc0)\\
       \EE[Y\pp|Z=0]&=\EE[\pp]\muc0+\EE[\pp^2](\muc1-\muc0)\\
@@
 \begin{equation}
-  \begin{split}
-    \hat{\eff0}&\equiv \hat{\beta}_2\\
-    \hat{\eff1}&\equiv \hat{\beta}_3-\hat{\beta}_2
-  \end{split}
-\end{equation}
-are M-estimators, with $\hat{\eff0}\rightarrow_p\eff0$ and $\hat{\eff1}\rightarrow_p\eff1$ as $n\rightarrow\infty$
+  %\begin{split}
+    \heff0\equiv \hat{\beta}_2\mbox{ and }
+    \heff1\equiv \hat{\beta}_3-\hat{\beta}_2
+%  \end{split}
+\end{equation}
+are M-estimators, with $\heff0\rightarrow_p\eff0$ and $\heff1\rightarrow_p\eff1$ as $n\rightarrow\infty$
 \end{prop}
@@
 
-If principal scores were known exactly, the M-estimation structure of $\hat{\eff0}$ and $\hat{\eff1}$ implies that their sampling variance could be estimated by the standard sandwich estimator for OLS \citep{stefanskiBoos}.
-This will rarely be the case.
-However, if the principal scores are modeled as \eqref{eq:prinScoreMod}, and parameters $\bm{\alpha}$ are estimated by M-estimation, with estimating equations $\sum_{i: Z_i=1}\Omega_i=\bm{0}$, then the estimating equations for the principal score model and principal effect estimation can be stacked \citep[c.f.][]{boosStefanskiBook}, as
+% If principal scores were known exactly, the M-estimation structure of $\heff0$ and $\heff1$ implies that their sampling variance could be estimated by the standard sandwich estimator for OLS \citep{stefanskiBoos}.
+% This will rarely be the case.
+%However, 
+If the principal scores are modeled as \eqref{eq:prinScoreMod}, and parameters $\bm{\alpha}$ are estimated by M-estimation, with estimating equations $\sum_{i: Z_i=1}\Omega_i=\bm{0}$, then the estimating equations for the principal score model and principal effect estimation can be stacked \citep[c.f.][]{boosStefanskiBook}, as
 \begin{equation}\label{eq:stacked}
@@
 
-Then, the standard error matrix for $[\bm{\hat{\alpha}},\bm{\hat{\beta}}$ can be estimated as \citep[][ch. 7]{boosStefanskiBook}:
+Then, the standard error matrix for $[\bm{\hat{\alpha}},\bm{\hat{\beta}}]$ can be estimated as \citep[][ch. 7]{boosStefanskiBook}:
 \begin{equation}\label{eq:sandwich}
@@
 \item Using OLS, fit model \eqref{eq:regression0}
-\item Estimate $\hat{\tau^0}=\hat{\beta}_2$ and $\hat{\tau^1}=\hat{\beta}_3-\hat{\beta}_2$
+\item Estimate $\heff0=\hat{\beta}_2$ and $\heff1=\hat{\beta}_3-\hat{\beta}_2$ with sampling variance \eqref{eq:sandwich}
 \end{enumerate}
@@
 Let $\bxy\in \mathbb{R}^k$, be a vector of covariates (possibly the same as $\bxs$); $\bxy$ can contain transformations and interactions of baseline covariates, allowing for a wide range of models.
-Then note that $\st$ is, essentially, a covariate, so by randomization, (Assumption \ref{ass:rand}), $\EE[\bxy|\st]=\EE[\bxy|\st,Z=0]=\EE[\bxy|\st,Z=1]$.
-Then information about the stratum-specific means of $\bxy$ can be extrapolated from the treatment to the control group.
+By randomization, (Assumption \ref{ass:rand}), $\EE[\bxy|\st]=\EE[\bxy|\st,Z=0]=\EE[\bxy|\st,Z=1]$, so 
+estimates of principal-stratum-specific means of $\bxy$ can be extrapolated from the treatment to the control group.
 This leads to the following proposition:
@@
   \begin{equation}\label{eq:interaction}
-    Y_i=\beta_0+\beta_1\sti+\beta_2 Z_i+\beta_3Z_i\sti+\bm{\gamma_1}'\bxy_i+\bm{\gamma_2}'\bxy_i\sti+
+    \begin{split}
+    Y_i=&\beta_0+\beta_1\sti+\beta_2 Z_i+\beta_3Z_i\sti\\
+    &+\bm{\gamma_1}'\bxy_i+\bm{\gamma_2}'\bxy_i\sti+
     \bm{\gamma_3}'\bxy Z_i+\bm{\gamma_4}'\bxy_i Z_i\sti+\epsilon_i
-  \end{equation}
-  with $\EE[\epsilon_i]=0$, $\{Z_i,\sti,\bxy_i\}\independent \epsilon_i$, and $\epsilon_i\independent\epsilon_j$ when $i\ne j$.
+    \end{split}
+  \end{equation}
+  with $\EE[\epsilon_i|\bxi,Z_i]=\EE[\epsilon_i|\bxi,Z_i,\sti]=0$ and $\epsilon_i\independent\epsilon_j$ when $i\ne j$.
 
@@
   \begin{split}
-    \hat{\eff0}_{int}&\equiv \hat{\beta}_2+\bm{\hat{\gamma}_3}'\overline{\bxy}_{Z=1,S=0}\\
-    \hat{\eff1}_{int}&\equiv \hat{\beta}_2+\hat{\beta}_3+(\bm{\hat{\gamma}_3}+\bm{\hat{\gamma}_4})'\overline{\bxy}_{Z=1,S=1}
+    \heff0_{int}&\equiv \hat{\beta}_2+\bm{\hat{\gamma}_3}'\overline{\bxy}_{Z=1,S=0}\\
+    \heff1_{int}&\equiv \hat{\beta}_2+\hat{\beta}_3+(\bm{\hat{\gamma}_3}+\bm{\hat{\gamma}_4})'\overline{\bxy}_{Z=1,S=1}
   \end{split}
@@
   \end{enumerate}
-  Then $\hat{\eff0}_{G}$ and $\hat{\eff1}_{int}$ are M-estimators, with $\hat{\eff0}_{int}\rightarrow_p\eff0$ and $\hat{\eff1}_{G}\rightarrow_p\eff1$ as $n\rightarrow\infty$.
+  Then $\heff0_{G}$ and $\heff1_{int}$ are M-estimators, with $\heff0_{int}\rightarrow_p\eff0$ and $\heff1_{G}\rightarrow_p\eff1$ as $n\rightarrow\infty$.
   Under suitible regularity conditions, they are jointly asymptotically normal, with a variance of the form \eqref{eq:sandwich}.
@@
 In short, researchers can estimate principal effects under strong monotonicity by simply imputing missing $\st$ values with estimated principal scores, and fitting an OLS regression with appropriate interaction terms---just as if $\st$ were completely observed.
+
+As in the simpler case discussed in \S~\ref{sec:ci}, since the estimators in \eqref{eq:prinEffEst} are M-Estimators, their sampling variance can be estimated with a consistent sandwich estimator.
+Uncertainty in principal score estimation can be incorporated by stacking the estimating equations for the principal score model \eqref{eq:prinScoreMod} and estimating the combined covariance matrix \eqref{eq:sandwich}, formulas for which are given in the appendix.
 
@@
 Such a model relies on what \citet{jo2002} calls ``Additivity of Treatment Assignment Effect,'' that $\bm{\gamma_3}=\bm{\gamma_4}=0$.
-In other words, the only interactions between covariates and treatment assignment are due to differences between principal effects---a fairly strong assumpton.
+In other words, the only interactions between covariates and treatment assignment are due to differences between principal effects---a fairly strong assumption.
 However, there are two mitigating points.
@@
 Second, simulation results in the next section suggest that some models excluding important interactions may still yield reliable inferences.
-
-As in the simpler case discussed in \S~\ref{sec:ci}, since the estimators in \eqref{eq:prinEffEst} are M-Estimators, their sampling variance can be estimated with a consistent sandwich estimator.
-Uncertainty in principal score estimation can be incorporated by stacking the estimating equations for the principal score model \eqref{eq:prinScoreMod} and estimating the combined covariance matrix \eqref{eq:sandwich}, formulas for which are given in the appendix.
 
@@
 where $\alpha$ is a manipulated factor.
-When $\alpha$ was higher, $\st$ was more easily predicted by covariates; when $\alpha=0$, Assumption \ref{ass:vps}, was violated, i.e. $Var(\ppi)=0$.
+When $\alpha$ was higher, $\st$ was more easily predicted by covariates; when $\alpha=0$, Assumption \ref{ass:vps}, was violated, i.e. $Var(\pp)=0$.
 Then principal stratum $\sti\in \{0,1\}$ was simulated as $S_i\sim Bern(\ppi)$
@@
 \begin{equation}\label{eq:y-sim}
-  \begin{split}
-  \yci&=\beta_1\sti+(\gamma_1+\gamma_2\sti)(x_{1i}+x_{2i})+\frac{1}{\sqrt{6}}x_{3i}+\epsilon_i\\
-  \yti&=0.3\sti+(\gamma_1+\gamma_2\sti+\gamma_3Z_i)(x_{1i}+x_{2i})+\frac{1}{\sqrt{6}}x_{3i}+\epsilon_i\\
-  Y_i&=Z_i\yti+(1-Z_i)\yci
-  \end{split}
-\end{equation}
-Hence, $\beta_0=\beta_2=0$, and $\beta_3=0.3-\beta_1$.
-When $\gamma_3=0$ (i.e.
+Y_i=\beta_1\sti+\beta_3Z_i\sti+(\gamma_1+\gamma_2\sti+\gamma_3Z_i)(x_{1i}+x_{2i})+\frac{1}{\sqrt{6}}x_{3i}+\epsilon_i%\\
+\end{equation}
+%   \begin{split}
+%   \yci&=\beta_1\sti+(\gamma_1+\gamma_2\sti)(x_{1i}+x_{2i})+\frac{1}{\sqrt{6}}x_{3i}+\epsilon_i\\
+%   \yti&=0.3\sti+(\gamma_1+\gamma_2\sti+\gamma_3Z_i)(x_{1i}+x_{2i})+\frac{1}{\sqrt{6}}x_{3i}+\epsilon_i\\
+%   Y_i&=Z_i\yti+(1-Z_i)\yci
+%   \end{split}
+% \end{equation}
+with $\beta_3=0.3-\beta_1$.
+Hence, $\beta_0=\beta_2=\gamma_4=0$, when $\gamma_3=0$ (i.e.
 no interaction between $Z$ and covariates), there is no average treatment effect in the $\st=0$ stratum, the coefficient for the unobserved covariate $x_3$ is constant across treatment groups and principal strata, and we do not consider three-way interactions between $x_1$ or $x_2$, $\st$, and $Z$.
@@
 
-For the Bayesian mixture model and GEEPERs we estimated treatment effects as $\hat{\eff0}=\hat{\beta}_2$ and $\hat{\eff1}=\hat{\beta}_2+\hat{\beta}_3$.
+For the Bayesian mixture model and GEEPERs we estimated treatment effects as $\heff0=\hat{\beta}_2$ and $\heff1=\hat{\beta}_2+\hat{\beta}_3$.
 
 For principal score weighting we extracted the estimated probabilities from \eqref{eq:psModsim}---the principal scores---plugged them in to \eqref{eq:psw} to estimate $\muc0$ and $\muc1$, and estimated $\mut0$ and $\mut1$ with the sample means of subjects with $Z=1$ and $S=0$ or $S=1$, respectively.
-Since the PSW estimate did not use an outcome model, we estimated $\hat{\eff0}=\hat{\mut0}-\hat{\muc0}$ and $\hat{\eff1}=\hat{\mut1}-\hat{\muc1}$.
+Since the PSW estimate did not use an outcome model, we estimated $\heff0=\hat{\mut0}-\hat{\muc0}$ and $\heff1=\hat{\mut1}-\hat{\muc1}$.
 
@@
 
-[NOTE: I NEED TO MAKE SURE THE NOTATION HERE MATCHES ABOVE]
 \section{Proofs and Calculations}
@@
 \begin{equation*}
-  \displaystyle\sum_i \begin{pmatrix}
+\begin{split}
+  &\psi_i=\\ &\begin{pmatrix}
   Y_i\\
@@
   Y_i\bxy_iZ_i\ri\\
-  \phantom{}\end{pmatrix} - \begin{pmatrix}
-    \beta_0+\beta_1R_i+\beta_2Z_i+\beta_3Z_iR_i\phantom{-----}\\
-    \hfill+\left\{\bm{\gamma_1}'+R_i\bm{\gamma_2}'+Z_i\bm{\gamma_3}'+R_iZ_i\bm{\gamma_4}'\right\}\bxy\\
-    \beta_0\ri+\beta_1R_i^2+\beta_2Z_i\ri+\beta_3Z_iR_i^2\phantom{-----}\\
-    \hfill+\ri\left\{\bm{\gamma_1}'+R_i\bm{\gamma_2}'+Z_i\bm{\gamma_3}'+R_iZ_i\bm{\gamma_4}'\right\}\bxy\\
-    (\beta_0+\beta_2)Z_i+(\beta_1+\beta_3)R_iZ_i\phantom{-----}\\
-    \hfill+Z_i\left\{\bm{\gamma_1}'+R_i\bm{\gamma_2}'+Z_i\bm{\gamma_3}'+R_iZ_i\bm{\gamma_4}'\right\}\bxy\\
-    \beta_0Z_i\ri+(\beta_1+\beta_3)Z_iR_i^2+\beta_2Z_i\ri\phantom{-----}\\
-    \hfill+Z_iR_i\left\{\bm{\gamma_1}'+\bm{\gamma_3}'+R_i(\bm{\gamma_2}'+\bm{\gamma_4}')\right\}\bxy \hfill\\
-    \beta_0\bxyt+\beta_1R_i\bxyt+\beta_2Z_i\bxyt+\beta_3Z_iR_i\bxyt\phantom{---}\\
-    \hfill {}+\left\{\bm{\gamma_1}'+R_i\bm{\gamma_2}'+Z_i\bm{\gamma_3}'+R_iZ_i\bm{\gamma_4}'\right\}\bxy\bxyt\\
-    \beta_0\ri\bxyt+\beta_1R_i^2\bxyt+\beta_2Z_i\ri\bxyt+\beta_3Z_iR_i^2\bxyt\phantom{---}\\
-    \hfill {}+\ri\left\{\bm{\gamma_1}'+R_i\bm{\gamma_2}'+Z_i\bm{\gamma_3}'+R_iZ_i\bm{\gamma_4}'\right\}\bxy\bxyt\\
-        (\beta_0+\beta_2)Z_i\bxyt+(\beta_1+\beta_3)R_iZ_i\bxyt\phantom{-----}\\
-    \hfill {}+Z_i\left\{\bm{\gamma_1}'+R_i\bm{\gamma_2}'+Z_i\bm{\gamma_3}'+R_iZ_i\bm{\gamma_4}'\right\}\bxy\bxyt\\
-\beta_0\ri Z_i\bxyt+\beta_1R_i^2Z_i\bxyt+\beta_2\ri Z_i\bxyt+\beta_3Z_iR_i^2\bxyt\phantom{---}\\
-    \hfill {}+\ri Z_i\left\{\bm{\gamma_1}'+R_i\bm{\gamma_2}'+Z_i\bm{\gamma_3}'+R_iZ_i\bm{\gamma_4}'\right\}\bxy\bxyt\\
-  \end{pmatrix}=\bm{0}
-\end{equation*}
+  \phantom{}\end{pmatrix} - \begin{pmatrix*}[l]
+    \beta_0+\beta_1R_i+\beta_2Z_i+\beta_3Z_iR_i\\
+    \quad+\left\{\bm{\gamma_1}'+R_i\bm{\gamma_2}'+Z_i\bm{\gamma_3}'+R_iZ_i\bm{\gamma_4}'\right\}\bxy\\
+    \beta_0\ri+\beta_1R_i^2+\beta_2Z_i\ri+\beta_3Z_iR_i^2\\
+    \quad+\ri\left\{\bm{\gamma_1}'+R_i\bm{\gamma_2}'+Z_i\bm{\gamma_3}'+R_iZ_i\bm{\gamma_4}'\right\}\bxy\\
+    (\beta_0+\beta_2)Z_i+(\beta_1+\beta_3)R_iZ_i\\
+    \quad+Z_i\left\{\bm{\gamma_1}'+R_i\bm{\gamma_2}'+Z_i\bm{\gamma_3}'+R_iZ_i\bm{\gamma_4}'\right\}\bxy\\
+    \beta_0Z_i\ri+(\beta_1+\beta_3)Z_iR_i^2+\beta_2Z_i\ri\\
+    \quad+Z_iR_i\left\{\bm{\gamma_1}'+\bm{\gamma_3}'+R_i(\bm{\gamma_2}'+\bm{\gamma_4}')\right\}\bxy \\
+    \beta_0\bxyt+\beta_1R_i\bxyt+\beta_2Z_i\bxyt+\beta_3Z_iR_i\bxyt\\
+    \quad {}+\left\{\bm{\gamma_1}'+R_i\bm{\gamma_2}'+Z_i\bm{\gamma_3}'+R_iZ_i\bm{\gamma_4}'\right\}\bxy\bxyt\\
+    \beta_0\ri\bxyt+\beta_1R_i^2\bxyt+\beta_2Z_i\ri\bxyt+\beta_3Z_iR_i^2\bxyt\\
+    \quad {}+\ri\left\{\bm{\gamma_1}'+R_i\bm{\gamma_2}'+Z_i\bm{\gamma_3}'+R_iZ_i\bm{\gamma_4}'\right\}\bxy\bxyt\\
+        (\beta_0+\beta_2)Z_i\bxyt+(\beta_1+\beta_3)R_iZ_i\bxyt\\
+    \quad {}+Z_i\left\{\bm{\gamma_1}'+R_i\bm{\gamma_2}'+Z_i\bm{\gamma_3}'+R_iZ_i\bm{\gamma_4}'\right\}\bxy\bxyt\\
+\beta_0\ri Z_i\bxyt+\beta_1R_i^2Z_i\bxyt+\beta_2\ri Z_i\bxyt+\beta_3Z_iR_i^2\bxyt\\
+    \quad {}+\ri Z_i\left\{\bm{\gamma_1}'+R_i\bm{\gamma_2}'+Z_i\bm{\gamma_3}'+R_iZ_i\bm{\gamma_4}'\right\}\bxy\bxyt\\
+  \end{pmatrix*}
+  \end{split}
+\end{equation*}
+with $\EE[\psi_i]=0$.
 These are the estimating equations for the regression model \eqref{eq:interaction}, with $R$ replacing $\st$, fit by OLS. 
@@
 
+\sloppy
 Following \eqref{eq:stacked}, let $\blam(Z_i,S_i,\bxsi,\bxy_i,Y_i;\bm{\alpha},\bm{\beta})=\begin{pmatrix} Z_i\Omega(\bxsi,S_i;\bm{\alpha})\\ \Psi(\bxy_i,\bxsi,Y_i,Z_i,\ri(\bm{\alpha});\bm{\beta})\end{pmatrix}$, the stacked estimating equations of \eqref{eq:psMod} and \eqref{eq:regression}.
-Going forward, for the sake of brevety, we will write $\blam_i(\bm{\alpha},\bm{\beta})=\blam(Z_i,S_i,\bxsi,\bxy_i,Y_i;\bm{\hat{\alpha}},\bm{\hat{\beta}})$, where dependence on the data for $i$ is captured in the subscript $i$, with similiar meanings for $\Psi_i(\bm{\alpha},\bm{\beta})$ and $\blam_i(\bm{\alpha},\bm{\beta})$.
+Going forward, for the sake of brevety, we will write $\blam_i(\bm{\alpha},\bm{\beta})=\blam(Z_i,S_i,\bxsi,\bxy_i,Y_i;\bm{\hat{\alpha}},\bm{\hat{\beta}})$, where dependence on the data for $i$ is captured in the subscript $i$, with similar meanings for $\Psi_i(\bm{\alpha},\bm{\beta})$ and $\omega_i(\bm{\alpha})$.
 %Then let $\blamh_i=\blam_i(Z_i,S_i,\bxsi,\bxy_i,Y_i;\bm{\hat{\alpha}},\bm{\hat{\beta}})$ be the estimating equations evaluated at the estimated parameters, and let 

### 2022-12-13 code/fh2t/analysisGEE.r
add ATEs to compare methods plot

@@ -63,8 +63,11 @@ estimates0 <- lapply(alts,
                           covFormU=formula(psMod7)[-2])
                     )
 
+save(ates,estimates0,estimates3,estimatesInt1,file='results/geeResults.RData')
+
 sapply(estimates0,effsFromFit,simplify=F)
 
+### model checking
 pdf('results/outcomeModels0.pdf')
 iwalk(estimates0,~plot(.x$outMod,which=1,main=.y))
 dev.off()
@@ -84,32 +87,32 @@ iwalk(estimates1,
       ~print(ggplot(aes(fitted(.x$outMod),resid(.x$outMod)))+geom_jitter(height=0.3)+ggtitle(.y)))#",main=.y)))
 dev.off()
 
-estimates2 <- lapply(setNames(alts,alts),
-                    function(alt)
-                      est(getDat(alt),#droplevels(filter(psdat,trt%in%c('ASSISTments',alt))),
-                          covFormU=formula(psMod7)[-2],
-                          covFormY=update(formula(psMod7)[-2],.~.-SchIDPre+pred+
-                                                                splines::ns(pre.total_math_score,3)+
-                                                                raceEth+
-                                                                Gender+
-                                                                GIFTED
-                                          ),
-                          block='ClaIDPre')
-                    )
+# estimates2 <- lapply(setNames(alts,alts),
+#                     function(alt)
+#                       est(getDat(alt),#droplevels(filter(psdat,trt%in%c('ASSISTments',alt))),
+#                           covFormU=formula(psMod7)[-2],
+#                           covFormY=update(formula(psMod7)[-2],.~.-SchIDPre+pred+
+#                                                                 splines::ns(pre.total_math_score,3)+
+#                                                                 raceEth+
+#                                                                 Gender+
+#                                                                 GIFTED
+#                                           ),
+#                           block='ClaIDPre')
+#                     )
 
 
-iwalk(estimates2,
[…]

### 2022-12-13 code/fh2t/plots.r

@@ -1,4 +1,8 @@
+library(tidyverse)
 library(coefplot)
+
+load('results/geeResults.RData')
+load('results/psw.RData')
 
 ## ps models
 coefplot(estimates3$BAU$psMod, predictors=names(model.frame(estimates3$BAU$psMod))[-c(1:2)],
@@ -16,7 +20,7 @@ ggsave('figure/psModCoef.jpg',width=5,height=3,units='in')
 
 ### plot estimates
 
-ateDF <- imap_dfr(ates,~data.frame(EFF=.25,Alternative=ifelse(.y=='Dragon','DragonBox',.y),estimates=.x$coefficients['Z'],
+ateDF <- imap_dfr(ates,~data.frame(EFF=.25,Alternative=paste('vs.',ifelse(.y=='Dragon','DragonBox',.y)),estimates=.x$coefficients['Z'],
                               ymin=.x$conf.low['Z'],ymax=.x$conf.high['Z']))
 
 bind_rows(
@@ -31,11 +35,11 @@ bind_rows(
 )%>%
   filter(startsWith(eff,'eff'))%>%
   mutate(
-    Alternative=ifelse(Alternative=='Dragon','DragonBox',Alternative),
+    Alternative=paste("vs.",ifelse(Alternative=='Dragon','DragonBox',Alternative)),
     EFF=ifelse(eff=='eff0',ifelse(int=="None",-.05,.05),ifelse(int=="None",.45,.55)),
     ymin=estimates-2*SE,
     ymax=estimates+2*SE
-  )%>%
+  ) %>%
   ggplot(aes(EFF,estimates,color=int,ymin=ymin,ymax=ymax))+
   geom_point()+geom_errorbar(width=0)+
   geom_point(data=ateDF,aes(EFF,estimates),color='black',inherit.aes=FALSE)+
@@ -69,16 +73,18 @@ bind_rows(
 )%>%
   filter(startsWith(eff,'eff'))%>%
   mutate(
-    Alternative=ifelse(Alternative=='Dragon','DragonBox',Alternative),
+    Alternative=paste("vs.",ifelse(Alternative=='Dragon','DragonBox',Alternative)),
     EFF=ifelse(eff=='eff0',ifelse(method=="GEEPERS",-.07,ifelse(method=='Mixture',0,.07)),ifelse(method=="GEEPERS",.43,ifelse(method=='Mixture',.5,.57))),
     ymin=estimates-2*SE,
     ymax=estimates+2*SE
[…]

### 2022-12-13 writeUps/paper.tex

@@
 \begin{lemma}\label{lemma:expectation}
-  Under Assumptions \ref{ass:sm}, \ref{ass:rand}, \ref{ass:vps}, and \ref{ass:ci}, %if principal scores $\ppi$, for $i:Z_i=0$ are known exactly, then:
-  \begin{equation*}
-    \begin{split}
-      \EE[Y|Z=0]&=\muc0+\EE\pp(\muc1-\muc0)\\
-      \EE[Y\pp|Z=0]&=\EE[\pp]\muc0+\EE[\pp^2](\muc1-\muc0)\\
-      \EE[Y|Z=1]&=\mut0+\EE[\st](\mut1-\mut0)\\
-      \EE[YS|Z=1]&=\EE[\st]\mut0+\EE[\st^2](\mut1-\mut0)
-    \end{split}
-  \end{equation*}
-\end{lemma}
-Since $\st=\st^2$, the expression for $\EE[\st Y_T]$ in Lemma \ref{lemma:expectation} can be simplified to $\EE[\st]\mut1$; however, we retain the unsimplified version, since its similarity with the analogous expression when $Z_i=0$ simplifies the reasoning below.
-
-Lemma \ref{lemma:expectation} in turn implies a set of estimating equations $\sum_{i=1}^n\tilde{\Psi}_i=\bm{0}$, where
-\begin{equation}\label{eq:estEq0}
+Let \begin{equation}\label{eq:estEq0}
 \tilde{\Psi}_i=\begin{pmatrix}
@@
 \end{equation}
-Note that Lemma \ref{lemma:expectation} implies that $\EE[\Psi_i]=\bm{0}$.
+Then under Assumptions \ref{ass:sm}, \ref{ass:rand}, \ref{ass:vps}, and \ref{ass:ci}, $\EE[\Psi_i]=\bm{0}$.
+\end{lemma}
+
+Lemma \ref{lemma:expectation} implies a set of estimating equations $\sum_{i=1}^n\tilde{\Psi}_i=\bm{0}$.
 
@@
 \end{equation}
-Note that, in general, $(1-Z_i)\ppi=(1-Z_i)\ri$ and $Z_i\sti=Z_i\ri$, allowing $\ri$ to take the place of $\ppi$ and $S_i$ throughout $\tilde{\Psi}_i$.
+In general, $(1-Z_i)\ppi=(1-Z_i)\ri$ and $Z_i\sti=Z_i\ri$, allowing $\ri$ to take the place of $\ppi$ and $S_i$ throughout $\tilde{\Psi}_i$.
 
@@
   %\begin{split}
-    \heff0\equiv \hat{\beta}_2\mbox{ and }
-    \heff1\equiv \hat{\beta}_3-\hat{\beta}_2
+    \heff0_{CI}\equiv \hat{\beta}_2\mbox{ and }
+    \heff1_{CI}\equiv \hat{\beta}_3+\hat{\beta}_2
 %  \end{split}
 \end{equation}
-are M-estimators, with $\heff0\rightarrow_p\eff0$ and $\heff1\rightarrow_p\eff1$ as $n\rightarrow\infty$
+are M-estimators, with $\heff0_{CI}\rightarrow_p\eff0$ and $\heff1_{CI}\rightarrow_p\eff1$ as $n\rightarrow\infty$
 \end{prop}
@@
 
-To summarize, under strong monotonicity, randomization, and covariate ignorability \eqref{eq:assumption}, principal effects can be estimated with the following procedure:
-\begin{enumerate}
-\item Estimate parameters $\bm{\alpha}$ as $\bm{\hat{\alpha}}$ by fitting model \eqref{eq:prinScoreMod} using data from subjects with $Z=1$.
-\item Estimate principal scores for subjects with $Z_i=0$ as $\ppip{\bm{\hat{\alpha}}}$
-\item Set $\ri=Z_iS_i+(1-Z_i)\ppip{\bm{\hat{\alpha}}}$
-\item Using OLS, fit model \eqref{eq:regression0}
-\item Estimate $\heff0=\hat{\beta}_2$ and $\heff1=\hat{\beta}_3-\hat{\beta}_2$ with sampling variance \eqref{eq:sandwich}
-\end{enumerate}
-
 \subsection{Relaxing Covariate Ignorability with Outcome Modeling}\label{sec:regression}
@@
 Fortunately, CI may be relaxed or obviated by regression (see \citealt[][\S 3.4]{jiangDing2021} for an analogous identification result).
-Let $\bxy\in \mathbb{R}^k$, be a vector of covariates (possibly the same as $\bxs$); $\bxy$ can contain transformations and interactions of baseline covariates, allowing for a wide range of models.
-By randomization, (Assumption \ref{ass:rand}), $\EE[\bxy|\st]=\EE[\bxy|\st,Z=0]=\EE[\bxy|\st,Z=1]$, so 
-estimates of principal-stratum-specific means of $\bxy$ can be extrapolated from the treatment to the control group.
-This leads to the following proposition:
-\begin{prop}\label{prop:interactions}
-  Say, for $i=1,\dots,n$, principal scores $\ppi$ are generated as \eqref{eq:pscore}, with parameters $\bm{\alpha}$ identified and estimable with M-estimation, and outcomes are generated as
-  \begin{equation}\label{eq:interaction}
+
+The assumption we'll state here is stronger than necessary, but leads to an attractively simple method; a weaker version of the assumption and corresponding method, along with the proof, can be found in an online appendix.
+
+\begin{ass}[Residualized Covariate Ignorability]\label{ass:rci}
+There exists $\bxy$, a (known) transformation of covariates $\bx$, and an (unknown) vector of coefficients $\bm{\gamma}$ such that
+\begin{equation}\label{eq:ass:rci}
 \begin{split}
-    Y_i=&\beta_0+\beta_1\sti+\beta_2 Z_i+\beta_3Z_i\sti\\
-    &+\bm{\gamma_1}'\bxy_i+\bm{\gamma_2}'\bxy_i\sti+
-    \bm{\gamma_3}'\bxy Z_i+\bm{\gamma_4}'\bxy_i Z_i\sti+\epsilon_i
+\EE[\yc-\bm{\gamma}'\bxy|\bxs,\st]&=\EE[\yc-\bm{\gamma}'\bxy|\st]\\
+\EE[\yt-\bm{\gamma}'\bxy|\bxs,\st]&=\EE[\yt-\bm{\gamma}'\bxy|\st]
 \end{split}
 \end{equation}
-  with $\EE[\epsilon_i|\bxi,Z_i]=\EE[\epsilon_i|\bxi,Z_i,\sti]=0$ and $\epsilon_i\independent\epsilon_j$ when $i\ne j$.
-
-  Then, under Assumptions \ref{ass:sm}, \ref{ass:rand}, and \ref{ass:vps}, if $\ppi$ is linearly independent of $\bxy$, a researcher may follow the following procedure to estimate principal effects:
-  \begin{enumerate}
-  \item Estimate principal scores by fitting model \eqref{eq:pscore} to data from the treatment group
-  \item Replace $\sti$ with $\ri$ (as defined in \ref{eq:ri}) in model \eqref{eq:interaction} and fit with OLS
-  \item Estimate principal effects as:
+\end{ass}
+That is, CI may not hold for potential outcomes themselves, but an analogous assumption holds for \emph{residualized} potential outcomes---i.e.
+$\yc$ and $\yt$ after subtracting out a linear function of (possibly transformed) covariates.
+By forcing $\bm{\gamma}$ to be the same for both treatment groups, and (implicitly) for both principal strata, Assumption \ref{ass:rci} assumes that there are no interactions between columns of $\bxy$ and either condition or principal stratum.
+This is akin to the ``Additivity of Treatment Assignment Effect'' assumption in \citet{jo2002}.
+Proposition \ref{prop:interactions} in the online appendix relaxes these requirements, but some preliminary simulations suggest that standard errors from estimates allowing for these interactions will be prohibitively large.
+Simulation results in the following section address some cases in which interactions are present in the data generating model but not in the data analysis.
+
+In any event, Assumption \ref{ass:rci} allows principal effects to be estimated with an OLS model including $\bxy$:
+
+\begin{prop}{GEEPERs}\label{prop:reg2}
+
+Under Assumptions \ref{ass:sm}, \ref{ass:rand}, \ref{ass:vps}, and \ref{ass:rci}, let principal scores be estimated as in \eqref{eq:prinScoreMod} using data from the treatment group and assume they are linearly independent of $\bxy$.
+Then say the researcher fits the following model with OLS:
+\begin{equation}\label{eq:reg2}
+Y_i=\beta_0+\beta_1\ri+\beta_2 Z_i+\beta_3Z_i\ri+\bm{\gamma_1}'\bxy_i+\epsilon_i
+\end{equation}
+where $\ri$ is defined as in \eqref{eq:ri}. 
+Then let 
 \begin{equation}\label{eq:prinEffEst}
   \begin{split}
-    \heff0_{int}&\equiv \hat{\beta}_2+\bm{\hat{\gamma}_3}'\overline{\bxy}_{Z=1,S=0}\\
-    \heff1_{int}&\equiv \hat{\beta}_2+\hat{\beta}_3+(\bm{\hat{\gamma}_3}+\bm{\hat{\gamma}_4})'\overline{\bxy}_{Z=1,S=1}
+    \heff0_{RCI}&\equiv \hat{\beta}_2\\
+    \heff1_{RCI}&\equiv \hat{\beta}_2+\hat{\beta}_3
   \end{split}
    \end{equation}
-   where $\overline{\bxy}_{Z=1,S=0}$ and $\overline{\bxy}_{Z=1,S=0}$ are the vector of covariate sample means for the subsets of subjects with $Z=1$ and $S=0$ or $S=1$, respectively.
-  \end{enumerate}
-  Then $\heff0_{G}$ and $\heff1_{int}$ are M-estimators, with $\heff0_{int}\rightarrow_p\eff0$ and $\heff1_{G}\rightarrow_p\eff1$ as $n\rightarrow\infty$.
+  $\heff0_{RCI}$ and $\heff1_{RCI}$ are M-estimators, with $\heff0_{RCI}\rightarrow_p\eff0$ and $\heff1_{RCI}\rightarrow_p\eff1$ as $n\rightarrow\infty$.
   Under suitible regularity conditions, they are jointly asymptotically normal, with a variance of the form \eqref{eq:sandwich}.
 \end{prop}
-
-In short, researchers can estimate principal effects under strong monotonicity by simply imputing missing $\st$ values with estimated principal scores, and fitting an OLS regression with appropriate interaction terms---just as if $\st$ were completely observed.
-
-As in the simpler case discussed in \S~\ref{sec:ci}, since the estimators in \eqref{eq:prinEffEst} are M-Estimators, their sampling variance can be estimated with a consistent sandwich estimator.
-Uncertainty in principal score estimation can be incorporated by stacking the estimating equations for the principal score model \eqref{eq:prinScoreMod} and estimating the combined covariance matrix \eqref{eq:sandwich}, formulas for which are given in the appendix.
-
-Proposition \ref{prop:interactions} would seem to nearly solve the problem of estimating principal effects under strong monotonicity, or at least seem to reduce it to the difficulty of other common problems in causal modeling.
-Unfortunately, simulation studies presented in the appendix suggest that, although consistent estimates for principal effects may be available, these estimates are far too inefficient for practical use; standard errors will be prohibitively high in all but the largest sample sizes.
-For instance, if there is one standard normal covariate $x$, and $\pp_i=\left[1+exp\{-x_i/2\right]^{-1}$, then $Var\left(\hat{\muc0}_{int}\right)/Var(Y_C)\approx 8,500/n$.
-In contrast, an analogous simulation in which Assumption \ref{ass:ci} holds, and the estimation routine of Proposition \ref{prop:reg1} was used, $Var\left(\hat{\muc0}_{int}\right)/Var(Y_C)\approx 15/n$.
-
-Fortunately, simpler models can be fit much more efficiently.
-In particular, if interactions between covariates and treatment assignment are omitted, then data from the treatment group---in which $\st$ is fully observed---can be used to estimate the coefficients on covariates.
-Such a model relies on what \citet{jo2002} calls ``Additivity of Treatment Assignment Effect,'' that $\bm{\gamma_3}=\bm{\gamma_4}=0$.
-In other words, the only interactions between covariates and treatment assignment are due to differences between principal effects---a fairly strong assumption.
-However, there are two mitigating points.
-First, since there is no requirement that all relevant covariates be included in the model, a researcher can exclude covariates from a model if he or she thinks their relationship with $Y$ may vary substantially between treatment groups, even in the same principal stratum.
-Second, simulation results in the next section suggest that some models excluding important interactions may still yield reliable inferences.
-
-To summarize, incorporating covariates into an outcome regression can substantially relax covariate ignorability, but estimation is only practical if  if the regression omits interactions between covariates and $Z$.
+In short, researchers can estimate principal effects under strong monotonicity by simply imputing missing $\st$ values with estimated principal scores, fitting an OLS regression, and estimating standard errors with a sandwich formula.
+For the remainder of the paper, we will refer to principal effect estimators $\heff1_{RCI}$ and $\heff0_{RCI}$---our preferred estimators---as ``GEEPERs.''
 
@@
 
-The methods all used the same ``observed'' dataset, consisting of outcomes $Y$, treatment assignments $Z$, two covariates $x_1$ and $x_2$, and $S$ observed only when $Z=1$.
+The methods all used the same ``observed'' dataset, consisting of outcomes $Y$, treatment assignments $Z$, two covariates $\bxs=\bxy=\bx=[x_1,x_2]$, and $S$ observed only when $Z=1$.
 The third covariate used in the data generating model, $x_3$, was ``unobserved''--since $x_3$ was correlated with both $\st$ and $Y$, Principal Ignorability (Assumption \ref{ass:PI}) was violated.
 
-All three methods used the same principal score model, fit using data from the treatment group in which $S=\st$ was observed:
-\begin{equation}\label{eq:psModsim}
-  \ppf{\bm{\alpha}}{x_1,x_2}=logit^{-1}(\alpha_0+\alpha_1x_1+\alpha_2x_2)
-\end{equation}
-
-The mixture estimator then fit two outcome models.
-For the treatment group, a normal regression:
-\begin{equation}\label{eq:mix1-sim}
-  f_{Y|Z=1}(y_i)=\phi\left(\frac{y_i-[\beta_0+\beta_1S_i+\beta_2 +\bm{\gamma}_1^T\bm{x}_{i,1:2}]}{\sigma_T}\right)
-\end{equation}
-and for the control group, mixture model:
-\begin{equation}\label{eq:mix-sim}
-\begin{split}
-  f_{Y|Z=0}(y_i)&=\ppf{\bm{\alpha}}{\bm{x}_{i,1:2}}\phi\left(\frac{y_i-[\beta_0+\beta_1+\bm{\gamma}_1^T\bm{x}_{i,1:2}]}{\sigma_C}\right)\\
-  &+(1-\ppf{\bm{\alpha}}{\bm{x}_{i,1:2}})\phi\left(\frac{y_i-[\beta_0+\bm{\gamma}_1^T\bm{x}_{i,1:2}]}{\sigma_C}\right)
-\end{split}
-\end{equation}
-Note that the coefficients $\bm{\gamma}_1$ are the same across treatment groups and principal strata, while the residual variance differs across treatment groups, but not across principal strata.
-The models \eqref{eq:psModsim}, \eqref{eq:mix1-sim}, and \eqref{eq:mix-sim} were fit simultaneously using Bayesian Markov Chain Monte Carlo with Stan \citep{rstan}.
+All three methods used the same principal score model, a logistic regression of $S$ on $\bx$, and an intercept, fit using data from the treatment group in which $S=\st$ was observed, i.e.
+$\ppf{\bm{\alpha}}{\bx}=logit^{-1}(\alpha_0+\bm{\alpha}'\bx)$.
+
+The mixture estimator then fit two outcome models:
+For the treatment group, the regression model $Y=\beta_0^T+\beta_1^TS+\bm{\gamma}'\bx+\epsilon$ and for the control group, a mixture model of the form \eqref{eq:mixtureModelPS}, but with $\muc0$ and $\muc1$ replaced by regression functions $\beta_0^C+\bm{\gamma}'\bx$ and $\beta_0^C+\beta_1^C+\bm{\gamma}'\bx$, respectively.
+Residual variance were allowed to differ across treatment groups, but not across principal strata.
+These models were fit simultaneously using Bayesian Markov Chain Monte Carlo with Stan \citep{rstan}, and principal effects $\eff0$ and $\eff1$ estimated as $\beta_0^T-\beta_0^C$ and $\beta_0^T+\beta_1^T-\beta_0^C-\beta_1^T$, respectively.
 We interpreted the posterior mean as a point estimate and the posterior standard deviation as a standard error.
 
-In contrast, for GEEPERs and principal score weighting the principal score model \eqref{eq:psModsim} was fit by itself, to data from the treatment group, using standard maximum likelihood methods.
-For GEEPERs, we used the estimated principal scores in the control group, along with the observed $S$ in the treatment group, to construct $R$, and then fit the model
-\begin{equation}\label{eq:simReg}
-  Y_i=\beta0+\beta_1 \ri+\beta_2Z_i+\beta_3Z_iR_i+\bm{\gamma}_1^T\bm{x}_{i,1:2}+\epsilon_i
-\end{equation}
-with OLS.
+In contrast, for GEEPERs and principal score weighting the principal score model was fit by itself, to data from the treatment group, using standard maximum likelihood methods.
+For GEEPERs, we used the estimated principal scores in the control group, along with the observed $S$ in the treatment group, to construct $R$, and then fit the model \eqref{eq:reg2} with OLS. 
 To estimate standard errors, we used the \texttt{sandwich} package \citep{sandwich} to extract estimating equations from \eqref{eq:psModsim} and \eqref{eq:simReg} and computed the sandwich covariance matrix following the procedure detailed in the appendix.
 
-For the Bayesian mixture model and GEEPERs we estimated treatment effects as $\heff0=\hat{\beta}_2$ and $\heff1=\hat{\beta}_2+\hat{\beta}_3$.
-
-For principal score weighting we extracted the estimated probabilities from \eqref{eq:psModsim}---the principal scores---plugged them in to \eqref{eq:psw} to estimate $\muc0$ and $\muc1$, and estimated $\mut0$ and $\mut1$ with the sample means of subjects with $Z=1$ and $S=0$ or $S=1$, respectively.
-Since the PSW estimate did not use an outcome model, we estimated $\heff0=\hat{\mut0}-\hat{\muc0}$ and $\heff1=\hat{\mut1}-\hat{\muc1}$.
+For principal score weighting we plugged estimated principal scores in to \eqref{eq:psw} to estimate $\muc0$ and $\muc1$, estimated $\mut0$ and $\mut1$ with the sample means of subjects with $Z=1$ and $S=0$ or $S=1$, respectively, and estimated $\heff0=\hat{\mut0}-\hat{\muc0}$ and $\heff1=\hat{\mut1}-\hat{\muc1}$.
 
@@
 When $\alpha=0$, Assumption \ref{ass:vps} is violated, and M-estimates are highly erratic; however, as Table \ref{tab:coverage} shows, when other assumptions hold inference remains valid.
-For the sake of clarity, we will focus here on cases where $\alpha>0$; results for $\alpha=0$ are in Tables \ref{tab:rmse} and \ref{tab:coverage} and in the online appendix.
-
-Figures \ref{fig:n} and \ref{fig:alpha} show the empirical bias and standard error (SE) of the M-estimator and mixture estimator of $\eff1$ as sample size and $\alpha$ vary.
+
+Figure \ref{fig:alphan} shows the empirical bias and standard error (SE) of the GEEPERs and mixture estimator of $\eff1$ as sample size and $\alpha$ vary.
 Across $n$ and $\alpha$, GEEPERs has lower bias and higher SE than the mixture estimator.
@@
 
-\begin{figure}
+\begin{figure}[!ht]
   \centering
@@
 
-Figure \ref{fig:boxplots} shows violin plots, layered on top of jittered scatterplots, showing estimation error for GEEPERs, the mixture estimator, and the PSW estimator of $\eff1$ under various sets of conditions, and Table \ref{tab:rmse} shows root mean squared error (RMSE) estimates under a somewhat wider set of circumstances when $\muc1=0.3$.
-The leftmost panel of Figure \ref{fig:boxplots} and the top line of Tables \ref{tab:rmse} and \ref{tab:coverage} show results under ``standard'' conditions---normal residuals and no interactions in the data generating model between covariates and either $S$ or $Z$.
+Figure \ref{fig:boxplots} shows violin plots, layered on top of jittered scatterplots, showing estimation error for GEEPERs, the mixture estimator, and the PSW estimator of $\eff1$ under various sets of conditions, and Table \ref{tab:coverage} shows empirical covarage of nominal 95\% confidence interval estimates under a somewhat wider set of circumstances when $\muc1=0.3$.
+The leftmost panel of Figure \ref{fig:boxplots} and the top line of Table \ref{tab:coverage} show results conditions favorable to both the mixture model and to GEEPERs---normal residuals and no interactions in the data generating model between covariates and either $S$ or $Z$.
 In this case, GEEPERs is roughly unbiased, but less precise than the other two methods---when $\alpha=0.2$ it is much noisier, and when $\alpha=0.5$ it is only slightly less precise.
@@
 This is somewhat surprising, because when $\muc1=0$, there is no separation between the principal strata in the control group, and previous literature \citep{griffin2008application} has identified this scenario as a particular challenge for mixture modeling.
-The PSW estimator is also biased---due, presumably, to the violation of Principal Ignorability, Assumption \ref{ass:PI}--inherent in the data generating model.
+The PSW estimator is also biased---due, presumably, to the violation of Principal Ignorability, Assumption \ref{ass:PI}---inherent in the data generating model.
 On the other hand, the PSW estimator has easily the lowest sampling variance of the three estimators.
-In fact, Table \ref{tab:rmse} shows that despite its bias, the PSW estimator has lower RMSE than GEEPERs when $\alpha=0$ or 0.2.
-The middle panel of Figure \ref{fig:boxplots} and the fifth line of Table \ref{tab:rmse} shows a similar set of results for when the residuals in \eqref{eq:y-sim} are uniformly distributed.
+The middle panel of Figure \ref{fig:boxplots} and the fifth line of Table \ref{tab:coverage} show a similar set of results for when the residuals in \eqref{eq:y-sim} are uniformly distributed.
 In this case, GEEPERs and PSW estimator behave in a similiar fashion to the normal case, but the mixture estimator shows a bimodal pattern--sometimes over-estimating $\eff1$ and sometimes underestimating, but rarely estimating $\eff1$ accurately.
 
+\sloppy
 The rightmost panel of Figure \ref{fig:boxplots} returns to the case of normal residuals---and only shows results when $\muc1=0.3$ and $\alpha=0.5$---but introduces interactions in the data generating model between covariates and either $\st$, $Z$, or both.
 That is, the coefficients on the covariates in the outcome model vary between principal strata and/or between treatment groups.
-Corresponding RMSE results, along with analogous results for uniform residuals and other values for $\alpha$, are in Table \ref{tab:rmse}.
 Presence of interactions could undermine the GEEPERs by inducing a relationship between principal scores and potential outcomes, even after adjusting the outcomes for covariate effects.
@@
 
-%% \begin{figure}
-%%   \centring
-%%   \includegraphics{../simFigs/normalOutcomesNoInteractionPosB1.jpg}
-%%   \caption{Violin and scatterplots of 500 simulation estimation errors for estimators described in Section~\ref{sec:simMods} when the distribution of residuals in \eqref{eq:yc-sim} was normal, there were no interactions between covariates and $S$ or $Z$, and $\alpha=0.2,0.5$.
-Annotations at the upper and lower edges of the plot indicate the number of extreme outliers excluded from the plotting area.}
-%%   \label{fig:simNorm}
-%% \end{figure}
-
-%% \begin{figure}
-%%   \centring
-%%   \includegraphics{../simFigs/unifOutcomesNoInteractionPosB1.jpg}
-%%   \caption{Violin and scatterplots of 500 simulation estimation errors for estimators described in Section~\ref{sec:simMods} when the distribution of residuals in \eqref{eq:yc-sim} was uniform, there were no interactions between covariates and $S$ or $Z$, and $\alpha=0.2,0.5$.
-Annotations at the upper and lower edges of the plot indicate the number of extreme outliers excluded from the plotting area.}
-%%   \label{fig:simUnif}
-%% \end{figure}
-
-%% \begin{figure}
-%%   \centring
-%%   \includegraphics{../simFigs/InteractionPosB1.jpg}
-%%   \caption{Violin and scatterplots of 500 simulation estimation errors for estimators described in Section~\ref{sec:simMods} when the distribution of residuals in \eqref{eq:yc-sim} was normal, there were possibly interactions between covariates and $S$ or $Z$, $\muc1=0.3$, and $\alpha=0.5$.
-Annotations at the upper and lower edges of the plot indicate the number of extreme outliers excluded from the plotting area.}ana
-%%   \label{fig:simUnif}
-%% \end{figure}
-
-%   \begin{table}
-%     \centering
-%   \input{rmseTab}
-%   \caption{Empirical root-mean-squared error (RMSE) estimates for GEEPERs, Bayesian mixture modeling, and PSW under varying conditions.}
-%   \label{tab:rmse}
-% \end{table}
-
 \begin{table}
-\caption{Empirical coverage of nominal 95\% confidence intervals for GEEPERs and Bayesian mixture modeling under varying conditions.}
-  \label{tab:coverage}
+
+\caption{\label{tab:coverage}Empirical coverage of nominal 95\% confidence intervals for GEEPERs and Bayesian mixture modeling under varying conditions.}
 \centering
-  \input{coverageTab}
+\begin{tabular}{*{9}{c}}%*{6}{r}}
+\toprule
+\multicolumn{3}{c}{ } & \multicolumn{2}{c}{$\alpha=0$} & \multicolumn{2}{c}{$\alpha=0.2$} & \multicolumn{2}{c}{$\alpha=0.5$} \\
+\cmidrule(l{3pt}r{3pt}){4-5} \cmidrule(l{3pt}r{3pt}){6-7} \cmidrule(l{3pt}r{3pt}){8-9}
+\makecell[l]{Residual\\Dist.} & \makecell[c]{X:Z\\Int.?} & \makecell[r]{X:S\\Int.?} & \rotatebox[origin=c]{300}{GEEPERs} & \rotatebox[origin=c]{300}{Mixture}  & \rotatebox[origin=c]{300}{GEEPERs} & \rotatebox[origin=c]{300}{Mixture}& \rotatebox[origin=c]{300}{GEEPERs} & \rotatebox[origin=c]{300}{Mixture}\\
+\midrule
+ & No & No & 0.99 & 0.99  & 0.97 & 0.97 & 0.96 & 0.96\\
+
+ & Yes & No & 0.88 & 0.74  & 0.88 & 0.74 & 0.94 & 0.89\\
+
+ & No & Yes & 1.00 & 1.00  & 0.98 & 0.97 & 0.97 & 0.96\\
+\multirow{-4}{*}{\raggedright\arraybackslash Normal} & Yes & Yes & 0.90 & 0.77  & 0.90 & 0.77 & 0.94 & 0.88\\
+\cmidrule{1-9}
+ & No & No & 0.99 & 0.29  & 0.97 & 0.40 & 0.95 & 0.61\\
+
+ & Yes & No & 0.83 & 0.11  & 0.87 & 0.20 & 0.93 & 0.48\\
+
+ & No & Yes & 1.00 & 0.47  & 0.98 & 0.53 & 0.96 & 0.57\\
+
+\multirow{-4}{*}{\raggedright\arraybackslash Uniform} & Yes & Yes & 0.83 & 0.15 & 0.88 & 0.21 & 0.94 & 0.53\\
+\bottomrule
+
+\end{tabular}
 \end{table}
@@
 
-\subsection{Proof for Proposition \ref{prop:interactions}}
+\subsection{A Stronger Version of Proposition \ref{prop:reg2} and a Proof}
+
+\begin{prop}\label{prop:interactions}
+  Say, for $i=1,\dots,n$, principal scores $\ppi$ are generated as \eqref{eq:pscore}, with parameters $\bm{\alpha}$ identified and estimable with M-estimation, and there exist $\beta_0$, $\beta_1$, $\beta_2$, $\beta_3$, $\bm{\gamma_1}$, $\bm{\gamma_2}$, $\bm{\gamma_3}$ amd $\bm{\gamma_4}$ such that
+  \begin{equation}\label{eq:interaction}
+    \begin{split}
+    \EE[Y_i|\st,Z,\bx]=&\beta_0+\beta_1\sti+\beta_2 Z_i+\beta_3Z_i\sti\\
+    &+\bm{\gamma_1}'\bxy_i+\bm{\gamma_2}'\bxy_i\sti+
+    \bm{\gamma_3}'\bxy Z_i+\bm{\gamma_4}'\bxy_i Z_i\sti
+    \end{split}
+  \end{equation}
+
+  Then, under Assumptions \ref{ass:sm}, \ref{ass:rand}, and \ref{ass:vps}, if $\ppi$ is linearly independent of $\bxy$, a researcher may follow the following procedure to estimate principal effects:
+  \begin{enumerate}
+  \item Estimate principal scores by fitting model \eqref{eq:pscore} to data from the treatment group
+  \item Replace $\sti$ with $\ri$ (as defined in \ref{eq:ri}) in model \eqref{eq:interaction} and fit with OLS
+  \item Estimate principal effects as:
+   \begin{equation}\label{eq:prinEffEst}
+  \begin{split}
+    \heff0_{int}&\equiv \hat{\beta}_2+\bm{\hat{\gamma}_3}'\overline{\bxy}_{Z=1,S=0}\\
+    \heff1_{int}&\equiv \hat{\beta}_2+\hat{\beta}_3+(\bm{\hat{\gamma}_3}+\bm{\hat{\gamma}_4})'\overline{\bxy}_{Z=1,S=1}
+  \end{split}
+   \end{equation}
+   where $\overline{\bxy}_{Z=1,S=0}$ and $\overline{\bxy}_{Z=1,S=0}$ are the vector of covariate sample means for the subsets of subjects with $Z=1$ and $S=0$ or $S=1$, respectively.
+  \end{enumerate}
+  Then $\heff0_{G}$ and $\heff1_{int}$ are M-estimators, with $\heff0_{int}\rightarrow_p\eff0$ and $\heff1_{G}\rightarrow_p\eff1$ as $n\rightarrow\infty$.
+  Under suitible regularity conditions, they are jointly asymptotically normal, with a variance of the form \eqref{eq:sandwich}.
+\end{prop}
+
+Equation \eqref{eq:interaction} implies Assumption \ref{ass:rci} with $\bm{\gamma_2}=\bm{\gamma_3}=\bm{\gamma_4}=0$.
+
+\begin{proof}
 First of all, by \eqref{eq:interaction}, 
@@
 Under standard OLS regularity conditions, the estimated coefficients of that model are consistent, completing the proof.  
+\end{proof}
 

### 2022-12-13 writeUps/paper.tex

@@
 
-In each run of the simulation, we simulated three indpendent standard normal covariates, $x_k$ $k=1,\dots,3$.
+In each run of the simulation, we simulated three independent standard normal covariates, $x_k$ $k=1,\dots,3$.
 However, only the first two covariates were ``observed,'' i.e.
@@
 Interactions between covariates and $Z$ were controlled by $\gamma_3$: when there were interactions, $\gamma_3=1/(2sqrt{6})$, otherwise $\gamma_3=0$.
+Either type of interaction violates Assumption \ref{ass:rci} since if $\gamma_2\ne 0$ or $\gamma_3 \ne 0$, there is no single vector $\bm{\gamma}$ that can fully capture the dependence of $\yc$ and $\yt$ on $S$---instead, the dependence varies with $\st$ and/or with $Z$.
 
@@
   \centering
-  \includegraphics[width=\textwidth]{../simFigs/boxplots.pdf}
+  \includegraphics[width=\textwidth,clip]{../simFigs/boxplots.pdf}
   \caption{Violin and scatterplots of 500 simulation estimation errors for estimators described in Section~\ref{sec:simMods} under varying conditions.
 For all plots, $n=500$; there were 500 replications for each set of conditions.
-Annotations at the upper and lower edges of the plot indicate the number of extreme outliers excluded from the plotting area.}
+Annotations indicate the number of extreme outliers excluded from the plotting area.}
   \label{fig:boxplots}
@@
 The middle panel of Figure \ref{fig:boxplots} and the fifth line of Table \ref{tab:coverage} show a similar set of results for when the residuals in \eqref{eq:y-sim} are uniformly distributed.
-In this case, GEEPERs and PSW estimator behave in a similiar fashion to the normal case, but the mixture estimator shows a bimodal pattern--sometimes over-estimating $\eff1$ and sometimes underestimating, but rarely estimating $\eff1$ accurately.
+In this case, GEEPERs and PSW estimator behave in a similar fashion to the normal case, but the mixture estimator shows a bimodal pattern--sometimes over-estimating $\eff1$ and sometimes underestimating, but rarely estimating $\eff1$ accurately.
+Credible intervals from the mixture model exhibit severe undercoverage, while GEEPERs confidence intervals achieve at least their nominal level.
 
-\sloppy
+%\sloppy
 The rightmost panel of Figure \ref{fig:boxplots} returns to the case of normal residuals---and only shows results when $\muc1=0.3$ and $\alpha=0.5$---but introduces interactions in the data generating model between covariates and either $\st$, $Z$, or both.
-That is, the coefficients on the covariates in the outcome model vary between principal strata and/or between treatment groups.
-Presence of interactions could undermine the GEEPERs by inducing a relationship between principal scores and potential outcomes, even after adjusting the outcomes for covariate effects.
-Fortunately, the presence of interactions in the data generating model in this scenario appear to make little difference, though RMSEs were slightly higher when interactions between $Z$ and covariates are present.
-Table \ref{tab:rmse} shows that when $\alpha=.2$, GEEPERs RMSE can be substantially higher when $Z$ interactions are present than when they are not; plots in the online appendix suggest that this is due to an increase in sampling variance and kurtosis rather than bias.
-
-Table \ref{tab:coverage} shows estimates of the empirical coverage of nominal 95\% confidence intervals for GEEPERs and the mixture estimator of $\eff1$, to evaluate their respective standard error estimates and the associated statistical inference.
-When $\alpha=0.5$, GEEPERs confidence intervals consistently achieve their nominal levels\footnote{With 500 replications and true coverage probabilities of 0.95, the standard error for empirical coverage rates is approximately 1 percentage point.}, whereas confidence intervals from mixture estimators undercover in the presence of interactions between $Z$ and $X$, or when residuals are non-normal.
-When $\alpha=0$ or 0.2, GEEPERs confidence intervals undercover in the presence of interactions between $X$ and $Z$.
-Across all conditions in the table, the GEEPERs confidence intervals have higher empirical coverage than their mixture-model counterparts.
+That is, the coefficients on the covariates in the outcome model vary between principal strata and/or between treatment groups, violating Assumption \ref{ass:rci}..
+%Presence of interactions could undermine the GEEPERs by inducing a relationship between principal scores and potential outcomes, even after adjusting the outcomes for covariate effects.
+Fortunately, the presence of interactions in the data generating model in this scenario appear to make little difference when $\alpha=0.5$; however, Table \ref{tab:coverage} shows that when $\alpha=0$ or 0.2, and there are interactions between $\bx$ and $Z$ (i.e.
+$\gamma_3\ne 0$) GEEPERs 95\% confidence intervals tend to under-cover.
+Nevertheless, coverage of GEEPERs intervals is nearly always substantially higher than coverage of corresponding 95\% credible intervals from mixture models.%\footnote{}
 
@@
 \bottomrule
+\multicolumn{9}{p{0.9\textwidth}}{\footnotesize Standard error $\approx 1$ percentage point, based on 500 replications and true coverage probabilities of 0.95.}
 
@@
 In all four conditions, students worked on their assigned math program in class on 12 different occasions---a prior assessment, a mid-test, a post-test, and nine learning sessions---spaced throughout the school year.
-The four conditions ranged from a game---DragonBox---to the active control condition in which students simply used the computer to work on a sequence of math problems. 
-In the DragonBox game, students learned to solve algebraic equations by isolating a box containing a dragon; as students progressed in the game, pictures of monsters were replaced by mathematical symbols.
-Another condition, the From Here to There (FH2T) program, used visual cues such as color and spacing to teach algebraic concepts, and allowed students to solve equations by manipulating expressions via dragging and dropping.
-In the ASSISTments condition, students used computers to work through a series of math problems taken from the district's textbook.
-While working, students had access to a series hints for each problem, culminating in a ``bottom out'' hint that contained the answer.
-Students in the ASSISTments condition were given immediate feedback on their answers and had to enter the correct answer in order to proceed to the next problem.
-In the active control condition, labeled ``business as usual'' (BAU), the computer work mimicked traditional pencil-and-paper mathematics coursework---students worked through a series of math problems without access to hints and received feedback only after completing the assignment.
+The four conditions included two gamified programs---DragonBox and From Here to There (FH2T)---and two programs in which students used computers to work through a series of algebra problems taken from the district's textbook---``Assistments'' and an active control called ``Business as Usual'' (BAU).
+% the active control condition in which students simply used the computer to work on a sequence of math problems.
+%In the DragonBox game, students learned to solve algebraic equations by isolating a box containing a dragon; as students progressed in the game, pictures of monsters were replaced by mathematical symbols.
+%Another condition, the From Here to There (FH2T) program, used visual cues such as color and spacing to teach algebraic concepts, and allowed students to solve equations by manipulating expressions via dragging and dropping.
+%In the ASSISTments condition, students used computers to work through a series of math problems taken from the district's textbook. 
+Students in the ASSISTments condition had access to a series hints for each problem, culminating in a ``bottom out'' hint that contained the answer and were given immediate feedback on their answers.
+% and had to enter the correct answer in order to proceed to the next problem.
+In the BAU condition students had no access to hints and received feedback only after completing the assignment. 
 In this illustration, we will compare the ASSISTments condition (which we arbitrarily label ``treatment'')---the only condition featuring bottom-out hints---with each of the three others separately.
@@
 However, in this experiment, the ability to bottom out was only one of several differences between the treatment conditions.
-One way to disentangle the role of bottoming out in any treatment effect would be to estimate average treatment effects separately for subjects who would (if assigned to treatment) bottom out, and for subjects who would not.
-If hints (that do not supply the correct answer) and error messages are helpful, but the ability to bottom out is harmful to learning, then we might expect to estimate a positive effect of being assigned to the feedback condition for students who do not bottom out, but a negative effect for students who do.
-If hints are helpful for learning only if they can be read as worked examples, then we might expect no effect of condition on students who do not bottom out, but a positive effect for students who do.
+One way to disentangle the role of bottoming out in any treatment effect would be to estimate average treatment effects of being assigned to the ASSISTments condition, relative to any of the other three, separately for subjects who would (if assigned to treatment) bottom out frequently, and for subjects who would not.
+If the availability of bottom-out hints plays an important positive role in ASSISTments's efficacy, we might expect larger effects of condition on students who would bottom out frequently than for those who would not; if bottom-out hints are harmful, we might expect the opposite.
+%If hints (that do not supply the correct answer) and error messages are helpful, but the ability to bottom out is harmful to learning, then we might expect to estimate a positive effect of being assigned to the feedback condition for students who do not bottom out, but a negative effect for students who do.
+If hints are helpful for learning only if they can be read as worked examples, then we might expect no effect of condition on students who generally do not bottom out, but a positive effect for students who do.
 (These hypothetical results may be masked by other treatment effect heterogeneity between students who would or wouldn't bottom out, but the results could still be informative.)
 
+\sloppy
 For the purpose of demonstrating GEEPERs, we gathered dichotomized data on %which students in the treatment condition requested at least one bottom-out hint.
@@
 
-We also had access to data on a number of baseline covariates: total scores on a 10-item pretest, fifth-grade standardized test scores, race/ethnicity (white, Asian, Hispanic/Latino, or Other), indicator variables for if students have an early intervention program (EIP) or individualized educational plan (IEP) in place, if students are learning English as a second language (ESOL), students' average time on task for the pretest (logged), total scores from baseline evaluations of students' math anxiety, negative reaction to math, and numerical confidence, and an indicator for weather students began the school year in remote or in-person instruction.
+We also had access to data on a number of baseline covariates, including prior achievement, demographics, baseline measurements of student attitudes towards math and an indicator for whether students began the school year in remote or in-person instruction.
 Students with missing pretest scores were dropped from the analysis; missing data in other covariates was imputed with a Random Forest algorithm \citep{missForest}.
-%Table \ref{tab:tab1} describes each covariate--as well as the sucess of the randomForest imputation algorithm.
-%% \begin{comment}
-%% \begin{table}
-%%   \centering
-%%   \input{table1}
-%%   \caption{Descriptive statistics on baseline covariates, $S$, and $Y$ for the analysis data.
-Values are mean (SD) unless otherwise noted.
-Imputation error is the ``out of box'' error estimate from the Random Forest imputation algorithm---root mean squared error divided by the observed standard deviation for continuous variables and misclassification proportion for categorical variables.}
-%%   \label{tab:tab1}
-%% \end{table}
-%% \end{comment}
-%% u
 
@@
 
-\subsection{Principal Score Model}\label{sec:psMod}
+Due to privacy regulations, the data are not publicly available; however, interested analysts may access anonymized data by emailing the study's principal investigator at [REDACTED] and signing a privacy agreement.
+Replication code for the analysis is available at \url{www.github.com/[Redacted]}.
 
-\begin{figure}
-  \centering
-  \includegraphics{../figure/psModCoef.jpg}
-  \caption{Principal Score Model Coefficients}
-  \label{fig:psMod}
-\end{figure}
+\subsection{Data Analysis}
+\subsubsection{Principal Score Model}\label{sec:psMod}
 
-We considered several models for the principal score, and full results (which are largely similar to each other) can be found in the online appendix.
-Here we present results from our model of choice, a logistic regression of observed $S$ on baseline variables chosen with a combination of expert judgement and backwards selection based on the AIC.
-These included school fixed effects and a set of student-level demographic and educational data: fifth-grade mathematics standardized test scores, %Gender, Race/Ethnicity,
-indicator variables for special education status (IEP), whether the student was taking English as a second language (ESOL), whether the student was classified as gifted, and whether students attended a full year of school in 5th grade.
-Finally, we also included data collected from students at the beginning of the study: the natural logarithm of the  time spent on task during the pretest, scores on baseline measurements of mathematics anxiety and self-efficacy and perceptual sensitivity, along with a missingness indicator for these measurements.
-Binned residual plots \citep{gelman2006data} from an initial model fit indicated a curvilinear relationship between perceptual sensitivity measurements (modeled as linear) and model errors, so in the final model we included perceptual sensitivity as a quadratic polynomial.
+% \begin{figure}
+%   \centering
+%   \includegraphics{../figure/psModCoef.jpg}
+%   \caption{Principal Score Model Coefficients}
+%   \label{fig:psMod}
+% \end{figure}
+
+We estimated principal scores using logistic regression of observed $S$ on schol fixed effects and a set of baseline variables chosen with a combination of expert judgement and backwards selection based on the AIC, and evaluated model fit with binned residual plots \citep{gelman2006data}.
 The model was fit to data from students in the treatment group, for whom bottom-out status ($S$) was observed; hence, the same model fit was used for all three treatment contrasts.
-
 The model was fairly successful in distinguishing bottom-outers from non-bottom-outers---the AUC, evaluated using out-of-sample predictions in a 10-fold cross-validation, was roughly 0.795.
 That is, in roughly 80\% of bottom-outer/non-bottom-outer pairs of subjects, the bottom-outer will have a higher predicted probability from the model.
-Table \ref{fig:psMod} show the model's coefficient estimates.
-Notably, conditional on the other covariates, students with higher fifth-grade standardized test scores were less likely to request many bottom out hints, as were students who spent little time on task during the pretest, while IEP and ESOL students, and students without baseline measurement data, were relatively unlikely to be bottom-outers.
-The relationship between perceptual sensitivity and bottom out requests was not monotonic---students with average levels were more likely than those with more extreme high or low levels to be bottom-outers; bottom-outers tended to have lower math self-efficacy, but the data are also consistent with a small positive relationship, and there appears to be little to no relationship between math anxiety and bottom-outer status at all.
 
-\subsection{Outcome Regression and Principal Effects}\label{sec:outMod}
-\begin{figure}
-  \centering
-  \includegraphics{../figure/prinEffs.pdf}
-  \caption{Principal Effects}
-  \label{fig:effects}
-\end{figure}
+% \subsubsection{Outcome Regression and Principal Effects}\label{sec:outMod}
+% \begin{figure}
+%   \centering
+%   \includegraphics{../figure/prinEffs.pdf}
+%   \caption{Principal Effects}
+%   \label{fig:effects}
+% \end{figure}
 
-Figure \ref{fig:effects} plots the estimated principal effects for bottom-outers and non-bottom-outers, using two different models---alongside estimates of the overall average treatment effect, labeled ``ATE.'' In both models, effects were adjusted for all of the variables that were in the principal score model, in addition to students' pretest scores; also, the models substituted class fixed effects for school fixed effects.\footnote{Why use a coarser set of fixed effects for the principal score model than the outcome regression?
+% Figure \ref{fig:effects} plots the estimated principal effects for bottom-outers and non-bottom-outers, using two different models---alongside estimates of the overall average treatment effect, labeled ``ATE.'' In both models, effects were adjusted for all of the variables that were in the principal score model, in addition to students' pretest scores; also, the models substituted class fixed effects for school fixed effects.
+%\footnote{Why use a coarser set of fixed effects for the principal score model than the outcome regression?
 There were two reasons: first, the principal score model was fit to a smaller sample (only students in the ``Instant'' condition), so some classrooms only included one student, which can lead to poor fit.
 Second, the asymptotics of logistic regression are more restricted than OLS, and so logistic regression tends to perform poorly when the number of parameters increases with the sample size \citep{agresti}.}
-The two sets of estimated differed in that the first assumed that coefficients on covariates were equal across principal strata, whereas the second set of results allowed the coefficients on pretest and 5th-grade standardized test scores---the most important predictors of outcomes and principal strata membership, respectively---to vary between strata.
+% The two sets of estimated differed in that the first assumed that coefficients on covariates were equal across principal strata, whereas the second set of results allowed the coefficients on pretest and 5th-grade standardized test scores---the most important predictors of outcomes and principal strata membership, respectively---to vary between strata.
 
-The results suggest little difference, if any, between the effects in the two principal strata.
-Estimated effects compared to BAU or DragonBox were slightly lower for bottom-outers than for non-bottom-outers, and estimated effects compared to FH2T were higher for non-bottom-outers.
-However, none of these differences is statistically significant, so opposite patterns, or no difference between principal effects at all, would also be consistent with the data.
-In fact, 95\% confidence intervals for all principal effects include both positive and negative principal effects.
-There is no discernible difference between estimated principal effects when coefficients on pretest and grade-5 standardized tests were or were not allowed to vary between strata, though standard errors from the models including an interaction were somewhat higher.
+% The results suggest little difference, if any, between the effects in the two principal strata.
+% Estimated effects compared to BAU or DragonBox were slightly lower for bottom-outers than for non-bottom-outers, and estimated effects compared to FH2T were higher for non-bottom-outers.
+% However, none of these differences is statistically significant, so opposite patterns, or no difference between principal effects at all, would also be consistent with the data.
+% In fact, 95\% confidence intervals for all principal effects include both positive and negative principal effects.
+% There is no discernible difference between estimated principal effects when coefficients on pretest and grade-5 standardized tests were or were not allowed to vary between strata, though standard errors from the models including an interaction were somewhat higher.
 
-\subsection{Mixture Model and Principal Score Weighting}
 \begin{figure}
@@
 
-Figure \ref{fig:compare} shows principal effect estimates from GEEPERs (without the $S$-interactions) compared to analogous estimates using Bayesian mixture modeling and PSW.
+Figure \ref{fig:compare} shows principal effect estimates from GEEPERs compared to analogous estimates using Bayesian mixture modeling and PSW, alongside estimates of the overall average treatment effects, labeled ``ATE.''.
 All three sets of estimates relied on the same principal score model described above, in Section \ref{sec:psMod}, though in the mixture model it was fit simultaneously with effect estimation using a Markov Chain Monte Carlo algorithm \citep{rstan}.
+Outcome regressions in GEEPERs and the mixture model included all of the covariates that were in the principal score model, in addition to students' pretest scores; also, the models substituted class fixed effects for school fixed effects.
+
 The Bayesian mixture model assumed that outcomes were normally distributed, conditional on covariates (the same as in Section \ref{sec:outMod}) and principal stratum, with a residual standard deviation that varied between principal strata.
@@
 All model parameters with support in $\mathbb{R}$ were given standard normal priors, while standard deviation parameters were given half-standard normal priors.
+
 The PSW estimators did not include any covariate outcome modeling. Standard errors for the PSW estimates were estimated using the bootstrap, with resampling done within schools.
-For all three sets of estimators, approximate 95\% confidence intervals were estimated by subtracting twice the standard error from the point estimate (or twice the posterior standard deviation from the posterior mean, in the Bayesian mixture model).
 
-In this case, all three estimation strategies largely agreed.
+For all three sets of estimators, approximate 95\% confidence intervals were estimated by adding and subtracting twice the standard error from the point estimate (or twice the posterior standard deviation from the posterior mean, in the Bayesian mixture model).
+
+The GEEPERs results suggest little difference, if any, between the effects in the two principal strata.
+Estimated effects compared to BAU or DragonBox were slightly lower for bottom-outers than for non-bottom-outers, and estimated effects compared to FH2T were higher for non-bottom-outers.
+However, none of these differences is statistically significant, so opposite patterns, or no difference between principal effects at all, would also be consistent with the data.
+In fact, 95\% confidence intervals for all principal effects include both positive and negative principal effects.
+
+The other estimation strategies largely agreed with GEEPERs, and with each other. 
 This may suggest that the estimators' various assumptions---additivity of treatment assignment effect for GEEPERs and the mixture model, normal residuals for the mixture model, and principal ignorability for PSW---were all innocuous and may have held approximately.

### 2022-12-19 code/simulation/displaySim.r
simulation appendix

@@ -88,7 +88,8 @@ coverageN <- pdns%>%
    summarize(coverage=mean(CInormL<=pop & CInormU>=pop,na.rm=TRUE))%>%ungroup()
 
 
-bind_rows(
+
+tab1 <- bind_rows(
   biasN%>%filter(eff==1)%>%select(-eff)%>%
   pivot_wider(names_from=n,names_prefix="n=",values_from=bias)%>%mutate(measure="Bias"),
   seN%>%filter(eff==1)%>%select(-eff)%>%
@@ -97,7 +98,10 @@ bind_rows(
   pivot_wider(names_from=n,names_prefix="n=",values_from=rmse)%>%mutate(measure="RMSE"),
   coverageN%>%filter(eff==1)%>%select(-eff)%>%
   pivot_wider(names_from=n,names_prefix="n=",values_from=coverage)%>%mutate(measure="95% CI Coverage"))%>%
-  select(measure,everything())%>%
+  select(measure,everything())
+  
+  
+  #%>%
   kbl()%>%
   collapse_rows(columns=1,latex_hline="major",valign="middle")
 
@@ -111,7 +115,7 @@ ggsave('simFigs/byN.jpg',width=6,height=3)
 ### results across B1
 ###############################################################################
 
- pdb1 <- resultsB1s %>%
+ pdb1 <- reausultsB1s %>%
   group_by(b1)%>%
   mutate(AUCf=mean(auc,na.rm=TRUE))%>%
   ungroup()%>%
@@ -477,6 +481,48 @@ select(`GEEPERs`,`Mix.`=Mixture,PSW))%>%
   collapse_rows(columns=1,latex_hline="major",valign="middle")
 sink()
 
+#######################################################
+#### rmse appendix
+#######################################################
+
+rmse <- pd%>%
+  filter(rhat<1.1|estimator=='PSW')%>%
+  group_by(n,mu01,errDist,b1,intS,intZ,interactionS,interactionZ,eff,estimator)%>%
+   summarize(rmse=sqrt(mean(errP^2,na.rm=TRUE)))%>%ungroup()
+
+rmse%>%filter(n==500,eff==1,errDist!='mix',mu01==0.3,b1>0)%>%
+  ggplot(aes(b1,rmse,color=estimator,fill=estimator,group=estimator))+
[…]

### 2022-12-19 writeUps/paper.tex

@@
 \usepackage{numdef}
+\usepackage{multibib}
 
 \input{notation.tex}
+
+\newcites{supp}{Supplementary References}
 
@@
 Another route to identification and estimation of principal effects leads through auxiliary data, including baseline covariates or secondary outcomes \citep{mattei2013exploiting}.
-Of particular importance is the ``principal score,'' \citep{jo,ding2017principal,feller2017principal}, or the probability that a subject belongs to a particular principal stratum, as a function of pre-treatment covariates.
-Under a strong, untestable assumption called ``principal ignorability,'' which is akin to the no-confounding assumption underlying observational studies, principal scores can be transformed into weights, which in turn form an unbiased estimate of principal effects.
+Of particular importance is the ``principal score,'' \citep{jo,dingLu,feller2017principal}, or the probability that a subject belongs to a particular principal stratum, as a function of pre-treatment covariates.
+Under a strong, untestable assumption called ``principal ignorability,''  principal scores can be transformed into weights to compute form an unbiased estimate of principal effects.
 
@@
 The following section gives a formal introduction to principal stratification and briefly describes two common estimation techniques, parametric mixture modeling and principal score weighting.
-Next, section \ref{sec:geepers} develops GEEPERs estimation, and gives conditions for its consistency.
+Next, Section \ref{sec:geepers} develops GEEPERs estimation, and gives conditions for its consistency.
 Section \ref{sec:simulation} describes a simulation study examining the operating characteristics of GEEPERs estimates and comparing them to estimates from parametric mixture modeling and principal score weighting.
@@
 Then $S$ itself has potential values $\sti$ and $\sci$, which $i$ would exibit if $Z_i=1$ or $Z_i=0$, respectively.
-Assume ``strong monotonicity'' \citep[c.f.][]{dingLu} (also referred to as ``one-way noncompliance'' \citealt{wing2017can}):
+Assume ``strong monotonicity'' \citep[c.f.][]{dingLu} (also referred to as ``one-way noncompliance''):%\citealt{wing2017can}):
 \begin{ass}[Strong Monotonicity]\label{ass:sm}
@@
 \subsection{Principal Scores}
-The principal score \citep[e.g.][]{jo2009use} is defined as
+The principal score \citep[e.g.][]{jo} is defined as
 \begin{equation}\label{eq:pscore}
@@
 
-Alternatively, an analyst may assume ``principal ignorability'' \citep{jo2009use,dingLu}:
+Alternatively, an analyst may assume ``principal ignorability'' \citep{jo,dingLu}:
 \begin{ass}[Principal Ignorability]\label{ass:PI}
@@
 That is, the principal stratum is unrelated to control potential outcomes conditional on covariates.
-Principal ignorability \eqref{eq:pi} is reminiscent of ignorability assumptions typical in observational studies \citep[e.g.][]{rosenbaum2010design}.
+Principal ignorability \eqref{eq:pi} is reminiscent of ignorability assumptions typical in observational studies. % \citep[e.g.][]{rosenbaum2010design}.
 % Both require a rich set of covariates and background knowledge of the determinants of $S$; 
@@
 
-Principal score weighting does not require any distributional assumptions and (after estimating principal scores) very easy to implement.
+Principal score weighting does not require any distributional assumptions and (after estimating principal scores) is very easy to implement.
 On the other hand, the principal ignorability assumption is strong and restrictive.
 
-\citet{feller2017principal} recommends the case-resampling bootstrap to estimate the sampling variance of $\hat{\muc0}_w$, $\hat{\muc1}_w$, and their associated principal effect estimators.
+\citet{feller2017principal} recommends the case-resampling bootstrap to estimate the sampling variances of $\hat{\mu}_{Cw}^0$, $\hat{\mu}_{Cw}^1$, and their associated principal effect estimators.
 
@@
 The study was designed to answer three sets of overarching questions.
-First, how does the performance of the GEEPERs estimator vary under different conditions, including sample size, the extent to which $\bx$ predicts $\st$, the principal effects themselves vary, and when important interactions were omitted from the outcome model.
+First, how does the performance of the GEEPERs estimator vary under different conditions, including sample size, the extent to which $\bx$ predicts $\st$, whether and how the principal effects themselves vary, and when important interactions are omitted from the outcome model.
 Second, how does GEEPERs compare with mixture modeling?
@@
 where $\alpha$ is a manipulated factor.
-When $\alpha$ was higher, $\st$ was more easily predicted by covariates; when $\alpha=0$, Assumption \ref{ass:vps}, was violated, i.e. $Var(\pp)=0$.
-Then principal stratum $\sti\in \{0,1\}$ was simulated as $S_i\sim Bern(\ppi)$
+When $\alpha$ was higher, $\st$ was more easily predicted by covariates; when $\alpha=0$, Assumption \ref{ass:vps} was violated, i.e. $Var(\pp)=0$.
+Principal stratum $\sti\in \{0,1\}$ was simulated as $S_i\sim Bern(\ppi)$
 
-Finally, potential and observed outcomes were generated as in \eqref{eq:interaction}, with $\bxy=[\bm{x_1},\bm{x_2},\bm{x_3}]$,
-and with $\EE[\epsilon_i]=0$ and $Var(\epsilon_i)=1/2$.
-Specifically,
+Potential and observed outcomes were generated as 
 \begin{equation}\label{eq:y-sim}
@@
 % \end{equation}
-with $\beta_3=0.3-\beta_1$.
-Hence, $\beta_0=\beta_2=\gamma_4=0$, when $\gamma_3=0$ (i.e.
-no interaction between $Z$ and covariates), there is no average treatment effect in the $\st=0$ stratum, the coefficient for the unobserved covariate $x_3$ is constant across treatment groups and principal strata, and we do not consider three-way interactions between $x_1$ or $x_2$, $\st$, and $Z$.
+with $\beta_3=0.3-\beta_1$, $\EE[\epsilon_i]=0$, and $Var(\epsilon_i)=1/2$.
+When $\gamma_3=0$ (i.e.
+no interaction between $Z$ and covariates), there is no average treatment effect in the $\st=0$ stratum.
 
@@
 
-\textbf{\emph{Sample Size per condition $\bm{n}$:}} The guarantees of M-estimation are all asymptotic, so we examined the behavior of our M-estimator at a range of sample sizes to determine its finite-sample properties.
+\textbf{\emph{Sample Size per condition $\bm{n}$:}} The guarantees of M-estimation are all asymptotic, so we examined the behavior of our GEEPERs at a range of sample sizes to determine its finite-sample properties.
 
@@
 
-Following \citet[][p. 373]{carroll2006measurement}, we %separate the estimating equations $\Psi$ into $\phi$ and $\psi$, estimating equations for models \eqref{
+Following \citetsupp[][p. 373]{carroll2006measurement}, we %separate the estimating equations $\Psi$ into $\phi$ and $\psi$, estimating equations for models \eqref{
 can decompose the matrices into diagonal elements 
@@
 
+\bibliographystylesupp{plainnat}
+\bibliographysupp{MOM}
+
 \end{document}

### 2022-12-19 code/simulation/displaySim.r
appendix coverage table

@@ -562,26 +562,36 @@ sink()
 
 sink('writeUps/coverageTabAppendix.tex')
 cbind(
-coverage%>%filter(errDist!='mix',mu01==0.3,eff!="Diff",b1==.2,estimator!='PSW')%>%
+  coverage%>%filter(errDist!='mix',eff!="Diff",b1==0,estimator!='PSW',n==500)%>%
   transmute(`Residual\nDist.`=c(norm='Normal',unif='Uniform')[errDist],
          `X:Z\nInt.?`=ifelse(intZ,'Yes','No'),
          `X:S\nInt.?`=ifelse(intS,'Yes','No'),
          estimator=ifelse(estimator=='Mixture','Mix.',estimator),
-         n,
-         `\\\\beta_1`=mu01,
-         `Prin. Eff`=paste0('$\\\\tau^',eff,'$'),
+         `$\\beta_1$`=mu01,
+         `Prin.\nEff`=paste0('$\\tau^',eff,'$'),
          coverage)%>%
 pivot_wider(names_from=estimator,values_from=coverage),
-coverage%>%filter(errDist!='mix',mu01==0.3,eff!="Diff",b1==.5,estimator!='PSW')%>%
+coverage%>%filter(errDist!='mix',eff!="Diff",b1==.2,estimator!='PSW',n==500)%>%
+  transmute(`Residual\nDist.`=c(norm='Normal',unif='Uniform')[errDist],
+         `X:Z\nInt.?`=ifelse(intZ,'Yes','No'),
+         `X:S\nInt.?`=ifelse(intS,'Yes','No'),
+         estimator=ifelse(estimator=='Mixture','Mix.',estimator),
+         `$\\beta_1$`=mu01,
+         `Prin.\nEff`=paste0('$\\tau^',eff,'$'),
+         coverage)%>%
+pivot_wider(names_from=estimator,values_from=coverage)%>%
+select(`GEEPERs`,`Mix.`),
+coverage%>%filter(errDist!='mix',n==500,eff!="Diff",b1==.5,estimator!='PSW')%>%
   transmute(`Residual\nDist.`=c(norm='Normal',unif='Uniform')[errDist],
          `X:Z\nInt.?`=ifelse(intZ,'Yes','No'),
          `X:S\nInt.?`=ifelse(intS,'Yes','No'),
          estimator=ifelse(estimator=='Mixture','Mix.',estimator),
          n,
[…]

### 2022-12-19 writeUps/paper.tex

@@
 
+\clearpage
+\section{Simulation Results}
+
+\begin{figure}[!h]
+    \centering
+    \includegraphics[width=0.8\textwidth]{simFigs/alphaAUC.pdf}
+    \caption{Estimated AUC values from principal score models for each value of $\alpha$ in the simulation, with $n=500$.}
+    \label{fig:alphaAUC}
+\end{figure}
+
+%\begin{table}
+%\caption{Coverage of nominal 95\% confidence intervals}
+
+\begin{tabular}[t]{lllrrlrrlllrrlrr}
+\toprule
+\multicolumn{3}{c}{ } & \multicolumn{3}{c}{$\alpha=0$} & \multicolumn{3}{c}{$\alpha=0.2$} & \multicolumn{3}{c}{$\alpha=0.5$} \\
+\cmidrule(l{3pt}r{3pt}){4-6} \cmidrule(l{3pt}r{3pt}){7-9} \cmidrule(l{3pt}r{3pt}){10-12}
+\makecell[l]{Residual\\Dist.} & \makecell[c]{X:Z\\Int.?} & \makecell[r]{X:S\\Int.?} & n & $\beta_1$ & Prin.
+Eff & GEEPERs & Mix.
+& \makecell[r]{Residual\\Dist.} & \makecell[l]{X:Z\\Int.?} & \makecell[c]{X:S\\Int.?} & n & $\beta_1$ & Prin.
+Eff & GEEPERs & Mix.\\
+\midrule
+ & No & No & 500 & 0.3 & $\tau^0$ & 0.97 & 0.98 & Normal & No & No & 500 & 0.3 & $\tau^0$ & 0.95 & 0.96\\
+
+ & No & No & 500 & 0.3 & $\tau^1$ & 0.97 & 0.97 & Normal & No & No & 500 & 0.3 & $\tau^1$ & 0.96 & 0.96\\
+
+ & Yes & No & 500 & 0.3 & $\tau^0$ & 0.88 & 0.76 & Normal & Yes & No & 500 & 0.3 & $\tau^0$ & 0.94 & 0.89\\
+
+ & Yes & No & 500 & 0.3 & $\tau^1$ & 0.88 & 0.74 & Normal & Yes & No & 500 & 0.3 & $\tau^1$ & 0.94 & 0.89\\
+
+ & No & Yes & 500 & 0.3 & $\tau^0$ & 0.97 & 0.96 & Normal & No & Yes & 500 & 0.3 & $\tau^0$ & 0.96 & 0.97\\
+
+ & No & Yes & 500 & 0.3 & $\tau^1$ & 0.98 & 0.97 & Normal & No & Yes & 500 & 0.3 & $\tau^1$ & 0.97 & 0.96\\
+
+ & Yes & Yes & 500 & 0.3 & $\tau^0$ & 0.90 & 0.76 & Normal & Yes & Yes & 500 & 0.3 & $\tau^0$ & 0.93 & 0.89\\
+
+\multirow{-8}{*}{\raggedright\arraybackslash Normal} & Yes & Yes & 500 & 0.3 & $\tau^1$ & 0.90 & 0.77 & Normal & Yes & Yes & 500 & 0.3 & $\tau^1$ & 0.94 & 0.88\\
+\cmidrule{1-16}
+ & No & No & 500 & 0.3 & $\tau^0$ & 0.98 & 0.39 & Uniform & No & No & 500 & 0.3 & $\tau^0$ & 0.95 & 0.61\\
+
+ & No & No & 500 & 0.3 & $\tau^1$ & 0.97 & 0.40 & Uniform & No & No & 500 & 0.3 & $\tau^1$ & 0.95 & 0.61\\
+
+ & Yes & No & 500 & 0.3 & $\tau^0$ & 0.87 & 0.21 & Uniform & Yes & No & 500 & 0.3 & $\tau^0$ & 0.94 & 0.48\\
+
+ & Yes & No & 500 & 0.3 & $\tau^1$ & 0.87 & 0.20 & Uniform & Yes & No & 500 & 0.3 & $\tau^1$ & 0.93 & 0.48\\
+
+ & No & Yes & 500 & 0.3 & $\tau^0$ & 0.97 & 0.49 & Uniform & No & Yes & 500 & 0.3 & $\tau^0$ & 0.95 & 0.60\\
+
+ & No & Yes & 500 & 0.3 & $\tau^1$ & 0.98 & 0.53 & Uniform & No & Yes & 500 & 0.3 & $\tau^1$ & 0.96 & 0.57\\
+
+ & Yes & Yes & 500 & 0.3 & $\tau^0$ & 0.89 & 0.21 & Uniform & Yes & Yes & 500 & 0.3 & $\tau^0$ & 0.95 & 0.52\\
+
+\multirow{-8}{*}{\raggedright\arraybackslash Uniform} & Yes & Yes & 500 & 0.3 & $\tau^1$ & 0.88 & 0.21 & Uniform & Yes & Yes & 500 & 0.3 & $\tau^1$ & 0.94 & 0.53\\
+\cmidrule{1-16}
+ & No & No & 1000 & 0.3 & $\tau^0$ & 0.96 & 0.97 & Normal & No & No & 1000 & 0.3 & $\tau^0$ & 0.94 & 0.94\\
+
+ & No & No & 1000 & 0.3 & $\tau^1$ & 0.97 & 0.97 & Normal & No & No & 1000 & 0.3 & $\tau^1$ & 0.95 & 0.96\\
+
+ & Yes & No & 1000 & 0.3 & $\tau^0$ & 0.89 & 0.60 & Normal & Yes & No & 1000 & 0.3 & $\tau^0$ & 0.94 & 0.80\\
+
+ & Yes & No & 1000 & 0.3 & $\tau^1$ & 0.90 & 0.60 & Normal & Yes & No & 1000 & 0.3 & $\tau^1$ & 0.93 & 0.79\\
+
+ & No & Yes & 1000 & 0.3 & $\tau^0$ & 0.97 & 0.96 & Normal & No & Yes & 1000 & 0.3 & $\tau^0$ & 0.95 & 0.97\\
+
+ & No & Yes & 1000 & 0.3 & $\tau^1$ & 0.97 & 0.97 & Normal & No & Yes & 1000 & 0.3 & $\tau^1$ & 0.94 & 0.96\\
+
+ & Yes & Yes & 1000 & 0.3 & $\tau^0$ & 0.91 & 0.62 & Normal & Yes & Yes & 1000 & 0.3 & $\tau^0$ & 0.95 & 0.83\\
+
+\multirow{-8}{*}{\raggedright\arraybackslash Normal} & Yes & Yes & 1000 & 0.3 & $\tau^1$ & 0.91 & 0.62 & Normal & Yes & Yes & 1000 & 0.3 & $\tau^1$ & 0.91 & 0.81\\
+\cmidrule{1-16}
+ & No & No & 1000 & 0.3 & $\tau^0$ & 0.95 & 0.01 & Uniform & No & No & 1000 & 0.3 & $\tau^0$ & 0.96 & 0.31\\
+
+ & No & No & 1000 & 0.3 & $\tau^1$ & 0.95 & 0.01 & Uniform & No & No & 1000 & 0.3 & $\tau^1$ & 0.96 & 0.30\\
+
+ & Yes & No & 1000 & 0.3 & $\tau^0$ & 0.91 & 0.00 & Uniform & Yes & No & 1000 & 0.3 & $\tau^0$ & 0.91 & 0.17\\
+
+ & Yes & No & 1000 & 0.3 & $\tau^1$ & 0.90 & 0.00 & Uniform & Yes & No & 1000 & 0.3 & $\tau^1$ & 0.92 & 0.17\\
+
+ & No & Yes & 1000 & 0.3 & $\tau^0$ & 0.96 & 0.07 & Uniform & No & Yes & 1000 & 0.3 & $\tau^0$ & 0.97 & 0.41\\
+
+ & No & Yes & 1000 & 0.3 & $\tau^1$ & 0.95 & 0.07 & Uniform & No & Yes & 1000 & 0.3 & $\tau^1$ & 0.97 & 0.40\\
+
+ & Yes & Yes & 1000 & 0.3 & $\tau^0$ & 0.89 & 0.03 & Uniform & Yes & Yes & 1000 & 0.3 & $\tau^0$ & 0.95 & 0.22\\
+
+\multirow{-8}{*}{\raggedright\arraybackslash Uniform} & Yes & Yes & 1000 & 0.3 & $\tau^1$ & 0.87 & 0.03 & Uniform & Yes & Yes & 1000 & 0.3 & $\tau^1$ & 0.92 & 0.21\\
+\bottomrule
+\end{tabular}
+
+%\end{table}
+
 \bibliographystylesupp{plainnat}

### 2022-12-19 writeUps/outcomeRegAppendix.tex
application results appendix

  \providecommand{\huxb}[2]{\arrayrulecolor[RGB]{#1}\global\arrayrulewidth=#2pt}
  \providecommand{\huxvb}[2]{\color[RGB]{#1}\vrule width #2pt}
  \providecommand{\huxtpad}[1]{\rule{0pt}{#1}}
  \providecommand{\huxbpad}[1]{\rule[-#1]{0pt}{#1}}

\begin{table}[ht]
\begin{centerbox}
\begin{threeparttable}
 \setlength{\tabcolsep}{0pt}
\begin{tabular}{l l l l}

\hhline{>{\huxb{0, 0, 0}{0.8}}->{\huxb{0, 0, 0}{0.8}}->{\huxb{0, 0, 0}{0.8}}->{\huxb{0, 0, 0}{0.8}}-}
\arrayrulecolor{black}

\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}c!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\centering \hspace{6pt}  \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{c!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\centering \hspace{6pt} BAU \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{c!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\centering \hspace{6pt} FH2T \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{c!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\centering \hspace{6pt} DragonBox \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]

\hhline{>{\huxb{255, 255, 255}{0.4}}->{\huxb{0, 0, 0}{0.4}}->{\huxb{0, 0, 0}{0.4}}->{\huxb{0, 0, 0}{0.4}}-}
\arrayrulecolor{black}

\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} (Intercept) \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 1.732\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 2.834 *** \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 2.510 **\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]

\hhline{}
\arrayrulecolor{black}

\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt}  \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.916)\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.718)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.919)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]

\hhline{}
\arrayrulecolor{black}

\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} Z \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.294\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} -0.174\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} -0.197\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]

\hhline{}
\arrayrulecolor{black}

\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt}  \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.277)\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.223)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.268)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]

\hhline{}
\arrayrulecolor{black}

\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} Sp \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.031\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} -0.216\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.042\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]

\hhline{}
\arrayrulecolor{black}

\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt}  \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.522)\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.412)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.515)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]

\hhline{}
\arrayrulecolor{black}

\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} Scale.Score5 \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.207\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.453 *** \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.049\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]

\hhline{}
\arrayrulecolor{black}

\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt}  \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.126)\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.104)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.132)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]

\hhline{}
\arrayrulecolor{black}

\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} ESOL1 \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.091\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} -0.459\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} -0.493\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]

\hhline{}
\arrayrulecolor{black}

\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt}  \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.364)\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.268)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.351)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]

\hhline{}
\arrayrulecolor{black}

\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} IEP1 \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.148\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.234\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.096\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]

\hhline{}
\arrayrulecolor{black}

\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt}  \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.288)\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.255)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.277)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]

\hhline{}
\arrayrulecolor{black}

\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} pre.total\_time\_on\_tasks \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} -0.131\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.078\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} -0.003\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]

\hhline{}
\arrayrulecolor{black}

\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt}  \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.091)\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.073)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.096)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]

\hhline{}
\arrayrulecolor{black}

\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} pre\_MSE\_total\_score \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.201 *\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.120\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.350 *** \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]

\hhline{}
\arrayrulecolor{black}

\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt}  \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.082)\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.068)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.084)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]

\hhline{}
\arrayrulecolor{black}

\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} fullYear5TRUE \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} -0.129\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} -0.099\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.062\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]

\hhline{}
\arrayrulecolor{black}

\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt}  \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.328)\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.286)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.341)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]

\hhline{}
\arrayrulecolor{black}

\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} pre\_PS\_tasks\_total\_score \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.345 ** \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.372 *** \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.231\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]

\hhline{}
\arrayrulecolor{black}

\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt}  \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.116)\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.090)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.118)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]

\hhline{}
\arrayrulecolor{black}

\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} raceEthHispanic/Latino \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.275\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.327\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.225\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]

\hhline{}
\arrayrulecolor{black}

\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt}  \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.290)\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.213)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.283)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]

\hhline{}
\arrayrulecolor{black}

\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} raceEthAsian \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.662 *\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.619 **\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.823 **\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]

\hhline{}
\arrayrulecolor{black}

\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt}  \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.279)\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.225)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.298)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]

\hhline{}
\arrayrulecolor{black}

\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} raceEthOther \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.455\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.208\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.095\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]

\hhline{}
\arrayrulecolor{black}

\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt}  \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.292)\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.231)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.300)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]

\hhline{}
\arrayrulecolor{black}

\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} GenderM \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} -0.186\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} -0.337 **\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} -0.267\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]

\hhline{}
\arrayrulecolor{black}

\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt}  \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.155)\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.125)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.155)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]

\hhline{}
\arrayrulecolor{black}

\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} GIFTED1 \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.659 ** \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.429 *\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.336\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]

\hhline{}
\arrayrulecolor{black}

\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt}  \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.243)\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.192)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.260)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]

\hhline{}
\arrayrulecolor{black}

\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} pre.total\_math\_score \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.141 ** \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.241 *** \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.146 **\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]

\hhline{}
\arrayrulecolor{black}

\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt}  \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.052)\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.038)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.054)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]

\hhline{}
\arrayrulecolor{black}

\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} pre.total\_math\_scoreNATRUE \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} -0.350\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.183\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} -0.440\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]

\hhline{}
\arrayrulecolor{black}

\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt}  \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.449)\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.358)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.431)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]

\hhline{}
\arrayrulecolor{black}

\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} Z:Sp \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} -0.135\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.168\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} -0.171\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]

\hhline{}
\arrayrulecolor{black}

\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt}  \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.486)\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.387)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.485)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]

\hhline{>{\huxb{255, 255, 255}{0.4}}->{\huxb{0, 0, 0}{0.4}}->{\huxb{0, 0, 0}{0.4}}->{\huxb{0, 0, 0}{0.4}}-}
\arrayrulecolor{black}

\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} N \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 787\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 1193\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 771\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]

\hhline{}
\arrayrulecolor{black}

\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} R2 \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.605\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.600\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.626\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]

\hhline{}
\arrayrulecolor{black}

\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} logLik \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} -1570.888\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} -2427.557\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} -1528.742\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]

\hhline{}
\arrayrulecolor{black}

\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} AIC \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 3421.776\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 5141.115\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 3339.484\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]

\hhline{>{\huxb{0, 0, 0}{0.8}}->{\huxb{0, 0, 0}{0.8}}->{\huxb{0, 0, 0}{0.8}}->{\huxb{0, 0, 0}{0.8}}-}
\arrayrulecolor{black}

\multicolumn{4}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt}  *** p $<$ 0.001;  ** p $<$ 0.01;  * p $<$ 0.05. \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]

\hhline{}
\arrayrulecolor{black}
\end{tabular}
\end{threeparttable}\par\end{centerbox}

\end{table}

### 2022-12-19 code/fh2t/plots.r

@@ -18,6 +18,8 @@ coefplot(estimates3$BAU$psMod, predictors=names(model.frame(estimates3$BAU$psMod
 ggsave('figure/psModCoef.jpg',width=5,height=3,units='in')
 
 
+
+
 ### plot estimates
 
 ateDF <- imap_dfr(ates,~data.frame(EFF=.25,Alternative=paste('vs.',ifelse(.y=='Dragon','DragonBox',.y)),estimates=.x$coefficients['Z'],

### 2022-12-19 code/fh2t/tables.r

@@ -92,3 +92,15 @@ sink('writeUps/table1.tex')
 kbl(tab1,format='latex',booktabs=TRUE)%>% #c("p{1.5in}",rep("l",5)))%>%
   group_rows("Baseline",1,17,indent=FALSE)%>%group_rows("Post-Treatment",18,19,indent=FALSE)
 sink()
+
+
+load('results/geeResults.RData')
+
+
+sink('writeUps/outcomeRegAppendix.tex')
+outModTab <- huxreg(
+  BAU=estimates3$BAU$outMod,
+  FH2T=estimates3$FH2T$outMod,
+    DragonBox=estimates3$Dragon$outMod,
+    omit_coefs = paste0("ClaIDPre",1:218))%>%print_latex()
+sink()

### 2022-12-19 code/simulation/displaySim.r

@@ -498,28 +498,29 @@ rmse%>%filter(n==500,eff==1,errDist!='mix',mu01==0.3,b1>0)%>%
 sink('writeUps/rmseTabAppendix.tex')
 cbind(
 
-rmse%>%filter(errDist!='mix',mu01==0.3,eff!="Diff",b1==.2)%>%
+rmse%>%filter(errDist!='mix',n==500,eff!="Diff",b1==.2)%>%
   transmute(`Residual\nDist.`=c(norm='Normal',unif='Uniform')[errDist],
          `X:Z\nInt.?`=ifelse(intZ,'Yes','No'),
          `X:S\nInt.?`=ifelse(intS,'Yes','No'),
          estimator=ifelse(estimator=='Mixture','Mix.',estimator),
-         n,
-         `\\\\beta_1`=mu01,
-         `Prin. Eff`=paste0('$\\\\tau^',eff,'$'),
+         
+         `$\\beta_1$`=mu01,
+         `Prin.\nEff`=paste0('$\\tau^',eff,'$'),
          rmse)%>%
 pivot_wider(names_from=estimator,values_from=rmse),
 rmse%>%filter(errDist!='mix',mu01==0.3,eff!="Diff",b1==.5)%>%
-  transmute(`Residual\nDist.`=c(norm='Normal',unif='Uniform')[errDist],
+  transmute(`Res.\nDist.`=c(norm='Norm.',unif='Unif.')[errDist],
          `X:Z\nInt.?`=ifelse(intZ,'Yes','No'),
          `X:S\nInt.?`=ifelse(intS,'Yes','No'),
          estimator=ifelse(estimator=='Mixture','Mix.',estimator),
          n,
-         `\\\\beta_1`=mu01,
-         `Prin. Eff`=paste0('$\\\\tau^',eff,'$'),
+         `$\\beta_1$`=mu01,
+         `Prin.\nEff`=paste0('$\\tau^',eff,'$'),
          rmse)%>%
-pivot_wider(names_from=estimator,values_from=rmse))%>%
+pivot_wider(names_from=estimator,values_from=rmse)%>%
+select(GEEPERs,Mix.,PSW))%>%
   kbl('latex',booktabs=TRUE,col.names=linebreak(names(.)),escape=FALSE,digits=2)%>%
[…]

### 2022-12-19 writeUps/outcomeRegAppendix.tex
application results appendix

@@
  \setlength{\tabcolsep}{0pt}
-\begin{tabular}{l l l l}
+\begin{tabular}{l l l l l}
 
-\hhline{>{\huxb{0, 0, 0}{0.8}}->{\huxb{0, 0, 0}{0.8}}->{\huxb{0, 0, 0}{0.8}}->{\huxb{0, 0, 0}{0.8}}-}
+\hhline{>{\huxb{0, 0, 0}{0.8}}->{\huxb{0, 0, 0}{0.8}}->{\huxb{0, 0, 0}{0.8}}->{\huxb{0, 0, 0}{0.8}}->{\huxb{0, 0, 0}{0.8}}-}
 \arrayrulecolor{black}
@@
 \multicolumn{1}{!{\huxvb{0, 0, 0}{0}}c!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\centering \hspace{6pt}  \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{c!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\centering \hspace{6pt} psModel \hspace{6pt}\huxbpad{6pt}} &
 \multicolumn{1}{c!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\centering \hspace{6pt} BAU \hspace{6pt}\huxbpad{6pt}} &
@@
 
-\hhline{>{\huxb{255, 255, 255}{0.4}}->{\huxb{0, 0, 0}{0.4}}->{\huxb{0, 0, 0}{0.4}}->{\huxb{0, 0, 0}{0.4}}-}
+\hhline{>{\huxb{255, 255, 255}{0.4}}->{\huxb{0, 0, 0}{0.4}}->{\huxb{0, 0, 0}{0.4}}->{\huxb{0, 0, 0}{0.4}}->{\huxb{0, 0, 0}{0.4}}-}
 \arrayrulecolor{black}
@@
 \multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} (Intercept) \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} -2.143\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
 \multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 1.732\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
@@
 \multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt}  \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (1.257)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
 \multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.916)\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
@@
 
+\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} Scale.Score5 \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} -0.594 **\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.207\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.453 *** \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.049\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
+
+\hhline{}
+\arrayrulecolor{black}
+
+\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt}  \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.188)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.126)\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.104)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.132)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
+
+\hhline{}
+\arrayrulecolor{black}
+
+\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} ESOL1 \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} -0.825\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.091\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} -0.459\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} -0.493\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
+
+\hhline{}
+\arrayrulecolor{black}
+
+\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt}  \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.642)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.364)\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.268)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.351)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
+
+\hhline{}
+\arrayrulecolor{black}
+
+\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} IEP1 \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} -1.335 **\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.148\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.234\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.096\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
+
+\hhline{}
+\arrayrulecolor{black}
+
+\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt}  \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.463)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.288)\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.255)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.277)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
+
+\hhline{}
+\arrayrulecolor{black}
+
+\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} pre.total\_time\_on\_tasks \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} -0.447 **\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} -0.131\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.078\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} -0.003\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
+
+\hhline{}
+\arrayrulecolor{black}
+
+\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt}  \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.152)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.091)\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.073)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.096)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
+
+\hhline{}
+\arrayrulecolor{black}
+
+\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} pre\_MSE\_total\_score \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} -0.179\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.201 *\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.120\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.350 *** \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
+
+\hhline{}
+\arrayrulecolor{black}
+
+\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt}  \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.130)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.082)\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.068)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.084)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
+
+\hhline{}
+\arrayrulecolor{black}
+
+\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} fullYear5TRUE \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 1.405 *\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} -0.129\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} -0.099\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.062\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
+
+\hhline{}
+\arrayrulecolor{black}
+
+\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt}  \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.585)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.328)\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.286)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.341)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
+
+\hhline{}
+\arrayrulecolor{black}
+
+\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} pre\_MA\_total\_scoreNATRUE \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} -3.205 *** \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} \hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} \hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} \hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
+
+\hhline{}
+\arrayrulecolor{black}
+
+\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt}  \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.706)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} \hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} \hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} \hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
+
+\hhline{}
+\arrayrulecolor{black}
+
+\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} GenderM \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.169\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} -0.186\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} -0.337 **\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} -0.267\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
+
+\hhline{}
+\arrayrulecolor{black}
+
+\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt}  \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.253)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.155)\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.125)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.155)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
+
+\hhline{}
+\arrayrulecolor{black}
+
+\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} raceEthHispanic/Latino \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} -0.561\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.275\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.327\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.225\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
+
+\hhline{}
+\arrayrulecolor{black}
+
+\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt}  \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.460)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.290)\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.213)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.283)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
+
+\hhline{}
+\arrayrulecolor{black}
+
+\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} raceEthAsian \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} -0.431\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.662 *\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.619 **\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.823 **\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
+
+\hhline{}
+\arrayrulecolor{black}
+
+\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt}  \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.416)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.279)\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.225)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.298)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
+
+\hhline{}
+\arrayrulecolor{black}
+
+\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} raceEthOther \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} -0.131\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.455\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.208\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.095\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
+
+\hhline{}
+\arrayrulecolor{black}
+
+\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt}  \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.492)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.292)\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.231)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.300)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
+
+\hhline{}
+\arrayrulecolor{black}
+
+\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} GIFTED1 \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} -0.240\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.659 ** \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.429 *\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.336\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
+
+\hhline{}
+\arrayrulecolor{black}
+
+\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt}  \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.371)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.243)\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.192)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.260)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
+
+\hhline{}
+\arrayrulecolor{black}
+
+\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} poly(pre\_PS\_tasks\_total\_score, 2, raw = TRUE)1 \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.006\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} \hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} \hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} \hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
+
+\hhline{}
+\arrayrulecolor{black}
+
+\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt}  \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.191)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} \hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} \hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} \hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
+
+\hhline{}
+\arrayrulecolor{black}
+
+\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} poly(pre\_PS\_tasks\_total\_score, 2, raw = TRUE)2 \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} -0.407 **\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} \hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} \hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} \hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
+
+\hhline{}
+\arrayrulecolor{black}
+
+\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt}  \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.133)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} \hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} \hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} \hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
+
+\hhline{}
+\arrayrulecolor{black}
+
 \multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} Z \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} \hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
 \multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.294\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
@@
 \multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt}  \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} \hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
 \multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.277)\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
@@
 \multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} Sp \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} \hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
 \multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.031\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
@@
 \multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt}  \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} \hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
 \multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.522)\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
@@
 
-\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} Scale.Score5 \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.207\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.453 *** \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.049\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
-
-\hhline{}
-\arrayrulecolor{black}
-
-\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt}  \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.126)\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.104)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.132)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
-
-\hhline{}
-\arrayrulecolor{black}
-
-\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} ESOL1 \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.091\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} -0.459\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} -0.493\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
-
-\hhline{}
-\arrayrulecolor{black}
-
-\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt}  \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.364)\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.268)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.351)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
-
-\hhline{}
-\arrayrulecolor{black}
-
-\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} IEP1 \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.148\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.234\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.096\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
-
-\hhline{}
-\arrayrulecolor{black}
-
-\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt}  \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.288)\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.255)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.277)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
-
-\hhline{}
-\arrayrulecolor{black}
-
-\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} pre.total\_time\_on\_tasks \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} -0.131\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.078\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} -0.003\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
-
-\hhline{}
-\arrayrulecolor{black}
-
-\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt}  \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.091)\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.073)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.096)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
-
-\hhline{}
-\arrayrulecolor{black}
-
-\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} pre\_MSE\_total\_score \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.201 *\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.120\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.350 *** \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
-
-\hhline{}
-\arrayrulecolor{black}
-
-\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt}  \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.082)\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.068)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.084)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
-
-\hhline{}
-\arrayrulecolor{black}
-
-\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} fullYear5TRUE \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} -0.129\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} -0.099\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.062\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
-
-\hhline{}
-\arrayrulecolor{black}
-
-\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt}  \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.328)\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.286)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.341)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
-
-\hhline{}
-\arrayrulecolor{black}
-
 \multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} pre\_PS\_tasks\_total\_score \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} \hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
 \multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.345 ** \hspace{6pt}\huxbpad{6pt}} &
@@
 \multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt}  \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} \hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
 \multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.116)\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
@@
 
-\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} raceEthHispanic/Latino \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.275\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.327\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.225\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
-
-\hhline{}
-\arrayrulecolor{black}
-
-\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt}  \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.290)\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.213)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.283)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
-
-\hhline{}
-\arrayrulecolor{black}
-
-\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} raceEthAsian \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.662 *\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.619 **\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.823 **\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
-
-\hhline{}
-\arrayrulecolor{black}
-
-\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt}  \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.279)\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.225)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.298)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
-
-\hhline{}
-\arrayrulecolor{black}
-
-\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} raceEthOther \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.455\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.208\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.095\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
-
-\hhline{}
-\arrayrulecolor{black}
-
-\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt}  \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.292)\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.231)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.300)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
-
-\hhline{}
-\arrayrulecolor{black}
-
-\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} GenderM \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} -0.186\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} -0.337 **\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} -0.267\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
-
-\hhline{}
-\arrayrulecolor{black}
-
-\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt}  \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.155)\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.125)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.155)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
-
-\hhline{}
-\arrayrulecolor{black}
-
-\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} GIFTED1 \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.659 ** \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.429 *\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.336\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
-
-\hhline{}
-\arrayrulecolor{black}
-
-\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt}  \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.243)\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.192)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.260)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
-
-\hhline{}
-\arrayrulecolor{black}
-
 \multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} pre.total\_math\_score \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} \hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
 \multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.141 ** \hspace{6pt}\huxbpad{6pt}} &
@@
 \multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt}  \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} \hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
 \multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.052)\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
@@
 \multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} pre.total\_math\_scoreNATRUE \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} \hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
 \multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} -0.350\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
@@
 \multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt}  \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} \hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
 \multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.449)\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
@@
 \multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} Z:Sp \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} \hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
 \multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} -0.135\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
@@
 \multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt}  \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} \hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
 \multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.486)\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
@@
 \multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} N \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 402\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
 \multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 787\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
@@
 \multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} R2 \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} \hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
 \multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.605\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
@@
 \multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} logLik \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} -200.775\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
 \multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} -1570.888\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
@@
 \multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} AIC \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 447.550\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
 \multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 3421.776\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
@@
 
-\hhline{>{\huxb{0, 0, 0}{0.8}}->{\huxb{0, 0, 0}{0.8}}->{\huxb{0, 0, 0}{0.8}}->{\huxb{0, 0, 0}{0.8}}-}
+\hhline{>{\huxb{0, 0, 0}{0.8}}->{\huxb{0, 0, 0}{0.8}}->{\huxb{0, 0, 0}{0.8}}->{\huxb{0, 0, 0}{0.8}}->{\huxb{0, 0, 0}{0.8}}-}
 \arrayrulecolor{black}
 
-\multicolumn{4}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt}  *** p $<$ 0.001;  ** p $<$ 0.01;  * p $<$ 0.05.
+\multicolumn{5}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt}  *** p $<$ 0.001;  ** p $<$ 0.01;  * p $<$ 0.05.
 \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]

### 2022-12-19 writeUps/outcomeRegAppendix.tex
application results appendix

--- 2022-12-19 writeUps/outcomeRegAppendix.tex
@@
-  \providecommand{\huxb}[2]{\arrayrulecolor[RGB]{#1}\global\arrayrulewidth=#2pt}
-  \providecommand{\huxvb}[2]{\color[RGB]{#1}\vrule width #2pt}
-  \providecommand{\huxtpad}[1]{\rule{0pt}{#1}}
-  \providecommand{\huxbpad}[1]{\rule[-#1]{0pt}{#1}}
-
-\begin{table}[ht]
-\begin{centerbox}
-\begin{threeparttable}
- \setlength{\tabcolsep}{0pt}
-\begin{tabular}{l l l l}
-
-\hhline{>{\huxb{0, 0, 0}{0.8}}->{\huxb{0, 0, 0}{0.8}}->{\huxb{0, 0, 0}{0.8}}->{\huxb{0, 0, 0}{0.8}}-}
-\arrayrulecolor{black}
-
-\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}c!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\centering \hspace{6pt}  \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{c!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\centering \hspace{6pt} BAU \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{c!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\centering \hspace{6pt} FH2T \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{c!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\centering \hspace{6pt} DragonBox \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
-
-\hhline{>{\huxb{255, 255, 255}{0.4}}->{\huxb{0, 0, 0}{0.4}}->{\huxb{0, 0, 0}{0.4}}->{\huxb{0, 0, 0}{0.4}}-}
-\arrayrulecolor{black}
-
-\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} (Intercept) \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 1.732\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 2.834 *** \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 2.510 **\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
-
-\hhline{}
-\arrayrulecolor{black}
-
-\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt}  \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.916)\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.718)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.919)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
-
-\hhline{}
-\arrayrulecolor{black}
-
-\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} Z \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.294\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} -0.174\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} -0.197\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
-
-\hhline{}
-\arrayrulecolor{black}
-
-\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt}  \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.277)\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.223)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.268)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
-
-\hhline{}
-\arrayrulecolor{black}
-
-\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} Sp \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.031\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} -0.216\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.042\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
-
-\hhline{}
-\arrayrulecolor{black}
-
-\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt}  \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.522)\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.412)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.515)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
-
-\hhline{}
-\arrayrulecolor{black}
-
-\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} Scale.Score5 \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.207\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.453 *** \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.049\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
-
-\hhline{}
-\arrayrulecolor{black}
-
-\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt}  \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.126)\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.104)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.132)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
-
-\hhline{}
-\arrayrulecolor{black}
-
-\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} ESOL1 \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.091\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} -0.459\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} -0.493\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
-
-\hhline{}
-\arrayrulecolor{black}
-
-\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt}  \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.364)\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.268)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.351)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
-
-\hhline{}
-\arrayrulecolor{black}
-
-\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} IEP1 \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.148\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.234\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.096\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
-
-\hhline{}
-\arrayrulecolor{black}
-
-\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt}  \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.288)\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.255)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.277)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
-
-\hhline{}
-\arrayrulecolor{black}
-
-\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} pre.total\_time\_on\_tasks \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} -0.131\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.078\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} -0.003\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
-
-\hhline{}
-\arrayrulecolor{black}
-
-\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt}  \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.091)\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.073)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.096)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
-
-\hhline{}
-\arrayrulecolor{black}
-
-\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} pre\_MSE\_total\_score \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.201 *\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.120\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.350 *** \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
-
-\hhline{}
-\arrayrulecolor{black}
-
-\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt}  \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.082)\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.068)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.084)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
-
-\hhline{}
-\arrayrulecolor{black}
-
-\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} fullYear5TRUE \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} -0.129\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} -0.099\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.062\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
-
-\hhline{}
-\arrayrulecolor{black}
-
-\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt}  \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.328)\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.286)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.341)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
-
-\hhline{}
-\arrayrulecolor{black}
-
-\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} pre\_PS\_tasks\_total\_score \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.345 ** \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.372 *** \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.231\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
-
-\hhline{}
-\arrayrulecolor{black}
-
-\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt}  \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.116)\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.090)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.118)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
-
-\hhline{}
-\arrayrulecolor{black}
-
-\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} raceEthHispanic/Latino \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.275\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.327\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.225\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
-
-\hhline{}
-\arrayrulecolor{black}
-
-\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt}  \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.290)\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.213)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.283)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
-
-\hhline{}
-\arrayrulecolor{black}
-
-\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} raceEthAsian \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.662 *\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.619 **\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.823 **\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
-
-\hhline{}
-\arrayrulecolor{black}
-
-\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt}  \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.279)\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.225)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.298)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
-
-\hhline{}
-\arrayrulecolor{black}
-
-\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} raceEthOther \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.455\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.208\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.095\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
-
-\hhline{}
-\arrayrulecolor{black}
-
-\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt}  \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.292)\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.231)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.300)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
-
-\hhline{}
-\arrayrulecolor{black}
-
-\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} GenderM \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} -0.186\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} -0.337 **\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} -0.267\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
-
-\hhline{}
-\arrayrulecolor{black}
-
-\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt}  \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.155)\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.125)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.155)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
-
-\hhline{}
-\arrayrulecolor{black}
-
-\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} GIFTED1 \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.659 ** \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.429 *\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.336\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
-
-\hhline{}
-\arrayrulecolor{black}
-
-\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt}  \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.243)\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.192)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.260)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
-
-\hhline{}
-\arrayrulecolor{black}
-
-\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} pre.total\_math\_score \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.141 ** \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.241 *** \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.146 **\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
-
-\hhline{}
-\arrayrulecolor{black}
-
-\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt}  \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.052)\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.038)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.054)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
-
-\hhline{}
-\arrayrulecolor{black}
-
-\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} pre.total\_math\_scoreNATRUE \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} -0.350\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.183\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} -0.440\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
-
-\hhline{}
-\arrayrulecolor{black}
-
-\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt}  \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.449)\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.358)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.431)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
-
-\hhline{}
-\arrayrulecolor{black}
-
-\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} Z:Sp \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} -0.135\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.168\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} -0.171\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
-
-\hhline{}
-\arrayrulecolor{black}
-
-\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt}  \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.486)\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.387)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.485)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
-
-\hhline{>{\huxb{255, 255, 255}{0.4}}->{\huxb{0, 0, 0}{0.4}}->{\huxb{0, 0, 0}{0.4}}->{\huxb{0, 0, 0}{0.4}}-}
-\arrayrulecolor{black}
-
-\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} N \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 787\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 1193\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 771\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
-
-\hhline{}
-\arrayrulecolor{black}
-
-\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} R2 \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.605\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.600\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.626\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
-
-\hhline{}
-\arrayrulecolor{black}
-
-\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} logLik \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} -1570.888\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} -2427.557\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} -1528.742\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
-
-\hhline{}
-\arrayrulecolor{black}
-
-\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} AIC \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 3421.776\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 5141.115\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 3339.484\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
-
-\hhline{>{\huxb{0, 0, 0}{0.8}}->{\huxb{0, 0, 0}{0.8}}->{\huxb{0, 0, 0}{0.8}}->{\huxb{0, 0, 0}{0.8}}-}
-\arrayrulecolor{black}
-
-\multicolumn{4}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt}  *** p $<$ 0.001;  ** p $<$ 0.01;  * p $<$ 0.05.
-\hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
-
-\hhline{}
-\arrayrulecolor{black}
+\begin{table}
+\begin{center}
+\begin{tabular}{l c c c c}
+\hline
+ & psModel & BAU & FH2T & DragonBox \\
+\hline
+(Intercept)                                        & $-2.14$       & $1.73$      & $2.83^{***}$ & $2.51^{**}$  \\
+                                                   & $(1.26)$      & $(0.92)$    & $(0.72)$     & $(0.92)$     \\
+Scale.Score5                                       & $-0.59^{**}$  & $0.21$      & $0.45^{***}$ & $0.05$       \\
+                                                   & $(0.19)$      & $(0.13)$    & $(0.10)$     & $(0.13)$     \\
+ESOL1                                              & $-0.83$       & $0.09$      & $-0.46$      & $-0.49$      \\
+                                                   & $(0.64)$      & $(0.36)$    & $(0.27)$     & $(0.35)$     \\
+IEP1                                               & $-1.34^{**}$  & $0.15$      & $0.23$       & $0.10$       \\
+                                                   & $(0.46)$      & $(0.29)$    & $(0.25)$     & $(0.28)$     \\
+pre.total\_time\_on\_tasks                         & $-0.45^{**}$  & $-0.13$     & $0.08$       & $-0.00$      \\
+                                                   & $(0.15)$      & $(0.09)$    & $(0.07)$     & $(0.10)$     \\
+pre\_MSE\_total\_score                             & $-0.18$       & $0.20^{*}$  & $0.12$       & $0.35^{***}$ \\
+                                                   & $(0.13)$      & $(0.08)$    & $(0.07)$     & $(0.08)$     \\
+fullYear5TRUE                                      & $1.40^{*}$    & $-0.13$     & $-0.10$      & $0.06$       \\
+                                                   & $(0.58)$      & $(0.33)$    & $(0.29)$     & $(0.34)$     \\
+pre\_MA\_total\_scoreNATRUE                        & $-3.20^{***}$ &             &              &              \\
+                                                   & $(0.71)$      &             &              &              \\
+GenderM                                            & $0.17$        & $-0.19$     & $-0.34^{**}$ & $-0.27$      \\
+                                                   & $(0.25)$      & $(0.15)$    & $(0.13)$     & $(0.16)$     \\
+raceEthHispanic/Latino                             & $-0.56$       & $0.28$      & $0.33$       & $0.23$       \\
+                                                   & $(0.46)$      & $(0.29)$    & $(0.21)$     & $(0.28)$     \\
+raceEthAsian                                       & $-0.43$       & $0.66^{*}$  & $0.62^{**}$  & $0.82^{**}$  \\
+                                                   & $(0.42)$      & $(0.28)$    & $(0.22)$     & $(0.30)$     \\
+raceEthOther                                       & $-0.13$       & $0.45$      & $0.21$       & $0.10$       \\
+                                                   & $(0.49)$      & $(0.29)$    & $(0.23)$     & $(0.30)$     \\
+GIFTED1                                            & $-0.24$       & $0.66^{**}$ & $0.43^{*}$   & $0.34$       \\
+                                                   & $(0.37)$      & $(0.24)$    & $(0.19)$     & $(0.26)$     \\
+poly(pre\_PS\_tasks\_total\_score, 2, raw = TRUE)1 & $0.01$        &             &              &              \\
+                                                   & $(0.19)$      &             &              &              \\
+poly(pre\_PS\_tasks\_total\_score, 2, raw = TRUE)2 & $-0.41^{**}$  &             &              &              \\
+                                                   & $(0.13)$      &             &              &              \\
+Z                                                  &               & $0.29$      & $-0.17$      & $-0.20$      \\
+                                                   &               & $(0.28)$    & $(0.22)$     & $(0.27)$     \\
+Sp                                                 &               & $0.03$      & $-0.22$      & $0.04$       \\
+                                                   &               & $(0.52)$    & $(0.41)$     & $(0.52)$     \\
+pre\_PS\_tasks\_total\_score                       &               & $0.34^{**}$ & $0.37^{***}$ & $0.23$       \\
+                                                   &               & $(0.12)$    & $(0.09)$     & $(0.12)$     \\
+pre.total\_math\_score                             &               & $0.14^{**}$ & $0.24^{***}$ & $0.15^{**}$  \\
+                                                   &               & $(0.05)$    & $(0.04)$     & $(0.05)$     \\
+pre.total\_math\_scoreNATRUE                       &               & $-0.35$     & $0.18$       & $-0.44$      \\
+                                                   &               & $(0.45)$    & $(0.36)$     & $(0.43)$     \\
+Z:Sp                                               &               & $-0.14$     & $0.17$       & $-0.17$      \\
+                                                   &               & $(0.49)$    & $(0.39)$     & $(0.49)$     \\
+\hline
+AIC                                                & $447.55$      & $$          & $$           & $$           \\
+BIC                                                & $539.47$      & $$          & $$           & $$           \\
+Log Likelihood                                     & $-200.78$     & $$          & $$           & $$           \\
+Deviance                                           & $401.55$      & $$          & $$           & $$           \\
+Num. obs.                                          & $402$         & $787$       & $1193$       & $771$        \\
+R$^2$                                              & $$            & $0.60$      & $0.60$       & $0.63$       \\
+Adj. R$^2$                                         & $$            & $0.52$      & $0.55$       & $0.54$       \\
+\hline
+\multicolumn{5}{l}{\scriptsize{$^{***}p<0.001$; $^{**}p<0.01$; $^{*}p<0.05$}}
 \end{tabular}
-\end{threeparttable}\par\end{centerbox}
-
+\caption{Statistical models}
+\label{table:coefficients}
+\end{center}
 \end{table}

### 2022-12-19 code/fh2t/tables.r

@@ -6,6 +6,7 @@ library(kableExtra)
 library(xtable)
 source('code/regression.r')
 library(huxtable)
+library(texreg)
 select <- dplyr::select
 
 options(knitr.kable.NA = '')
@@ -98,9 +99,10 @@ load('results/geeResults.RData')
 
 
 sink('writeUps/outcomeRegAppendix.tex')
-outModTab <- huxreg(
+texreg(list(
+  psModel=estimates3$BAU$psMod,
   BAU=estimates3$BAU$outMod,
   FH2T=estimates3$FH2T$outMod,
-    DragonBox=estimates3$Dragon$outMod,
-    omit_coefs = paste0("ClaIDPre",1:218))%>%print_latex()
-sink()
+    DragonBox=estimates3$Dragon$outMod),
+    file='writeUps/outcomeRegAppendix.tex',
+    omit.coef = c('SchIDPre|ClaIDPre'))

### 2022-12-19 writeUps/paper.tex
Updates from Overleaf

--- 2022-12-19 writeUps/paper.tex
@@
 
-\title[GEEPERS]{GEEPERs: Principal Stratification using Principal Scores and Stacked Estimating Equations}
+\title[\textsc{geepers}]{\textsc{geepers}: Principal Stratification using Principal Scores and Stacked Estimating Equations}
 
@@
 %, just as in the pre-treatment subgroup case.
-We will refer to the estimators presented here as ``GEEPERs'': generalized estimating equations for principal effects using regressions.
-GEEPERs estimates rely on correct specification of regression functions, but unlike common fully-parametric methods, they make no assumptions about the shape of the distribution of regression errors.
-Unlike principal score weighting, GEEPERs does not assume principal ignorability, and admits to sandwich-style standard error estimates, which we provide in the appendix.
+We will refer to the estimators presented here as ``\textsc{geepers}'': generalized estimating equations for principal effects using regressions.
+\textsc{geepers} estimates rely on correct specification of regression functions, but unlike common fully-parametric methods, they make no assumptions about the shape of the distribution of regression errors.
+Unlike principal score weighting, \textsc{geepers} does not assume principal ignorability, and admits to sandwich-style standard error estimates, which we provide in the appendix.
 
@@
 The following section gives a formal introduction to principal stratification and briefly describes two common estimation techniques, parametric mixture modeling and principal score weighting.
-Next, Section \ref{sec:geepers} develops GEEPERs estimation, and gives conditions for its consistency.
-Section \ref{sec:simulation} describes a simulation study examining the operating characteristics of GEEPERs estimates and comparing them to estimates from parametric mixture modeling and principal score weighting.
-Under ideal circumstances, estimates from parametric mixture models and weighting estimators are somewhat more efficient than GEEPERs, but when assumptions are violated GEEPERs easily out-performs the other two methods.
-Section \ref{sec:fh2t} illustrates GEEPERs in analyzing data from a recent large-scale field trial of educational software, and Section \ref{sec:discussion} concludes.
+Next, Section \ref{sec:geepers} develops \textsc{geepers} estimation, and gives conditions for its consistency.
+Section \ref{sec:simulation} describes a simulation study examining the operating characteristics of \textsc{geepers} estimates and comparing them to estimates from parametric mixture modeling and principal score weighting.
+Under ideal circumstances, estimates from parametric mixture models and weighting estimators are somewhat more efficient than \textsc{geepers}, but when assumptions are violated \textsc{geepers} easily out-performs the other two methods.
+Section \ref{sec:fh2t} illustrates \textsc{geepers} in analyzing data from a recent large-scale field trial of educational software, and Section \ref{sec:discussion} concludes.
 
@@
 
-Principal score weighting does not require any distributional assumptions and (after estimating principal scores) is very easy to implement.
+Principal score weighting (\textsc{psw}) does not require any distributional assumptions and (after estimating principal scores) is very easy to implement.
 On the other hand, the principal ignorability assumption is strong and restrictive.
@@
 
-\section{GEEPERs}\label{sec:geepers}
+\section{\textsc{geepers}}\label{sec:geepers}
 
@@
 
-\begin{prop}{GEEPERs}\label{prop:reg2}
+\begin{prop}{\textsc{geepers}}\label{prop:reg2}
 
@@
 In short, researchers can estimate principal effects under strong monotonicity by simply imputing missing $\st$ values with estimated principal scores, fitting an OLS regression, and estimating standard errors with a sandwich formula.
-For the remainder of the paper, we will refer to principal effect estimators $\heff1_{RCI}$ and $\heff0_{RCI}$---our preferred estimators---as ``GEEPERs.''
+For the remainder of the paper, we will refer to principal effect estimators $\heff1_{RCI}$ and $\heff0_{RCI}$---our preferred estimators---as ``\textsc{geepers}.''
 
@@
 
-We conducted a simulation study to compare the performance of GEEPERs compared to a Bayesian mixture model and to a principal score weighting estimator, in both favorable and unfavorable circumstances.
+We conducted a simulation study to compare the performance of \textsc{geepers} compared to a Bayesian mixture model and to a principal score weighting estimator, in both favorable and unfavorable circumstances.
 The study was designed to answer three sets of overarching questions.
-First, how does the performance of the GEEPERs estimator vary under different conditions, including sample size, the extent to which $\bx$ predicts $\st$, whether and how the principal effects themselves vary, and when important interactions are omitted from the outcome model.
-Second, how does GEEPERs compare with mixture modeling?
+First, how does the performance of the \textsc{geepers} estimator vary under different conditions, including sample size, the extent to which $\bx$ predicts $\st$, whether and how the principal effects themselves vary, and when important interactions are omitted from the outcome model.
+Second, how does \textsc{geepers} compare with mixture modeling?
 Is it competitive in circumstances favorable to both techniques?
 Does it avoid the pitfalls of mixture modeling when the parametric assumptions of the normal mixture model fail?
-Third, are there ways or circumstances when the principal score weighting estimator outperforms GEEPERs even though principal ignorability does not hold?
+Third, are there ways or circumstances when the principal score weighting estimator outperforms \textsc{geepers} even though principal ignorability does not hold?
 
@@
 
-\textbf{\emph{Sample Size per condition $\bm{n}$:}} The guarantees of M-estimation are all asymptotic, so we examined the behavior of our GEEPERs at a range of sample sizes to determine its finite-sample properties.
+\textbf{\emph{Sample Size per condition $\bm{n}$:}} The guarantees of M-estimation are all asymptotic, so we examined the behavior of our \textsc{geepers} at a range of sample sizes to determine its finite-sample properties.
 
@@
 When neither interaction was present, $\gamma_2=\gamma_3=0$, then $\gamma_1=1/\sqrt{6}$.
-Then, since  $Var(\gamma_1(x_1+x2+x_3))=1/2$ and $Var(\epsilon_i)=1/2$, covariates explained half of the variance of $Y$ within treatment group and principal stratum, and observed covariates $x_1$ and $x_2$ explained 1/3 of that variance.
+Then, since  $Var(\gamma_1(x_1+x_2+x_3))=1/2$ and $Var(\epsilon_i)=1/2$, covariates explained half of the variance of $Y$ within treatment group and principal stratum, and observed covariates $x_1$ and $x_2$ explained 1/3 of that variance.
 When there was an interaction between covariates and $\sti$, then $\gamma_1=3/(4\sqrt{6})$ and $\gamma_2=1/(2\sqrt{6})$, so that the slopes for $x_1$ and $x_2$ varied by half their magnitude between principal strata, and on average they were equal to $1/\sqrt{6}$, as in the no-interaction case.
-Interactions between covariates and $Z$ were controlled by $\gamma_3$: when there were interactions, $\gamma_3=1/(2sqrt{6})$, otherwise $\gamma_3=0$.
+Interactions between covariates and $Z$ were controlled by $\gamma_3$: when there were interactions, $\gamma_3=1/(2\sqrt{6})$, otherwise $\gamma_3=0$.
 Either type of interaction violates Assumption \ref{ass:rci} since if $\gamma_2\ne 0$ or $\gamma_3 \ne 0$, there is no single vector $\bm{\gamma}$ that can fully capture the dependence of $\yc$ and $\yt$ on $S$---instead, the dependence varies with $\st$ and/or with $Z$.
@@
 
-Using each simulated dataset, we estimated $\eff0$ and $\eff1$ with three methods: a mixture model \eqref{eq:mixtureModelPS} including principal scores, GEEPERs, and third, the principal score weighting estimator \eqref{eq:psw}.
+Using each simulated dataset, we estimated $\eff0$ and $\eff1$ with three methods: a mixture model \eqref{eq:mixtureModelPS} including principal scores, \textsc{geepers}, and third, the principal score weighting estimator \eqref{eq:psw}.
 
@@
 
-In contrast, for GEEPERs and principal score weighting the principal score model was fit by itself, to data from the treatment group, using standard maximum likelihood methods.
-For GEEPERs, we used the estimated principal scores in the control group, along with the observed $S$ in the treatment group, to construct $R$, and then fit the model \eqref{eq:reg2} with OLS. 
-To estimate standard errors, we used the \texttt{sandwich} package \citep{sandwich} to extract estimating equations from \eqref{eq:psModsim} and \eqref{eq:simReg} and computed the sandwich covariance matrix following the procedure detailed in the appendix.
-
-For principal score weighting we plugged estimated principal scores in to \eqref{eq:psw} to estimate $\muc0$ and $\muc1$, estimated $\mut0$ and $\mut1$ with the sample means of subjects with $Z=1$ and $S=0$ or $S=1$, respectively, and estimated $\heff0=\hat{\mut0}-\hat{\muc0}$ and $\heff1=\hat{\mut1}-\hat{\muc1}$.
+In contrast, for \textsc{geepers} and principal score weighting the principal score model was fit by itself, to data from the treatment group, using standard maximum likelihood methods.
+For \textsc{geepers}, we used the estimated principal scores in the control group, along with the observed $S$ in the treatment group, to construct $R$, and then fit the model \eqref{eq:reg2} with OLS.
+To estimate standard errors, we computed the sandwich covariance matrix following the procedure detailed in the appendix.
+
+For principal score weighting we plugged estimated principal scores in to \eqref{eq:psw} to estimate $\muc0$ and $\muc1$, estimated $\mut0$ and $\mut1$ with the sample means of subjects with $Z=1$ and $S=0$ or $S=1$, respectively, and estimated $\heff0=\hat{\mu}_{Tw}^0-\hat{\mu}_{Cw}^0$ and $\heff1=\hat{\mu}_{Tw}^1-\hat{\mu}_{Cw}^1$.
 
@@
 \subsubsection{Varying $n$ and $\alpha$}
-When $\alpha=0$, Assumption \ref{ass:vps} is violated, and M-estimates are highly erratic; however, as Table \ref{tab:coverage} shows, when other assumptions hold inference remains valid.
-
-Figure \ref{fig:alphan} shows the empirical bias and standard error (SE) of the GEEPERs and mixture estimator of $\eff1$ as sample size and $\alpha$ vary.
-Across $n$ and $\alpha$, GEEPERs has lower bias and higher SE than the mixture estimator.
-GEEPERs is positively biased for low $n$ with bias converging to close to 0 (up to simulation error) for $n\ge 300$, and appears to be roughly unbiased (up to simulation error) for $\alpha \ge 0.2$, with $n$ fixed at 500.
-The SE of the GEEPERs decreases rapidly for $100\le n \le 500$ and $0.2\le \alpha \le 0.5$ and more gradually for higher values of $n$ and $\alpha$.
+When $\alpha=0$, Assumption \ref{ass:vps} was violated, and \textsc{geepers} estimates were highly erratic; however, as Table \ref{tab:coverage} shows, when other assumptions held inference remained valid.
+
+Figure \ref{fig:alphan} shows the empirical bias and standard error (SE) of the \textsc{geepers} and mixture estimator of $\eff1$ as sample size and $\alpha$ varied.
+Across $n$ and $\alpha$, \textsc{geepers} had lower bias and higher SE than the mixture estimator.
+\textsc{geepers} was positively biased for low $n$ with bias converging to close to 0 (up to simulation error) for $n\ge 300$, and was roughly unbiased (up to simulation error) for $\alpha \ge 0.2$, with $n$ fixed at 500.
+The SE of the \textsc{geepers} decreased rapidly for $100\le n \le 500$ and $0.2\le \alpha \le 0.5$ and more gradually for higher values of $n$ and $\alpha$.
 
@@
   \includegraphics[width=\textwidth]{../simFigs/biasSEbyB1N.jpg}
-  \caption{Empirical bias and standard error of GEEPERs and Mixture estimates of $\muc1$ as sample size per treatment group $n$ or $\alpha$ varies.
+  \caption{Empirical bias and standard error of \textsc{geepers} and Mixture estimates of $\muc1$ as sample size per treatment group $n$ or $\alpha$ varied.
 There were 500 replications for each $n$ or $\alpha$.
@@
 
-Figure \ref{fig:boxplots} shows violin plots, layered on top of jittered scatterplots, showing estimation error for GEEPERs, the mixture estimator, and the PSW estimator of $\eff1$ under various sets of conditions, and Table \ref{tab:coverage} shows empirical covarage of nominal 95\% confidence interval estimates under a somewhat wider set of circumstances when $\muc1=0.3$.
-The leftmost panel of Figure \ref{fig:boxplots} and the top line of Table \ref{tab:coverage} show results conditions favorable to both the mixture model and to GEEPERs---normal residuals and no interactions in the data generating model between covariates and either $S$ or $Z$.
-In this case, GEEPERs is roughly unbiased, but less precise than the other two methods---when $\alpha=0.2$ it is much noisier, and when $\alpha=0.5$ it is only slightly less precise.
-The mixture estimator's bias depends on $\muc1$---when $\muc1=0$, the bias is slight, and when $\muc1=0.3$ the bias is more substantial.
-This is somewhat surprising, because when $\muc1=0$, there is no separation between the principal strata in the control group, and previous literature \citep{griffin2008application} has identified this scenario as a particular challenge for mixture modeling.
-The PSW estimator is also biased---due, presumably, to the violation of Principal Ignorability, Assumption \ref{ass:PI}---inherent in the data generating model.
-On the other hand, the PSW estimator has easily the lowest sampling variance of the three estimators.
+Figure \ref{fig:boxplots} shows violin plots, layered on top of jittered scatterplots, showing estimation error for \textsc{geepers}, the mixture estimator, and the \textsc{psw} estimator of $\eff1$ under various sets of conditions, and Table \ref{tab:coverage} shows empirical coverage of nominal 95\% confidence interval estimates under a somewhat wider set of circumstances when $\muc1=0.3$.
+More complete results, including root-mean-squared-error estimates, are in the online appendix.
+The leftmost panel of Figure \ref{fig:boxplots} and the top line of Table \ref{tab:coverage} show results under conditions favorable to both the mixture model and to \textsc{geepers}---normal residuals and no interactions in the data generating model between covariates and either $S$ or $Z$.
+\textsc{geepers} was roughly unbiased, but less precise than the other two methods---when $\alpha=0.2$ it was much noisier, and when $\alpha=0.5$ it was only slightly less precise.
+The mixture estimator's bias depended on $\muc1$---when $\muc1=0$, the bias was slight, and when $\muc1=0.3$ the bias was more substantial.
+This is somewhat surprising, because when $\muc1=0$, there was no separation between the principal strata in the control group, and previous literature \citep{griffin2008application} has identified this scenario as a particular challenge for mixture modeling.
+The \textsc{psw} estimator was also biased---due, presumably, to the violation of Principal Ignorability, Assumption \ref{ass:PI}---inherent in the data generating model.
+On the other hand, the \textsc{psw} estimator had easily the lowest sampling variance of the three estimators.
 The middle panel of Figure \ref{fig:boxplots} and the fifth line of Table \ref{tab:coverage} show a similar set of results for when the residuals in \eqref{eq:y-sim} are uniformly distributed.
-In this case, GEEPERs and PSW estimator behave in a similar fashion to the normal case, but the mixture estimator shows a bimodal pattern--sometimes over-estimating $\eff1$ and sometimes underestimating, but rarely estimating $\eff1$ accurately.
-Credible intervals from the mixture model exhibit severe undercoverage, while GEEPERs confidence intervals achieve at least their nominal level.
+In this case, \textsc{geepers} and \textsc{psw} estimator, which do not assume normality, behaved in a similar fashion as in the normal case, but the mixture estimator showed a bimodal pattern--sometimes over-estimating $\eff1$ and sometimes underestimating, but rarely estimating $\eff1$ accurately.
+Credible intervals from the mixture model exhibited severe undercoverage, while \textsc{geepers} confidence intervals achieved at least their nominal level.
 
 %\sloppy
-The rightmost panel of Figure \ref{fig:boxplots} returns to the case of normal residuals---and only shows results when $\muc1=0.3$ and $\alpha=0.5$---but introduces interactions in the data generating model between covariates and either $\st$, $Z$, or both.
-That is, the coefficients on the covariates in the outcome model vary between principal strata and/or between treatment groups, violating Assumption \ref{ass:rci}..
-%Presence of interactions could undermine the GEEPERs by inducing a relationship between principal scores and potential outcomes, even after adjusting the outcomes for covariate effects.
-Fortunately, the presence of interactions in the data generating model in this scenario appear to make little difference when $\alpha=0.5$; however, Table \ref{tab:coverage} shows that when $\alpha=0$ or 0.2, and there are interactions between $\bx$ and $Z$ (i.e.
-$\gamma_3\ne 0$) GEEPERs 95\% confidence intervals tend to under-cover.
-Nevertheless, coverage of GEEPERs intervals is nearly always substantially higher than coverage of corresponding 95\% credible intervals from mixture models.%\footnote{}
+The rightmost panel of Figure \ref{fig:boxplots} returns to the case of normal residuals---and only shows results when $\muc1=0.3$ and $\alpha=0.5$---but introduces interactions in the data generating model between covariates and either $\st$, $Z$, or both, violating Assumption \ref{ass:rci}.
+%Presence of interactions could undermine the \textsc{geepers} by inducing a relationship between principal scores and potential outcomes, even after adjusting the outcomes for covariate effects.
+Fortunately, these interactions made little difference when $\alpha=0.5$; however, Table \ref{tab:coverage} shows that when $\alpha=0$ or 0.2, interactions between $\bx$ and $Z$ were present (i.e.
+$\gamma_3\ne 0$) \textsc{geepers} 95\% confidence intervals tended to under-cover.
+Nevertheless, coverage of \textsc{geepers} intervals was nearly always substantially higher than coverage of corresponding 95\% credible intervals from mixture models.%\footnote{}
 
@@
 
-\caption{\label{tab:coverage}Empirical coverage of nominal 95\% confidence intervals for GEEPERs and Bayesian mixture modeling under varying conditions.}
+\caption{\label{tab:coverage}Empirical coverage of nominal 95\% confidence intervals for estimates of $\eff1$.}
 \centering
@@
 \cmidrule(l{3pt}r{3pt}){4-5} \cmidrule(l{3pt}r{3pt}){6-7} \cmidrule(l{3pt}r{3pt}){8-9}
-\makecell[l]{Residual\\Dist.} & \makecell[c]{X:Z\\Int.?} & \makecell[r]{X:S\\Int.?} & \rotatebox[origin=c]{300}{GEEPERs} & \rotatebox[origin=c]{300}{Mixture}  & \rotatebox[origin=c]{300}{GEEPERs} & \rotatebox[origin=c]{300}{Mixture}& \rotatebox[origin=c]{300}{GEEPERs} & \rotatebox[origin=c]{300}{Mixture}\\
+\makecell[l]{Residual\\Dist.} & \makecell[c]{X:Z\\Int.?} & \makecell[r]{X:S\\Int.?} & \rotatebox[origin=c]{300}{\textsc{geepers}} & \rotatebox[origin=c]{300}{Mixture}  & \rotatebox[origin=c]{300}{\textsc{geepers}} & \rotatebox[origin=c]{300}{Mixture}& \rotatebox[origin=c]{300}{\textsc{geepers}} & \rotatebox[origin=c]{300}{Mixture}\\
 \midrule
@@
 \bottomrule
-\multicolumn{9}{p{0.9\textwidth}}{\footnotesize Standard error $\approx 1$ percentage point, based on 500 replications and true coverage probabilities of 0.95.}
+\multicolumn{9}{p{0.9\textwidth}}{\footnotesize Based on 500 replications. $n=500$, $\beta_1=0.3$. Simulation standard error $\approx 1$ percentage point.}
 
@@
 In a recent educational field experiment, middle school students completing their math schoolwork on a computer program were randomized between four conditions. 
-In all four conditions, students worked on their assigned math program in class on 12 different occasions---a prior assessment, a mid-test, a post-test, and nine learning sessions---spaced throughout the school year.
-The four conditions included two gamified programs---DragonBox and From Here to There (FH2T)---and two programs in which students used computers to work through a series of algebra problems taken from the district's textbook---``Assistments'' and an active control called ``Business as Usual'' (BAU).
+These included two gamified programs---DragonBox and From Here to There (FH2T)---and two programs in which students used computers to work through a series of algebra problems taken from the district's textbook---``Assistments'' and an active control called ``Business as Usual'' (BAU).
 % the active control condition in which students simply used the computer to work on a sequence of math problems.
@@
 %Another condition, the From Here to There (FH2T) program, used visual cues such as color and spacing to teach algebraic concepts, and allowed students to solve equations by manipulating expressions via dragging and dropping.
-%In the ASSISTments condition, students used computers to work through a series of math problems taken from the district's textbook. 
-Students in the ASSISTments condition had access to a series hints for each problem, culminating in a ``bottom out'' hint that contained the answer and were given immediate feedback on their answers.
+%In ASSISTments, students used computers to work through a series of math problems taken from the district's textbook. 
+Students in ASSISTments had access to a series hints for each problem, culminating in a ``bottom out'' hint that contained the answer and were given immediate feedback on their answers.
 % and had to enter the correct answer in order to proceed to the next problem.
 In the BAU condition students had no access to hints and received feedback only after completing the assignment. 
-In this illustration, we will compare the ASSISTments condition (which we arbitrarily label ``treatment'')---the only condition featuring bottom-out hints---with each of the three others separately.
-Our goal is to estimate different treatment effects for groups of students who would, if assigned to the ASSISTments condition, request few or many bottom out hints.
+In all four conditions, students worked on their assigned math program in class on 12 different occasions spaced throughout the school year: a prior assessment, a mid-test, a post-test, and nine learning sessions.
+In this illustration, we will compare ASSISTments---the only condition featuring bottom-out hints---with each of the three others separately.
+Our goal is to estimate different treatment effects for groups of students who would, if assigned to ASSISTments, request few or many bottom out hints.
 
@@
 On the other hand, they allow students to ``game the system,'' by simply requesting enough hints to see the right answer, and use the software without ever working to solve any problems \citep[e.g.][]{guo2008trying}.
-However, in this experiment, the ability to bottom out was only one of several differences between the treatment conditions.
-One way to disentangle the role of bottoming out in any treatment effect would be to estimate average treatment effects of being assigned to the ASSISTments condition, relative to any of the other three, separately for subjects who would (if assigned to treatment) bottom out frequently, and for subjects who would not.
+
+In this experiment, the ability to bottom out was only one of several differences between the treatment conditions.
+One way to disentangle the role of bottoming out in any treatment effect would be to estimate average treatment effects of being assigned to ASSISTments, relative to any of the other three, separately for subjects who would (if assigned to ASSISTments) bottom out frequently, and for subjects who would not.
 If the availability of bottom-out hints plays an important positive role in ASSISTments's efficacy, we might expect larger effects of condition on students who would bottom out frequently than for those who would not; if bottom-out hints are harmful, we might expect the opposite.
@@
 \sloppy
-For the purpose of demonstrating GEEPERs, we gathered dichotomized data on %which students in the treatment condition requested at least one bottom-out hint.
-whether students in the treatment condition requested more than the median number of bottom-out hints (11).
-That is, $\st=1$ for students who, if assigned to the treatment condition, would request at more than 11 bottom-out hints (``bottom-outers''), and $\st=0$ for students who would request 11 or fewer (``non-bottom-outers'').
+We gathered dichotomized data on %which students in the treatment condition requested at least one bottom-out hint.
+whether students in ASSISTments requested more than the median number of bottom-out hints (i.e. 11).
+That is, $\st=1$ for students who, if assigned to ASSISTments, would request at more than 11 bottom-out hints (``bottom-outers''), and $\st=0$ for students who would request 11 or fewer (``non-bottom-outers'').
 %least one bottom-out hint, and $\st=0$ for students who would not.
 The outcome of interest $Y$ in our example is students' total scores on a ten-item posttest completed within the online tutoring system; $Y$ is an integer ranging from 0 to 10.
-The goal of our analysis will be to determine the principal effects $\tau_1=\EE[\yt-\yc|\st=0]$ and $\tau_0=\EE[\yt-\yc|\st=0]$, the average effects of assignment to the treatment condition for bottom-outers and non-bottom-outers.
+The goal of our analysis will be to determine the principal effects $\tau_1=\EE[\yt-\yc|\st=0]$ and $\tau_0=\EE[\yt-\yc|\st=0]$, the average effects of assignment to ASSISTments (``$T$'' subscripts) versus the other conditions (``$C$'') for bottom-outers and non-bottom-outers.
 
@@
 
-The data analysis we present here is intended as a demonstration of GEEPERs, rather than for its substantive conclusions, and our description omits discussion of some important methodological considerations, including attrition and post-selection inference.
+The data analysis we present here is intended as a demonstration of \textsc{geepers}, rather than for its substantive conclusions, and our description omits discussion of some important methodological considerations, including attrition and post-selection inference.
 For a more complete discussion of the experimental design, the conditions being compared, and the impact analysis, please see \citet{impactPaper}.
@@
 
-\subsection{Data Analysis}
-\subsubsection{Principal Score Model}\label{sec:psMod}
+%\subsection{Data Analysis}
+%\subsubsection{Principal Score Model}\label{sec:psMod}
 
@@
 
-We estimated principal scores using logistic regression of observed $S$ on schol fixed effects and a set of baseline variables chosen with a combination of expert judgement and backwards selection based on the AIC, and evaluated model fit with binned residual plots \citep{gelman2006data}.
-The model was fit to data from students in the treatment group, for whom bottom-out status ($S$) was observed; hence, the same model fit was used for all three treatment contrasts.
+We estimated principal scores using logistic regression of observed $S$ on school fixed effects and a set of baseline variables chosen with a combination of expert judgement and backwards selection based on the AIC, and evaluated model fit with binned residual plots \citep{gelman2006data}.
+The model was fit to data from students in the ASSISTments group, for whom bottom-out status ($S$) was observed; hence, the same model fit was used for all three treatment contrasts.
 The model was fairly successful in distinguishing bottom-outers from non-bottom-outers---the AUC, evaluated using out-of-sample predictions in a 10-fold cross-validation, was roughly 0.795.
@@
 
-Figure \ref{fig:compare} shows principal effect estimates from GEEPERs compared to analogous estimates using Bayesian mixture modeling and PSW, alongside estimates of the overall average treatment effects, labeled ``ATE.''.
+Figure \ref{fig:compare} shows principal effect estimates from \textsc{geepers} compared to analogous estimates using Bayesian mixture modeling and \textsc{psw}, alongside estimates of the overall average treatment effects, labeled ``ATE.''.
 All three sets of estimates relied on the same principal score model described above, in Section \ref{sec:psMod}, though in the mixture model it was fit simultaneously with effect estimation using a Markov Chain Monte Carlo algorithm \citep{rstan}.
-Outcome regressions in GEEPERs and the mixture model included all of the covariates that were in the principal score model, in addition to students' pretest scores; also, the models substituted class fixed effects for school fixed effects.
+Outcome regressions in \textsc{geepers} and the mixture model included all of the covariates that were in the principal score model, in addition to students' pretest scores; also, the models substituted class fixed effects for school fixed effects.
 
@@
 
-The PSW estimators did not include any covariate outcome modeling. Standard errors for the PSW estimates were estimated using the bootstrap, with resampling done within schools.
+The \textsc{psw} estimators did not include any covariate outcome modeling. Standard errors for the \textsc{psw} estimates were estimated using the bootstrap, with resampling done within schools.
 
@@
 
-The GEEPERs results suggest little difference, if any, between the effects in the two principal strata.
+The \textsc{geepers} results suggest little difference, if any, between the effects in the two principal strata.
 Estimated effects compared to BAU or DragonBox were slightly lower for bottom-outers than for non-bottom-outers, and estimated effects compared to FH2T were higher for non-bottom-outers.
@@
 
-The other estimation strategies largely agreed with GEEPERs, and with each other. 
-This may suggest that the estimators' various assumptions---additivity of treatment assignment effect for GEEPERs and the mixture model, normal residuals for the mixture model, and principal ignorability for PSW---were all innocuous and may have held approximately.
+The other estimation strategies largely agreed with \textsc{geepers}, and with each other. 
+This may suggest that the estimators' various assumptions---additivity of treatment assignment effect for \textsc{geepers} and the mixture model, normal residuals for the mixture model, and principal ignorability for \textsc{psw}---were all innocuous and may have held approximately.
 Alternatively, violations of the different assumptions may have led to similar errors.
@@
 \section{Discussion}\label{sec:discussion}
-GEEPERs is a straightforward approach to principal effect estimation under strong monotonicity, built on widely-used regression models, which is more robust---though less precise---than alternative approaches under a wide array of scenarios.
-
-There is good reason to hope that extensions to GEEPERs may be straightforward.
-First, GEEPERs may be extended to situations in which there is two-way noncompliance, and only a weaker form of monotonicity (Assumption \ref{ass:sm}) holds.
+\textsc{geepers} is a straightforward approach to principal effect estimation under strong monotonicity, built on widely-used regression models, which is more robust---though less precise---than alternative approaches under a wide array of scenarios.
+
+There is good reason to hope that extensions to \textsc{geepers} may be straightforward.
+First, \textsc{geepers} may be extended to situations in which there is two-way noncompliance, and only a weaker form of monotonicity (Assumption \ref{ass:sm}) holds.
 For instance, in truncation by death problems \citep[e.g.][]{zhangRubin,ding2011}, $S_i=1$ if participant $i$ survives (or, more generally, if the outcome $Y_i$ is measured) and 0 otherwise, and interest is is in the principal effect for the ``always survive'' principal stratum in which $\sti=S_{ci}=1$.
@@
 Another direction of extension involved the principal score model \eqref{eq:pscore}.
-The performance of GEEPERs in the simulation study of Section \ref{sec:simulation} depended heavily on the factor $\alpha$, which controlled the extent to which covariates could predict $S$.
-That suggests that when covariates are high-dimensional, GEEPERs' performance in applications could be optimized with a high-dimensional semi- or non-parametric model.
+The performance of \textsc{geepers} in the simulation study of Section \ref{sec:simulation} depended heavily on the factor $\alpha$, which controlled the extent to which covariates could predict $S$.
+That suggests that when covariates are high-dimensional, \textsc{geepers}' performance in applications could be optimized with a high-dimensional semi- or non-parametric model.
 If so, several further questions emerge.
@@
 
-GEEPERs is a flexible, easily-implementable method for principal effect estimation for one-way noncompliance, with predictive covariates; extensions to a broader set of circumstances could be a boon to causal modeling.
+\textsc{geepers} is a flexible, easily-implementable method for principal effect estimation for one-way noncompliance, with predictive covariates; extensions to a broader set of circumstances could be a boon to causal modeling.
 
@@
 
-Here we will derive the sandwich variance-covariance matrix for the GEEPERs estimate without interactions between $\bx$ and either $Z$ or $\st$---i.e.
+Here we will derive the sandwich variance-covariance matrix for the \textsc{geepers} estimate without interactions between $\bx$ and either $Z$ or $\st$---i.e.
 with $\bm{\gamma_2}=\bm{\gamma_3}=\bm{\gamma_4}=0$ in the notation of \eqref{eq:interaction}---and estimating principal scores using a generalized linear model.
@@
 
+\clearpage
+\section{Simulation Results}
+
+\begin{figure}[!h]
+    \centering
+    \includegraphics[width=0.8\textwidth]{simFigs/alphaAUC.pdf}
+    \caption{Estimated AUC values from principal score models for each value of $\alpha$ in the simulation, with $n=500$.}
+    \label{fig:alphaAUC}
+\end{figure}
+
+%\begin{table}
+%\caption{Coverage of nominal 95\% confidence intervals}
+\subsection{Full Empirical 95\% Interval Coverage Results}
+\input{writeUps/coverageTabAppendix.tex}
+
+\clearpage
+\subsection{Full RMSE Results}
+\input{writeUps/rmseTabAppendix}
+
+%\end{table}
+
+\section{More Results from Application}
+
+\begin{figure}[!h]
+  \centering
+  \includegraphics{../figure/psModCoef.jpg}
+  \caption{Principal Score Model Coefficients}
+  \label{fig:psMod}
+\end{figure}
+
 \bibliographystylesupp{plainnat}

### 2022-12-19 writeUps/paper.tex

@@
 
-\begin{figure}[!h]
-  \centering
-  \includegraphics{../figure/psModCoef.jpg}
-  \caption{Principal Score Model Coefficients}
-  \label{fig:psMod}
-\end{figure}
+\subsection{Summary Statistics}
+
+\input{table1}
+
+\subsection{Regression Results}
+Regression estimates from principal score logit model and three outcome regressions for \textsc{geepers} estimates.
+Standard errors shown are nominal regression errors, not sandwich corrected.
+Fixed-effect estimates for school (PS-model) or classroom (outcome models) are omitted.
+
+\input{writeUps/outcomeRegAppendix}
 

### 2022-12-19 writeUps/paper.tex
submit v1

@@
 \usepackage{numdef}
-\usepackage{multibib}
+%\usepackage{multibib}
 
@@
 
-\newcites{supp}{Supplementary References}
+%\newcites{supp}{Supplementary References}
 
@@
 
-\author[Sales and Vanacore]{Adam C Sales}
-
-\author{Kirk Vanacore}
+\author[Sales et al.]{Adam C Sales}
+
+\author{Kirk P Vanacore}
+\author{Erin R Ottmar}
+
 \address{Worcester Polytechnic Institute,
@@
 
-In a separate, if related, line of research, \citet{jo2002}, \citet{ding2011}, \citet{jiangDing2021}, and others have shown that, under certain circumstances, researchers can use baseline covariates to identify and estimate principal effects without relying on principal ignorability.
+On the other hand, \citet{jo2002}, \citet{ding2011}, \citet{jiangDing2021}, and others have shown that, under certain circumstances, researchers can use baseline covariates to identify and estimate principal effects without relying on principal ignorability.
 This paper expands on those identification results by presenting a new set of principal effect estimators, and associated standard error estimates, based on a general estimating equations (or M-estimation or method-of-moments) \citep{stefanskiBoos}.
@@
 Let $S_i\in\{0,1\}$ be a measurement on subject $i$ taken subsequent to treatment assignment, so that $S$ may be affected by $Z$.
-Then $S$ itself has potential values $\sti$ and $\sci$, which $i$ would exibit if $Z_i=1$ or $Z_i=0$, respectively.
+Then $S$ itself has potential values $\sti$ and $\sci$, which $i$ would exhibit if $Z_i=1$ or $Z_i=0$, respectively.
 Assume ``strong monotonicity'' \citep[c.f.][]{dingLu} (also referred to as ``one-way noncompliance''):%\citealt{wing2017can}):
@@
 
-Under Assumptions \ref{ass:sm} and \ref{ass:rand}, conditional means $\mut0$ and $\mut1$ are non-parametrically identified.
+Under Assumptions \ref{ass:sm} and \ref{ass:rand}, conditional means $\mut0$ and $\mut1$ are nonparametrically  identified.
 In particular, the means of observed outcomes for subjects with $Z=1$ and $S=0$ or $S=1$ are unbiased for $\mut0$ and $\mut1$, respectively, since when $Z_i=1$, $Y_i=\yti$ and $S_i=\sti$.
@@
 \subsection{Simulation Design}
-The simulation study was conducted in \texttt{R} \citep{rcite} and Stan \citep{rstan}, and full replication code is available at \url{www.github.com/[Redacted]}.
+The simulation study was conducted in \texttt{R} \citep{rcite} and Stan \citep{rstan}, and full replication code is available at \url{https://github.com/adamSales/psGee}.
 \subsubsection{Data Generation}
@@
   \includegraphics[width=\textwidth,clip]{../simFigs/boxplots.pdf}
-  \caption{Violin and scatterplots of 500 simulation estimation errors for estimators described in Section~\ref{sec:simMods} under varying conditions.
+  \caption{Violin and scatter plots of 500 simulation estimation errors for estimators described in Section~\ref{sec:simMods} under varying conditions.
 For all plots, $n=500$; there were 500 replications for each set of conditions.
@@
 
-Figure \ref{fig:boxplots} shows violin plots, layered on top of jittered scatterplots, showing estimation error for \textsc{geepers}, the mixture estimator, and the \textsc{psw} estimator of $\eff1$ under various sets of conditions, and Table \ref{tab:coverage} shows empirical coverage of nominal 95\% confidence interval estimates under a somewhat wider set of circumstances when $\muc1=0.3$.
+Figure \ref{fig:boxplots} shows violin plots, layered on top of jittered scatter plots, showing estimation error for \textsc{geepers}, the mixture estimator, and the \textsc{psw} estimator of $\eff1$ under various sets of conditions, and Table \ref{tab:coverage} shows empirical coverage of nominal 95\% confidence interval estimates under a somewhat wider set of circumstances when $\muc1=0.3$.
 More complete results, including root-mean-squared-error estimates, are in the online appendix.
@@
 One way to disentangle the role of bottoming out in any treatment effect would be to estimate average treatment effects of being assigned to ASSISTments, relative to any of the other three, separately for subjects who would (if assigned to ASSISTments) bottom out frequently, and for subjects who would not.
-If the availability of bottom-out hints plays an important positive role in ASSISTments's efficacy, we might expect larger effects of condition on students who would bottom out frequently than for those who would not; if bottom-out hints are harmful, we might expect the opposite.
+If the availability of bottom-out hints plays an important positive role in ASSISTments' efficacy, we might expect larger effects of condition on students who would bottom out frequently than for those who would not; if bottom-out hints are harmful, we might expect the opposite.
 %If hints (that do not supply the correct answer) and error messages are helpful, but the ability to bottom out is harmful to learning, then we might expect to estimate a positive effect of being assigned to the feedback condition for students who do not bottom out, but a negative effect for students who do.
@@
 Students with missing pretest scores were dropped from the analysis; missing data in other covariates was imputed with a Random Forest algorithm \citep{missForest}.
+Summary statistics are displayed in the online appendix.
 
@@
 
-Due to privacy regulations, the data are not publicly available; however, interested analysts may access anonymized data by emailing the study's principal investigator at [REDACTED] and signing a privacy agreement.
-Replication code for the analysis is available at \url{www.github.com/[Redacted]}.
+Due to privacy regulations, the data are not publicly available; however, interested analysts may access anonymized data by emailing the study's principal investigator at erottmar@wpi.edu and signing a privacy agreement.
+Replication code for the analysis is available at \url{https://github.com/adamSales/psGee}.
 
@@
 That is, in roughly 80\% of bottom-outer/non-bottom-outer pairs of subjects, the bottom-outer will have a higher predicted probability from the model.
+Coefficient estimates are displayed in the online appendix.
 
@@
 Figure \ref{fig:compare} shows principal effect estimates from \textsc{geepers} compared to analogous estimates using Bayesian mixture modeling and \textsc{psw}, alongside estimates of the overall average treatment effects, labeled ``ATE.''.
-All three sets of estimates relied on the same principal score model described above, in Section \ref{sec:psMod}, though in the mixture model it was fit simultaneously with effect estimation using a Markov Chain Monte Carlo algorithm \citep{rstan}.
+%All three sets of estimates relied on the same principal score model described above, in Section \ref{sec:psMod}, though in the mixture model it was fit simultaneously with effect estimation using a Markov Chain Monte Carlo algorithm \citep{rstan}.
 Outcome regressions in \textsc{geepers} and the mixture model included all of the covariates that were in the principal score model, in addition to students' pretest scores; also, the models substituted class fixed effects for school fixed effects.
-
-The Bayesian mixture model assumed that outcomes were normally distributed, conditional on covariates (the same as in Section \ref{sec:outMod}) and principal stratum, with a residual standard deviation that varied between principal strata.
+Coefficient estimates are displayed in the online appendix.
+
+The Bayesian mixture model assumed that outcomes were normally distributed, conditional on covariates and principal stratum, with a residual standard deviation that varied between principal strata.
 This modeling assumption is necessarily false, since post-test scores were integers between 0 and 10.
 All model parameters with support in $\mathbb{R}$ were given standard normal priors, while standard deviation parameters were given half-standard normal priors.
+The principal score and outcome models were fit simultaneously using a Markov Chain Monte Carlo algorithm \citep{rstan}.
 
@@
 \section{Discussion}\label{sec:discussion}
-\textsc{geepers} is a straightforward approach to principal effect estimation under strong monotonicity, built on widely-used regression models, which is more robust---though less precise---than alternative approaches under a wide array of scenarios.
-
-There is good reason to hope that extensions to \textsc{geepers} may be straightforward.
-First, \textsc{geepers} may be extended to situations in which there is two-way noncompliance, and only a weaker form of monotonicity (Assumption \ref{ass:sm}) holds.
-For instance, in truncation by death problems \citep[e.g.][]{zhangRubin,ding2011}, $S_i=1$ if participant $i$ survives (or, more generally, if the outcome $Y_i$ is measured) and 0 otherwise, and interest is is in the principal effect for the ``always survive'' principal stratum in which $\sti=S_{ci}=1$.
+\textsc{geepers} is a straightforward approach to principal effect estimation under strong monotonicity, built on widely-used regression models, which is more robust---though sometimes less precise---than alternative approaches under a wide array of scenarios.
+
+There is good reason to hope that extensions to \textsc{geepers}, including cases in which $S$ takes more than two values, may be straightforward.
+For instance, in truncation by death problems \citep[e.g.][]{zhangRubin,ding2011} with weak monotonicity, $S_i=1$ if participant $i$ survives (or, more generally, if the outcome $Y_i$ is measured) and 0 otherwise, interest is typically in the principal effect for the ``always survive'' principal stratum in which $\sti=S_{ci}=1$.
 If, say, $\sti\ge S_{Ci}$, then every subject in the control condition with $S_i=1$ is in the always survive stratum, while those subjects in the treatment condition with $S_i=1$ are a mixture of the always survive and $\sti=1;\;S_{Ci}=0$ stratum.
-The scenario of truncation by death with (weak) monotonicity is broadly similar to the strong monotonicity, one-way noncompliance situation that this paper discussed.
+This scenario is broadly similar to the strong monotonicity, one-way noncompliance situation that this paper discussed.
 On the other hand, when no monotonicity assumption holds, both $\st$ and $S_C$ will have to be imputed for every subject; further research is necessary to determine the appropriate way to do so.
@@
 \bibliography{MOM}
+
+\clearpage
+\setcounter{page}{1}
+
+\begin{center}
+\large
+    \textbf{Online Appendices for ``GEEPERs: Principal Stratification using Principal Scores and Stacked Estimating Equations''}
+\end{center}
+
 \appendix
@@
 
-Replacing $\sti$ and $\pp$ in \eqref{eq:estEq0} with $\ri$, as in \eqref{eq:ri}, and replacing $\tilde{\Psi}_i$ with $\Psi_i=\begin{psmallmatrix} 1 & 0&1&0\\ 0&1&0&1\\0&0&1&0\\0&0&0&1\end{psmallmatrix}\tilde{\Psi}_i$ gives an equivalent set of estimating equations $\sum_{i=1}^{n_C}\Psi_i=\bm{0}$ or
+Replacing $\sti$ and $\pp$ in \eqref{eq:estEq0} with $\ri$, as in \eqref{eq:ri}, and replacing $\tilde{\Psi}_i$ with $\Psi_i=\begin{psmallmatrix} 1 & 0&1&0\\ 0&1&0&1\\0&0&1&0\\0&0&0&1\end{psmallmatrix}\tilde{\Psi}_i$ gives an equivalent set of estimating equations $\sum_{i=1}^{n_C}\Psi_i=\bm{0}$ with $\Psi_i=$
 \begin{equation}\label{eq:estEq1}
-\sum_{i=1}^{n_C}\begin{pmatrix}
+\begin{pmatrix}
     Y_i-\muc0-\ri(\muc1-\muc0)-Z_i(\mut0-\muc0)-Z_i\ri(\mut1-\mut0-\muc1+\muc0)\\
@@
 
-\end{pmatrix}=\begin{pmatrix} 0\\0\\0\\0\end{pmatrix}
+\end{pmatrix}
 \end{equation}
@@
   \item Estimate principal effects as:
-   \begin{equation}\label{eq:prinEffEst}
+   \begin{equation}\label{eq:prinEffEstApp}
   \begin{split}
@@
 \begin{equation*}
-  \EE[Y_T-Y_C|\st=0]=\EE[Y|Z=1,\st=0]-\EE[Y|Z=1,\st=0]=\beta_2+\bm{\gamma_3}'\EE[\bxy|\st=0]
+\begin{split}
+  \EE[Y_T-Y_C|\st=0]&\\
+  =&\EE[Y|Z=1,\st=0]-\EE[Y|Z=1,\st=0]\\
+  =&\beta_2+\bm{\gamma_3}'\EE[\bxy|\st=0]
+\end{split}
 \end{equation*}
@@
 \begin{equation*}
-  \EE[Y_T-Y_C|\st=1]=\EE[Y|Z=1,\st=1]-\EE[Y|Z=1,\st=1]=\beta_2+\beta_3+(\bm{\gamma_3}'+\bm{\gamma_4}')\EE[\bxy|\st=1]
+\begin{split}
+  \EE[Y_T-Y_C|\st=1]&\\
+  =&\EE[Y|Z=1,\st=1]-\EE[Y|Z=1,\st=1]\\
+  =&\beta_2+\beta_3+(\bm{\gamma_3}'+\bm{\gamma_4}')\EE[\bxy|\st=1]
+\end{split}
 \end{equation*}
@@
 \begin{align*}
-  \EE[Y]&=\EE[Y|Z=0]+\EE[Z]\left\{\EE[Y|Z=1]-\EE[Y|Z=0]\right\}\\
-  &=\beta_0+\beta_1\EE[R]+\beta_2\EE[Z]+\beta_3\EE[ZR]+\bm{\gamma_1}'\EE[\bxy]+\bm{\gamma_2}'\EE[\bxy R]+\bm{\gamma_3}'\EE[\bxy Z]+\bm{\gamma_4}\EE[\bxy RZ]
+  \EE[Y]=&\EE[Y|Z=0]+\EE[Z]\left\{\EE[Y|Z=1]-\EE[Y|Z=0]\right\}\\
+  =&\beta_0+\beta_1\EE[R]+\beta_2\EE[Z]+\beta_3\EE[ZR]+\bm{\gamma_1}'\EE[\bxy]+\bm{\gamma_2}'\EE[\bxy R]\\
+  &+\bm{\gamma_3}'\EE[\bxy Z]+\bm{\gamma_4}\EE[\bxy RZ]
 \end{align*}
@@
 
-Following \citetsupp[][p. 373]{carroll2006measurement}, we %separate the estimating equations $\Psi$ into $\phi$ and $\psi$, estimating equations for models \eqref{
+Following \citet[][p. 373]{carroll2006measurement}, we %separate the estimating equations $\Psi$ into $\phi$ and $\psi$, estimating equations for models \eqref{
 can decompose the matrices into diagonal elements 
@@
 
-\clearpage
 \section{Simulation Results}
-
-\begin{figure}[!h]
-    \centering
+\subsection{Plot of  AUC versus $\alpha$}
+
+    \begin{center}
     \includegraphics[width=0.8\textwidth]{simFigs/alphaAUC.pdf}
-    \caption{Estimated AUC values from principal score models for each value of $\alpha$ in the simulation, with $n=500$.}
-    \label{fig:alphaAUC}
-\end{figure}
-
+\end{center}
 %\begin{table}
@@
 \subsection{Summary Statistics}
-
+\small
 \input{table1}
@@
 
+\small
 \input{writeUps/outcomeRegAppendix}
 
-\bibliographystylesupp{plainnat}
-\bibliographysupp{MOM}
+% \bibliographystylesupp{plainnat}
+% \bibliographysupp{MOM}
 

### 2023-11-03 writeUps/outcomeRegAppendix.tex
add class FE, replicable appendix tables

@@
-\begin{table}
+\usepackage{longtable}
+
 \begin{center}
-\begin{tabular}{l c c c c}
+\begin{longtable}{l c c c c}
 \hline
@@
 \hline
-(Intercept)                                        & $-2.14$       & $1.73$      & $2.83^{***}$ & $2.51^{**}$  \\
-                                                   & $(1.26)$      & $(0.92)$    & $(0.72)$     & $(0.92)$     \\
-Scale.Score5                                       & $-0.59^{**}$  & $0.21$      & $0.45^{***}$ & $0.05$       \\
-                                                   & $(0.19)$      & $(0.13)$    & $(0.10)$     & $(0.13)$     \\
-ESOL1                                              & $-0.83$       & $0.09$      & $-0.46$      & $-0.49$      \\
-                                                   & $(0.64)$      & $(0.36)$    & $(0.27)$     & $(0.35)$     \\
-IEP1                                               & $-1.34^{**}$  & $0.15$      & $0.23$       & $0.10$       \\
-                                                   & $(0.46)$      & $(0.29)$    & $(0.25)$     & $(0.28)$     \\
-pre.total\_time\_on\_tasks                         & $-0.45^{**}$  & $-0.13$     & $0.08$       & $-0.00$      \\
-                                                   & $(0.15)$      & $(0.09)$    & $(0.07)$     & $(0.10)$     \\
-pre\_MSE\_total\_score                             & $-0.18$       & $0.20^{*}$  & $0.12$       & $0.35^{***}$ \\
-                                                   & $(0.13)$      & $(0.08)$    & $(0.07)$     & $(0.08)$     \\
-fullYear5TRUE                                      & $1.40^{*}$    & $-0.13$     & $-0.10$      & $0.06$       \\
-                                                   & $(0.58)$      & $(0.33)$    & $(0.29)$     & $(0.34)$     \\
-pre\_MA\_total\_scoreNATRUE                        & $-3.20^{***}$ &             &              &              \\
-                                                   & $(0.71)$      &             &              &              \\
-GenderM                                            & $0.17$        & $-0.19$     & $-0.34^{**}$ & $-0.27$      \\
-                                                   & $(0.25)$      & $(0.15)$    & $(0.13)$     & $(0.16)$     \\
-raceEthHispanic/Latino                             & $-0.56$       & $0.28$      & $0.33$       & $0.23$       \\
-                                                   & $(0.46)$      & $(0.29)$    & $(0.21)$     & $(0.28)$     \\
-raceEthAsian                                       & $-0.43$       & $0.66^{*}$  & $0.62^{**}$  & $0.82^{**}$  \\
-                                                   & $(0.42)$      & $(0.28)$    & $(0.22)$     & $(0.30)$     \\
-raceEthOther                                       & $-0.13$       & $0.45$      & $0.21$       & $0.10$       \\
-                                                   & $(0.49)$      & $(0.29)$    & $(0.23)$     & $(0.30)$     \\
-GIFTED1                                            & $-0.24$       & $0.66^{**}$ & $0.43^{*}$   & $0.34$       \\
-                                                   & $(0.37)$      & $(0.24)$    & $(0.19)$     & $(0.26)$     \\
-poly(pre\_PS\_tasks\_total\_score, 2, raw = TRUE)1 & $0.01$        &             &              &              \\
-                                                   & $(0.19)$      &             &              &              \\
-poly(pre\_PS\_tasks\_total\_score, 2, raw = TRUE)2 & $-0.41^{**}$  &             &              &              \\
-                                                   & $(0.13)$      &             &              &              \\
-Z                                                  &               & $0.29$      & $-0.17$      & $-0.20$      \\
-                                                   &               & $(0.28)$    & $(0.22)$     & $(0.27)$     \\
-Sp                                                 &               & $0.03$      & $-0.22$      & $0.04$       \\
-                                                   &               & $(0.52)$    & $(0.41)$     & $(0.52)$     \\
-pre\_PS\_tasks\_total\_score                       &               & $0.34^{**}$ & $0.37^{***}$ & $0.23$       \\
-                                                   &               & $(0.12)$    & $(0.09)$     & $(0.12)$     \\
-pre.total\_math\_score                             &               & $0.14^{**}$ & $0.24^{***}$ & $0.15^{**}$  \\
-                                                   &               & $(0.05)$    & $(0.04)$     & $(0.05)$     \\
-pre.total\_math\_scoreNATRUE                       &               & $-0.35$     & $0.18$       & $-0.44$      \\
-                                                   &               & $(0.45)$    & $(0.36)$     & $(0.43)$     \\
-Z:Sp                                               &               & $-0.14$     & $0.17$       & $-0.17$      \\
-                                                   &               & $(0.49)$    & $(0.39)$     & $(0.49)$     \\
+\endfirsthead
 \hline
-AIC                                                & $447.55$      & $$          & $$           & $$           \\
-BIC                                                & $539.47$      & $$          & $$           & $$           \\
-Log Likelihood                                     & $-200.78$     & $$          & $$           & $$           \\
-Deviance                                           & $401.55$      & $$          & $$           & $$           \\
-Num. obs.                                          & $402$         & $787$       & $1193$       & $771$        \\
-R$^2$                                              & $$            & $0.60$      & $0.60$       & $0.63$       \\
-Adj. R$^2$                                         & $$            & $0.52$      & $0.55$       & $0.54$       \\
+ & psModel & BAU & FH2T & DragonBox \\
 \hline
-\multicolumn{5}{l}{\scriptsize{$^{***}p<0.001$; $^{**}p<0.01$; $^{*}p<0.05$}}
-\end{tabular}
+\endhead
+\hline
+\endfoot
+\hline
+\multicolumn{5}{l}{\scriptsize{$^{***}p<0.001$; $^{**}p<0.01$; $^{*}p<0.05$}}\\
 \caption{Statistical models}
 \label{table:coefficients}
+\endlastfoot \\
+(Intercept)                                        & $-7.75$      & $5.64^{*}$   & $7.43^{***}$ & $3.65$       \\
+                                                   & $(5.04)$     & $(2.47)$     & $(1.82)$     & $(2.35)$     \\
+GenderM                                            & $0.21$       & $-0.26$      & $-0.30^{*}$  & $-0.16$      \\
+                                                   & $(0.28)$     & $(0.15)$     & $(0.13)$     & $(0.16)$     \\
+raceEthHispanic/Latino                             & $-0.54$      & $-0.02$      & $0.43^{*}$   & $0.41$       \\
+                                                   & $(0.50)$     & $(0.28)$     & $(0.21)$     & $(0.28)$     \\
+raceEthAsian                                       & $-0.36$      & $0.54^{*}$   & $0.79^{***}$ & $0.63^{*}$   \\
+                                                   & $(0.45)$     & $(0.25)$     & $(0.20)$     & $(0.26)$     \\
+raceEthOther                                       & $-0.09$      & $0.37$       & $0.35$       & $-0.06$      \\
+                                                   & $(0.54)$     & $(0.29)$     & $(0.23)$     & $(0.30)$     \\
+Performance.Level5Developing Learner               & $-0.79$      & $-0.62$      & $-0.52$      & $-0.14$      \\
+                                                   & $(0.95)$     & $(0.53)$     & $(0.39)$     & $(0.48)$     \\
+Performance.Level5Distinguished Learner            & $-1.08$      & $-0.41$      & $0.11$       & $0.34$       \\
+                                                   & $(1.23)$     & $(0.68)$     & $(0.54)$     & $(0.67)$     \\
+Performance.Level5Proficient Learner               & $-1.08$      & $-0.27$      & $-0.20$      & $-0.09$      \\
+                                                   & $(1.04)$     & $(0.58)$     & $(0.43)$     & $(0.53)$     \\
+EIP1                                               & $0.50$       & $0.26$       & $0.09$       & $0.33$       \\
+                                                   & $(0.73)$     & $(0.40)$     & $(0.31)$     & $(0.39)$     \\
+Scale.Score5                                       & $-0.57$      & $0.48^{**}$  & $0.47^{**}$  & $0.28$       \\
+                                                   & $(0.35)$     & $(0.18)$     & $(0.15)$     & $(0.19)$     \\
+ESOL1                                              & $-1.01$      & $0.40$       & $-0.40$      & $-0.38$      \\
+                                                   & $(0.70)$     & $(0.35)$     & $(0.26)$     & $(0.34)$     \\
+GIFTED1                                            & $-0.26$      & $0.79^{***}$ & $0.62^{***}$ & $0.78^{**}$  \\
+                                                   & $(0.41)$     & $(0.23)$     & $(0.18)$     & $(0.25)$     \\
+IEP1                                               & $-1.46^{**}$ & $0.18$       & $0.24$       & $0.16$       \\
+                                                   & $(0.51)$     & $(0.27)$     & $(0.24)$     & $(0.27)$     \\
+IST1                                               & $-0.55$      & $-0.06$      & $0.14$       & $-0.46$      \\
+                                                   & $(0.67)$     & $(0.36)$     & $(0.29)$     & $(0.38)$     \\
+SECTION5041                                        & $0.42$       & $0.73$       & $0.58$       & $0.51$       \\
+                                                   & $(0.92)$     & $(0.47)$     & $(0.39)$     & $(0.48)$     \\
+SST1                                               & $-0.61$      & $-0.40$      & $0.14$       & $-0.77$      \\
+                                                   & $(1.11)$     & $(0.58)$     & $(0.41)$     & $(0.51)$     \\
+AbsentDays5                                        & $0.25$       & $-0.12$      & $-0.02$      & $-0.12$      \\
+                                                   & $(0.27)$     & $(0.15)$     & $(0.12)$     & $(0.16)$     \\
+UnexcusedDays5                                     & $-0.48$      & $0.24$       & $-0.00$      & $0.19$       \\
+                                                   & $(0.40)$     & $(0.21)$     & $(0.18)$     & $(0.23)$     \\
+AbsentDays6                                        & $-0.30$      & $0.04$       & $0.12$       & $0.04$       \\
+                                                   & $(0.31)$     & $(0.17)$     & $(0.14)$     & $(0.17)$     \\
+UnexcusedDays6                                     & $0.57$       &              &              &              \\
+                                                   & $(0.32)$     &              &              &              \\
+pre.total\_math\_score                             & $-0.05$      &              &              &              \\
+                                                   & $(0.16)$     &              &              &              \\
+pre.sub\_P\_score                                  & $-0.04$      & $-0.32^{*}$  & $-0.35^{**}$ & $-0.25$      \\
+                                                   & $(0.24)$     & $(0.14)$     & $(0.11)$     & $(0.14)$     \\
+pre.sub\_F\_score                                  & $0.12$       & $-0.38^{**}$ & $-0.28^{*}$  & $-0.32^{*}$  \\
+                                                   & $(0.24)$     & $(0.14)$     & $(0.11)$     & $(0.14)$     \\
+pre.math\_completed\_num                           & $0.73$       & $-0.15$      & $-0.06$      & $-0.10$      \\
+                                                   & $(0.46)$     & $(0.19)$     & $(0.12)$     & $(0.16)$     \\
+pre.total\_time\_on\_tasks                         & $-0.52^{**}$ & $-0.18$      & $0.11$       & $-0.07$      \\
+                                                   & $(0.19)$     & $(0.10)$     & $(0.08)$     & $(0.10)$     \\
+pre\_MA\_total\_score                              & $-0.01$      &              &              &              \\
+                                                   & $(0.03)$     &              &              &              \\
+pre\_MSE\_total\_score                             & $-0.15$      &              &              &              \\
+                                                   & $(0.16)$     &              &              &              \\
+pre\_PS\_part2E\_score                             & $0.17$       & $0.04$       & $0.06$       & $0.12$       \\
+                                                   & $(0.20)$     & $(0.12)$     & $(0.09)$     & $(0.12)$     \\
+pre\_PS\_part2NE\_score                            & $0.04$       & $0.24^{*}$   & $0.20^{*}$   & $0.04$       \\
+                                                   & $(0.19)$     & $(0.11)$     & $(0.09)$     & $(0.11)$     \\
+pre\_PS\_completed\_num                            & $-0.11$      & $-0.11^{*}$  & $-0.09^{*}$  & $-0.04$      \\
+                                                   & $(0.09)$     & $(0.05)$     & $(0.04)$     & $(0.05)$     \\
+pre\_PS\_total\_RT\_sec                            & $0.07$       & $0.17$       & $0.02$       & $0.21$       \\
+                                                   & $(0.23)$     & $(0.13)$     & $(0.10)$     & $(0.13)$     \\
+fullYear5TRUE                                      & $1.54^{*}$   & $-0.41$      & $-0.08$      & $-0.09$      \\
+                                                   & $(0.62)$     & $(0.32)$     & $(0.28)$     & $(0.34)$     \\
+fullYear6TRUE                                      & $0.65$       & $0.17$       & $-0.15$      & $0.50$       \\
+                                                   & $(0.84)$     & $(0.42)$     & $(0.32)$     & $(0.42)$     \\
+noUnexcused5TRUE                                   & $-0.23$      & $0.05$       & $-0.21$      & $-0.27$      \\
+                                                   & $(0.52)$     & $(0.28)$     & $(0.23)$     & $(0.30)$     \\
+Performance.Level5NATRUE                           & $0.15$       & $-0.22$      & $-0.25$      & $-0.05$      \\
+                                                   & $(0.61)$     & $(0.33)$     & $(0.24)$     & $(0.32)$     \\
+AbsentDays6NATRUE                                  & $-0.24$      & $-0.72$      & $-0.18$      & $0.11$       \\
+                                                   & $(0.80)$     & $(0.45)$     & $(0.35)$     & $(0.45)$     \\
+pre.total\_math\_scoreNATRUE                       & $3.23$       & $-1.66$      & $-0.99$      & $-0.77$      \\
+                                                   & $(3.71)$     & $(1.12)$     & $(0.74)$     & $(1.10)$     \\
+pre\_MA\_total\_scoreNATRUE                        & $-1.58$      & $-1.84^{*}$  & $-1.05$      & $-1.22$      \\
+                                                   & $(1.59)$     & $(0.87)$     & $(0.56)$     & $(0.70)$     \\
+poly(pre\_PS\_tasks\_total\_score, 2, raw = TRUE)1 & $-0.18$      & $0.12$       & $0.20$       & $0.00$       \\
+                                                   & $(0.32)$     & $(0.18)$     & $(0.14)$     & $(0.18)$     \\
+poly(pre\_PS\_tasks\_total\_score, 2, raw = TRUE)2 & $-0.41^{**}$ & $-0.05$      & $-0.07$      & $0.04$       \\
+                                                   & $(0.14)$     & $(0.08)$     & $(0.06)$     & $(0.07)$     \\
+Z                                                  &              & $0.33$       & $-0.17$      & $-0.33$      \\
+                                                   &              & $(0.27)$     & $(0.23)$     & $(0.27)$     \\
+Sp                                                 &              & $0.08$       & $-0.12$      & $0.04$       \\
+                                                   &              & $(0.53)$     & $(0.45)$     & $(0.54)$     \\
+ns(UnexcusedDays6, 3)1                             &              & $-1.17^{**}$ & $-1.03^{**}$ & $-0.95^{*}$  \\
+                                                   &              & $(0.45)$     & $(0.37)$     & $(0.44)$     \\
+ns(UnexcusedDays6, 3)2                             &              & $-0.57$      & $-0.29$      & $-0.70$      \\
+                                                   &              & $(0.58)$     & $(0.51)$     & $(0.60)$     \\
+ns(UnexcusedDays6, 3)3                             &              & $-0.70$      & $0.14$       & $-0.08$      \\
+                                                   &              & $(0.60)$     & $(0.59)$     & $(0.64)$     \\
+ns(pre\_MA\_total\_score, 3)1                      &              & $-0.16$      & $-0.42$      & $-0.15$      \\
+                                                   &              & $(0.37)$     & $(0.29)$     & $(0.39)$     \\
+ns(pre\_MA\_total\_score, 3)2                      &              & $-0.34$      & $-0.90$      & $1.20$       \\
+                                                   &              & $(1.09)$     & $(0.85)$     & $(1.09)$     \\
+ns(pre\_MA\_total\_score, 3)3                      &              & $-0.05$      & $0.02$       & $0.38$       \\
+                                                   &              & $(0.49)$     & $(0.41)$     & $(0.54)$     \\
+pre\_MSE\_total\_score $<$ -2TRUE                  &              & $0.93$       & $-0.79$      & $-0.21$      \\
+                                                   &              & $(0.72)$     & $(0.58)$     & $(0.80)$     \\
+ns(pre\_MSE\_total\_score, 3)1                     &              & $1.34$       & $-0.10$      & $0.96$       \\
+                                                   &              & $(0.79)$     & $(0.65)$     & $(0.82)$     \\
+ns(pre\_MSE\_total\_score, 3)2                     &              & $3.56$       & $-1.41$      & $2.29$       \\
+                                                   &              & $(2.57)$     & $(2.20)$     & $(2.68)$     \\
+ns(pre\_MSE\_total\_score, 3)3                     &              & $1.13$       & $-0.03$      & $1.00$       \\
+                                                   &              & $(0.58)$     & $(0.45)$     & $(0.57)$     \\
+ns(pre.total\_math\_score, 4)1                     &              & $1.04$       & $1.71^{***}$ & $0.84$       \\
+                                                   &              & $(0.61)$     & $(0.47)$     & $(0.63)$     \\
+ns(pre.total\_math\_score, 4)2                     &              & $2.62^{***}$ & $2.57^{***}$ & $2.77^{***}$ \\
+                                                   &              & $(0.69)$     & $(0.53)$     & $(0.71)$     \\
+ns(pre.total\_math\_score, 4)3                     &              & $4.54^{***}$ & $5.68^{***}$ & $5.28^{***}$ \\
+                                                   &              & $(1.29)$     & $(1.04)$     & $(1.39)$     \\
+ns(pre.total\_math\_score, 4)4                     &              & $3.47^{***}$ & $4.35^{***}$ & $3.19^{***}$ \\
+                                                   &              & $(0.84)$     & $(0.64)$     & $(0.82)$     \\
+Z:Sp                                               &              & $-0.16$      & $0.14$       & $-0.06$      \\
+                                                   &              & $(0.48)$     & $(0.40)$     & $(0.49)$     \\
+\hline
+AIC                                                & $484.78$     & $$           & $$           & $$           \\
+BIC                                                & $676.61$     & $$           & $$           & $$           \\
+Log Likelihood                                     & $-194.39$    & $$           & $$           & $$           \\
+Deviance                                           & $388.78$     & $$           & $$           & $$           \\
+Num. obs.                                          & $402$        & $787$        & $1193$       & $771$        \\
+R$^2$                                              & $$           & $0.54$       & $0.57$       & $0.55$       \\
+Adj. R$^2$                                         & $$           & $0.50$       & $0.55$       & $0.51$       \\
+\end{longtable}
 \end{center}
-\end{table}

### 2023-11-03 code/fh2t/analysisGEE.r

@@ -144,6 +144,7 @@ estimates1all <-
            est(getDat(alt),psMod=psModAll,
                covFormY=update(formula(psModAll)[-2],
                                .~.
+                               -SchIDPre+ClaIDPre
                                -UnexcusedDays6+ns(UnexcusedDays6,3)
                                -pre_MA_total_score+ns(pre_MA_total_score,3)
                                -pre_MSE_total_score+I(pre_MSE_total_score< -2)+ns(pre_MSE_total_score,3)
@@ -174,6 +175,7 @@ estimates1rest <-
            est(getDat(alt),psMod=psModRest,
                covFormY=update(formula(psModAll)[-2],
                                .~.
+                               -SchIDPre+ClaIDPre
                                -UnexcusedDays6+ns(UnexcusedDays6,3)
                                -pre_MA_total_score+ns(pre_MA_total_score,3)
                                -pre_MSE_total_score+I(pre_MSE_total_score< -2)+ns(pre_MSE_total_score,3)

### 2023-11-03 code/fh2t/data.r

@@ -106,7 +106,7 @@ names(imp$OOBerror) <- names(impDat)
 
 r2imp=1-imp$OOBerror[mse]/sapply(impDat[,mse],var,na.rm=TRUE)
 
-save(imp,mse,r2imp,file='data/imputations.RData')
+save(imp,mse,r2imp,impDat,file='data/imputations.RData')
 
 #names(imp$ximp) <- paste0(names(imp$ximp),'Imp')
 

### 2023-11-03 code/fh2t/estimates.r

@@ -66,6 +66,7 @@ estimates1all <-
            est(getDat(alt),psMod=psModAll,
                covFormY=update(formula(psModAll)[-2],
                                .~.
+                               -SchIDPre+ClaIDPre
                                -UnexcusedDays6+ns(UnexcusedDays6,3)
                                -pre_MA_total_score+ns(pre_MA_total_score,3)
                                -pre_MSE_total_score+I(pre_MSE_total_score< -2)+ns(pre_MSE_total_score,3)
@@ -81,6 +82,7 @@ estimates1rest <-
            est(getDat(alt),psMod=psModRest,
                covFormY=update(formula(psModAll)[-2],
                                .~.
+                               -SchIDPre+ClaIDPre
                                -UnexcusedDays6+ns(UnexcusedDays6,3)
                                -pre_MA_total_score+ns(pre_MA_total_score,3)
                                -pre_MSE_total_score+I(pre_MSE_total_score< -2)+ns(pre_MSE_total_score,3)
@@ -89,6 +91,8 @@ estimates1rest <-
                                )
                )
          )
+
+save(atesAll,estimates3,estimates1all,estimates1rest,file='results/geeResults.RData')
 
 #######################################################################
 ### estimate principal effects with PSW

### 2023-11-03 code/fh2t/tables.r

@@ -1,86 +1,137 @@
 options(knitr.kable.NA = '')
 
 print(load('data/psdat.RData'))
-print(load('data/imputation.RData'))
+print(load('data/imputations.RData'))
 
 tab1dat <-
-  psdat%>%
-  #filter(Z%in%c('ASSISTments','BAU',''))%>%
-  select(
-    pretest , Scale.Score5 , MALE , race ,
-    virtual , EIP , IEP , ESOL , GIFTED , pre.avg_time_on_tasks ,
-    pre_MA_total_score , pre_negative_reaction_score ,
-    pre_numerical_confindence_score,anyBottom,post.total_math_score,Z)%>%
-  mutate(pre.avg_time_on_tasks=log(pre.avg_time_on_tasks),
-         MALE=ifelse(MALE==1,'Male','Female'),
-         virtual=ifelse(virtual==1,'Remote','In-Person'),
-         anyBottom=anyBottom*1)
-
-
-oob=as.data.frame(rbind(imp$OOBerror))%>%  select(
+  impDat[rownames(psdat),c(intersect(names(impDat),names(psdat)),'nbo')]%>%
+  mutate(
+    Condition=psdat$trt,
+    `Bottom-Outer`=as.factor(ifelse(is.na(psdat$S),0,psdat$S*1)),
+    Posttest=psdat$Y,
+    Gender=ifelse(Gender=="M",'Male','Female'),
+    virtual=ifelse(virtual==1,'Remote','In-Person'),
+    across(where(~length(dim(.))>0),~.[,1]))%>%
+  select(-contains("ID",ignore.case=FALSE))%>%
+  rename(
+    `# Bottom-Out`=nbo,
     `Has EIP`=EIP,
     `Has IEP`=IEP,
-    ESOL,
+    Modality=virtual,
     `Gifted`=GIFTED,
-        Pretest=pretest,
+    Pretest=pre.total_math_score,
     `Grade 5 Stand. Test`=Scale.Score5,
-    Gender=MALE,
-    `Race/Ethnicity`=race,
-    `log(Pretest ToT)`=pre.avg_time_on_tasks,
+    `Grade 5 Perf. Lev.`=Performance.Level5,
+    `Race/Ethnicity`=raceEth,
+    `log(Pretest ToT)`=pre.total_time_on_tasks,
     `Math Anxiety`=pre_MA_total_score,
-    `Math Neg. React.`=pre_negative_reaction_score,
[…]

### 2023-11-03 code/fh2t/data.r
part 2--had forgotten to save!

@@ -106,6 +106,8 @@ names(imp$OOBerror) <- names(impDat)
 
 r2imp=1-imp$OOBerror[mse]/sapply(impDat[,mse],var,na.rm=TRUE)
 
+impDat$nbo <- dat$nbo
+
 save(imp,mse,r2imp,impDat,file='data/imputations.RData')
 
 #names(imp$ximp) <- paste0(names(imp$ximp),'Imp')

### 2023-11-03 code/fh2t/replicateFH2T.r

@@ -1,12 +1,15 @@
 library(tidyverse)
 library(arm)
 library(randomForest)
+library(missForest)
 library(estimatr)
-select <- dplyr::select
 library(splines)
 library(rstan)
 library(kableExtra)
 library(texreg)
+library(tableone)
+library(xtable)
+select <- dplyr::select
 
 
 source('code/regression.r')
@@ -67,3 +70,5 @@ source('code/fh2t/plots.r')
 ############################################################
 ### make tables (for appendix)
 ############################################################
+
+source('code/fh2t/tables.r')

### 2023-11-03 code/fh2t/tables.r

@@ -137,13 +137,74 @@ xtable(tab1oth,
 load('results/geeResults.RData')
 
 
-#sink('writeUps/outcomeRegAppendix.tex')
-texreg(
+                                        #sink('writeUps/outcomeRegAppendix.tex')
+
+
+mods <-
   list(
     psModel=estimates1all$BAU$psMod,
     BAU=estimates1all$BAU$outMod,
     FH2T=estimates1all$FH2T$outMod,
-    DragonBox=estimates1all$Dragon$outMod),
-  file='writeUps/outcomeRegAppendix.tex',
-  longtable=TRUE,
-  omit.coef = c('SchIDPre|ClaIDPre'))
+    DragonBox=estimates1all$Dragon$outMod)
+
+cnames <- mods%>%
+  lapply(coef)%>%
+  lapply(names)%>%
+  do.call('c',.)%>%
+  unique()
+
+cnames <- cnames[-grep("ID",cnames)]
+cnames <- c('(Intercept)','Z','Sp','Z:Sp')%>%c(.,setdiff(cnames,.))
+
+
+ccm <- rep(NA,length(cnames))
+ccm[endsWith(cnames,'TRUE')] <-
+  substr(cnames[endsWith(cnames,'TRUE')],1,nchar(cnames[endsWith(cnames,'TRUE')])-4)
+ccm[endsWith(cnames,'1')] <-
+  substr(cnames[endsWith(cnames,'1')],1,nchar(cnames[endsWith(cnames,'1')])-1)
+names(ccm) <- cnames
+ccm <- as.list(ccm)
+
+regTab <-
+  texreg(mods,
+         longtable=TRUE,
+         custom.coef.map = ccm,
+         custom.gof.rows=
+           list(
+             "School Effs"=lapply(mods,\(x) any(grepl("SchIDPre",names(coef(x))))),
+             "Class Effs"=lapply(mods,\(x) any(grepl("ClaIDPre",names(coef(x)))))
+           ),
+         caption="Coefficient estimates from main PS and outcome models; school and classroom fixed effects are omitted",
+         label="tab:regTab"
+)
+#  omit.coef = c('SchIDPre|ClaIDPre'))
+
+covnames <-    c( `Has EIP`="EIP",
+    `Has IEP`="IEP",
+    Modality="virtual",
+    `Gifted`="GIFTED",
+    Pretest="pre.total_math_score",
[…]

### 2024-12-03 writeUps/paper.tex
Updates from Overleaf

@@
-\documentclass{statsoc} % use larger type; default would be 10pt
+\documentclass[]{interact}
+
+\RequirePackage{amsthm,amsmath,amsfonts,amssymb}
+%\RequirePackage[authoryear]{natbib}
+\usepackage{longtable}
+\usepackage[notablist,tablesfirst]{endfloat}% To `separate' figures and tables from text if required
+
+\usepackage{epstopdf}% To incorporate .eps illustrations using PDFLaTeX, etc.
+\usepackage[caption=false]{subfig}% Support for small, `sub' figures and tables
+\usepackage[doublespacing]{setspace}% To produce a `double spaced' document if required
+%\setlength\parindent{24pt}% To increase paragraph indentation when line spacing is doubled
+
+\usepackage[longnamesfirst,sort]{natbib}% Citation support using natbib.sty
+\bibpunct[, ]{(}{)}{;}{a}{,}{,}% Citation support using natbib.sty
+\renewcommand\bibfont{\fontsize{10}{12}\selectfont}% To set the list of references in 10 point font using natbib.sty
+
+\theoremstyle{plain}% Theorem-like structures provided by asthma.sty
+\newtheorem{theorem}{Theorem}[section]
+\newtheorem{lemma}[theorem]{Lemma}
+\newtheorem{corollary}[theorem]{Corollary}
+\newtheorem{prop}[theorem]{Proposition}
+
+\theoremstyle{definition}
+\newtheorem{definition}[theorem]{Definition}
+\newtheorem{example}[theorem]{Example}
+
+\theoremstyle{remark}
+\newtheorem{remark}{Remark}
+\newtheorem{notation}{Notation}
 
 \usepackage[utf8]{inputenc} % set input encoding (not needed with XeLaTeX)
-\usepackage{amsmath,amssymb}
 \usepackage{bm}
-\usepackage{natbib}
 \usepackage[hidelinks]{hyperref}
@@
 \usepackage{mathtools}
-\usepackage{fullpage}
+%\usepackage{fullpage}
 \usepackage{makecell}
@@
 %\usepackage{multibib}
+\usepackage{rotating}
+
+%\startlocaldefs
 
@@
 
+%\endlocaldefs
 %\newcites{supp}{Supplementary References}
-
-%%% TO DO
-%% Add detail to application section
-%% fill in missing citations
-%% Appendix
-%%   clean up/fix notation for math
-%%   add in full simulation results
-%%   small interaction simulation
-%%   look for what else I promised and either put it in or delete promise
-%% proofread
-
-\title[\textsc{geepers}]{\textsc{geepers}: Principal Stratification using Principal Scores and Stacked Estimating Equations}
-
-\author[Sales et al.]{Adam C Sales}
-
-\author{Kirk P Vanacore}
-\author{Erin R Ottmar}
-
-\address{Worcester Polytechnic Institute,
-Worcester, MA,
-USA.}
-\email{asales@wpi.edu}
-
 \begin{document}
+
+%\begin{frontmatter}
+
+\title{GEEPERs: Principal Stratification using Principal Scores and Stacked Estimating Equations}
+%\runtitle{\textsc{geepers}}
+%\maketitle
+
+%\begin{aug}
+\author{
+\name{Adam C. Sales,\thanks{CONTACT Adam C. Sales. Email: asales@wpi.edu} Kirk P. Vanacore, and Erin R. Ottmar}
+\affil{Worcester Polytechnic Institute, Worcester, MA, USA}
+}
+
 \maketitle
@@
 Principal stratification is a framework for making sense of causal effects conditioned on variables that themselves may have been affected by treatment.
-Most principal stratification estimators rely on strong structural or modeling assumptions, and many require advanced statistical training to fit and to check.
+For instance, one component of an educational computer application is the availability of “bottom-out” hints that provide the answer.
+In evaluating a recent experimental evaluation against alternative programs without bottom-out hints, researchers may be interested in estimating separate average treatment effects for students who, if given the opportunity, would request bottom-out hints frequently, and for students who would not.
+Most principal stratification estimators rely on strong structural or modeling assumptions, and many require advanced statistical training to fit and check.
 In this paper, we introduce a new M-estimation principal effect estimator for one-way noncompliance based on a binary indicator.
 Estimates may be computed using conventional regressions (though the standard errors require a specialized sandwich formula) and do not rely on distributional assumptions.
-We illustrate the new technique in an analysis of student log data from a recent educational technology field experiment.
+We present a simulation study that shows that the novel method is more robust than popular alternatives and illustrate the method in an analysis of data on bottom-out hint requests.
 \end{abstract}
-\keywords{Causal Inference; Principal Stratification; Educational Technology}
+\begin{keywords}
+Causal Inference; Principal Stratification; Educational Technology
+\end{keywords}
+%\end{frontmatter}
+
+\clearpage
 
@@
 
-Randomized experiments easily admit subgroup analysis---estimation of treatment effects for a subset of the study population---when those subgroups are defined at baseline, and not themselves a function of treatment assignment.
+%Randomized experiments easily admit subgroup analysis---estimation of treatment effects for a subset of the study population---when those subgroups are defined at baseline, and not themselves a function of treatment assignment.
 %A simple, popular approach to estimating subgroup effects is to use ordinary least squares (OLS) regression, regressing the outcome on a treatment indicator interacted with an indicator for subgroup membership.
-However, when subgroup membership is a function of a post-treatment variable, then it may itself be affected by the treatment.
-Contrasts between treatment and control conditions within those subgroups may not have a causal interpretation at all---and when they do, the same techniques that work for subgroups defined at baseline %(such as OLS with interactions)
-will often produce inconsistent estimates.
-
-Principal stratification \citep{frangakis} is a broad framework for defining subgroup effects for post-treatment subgroups.
-It relies on the notion of \emph{potential} subgroups called principal strata: groups of subjects defined based on the values a post-treatment variable would potentially take if the subject were assigned to one treatment condition or another.
-Since principal strata membership is based on subjects' responses to both factual and counter-factual treatment assignments, it is in general unobserved.
-Nonetheless, researchers have devised estimators of principal effects, as average effects within principal strata are called, to address a broad range of applied statistical problems, including non-compliance in randomized experiments \citep{air}, the evaluation of surrogate outcomes \citep{li2010bayesian}, the investigation of causal mechanisms \citep{lidsayPage}, and sample attrition \citep{zhangRubin, ding2011}.
-
-Because principal strata membership is generally unobserved, principal effects are not always identified, and when they are, estimation can still be difficult.
+%However, when subgroup membership is a function of a post-treatment variable, then it may itself be affected by the treatment.
+%Contrasts between treatment and control conditions within those subgroups may not have a causal interpretation at all---and when they do, the same techniques that work for subgroups defined at baseline %(such as OLS with interactions)
+%will often produce inconsistent estimates.
+
+%%% open with hook for ed researchers
+
+When students get stuck on math problems during homework or classwork, it may be helpful to offer them hints or other ``just-in-time'' support \cite[e.g][]{jit}, possibly helping them overcome specific misconceptions or other sources of confusion and allowing them to move on in the assignment.
+The availability of problem-specific support for struggling students is one potential advantage of educational technology over traditional pencil-and-paper practice. 
+On the other hand, the ready availability of hints may discourage students from working through difficult problems on their own, and so may inhibit learning.
+
+The literature on hints in online learning is mixed \citep[see, e.g.,][]{aleven2016help,goldin2012learner,sales2021student} in particular when it comes to requesting a ``bottom-out'' hint.
+In some online mathematics programs, sequences of hints, which a student can request one at a time, end with a bottom-out hint providing the answer to the problem.
+On the one hand, hints that include the correct answer are essentially worked examples, which can be beneficial to learning \citep[e.g.]{sweller1985use}.
+On the other hand, they allow students to ``game the system,'' by simply requesting enough hints to see the right answer, and use the software without ever working to solve any problems \citep[e.g.][]{guo2008trying}.
+
+Data from the randomized controlled trial (RCT) reported in \citet{impactPaper} may provide evidence for one or another side of this debate. 
+That RCT contrasted four different applications for computer-based math practice, one of which included bottom-out hints. 
+If that condition benefited students who frequently requested bottom-out hints (``bottom-outers'') more than those who didn't, that may suggest that bottom-out hints can be helpful. 
+If students who avoided bottom-out hints benefited more, it may be because bottom-out hints were harmful.
+(There are certainly other factors at play as well, so this analysis would not be conclusive.)
+
+Estimating an effect for bottom-outers is a tricky proposition. 
+Causal inference in statistics typically depends on comparisons of outcomes between similar groups of subjects.
+However, there is no readily identifiable comparison group for bottom-outers, who are only represented in one of the four treatment conditions, and who presumably struggled more frequently, on average, than their non-bottom-outer peers.
+We will formulate a solution using Principal stratification \citep{frangakis}, a  framework for defining subgroup effects for subgroups that may themselves be affected by treatment assignment.
+It relies on the notion of \emph{potential} subgroups called principal strata, grouping subjects based on the values a post-treatment variable would potentially take if the subject were assigned to one treatment condition or another.
+For example, the appropriate comparison group for bottom-outers is the set of students, assigned to the other treatment arms, who \emph{would have} requested bottom-out hints if only given that option.
+
+Since principal strata membership is based on subjects' responses to both factual and counter-factual treatment assignments, it is unobserved---we don't observe which students in the conditions without bottom-out hints would have bottomed out if given the opportunity.
+Nonetheless, researchers have devised estimators of principal effects, as average effects within principal strata are called, to address a broad range of applied statistical problems.
+For instance, \citet{air} introduced the framework (or an antecedent thereof) for the case of non-compliance, in which treatment assignment and treatment receipt don't match.
+They applied it to estimate the effects of military service in the American war in Vietnam using the draft lottery, even though some draftees didn't serve and some people served without being drafted, by focusing on the principal stratum of Americans who would serve only if drafted.
+\citet{fellerEtAl2016} estimated differential effects of random assignment to the Head Start early childhood care program for children who, if assigned to control, would receive home-based care, and for those who would attend a different early childhood care program.
+Other methodologists have developed estimators for the evaluation of surrogate outcomes \citep{li2010bayesian}, the investigation of causal mechanisms \citep{lidsayPage}, sample attrition \citep{zhangRubin, ding2011}, and other problems.
+
+Because principal strata membership is unobserved, principal effects are not always identified, and when they are, estimation can still be difficult.
 Typically, principal effects are identified and estimated based on strong structural assumptions \citep[e.g.][]{air} and/or parametric modeling \citep[e.g.][]{imbens1997bayesian}.
 Parametric models for principal effects are often hard to specify and hard or impossible to test.
-Even when they are well-specified they can yield estimates that are severely biased and misleading \citep{griffin2008application,feller2016principal}.
+Even when they are well-specified they can yield estimates that are severely biased  \citep{griffin2008application,feller2016principal}.
+The most common type of parametric model, the normal mixture model, is especially dubious for studying bottom-outers in \citet{impactPaper}'s RCT, whose outcome, the total score on a 10-item posttest, is not normally distributed.
+
 Another route to identification and estimation of principal effects leads through auxiliary data, including baseline covariates or secondary outcomes \citep{mattei2013exploiting}.
 Of particular importance is the ``principal score,'' \citep{jo,dingLu,feller2017principal}, or the probability that a subject belongs to a particular principal stratum, as a function of pre-treatment covariates.
-Under a strong, untestable assumption called ``principal ignorability,''  principal scores can be transformed into weights to compute form an unbiased estimate of principal effects.
-
-On the other hand, \citet{jo2002}, \citet{ding2011}, \citet{jiangDing2021}, and others have shown that, under certain circumstances, researchers can use baseline covariates to identify and estimate principal effects without relying on principal ignorability.
-This paper expands on those identification results by presenting a new set of principal effect estimators, and associated standard error estimates, based on a general estimating equations (or M-estimation or method-of-moments) \citep{stefanskiBoos}.
-It turns out that, under certain general conditions, principal effects can be estimated with a simple, two-step procedure: first, estimate principal scores with logistic regression or another M-estimator; then, impute unknown principal subgroup membership indicators with principal scores and estimate principal effects using ordinary least squares (OLS) regression with interactions.
+Under a strong, untestable assumption called ``principal ignorability,''  principal scores can be transformed into weights to compute an unbiased estimate of principal effects.
+
+%%% put in paragraph here or above with PS in ed research (lindsay page, feller, my papers)
+
+\citet{jo2002}, \citet{ding2011}, \citet{jiangDing2021}, and others have shown that, under certain circumstances, researchers can use baseline covariates to identify and estimate principal effects without relying on principal ignorability.
+This paper expands on those identification results by presenting a new set of principal effect estimators, and associated standard error estimates, based on general estimating equations (or M-estimation or method-of-moments) \citep{stefanskiBoos}.
+Under certain general conditions, principal effects can be estimated with a simple, three-step procedure: first, estimate principal scores with logistic regression or another M-estimator; then, impute unknown principal subgroup membership indicators with principal scores, and finally estimate principal effects using ordinary least squares (OLS) regression with interactions.
 %, just as in the pre-treatment subgroup case.
-We will refer to the estimators presented here as ``\textsc{geepers}'': generalized estimating equations for principal effects using regressions.
-\textsc{geepers} estimates rely on correct specification of regression functions, but unlike common fully-parametric methods, they make no assumptions about the shape of the distribution of regression errors.
-Unlike principal score weighting, \textsc{geepers} does not assume principal ignorability, and admits to sandwich-style standard error estimates, which we provide in the appendix.
-
-This paper focuses on a special case of principal stratification---one-way, binary non-compliance---that is, there are only two principal strata, and principal stratum membership is observable in one treatment arm, but not in the other.
-This scenario would occur if, say, subjects in the control arm of a study had no access to the treatment, and not all subjects assigned to the treatment arm actually received the full treatment.
+We will refer to the estimators presented here as ``\textsc{geepers}'': general estimating equations for principal effects using regressions.
+\textsc{geepers} estimates rely on the correct specification of regression functions, but unlike common fully parametric methods, they make no assumptions about the shape of the distribution of regression errors.
+Unlike principal score weighting, \textsc{geepers} does not assume principal ignorability and admits to sandwich-style standard error estimates, which we provide in the appendix.
+
+This work advances the literature in three ways: first, it makes an explicit connection between principal score methods, M-estimation, and semi-parametric estimation of principal effects using auxiliary data.
+Second, it provides a general estimator that relies on two of the most familiar tools in the applied education researcher's toolbox: linear and logistic regression.
+The \textsc{geepers} method can, therefore, make principal stratification methods accessible for a much broader range of applied education researchers.
+Lastly, this work includes an extensive simulation study that compares \textsc{geepers} to parametric mixture modeling and principal score weighting under a range of favorable and challenging scenarios.
+The simulation demonstrates the strengths and weaknesses of each method and the tradeoffs between them.
+
+This paper focuses on a special case of principal stratification---one-way, binary non-compliance---that is, there are only two principal strata and principal stratum membership is observable in one treatment arm, but not in the other.
+This scenario would occur if, say, subjects in the control arm of a study had no access to the treatment, and not all subjects assigned to the treatment arm received the full treatment.
 While such a setup may represent the simplest and easiest principal stratification problem, we believe it is a common scenario.
-Moreover, we believe that the methods developed here can extend to more complex scenarios, which we leave to future work.
+For example, \textsc{geepers} could be used to estimate effects from an RCT for participants categorized as low- or high-implementers.
+It could also be used to assess more granular questions about effect heterogeneity, such as whether an intervention's impact is contingent upon specific behaviors displayed by either practitioners or students.
 
 The following section gives a formal introduction to principal stratification and briefly describes two common estimation techniques, parametric mixture modeling and principal score weighting.
-Next, Section \ref{sec:geepers} develops \textsc{geepers} estimation, and gives conditions for its consistency.
-Section \ref{sec:simulation} describes a simulation study examining the operating characteristics of \textsc{geepers} estimates and comparing them to estimates from parametric mixture modeling and principal score weighting.
-Under ideal circumstances, estimates from parametric mixture models and weighting estimators are somewhat more efficient than \textsc{geepers}, but when assumptions are violated \textsc{geepers} easily out-performs the other two methods.
-Section \ref{sec:fh2t} illustrates \textsc{geepers} in analyzing data from a recent large-scale field trial of educational software, and Section \ref{sec:discussion} concludes.
+Next, Section \ref{sec:geepers} develops \textsc{geepers} estimation and gives conditions for its consistency.
+Section \ref{sec:simulation} describes the simulation study.
+% examining the operating characteristics of \textsc{geepers} estimates and comparing them to estimates from parametric mixture modeling and principal score weighting.
+%Under ideal circumstances, estimates from parametric mixture models and weighting estimators are somewhat more efficient than \textsc{geepers}, but when assumptions are violated \textsc{geepers} easily out-performs the other two methods.
+Section \ref{sec:fh2t} illustrates \textsc{geepers} by returning to the question of bottom-out hints: does the effect of being randomized to a condition that includes bottom-out hints depend on how often they are requested?
+Section \ref{sec:discussion} concludes.
 
 \section{Background}
-In a randomized experiment with two conditions, let $Z_i=0$ if subject $i$, $i=1,\dots,n$ is assigned to the treatment condition, and let $Z_i=0$ otherwise.
+In a randomized experiment with two conditions, let $Z_i=1$ if subject $i$, $i=1,\dots,n$ is assigned to the treatment condition, and let $Z_i=0$ otherwise.
 If interest is in the effect of $Z$ on an outcome $Y$, under the ``stable unit treatment value assumption'' \citep{air} of no interference between units and no hidden versions of the treatment, define potential outcomes \citep{splawa1990application} $\yti$ as the value of $Y$ that $i$ would exhibit if $Z_i=1$, and let $\yci$ be the value if $Z_i=0$.
@@
 
-Let $S_i\in\{0,1\}$ be a measurement on subject $i$ taken subsequent to treatment assignment, so that $S$ may be affected by $Z$.
+Let $S_i\in\{0,1\}$ be a measurement on subject $i$ taken after treatment assignment, so that $S$ may be affected by $Z$.
 Then $S$ itself has potential values $\sti$ and $\sci$, which $i$ would exhibit if $Z_i=1$ or $Z_i=0$, respectively.
@@
 Under strong monotonicity, since $S$ is binary, every subject belongs to one of two ``principal strata,'' $\sti=1$ or $\sti=0$.
-The average treatment effect in each principal stratum, $\eff1=\EE[\tau|\st=1]$ or $\eff1=\EE[\tau|\st=0]$, is called a principal effect. %
-The challenge in estimating principal effects is due to the fact that strata membership $\sti$ is only observed when $Z_i=1$, in which case $\sti=S_i$, the observed value of $S$.
+The average treatment effect in each principal stratum, $\eff1=\EE[\tau|\st=1]$ or $\eff0=\EE[\tau|\st=0]$, is called a principal effect. %
+The challenge in estimating principal effects is because strata membership $\sti$ is only observed when $Z_i=1$, in which case $\sti=S_i$, the observed value of $S$.
 
@@
 \end{ass}
-where $\bm{x}_i$ is a vector of pre-treatment covariates.
+where $\bm{x}_i$ is a vector of pre-treatment covariates, and $\independent$ denotes independence.
 
@@
 
-This approach has a number of setbacks.
+This approach has several problems.
 First, the conditional normality assumption is untestable and limiting.
@@
 with parameter vector $\bm{\alpha}$.
-An analyst may estimate $\bm{\alpha}$ by fitting model \eqref{eq:prinScoreMod} using observed $\bm{S}$ and $\bxs$ for subjects with $Z=1$, and then compute estimated principal scores $\ppip{\bm{\hat{\alpha}}}$ for subjects with $Z_i=1$.
+An analyst may estimate $\bm{\alpha}$ by fitting model \eqref{eq:prinScoreMod} using observed $\bm{S}$ and $\bxs$ for subjects with $Z=1$, and then compute estimated principal scores $\ppip{\bm{\hat{\alpha}}}$ for subjects with $Z_i=0$.
 
@@
 
-In order for principal scores to be useful, they need to vary with $\bm{x}$. Otherwise, $\bm{x}$ is uninformative about $\st$. In particular, we will assume the following:
+For principal scores to be useful, they need to vary with $\bm{x}$. Otherwise, $\bm{x}$ is uninformative about $\st$. In particular, we will assume the following:
 \begin{ass}[Variable Principal Scores]\label{ass:vps}
@@
 
-\section{\textsc{geepers}}\label{sec:geepers}
+\section{GEEPERS }\label{sec:geepers}
 
 The approach we introduce here incorporates principal scores into an M-Estimator for the mixture model \eqref{eq:mixtureModel}.
-Like principal score weighting, our M-estimator does not require distributional assumptions, but instead requires a conditional independence assumption.
-We will begin by describing a stronger-than-necessary conditional independence assumption in order to build intuition; in \S~\ref{sec:regression} we will present a considerably weaker alternative.
+Like principal score weighting, our M-estimator does not require distributional assumptions but instead requires a conditional independence assumption.
+We will begin by describing a stronger-than-necessary conditional independence assumption to build intuition; in \S~\ref{sec:regression} we will present a considerably weaker alternative.
 
@@
 
-We begin by introducing an assumption that we term ``Covariate Ignorability,'' or CI. It is a slightly weaker version of \citet{jiangDing2021}'s ``auxiliary independence'' assumption.
-\begin{ass}[Covariate Ignorability]\label{ass:ci}
+We begin by introducing an assumption that we term ``Covariate Independence,'' or CI. It is a slightly weaker version of \citet{jiangDing2021}'s ``auxiliary independence'' assumption.
+\begin{ass}[Covariate Independence]\label{ass:ci}
 \begin{equation}\label{eq:assumption}
@@
 As stated above, CI will rarely be plausible.
-However, in some circumstances researchers can identify a subset of observed covariates that satisfy CI, and use those covariates in estimation.
+However, in some circumstances, researchers can identify a subset of observed covariates that satisfy CI, and use those covariates in estimation.
 
 It turns out that under strong monotonicity, randomization, and CI, and given a set of principal scores, the two principal effects $\eff0$ and $\eff1$ can be estimated by a simple OLS regression.
-To see how, we'll %first
+To see how we'll %first
 derive the estimating equations for $\muc0$, $\muc1$, $\mut0$, and $\mut1$, and show that they are equivalent to estimating equations for a model fit by OLS.
@@
 \begin{prop}\label{prop:reg1}
-  Under Assumptions \ref{ass:sm}, \ref{ass:rand}, \ref{ass:vps}, \ref{ass:ci}, and suitible regularity conditions \citep[e.g.][p. 327]{boosStefanskiBook}, if %principal scores are known exactly,
+  Under Assumptions \ref{ass:sm}, \ref{ass:rand}, \ref{ass:vps}, \ref{ass:ci}, and suitible regularity conditions, if %principal scores are known exactly,
   $\ri$ is defined as in \eqref{eq:ri} and the model
@@
 
-\subsection{Relaxing Covariate Ignorability with Outcome Modeling}\label{sec:regression}
+\subsection{Relaxing Covariate Independence with Outcome Modeling}\label{sec:regression}
 In most applications, identifying a set of covariates $\bxs$ satisfying CI will be difficult or impossible.
@@
 
-The assumption we'll state here is stronger than necessary, but leads to an attractively simple method; a weaker version of the assumption and corresponding method, along with the proof, can be found in an online appendix.
-
-\begin{ass}[Residualized Covariate Ignorability]\label{ass:rci}
+The assumption we'll state here is stronger than necessary, but leads to an attractively simple method; a weaker version of the assumption and corresponding method, along with the proof, can be found in an appendix.
+
+\begin{ass}[Residualized Covariate Independence]\label{ass:rci}
 There exists $\bxy$, a (known) transformation of covariates $\bx$, and an (unknown) vector of coefficients $\bm{\gamma}$ such that
@@
 This is akin to the ``Additivity of Treatment Assignment Effect'' assumption in \citet{jo2002}.
-Proposition \ref{prop:interactions} in the online appendix relaxes these requirements, but some preliminary simulations suggest that standard errors from estimates allowing for these interactions will be prohibitively large.
-Simulation results in the following section address some cases in which interactions are present in the data generating model but not in the data analysis.
+Proposition \ref{prop:interactions} in the appendix relaxes these requirements, but some preliminary simulations suggest that standard errors from estimates allowing for these interactions will be prohibitively large.
+Simulation results in the following section address some cases in which interactions are present in the data-generating model but not in the data analysis.
 
@@
   $\heff0_{RCI}$ and $\heff1_{RCI}$ are M-estimators, with $\heff0_{RCI}\rightarrow_p\eff0$ and $\heff1_{RCI}\rightarrow_p\eff1$ as $n\rightarrow\infty$.
-  Under suitible regularity conditions, they are jointly asymptotically normal, with a variance of the form \eqref{eq:sandwich}.
+  Under suitable regularity conditions, they are jointly asymptotically normal, with a variance of the form \eqref{eq:sandwich}.
 \end{prop}
@@
 Does it avoid the pitfalls of mixture modeling when the parametric assumptions of the normal mixture model fail?
-Third, are there ways or circumstances when the principal score weighting estimator outperforms \textsc{geepers} even though principal ignorability does not hold?
+Third, are there ways or circumstances when the principal score weighting estimator outperforms \textsc{geepers} or a mixture estimator even though principal ignorability does not hold?
 
 \subsection{Simulation Design}
-The simulation study was conducted in \texttt{R} \citep{rcite} and Stan \citep{rstan}, and full replication code is available at \url{https://github.com/adamSales/psGee}.
+The simulation study was conducted in \texttt{R} \citep{rcite} and Stan \citep{rstan}, and full replication code is available at %[Redacted]. %
+\url{https://github.com/adamSales/psGee}.
 \subsubsection{Data Generation}
@@
 When $\alpha$ was higher, $\st$ was more easily predicted by covariates; when $\alpha=0$, Assumption \ref{ass:vps} was violated, i.e. $Var(\pp)=0$.
-Principal stratum $\sti\in \{0,1\}$ was simulated as $S_i\sim Bern(\ppi)$
+Principal stratum $\sti\in \{0,1\}$ was simulated as $S_i\sim Bern(\ppi)$.
 
@@
 
-The distribution of $\epsilon_i$, $\beta_1$, and coefficients $\gamma_1$, $\gamma_2$, and $\gamma_3$ were manipulated factors in the study.
+The distribution of $\epsilon_i$, $\beta_1$, and values of coefficients $\gamma_1$, $\gamma_2$, and $\gamma_3$ were manipulated factors in the study.
 
@@
 
-Using each simulated dataset, we estimated $\eff0$ and $\eff1$ with three methods: a mixture model \eqref{eq:mixtureModelPS} including principal scores, \textsc{geepers}, and third, the principal score weighting estimator \eqref{eq:psw}.
+Using each simulated dataset, we estimated $\eff0$ and $\eff1$ with three methods: a mixture model \eqref{eq:mixtureModelPS} including principal scores, \textsc{geepers}, and the principal score weighting estimator \eqref{eq:psw}.
 
@@
 
-The mixture estimator then fit two outcome models:
+The mixture estimator then fits two outcome models:
 For the treatment group, the regression model $Y=\beta_0^T+\beta_1^TS+\bm{\gamma}'\bx+\epsilon$ and for the control group, a mixture model of the form \eqref{eq:mixtureModelPS}, but with $\muc0$ and $\muc1$ replaced by regression functions $\beta_0^C+\bm{\gamma}'\bx$ and $\beta_0^C+\beta_1^C+\bm{\gamma}'\bx$, respectively.
@@
 
-In contrast, for \textsc{geepers} and principal score weighting the principal score model was fit by itself, to data from the treatment group, using standard maximum likelihood methods.
+For \textsc{geepers} and principal score weighting the principal score model was fit by itself, to data from the treatment group, using standard maximum likelihood methods.
 For \textsc{geepers}, we used the estimated principal scores in the control group, along with the observed $S$ in the treatment group, to construct $R$, and then fit the model \eqref{eq:reg2} with OLS.
@@
 
-For principal score weighting we plugged estimated principal scores in to \eqref{eq:psw} to estimate $\muc0$ and $\muc1$, estimated $\mut0$ and $\mut1$ with the sample means of subjects with $Z=1$ and $S=0$ or $S=1$, respectively, and estimated $\heff0=\hat{\mu}_{Tw}^0-\hat{\mu}_{Cw}^0$ and $\heff1=\hat{\mu}_{Tw}^1-\hat{\mu}_{Cw}^1$.
+For principal score weighting we plugged estimated principal scores into \eqref{eq:psw} to estimate $\muc0$ and $\muc1$, estimated $\mut0$ and $\mut1$ with the sample means of subjects with $Z=1$ and $S=0$ or $S=1$, respectively, and estimated $\heff0=\hat{\mu}_{Tw}^0-\hat{\mu}_{Cw}^0$ and $\heff1=\hat{\mu}_{Tw}^1-\hat{\mu}_{Cw}^1$.
 
@@
 \subsubsection{Varying $n$ and $\alpha$}
-When $\alpha=0$, Assumption \ref{ass:vps} was violated, and \textsc{geepers} estimates were highly erratic; however, as Table \ref{tab:coverage} shows, when other assumptions held inference remained valid.
+When $\alpha=0$, Assumption \ref{ass:vps} was violated, and \textsc{geepers} estimates were highly erratic; however, as Table \ref{tab:coverage} shows, when other assumptions held, inference remained valid.
 
@@
   \centering
-  \includegraphics[width=\textwidth]{../simFigs/biasSEbyB1N.jpg}
-  \caption{Empirical bias and standard error of \textsc{geepers} and Mixture estimates of $\muc1$ as sample size per treatment group $n$ or $\alpha$ varied.
+  \includegraphics[width=4.5in]{../simFigs/biasSEbyB1N.jpg}
+  \caption{Empirical bias and standard error of \textsc{geepers} and Mixture estimates of $\mu_C^1$ as sample size per treatment group $n$ or $\alpha$ varied.
 There were 500 replications for each $n$ or $\alpha$.
@@
   \centering
-  \includegraphics[width=\textwidth,clip]{../simFigs/boxplots.pdf}
+  \includegraphics[width=5.5in,clip]{../simFigs/boxplots.pdf}
   \caption{Violin and scatter plots of 500 simulation estimation errors for estimators described in Section~\ref{sec:simMods} under varying conditions.
@@
 This is somewhat surprising, because when $\muc1=0$, there was no separation between the principal strata in the control group, and previous literature \citep{griffin2008application} has identified this scenario as a particular challenge for mixture modeling.
-The \textsc{psw} estimator was also biased---due, presumably, to the violation of Principal Ignorability, Assumption \ref{ass:PI}---inherent in the data generating model.
+The \textsc{psw} estimator was also biased, due, presumably, to the violation of Principal Ignorability, Assumption \ref{ass:PI}, inherent in the data generating model.
 On the other hand, the \textsc{psw} estimator had easily the lowest sampling variance of the three estimators.
-The middle panel of Figure \ref{fig:boxplots} and the fifth line of Table \ref{tab:coverage} show a similar set of results for when the residuals in \eqref{eq:y-sim} are uniformly distributed.
+
+The middle panel of Figure \ref{fig:boxplots} and the fifth line of Table \ref{tab:coverage} show results for when the residuals in \eqref{eq:y-sim} are uniformly distributed.
 In this case, \textsc{geepers} and \textsc{psw} estimator, which do not assume normality, behaved in a similar fashion as in the normal case, but the mixture estimator showed a bimodal pattern--sometimes over-estimating $\eff1$ and sometimes underestimating, but rarely estimating $\eff1$ accurately.
@@
 %Presence of interactions could undermine the \textsc{geepers} by inducing a relationship between principal scores and potential outcomes, even after adjusting the outcomes for covariate effects.
-Fortunately, these interactions made little difference when $\alpha=0.5$; however, Table \ref{tab:coverage} shows that when $\alpha=0$ or 0.2, interactions between $\bx$ and $Z$ were present (i.e.
-$\gamma_3\ne 0$) \textsc{geepers} 95\% confidence intervals tended to under-cover.
+Fortunately, these interactions made little difference when $\alpha=0.5$.
+However, Table \ref{tab:coverage} shows that when $\alpha=0$ or 0.2, and interactions between $\bx$ and $Z$ were present (i.e.
+$\gamma_3\ne 0$), \textsc{geepers} 95\% confidence intervals tended to under-cover.
 Nevertheless, coverage of \textsc{geepers} intervals was nearly always substantially higher than coverage of corresponding 95\% credible intervals from mixture models.%\footnote{}
 
-\begin{table}
-
-\caption{\label{tab:coverage}Empirical coverage of nominal 95\% confidence intervals for estimates of $\eff1$.}
+\begin{table}%[!ht]
+
+\caption{\label{tab:coverage}Empirical coverage of nominal 95\% confidence intervals for estimates of $\tau^1$.}
 \centering
-\begin{tabular}{*{9}{c}}%*{6}{r}}
+\begin{tabular}{ccccccccc}%*{6}{r}}
 \toprule
@@
 \bottomrule
-\multicolumn{9}{p{0.9\textwidth}}{\footnotesize Based on 500 replications. $n=500$, $\beta_1=0.3$. Simulation standard error $\approx 1$ percentage point.}
+\multicolumn{9}{p{3.5in}}{\footnotesize Based on 500 replications. $n=500$, $\beta_1=0.3$. Simulation standard error $\approx 1$ percentage point.}
 
@@
 
-\section{Application: Estimating Treatment Effects for Implementation Subgroups}\label{sec:fh2t}
-
-In a recent educational field experiment, middle school students completing their math schoolwork on a computer program were randomized between four conditions. 
-These included two gamified programs---DragonBox and From Here to There (FH2T)---and two programs in which students used computers to work through a series of algebra problems taken from the district's textbook---``Assistments'' and an active control called ``Business as Usual'' (BAU).
+\section{Application: Estimating Treatment Effects for Implementation Subgroups: Students who Frequently or Rarely Request Bottom-Out Hints}\label{sec:fh2t}
+
+In the educational field experiment reported in \citet{impactPaper}, middle school students completing their math schoolwork on a computer program were randomized between four conditions. 
+These included two gamified programs---DragonBox and From Here to There (FH2T)---and two programs in which students used computers to work through a series of algebra problems taken from the district's textbook---``ASSISTments'' and an active control called ``Business as Usual'' (BAU).
 % the active control condition in which students simply used the computer to work on a sequence of math problems.
-%In the DragonBox game, students learned to solve algebraic equations by isolating a box containing a dragon; as students progressed in the game, pictures of monsters were replaced by mathematical symbols.
-%Another condition, the From Here to There (FH2T) program, used visual cues such as color and spacing to teach algebraic concepts, and allowed students to solve equations by manipulating expressions via dragging and dropping.
+In the DragonBox game, students learned to solve algebraic equations by isolating a box containing a dragon; as students progressed in the game, pictures of monsters were replaced by mathematical symbols.
+Another condition, the From Here to There (FH2T) program, used visual cues such as color and spacing to teach algebraic concepts and allowed students to solve equations by manipulating expressions via dragging and dropping.
 %In ASSISTments, students used computers to work through a series of math problems taken from the district's textbook. 
-Students in ASSISTments had access to a series hints for each problem, culminating in a ``bottom out'' hint that contained the answer and were given immediate feedback on their answers.
-% and had to enter the correct answer in order to proceed to the next problem.
+Students in the ASSISTments condition had access to a series of hints for each problem, culminating in a ``bottom out'' hint that contained the answer.
+They received immediate feedback on their answers and had to enter the correct answer to proceed to the next problem.
 In the BAU condition students had no access to hints and received feedback only after completing the assignment.
-In all four conditions, students worked on their assigned math program in class on 12 different occasions spaced throughout the school year: a prior assessment, a mid-test, a post-test, and nine learning sessions.
+
+In all four conditions, students worked in class on their assigned math program on 12 different occasions spaced throughout the school year: a prior assessment, a mid-test, a post-test, and nine learning sessions.
 In this illustration, we will compare ASSISTments---the only condition featuring bottom-out hints---with each of the three others separately.
-Our goal is to estimate different treatment effects for groups of students who would, if assigned to ASSISTments, request few or many bottom out hints.
-
-The literature on hints in online learning is mixed \citep[see, e.g.,][]{aleven2016help,goldin2012learner,sales2021student} in particular when it comes to requesting a bottom-out hint (called ``bottoming out'').
-On the one hand, hints that include the correct answer are essentially worked examples, which can be beneficial to learning \citep[e.g.]{sweller1985use}.
-On the other hand, they allow students to ``game the system,'' by simply requesting enough hints to see the right answer, and use the software without ever working to solve any problems \citep[e.g.][]{guo2008trying}.
-
-In this experiment, the ability to bottom out was only one of several differences between the treatment conditions.
+Our goal is to estimate different treatment effects for groups of students who would, if assigned to ASSISTments, request few or many bottom-out hints.
+
+The ability to bottom out was only one of several differences between the treatment conditions.
 One way to disentangle the role of bottoming out in any treatment effect would be to estimate average treatment effects of being assigned to ASSISTments, relative to any of the other three, separately for subjects who would (if assigned to ASSISTments) bottom out frequently, and for subjects who would not.
 If the availability of bottom-out hints plays an important positive role in ASSISTments' efficacy, we might expect larger effects of condition on students who would bottom out frequently than for those who would not; if bottom-out hints are harmful, we might expect the opposite.
-%If hints (that do not supply the correct answer) and error messages are helpful, but the ability to bottom out is harmful to learning, then we might expect to estimate a positive effect of being assigned to the feedback condition for students who do not bottom out, but a negative effect for students who do.
+If hints (that do not supply the correct answer) and error messages are helpful, but the ability to bottom out is harmful to learning, then we might expect to estimate a positive effect of being assigned to the feedback condition for students who do not bottom out, but a negative effect for students who do.
 If hints are helpful for learning only if they can be read as worked examples, then we might expect no effect of condition on students who generally do not bottom out, but a positive effect for students who do.
@@
 whether students in ASSISTments requested more than the median number of bottom-out hints (i.e. 11).
-That is, $\st=1$ for students who, if assigned to ASSISTments, would request at more than 11 bottom-out hints (``bottom-outers''), and $\st=0$ for students who would request 11 or fewer (``non-bottom-outers'').
+That is, $\st=1$ for students who, if assigned to ASSISTments, would request more than 11 bottom-out hints (``bottom-outers''), and $\st=0$ for students who would request 11 or fewer (``non-bottom-outers'').
 %least one bottom-out hint, and $\st=0$ for students who would not.
 The outcome of interest $Y$ in our example is students' total scores on a ten-item posttest completed within the online tutoring system; $Y$ is an integer ranging from 0 to 10.
-The goal of our analysis will be to determine the principal effects $\tau_1=\EE[\yt-\yc|\st=0]$ and $\tau_0=\EE[\yt-\yc|\st=0]$, the average effects of assignment to ASSISTments (``$T$'' subscripts) versus the other conditions (``$C$'') for bottom-outers and non-bottom-outers.
-
-We also had access to data on a number of baseline covariates, including prior achievement, demographics, baseline measurements of student attitudes towards math and an indicator for whether students began the school year in remote or in-person instruction.
+The goal of our analysis will be to determine the principal effects $\eff1=\EE[\yt-\yc|\st=0]$ and $\eff0=\EE[\yt-\yc|\st=0]$, the average effects of assignment to ASSISTments (``$T$'' subscripts) versus the other conditions (``$C$'') for bottom-outers and non-bottom-outers.
+
+We also had access to data on several baseline covariates, including prior achievement, demographics, baseline measurements of student attitudes toward math, and an indicator for whether students began the school year in remote or in-person instruction.
 Students with missing pretest scores were dropped from the analysis; missing data in other covariates was imputed with a Random Forest algorithm \citep{missForest}.
-Summary statistics are displayed in the online appendix.
-
-The data analysis we present here is intended as a demonstration of \textsc{geepers}, rather than for its substantive conclusions, and our description omits discussion of some important methodological considerations, including attrition and post-selection inference.
-For a more complete discussion of the experimental design, the conditions being compared, and the impact analysis, please see \citet{impactPaper}.
-
-Due to privacy regulations, the data are not publicly available; however, interested analysts may access anonymized data by emailing the study's principal investigator at erottmar@wpi.edu and signing a privacy agreement.
-Replication code for the analysis is available at \url{https://github.com/adamSales/psGee}.
-
-%\subsection{Data Analysis}
+Summary statistics are displayed in appendix tables \ref{table:tab1fac} and \ref{table:tab1num}.
+
+%The data analysis we present here is intended as a demonstration of \textsc{geepers}, rather than for its substantive conclusions, and our description omits discussion of some important methodological considerations, including attrition and post-selection inference.
+For details on the experimental design, the conditions being compared, sample attrition---which was unusually high, due to the COVID-19 pandemic---and the impact analysis, please see \citet{impactPaper}.
+
+For a description of the dataset and an explanation of how to access it, see \citet{ottmar2023data} and \url{https://osf.io/r3nf2/}. 
+Replication code for the analysis is available at %[Redacted]. 
+\url{https://github.com/adamSales/psGee}.
+
+\subsection{Data Analysis}
+We estimated principal scores using logistic regression of observed $S$ on school fixed effects and a set of baseline variables $\bxs$.
+Three criteria governed our choice of covariates and regression specifications for estimating principal scores and effects.
+First, Proposition \ref{prop:reg2} requires consistent estimation of principal scores and (via Assumption \ref{ass:rci}) correct specification of the outcome model, that is, the model for $\EE[Y|\bxy,Z,\st]$.
+Hence, it is essential to choose covariates and transformations $\bxs$ and $\bxy$ that lead to good model fit.
+
+Second, Assumption \ref{ass:vps} requires principal scores to vary with at least some element $\bxs$.
+Moreover, the results of the simulation study suggest that when $\st$ can be predicted more precisely as a function of $\bxs$, \textsc{geeper}s principal effect estimators are more precise and more robust to model misspecification.
+Those considerations militate in favor of choosing covariates $\bxs$ to optimize out-of-sample principal score prediction accuracy. 
+There is good reason to believe that analogous arguments apply to selection of outcome covariates $\bxy$.
+
+The first two criteria motivated our choice of two strategies for choosing $\bxs$ and $\bxy$.
+In our main model, we included every available covariate in the dataset and chose transformations of those covariates based on the results of residual and binned residual plots \citep{arm}.
+Including all possible covariates maximizes the model's opportunities to detect covariate relationships that would predict $Y$ or $\st$. 
+However, this strategy can also lead to overfitting, which would decrease out-of-sample predictive accuracy. 
+In an alternative analysis, we chose a subset of covariates to include in $\bxs$ and $\bxy$ with a backward stepwise selection algorithm minimizing the AIC \citep{aic} of the principal score model and then modified the resulting models based on residual plots.
+The model selection process is recorded in the GitHub repository cited above.
+
+Lastly, Assumption \ref{ass:ci} implies that there is no interaction between $\bxs$ and $\st$ in the outcome model.
+Further, for the sake of statistical efficiency, we did not consider outcome models with interactions between covariates and $Z$.
+\emph{A priori}, the student attribute that poses the greatest threat to these modeling choices is prior achievement, that is, scores on the pretest and the fifth-grade state test. 
+This is both because they measure a similar construct to the posttest, and because \cite{impactPaper} found a pretest-treatment interaction in a different treatment contrast. 
+Following that reasoning, we conducted a third parallel analysis excluding prior achievement covariates in $\bxs$ but including all covariates in $\bxy$.
+
+All three principal score models were fit to data from students in the ASSISTments group, for whom bottom-out status ($S$) was observed; hence, the same model fit was used for all three treatment contrasts.
+
+Outcome models for all analyses adjusted for all of the variables that were in the corresponding principal score models, in addition to measurements of students' prior knowledge; also, the models substituted class fixed effects for school fixed effects.\footnote{Why use a coarser set of fixed effects for the principal score model than the outcome regression?
+First, the principal score model was fit to a smaller sample (only students in the ``Instant'' condition), so some classrooms only included one student, which can lead to poor fit.
+Second, the asymptotics of logistic regression are more restricted than OLS, so logistic regression tends to perform poorly when the number of parameters increases with the sample size \citep{agresti}.}
+As with principal score models, we checked model fit with residual plots and modified model specifications accordingly. 
 %\subsubsection{Principal Score Model}\label{sec:psMod}
@@
 
-We estimated principal scores using logistic regression of observed $S$ on school fixed effects and a set of baseline variables chosen with a combination of expert judgement and backwards selection based on the AIC, and evaluated model fit with binned residual plots \citep{gelman2006data}.
-The model was fit to data from students in the ASSISTments group, for whom bottom-out status ($S$) was observed; hence, the same model fit was used for all three treatment contrasts.
-The model was fairly successful in distinguishing bottom-outers from non-bottom-outers---the AUC, evaluated using out-of-sample predictions in a 10-fold cross-validation, was roughly 0.795.
-That is, in roughly 80\% of bottom-outer/non-bottom-outer pairs of subjects, the bottom-outer will have a higher predicted probability from the model.
-Coefficient estimates are displayed in the online appendix.
+\subsection{Results}
+
+The three principal score models were fairly successful in distinguishing bottom-outers from non-bottom-outers---the AUC, evaluated using out-of-sample predictions in a 10-fold cross-validation, was roughly 0.795 for the AIC-optimal model, and about 0.745 for the two others.
+That is, in roughly 75--80\% of bottom-outer/non-bottom-outer pairs of subjects, the bottom-outer will have a higher predicted probability from the model.
+Coefficient estimates are displayed in table \ref{tab:psTab} in the appendix.
+
+Figure \ref{fig:effects} shows principal effect point estimates and 95\% confidence intervals from analyses based on the three principal score models.
+The three covariate selection strategies led to similar point and interval estimates. 
+That said, in nearly all cases the AIC-optimal model led to the smallest standard errors.
+Confidence intervals for all nine principal effect estimates were about twice as wide as for corresponding ATE estimates.
+In no case was there evidence at $\alpha=0.05$---or any other conventional level---for non-zero principal effects or differences between principal effects.
+Table \ref{tab:regTab}, in the appendix, displays the full set of regression coefficients for the outcome models based on the "All Covariates" principal score model, along with nominal standard error estimates.
+
+A comparison of point estimates suggests that assignment to ASSISTments rather than BAU benefits bottom-outers less than non-bottom-outers. 
+We speculate that, if true, this pattern may be because many students use bottom-out hints to game the system---entering the provided correct answers rather than attempting to figure out the problems on their own.
+Removing access to hints for bottom-outers forces them to attempt the problems rather than game the system. 
+At the same time, hints and instant feedback are helpful for non-bottom-outers, who are less likely to game the system.
+
+On the other hand, point estimates suggest that assignment to ASSISTments, relative to FH2T, hurts non-bottom-outers but not non-bottom-outers. 
+Perhaps non-bottom-outers tend to have higher prior knowledge, and, as \cite{impactPaper} shows, FH2T benefits such students more than students with lower prior knowledge.
+
+Perhaps the simplest explanation for observed differences in point estimates is sampling error, which cannot be ruled out.
+
+\begin{figure}
+  \centering
+  \includegraphics{prinEffs.pdf}
+  \caption{Principal effect estimates with 95\% confidence intervals using three different principal score models, alongside ATE estimates and 95\% confidence intervals, for contrasts between the ASSISTments condition and each of the three other conditions in the study.}
+  \label{fig:effects}
+\end{figure}
+
+\subsubsection{Comparing Principal Stratification Methods}
 
 % \subsubsection{Outcome Regression and Principal Effects}\label{sec:outMod}
-% \begin{figure}
-%   \centering
-%   \includegraphics{../figure/prinEffs.pdf}
-%   \caption{Principal Effects}
-%   \label{fig:effects}
-% \end{figure}
-
-% Figure \ref{fig:effects} plots the estimated principal effects for bottom-outers and non-bottom-outers, using two different models---alongside estimates of the overall average treatment effect, labeled ``ATE.'' In both models, effects were adjusted for all of the variables that were in the principal score model, in addition to students' pretest scores; also, the models substituted class fixed effects for school fixed effects.
-%\footnote{Why use a coarser set of fixed effects for the principal score model than the outcome regression?
-There were two reasons: first, the principal score model was fit to a smaller sample (only students in the ``Instant'' condition), so some classrooms only included one student, which can lead to poor fit.
-Second, the asymptotics of logistic regression are more restricted than OLS, and so logistic regression tends to perform poorly when the number of parameters increases with the sample size \citep{agresti}.}
+
+% Figure \ref{fig:effects} plots the estimated principal effects for bottom-outers and non-bottom-outers, using two different models---alongside estimates of the overall average treatment effect, labeled ``ATE.''
 % The two sets of estimated differed in that the first assumed that coefficients on covariates were equal across principal strata, whereas the second set of results allowed the coefficients on pretest and 5th-grade standardized test scores---the most important predictors of outcomes and principal strata membership, respectively---to vary between strata.
@@
   \centering
-  \includegraphics{../figure/compareMethods.pdf}
-  \caption{Principal Effects}
+  \includegraphics{compareMethods.pdf}
+  \caption{Principal effect estimates with 95\% confidence or credible intervals using three different principal stratification methods, alongside ATE estimates and 95\% confidence intervals, for contrasts between the ASSISTments condition and each of the three other conditions in the study.
+The \textsc{geepers} estimates used the ``All Covariates'' principal score model.}
   \label{fig:compare}
@@
 
-Figure \ref{fig:compare} shows principal effect estimates from \textsc{geepers} compared to analogous estimates using Bayesian mixture modeling and \textsc{psw}, alongside estimates of the overall average treatment effects, labeled ``ATE.''.
-%All three sets of estimates relied on the same principal score model described above, in Section \ref{sec:psMod}, though in the mixture model it was fit simultaneously with effect estimation using a Markov Chain Monte Carlo algorithm \citep{rstan}.
-Outcome regressions in \textsc{geepers} and the mixture model included all of the covariates that were in the principal score model, in addition to students' pretest scores; also, the models substituted class fixed effects for school fixed effects.
-Coefficient estimates are displayed in the online appendix.
+Figure \ref{fig:compare} shows principal effect estimates from \textsc{geepers} (using the All Covariates principal score model) compared to analogous estimates using Bayesian mixture modeling and \textsc{psw}, alongside estimates of the overall average treatment effects, labeled ``ATE.''
+The \textsc{psw} model used the same principal score model as \textsc{geepers}, while the Bayesian model used the same principal score and outcome model specifications.
 
 The Bayesian mixture model assumed that outcomes were normally distributed, conditional on covariates and principal stratum, with a residual standard deviation that varied between principal strata.
-This modeling assumption is necessarily false, since post-test scores were integers between 0 and 10.
+This modeling assumption is necessarily false since post-test scores were integers between 0 and 10.
 All model parameters with support in $\mathbb{R}$ were given standard normal priors, while standard deviation parameters were given half-standard normal priors.
@@
 
-The \textsc{geepers} results suggest little difference, if any, between the effects in the two principal strata.
-Estimated effects compared to BAU or DragonBox were slightly lower for bottom-outers than for non-bottom-outers, and estimated effects compared to FH2T were higher for non-bottom-outers.
-However, none of these differences is statistically significant, so opposite patterns, or no difference between principal effects at all, would also be consistent with the data.
-In fact, 95\% confidence intervals for all principal effects include both positive and negative principal effects.
-
-The other estimation strategies largely agreed with \textsc{geepers}, and with each other. 
-This may suggest that the estimators' various assumptions---additivity of treatment assignment effect for \textsc{geepers} and the mixture model, normal residuals for the mixture model, and principal ignorability for \textsc{psw}---were all innocuous and may have held approximately.
-Alternatively, violations of the different assumptions may have led to similar errors.
+The other estimation strategies largely agreed with \textsc{geepers} and with each other in the contrast between ASSISTments and BAU, and, to a lesser extent, in teh contrast between ASSISTments and DragonBox.
+The biggest difference between the methods was in the contrast with FH2T, in which the Bayesian mixture model resulted in a larger difference between the two principal effects, as well as the largest standard errors, while the two \textsc{psw} principal effect estimates were quite similar.
+However, these differences are difficult to interpret due to the large uncertainty around all of the point estimates.
 
@@
 If so, several further questions emerge.
-First, can the parameter $\alpha$ be extended to a more general parameter measuring the prediction accuracy of a non-parametric model.
+First, can the parameter $\alpha$ be extended to a more general parameter measuring the prediction accuracy of a non-parametric model?
 Second, if the principal score model cannot be formulated as the solution to a set of estimating equations, how should the standard error be computed?
@@
 
+Sampling error prevents firm conclusions regarding principal effects for students who would, if given the opportunity, request frequent bottom-out hints.
+However, the results are consistent with the hypothesis that ASSISTments is less effective for frequent bottom-out requesters when compared to a similar program without hints or immediate feedback.
+We hypothesize that many students who request bottom-out hints frequently are gaming the system, abetted by the availability of bottom-out hints.
+
 \textsc{geepers} is a flexible, easily-implementable method for principal effect estimation for one-way noncompliance, with predictive covariates; extensions to a broader set of circumstances could be a boon to causal modeling.
@@
 
-\clearpage
-\setcounter{page}{1}
-
-\begin{center}
-\large
-    \textbf{Online Appendices for ``GEEPERs: Principal Stratification using Principal Scores and Stacked Estimating Equations''}
-\end{center}
-
 \appendix
 
-\section{Proofs and Calculations}
-\subsection{Proof for Lemma \ref{lemma:expectation}}
+\section*{Appendix: Proofs and Calculations}
+\subsection*{Proof for Lemma \ref{lemma:expectation}}
 As a preliminary, note that
@@
 
-\subsection{Proof for Proposition \ref{prop:reg1}}
+\subsection*{Proof for Proposition \ref{prop:reg1}}
 
@@
 
-\subsection{A Stronger Version of Proposition \ref{prop:reg2} and a Proof}
+\subsection*{A Stronger Version of Proposition \ref{prop:reg2} and a Proof}
 
 \begin{prop}\label{prop:interactions}
-  Say, for $i=1,\dots,n$, principal scores $\ppi$ are generated as \eqref{eq:pscore}, with parameters $\bm{\alpha}$ identified and estimable with M-estimation, and there exist $\beta_0$, $\beta_1$, $\beta_2$, $\beta_3$, $\bm{\gamma_1}$, $\bm{\gamma_2}$, $\bm{\gamma_3}$ amd $\bm{\gamma_4}$ such that
+  Say, for $i=1,\dots,n$, principal scores $\ppi$ are generated as \eqref{eq:pscore}, with parameters $\bm{\alpha}$ identified and consistently estimable with M-estimation, and there exist $\beta_0$, $\beta_1$, $\beta_2$, $\beta_3$, $\bm{\gamma_1}$, $\bm{\gamma_2}$, $\bm{\gamma_3}$ and $\bm{\gamma_4}$ such that $\{Y_i,Z_i,\sti,\bxy_i\}_{i=1}^n$ are independent and identically distributed with
   \begin{equation}\label{eq:interaction}
@@
   \end{enumerate}
-  Then $\heff0_{G}$ and $\heff1_{int}$ are M-estimators, with $\heff0_{int}\rightarrow_p\eff0$ and $\heff1_{G}\rightarrow_p\eff1$ as $n\rightarrow\infty$.
-  Under suitible regularity conditions, they are jointly asymptotically normal, with a variance of the form \eqref{eq:sandwich}.
+  Then $\heff0_{G}$ and $\heff1_{int}$ are M-estimators.
+If the estimating equations for \eqref{eq:prinEffEstApp} are each bounded by an integrable function of $\{\bm{Y},\bxy, \bm{S},\pp,\bm{Z}\}$ that does not depend on $\{\bm{\beta},\bm{\gamma}\}$, then $\heff0_{int}\rightarrow_p\eff0$ and $\heff1_{int}\rightarrow_p\eff1$ as $n\rightarrow\infty$.
+
+  If the parameter estimates of the principal score model are asymptotically normal, second partial derivatives of the estimating equations for \eqref{eq:prinEffEstApp} are bounded by an integrable function of the data for values of $\{\bm{\beta},\bm{\gamma}\}$ in a neighborhood of their probability limits, and the sandwich components of \eqref{eq:sandwich}, $A$ and $B$, exist and are finite, and if $B$ is non-singular, then $\heff0_{int}$ and $\heff1_{int}$ are jointly asymptotically normal, with a variance of the form \eqref{eq:sandwich}.
 \end{prop}
@@
 These are the estimating equations for the regression model \eqref{eq:interaction}, with $R$ replacing $\st$, fit by OLS. 
-Under standard OLS regularity conditions, the estimated coefficients of that model are consistent, completing the proof.  
+Consistency and asymptotic normality follow from theorems 7.8.1 and 7.8.2, respectively, of \citet{boosStefanskiBook} 
 \end{proof}
 
-\subsection{Sandwich Matrix Calculations}
+\subsection*{Sandwich Matrix Calculations}
 
@@
 Following \eqref{eq:stacked}, let $\blam(Z_i,S_i,\bxsi,\bxy_i,Y_i;\bm{\alpha},\bm{\beta})=\begin{pmatrix} Z_i\Omega(\bxsi,S_i;\bm{\alpha})\\ \Psi(\bxy_i,\bxsi,Y_i,Z_i,\ri(\bm{\alpha});\bm{\beta})\end{pmatrix}$, the stacked estimating equations of \eqref{eq:psMod} and \eqref{eq:regression}.
-Going forward, for the sake of brevety, we will write $\blam_i(\bm{\alpha},\bm{\beta})=\blam(Z_i,S_i,\bxsi,\bxy_i,Y_i;\bm{\hat{\alpha}},\bm{\hat{\beta}})$, where dependence on the data for $i$ is captured in the subscript $i$, with similar meanings for $\Psi_i(\bm{\alpha},\bm{\beta})$ and $\omega_i(\bm{\alpha})$.
+Going forward, for the sake of brevity, we will write $\blam_i(\bm{\alpha},\bm{\beta})=\blam(Z_i,S_i,\bxsi,\bxy_i,Y_i;\bm{\hat{\alpha}},\bm{\hat{\beta}})$, where dependence on the data for $i$ is captured in the subscript $i$, with similar meanings for $\Psi_i(\bm{\alpha},\bm{\beta})$ and $\omega_i(\bm{\alpha})$.
 %Then let $\blamh_i=\blam_i(Z_i,S_i,\bxsi,\bxy_i,Y_i;\bm{\hat{\alpha}},\bm{\hat{\beta}})$ be the estimating equations evaluated at the estimated parameters, and let 
@@
  \end{equation*}
- which capture to the dependence of model \eqref{eq:regression} on the parameters $\bm{\alpha}$ from \eqref{eq:psMod} and the covariance between the estimating equations of the two models. 
+ which capture the dependence of model \eqref{eq:regression} on the parameters $\bm{\alpha}$ from \eqref{eq:psMod} and the covariance between the estimating equations of the two models. 
 The sub-matrix $A_{12}=\sum_i \partial \Omega_i/\partial \bm{\beta}=0$, since \eqref{eq:psMod} does not depend on $\bm{\beta}$.
@@
 
-\subsection{$A_{21}$ Matrix}
+\subsection*{$A_{21}$ Matrix}
 The estimating equations for the regression \eqref{eq:regression} are
@@
 
-\section{Simulation Results}
+\clearpage
+
+\processdelayedfloats
+
+\setcounter{page}{1}
+
+\begin{center}
+\large
+    \textbf{Online Appendices for ``GEEPERs: Principal Stratification using Principal Scores and Stacked Estimating Equations''}
+\end{center}
+\section{Additional Simulation Results}
 \subsection{Plot of  AUC versus $\alpha$}
@@
 \begin{center}
-    \includegraphics[width=0.8\textwidth]{simFigs/alphaAUC.pdf}
+    \includegraphics[width=5.5in]{simFigs/alphaAUC.pdf}
 \end{center}
@@
 %\caption{Coverage of nominal 95\% confidence intervals}
+\clearpage
+
 \subsection{Full Empirical 95\% Interval Coverage Results}
+The following table gives the empirical coverage of nominal 95\% intervals for \textsc{geepers} and mixture model principal effect estimates under varying data generating models.\\
 \input{writeUps/coverageTabAppendix.tex}
@@
 \subsection{Full RMSE Results}
+The following table gives the root mean squared error (RMSE) for \textsc{geepers}, mixture model, and principal score weighting principal effect estimates under varying data generating models.\\
 \input{writeUps/rmseTabAppendix}
@@
 %\end{table}
-
-\section{More Results from Application}
+\clearpage
+
+\section{Additional Results from the Application}
+\singlespacing
 
 \subsection{Summary Statistics}
+The following table gives summary statistics for covariates and post-treatment outcomes in two of the conditions from the empirical study.\\
 \small
-\input{table1}
+\input{tab1fac}
+\clearpage
+
+\small
+\input{writeUps/tab1num}
 
 \subsection{Regression Results}
-Regression estimates from principal score logit model and three outcome regressions for \textsc{geepers} estimates.
+Regression estimates from three principal score logit models and three outcome regressions, based on the "All Covariates" principal score model, for \textsc{geepers} estimates.
 Standard errors shown are nominal regression errors, not sandwich corrected.
-Fixed-effect estimates for school (PS-model) or classroom (outcome models) are omitted.
+Fixed-effect estimates for school (PS-models) or classroom (outcome models) are omitted.
 
 \small
-\input{writeUps/outcomeRegAppendix}
+%\input{writeUps/outcomeRegAppendix}
+\input{writeUps/psTab}
+
+\input{writeUps/regTab}
 

### 2024-12-29 writeUps/paper.tex
update

@@
-\documentclass[]{interact}
+\documentclass[]{article}
 
@@
 \bibpunct[, ]{(}{)}{;}{a}{,}{,}% Citation support using natbib.sty
-\renewcommand\bibfont{\fontsize{10}{12}\selectfont}% To set the list of references in 10 point font using natbib.sty
-
-\theoremstyle{plain}% Theorem-like structures provided by asthma.sty
-\newtheorem{theorem}{Theorem}[section]
-\newtheorem{lemma}[theorem]{Lemma}
-\newtheorem{corollary}[theorem]{Corollary}
-\newtheorem{prop}[theorem]{Proposition}
-
-\theoremstyle{definition}
-\newtheorem{definition}[theorem]{Definition}
-\newtheorem{example}[theorem]{Example}
-
-\theoremstyle{remark}
-\newtheorem{remark}{Remark}
-\newtheorem{notation}{Notation}
+\newcommand\bibfont{\fontsize{10}{12}\selectfont}% To set the list of references in 10 point font using natbib.sty
+
+% \theoremstyle{plain}% Theorem-like structures provided by asthma.sty
+% \newtheorem{theorem}{Theorem}[section]
+% \newtheorem{lemma}[theorem]{Lemma}
+% \newtheorem{corollary}[theorem]{Corollary}
+% \newtheorem{prop}[theorem]{Proposition}
+
+% \theoremstyle{definition}
+% \newtheorem{definition}[theorem]{Definition}
+% \newtheorem{example}[theorem]{Example}
+
+% \theoremstyle{remark}
+% \newtheorem{remark}{Remark}
+% \newtheorem{notation}{Notation}
 
@@
 %\begin{aug}
-\author{
-\name{Adam C. Sales,\thanks{CONTACT Adam C. Sales. Email: asales@wpi.edu} Kirk P. Vanacore, and Erin R. Ottmar}
-\affil{Worcester Polytechnic Institute, Worcester, MA, USA}
-}
-
-\maketitle
+% \author{
+% \name{Adam C. Sales,\thanks{CONTACT Adam C. Sales. Email: asales@wpi.edu} Kirk P. Vanacore, and Erin R. Ottmar}
+% \affil{Worcester Polytechnic Institute, Worcester, MA, USA}
+% }
+
+% \maketitle
 
@@
 \end{abstract}
-\begin{keywords}
-Causal Inference; Principal Stratification; Educational Technology
-\end{keywords}
+% \begin{keywords}
+% Causal Inference; Principal Stratification; Educational Technology
+% \end{keywords}
 %\end{frontmatter}
@@
 That RCT contrasted four different applications for computer-based math practice, one of which included bottom-out hints.
-If that condition benefited students who frequently requested bottom-out hints (``bottom-outers'') more than those who didn't, that may suggest that bottom-out hints can be helpful. 
-If students who avoided bottom-out hints benefited more, it may be because bottom-out hints were harmful.
+If that condition benefited students who frequently requested bottom-out hints (``bottom-outers'') more than those who didn't, that may suggest that the availability of bottom-out hints played a role in the condition's effectiveness.
+If students who avoided bottom-out hints benefited more, it may be because the availability of bottom-out impeded the conditions effectiveness.
 (There are certainly other factors at play as well, so this analysis would not be conclusive.)
@@
 Nonetheless, researchers have devised estimators of principal effects, as average effects within principal strata are called, to address a broad range of applied statistical problems.
-For instance, \citet{air} introduced the framework (or an antecedent thereof) for the case of non-compliance, in which treatment assignment and treatment receipt don't match.
-They applied it to estimate the effects of military service in the American war in Vietnam using the draft lottery, even though some draftees didn't serve and some people served without being drafted, by focusing on the principal stratum of Americans who would serve only if drafted.
-\citet{fellerEtAl2016} estimated differential effects of random assignment to the Head Start early childhood care program for children who, if assigned to control, would receive home-based care, and for those who would attend a different early childhood care program.
+
+The conceptual basis for principal stratification---though not the terminology---emerged from the study of randomized experiments with noncompliance.
+\citet{air}, which gave a novel interpretation of the econometric instrumental variables estimator as the average effect of the treatment on the principal stratum of ``compliers''---those subjects who would receive the treatment if and only if randomized to the treatment condition.
+They showed that the average effect for compliers is nonparametrically identified if a random treatment assignment only has an effect for compliers---``exclusion restriction''---and if treatment assignment never prevents any subjects from receiving treatment (``monotonicity'').
+In a follow-up analysis, \citet{imbens1997bayesian} relaxed the exclusion restriction by way of parametric Bayesian finite mixture modeling, which remains a popular approach.
+
+\citet{frangakis} introduced the terminology of principal stratification and extended the framework well beyond noncompliance, inspiring a broad literature introducing (and debating) novel applications and estimators.
+For instance, \citet{fellerEtAl2016} estimated differential effects of random assignment to the Head Start early childhood care program for children who, if assigned to control, would receive home-based care, and for those who would attend a different early childhood care program.
 Other methodologists have developed estimators for the evaluation of surrogate outcomes \citep{li2010bayesian}, the investigation of causal mechanisms \citep{lidsayPage}, sample attrition \citep{zhangRubin, ding2011}, and other problems.
-
-Because principal strata membership is unobserved, principal effects are not always identified, and when they are, estimation can still be difficult.
-Typically, principal effects are identified and estimated based on strong structural assumptions \citep[e.g.][]{air} and/or parametric modeling \citep[e.g.][]{imbens1997bayesian}.
-Parametric models for principal effects are often hard to specify and hard or impossible to test.
-Even when they are well-specified they can yield estimates that are severely biased  \citep{griffin2008application,feller2016principal}.
+Since interest is often in comparing effects between principal strata---a goal that runs counter to the exclusion restriction---the earliest principal stratification estimators built on \citet{imbens1997}'s parametric mixture models \citep[e.g.][]{Barnard01062003,mealli2004analyzing}, and that approach remains popular \citep{lidsayPage,fellerEtaAl2016}.
+However, these models are often hard to specify and hard or impossible to test.
+Even when they are well-specified they can yield estimates that are severely biased and misleading \citep{griffin2008application,feller2016principal}.
 The most common type of parametric model, the normal mixture model, is especially dubious for studying bottom-outers in \citet{impactPaper}'s RCT, whose outcome, the total score on a 10-item posttest, is not normally distributed.
 
-Another route to identification and estimation of principal effects leads through auxiliary data, including baseline covariates or secondary outcomes \citep{mattei2013exploiting}.
+These concerns have prompted the development of alternative approaches, including bounding \citep{bounding} and randomization inference \citep{nolen2011randomization}.
 Of particular importance is the ``principal score,'' \citep{jo,dingLu,feller2017principal}, or the probability that a subject belongs to a particular principal stratum, as a function of pre-treatment covariates.
-Under a strong, untestable assumption called ``principal ignorability,''  principal scores can be transformed into weights to compute an unbiased estimate of principal effects.
-
-%%% put in paragraph here or above with PS in ed research (lindsay page, feller, my papers)
-
-\citet{jo2002}, \citet{ding2011}, \citet{jiangDing2021}, and others have shown that, under certain circumstances, researchers can use baseline covariates to identify and estimate principal effects without relying on principal ignorability.
-This paper expands on those identification results by presenting a new set of principal effect estimators, and associated standard error estimates, based on general estimating equations (or M-estimation or method-of-moments) \citep{stefanskiBoos}.
-Under certain general conditions, principal effects can be estimated with a simple, three-step procedure: first, estimate principal scores with logistic regression or another M-estimator; then, impute unknown principal subgroup membership indicators with principal scores, and finally estimate principal effects using ordinary least squares (OLS) regression with interactions.
+Under a strong, untestable assumption called ``principal ignorability,''  principal scores can be transformed into weights to compute form an unbiased estimate of principal effects.
+Principal ignorability is a direct analogue of the strong ignorability assumption in causal observational studies \citep[e.g.][]{rosenbaum1986central}---that all covariates that predict both principal stratum membership and outcomes have been accounted for---in other words, the precise type of assumption that randomized trials are designed to avoid.
+A separate set of results in \citet{jo2002}, \citet{ding2011}, \citet{jiangDing2021}, and others have shown that, under certain circumstances, researchers can use baseline covariates to identify and estimate principal effects without relying on principal ignorability.
+
+This paper expands on those latter results by presenting a new principal effect estimator, that boils down to a simple, three-step procedure: first, estimate principal scores, for instance, with logistic regression; then, impute unknown principal subgroup membership indicators with principal scores; and finally estimate principal effects using ordinary least squares (OLS) regression with interactions.
 %, just as in the pre-treatment subgroup case.
+We show that this estimator is the solution to a set of estimating equations based on the combined moments of principal scores, indicators for treatment assignment and observed principal stratum membership in the treatment group, outcomes and (optionally) covariates, and give a set of assumptions under which it is consistent for the true principal effects.
 We will refer to the estimators presented here as ``\textsc{geepers}'': general estimating equations for principal effects using regressions.
 \textsc{geepers} estimates rely on the correct specification of regression functions, but unlike common fully parametric methods, they make no assumptions about the shape of the distribution of regression errors.
-Unlike principal score weighting, \textsc{geepers} does not assume principal ignorability and admits to sandwich-style standard error estimates, which we provide in the appendix.
+Unlike principal score weighting, \textsc{geepers} does not assume principal ignorability and admits to sandwich-style standard error estimates \citep{stefanskiBoos}, which we provide in the appendix.
+
+After the initial submission of this manuscript, we became aware of \citet{richardson2023estimating}, which suggests a similar, though not identical, estimator to ours, using G-estimation of a structural nested mean model \citep[e.g.][]{wallace2017r}.
+Our development of \textsc{geepers} builds on \citet{richardson2023estimating} by emphasizing connections to both parametric mixture modeling and principal scores analysis, and comparing \textsc{geepers} to those methods in a simulation study.
+Moreover, \textsc{geepers}, which, in practice, is a combination of logigistic regression and OLS, is more accessible to applied researchers while yielding very similar results.
 
@@
 %Under ideal circumstances, estimates from parametric mixture models and weighting estimators are somewhat more efficient than \textsc{geepers}, but when assumptions are violated \textsc{geepers} easily out-performs the other two methods.
-Section \ref{sec:fh2t} illustrates \textsc{geepers} by returning to the question of bottom-out hints: does the effect of being randomized to a condition that includes bottom-out hints depend on how often they are requested?
+Section \ref{sec:fh2t} illustrates \textsc{geepers} in two applied examples: first, we re-analyze the applied example from \citet{richardson2023estimating}, a randomized trial of a periodontal treatment for pregnant women in which there was significant one-way noncompliance.
+Second, we return to the question of bottom-out hints: does the effect of being randomized to a condition that includes bottom-out hints depend on how often they are requested?
 Section \ref{sec:discussion} concludes.
@@
 
+\subsection{M-Estimation}
+M-estimation (also generalized method of moments or generalized estimating equations) is a general approach to deriving consistent statistical estimators and their asymptotic distributions.
+In the following section, we will describe \textsc{geepers}, a novel pricipal effect estimator, and derive its properties using the principals of M-estimation.
+Here, we will give a very brief overview of M-estimation, drawn mostly from the excellent treatment in \citet{stefanskiBoos}.
+
+Let $\bm{\mathcal{O}}_i$ be a vector of observed data for subject $i=1,\dots,n$, with $\bm{\mathcal{O}}_i\independent \bm{\mathcal{O}}_{i'}$ if $i\ne i'$, and let $\bm{\theta}$ be a vector of parameters.
+Then, an M-estimator $\bm{\hat{\theta}}$ satisfies ``estimating equations'': for $j=1,\dots,J$, $\sum_i \psi_j\left(\bm{\mathcal{O}}_i,\bm{\hat{\theta}}\right)=0$. 
+When data $\bm{\mathcal{O}}_i$, $i=1,\dots, n$ are identically distributed, $\bm{\hat{\theta}}$ is consistent for $\bm{\theta}$ so long as $\EE\left[\psi_j(\bm{\mathcal{O}}_1,\bm{\theta})\right]=0$ for $j=1,\dots,J$.
+Some common examples are maximum likelihood estimatiors, where $\bm{\psi}(\bm{\mathcal{O}},\bm{\theta})$ is the score function, with $\psi_j(\bm{\mathcal{O}},\bm{\theta})=\partial log f(\bm{\mathcal{O}},\bm{\theta})/\partial \theta_j$ and $f(\cdot,\bm{theta})$ is the density of $\bm{\mathcal{O}}$, or method of moments estimators, in which, if data $\mathcal{O}$ are unidimensional, $\psi_j(\mathcal{O},\bm{\theta})=\mathcal{O}^j-\EE\left[\mathcal{O}^j\right]$.
+
+Under suitable regularity conditions, an M-estimator $\bm{\hat{\theta}}$ is asymptotically normal and its sampling variance-covariance matrix may be estimated using the formula \citep[][ch.
+7]{boosStefanskiBook}:
+\begin{equation}\label{eq:sandwich}
+  \widehat{var}(\bm{\hat{\theta}})=ABA'.
+\end{equation}
+Because of its form, \eqref{eq:sandwich} is sometimes called a ``sandwich'' formula, where $A$ is the ``bread'' and $B$ is the ``meat'' \citep[e.g.][]{sandwich}. 
+To define the matrices $A$ and $B$, let the vector $\bm{\psi}_i=\left\{\psi_j(\bm{\mathcal{O}}_i,\bm{\hat{\theta}})\right\}_{j=1}^J$. Then,
+\begin{equation*}
+  A^{-1}=\sum_i\frac{\partial}{\partial \bm{theta}'} \bm{\psi}_i'
+\end{equation*}
+and
+\begin{equation*}%\label{eq:Bmat}
+  B=\sum_i \bm{\psi}_i\bm{\psi}_i'
+\end{equation*}
+In general, the conditions for asymptotic normality may be difficult to prove; fortunately, they are met for OLS and common generalized linear models such as logistic regression, so long as the dimension of $\bm{\theta}$ does not increase with $n$.
+
+Multiple M-estimators can be combined by ``stacking'' their estimating equations.
+In particular, say an M-estimator with estimating equations $\bm{\psi}^2(\bm{\mathcal{O}},\bm{\hat\theta^2})=\bm{0}$ takes, as input, the results from a previous M-estimator, with estimating equations $\bm{\psi}^1(\bm{\mathcal{O}},\bm{\hat\theta^1})=\bm{0}$.
+The resulting estimator is the solution to the combined estimating equations
+\begin{equation}\label{eq:stacked}
+  \psi^s(\bm{\mathcal{O}},[\bm{\hat\theta^1},\bm{\hat{\theta}^2}])=[\psi^1(\bm{\mathcal{O}},\bm{\hat\theta^1}),\psi^2(\bm{\mathcal{O}},\bm{\hat\theta^2})]=\bm{0},
+\end{equation}
+where the superscript $s$ stands for ``stacked.''
+A variance-covariance matrix estimated using the sandwich formula based on $\psi^s$ accounts for uncertainty from both steps of the estimation procedure.
+
 \section{GEEPERS }\label{sec:geepers}
@@
 
-Then, the standard error matrix for $[\bm{\hat{\alpha}},\bm{\hat{\beta}}]$ can be estimated as \citep[][ch. 7]{boosStefanskiBook}:
-\begin{equation}\label{eq:sandwich}
-  \widehat{var}(\bm{\hat{\beta}})=A^{-1}BA'
-\end{equation}
-Where
-\begin{equation*}
-  A=\sum_i\frac{\partial}{\partial [\alpha,\beta]'} [Z_i\Omega_i,\Psi_i]'
-\end{equation*}
-and
-\begin{equation*}%\label{eq:Bmat}
-  B=\sum_i [Z_i\Omega_i, \Psi_i]'[Z_i\Omega_i, \Psi_i]
-\end{equation*}
-where $[Z_i \Omega_i, \Psi_i]$ is the stacked vector of estimating equations evaluated at $Z_i$, $\bxsi$, $S_i$, and $Y_i$.
-The appendix gives details of this calculation, giving a consistent variance estimator.
-
 \subsection{Relaxing Covariate Independence with Outcome Modeling}\label{sec:regression}
@@
 \url{https://github.com/adamSales/psGee}.
+We replicated each set of conditions 500 times.
+
 \subsubsection{Data Generation}
@@
 with $\beta_3=0.3-\beta_1$, $\EE[\epsilon_i]=0$, and $Var(\epsilon_i)=1/2$.
-When $\gamma_3=0$ (i.e.
-no interaction between $Z$ and covariates), there is no average treatment effect in the $\st=0$ stratum.
-
 The distribution of $\epsilon_i$, $\beta_1$, and values of coefficients $\gamma_1$, $\gamma_2$, and $\gamma_3$ were manipulated factors in the study.
+As described below, in most simulated scenarios, $\gamma_1=1/\sqrt{6}$ and $\gamma_2=\gamma_3=0$.
+In these cases there is no average treatment effect in the $\st=0$ stratum, and, since  $Var(\gamma_1(x_1+x_2+x_3))=1/2$ and $Var(\epsilon_i)=1/2$, covariates explained half of the variance of $Y$ within treatment group and principal stratum, and observed covariates $x_1$ and $x_2$ explained 1/3 of that variance.
+These values are in line with estimates from outcome models in Section \ref{sec:fh26} and are typical in education contexts \cite{hedgesHedberg}.
+
+Treatment assignment $Z$ was simulated as independent of all the other variables with half of subjects assigned to treatment $Z=1$ and the rest to control $Z=0$.
 
@@
 \textbf{\emph{Sample Size per condition $\bm{n}$:}} The guarantees of M-estimation are all asymptotic, so we examined the behavior of our \textsc{geepers} at a range of sample sizes to determine its finite-sample properties.
+The range of 100--1000 was chosen to represent typical student-level randomized trials in education, for instance, \citet{growthMindsetRuralBurnette} with $n=115$ in the treatment group and $n=107$ in the control group, or \citet{impactPaper}, which initially randomized an average of 1000 students to each condition.
 
@@
 When neither interaction was present, $\gamma_2=\gamma_3=0$, then $\gamma_1=1/\sqrt{6}$.
-Then, since  $Var(\gamma_1(x_1+x_2+x_3))=1/2$ and $Var(\epsilon_i)=1/2$, covariates explained half of the variance of $Y$ within treatment group and principal stratum, and observed covariates $x_1$ and $x_2$ explained 1/3 of that variance.
 When there was an interaction between covariates and $\sti$, then $\gamma_1=3/(4\sqrt{6})$ and $\gamma_2=1/(2\sqrt{6})$, so that the slopes for $x_1$ and $x_2$ varied by half their magnitude between principal strata, and on average they were equal to $1/\sqrt{6}$, as in the no-interaction case.
@@
 
-\section{Application: Estimating Treatment Effects for Implementation Subgroups: Students who Frequently or Rarely Request Bottom-Out Hints}\label{sec:fh2t}
+\section{Applications}
+To illustrate \textsc{geepers} and compare it to the other discussed methods, we include two applied examples.
+First, we briefly replicate an analysis from \citet{richardson2023estimating} using the \textsc{geepers} and the other methods discussed above.
+This is a relatively straightforward example of an experiment with imperfect compliance in which \citet{richardson2023estimating} estimated stastistically significant effects overall in both principal strata.
+Next, we present a novel analysis of data from the RCT of computer-based math applications reported in \citet{impactPaper}, and look into the role of bottom-out hints, as described in the introduction.
+This analysis is intended to shed light on an important question in education research, and provides an opportunity to raise questions of model and covariate selection and the plausability of identification assumptions.
+
+\subsection{Accounting for Imperfect Compliance: Re-Analysis of the Obstetrics and Periodontal Therapy Study}
+Peridontal disease, a weakening of bone and connective tissue supporting teeth, is highly prevelant among American adults, and is associated with adverse pregnancy outcomes, including preterm birth \citep{periodontalEpi}.
+The Obstetrics and Periodontal Therapy Study \citep[(OPTS)][]{michalowicz2006treatment} investigated the efficacy of nonsurgical periodontal treatment to reduce the preterm birth rate; replication data are available at \url{https://causeweb.org/tshs/obstetrics-and-periodontal-therapy/}.
+Pregnant patients at four different hospitals who were at least 16 years old, at less than 16 weeks and 6 days of gestation, and diagnosed with peridontal disease were individually randomized to either receive peridontal treatment during the first 21 weeks of pregnancy ($Z=1$, $n_T=413$) or after delivery ($Z=0$, $n_C=410$).
+The primary prespecified outcome of the study was babies' gestational age at birth, but the study team also collected data on several secondary birth outcomes, as well as measures of patients' peridontal disease at follow-up.
+
+The main component of the intervention took place over the course of up to four visits, followed by monthly tooth polishing and other dental care as needed, and not all participants randomized to the treatment condition completed the entire treatment plan.
+However, 96\% of patients in the treatment group received at least some periodontal care during pregnancy---therefore, there is reason to suspect nonzero treatment effects even for those patients who did not (or would not) fully comply.
+Patients randomized to the control condition did not have access to the intervention.
+\citet{richardson2023estimating} used a method they refer to as ``bespoke instrumental variable'' (BSIV) to estimate the average effects for the principal stratum of subjects who, if randomized to treatment,  would complete the treatment plan and for the principal stratum of subjects who would not.
+Here, we will use \textsc{geepers}, a Bayesian mixture model, and principal score weighting to estimate the same quantities, following the same analysis decisions as \citet{richardson2023estimating}.
+Replication code for our analysis can be found at \url{}.
+
+Following \citep{richardson2023estimating}, we will estimate the principal effects of randomization $Z$ on $Y$, the percentage of probed sites (6 sites per tooth) that were bleeding at visit 5, which was at 29--32 weeks gestation, conditioning on two covariates, Serum measures of endotoxin level and fibrinogen at baseline.
+We omit the 164 cases in which $Y$ was missing as well as 19 additional cases in which one or both of the covariates were missing, leaving a total sample size of 640, with 314 randomized to the treatment condition and 326 randomized to control.
+The compliance variable in the dataset had three categories: ``Yes,'' for subjects in the treatment group who completed the treatment plan (157 subjects), ``No'' for those who did not (4 subjects), and ``Und'' for those who received some therapy, but it is unknown whether they completed the treatment (157 subjects).
+We let $S=1$ when compliance was ``yes''---i.e. when subjects in the treatment group completed the treatment---and $S=0$ otherwise.
+
+For all three estimation methods, we estimated principal scores with logistic regression of $\st$ on the two covariates.
+For \textsc{geepers} and \textsc{psw}, we used standard maximum likelihood methods and data from the treatment group, achieving an AUC of 0.68; for the mixture model we incorporated the logistic regression as a sub-model within a larger Bayesian model fit in Stan.
+
+The results are displayed in Figure \ref{fig:opt} and Table \ref{tab:optEffs}: both \textsc{geepers} and the Bayesian mixture model gave similar results to those reported in \citet{richardson2023estimating} from BSIV.
+For the principal stratum of subjects who would not fully complete the treatment, those three methods estimated that assignment to the treatment condition reduced percentage of dental sites with bleeding by 31--34 percentage points; for subjects who would complete the treatment, the three estimated a reduction of 14-16 percentage points.
+These methods' conclusion is somewhat surprising: that the effect was larger for subjects who would not complete the treatment than for those who would (the two strata had similar bleeding percentages at baseline, and a logit transformation of the outcome leave the results qualitatively unchanged).
+One possible explanation is the conjecture that subjects stop showing up for treatment once their periodontal disease has abated.
+
+In contrast, prinicpal score weighting resulted in very similar effects for both strata---an reduction of 23--25 percentage points, which is close to the estimated intent to treat effect of 25 percentage points.
+
+\subsection{Estimating Treatment Effects for Implementation Subgroups: Students who Frequently or Rarely Request Bottom-Out Hints}\label{sec:fh2t}
 
@@
 
+\section{Additional Results from the OPT Study}
+\singlespacing
+\subsection{Summary Statistics}
+The following table gives summary statistics for covariates and post-treatment outcomes in two of the conditions from the empirical study.\\
+
+\begin{table}
+
+\caption{\label{tab:tab:optTab1}Descriptive statistics---mean and standard deviation or count and percent---for study variables in the full OPT dataset and in the analysis sample (i.e.
+complete cases)}
+\centering
+\begin{tabular}[t]{lllll}
+\toprule
+\multicolumn{1}{c}{ } & \multicolumn{2}{c}{Full Data} & \multicolumn{2}{c}{Complete Cases} \\
+\cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5}
+  & Control & Treatment & Control & Treatment\\
+\midrule
+n & 410 & 413 & 326 & 314\\
+Fibrinogen & 21.3 (14.9) & 20.2 (13.3) & 21.5 (15.4) & 20 (13.1)\\
+Endotoxin & 1.8 (1.1) & 1.8 (1.1) & 1.8 (1.1) & 1.8 (1.1)\\
+\% Sites Bleeding & 67.3 (21.1) & 43.9 (20.4) & 67.6 (21.3) & 43.5 (20.3)\\
+Trt. Completed: No & - & 14 (3.4\%) & - & 4 (1.3\%)\\
+\addlinespace
+Trt. Completed: Und & - & 196 (47.5\%) & - & 153 (48.7\%)\\
+Trt. Completed: Yes & - & 185 (44.8\%) & - & 157 (50\%)\\
+\bottomrule
+\end{tabular}
+\end{table}
+
 \section{Additional Results from the Application}
 \singlespacing
+The following table gives summary statistics for covariates and post-treatment outcomes in two of the conditions from the empirical study.\\
 

### 2024-12-29 writeUps/regression.tex
update

@@
 
-Let $p_i=Pr(S_i=1|\bm{x}_i)$.
+Let $p_i=Pr(S_i=1|\bm{w}_i)$, where $\bm{w}_i$ is a vector of covariates that (putatively) predict $S_i$.
+(Really this should be $S_{Ti}$, cuz we're doing principal stratification; we'll drop the $T$ subscript for clarity.)
+
 Then the original estimating equations were:
@@
 \section{Missing $S$ in Treatment Group}
-When $S$ is missing at random (conditional on $\bm{x}$) for some observations in the treatment group,
+When $S$ is missing at random (conditional on $\bm{w}$) for some observations in the treatment group,
 and $\mathbb{E}[Y|Z=1,S,p]=\mathbb{E}[Y|Z=1,S]$,
@@
 \begin{equation}\label{eq:psMod}
-  Pr(S_{Ti}=1|\bm{\tilde{x}}_i)=f(\bm{\alpha}'\bm{\tilde{x}}_i)
+  Pr(S_{Ti}=1|\bm{\tilde{w}}_i)=f(\bm{\alpha}'\bm{\tilde{w}}_i)
 \end{equation}
-for some inverse link function $f(\cdot)$, where $\bm{\tilde{x}}_i=[1,\bm{x}_i]$, using (observed) values from the treatment group, and estimating $\hat{\alpha}$.
+for some inverse link function $f(\cdot)$, where $\bm{\tilde{x}}_i=[1,\bm{w}_i]$, using (observed) values from the treatment group, and estimating $\hat{\alpha}$.
 Then let
 \begin{equation}\label{eq:ps}
-  \hat{p}_i=f(\bm{\hat{\alpha}}'\bm{\tilde{x}}_i)
+  \hat{p}_i=f(\bm{\hat{\alpha}}'\bm{\tilde{w}}_i)
 \end{equation}
@@
 \end{equation}
-to estimate $\bm{\beta}$ and hence principal effects.
+to estimate $\bm{\beta}$ and hence principal effects, where $\bm{x}_i$ is a set of covariates predictive of $Y$ within principal strata.
 [Note that \eqref{eq:regression} is just \eqref{eq:regression0} with $\bm{\gamma}^t\bm{x}_i$ added on the end].
+$\bm{x}_i$ should, generally, contain $\bm{w}_i$; the two may also be equivalent.
 Let $\bm{\beta}=[\beta_0,\beta_1,\beta_2,\beta_3,\bm{\gamma}^t]^t$.
@@
 \begin{equation}\label{eq:Amat}
-  A=\sum_i\frac{\partial}{\partial [\alpha,\beta]^t} \Psi(Y_i;\alpha,\beta,\bm{x}_i)
+  A=\sum_i\frac{\partial}{\partial [\alpha,\beta]^t} \Psi(Y_i;\alpha,\beta,\bm{x}_i,\bm{w}_i)
 \end{equation}
@@
 \begin{equation}\label{eq:Bmat}
-  B=\sum_i \Psi(Y_i;\alpha,\beta,\bm{x}_i)\Psi(Y_i;\alpha,\beta,\bm{x}_i)^t
+  B=\sum_i \Psi(Y_i;\alpha,\beta,\bm{x}_i)\Psi(Y_i;\alpha,\beta,\bm{x}_i,\bm{w}_i)^t
 \end{equation}
@@
 \begin{equation}\label{eq:eeOLS}
-  \psi(Y_i,\bm{x}_i,\beta,\alpha)=X_iY_i-X_iX_i^t\beta
+  \psi(Y_i,\bm{x}_i,\bm{w}_i,\beta,\alpha)=X_iY_i-X_iX_i^t\beta
 \end{equation}
 Where $X_i=[1,r_i,Z_i,r_iZ_i,\bm{x}_i^t]^t$.
+In other words,
+\begin{align*}
+  \psi(Y_i,&\bm{x}_i,\bm{w}_i,\beta,\alpha)=\\
+  &\left\{
+  \begin{array}{l}
+    Y_i-\left(\beta_0+\beta_1r_i+\beta_2Z_i+\beta_3Z_ir_i+\bm{\gamma}'\bm{x}_i\right)\\
+    r_iY_i-r_i\left(\beta_0+\beta_2Z_i+\bm{\gamma}'\bm{x}_i\right)-r_i^2\left(\beta_1+\beta_3Z_i\right)\\
+    Z_iY_i-Z_i\left(\beta_0+\beta_1r_i+\beta_2Z_i+\beta_3Z_ir_i+\bm{\gamma}'\bm{x}_i\right)\\
+    Z_ir_iY_i-Z_ir_i\left(\beta_0+\beta_2Z_i+\bm{\gamma}'\bm{x}_i\right)-Z_ir_i^2\left(\beta_1+\beta_3Z_i\right)\\
+    \bm{x}_iY_i-\bm{x}_i\left(\beta_0+\beta_1r_i+\beta_2Z_i+\beta_3Z_ir_i+\bm{\gamma}'\bm{x}_i\right)
+  \end{array}
+  \right\}
+\end{align*}
+(noting that $Z^2=Z$).
 These depend on $\alpha$ when $r_i=p_i$, i.e. when $Z_i=0$ or $S_i$ is missing.
-Then note that, following \eqref{eq:ps}, and letting $\eta_i=\alpha^t\bm{\tilde{x}}_i$
+Then note that, following \eqref{eq:ps}, and letting $\eta_i=\alpha^t\bm{\tilde{w}}_i$
 \begin{equation}\label{eq:derivP}
-  \frac{\partial p_i}{\partial \alpha^t}=f'(\eta_i)\frac{\partial \eta_i}{\partial \alpha^t}=f'(\eta_i)\bm{\tilde{x}}_i^t
+  \frac{\partial p_i}{\partial \alpha^t}=f'(\eta_i)\frac{\partial \eta_i}{\partial \alpha^t}=f'(\eta_i)\bm{\tilde{w}}_i^t
 \end{equation}
@@
 \begin{equation}
-  \frac{\partial p_i^2}{\partial \alpha^t}=2p\frac{\partial p_i}{\partial \alpha^t}=2f(\eta)f'(\eta_i)\bm{\tilde{x}}_i^t
+  \frac{\partial p_i^2}{\partial \alpha^t}=2p_i\frac{\partial p_i}{\partial \alpha^t}=2f(\eta)f'(\eta_i)\bm{\tilde{w}}_i^t=2p_if'(\eta_i)\bm{\tilde{w}}_i^t
 \end{equation}
@@
   &\left[\begin{array}{c}
-          -\beta_1\frac{\partial p_i}{\partial \alpha^t}\\
+          -(\beta_1+\beta_3Z_i)\frac{\partial p_i}{\partial \alpha^t}\\
           (Y_i-\beta_0-\beta_2Z_i-\bm{\gamma}^t\bm{x}_{i})\partial p/\partial\alpha^t-(\beta_1+\beta_3Z_i)\partial p^2/\partial \alpha^t\\
-          -\beta_1\frac{\partial p_i}{\partial \alpha^t}Z_i\\
+          -(\beta_1+\beta_3Z_i)\frac{\partial p_i}{\partial \alpha^t}Z_i\\
           Z_i\left[(Y_i-\beta_0-\beta_2Z_i-\bm{\gamma}^t\bm{x}_{i})\partial p/\partial\alpha^t-(\beta_1+\beta_3Z_i)\partial p^2/\partial \alpha^t\right]\\
-          -\beta_1\frac{\partial p_i}{\partial \alpha^t}\bm{x}_i
+          -(\beta_1+\beta_3Z_i)\frac{\partial p_i}{\partial \alpha^t}\bm{x}_i
         \end{array}\right]\\
   =& \left[\begin{array}{c}
-          -\beta_1f'(\eta_i)\bm{\tilde{x}}_i^t\\
-          \left[Y_i-X_i^t\beta-p_i(\beta_1+\beta_3Z_i)\right]f'(\eta_i)\bm{\tilde{x}}_i\\
-          -Z_i\beta_1f'(\eta_i)\bm{\tilde{x}}_i^t\\
-          Z_i\left[Y_i-X_i^t\beta-p_i(\beta_1+\beta_3)\right]f'(\eta_i)\bm{\tilde{x}}_i\\
-          -\beta_1f'(\eta_i)\bm{x}_i\bm{x}_i^t
+          -(\beta_1+\beta_3Z_i)f'(\eta_i)\bm{\tilde{w}}_i^t\\
+          \left[Y_i-\beta_0-\beta_2Z_i-\bm{x}_i^t\bm{\gamma}-2p_i(\beta_1+\beta_3Z_i)\right]f'(\eta_i)\bm{\tilde{w}}_i^t\\
+          -Z_i\beta_1f'(\eta_i)\bm{\tilde{w}}_i^t\\
+          Z_i\left[Y_i-\beta_0-\beta_2Z_i-\bm{x}_i^t\bm{\gamma}-2p_i(\beta_1+\beta_3)\right]f'(\eta_i)\bm{\tilde{w}}_i^t\\
+          -(\beta_1+\beta_3Z_i)f'(\eta_i)\bm{x}_i\bm{\tilde{w}}_i^t
     \end{array}\right]
 \end{align*}
-Note that unless there is missing $S$ in the treatment group, $Z_i=0$ whenever $r_i=p_i$, so the 3rd and 4th lines of the matrix are all 0.
 
-If $r_i=S_i$, $\frac{\partial}{\partial \alpha^t}\text{ \& }\psi(Y_i,\bm{x}_i,\beta,\alpha)=0$.
+When $S$ is observed for all members of the treatment group, $Z_i=0$ whenever $r_i=p_i$, so the latter expression reduces to
+\begin{align*}
+  \frac{\partial}{\partial \alpha^t}&\psi(Y_i,\bm{x}_i,\beta,\alpha)=\\
+  & \left[\begin{array}{c}
+          -(\beta_1)f'(\eta_i)\bm{\tilde{w}}_i^t\\
+          \left[Y_i-X_i^t\beta-2p_i(\beta_1)\right]f'(\eta_i)\bm{\tilde{w}}_i^t\\
+          0\\
+          0\\
+          -\beta_1f'(\eta_i)\bm{x}_i\bm{\tilde{w}}_i^t
+    \end{array}\right]
+\end{align*}
+
+If $r_i=S_i$, $\frac{\partial}{\partial \alpha^t}\psi(Y_i,\bm{x}_i,\beta,\alpha)=0$.
 

### 2024-12-29 writeUps/rmseTab.tex
update

@@
 \cmidrule(l{3pt}r{3pt}){4-6} \cmidrule(l{3pt}r{3pt}){7-9} \cmidrule(l{3pt}r{3pt}){10-12}
-\makecell[l]{Residual\\Dist.} & \makecell[c]{X:Z\\Int.?} & \makecell[r]{X:S\\Int.?} & GEEPERs & Mix. & PSW & GEEPERs & Mix. & PSW & GEEPERs & Mix. & PSW\\
+\makecell[l]{Residual\\Dist.} & \makecell[c]{X:Z\\Int.?} & \makecell[r]{X:S\\Int.?} & \rotatebox[origin=c]{300}{GEEPERs} & \rotatebox[origin=c]{300}{Mixture} & \rotatebox[origin=c]{300}{PSW} & \rotatebox[origin=c]{300}{GEEPERs} & \rotatebox[origin=c]{300}{Mixture} & \rotatebox[origin=c]{300}{PSW}&\rotatebox[origin=c]{300}{GEEPERs} & \rotatebox[origin=c]{300}{Mixture} & \rotatebox[origin=c]{300}{PSW}\\
 \midrule

### 2024-12-29 code/regression.r

@@ -127,22 +127,22 @@ EFps <- function(psMod,data){
 sandwichMats <- function(psMod,outMod,data,clust=NULL,int=any(grepl(":x",names(coef(outMod))))){
     #data <- model.frame(outMod)
     ### estimating equations
-    efPS <- EFps(psMod,data)
-    efOut <- estfun(outMod)
+    efPS <- EFps(psMod,data) ## psi1
+    efOut <- estfun(outMod) ## psi2
 
     out <- list(
-        a11inv = bread(psMod),#/sum(data$Z),
-        a22inv = bread(outMod),#/nrow(data),
-        a21 = A21(psMod,outMod,data),#/nrow(data),
-        b11=(if(is.null(clust)) meat(psMod) else meatCL(psMod,cluster=clust[data$Z==1])),#*nrow(model.frame(psMod)),
-        b22 = (if(is.null(clust)) meat(outMod) else meatCL(outMod,cluster=clust))#*nrow(data)#,adjust=TRUE)
+        a11inv = bread(psMod)/sum(data$Z),
+        a22inv = bread(outMod)/nrow(data),
+        a21 = A21(psMod,outMod,data)/nrow(data),
+        b11=(if(is.null(clust)) meat(psMod) else meatCL(psMod,cluster=clust[data$Z==1]))*nrow(model.frame(psMod)),
+        b22 = (if(is.null(clust)) meat(outMod) else meatCL(outMod,cluster=clust))*nrow(data)#,adjust=TRUE)
     )
     out <- within(out,b12 <-
         if(int){
             matrix(0,nrow(b11),ncol(b22))
         } else if(is.null(clust)) crossprod(efPS,efOut)/nrow(data)
         else crossprod(apply(efPS,2L,rowsum,clust),
-                       apply(efOut,2L,rowsum,clust))#/sum(data$Z==0)
+                       apply(efOut,2L,rowsum,clust))/sum(data$Z==0)
         )
     out
 }
@@ -175,7 +175,7 @@ vcvPS <- function(psMod,outMod,data,clust=NULL,int=any(grepl(":x",names(coef(out
 
     n <- nrow(model.frame(outMod))
 
-    vcvFull <- solve(A)%*%B%*%t(solve(A))/nrow(data)
[…]

### 2024-12-29 writeUps/paper.tex
update2

@@
   \centering
-  \includegraphics{prinEffs.pdf}
+  \includegraphics{figure/prinEffs.pdf}
   \caption{Principal effect estimates with 95\% confidence intervals using three different principal score models, alongside ATE estimates and 95\% confidence intervals, for contrasts between the ASSISTments condition and each of the three other conditions in the study.}
@@
   \centering
-  \includegraphics{compareMethods.pdf}
+  \includegraphics{figure/compareMethods.pdf}
   \caption{Principal effect estimates with 95\% confidence or credible intervals using three different principal stratification methods, alongside ATE estimates and 95\% confidence intervals, for contrasts between the ASSISTments condition and each of the three other conditions in the study.
@@
 \begin{center}
-    \includegraphics[width=5.5in]{simFigs/alphaAUC.pdf}
+    \includegraphics[width=5.5in]{../simFigs/alphaAUC.pdf}
 \end{center}
@@
 The following table gives the empirical coverage of nominal 95\% intervals for \textsc{geepers} and mixture model principal effect estimates under varying data generating models.\\
-\input{writeUps/coverageTabAppendix.tex}
+\input{coverageTabAppendix.tex}
 
@@
 The following table gives the root mean squared error (RMSE) for \textsc{geepers}, mixture model, and principal score weighting principal effect estimates under varying data generating models.\\
-\input{writeUps/rmseTabAppendix}
+\input{rmseTabAppendix}
 
@@
 
-\section{Additional Results from the Application}
-\singlespacing
-The following table gives summary statistics for covariates and post-treatment outcomes in two of the conditions from the empirical study.\\
-
-\subsection{Summary Statistics}
-The following table gives summary statistics for covariates and post-treatment outcomes in two of the conditions from the empirical study.\\
-\small
-\input{tab1fac}
-\clearpage
-
-\small
-\input{writeUps/tab1num}
-
-\subsection{Regression Results}
-Regression estimates from three principal score logit models and three outcome regressions, based on the "All Covariates" principal score model, for \textsc{geepers} estimates.
+% \section{Additional Results from the Application}
+% \singlespacing
+% The following table gives summary statistics for covariates and post-treatment outcomes in two of the conditions from the empirical study.\\
+
+% \subsection{Summary Statistics}
+% The following table gives summary statistics for covariates and post-treatment outcomes in two of the conditions from the empirical study.\\
+% \small
+% \input{tab1fac}
+% \clearpage
+
+% \small
+% \input{tab1num}
+
+% \subsection{Regression Results}
+% Regression estimates from three principal score logit models and three outcome regressions, based on the "All Covariates" principal score model, for \textsc{geepers} estimates.
 Standard errors shown are nominal regression errors, not sandwich corrected.
@@
 
-\small
-%\input{writeUps/outcomeRegAppendix}
-\input{writeUps/psTab}
-
-\input{writeUps/regTab}
+% \small
+% %\input{writeUps/outcomeRegAppendix}
+% \input{psTab}
+
+% \input{regTab}
 

### 2024-12-29 code/OPTS/opts.r

@@ -1,0 +1,249 @@
+library(rstan)
+library(dplyr)
+library(tibble)
+library(ggplot2)
+library(sandwich)
+library(lmtest)
+library(tableone)
+library(purrr)
+library(kableExtra)
+library(tidyr)
+
+rstan_options(auto_write = FALSE)
+
+
+source('code/regression.r')
+options(mc.cores = 2)
+
+auc <- function(x,y)
+  if(length(unique(y))==2 & length(y)==length(x)){
+    wilcox.test(x~y)$statistic/(sum(y)*(sum(1-y)))
+  }else
+    wilcox.test(x,y)$statistic/(legnth(x)*length(y))
+
+aucMod <- function(mod)
+  auc(mod$linear,1-mod$y)
+
+
+print(load('code/OPTS/OPT_Study_PersonLevel_Data.RData'))
+## downloaded from https://www.causeweb.org/tshs/obstetrics-and-periodontal-therapy/
+
+
+
+opt <- opt%>%
+  mutate(OFIBRIN1=as.numeric(as.character(OFIBRIN1)),
+         ETXU_CAT1=as.numeric(as.character(ETXU_CAT1)))
+
+levels(opt$Group)=c(T="Treatment",C="Control")[levels(opt$Group)]
+
+xtabs(~is.na(OFIBRIN1)+is.na(ETXU_CAT1)+is.na(V5..BOP),data=opt)
+
+sum(is.na(opt$V5..BOP))
+sum((is.na(opt$ETXU_CAT1)|is.na(opt$OFIBRIN1))&!is.na(opt$V5..BOP))
+
+cca <- opt%>%filter(!is.na(OFIBRIN1),!is.na(ETXU_CAT1),!is.na(V5..BOP))%>%
+    #select(Group, OFIBRIN1,ETXU_CAT1,V5..BOP,Tx.comp.,Completed.EDC,BL..BOP)%>%
+    mutate(S=Tx.comp.=='Yes',
+         Y=V5..BOP/100,
+         Z=Group=='Treatment')
+
+tab1 <- map(list(`Full Data`=opt,`Complete Cases`=cca),
+            function(dat)
+                dat%>%
+                mutate(Data=if(any(is.na(V5..BOP))) "Full Data" else "Complete Cases")%>%
+    #rename("Trt. Completed"=Tx.comp.)%>%
+  group_by(Data,Group)%>%
[…]

### 2025-01-09 writeUps/paper.tex
paper edits

--- 2024-12-29 writeUps/paper.tex
@@
 \documentclass[]{article}
+\usepackage[letterpaper]{geometry}
 
@@
 \bibpunct[, ]{(}{)}{;}{a}{,}{,}% Citation support using natbib.sty
-\newcommand\bibfont{\fontsize{10}{12}\selectfont}% To set the list of references in 10 point font using natbib.sty
+\renewcommand\bibfont{\fontsize{10}{12}\selectfont}% To set the list of references in 10 point font using natbib.sty
 
@@
 
+\maketitle
+
 \begin{abstract}
@@
 % \end{keywords}
-%\end{frontmatter}
 
@@
 \section{Introduction}
-
-%Randomized experiments easily admit subgroup analysis---estimation of treatment effects for a subset of the study population---when those subgroups are defined at baseline, and not themselves a function of treatment assignment.
-%A simple, popular approach to estimating subgroup effects is to use ordinary least squares (OLS) regression, regressing the outcome on a treatment indicator interacted with an indicator for subgroup membership.
-%However, when subgroup membership is a function of a post-treatment variable, then it may itself be affected by the treatment.
-%Contrasts between treatment and control conditions within those subgroups may not have a causal interpretation at all---and when they do, the same techniques that work for subgroups defined at baseline %(such as OLS with interactions)
-%will often produce inconsistent estimates.
-
-%%% open with hook for ed researchers
 
@@
 Other methodologists have developed estimators for the evaluation of surrogate outcomes \citep{li2010bayesian}, the investigation of causal mechanisms \citep{lidsayPage}, sample attrition \citep{zhangRubin, ding2011}, and other problems.
-Since interest is often in comparing effects between principal strata---a goal that runs counter to the exclusion restriction---the earliest principal stratification estimators built on \citet{imbens1997}'s parametric mixture models \citep[e.g.][]{Barnard01062003,mealli2004analyzing}, and that approach remains popular \citep{lidsayPage,fellerEtaAl2016}.
+Since interest is often in comparing effects between principal strata---a goal that runs counter to the exclusion restriction---the earliest principal stratification estimators built on \citet{imbens1997bayesian}'s parametric mixture models \citep[e.g.][]{Barnard01062003,mealli2004analyzing}, and that approach remains popular \citep{lidsayPage,fellerEtAl2016}.
 However, these models are often hard to specify and hard or impossible to test.
@@
 Under a strong, untestable assumption called ``principal ignorability,''  principal scores can be transformed into weights to compute form an unbiased estimate of principal effects.
-Principal ignorability is a direct analogue of the strong ignorability assumption in causal observational studies \citep[e.g.][]{rosenbaum1986central}---that all covariates that predict both principal stratum membership and outcomes have been accounted for---in other words, the precise type of assumption that randomized trials are designed to avoid.
+Principal ignorability is a direct analogue of the strong ignorability assumption in causal observational studies \citep[e.g.][]{rosenbaum1983central}---that all covariates that predict both principal stratum membership and outcomes have been accounted for---in other words, the precise type of assumption that randomized trials are designed to avoid.
 A separate set of results in \citet{jo2002}, \citet{ding2011}, \citet{jiangDing2021}, and others have shown that, under certain circumstances, researchers can use baseline covariates to identify and estimate principal effects without relying on principal ignorability.
@@
 The simulation demonstrates the strengths and weaknesses of each method and the tradeoffs between them.
-
-This paper focuses on a special case of principal stratification---one-way, binary non-compliance---that is, there are only two principal strata and principal stratum membership is observable in one treatment arm, but not in the other.
-This scenario would occur if, say, subjects in the control arm of a study had no access to the treatment, and not all subjects assigned to the treatment arm received the full treatment.
-While such a setup may represent the simplest and easiest principal stratification problem, we believe it is a common scenario.
-For example, \textsc{geepers} could be used to estimate effects from an RCT for participants categorized as low- or high-implementers.
-It could also be used to assess more granular questions about effect heterogeneity, such as whether an intervention's impact is contingent upon specific behaviors displayed by either practitioners or students.
 
@@
 In general, $(1-Z_i)\ppi=(1-Z_i)\ri$ and $Z_i\sti=Z_i\ri$, allowing $\ri$ to take the place of $\ppi$ and $S_i$ throughout $\tilde{\Psi}_i$.
+Denote estimating equations based on $\ri$ as $\sum_i \psi_i=\bm{0}$.
 
@@
 %However,
-If the principal scores are modeled as \eqref{eq:prinScoreMod}, and parameters $\bm{\alpha}$ are estimated by M-estimation, with estimating equations $\sum_{i: Z_i=1}\Omega_i=\bm{0}$, then the estimating equations for the principal score model and principal effect estimation can be stacked \citep[c.f.][]{boosStefanskiBook}, as
-\begin{equation}\label{eq:stacked}
-  \sum_{i=1}^n \begin{pmatrix}
-    Z_i\Omega_i(\bxsi,S_i;\bm{\alpha})\\
-    \Psi_i\left(Y_i,Z_i,\ri(\bm{\alpha}); \bm{\beta}\right)\end{pmatrix}=\begin{pmatrix} \bm{0}\\\bm{0}\end{pmatrix}
-\end{equation}
-where $\ri\left(\bm{\alpha}\right)=S_iZ_i+\ppip{\bm{\alpha}}(1-Z_i)$ emphasizes $\ppi$'s dependence on $\bm{\alpha}$.
-In practice, the principal score model \eqref{eq:prinScoreMod} could be fit first, yielding estimates $\bm{\hat{\alpha}}$, and model \eqref{eq:regression0} could be fit next, using $\ri=\ri\left(\bm{\hat{\alpha}}\right)$.
-Together, the vector of estimates $[\bm{\hat{\alpha}},\bm{\hat{\beta}}]$ represents a zero of the stacked estimating equations \eqref{eq:stacked}.
+If the principal scores are modeled as \eqref{eq:prinScoreMod}, and parameters $\bm{\alpha}$ are estimated by M-estimation, with estimating equations $\sum_i Z_i\Omega_i=\bm{0}$, then the estimating equations for the principal score model and principal effect estimation can be stacked as in \eqref{eq:stacked}.
+Sandwich standard error estimates of the form \eqref{sandwich} based on $\left[Z_i\omega_i,\psi_i\right]$ will incorporate uncertainty from both principal score modeling and principal effect estimation conditional on principal scores.
 
@@
 In any event, Assumption \ref{ass:rci} allows principal effects to be estimated with an OLS model including $\bxy$:
-
 \begin{prop}{\textsc{geepers}}\label{prop:reg2}
@@
 In these cases there is no average treatment effect in the $\st=0$ stratum, and, since  $Var(\gamma_1(x_1+x_2+x_3))=1/2$ and $Var(\epsilon_i)=1/2$, covariates explained half of the variance of $Y$ within treatment group and principal stratum, and observed covariates $x_1$ and $x_2$ explained 1/3 of that variance.
-These values are in line with estimates from outcome models in Section \ref{sec:fh26} and are typical in education contexts \cite{hedgesHedberg}.
+These values are in line with estimates from outcome models in Section \ref{sec:fh2t} and are typical in education contexts \cite{hedgesHedberg}.
 
@@
 
-\subsection{Data Analysis}
+\subsubsection{Data Analysis}
 We estimated principal scores using logistic regression of observed $S$ on school fixed effects and a set of baseline variables $\bxs$.
@@
 Those considerations militate in favor of choosing covariates $\bxs$ to optimize out-of-sample principal score prediction accuracy.
+
 There is good reason to believe that analogous arguments apply to selection of outcome covariates $\bxy$.
@@
 In our main model, we included every available covariate in the dataset and chose transformations of those covariates based on the results of residual and binned residual plots \citep{arm}.
+
 Including all possible covariates maximizes the model's opportunities to detect covariate relationships that would predict $Y$ or $\st$.
@@
 This is both because they measure a similar construct to the posttest, and because \cite{impactPaper} found a pretest-treatment interaction in a different treatment contrast.
+
 Following that reasoning, we conducted a third parallel analysis excluding prior achievement covariates in $\bxs$ but including all covariates in $\bxy$.
@@
 As with principal score models, we checked model fit with residual plots and modified model specifications accordingly.
-%\subsubsection{Principal Score Model}\label{sec:psMod}
 
@@
 
-\subsection{Results}
+\subsubsection{Results}
 
@@
 
-% \subsubsection{Outcome Regression and Principal Effects}\label{sec:outMod}
-
-% Figure \ref{fig:effects} plots the estimated principal effects for bottom-outers and non-bottom-outers, using two different models---alongside estimates of the overall average treatment effect, labeled ``ATE.''
-% The two sets of estimated differed in that the first assumed that coefficients on covariates were equal across principal strata, whereas the second set of results allowed the coefficients on pretest and 5th-grade standardized test scores---the most important predictors of outcomes and principal strata membership, respectively---to vary between strata.
-
-% The results suggest little difference, if any, between the effects in the two principal strata.
-% Estimated effects compared to BAU or DragonBox were slightly lower for bottom-outers than for non-bottom-outers, and estimated effects compared to FH2T were higher for non-bottom-outers.
-% However, none of these differences is statistically significant, so opposite patterns, or no difference between principal effects at all, would also be consistent with the data.
-% In fact, 95\% confidence intervals for all principal effects include both positive and negative principal effects.
-% There is no discernible difference between estimated principal effects when coefficients on pretest and grade-5 standardized tests were or were not allowed to vary between strata, though standard errors from the models including an interaction were somewhat higher.
-
 \begin{figure}
   \centering
-  \includegraphics{compareMethods.pdf}
+  \includegraphics{figure/compareMethods.pdf}
   \caption{Principal effect estimates with 95\% confidence or credible intervals using three different principal stratification methods, alongside ATE estimates and 95\% confidence intervals, for contrasts between the ASSISTments condition and each of the three other conditions in the study.
@@
 Figure \ref{fig:compare} shows principal effect estimates from \textsc{geepers} (using the All Covariates principal score model) compared to analogous estimates using Bayesian mixture modeling and \textsc{psw}, alongside estimates of the overall average treatment effects, labeled ``ATE.''
+
 The \textsc{psw} model used the same principal score model as \textsc{geepers}, while the Bayesian model used the same principal score and outcome model specifications.
@@
 %Then let $\blamh_i=\blam_i(Z_i,S_i,\bxsi,\bxy_i,Y_i;\bm{\hat{\alpha}},\bm{\hat{\beta}})$ be the estimating equations evaluated at the estimated parameters, and let
+
 %\begin{equation*}
@@
  which capture the dependence of model \eqref{eq:regression} on the parameters $\bm{\alpha}$ from \eqref{eq:psMod} and the covariance between the estimating equations of the two models.
+
  The sub-matrix $A_{12}=\sum_i \partial \Omega_i/\partial \bm{\beta}=0$, since \eqref{eq:psMod} does not depend on $\bm{\beta}$.
@@
 \begin{center}
-    \includegraphics[width=5.5in]{simFigs/alphaAUC.pdf}
+
+    \includegraphics[width=5.5in]{../simFigs/alphaAUC.pdf}
+
 \end{center}
@@
 The following table gives the empirical coverage of nominal 95\% intervals for \textsc{geepers} and mixture model principal effect estimates under varying data generating models.\\
-\input{writeUps/coverageTabAppendix.tex}
+
+\input{coverageTabAppendix.tex}
 
@@
 The following table gives the root mean squared error (RMSE) for \textsc{geepers}, mixture model, and principal score weighting principal effect estimates under varying data generating models.\\
-\input{writeUps/rmseTabAppendix}
+
+\input{rmseTabAppendix}
 
@@
 \small
-\input{writeUps/tab1num}
+\input{tab1num}
 
@@
 %\input{writeUps/outcomeRegAppendix}
-\input{writeUps/psTab}
-
-\input{writeUps/regTab}
+\input{psTab}
+
+\input{regTab}
 

### 2025-06-06 writeUps/paper.tex
more writing
updates 6/6/25

@@
 
-\RequirePackage{amsthm,amsmath,amsfonts,amssymb}
+\RequirePackage{amsmath,amsfonts,amssymb}
 %\RequirePackage[authoryear]{natbib}
@@
 
+\usepackage[dvipsnames]{xcolor}
+
 %\startlocaldefs
+
+\newcommand{\rd}[1]{\textcolor{BrickRed}{#1}}
 
@@
 (There are certainly other factors at play as well, so this analysis would not be conclusive.)
-
-Estimating an effect for bottom-outers is a tricky proposition.
+Indeed, if the use of bottom-out hints is considered part of the treatment, then those students who rarely used them are partial compliers---could the treatment still benefit them?
+
+Estimating an effect for bottom-outers, or other partial compliers, is a tricky proposition.
 Causal inference in statistics typically depends on comparisons of outcomes between similar groups of subjects.
@@
 \begin{equation*}
-  A^{-1}=\sum_i\frac{\partial}{\partial \bm{theta}'} \bm{\psi}_i'
+  A^{-1}=\sum_i\frac{\partial}{\partial \bm{\theta}'} \bm{\psi}_i'
 \end{equation*}
@@
   \centering
-  \includegraphics[width=5.5in,clip]{../simFigs/boxplots.pdf}
+  \includegraphics[width=5.5in,clip]{../simFigs/boxplotsNew.pdf}
   \caption{Violin and scatter plots of 500 simulation estimation errors for estimators described in Section~\ref{sec:simMods} under varying conditions.
@@
 
-Figure \ref{fig:boxplots} shows violin plots, layered on top of jittered scatter plots, showing estimation error for \textsc{geepers}, the mixture estimator, and the \textsc{psw} estimator of $\eff1$ under various sets of conditions, and Table \ref{tab:coverage} shows empirical coverage of nominal 95\% confidence interval estimates under a somewhat wider set of circumstances when $\muc1=0.3$.
-More complete results, including root-mean-squared-error estimates, are in the online appendix.
-The leftmost panel of Figure \ref{fig:boxplots} and the top line of Table \ref{tab:coverage} show results under conditions favorable to both the mixture model and to \textsc{geepers}---normal residuals and no interactions in the data generating model between covariates and either $S$ or $Z$.
-\textsc{geepers} was roughly unbiased, but less precise than the other two methods---when $\alpha=0.2$ it was much noisier, and when $\alpha=0.5$ it was only slightly less precise.
-The mixture estimator's bias depended on $\muc1$---when $\muc1=0$, the bias was slight, and when $\muc1=0.3$ the bias was more substantial.
-This is somewhat surprising, because when $\muc1=0$, there was no separation between the principal strata in the control group, and previous literature \citep{griffin2008application} has identified this scenario as a particular challenge for mixture modeling.
-The \textsc{psw} estimator was also biased, due, presumably, to the violation of Principal Ignorability, Assumption \ref{ass:PI}, inherent in the data generating model.
-On the other hand, the \textsc{psw} estimator had easily the lowest sampling variance of the three estimators.
-
-The middle panel of Figure \ref{fig:boxplots} and the fifth line of Table \ref{tab:coverage} show results for when the residuals in \eqref{eq:y-sim} are uniformly distributed.
-In this case, \textsc{geepers} and \textsc{psw} estimator, which do not assume normality, behaved in a similar fashion as in the normal case, but the mixture estimator showed a bimodal pattern--sometimes over-estimating $\eff1$ and sometimes underestimating, but rarely estimating $\eff1$ accurately.
-Credible intervals from the mixture model exhibited severe undercoverage, while \textsc{geepers} confidence intervals achieved at least their nominal level.
-
-%\sloppy
-The rightmost panel of Figure \ref{fig:boxplots} returns to the case of normal residuals---and only shows results when $\muc1=0.3$ and $\alpha=0.5$---but introduces interactions in the data generating model between covariates and either $\st$, $Z$, or both, violating Assumption \ref{ass:rci}.
-%Presence of interactions could undermine the \textsc{geepers} by inducing a relationship between principal scores and potential outcomes, even after adjusting the outcomes for covariate effects.
-Fortunately, these interactions made little difference when $\alpha=0.5$.
-However, Table \ref{tab:coverage} shows that when $\alpha=0$ or 0.2, and interactions between $\bx$ and $Z$ were present (i.e.
-$\gamma_3\ne 0$), \textsc{geepers} 95\% confidence intervals tended to under-cover.
-Nevertheless, coverage of \textsc{geepers} intervals was nearly always substantially higher than coverage of corresponding 95\% credible intervals from mixture models.%\footnote{}
+Figure \ref{fig:boxplots} shows violin plots, layered on top of jittered scatter plots, showing estimation error for \textsc{geepers}, the parametric mixture model, and the \textsc{psw} estimator of $\eff1$ for, and Table \ref{tab:coverage} shows empirical coverage of nominal 95\% confidence interval estimates for \textsc{geepers} and a parametric mixture models; when at least one of an estimator's assumptions is violated, the result is colored red.
+
+The first row of Table \ref{tab:coverage} and the first column of Figure \ref{fig:boxplots} show similar results as Figure \ref{fig:alphan}---when all assumptions of both methods are met, \textsc{geepers} is slightly less biased than the parametric mixture model; when $\alpha=0.2$, \textsc{geepers} is much more variable and $\alpha=0.5$ the difference in variance is slight.
+\textsc{psw} has by far the lowest variance and the hightest bias; the latter is due to the violation of principal ignorability.
+
+The remainder of the results reflect violations modeling assumptions for either \geepers or the parametric mixture model or both: interactions in the data generating model between covariates $X$ and treatment assignment $Z$ and/or principal stratum $S$, either of which would violate assumption \ref{ass:rci} and the linear additive model of the \textsc{pmm}, and/or residuals drawn from a uniform distribution, which violates the normality assumption of \textsc{pmm}.
+For \geepers, when $\alpha=0.2$ the presence of interactions in the data generating model increases sampling variance, but only interactions between $X$ and $Z$ reduce confidence interval coverage.
+When $\alpha=0.5$, the presence of interactions has little to no effect on the performance of the \geepers estimator, and all 95\% confidence intervals approximately acheive their nominal levels.
+\geepers makes no assumptions about the shape of the residual distribution and behaves similarly under either normal or uniform residuals.
+
+Like \geepers, \pmm is insensitive to interactions between $X$ and $S$.
+However, interactions between $X$ and $Z$ in the data generating model skew the \pmm sampling distribution and reduce the coverage of its 95\% confidence intervals, even when $\alpha=0.5$.
+When the residual distribution was uniformly distributed, the \pmm sampling distribution appeared bimodal, and its credible intervals exhibited severe undercoverage.
+
+In sum, \psw is the least variable of all three estimator, and does not depend on outcome modeling, but is vulnerable to bias from an unmeasured covariate that predicts both $S$ and $Y$.
+When all assumptions of both methods are met, \geepers is less efficient, but also less biased, than \pmm, but differences in efficiency decrease with $n$ and $\alpha$.
+Neither \psw nor \pmm are vulnerable to interactions beetween $X$ and $S$.
+\pmm is highly vulnerable to violations of normality and to interactions between $X$ and $Z$, while \psw is robust to non-normality and robust to $X$-$Z$ interactions when $\alpha-0.5$.
 
@@
 \centering
-\begin{tabular}{ccccccccc}%*{6}{r}}
+\begin{tabular}[t]{lllllll}
 \toprule
-\multicolumn{3}{c}{ } & \multicolumn{2}{c}{$\alpha=0$} & \multicolumn{2}{c}{$\alpha=0.2$} & \multicolumn{2}{c}{$\alpha=0.5$} \\
-\cmidrule(l{3pt}r{3pt}){4-5} \cmidrule(l{3pt}r{3pt}){6-7} \cmidrule(l{3pt}r{3pt}){8-9}
-\makecell[l]{Residual\\Dist.} & \makecell[c]{X:Z\\Int.?} & \makecell[r]{X:S\\Int.?} & \rotatebox[origin=c]{300}{\textsc{geepers}} & \rotatebox[origin=c]{300}{Mixture}  & \rotatebox[origin=c]{300}{\textsc{geepers}} & \rotatebox[origin=c]{300}{Mixture}& \rotatebox[origin=c]{300}{\textsc{geepers}} & \rotatebox[origin=c]{300}{Mixture}\\
+\multicolumn{3}{c}{ } & \multicolumn{2}{c}{$\alpha=0.2$} & \multicolumn{2}{c}{$\alpha=0.5$} \\
+\cmidrule(l{3pt}r{3pt}){4-5} \cmidrule(l{3pt}r{3pt}){6-7}
+\makecell[l]{Residual\\Dist.} & \makecell[l]{X:Z\\Int.?} & \makecell[l]{X:S\\Int.?} & \geepers & \pmm & \geepers & \pmm\\
 \midrule
- & No & No & 0.99 & 0.99  & 0.97 & 0.97 & 0.96 & 0.96\\
-
- & Yes & No & 0.88 & 0.74  & 0.88 & 0.74 & 0.94 & 0.89\\
-
- & No & Yes & 1.00 & 1.00  & 0.98 & 0.97 & 0.97 & 0.96\\
-\multirow{-4}{*}{\raggedright\arraybackslash Normal} & Yes & Yes & 0.90 & 0.77  & 0.90 & 0.77 & 0.94 & 0.88\\
-\cmidrule{1-9}
- & No & No & 0.99 & 0.29  & 0.97 & 0.40 & 0.95 & 0.61\\
-
- & Yes & No & 0.83 & 0.11  & 0.87 & 0.20 & 0.93 & 0.48\\
-
- & No & Yes & 1.00 & 0.47  & 0.98 & 0.53 & 0.96 & 0.57\\
-
-\multirow{-4}{*}{\raggedright\arraybackslash Uniform} & Yes & Yes & 0.83 & 0.15 & 0.88 & 0.21 & 0.94 & 0.53\\
+ & No & No & 0.97 & 0.97 & 0.96 & 0.96\\
+
+ & Yes & No & \rd{0.88} & \rd{0.74} & \rd{0.94} & \rd{0.89}\\
+
+ & No & Yes & \rd{0.98} & \rd{0.97} & \rd{0.97} & \rd{0.96}\\
+
+\multirow{-4}{*}{\raggedright\arraybackslash Normal} & Yes & Yes & \rd{0.9} & \rd{0.77} & \rd{0.94} & \rd{0.88}\\
+\cmidrule{1-7}
+ & No & No & 0.97 & \rd{0.4} & 0.95 & \rd{0.61}\\
+
+ & Yes & No & \rd{0.87} & \rd{0.2} & \rd{0.93} & \rd{0.48}\\
+
+ & No & Yes & \rd{0.98} & \rd{0.53} & \rd{0.96} & \rd{0.57}\\
+
+\multirow{-4}{*}{\raggedright\arraybackslash Uniform} & Yes & Yes & \rd{0.88} & \rd{0.21} & \rd{0.94} & \rd{0.53}\\
 \bottomrule
-\multicolumn{9}{p{3.5in}}{\footnotesize Based on 500 replications. $n=500$, $\beta_1=0.3$. Simulation standard error $\approx 1$ percentage point.}
+\multicolumn{7}{p{3.5in}}{\footnotesize Based on 500 replications.
+$n=500$, $\beta_1=0.3$.
+Simulation standard error $\approx 1$ percentage point.
+Estimates colored \rd{red} indicate cases where the assumptions of the model are not met.}
 
@@
 For \textsc{geepers} and \textsc{psw}, we used standard maximum likelihood methods and data from the treatment group, achieving an AUC of 0.68; for the mixture model we incorporated the logistic regression as a sub-model within a larger Bayesian model fit in Stan.
+
+\begin{figure}
+  \centering
+  \includegraphics{../figure/opts.pdf}
+  \caption{Estimated principal effects and 95\% confidence or credible intervals for the Obstetrics and Periodontal Therapy Study, using four estimation methods.
+Results for BSIV were taken from \citet{richardson2023estimating}}
+  \label{fig:opt}
+\end{figure}
 
@@
 Our goal is to estimate different treatment effects for groups of students who would, if assigned to ASSISTments, request few or many bottom-out hints.
+For the sake of the illustration, we assert that in order to receive the full dose of the ASSISTments condition, a student would have to request bottom-out hints
 
 The ability to bottom out was only one of several differences between the treatment conditions.
-One way to disentangle the role of bottoming out in any treatment effect would be to estimate average treatment effects of being assigned to ASSISTments, relative to any of the other three, separately for subjects who would (if assigned to ASSISTments) bottom out frequently, and for subjects who would not.
+One way to disentangle the role of bottoming out in any treatment effect would be to estimate average treatment effects of being assigned to ASSISTments, relative to each of the other three, separately for subjects who would (if assigned to ASSISTments) bottom out frequently, and for subjects who would not.
 If the availability of bottom-out hints plays an important positive role in ASSISTments' efficacy, we might expect larger effects of condition on students who would bottom out frequently than for those who would not; if bottom-out hints are harmful, we might expect the opposite.
@@
 (These hypothetical results may be masked by other treatment effect heterogeneity between students who would or wouldn't bottom out, but the results could still be informative.)
+
+\subsubsection{Principal Stratification and Alternatives}
+The goal of principal strafication in this context is to determine if assignment to the ASSISTments condition affected students who tend to request frequent bottom-out hints differently than students who tended to request fewer bottom-out hints.
+Is this the right question to ask?
+
+In a similar context, \citet{sales2021student} suggests two alternative appproaches: the average effect of requesting bottom-out hints on posttest scores for students in the ASSISTments condition, and estimating natural direct and indirect effects of assignment to the ASSISTments condition, treating bottom-out hint requests as a mediator.
+The former approach is an observational study: although students were randomized to the ASSISTments condition, bottom-out hint requests were not randomized.
+Therefore, estimating the effect of bottom-out hints requires careful control of confounding.
+Unfortunately, confounding is a major concern in this case, since, presumably, students only requested bottom-out hints if they were struggling with the material, and students who struggled with practice problems are likely to score lower on the posttest, on average, than those who needed less help during practice.
+The best available measure of students' baseline ability, their score on the 10-item pretest, is likely insufficient.
+That said, a plausible estimate of the average effect of bottom-out hint requests would be a valuable contribution to the literature, and a suitably innovative observational design \citep[e.g.][]{beck} may bear fruit.
+
+Another arguable disadvantage of the observational study approach is that it ignores the larger context of the experiment, asking whether bottom-out hints affect posttest scores, rather than asking what role they play in ASSISTments' overall effectiveness.
+A second approach, mediation analysis, contextualizes the bottom-out effect within the larger experiment.
+It, too, relies on the same non-confounding assumption as an observational study, and also suffers from an additional challenge: mediation analysis relies on a ``common support'' condition that ``the mediator state must not be a deterministic function of the treatment'' \citep[p.
+225]{celli2022causal}.
+However, bottom-out hints are unavailable in the other three arms of the study, violating this condition (\citealt{sales2021student} offers a solution, which we find unconvincing).
+
+A principal stratification approach treats bottom-out hint requests as a type of compliance.
+In other words, if we consider full implementation of the ASSISTments condition to include frequent bottom-out hint requests, students assigned to ASSISTments but requests few bottom-out hints is only partially complying with their treatment assignment.
+Then the question is if students who do not fully comply with their treatment assignment still experiance an effect.
+Conversely, might students who only partially comply experience a larger positive effect?
+That would be the case if either bottom-out hints were harmful, or if the treatment was better targeted toward the type of student who tends to need them less.
+
+Principal score weighting depends on the assumption of principal ignorability, which is equivalent to the no-confounding assumption of the observational study and mediation analysis.
+The mixture model and \textsc{geepers} estimators do not rely on that assumption, instead relying on parametric models and residualized covariate independence, respectively.
+Finally, the instrumental variables estimator of \citet{air} relies on the ``exclusion restriction'' that, for students who request few bottom-out hints, there is no effect of randomization to the ASSISTments condition.
+Since bottom-out hints are only one of several features of ASSISTments, this assumption is untenable.
+
+\subsubsection{Data on Bottom-Out Hints and Covariates}
 
@@
 This is both because they measure a similar construct to the posttest, and because \cite{impactPaper} found a pretest-treatment interaction in a different treatment contrast.
-
-Following that reasoning, we conducted a third parallel analysis excluding prior achievement covariates in $\bxs$ but including all covariates in $\bxy$.
+Following that reasoning, we conducted a third parallel analysis excluding prior achievement covariates.% in $\bxs$ but including all covariates in $\bxy$.
 
@@
 The three covariate selection strategies led to similar point and interval estimates.
-That said, in nearly all cases the AIC-optimal model led to the smallest standard errors.
+That said, in most cases the AIC-optimal model led to the smallest standard errors.
 Confidence intervals for all nine principal effect estimates were about twice as wide as for corresponding ATE estimates.
@@
 
-\section{Additional Results from the Application}
+\section{Additional Results from the Applications}
 \singlespacing
@@
 
+\clearpage
+
+\section{Responses to Reviewers}
+We are grateful for the reviewers' and the editor's careful reading
+and comments. We believe that the the substantial changes we made
+chages to the manuscript in response to those comments greatly
+improved the paper. Most significantly, we clarified our presentation
+and added a second application.
+We provide point-by-point responses below.
+
+\subsection{Reviewer 1}
+\begin{enumerate}
+
+  \item \begin{quote}The rationale for using Principal Stratification to estimate treatment
+effects for bottom-outers is insufficient. Additionally, it is
+advisable to compare the Principal Stratification method to other
+causal inference methods, such as instrumental variable and propensity score analysis, to provide a comprehensive understanding of the chosen
+approach.
+\end{quote}
+\item \begin{quote}
+    While the authors effectively describe the novelty of their study, the literature review lacks the historical
+development of Principal Stratification. Including more in-depth reviews and relevant citations would enhance the
+background information and context of the study.
+\end{quote}
+\item
+\begin{quote}
+  The simulation details about how variable Z is simulated and the number of replications for each condition
+are not clearly explained. Providing more information on these aspects would contribute to the transparency and replicability
+of the study.
+\end{quote}
+\item \begin{quote}
+    The rationale for selecting specific values for manipulated factors, such as choosing a sample size range of
+100 to 1000 and using N(0,1/2) instead of N(0,1), is not adequately justified. Including evidence and citations to support
+these choices would strengthen the methodological foundation of the
+study.
+\end{quote}
+\item \begin{quote}
+    A section detailing the evaluation metrics and criteria used to assess simulation results should precede the
+presentation of results. This would provide readers with a clear understanding of the benchmarks against which the findings
+are measured.
+\end{quote}
+\item
+ \begin{quote} It is highly recommended that the authors employ a series of ANOVAs to analyze the impact of various
+factors on the accuracy of effect estimation to guide results interpretation. Additionally, a concise paragraph summarizing key
+findings before the application example would enhance the clarity and
+interpretation of the results.
+\end{quote}
+\item
+  \begin{quote}
+    APA Citations and References: The paper does not adhere to the APA 7th style. Authors are advised to consult the APA
+    Formatting and Style Guide %(https://owl.purdue.edu/owl/research_and_citation/apa_style/apa_formatting_and_style_guide/index.html)
+    for proper citation and reference formatting throughout the manuscript. Consistency in style is crucial
+for scholarly publications.
+\end{quote}
+
+\end{enumerate}
+
+\subsection{Reviewer 2}
+\begin{quote}
+Great work. Well written. Please find some of my comments below. I
+hope they are useful.
+\end{quote}
+
+\begin{itemize}
+\item
+  \begin{quote} Please provide more background on M-estimators especially to the extent it evolves to become relevant to the
+    GEEPERs.
+    \end{quote}
+\item
+  \begin{quote}
+I am wondering if doubly robust PSW estimator would reduce the remaining bias difference between GEEPERs and
+PSW, and Mixture and PSW.
+    \end{quote}
+
+\item
+  \begin{quote}
+    GEEPERs seems great, but also, surprised by how efficient Mixture can be. Are there any mechanisms that link M-
+    estimators to Mixture modeling? Just wondering.
+        \end{quote}
+
+\item
+  \begin{quote}
+    Would it be possible to offer some evaluation metric for GEEPERs,
+    which would indicate, under some conditions researchers could
+    possibly opt or seek other estimators
+  \end{quote}
+
+\item
+  \begin{quote}
+    Quickly checked Frangakis and Rubin (2002). Do you think the effect of bottom-outers (or similar experimental
+designs) could be partitioned into associative and dissociative
+effects with GEEPERs?
+    \end{quote}
+
+\item
+  \begin{quote}
+    It would be great if you could share the code for estimation and simulation in the next round of review (if there is). Also
+make sure you have comments in the code so someone reading the article
+can match the narrative with the code.
+    \end{quote}
+
+\end{itemize}
+
+\subsection{Reviewer: 3}
+Comments to the Author
+This paper develops a new estimator for principal stratification based on M-estimation and stacked equations. The proposed
+method is designed to be more robust and simpler to use than standard methods. The method is, however, tailored to a very
+specific set up: one-way noncompliance. In general, principal stratification is a much more general framework that can be
+used for a wide variety of settings where there is conditioning on some post-treatment quantity.
+The method is outlined, and then evaluated through a series of simulations, and then an application is provided. My primary
+criticism is that the new method doesn't appear to outperform extant methods to a compelling degree. I take up this point in
+more detail first for the simulations and second for the empirical
+application.
+
+\subsubsection{Simulations}
+\begin{quote}
+  In my view, the simulations should show the following:
+  \begin{itemize}
+    \item When the assumptions of extant methods are met, GEEPERS should do no worse.
+    \item When the assumptions of extend methods are not met, GEEPERS
+      should do better.
+  \end{itemize}
+Given this criteria, it seems to me that the simulation should focus on the last three factors in the design. That is, I would
+hold the sample size and alpha fixed. Currently too many factors are varied at the same time to give a clear picture. I would
+alter Table 1, to indicate which factors yield assumption violations and then violate these assumptions in isolation instead of
+relaxing them simultaneously. As it currently stands, it is quite hard to understand the relative performance of GEEPERs
+compared to extant methods. Careful relaxation of each assumption allows the reader to see how much bias results from a
+single factor.
+\end{quote}
+
+\subsubsection{Application.}
+I thought there are a number of less than desirable parts to the application. First, the focus on bottoming out is less than
+ideal. This strikes me as a fairly rare phenomenon. I didn't find the concept very intuitive and had to read just the descriptive
+section several times to understand why one might be interested in a principal stratification analysis here. This makes the
+methods seem not particularly general. Why not focus on an application with straightforward oneway noncompliance? This
+is an area where principal stratification is often applied, and the concepts are well understood.
+In addition, it isn't ideal that the ATE in each condition is not estimated very precisely, that is the CI appears to cover zero in
+all cases. Next, all the principal stratification estimates are all pretty close to the same value as the ATE. That serves as
+another reason this application isn't compelling. It is hard to care about an application where principal stratification doesn't
+provide much in the way of insights relative to the ATE. Next, the performance of GEEPERs seems to be quite similar to
+extant methods in each scenario. Overall, to my thinking, there is little to recommend the current application. Again, there is
+little here that demonstrates the value of GEEPERs.
+To my mind, these problems could both be fixed. That is, the simulations could be redesigned to provide more compelling
+evidence. Next, the authors may be able to identify are better application. Therefore, the manuscript should be revised and
+resubmitted.
+
 \end{document}

### 2025-06-06 code/OPTS/opts.r

@@ -142,7 +142,7 @@ sdat <- list(
     sdat$ncovY=ncol(Xout)
 
 
-  psStan <- stan(file = 'code/fh2t/psModSimp.stan')#,data=sdat)
+  psStan <- stan(file = 'code/fh2t/psModSimp.stan',data=sdat)
 save(psStan,file='code/OPTS/psStanSimp.RData')
 load('code/OPTS/psStanSimp.RData')
 

### 2025-06-06 code/fh2t/analysisGEE.r

@@ -86,8 +86,8 @@ estimates0all <- lapply(alts,
 
 estimates0rest <- lapply(alts,
                         function(alt)
-                          est(getDat(alt),psMod=psModRest,
-                              covFormY=formula(psModAll)[-2])
+                          est(getDat(alt),psMod=psModRest)#,
+#                              covFormY=formula(psModAll)[-2])
                     )
 
 save(ates,estimates0,estimates3,estimatesInt1,file='results/geeResults.RData')

### 2025-06-06 code/simulation/displaySim.r

@@ -99,8 +99,8 @@ tab1 <- bind_rows(
   coverageN%>%filter(eff==1)%>%select(-eff)%>%
   pivot_wider(names_from=n,names_prefix="n=",values_from=coverage)%>%mutate(measure="95% CI Coverage"))%>%
   select(measure,everything())
-  
-  
+
+
   #%>%
   kbl()%>%
   collapse_rows(columns=1,latex_hline="major",valign="middle")
@@ -221,11 +221,49 @@ pd <- results %>%
     dist=paste(c(mix='Mixture',unif='Unform',norm='Normal')[errDist],'Errors'),
     estimator=c(bayes='Mixture',mest='GEEPERs',psw='PSW')[estimator],
     interactionZ=ifelse(intZ,"Z\ninteraction","No Z\ninteraction"),
-    interactionS=ifelse(intS,"S\ninteraction","No S\ninteraction")
+    interactionS=ifelse(intS,"S\ninteraction","No S\ninteraction"),
+    intAll=ifelse(intZ,ifelse(intS,"X:S & X:Z","X:Z"),ifelse(intS,"X:S","No inter.")),
+    intAll=factor(intAll,levels=c("No inter.","X:Z", "X:S","X:S & X:Z"))
   ) %>%
   filter(estimator=='PSW'|rhat<1.1)
 
 ### figure for paper
+
+norm<-bp(pd,
+   subset=b1 > 0& errDist=='norm'& eff==1&mu01==0.3&n==500&PE=='Stratum 1',
+   title="Normal Residuals",ylim=c(-1.5,1.5),#c(-1.5,1.5),
+   facet=B1~intAll,#interactionZ+interactionS,
+   Labeller=labeller(B1="none",#label_parsed,
+                     intAll=label_value),labSize=2.5
+   )+
+    labs(y=bquote("Estimation Error for "~tau^1),#subtitle="No Interactions",
+         x=NULL)+
+  #theme_bw()+
+    theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1),
+          strip.text.y=element_blank(),strip.background.y=element_blank(),
+                                        #plot.subtitle = element_text(size=10),
+        legend.position="none")+
+  scale_fill_manual(values=c('#1b9e77','#d95f02','#7570b3'))+
[…]

### 2025-06-06 code/simulation/readSimFuncs.r

@@ -168,8 +168,8 @@ bp <- function(pd,subset,facet,title=deparse(substitute(subset)),ylim,Labeller="
   ##                             mapping=aes(estimator,y,label=lab),size=labSize,label.padding=unit(0.1, "lines"))
 
   if(!is.null(ylim))
-    p=p+stat_summary(aes(estimator,errP),geom='label', fun.data=function(xx) data.frame(y=if(sum(xx<ylim[1])>0) ylim[1] else ylim[1]-100,label=paste('+',sum(xx <ylim[1]))),inherit.aes=FALSE)+
-      stat_summary(aes(estimator,errP),geom='label', fun.data=function(xx) data.frame(y=if(sum(xx>ylim[2])>0) ylim[2] else ylim[2]+100,label=paste('+',sum(xx >ylim[2]))),inherit.aes=FALSE)
+    p=p+stat_summary(aes(estimator,errP),geom='label', fun.data=function(xx) data.frame(y=if(sum(xx<ylim[1])>0) ylim[1] else ylim[1]-100,label=paste('+',sum(xx <ylim[1]))),inherit.aes=FALSE,size=labSize)+
+      stat_summary(aes(estimator,errP),geom='label', fun.data=function(xx) data.frame(y=if(sum(xx>ylim[2])>0) ylim[2] else ylim[2]+100,label=paste('+',sum(xx >ylim[2]))),inherit.aes=FALSE,size=labSize)
   p
 }
 

### 2025-06-12 writeUps/paper.tex
overleaf update

--- 2025-01-09 writeUps/paper.tex
@@
 \usepackage{longtable}
-\usepackage[notablist,tablesfirst]{endfloat}% To `separate' figures and tables from text if required
+%\usepackage[notablist,tablesfirst]{endfloat}% To `separate' figures and tables from text if required
 
@@
 %\begin{aug}
-% \author{
-% \name{Adam C. Sales,\thanks{CONTACT Adam C. Sales. Email: asales@wpi.edu} Kirk P. Vanacore, and Erin R. Ottmar}
-% \affil{Worcester Polytechnic Institute, Worcester, MA, USA}
-% }
+ \author{
+ Adam C. Sales\thanks{CONTACT Adam C. Sales. Email: asales@wpi.edu}, Kirk P. Vanacore, and Erin R. Ottmar%\\
+ %\affil{Worcester Polytechnic Institute, Worcester, MA, USA}
+}
 
@@
 Estimates may be computed using conventional regressions (though the standard errors require a specialized sandwich formula) and do not rely on distributional assumptions.
-We present a simulation study that shows that the novel method is more robust than popular alternatives and illustrate the method in an analysis of data on bottom-out hint requests.
+We present a simulation study that demonstrates the novel method's greater robustness compared to popular alternatives and illustrate the method through two real-data analyses.
 \end{abstract}
@@
 The availability of problem-specific support for struggling students is one potential advantage of educational technology over traditional pencil-and-paper practice.
-On the other hand, the ready availability of hints may discourage students from working through difficult problems on their own, and so may inhibit learning.
-
-The literature on hints in online learning is mixed \citep[see, e.g.,][]{aleven2016help,goldin2012learner,sales2021student} in particular when it comes to requesting a ``bottom-out'' hint.
+On the other hand, the ready availability of hints may discourage students from working through difficult problems on their own, thereby inhibiting their learning.
+
+The literature on hints in online learning is mixed \citep[see, e.g.,][]{aleven2016help,goldin2012learner,sales2021student}, particularly when it comes to requesting a ``bottom-out'' hint.
 In some online mathematics programs, sequences of hints, which a student can request one at a time, end with a bottom-out hint providing the answer to the problem.
 On the one hand, hints that include the correct answer are essentially worked examples, which can be beneficial to learning \citep[e.g.]{sweller1985use}.
-On the other hand, they allow students to ``game the system,'' by simply requesting enough hints to see the right answer, and use the software without ever working to solve any problems \citep[e.g.][]{guo2008trying}.
+On the other hand, they allow students to ``game the system,'' by simply requesting enough hints to see the right answer and use the software without ever working to solve any problems \citep[e.g.][]{guo2008trying}.
 
@@
 If that condition benefited students who frequently requested bottom-out hints (``bottom-outers'') more than those who didn't, that may suggest that the availability of bottom-out hints played a role in the condition's effectiveness.
-If students who avoided bottom-out hints benefited more, it may be because the availability of bottom-out impeded the conditions effectiveness.
+Alternatively, if students who avoided bottom-out hints benefited more, it may be because the availability of bottom-out hints impeded the condition's effectiveness.
 (There are certainly other factors at play as well, so this analysis would not be conclusive.)
@@
 However, there is no readily identifiable comparison group for bottom-outers, who are only represented in one of the four treatment conditions, and who presumably struggled more frequently, on average, than their non-bottom-outer peers.
-We will formulate a solution using Principal stratification \citep{frangakis}, a  framework for defining subgroup effects for subgroups that may themselves be affected by treatment assignment.
-It relies on the notion of \emph{potential} subgroups called principal strata, grouping subjects based on the values a post-treatment variable would potentially take if the subject were assigned to one treatment condition or another.
+We will formulate a solution using Principal stratification \citep{frangakis}, a framework for defining and estimating effects for post-treatment subgroups, that is, subgroups that may themselves be a function of treatment assignment.
+It relies on the notion of \emph{potential} subgroups called principal strata: which subgroup would a subject be in if they were (perhaps counterfactually) assigned to one treatment condition or another.
 For example, the appropriate comparison group for bottom-outers is the set of students, assigned to the other treatment arms, who \emph{would have} requested bottom-out hints if only given that option.
 
-Since principal strata membership is based on subjects' responses to both factual and counter-factual treatment assignments, it is unobserved---we don't observe which students in the conditions without bottom-out hints would have bottomed out if given the opportunity.
-Nonetheless, researchers have devised estimators of principal effects, as average effects within principal strata are called, to address a broad range of applied statistical problems.
+Since principal strata membership is based on subjects' responses to both factual and counterfactual treatment assignments, it is unobserved---we don't observe which students in the conditions without bottom-out hints would have bottomed out if given the opportunity.
+Nonetheless, researchers have devised estimators of principal effects---average effects within principal strata---to address a broad range of applied statistical problems.
 
 The conceptual basis for principal stratification---though not the terminology---emerged from the study of randomized experiments with noncompliance.
-\citet{air}, which gave a novel interpretation of the econometric instrumental variables estimator as the average effect of the treatment on the principal stratum of ``compliers''---those subjects who would receive the treatment if and only if randomized to the treatment condition.
-They showed that the average effect for compliers is nonparametrically identified if a random treatment assignment only has an effect for compliers---``exclusion restriction''---and if treatment assignment never prevents any subjects from receiving treatment (``monotonicity'').
+Specifically, \citet{air} showed that the econometric instrumental variables estimand is the average effect of the treatment on the principal stratum of ``compliers''---those subjects who would receive the treatment if and only if randomized to the treatment condition.
+Moreover, the average effect for compliers is nonparametrically identified if a random treatment assignment only has an effect for compliers---``exclusion restriction''---and if treatment assignment never \emph{prevents} any subjects from receiving treatment (``monotonicity'').
 In a follow-up analysis, \citet{imbens1997bayesian} relaxed the exclusion restriction by way of parametric Bayesian finite mixture modeling, which remains a popular approach.
@@
 Of particular importance is the ``principal score,'' \citep{jo,dingLu,feller2017principal}, or the probability that a subject belongs to a particular principal stratum, as a function of pre-treatment covariates.
-Under a strong, untestable assumption called ``principal ignorability,''  principal scores can be transformed into weights to compute form an unbiased estimate of principal effects.
+Under a strong, untestable assumption called ``principal ignorability,''  principal scores can be transformed into weights to compute an unbiased estimate of principal effects.
 Principal ignorability is a direct analogue of the strong ignorability assumption in causal observational studies \citep[e.g.][]{rosenbaum1983central}---that all covariates that predict both principal stratum membership and outcomes have been accounted for---in other words, the precise type of assumption that randomized trials are designed to avoid.
@@
 %, just as in the pre-treatment subgroup case.
-We show that this estimator is the solution to a set of estimating equations based on the combined moments of principal scores, indicators for treatment assignment and observed principal stratum membership in the treatment group, outcomes and (optionally) covariates, and give a set of assumptions under which it is consistent for the true principal effects.
+We show that this estimator is the solution to a set of estimating equations based on the combined moments of principal scores, indicators for treatment assignment and observed principal stratum membership in the treatment group, outcomes, and (optionally) covariates, and give a set of assumptions under which it is consistent for the true principal effects.
 We will refer to the estimators presented here as ``\textsc{geepers}'': general estimating equations for principal effects using regressions.
@@
 Our development of \textsc{geepers} builds on \citet{richardson2023estimating} by emphasizing connections to both parametric mixture modeling and principal scores analysis, and comparing \textsc{geepers} to those methods in a simulation study.
-Moreover, \textsc{geepers}, which, in practice, is a combination of logigistic regression and OLS, is more accessible to applied researchers while yielding very similar results.
+Moreover, \textsc{geepers}, which, in practice, is a combination of logistic regression and OLS, is more accessible to applied researchers while yielding very similar results.
 
@@
 \end{equation*}
-estimating principal effects requires estimating the four conditional means.
+estimating principal effects requires estimating $\muc0$, $\mut0$, $\muc1$, and $\mut1$.
 
@@
 
-\subsubsection{Normal Mixture Modeling}
-The normal mixture modeling approach to estimating $\muc0$ and $\muc1$ \citep{imbens1997bayesian} assumes that, conditional on $\st$, $Y_C$ is normally distributed.
+\subsubsection{Paremetric Mixture Modeling}
+%The normal mixture modeling approach to estimating $\muc0$ and $\muc1$ \citep{imbens1997bayesian} assumes that, conditional on $\st$, $Y_C$ is normally distributed.
+Say an analyst assumes that, within principal strata, $Y_C$ is drawn from a parametric model with parameter vector $\bm{\delta}^s$, i.e., $f_{Y_C}(y|\st=s)=g^s(y)$.
+For instance, \citep{imbens1997bayesian} assumes $Y_C$ is normally distributed within strata, so $g^s(y)=\phi\left(\frac{y-\mucs}{\sigma_C^s}\right)$, where $\phi(\cdot)$ is the standard normal density function, and $\sigma_C^0>0$ and $\sigma_C^1>0$ are standard deviations.
 Then the probability density of $Y_C$ may be written as
 \begin{equation}\label{eq:mixtureModel}
-  f_{Y_C}(y)=Pr(\st=0)\phi\left(\frac{y-\muc0}{\sigma_C^0}\right)+Pr(\st=1)\phi\left(\frac{y-\muc1}{\sigma_C^1}\right)
-\end{equation}
-where $\phi(\cdot)$ is the standard normal density function, and $\sigma_C^0>0$ and $\sigma_C^1>0$ are standard deviations, which are sometimes assumed equal.
+ % f_{Y_C}(y)=Pr(\st=0)\phi\left(\frac{y-\muc0}{\sigma_C^0}\right)+Pr(\st=1)\phi\left(\frac{y-\muc1}{\sigma_C^1}\right)
+ f_{Y_C}(y)=Pr(\st=0)g^0(y)+Pr(\st=1)g^1(y).
+\end{equation}
 The probabilities $Pr(\st=0)$ and $Pr(\st=1)$ may be estimated first using data from the treatment group, since Assumption \ref{ass:rand} implies that $Pr(\st=1|Z=1)=Pr(\st=1)$.
@@
 This approach has several problems.
-First, the conditional normality assumption is untestable and limiting.
+First, the assumption of a conditional parametric model %normality assumption 
+is untestable and limiting.
 Unfortunately, even when it holds, mixture model estimates can be unreliable and, in some cases, severely biased \citep{griffin2008application,feller2017principal}.
@@
 \begin{equation}\label{eq:pscore}
-  \ppi=Pr(\sti=1|\bxsi)
-\end{equation}
-where $\bxsi\subseteq\bm{x}_i$, the set of covariates used to model $S$.
+  \ppi=Pr(\sti=1|\bxsi),
+\end{equation}
+%where $\bxsi\subseteq\bm{x}_i$, the set of covariates used to model $S$.
+the probability of a subject with covariates $\bxsi$ being in the $\sti=1$ principal stratum.
 Note that under randomization, $Pr(\sti=1|\bxsi,Z_i)=Pr(\sti=1|\bxsi)$, so a model for principal scores can be estimated using data from the treatment group and extrapolated to the control group.
@@
 
-For principal scores to be useful, they need to vary with $\bm{x}$. Otherwise, $\bm{x}$ is uninformative about $\st$. In particular, we will assume the following:
+If $\bm{x}$ is informative about $\st$, principal scores will vary between subjects. We will assume this is the case: %the following:
 \begin{ass}[Variable Principal Scores]\label{ass:vps}
@@
 \begin{equation}\label{eq:mixtureModelPS}
-  f_{Y_C}(y)=(1-\ppip{\bm{\alpha}})\phi\left(\frac{y-\muc0}{\sigma_C^0}\right)+\ppip{\bm{\alpha}}\phi\left(\frac{y-\muc1}{\sigma_C^1}\right)
-\end{equation}
+  f_{Y_C}(y|\bm{x})=(1-\ppip{\bm{\alpha}})%\phi\left(\frac{y-\muc0}{\sigma_C^0}\right)
+g^0(y|\bm{x})+\ppip{\bm{\alpha}}%\phi\left(\frac{y-\muc1}{\sigma_C^1}\right)
+g^1(y|\bm{x})
+\end{equation}
+where $f_{Y_C}(y|\bm{x})$ and $g^s(y|\bm{x})$, $s=0,1$, are conditional densities which must be specified.
+Most commonly, researchers assume a linear normal model, i.e., $g^s(y|\bm{x})=\phi\left(\frac{y-\mucs-\bm{\gamma}'\bm{x}}{\sigma_C^s}\right)$, where the columns of $\bm{x}$ are centered so that $\mathbb{E}[\bm{x}]=\bm{0}$.
 %Ideally, a statistician will estimate parameters $\muc0$, $\muc1$, $\mut0$, $\mut1$, $\sigma^0_C$, $\sigma^1_C$, and $\bm{\alpha}$ simultaneously, using maximum likelihood or Bayesian techniques.
@@
 M-estimation (also generalized method of moments or generalized estimating equations) is a general approach to deriving consistent statistical estimators and their asymptotic distributions.
-In the following section, we will describe \textsc{geepers}, a novel pricipal effect estimator, and derive its properties using the principals of M-estimation.
-Here, we will give a very brief overview of M-estimation, drawn mostly from the excellent treatment in \citet{stefanskiBoos}.
-
-Let $\bm{\mathcal{O}}_i$ be a vector of observed data for subject $i=1,\dots,n$, with $\bm{\mathcal{O}}_i\independent \bm{\mathcal{O}}_{i'}$ if $i\ne i'$, and let $\bm{\theta}$ be a vector of parameters.
-Then, an M-estimator $\bm{\hat{\theta}}$ satisfies ``estimating equations'': for $j=1,\dots,J$, $\sum_i \psi_j\left(\bm{\mathcal{O}}_i,\bm{\hat{\theta}}\right)=0$. 
-When data $\bm{\mathcal{O}}_i$, $i=1,\dots, n$ are identically distributed, $\bm{\hat{\theta}}$ is consistent for $\bm{\theta}$ so long as $\EE\left[\psi_j(\bm{\mathcal{O}}_1,\bm{\theta})\right]=0$ for $j=1,\dots,J$.
-Some common examples are maximum likelihood estimatiors, where $\bm{\psi}(\bm{\mathcal{O}},\bm{\theta})$ is the score function, with $\psi_j(\bm{\mathcal{O}},\bm{\theta})=\partial log f(\bm{\mathcal{O}},\bm{\theta})/\partial \theta_j$ and $f(\cdot,\bm{theta})$ is the density of $\bm{\mathcal{O}}$, or method of moments estimators, in which, if data $\mathcal{O}$ are unidimensional, $\psi_j(\mathcal{O},\bm{\theta})=\mathcal{O}^j-\EE\left[\mathcal{O}^j\right]$.
-
-Under suitable regularity conditions, an M-estimator $\bm{\hat{\theta}}$ is asymptotically normal and its sampling variance-covariance matrix may be estimated using the formula \citep[][ch.
+In the following section, we will describe \textsc{geepers}, a novel principal effect estimator, and derive its properties using the principles of M-estimation.
+Here, we provide a brief overview of M-estimation, primarily drawn from the excellent treatment in \citet{stefanskiBoos}.
+
+Let $\bm{\mathcal{O}}_i$ be a vector of observed data for subject $i=1,\dots,n$, with $\bm{\mathcal{O}_1},\dots,\bm{\mathcal{O}}_n\overset{\mathrm{iid}}{\sim} P$ for some distribution $P$, %\bm{\mathcal{O}}_i\independent \bm{\mathcal{O}}_{i'}$ if $i\ne i'$,
+let $\bm{\theta}$ be a vector of parameters, 
+and let $\bm{\psi}(\bm{\mathcal{O}},\bm{\theta})$ be a vector field such that $\mathbb{E}_P[\bm{\psi}(\bm{\mathcal{O}},\bm{\theta}_0)]=\bm{0}$ for the true parameter value $\bm{\theta}_0$. 
+Then, an M-estimator $\bm{\hat{\theta}}$ satisfies ``estimating equations'' $\sum_i \psi_j\left(\bm{\mathcal{O}}_i,\bm{\hat{\theta}}\right)=0$. 
+Some common examples are maximum likelihood estimators, where $\bm{\psi}(\bm{\mathcal{O}},\bm{\theta})$ is the score function %, with $\psi_j(\bm{\mathcal{O}},\bm{\theta})=\partial log f(\bm{\mathcal{O}},\bm{\theta})/\partial \theta_j$ and $f(\cdot,\bm{theta})$ is the density of $\bm{\mathcal{O}}$,
+or method of moments estimators, in which, if data $\mathcal{O}$ are unidimensional, $\psi(\mathcal{O},\bm{\theta})=\mathcal{O}-\EE_\theta\left[\mathcal{O}\right]$.
+
+Under suitable regularity conditions, an M-estimator $\bm{\hat{\theta}}$ is asymptotically normal, and its sampling variance-covariance matrix may be estimated using the formula \citep[][ch.
 7]{boosStefanskiBook}:
 \begin{equation}\label{eq:sandwich}
-  \widehat{var}(\bm{\hat{\theta}})=ABA'.
-\end{equation}
+  \widehat{var}(\bm{\hat{\theta}})=ABA'
+\end{equation}
+%\citep[][ch. 7]{boosStefanskiBook}.
+with matrices
+\begin{equation*}
+    A=\left\{\frac{1}{n}\displaystyle\sum_i\left.
+\frac{\partial \bm{\psi}(\bm{\mathcal{O}_i},\bm{\theta})'}{\partial \bm{\theta}'}\right|_{\bm{\theta}=\bm{\hat{\theta}}}\right\}^{-1}\mbox{ and }B=\frac{1}{n}\sum_i \bm{\psi}(\bm{\mathcal{O}_i},\bm{\hat\theta})\bm{\psi}(\bm{\mathcal{O}_i},\bm{\hat\theta})'.
+\end{equation*}
 Because of its form, \eqref{eq:sandwich} is sometimes called a ``sandwich'' formula, where $A$ is the ``bread'' and $B$ is the ``meat'' \citep[e.g.][]{sandwich}. 
-To define the matrices $A$ and $B$, let the vector $\bm{\psi}_i=\left\{\psi_j(\bm{\mathcal{O}}_i,\bm{\hat{\theta}})\right\}_{j=1}^J$. Then,
-\begin{equation*}
-  A^{-1}=\sum_i\frac{\partial}{\partial \bm{theta}'} \bm{\psi}_i'
-\end{equation*}
-and
-\begin{equation*}%\label{eq:Bmat}
-  B=\sum_i \bm{\psi}_i\bm{\psi}_i'
-\end{equation*}
+%To define the matrices $A$ and $B$, let the vector $\bm{\psi}_i=\left\{\psi_j(\bm{\mathcal{O}}_i,\bm{\hat{\theta}})\right\}_{j=1}^J$. Then,
+% \begin{equation*}
+%   A^{-1}=\frac{1}{n}\sum_i\left. \frac{\partial \bm{\psi}_i'}{\partial \bm{\theta}'}\right|_{\bm{\theta}=\bm{\hat{\theta}}}
+% \end{equation*}
+% and
+% \begin{equation*}%\label{eq:Bmat}
+%   B=\frac{1}{n}\sum_i \bm{\psi}_i\bm{\psi}_i'
+% \end{equation*}
 In general, the conditions for asymptotic normality may be difficult to prove; fortunately, they are met for OLS and common generalized linear models such as logistic regression, so long as the dimension of $\bm{\theta}$ does not increase with $n$.
@@
 \begin{equation}\label{eq:stacked}
-  \psi^s(\bm{\mathcal{O}},[\bm{\hat\theta^1},\bm{\hat{\theta}^2}])=[\psi^1(\bm{\mathcal{O}},\bm{\hat\theta^1}),\psi^2(\bm{\mathcal{O}},\bm{\hat\theta^2})]=\bm{0},
-\end{equation}
-where the superscript $s$ stands for ``stacked.''
+  \psi(\bm{\mathcal{O}},[\bm{\hat\theta^1},\bm{\hat{\theta}^2}])=[\psi^1(\bm{\mathcal{O}},\bm{\hat\theta^1}),\psi^2(\bm{\mathcal{O}},\bm{\hat\theta^2})]=\bm{0},
+\end{equation}
+%where the superscript $s$ stands for ``stacked.''
 A variance-covariance matrix estimated using the sandwich formula based on $\psi^s$ accounts for uncertainty from both steps of the estimation procedure.
@@
 \end{ass}
-i.e. $Y_C$ is mean-independent of $\bxs$ conditional on $\st$.
+i.e., $Y_C$ is mean-independent of $\bxs$ conditional on $\st$.
 Under CI, covariates are not informative of the mean of $Y_C$ within principal strata.
@@
 If the principal scores are modeled as \eqref{eq:prinScoreMod}, and parameters $\bm{\alpha}$ are estimated by M-estimation, with estimating equations $\sum_i Z_i\Omega_i=\bm{0}$, then the estimating equations for the principal score model and principal effect estimation can be stacked as in \eqref{eq:stacked}.
-Sandwich standard error estimates of the form \eqref{sandwich} based on $\left[Z_i\omega_i,\psi_i\right]$ will incorporate uncertainty from both principal score modeling and principal effect estimation conditional on principal scores.
+Sandwich standard error estimates of the form \eqref{eq:sandwich} based on $\left[Z_i\omega_i,\psi_i\right]$ will incorporate uncertainty from both principal score modeling and principal effect estimation conditional on principal scores.
 
@@
 \end{ass}
-That is, CI may not hold for potential outcomes themselves, but an analogous assumption holds for \emph{residualized} potential outcomes---i.e.
-$\yc$ and $\yt$ after subtracting out a linear function of (possibly transformed) covariates.
+That is, CI may not hold for potential outcomes themselves, but an analogous assumption holds for \emph{residualized} potential outcomes---i.e., $\yc$ and $\yt$ after subtracting out a linear function of (possibly transformed) covariates.
 By forcing $\bm{\gamma}$ to be the same for both treatment groups, and (implicitly) for both principal strata, Assumption \ref{ass:rci} assumes that there are no interactions between columns of $\bxy$ and either condition or principal stratum.
@@
 Proposition \ref{prop:interactions} in the appendix relaxes these requirements, but some preliminary simulations suggest that standard errors from estimates allowing for these interactions will be prohibitively large.
-Simulation results in the following section address some cases in which interactions are present in the data-generating model but not in the data analysis.
+The simulation results in the following section address cases in which interactions are present in the data-generating model but not in the data analysis.
 
@@
 Under Assumptions \ref{ass:sm}, \ref{ass:rand}, \ref{ass:vps}, and \ref{ass:rci}, let principal scores be estimated as in \eqref{eq:prinScoreMod} using data from the treatment group and assume they are linearly independent of $\bxy$.
-Then say the researcher fits the following model with OLS:
+Then, say the researcher fits the following model with OLS:
 \begin{equation}\label{eq:reg2}
@@
 
-We conducted a simulation study to compare the performance of \textsc{geepers} compared to a Bayesian mixture model and to a principal score weighting estimator, in both favorable and unfavorable circumstances.
+We conducted a simulation study to compare the performance of \textsc{geepers} with that of a Bayesian mixture model and a principal score weighting estimator, in both favorable and unfavorable circumstances.
 The study was designed to answer three sets of overarching questions.
-First, how does the performance of the \textsc{geepers} estimator vary under different conditions, including sample size, the extent to which $\bx$ predicts $\st$, whether and how the principal effects themselves vary, and when important interactions are omitted from the outcome model.
+First, how does the performance of the \textsc{geepers} estimator vary under different conditions, including sample size, the extent to which $\bx$ predicts $\st$, whether and how the principal effects themselves vary, and when important interactions are omitted from the outcome model?
 Second, how does \textsc{geepers} compare with mixture modeling?
@@
 In each run of the simulation, we simulated three independent standard normal covariates, $x_k$ $k=1,\dots,3$.
-However, only the first two covariates were ``observed,'' i.e.
-were included in the analysis model.
-Given the covariates, the (true) principal scores were defined as
+However, only the first two covariates were ``observed,'' i.e., were included in the analysis model.
+Given all three covariates, the principal scores were set as
 \begin{equation*}
@@
 where $\alpha$ is a manipulated factor.
-When $\alpha$ was higher, $\st$ was more easily predicted by covariates; when $\alpha=0$, Assumption \ref{ass:vps} was violated, i.e. $Var(\pp)=0$.
+When $\alpha$ was higher, $\st$ was more easily predicted by covariates; when $\alpha=0$, Assumption \ref{ass:vps} was violated, i.e., $Var(\pp)=0$.
 Principal stratum $\sti\in \{0,1\}$ was simulated as $S_i\sim Bern(\ppi)$.
@@
 
-Treatment assignment $Z$ was simulated as independent of all the other variables with half of subjects assigned to treatment $Z=1$ and the rest to control $Z=0$.
+Treatment assignment $Z$ was simulated as independent of all the other variables, with half of the subjects assigned to treatment $Z=1$ and the rest to control $Z=0$.
 
@@
 
-$\bm{\alpha}$\textbf{:} As mentioned above, $\alpha$ controls the variance of the principal scores, i.e.
-the extent to which $\st$ is predictable as a function of covariates.
+$\bm{\alpha}$\textbf{:} As mentioned above, $\alpha$ controls the variance of the principal scores, i.e., the extent to which $\st$ is predictable as a function of covariates.
 When $\alpha=0$, the covariates are unrelated to $\st$ and principal scores do not vary between participants, violating \eqref{ass:vps}.
@@
 
-$\bm{\beta_1}$\textbf{:} When $\beta_1=0$ there is only one mixture component in the control
+$\bm{\beta_1}$\textbf{:} When $\beta_1=0$, there is only one mixture component in the control
 group, which can be problematic for mixture modeling \citep{griffin2008application,feller2016principal}.
-Conversely, when $\beta_1=0.3$, unless there is an interaction between $Z$ and covariates (i.e.
-$\gamma_3\ne 0$), $\eff0=\eff1=0$, so there is no average treatment effect for either principal stratum.
+Conversely, when $\beta_1=0.3$, unless there is an interaction between $Z$ and covariates (i.e., $\gamma_3\ne 0$), $\eff0=\eff1=0$, so there is no average treatment effect for either principal stratum.
 
@@
 
-All three methods used the same principal score model, a logistic regression of $S$ on $\bx$, and an intercept, fit using data from the treatment group in which $S=\st$ was observed, i.e.
-$\ppf{\bm{\alpha}}{\bx}=logit^{-1}(\alpha_0+\bm{\alpha}'\bx)$.
+All three methods used the same principal score model, a logistic regression of $S$ on $\bx$, and an intercept, fit using data from the treatment group in which $S=\st$ was observed, i.e., $\ppf{\bm{\alpha}}{\bx}=logit^{-1}(\alpha_0+\bm{\alpha}'\bx)$.
 
@@
 
-For \textsc{geepers} and principal score weighting the principal score model was fit by itself, to data from the treatment group, using standard maximum likelihood methods.
+For \textsc{geepers} and principal score weighting, the principal score model was fit by itself, to data from the treatment group, using standard maximum likelihood methods.
 For \textsc{geepers}, we used the estimated principal scores in the control group, along with the observed $S$ in the treatment group, to construct $R$, and then fit the model \eqref{eq:reg2} with OLS.
@@
 \subsection{Results}
-In this section we present the subset of simulation results that we think is the most telling and interesting; a fuller set of results is available in the online appendix.
+In this section, we present the subset of simulation results that we think is the most telling and interesting; a fuller set of results is available in the online appendix.
 
@@
 \textsc{geepers} was roughly unbiased, but less precise than the other two methods---when $\alpha=0.2$ it was much noisier, and when $\alpha=0.5$ it was only slightly less precise.
-The mixture estimator's bias depended on $\muc1$---when $\muc1=0$, the bias was slight, and when $\muc1=0.3$ the bias was more substantial.
-This is somewhat surprising, because when $\muc1=0$, there was no separation between the principal strata in the control group, and previous literature \citep{griffin2008application} has identified this scenario as a particular challenge for mixture modeling.
+The mixture estimator's bias depended on $\muc1$---when $\muc1=0$, the bias was slight, and when $\muc1=0.3$, the bias was more substantial.
+This result is somewhat surprising, because when $\muc1=0$, there was no separation between the principal strata in the control group, and previous literature \citep{griffin2008application} has identified this scenario as a particular challenge for mixture modeling.
 The \textsc{psw} estimator was also biased, due, presumably, to the violation of Principal Ignorability, Assumption \ref{ass:PI}, inherent in the data generating model.
@@
 The middle panel of Figure \ref{fig:boxplots} and the fifth line of Table \ref{tab:coverage} show results for when the residuals in \eqref{eq:y-sim} are uniformly distributed.
-In this case, \textsc{geepers} and \textsc{psw} estimator, which do not assume normality, behaved in a similar fashion as in the normal case, but the mixture estimator showed a bimodal pattern--sometimes over-estimating $\eff1$ and sometimes underestimating, but rarely estimating $\eff1$ accurately.
+In this case, \textsc{geepers} and \textsc{psw} estimators, which do not assume normality, behaved in a similar fashion as in the normal case, but the mixture estimator showed a bimodal pattern--sometimes over-estimating $\eff1$ and sometimes underestimating, but rarely estimating $\eff1$ accurately.
 Credible intervals from the mixture model exhibited severe undercoverage, while \textsc{geepers} confidence intervals achieved at least their nominal level.
@@
 Fortunately, these interactions made little difference when $\alpha=0.5$.
-However, Table \ref{tab:coverage} shows that when $\alpha=0$ or 0.2, and interactions between $\bx$ and $Z$ were present (i.e.
-$\gamma_3\ne 0$), \textsc{geepers} 95\% confidence intervals tended to under-cover.
+However, Table \ref{tab:coverage} shows that when $\alpha=0$ or 0.2, and interactions between $\bx$ and $Z$ were present (i.e., $\gamma_3\ne 0$), \textsc{geepers} 95\% confidence intervals tended to under-cover.
 Nevertheless, coverage of \textsc{geepers} intervals was nearly always substantially higher than coverage of corresponding 95\% credible intervals from mixture models.%\footnote{}
@@
 First, we briefly replicate an analysis from \citet{richardson2023estimating} using the \textsc{geepers} and the other methods discussed above.
-This is a relatively straightforward example of an experiment with imperfect compliance in which \citet{richardson2023estimating} estimated stastistically significant effects overall in both principal strata.
+This is a relatively straightforward example of an experiment with imperfect compliance in which \citet{richardson2023estimating} estimated statistically significant effects overall in both principal strata.
 Next, we present a novel analysis of data from the RCT of computer-based math applications reported in \citet{impactPaper}, and look into the role of bottom-out hints, as described in the introduction.
-This analysis is intended to shed light on an important question in education research, and provides an opportunity to raise questions of model and covariate selection and the plausability of identification assumptions.
+This analysis aims to shed light on an important question in education research, illustrates model and covariate selection, and includes a discussion on the plausibility of identification assumptions.
 
 \subsection{Accounting for Imperfect Compliance: Re-Analysis of the Obstetrics and Periodontal Therapy Study}
-Peridontal disease, a weakening of bone and connective tissue supporting teeth, is highly prevelant among American adults, and is associated with adverse pregnancy outcomes, including preterm birth \citep{periodontalEpi}.
+Periodontal disease, a weakening of bone and connective tissue supporting teeth, is highly prevalent among American adults and is associated with adverse pregnancy outcomes, including preterm birth \citep{periodontalEpi}.
 The Obstetrics and Periodontal Therapy Study \citep[(OPTS)][]{michalowicz2006treatment} investigated the efficacy of nonsurgical periodontal treatment to reduce the preterm birth rate; replication data are available at \url{https://causeweb.org/tshs/obstetrics-and-periodontal-therapy/}.
-Pregnant patients at four different hospitals who were at least 16 years old, at less than 16 weeks and 6 days of gestation, and diagnosed with peridontal disease were individually randomized to either receive peridontal treatment during the first 21 weeks of pregnancy ($Z=1$, $n_T=413$) or after delivery ($Z=0$, $n_C=410$).
-The primary prespecified outcome of the study was babies' gestational age at birth, but the study team also collected data on several secondary birth outcomes, as well as measures of patients' peridontal disease at follow-up.
-
-The main component of the intervention took place over the course of up to four visits, followed by monthly tooth polishing and other dental care as needed, and not all participants randomized to the treatment condition completed the entire treatment plan.
+Pregnant patients at four different hospitals who were at least 16 years old, at less than 16 weeks and 6 days of gestation, and diagnosed with periodontal disease were individually randomized to either receive periodontal treatment during the first 21 weeks of pregnancy ($Z=1$, $n_T=413$) or after delivery ($Z=0$, $n_C=410$).
+The primary prespecified outcome of the study was the babies' gestational age at birth; however, the study team also collected data on several secondary birth outcomes, as well as measures of patients' periodontal disease at follow-up.
+
+The main component of the intervention consisted of up to four visits, followed by monthly tooth polishing and other dental care as needed.
+However, not all participants randomized to the treatment condition completed the entire treatment plan.
 However, 96\% of patients in the treatment group received at least some periodontal care during pregnancy---therefore, there is reason to suspect nonzero treatment effects even for those patients who did not (or would not) fully comply.
@@
 Here, we will use \textsc{geepers}, a Bayesian mixture model, and principal score weighting to estimate the same quantities, following the same analysis decisions as \citet{richardson2023estimating}.
-Replication code for our analysis can be found at \url{}.
+Replication code for our analysis can be found at \url{ https://github.com/adamSales/psGee}.
 
@@
 The compliance variable in the dataset had three categories: ``Yes,'' for subjects in the treatment group who completed the treatment plan (157 subjects), ``No'' for those who did not (4 subjects), and ``Und'' for those who received some therapy, but it is unknown whether they completed the treatment (157 subjects).
-We let $S=1$ when compliance was ``yes''---i.e. when subjects in the treatment group completed the treatment---and $S=0$ otherwise.
+We let $S=1$ when compliance was ``yes''---i.e., when subjects in the treatment group completed the treatment---and $S=0$ otherwise.
 
 For all three estimation methods, we estimated principal scores with logistic regression of $\st$ on the two covariates.
-For \textsc{geepers} and \textsc{psw}, we used standard maximum likelihood methods and data from the treatment group, achieving an AUC of 0.68; for the mixture model we incorporated the logistic regression as a sub-model within a larger Bayesian model fit in Stan.
-
-The results are displayed in Figure \ref{fig:opt} and Table \ref{tab:optEffs}: both \textsc{geepers} and the Bayesian mixture model gave similar results to those reported in \citet{richardson2023estimating} from BSIV.
+For \textsc{geepers} and \textsc{psw}, we used standard maximum likelihood methods and data from the treatment group, achieving an AUC of 0.68.
+For the mixture model, we incorporated the logistic regression specification as a sub-model within a larger Bayesian model fit in Stan.
+
+\begin{figure}
+  \centering
+  \includegraphics{figure/opts.pdf}
+  \caption{Estimated principal effects and 95\% confidence or credible intervals for the Obstetrics and Periodontal Therapy Study, using four estimation methods.
+Results for BSIV were taken from \citet{richardson2023estimating}.}
+  \label{fig:opt}
+\end{figure}
+
+The results are displayed in Figure \ref{fig:opt}% and Table \ref{tab:optEffs}
+: both \textsc{geepers} and the Bayesian mixture model gave similar results to those reported in \citet{richardson2023estimating} from BSIV.
 For the principal stratum of subjects who would not fully complete the treatment, those three methods estimated that assignment to the treatment condition reduced percentage of dental sites with bleeding by 31--34 percentage points; for subjects who would complete the treatment, the three estimated a reduction of 14-16 percentage points.
-These methods' conclusion is somewhat surprising: that the effect was larger for subjects who would not complete the treatment than for those who would (the two strata had similar bleeding percentages at baseline, and a logit transformation of the outcome leave the results qualitatively unchanged).
+The conclusion of these methods is somewhat surprising: that the effect was larger for subjects who would not complete the treatment than for those who would (the two strata had similar bleeding percentages at baseline, and a logit transformation of the outcome leaves the results qualitatively unchanged).
 One possible explanation is the conjecture that subjects stop showing up for treatment once their periodontal disease has abated.
 
-In contrast, prinicpal score weighting resulted in very similar effects for both strata---an reduction of 23--25 percentage points, which is close to the estimated intent to treat effect of 25 percentage points.
+In contrast, principal score weighting resulted in very similar effects for both strata---a reduction of 23--25 percentage points, which is close to the estimated intent-to-treat effect of 25 percentage points.
 
@@
 
-In the educational field experiment reported in \citet{impactPaper}, middle school students completing their math schoolwork on a computer program were randomized between four conditions.
+In the educational field experiment reported in \citet{impactPaper}, middle school students completing their math schoolwork on a computer program were randomly assigned to four conditions.
 These included two gamified programs---DragonBox and From Here to There (FH2T)---and two programs in which students used computers to work through a series of algebra problems taken from the district's textbook---``ASSISTments'' and an active control called ``Business as Usual'' (BAU).
@@
 They received immediate feedback on their answers and had to enter the correct answer to proceed to the next problem.
-In the BAU condition students had no access to hints and received feedback only after completing the assignment.
-
-In all four conditions, students worked in class on their assigned math program on 12 different occasions spaced throughout the school year: a prior assessment, a mid-test, a post-test, and nine learning sessions.
+In the BAU condition, students had no access to hints and received feedback only after completing the assignment.
+
+In all four conditions, students worked on their assigned math program in class on 12 different occasions spaced throughout the school year: a prior assessment, a mid-test, a post-test, and nine learning sessions.
 In this illustration, we will compare ASSISTments---the only condition featuring bottom-out hints---with each of the three others separately.
-Our goal is to estimate different treatment effects for groups of students who would, if assigned to ASSISTments, request few or many bottom-out hints.
+Our goal is to estimate different treatment effects for groups of students who, if assigned to ASSISTments, would request few or many bottom-out hints.
 
@@
 One way to disentangle the role of bottoming out in any treatment effect would be to estimate average treatment effects of being assigned to ASSISTments, relative to any of the other three, separately for subjects who would (if assigned to ASSISTments) bottom out frequently, and for subjects who would not.
-If the availability of bottom-out hints plays an important positive role in ASSISTments' efficacy, we might expect larger effects of condition on students who would bottom out frequently than for those who would not; if bottom-out hints are harmful, we might expect the opposite.
+If the availability of bottom-out hints plays a positive role in ASSISTments' efficacy, we might expect larger effects of the condition on students who frequently bottom out than on those who do not; if bottom-out hints are harmful, we might expect the opposite.
 If hints (that do not supply the correct answer) and error messages are helpful, but the ability to bottom out is harmful to learning, then we might expect to estimate a positive effect of being assigned to the feedback condition for students who do not bottom out, but a negative effect for students who do.
@@
 We gathered dichotomized data on %which students in the treatment condition requested at least one bottom-out hint.
-whether students in ASSISTments requested more than the median number of bottom-out hints (i.e. 11).
+whether students in ASSISTments requested more than the median number of bottom-out hints (i.e., 11).
 That is, $\st=1$ for students who, if assigned to ASSISTments, would request more than 11 bottom-out hints (``bottom-outers''), and $\st=0$ for students who would request 11 or fewer (``non-bottom-outers'').
@@
 We also had access to data on several baseline covariates, including prior achievement, demographics, baseline measurements of student attitudes toward math, and an indicator for whether students began the school year in remote or in-person instruction.
-Students with missing pretest scores were dropped from the analysis; missing data in other covariates was imputed with a Random Forest algorithm \citep{missForest}.
+Students with missing pretest scores were dropped from the analysis; missing data in other covariates were imputed with a Random Forest algorithm \citep{missForest}.
 Summary statistics are displayed in appendix tables \ref{table:tab1fac} and \ref{table:tab1num}.
@@
 %The data analysis we present here is intended as a demonstration of \textsc{geepers}, rather than for its substantive conclusions, and our description omits discussion of some important methodological considerations, including attrition and post-selection inference.
-For details on the experimental design, the conditions being compared, sample attrition---which was unusually high, due to the COVID-19 pandemic---and the impact analysis, please see \citet{impactPaper}.
+For details on the experimental design, the conditions being compared, sample attrition---which was unusually high due to the COVID-19 pandemic---and the impact analysis, please see \citet{impactPaper}.
 
@@
 First, Proposition \ref{prop:reg2} requires consistent estimation of principal scores and (via Assumption \ref{ass:rci}) correct specification of the outcome model, that is, the model for $\EE[Y|\bxy,Z,\st]$.
-Hence, it is essential to choose covariates and transformations $\bxs$ and $\bxy$ that lead to good model fit.
+Hence, it is essential to choose covariates and transformations $\bxs$ and $\bxy$ that lead to a good model fit.
 
 Second, Assumption \ref{ass:vps} requires principal scores to vary with at least some element $\bxs$.
-Moreover, the results of the simulation study suggest that when $\st$ can be predicted more precisely as a function of $\bxs$, \textsc{geeper}s principal effect estimators are more precise and more robust to model misspecification.
+Moreover, the results of the simulation study suggest that when $\st$ can be predicted more precisely as a function of $\bxs$, \textsc{geepers} principal effect estimators are more precise and more robust to model misspecification.
 Those considerations militate in favor of choosing covariates $\bxs$ to optimize out-of-sample principal score prediction accuracy.
 
-There is good reason to believe that analogous arguments apply to selection of outcome covariates $\bxy$.
+There is good reason to believe that analogous arguments apply to the selection of outcome covariates $\bxy$.
 
@@
 That is, in roughly 75--80\% of bottom-outer/non-bottom-outer pairs of subjects, the bottom-outer will have a higher predicted probability from the model.
-Coefficient estimates are displayed in table \ref{tab:psTab} in the appendix.
+Coefficient estimates are displayed in Table \ref{tab:psTab} in the appendix.
 
@@
 The three covariate selection strategies led to similar point and interval estimates.
-That said, in nearly all cases the AIC-optimal model led to the smallest standard errors.
+That said, in nearly all cases, the AIC-optimal model led to the smallest standard errors.
 Confidence intervals for all nine principal effect estimates were about twice as wide as for corresponding ATE estimates.
@@
 
-The other estimation strategies largely agreed with \textsc{geepers} and with each other in the contrast between ASSISTments and BAU, and, to a lesser extent, in teh contrast between ASSISTments and DragonBox.
+The other estimation strategies largely agreed with \textsc{geepers} and with each other in the contrast between ASSISTments and BAU, and, to a lesser extent, in the contrast between ASSISTments and DragonBox.
 The biggest difference between the methods was in the contrast with FH2T, in which the Bayesian mixture model resulted in a larger difference between the two principal effects, as well as the largest standard errors, while the two \textsc{psw} principal effect estimates were quite similar.
@@
 
-Here we will derive the sandwich variance-covariance matrix for the \textsc{geepers} estimate without interactions between $\bx$ and either $Z$ or $\st$---i.e.
-with $\bm{\gamma_2}=\bm{\gamma_3}=\bm{\gamma_4}=0$ in the notation of \eqref{eq:interaction}---and estimating principal scores using a generalized linear model.
+Here we will derive the sandwich variance-covariance matrix for the \textsc{geepers} estimate without interactions between $\bx$ and either $Z$ or $\st$---i.e., with $\bm{\gamma_2}=\bm{\gamma_3}=\bm{\gamma_4}=0$ in the notation of \eqref{eq:interaction}---and estimating principal scores using a generalized linear model.
 
@@
 (noting that $Z^2=Z$).
-These depend on $\bm{\alpha}$ when $r_i=p_i$, i.e. when $Z_i=0$.% or $S_i$ is missing.
+These depend on $\bm{\alpha}$ when $r_i=p_i$, i.e., when $Z_i=0$.% or $S_i$ is missing.
 Then note that, following \eqref{eq:ps}, and letting $\eta_i=\bm{\alpha}'\bxsit$
@@
 
-\processdelayedfloats
+%\processdelayedfloats
 
@@
 
-\caption{\label{tab:tab:optTab1}Descriptive statistics---mean and standard deviation or count and percent---for study variables in the full OPT dataset and in the analysis sample (i.e.
-complete cases)}
+\caption{\label{tab:optTab1}Descriptive statistics---mean and standard deviation or count and percent---for study variables in the full OPT dataset and in the analysis sample (i.e., complete cases)}
 \centering

### 2025-06-12 writeUps/paper.tex

@@
 
-Figure \ref{fig:boxplots} shows violin plots, layered on top of jittered scatter plots, showing estimation error for \textsc{geepers}, the mixture estimator, and the \textsc{psw} estimator of $\eff1$ under various sets of conditions, and Table \ref{tab:coverage} shows empirical coverage of nominal 95\% confidence interval estimates under a somewhat wider set of circumstances when $\muc1=0.3$.
+%<<<<<<< HEAD
+Figure \ref{fig:boxplots} shows violin plots, layered on top of jittered scatter plots, showing estimation error for \textsc{geepers}, the parametric mixture model, and the \textsc{psw} estimator of $\eff1$ for, and Table \ref{tab:coverage} shows empirical coverage of nominal 95\% confidence interval estimates for \textsc{geepers} and a parametric mixture models; when at least one of an estimator's assumptions is violated, the result is colored red.
+
+The first row of Table \ref{tab:coverage} and the first column of Figure \ref{fig:boxplots} show similar results as Figure \ref{fig:alphan}---when all assumptions of both methods are met, \textsc{geepers} is slightly less biased than the parametric mixture model; when $\alpha=0.2$, \textsc{geepers} is much more variable and $\alpha=0.5$ the difference in variance is slight.
+\textsc{psw} has by far the lowest variance and the hightest bias; the latter is due to the violation of principal ignorability.
+
+The remainder of the results reflect violations modeling assumptions for either \geepers or the parametric mixture model or both: interactions in the data generating model between covariates $X$ and treatment assignment $Z$ and/or principal stratum $S$, either of which would violate assumption \ref{ass:rci} and the linear additive model of the \textsc{pmm}, and/or residuals drawn from a uniform distribution, which violates the normality assumption of \textsc{pmm}.
+For \geepers, when $\alpha=0.2$ the presence of interactions in the data generating model increases sampling variance, but only interactions between $X$ and $Z$ reduce confidence interval coverage.
+When $\alpha=0.5$, the presence of interactions has little to no effect on the performance of the \geepers estimator, and all 95\% confidence intervals approximately acheive their nominal levels.
+\geepers makes no assumptions about the shape of the residual distribution and behaves similarly under either normal or uniform residuals.
+
+Like \geepers, \pmm is insensitive to interactions between $X$ and $S$.
+However, interactions between $X$ and $Z$ in the data generating model skew the \pmm sampling distribution and reduce the coverage of its 95\% confidence intervals, even when $\alpha=0.5$.
+When the residual distribution was uniformly distributed, the \pmm sampling distribution appeared bimodal, and its credible intervals exhibited severe undercoverage.
+
+In sum, \psw is the least variable of all three estimator, and does not depend on outcome modeling, but is vulnerable to bias from an unmeasured covariate that predicts both $S$ and $Y$.
+When all assumptions of both methods are met, \geepers is less efficient, but also less biased, than \pmm, but differences in efficiency decrease with $n$ and $\alpha$.
+Neither \psw nor \pmm are vulnerable to interactions beetween $X$ and $S$.
+\pmm is highly vulnerable to violations of normality and to interactions between $X$ and $Z$, while \psw is robust to non-normality and robust to $X$-$Z$ interactions when $\alpha-0.5$.
+%=======
+% Figure \ref{fig:boxplots} shows violin plots, layered on top of jittered scatter plots, showing estimation error for \textsc{geepers}, the mixture estimator, and the \textsc{psw} estimator of $\eff1$ under various sets of conditions, and Table \ref{tab:coverage} shows empirical coverage of nominal 95\% confidence interval estimates under a somewhat wider set of circumstances when $\muc1=0.3$.
 More complete results, including root-mean-squared-error estimates, are in the online appendix.
-The leftmost panel of Figure \ref{fig:boxplots} and the top line of Table \ref{tab:coverage} show results under conditions favorable to both the mixture model and to \textsc{geepers}---normal residuals and no interactions in the data generating model between covariates and either $S$ or $Z$.
-\textsc{geepers} was roughly unbiased, but less precise than the other two methods---when $\alpha=0.2$ it was much noisier, and when $\alpha=0.5$ it was only slightly less precise.
-The mixture estimator's bias depended on $\muc1$---when $\muc1=0$, the bias was slight, and when $\muc1=0.3$, the bias was more substantial.
+% The leftmost panel of Figure \ref{fig:boxplots} and the top line of Table \ref{tab:coverage} show results under conditions favorable to both the mixture model and to \textsc{geepers}---normal residuals and no interactions in the data generating model between covariates and either $S$ or $Z$.
+% \textsc{geepers} was roughly unbiased, but less precise than the other two methods---when $\alpha=0.2$ it was much noisier, and when $\alpha=0.5$ it was only slightly less precise.
+% The mixture estimator's bias depended on $\muc1$---when $\muc1=0$, the bias was slight, and when $\muc1=0.3$, the bias was more substantial.
 This result is somewhat surprising, because when $\muc1=0$, there was no separation between the principal strata in the control group, and previous literature \citep{griffin2008application} has identified this scenario as a particular challenge for mixture modeling.
-The \textsc{psw} estimator was also biased, due, presumably, to the violation of Principal Ignorability, Assumption \ref{ass:PI}, inherent in the data generating model.
+% The \textsc{psw} estimator was also biased, due, presumably, to the violation of Principal Ignorability, Assumption \ref{ass:PI}, inherent in the data generating model.
 On the other hand, the \textsc{psw} estimator had easily the lowest sampling variance of the three estimators.
 
-The middle panel of Figure \ref{fig:boxplots} and the fifth line of Table \ref{tab:coverage} show results for when the residuals in \eqref{eq:y-sim} are uniformly distributed.
-In this case, \textsc{geepers} and \textsc{psw} estimators, which do not assume normality, behaved in a similar fashion as in the normal case, but the mixture estimator showed a bimodal pattern--sometimes over-estimating $\eff1$ and sometimes underestimating, but rarely estimating $\eff1$ accurately.
-Credible intervals from the mixture model exhibited severe undercoverage, while \textsc{geepers} confidence intervals achieved at least their nominal level.
-
-%\sloppy
-The rightmost panel of Figure \ref{fig:boxplots} returns to the case of normal residuals---and only shows results when $\muc1=0.3$ and $\alpha=0.5$---but introduces interactions in the data generating model between covariates and either $\st$, $Z$, or both, violating Assumption \ref{ass:rci}.
-%Presence of interactions could undermine the \textsc{geepers} by inducing a relationship between principal scores and potential outcomes, even after adjusting the outcomes for covariate effects.
-Fortunately, these interactions made little difference when $\alpha=0.5$.
+% The middle panel of Figure \ref{fig:boxplots} and the fifth line of Table \ref{tab:coverage} show results for when the residuals in \eqref{eq:y-sim} are uniformly distributed.
+% In this case, \textsc{geepers} and \textsc{psw} estimators, which do not assume normality, behaved in a similar fashion as in the normal case, but the mixture estimator showed a bimodal pattern--sometimes over-estimating $\eff1$ and sometimes underestimating, but rarely estimating $\eff1$ accurately.
+% Credible intervals from the mixture model exhibited severe undercoverage, while \textsc{geepers} confidence intervals achieved at least their nominal level.
+
+% %\sloppy
+% The rightmost panel of Figure \ref{fig:boxplots} returns to the case of normal residuals---and only shows results when $\muc1=0.3$ and $\alpha=0.5$---but introduces interactions in the data generating model between covariates and either $\st$, $Z$, or both, violating Assumption \ref{ass:rci}.
+% %Presence of interactions could undermine the \textsc{geepers} by inducing a relationship between principal scores and potential outcomes, even after adjusting the outcomes for covariate effects.
+% Fortunately, these interactions made little difference when $\alpha=0.5$.
 However, Table \ref{tab:coverage} shows that when $\alpha=0$ or 0.2, and interactions between $\bx$ and $Z$ were present (i.e., $\gamma_3\ne 0$), \textsc{geepers} 95\% confidence intervals tended to under-cover.
 Nevertheless, coverage of \textsc{geepers} intervals was nearly always substantially higher than coverage of corresponding 95\% credible intervals from mixture models.%\footnote{}
+% >>>>>>> overleaf-2025-06-12-1948
 
@@
 
+\subsubsection{Principal Stratification and Alternatives}
+The goal of principal strafication in this context is to determine if assignment to the ASSISTments condition affected students who tend to request frequent bottom-out hints differently than students who tended to request fewer bottom-out hints.
+Is this the right question to ask?
+
+In a similar context, \citet{sales2021student} suggests two alternative appproaches: the average effect of requesting bottom-out hints on posttest scores for students in the ASSISTments condition, and estimating natural direct and indirect effects of assignment to the ASSISTments condition, treating bottom-out hint requests as a mediator.
+The former approach is an observational study: although students were randomized to the ASSISTments condition, bottom-out hint requests were not randomized.
+Therefore, estimating the effect of bottom-out hints requires careful control of confounding.
+Unfortunately, confounding is a major concern in this case, since, presumably, students only requested bottom-out hints if they were struggling with the material, and students who struggled with practice problems are likely to score lower on the posttest, on average, than those who needed less help during practice.
+The best available measure of students' baseline ability, their score on the 10-item pretest, is likely insufficient.
+That said, a plausible estimate of the average effect of bottom-out hint requests would be a valuable contribution to the literature, and a suitably innovative observational design \citep[e.g.][]{beck} may bear fruit.
+
+Another arguable disadvantage of the observational study approach is that it ignores the larger context of the experiment, asking whether bottom-out hints affect posttest scores, rather than asking what role they play in ASSISTments' overall effectiveness.
+A second approach, mediation analysis, contextualizes the bottom-out effect within the larger experiment.
+It, too, relies on the same non-confounding assumption as an observational study, and also suffers from an additional challenge: mediation analysis relies on a ``common support'' condition that ``the mediator state must not be a deterministic function of the treatment'' \citep[p.
+225]{celli2022causal}.
+However, bottom-out hints are unavailable in the other three arms of the study, violating this condition (\citealt{sales2021student} offers a solution, which we find unconvincing).
+
+A principal stratification approach treats bottom-out hint requests as a type of compliance.
+In other words, if we consider full implementation of the ASSISTments condition to include frequent bottom-out hint requests, students assigned to ASSISTments but requests few bottom-out hints is only partially complying with their treatment assignment.
+Then the question is if students who do not fully comply with their treatment assignment still experiance an effect.
+Conversely, might students who only partially comply experience a larger positive effect?
+That would be the case if either bottom-out hints were harmful, or if the treatment was better targeted toward the type of student who tends to need them less.
+
+Principal score weighting depends on the assumption of principal ignorability, which is equivalent to the no-confounding assumption of the observational study and mediation analysis.
+The mixture model and \textsc{geepers} estimators do not rely on that assumption, instead relying on parametric models and residualized covariate independence, respectively.
+Finally, the instrumental variables estimator of \citet{air} relies on the ``exclusion restriction'' that, for students who request few bottom-out hints, there is no effect of randomization to the ASSISTments condition.
+Since bottom-out hints are only one of several features of ASSISTments, this assumption is untenable.
+
+\subsubsection{Data on Bottom-Out Hints and Covariates}
+
 \sloppy

### 2025-06-17 code/regression.r
fixed interaction bug

@@ -72,7 +72,7 @@ datNames <- function(data,trt='Z',out='Y',use='S',block=NULL){
 }
 
 ### computes/returns regression models
-pointEst <- function(data,covFormU=~x1+x2,covFormY=covFormU,intS=NULL,psMod=NULL){
+pointEst <- function(data,covFormU=~x1+x2,covFormY=covFormU,intSx=NULL,psMod=NULL){
 
   data$S[data$Z==0] <- NA
 
@@ -94,8 +94,8 @@ pointEst <- function(data,covFormU=~x1+x2,covFormY=covFormU,intS=NULL,psMod=NULL
   outForm <- update(covFormY,Y~Z*Sp+.)
 #  outForm <- update(covFormY,Y~Z*Sp+.)
 
-  if(!is.null(intS))
-    outForm <- update(outForm,as.formula(paste0('.~.+Sp:(',paste(intS,collapse='+'),')')))
+  if(!is.null(intSx))
+    outForm <- update(outForm,as.formula(paste0('.~.+Sp:(',paste(intSx,collapse='+'),')')))
 
 
   if('block'%in%names(data)) if(!is.null(data$block))
@@ -195,14 +195,14 @@ vcvPS <- function(psMod,outMod,data,clust=NULL,int=any(grepl(":x",names(coef(out
 
 
 ### wrapper function for estimating regressions+vcov
-est <- function(data,covFormU=~x1+x2,covFormY=NULL,psMod=NULL,clust=NULL,intS=NULL,
+est <- function(data,covFormU=~x1+x2,covFormY=NULL,psMod=NULL,clust=NULL,intSx=NULL,
                 trt='Z',out='Y',use='S',block=NULL){
 
   #if(!missing(psMod))
 
   data <- datNames(data,trt=trt,out=out,use=use,block=block)
 
-  Attach(pointEst(data=data,covFormU=covFormU,covFormY=covFormY,intS=intS,psMod=psMod))
+  Attach(pointEst(data=data,covFormU=covFormU,covFormY=covFormY,intSx=intSx,psMod=psMod))
 
   vcv <- vcvPS(psMod,outMod,data=data,clust=clust)
 
@@ -235,10 +235,10 @@ effsFromFit <- function(ests){
 }
 
 ### estimates effects of interest, starting from data
[…]

### 2025-06-17 code/simulation/runSimBayes.r

@@ -60,3 +60,5 @@ fullsim(500,ns=c(200,300,400,600,700,800),errDist='norm',b1s=.5,intS=FALSE,intZ=
 
 
 fullsim(500,ns=500,errDist='norm',b1s=c(.3,.4,seq(.6,1,.1)),intS=FALSE,intZ=FALSE,mu01=.3,mu10=0,ext='b1s')
+
+fullsim(500,ns=500,errDist=c("norm","unif"),b1s=c(0.3,0.4),intZ=TRUE,mu10=0,mu01=0.3,ext="alpha34")

### 2025-06-17 code/simulation/simFuncs.r

@@ -133,6 +133,8 @@ fullsim <- function(nsim,
                     errDist=c('norm','mix','unif'),
                     b1s=c(0,0.2,0.5),
                     ext='',
+                    intS=c(TRUE,FALSE),
+                    intZ=c(TRUE,FALSE),
                     se=TRUE,
                     ncores=8,
 		    cl=NULL,
@@ -145,9 +147,9 @@ fullsim <- function(nsim,
 		mu10=mu10,
 		mu11=mu11,
 		errDist=errDist,
-		b1=b1s,
-		intS=c(TRUE,FALSE),
-		intZ=c(TRUE,FALSE),
+                b1=b1s,
+                intS=intS,
+                intZ=intZ,
 		stringsAsFactors=FALSE)
 
     cat(nrow(cases),' conditions\n')

### 2025-06-17 writeUps/paper.tex

@@
 \renewcommand\bibfont{\fontsize{10}{12}\selectfont}% To set the list of references in 10 point font using natbib.sty
+
+\usepackage{xspace}
 
@@
 
-The conceptual basis for principal stratification---though not the terminology---emerged from the study of randomized experiments with noncompliance.
+\label{background}The conceptual basis for principal stratification---though not the terminology---emerged from the study of randomized experiments with noncompliance.
 Specifically, \citet{air} showed that the econometric instrumental variables estimand is the average effect of the treatment on the principal stratum of ``compliers''---those subjects who would receive the treatment if and only if randomized to the treatment condition.
@@
 
-\subsection{M-Estimation}
+\subsection{M-Estimation}\label{sec:mest}
 M-estimation (also generalized method of moments or generalized estimating equations) is a general approach to deriving consistent statistical estimators and their asymptotic distributions.
@@
 
-\subsubsection{Data Generation}
+\subsubsection{Data Generation}\label{sec:dataGeneration}
 
@@
 
-Potential and observed outcomes were generated as
+\label{potential}Potential and observed outcomes were generated as
 \begin{equation}\label{eq:y-sim}
@@
 
-\subsubsection{Manipulated Factors}
-We manipulated six factors in the simulation, listed in Table \ref{tab:factor}.
-The six factors are not completely crossed---in particular, we let $n$ and $\alpha$ each vary across all the levels listed while holding the other five factors fixed.
+\subsubsection{Manipulated Factors}\label{sec:manipulatedFactors}
+We manipulated five factors in the simulation, listed in Table \ref{tab:factor}.
+The five factors are not completely crossed---in particular, we let $n$ and $\alpha$ each vary across all the levels listed while holding the other four factors fixed.
 
@@
   \centering
-\begin{tabular}{*{2}{c}}
+\begin{tabular}{r|rl}%{*{3}{c}}
   \hline
-  Factor &Levels\\
+  Factor &\multicolumn{2}{l}{Levels}\\
   \hline
-Sample size per condition $n$ &$100,200,\dots,1000$\\
-$\alpha$ &$0,0.2,0.3,\dots,1$\\
-The distribution of $\epsilon_i$& Normal, Uniform\\
-$\beta_1$ & 0, 0.3\\
-Interactions between $\st$ and covariates & No, Yes\\
-Interactions between $Z$ and covariates& No, Yes\\
+Sample size per condition $n$ &\multicolumn{2}{l}{$100,200,\dots,900,1000$}\\
+$\alpha$ &\multicolumn{2}{l}{$0,0.2,0.3,\dots,0.9,1$}\\
+The distribution of $\epsilon_i$&\multicolumn{2}{l}{ Normal, Uniform}\\
+$\beta_1$ & \multicolumn{2}{l}{0, 0.3}\\
+Interactions in \eqref{eq:y-sim} & None & $\gamma_1=1/\sqrt{6};\gamma_2=\gamma_3=0$\\ 
+& $\bm{x}:\st$ & $\gamma_1=3/(4\sqrt{6})$; $\gamma_2=1/(2\sqrt{6})$; $\gamma_3=0$\\
+& $\bm{x}:Z$&$\gamma_1=1/\sqrt{6};\gamma_2=0$; $\gamma_3=1/(2\sqrt{6})$\\
+& $\bm{x}:\st$ \& $\bm{x}:Z$& $\gamma_1=3/(4\sqrt{6})$; $\gamma_2=\gamma_3=1/(2\sqrt{6})$\\
+% between $\st$ and covariates & No, Yes\\
+% Interactions between $Z$ and covariates& No, Yes\\
 \hline
@@
 
-\textbf{\emph{Sample Size per condition $\bm{n}$:}} The guarantees of M-estimation are all asymptotic, so we examined the behavior of our \textsc{geepers} at a range of sample sizes to determine its finite-sample properties.
+\textbf{\emph{Sample Size per condition $\bm{n}$:}} The guarantees of M-estimation are all asymptotic, so we examined the behavior of \textsc{geepers} at a range of sample sizes to determine its finite-sample properties.
 The range of 100--1000 was chosen to represent typical student-level randomized trials in education, for instance, \citet{growthMindsetRuralBurnette} with $n=115$ in the treatment group and $n=107$ in the control group, or \citet{impactPaper}, which initially randomized an average of 1000 students to each condition.
@@
 For the treatment group, the regression model $Y=\beta_0^T+\beta_1^TS+\bm{\gamma}'\bx+\epsilon$ and for the control group, a mixture model of the form \eqref{eq:mixtureModelPS}, but with $\muc0$ and $\muc1$ replaced by regression functions $\beta_0^C+\bm{\gamma}'\bx$ and $\beta_0^C+\beta_1^C+\bm{\gamma}'\bx$, respectively.
-Residual variance were allowed to differ across treatment groups, but not across principal strata.
+Residual variance was allowed to differ across treatment groups, but not across principal strata.
 These models were fit simultaneously using Bayesian Markov Chain Monte Carlo with Stan \citep{rstan}, and principal effects $\eff0$ and $\eff1$ estimated as $\beta_0^T-\beta_0^C$ and $\beta_0^T+\beta_1^T-\beta_0^C-\beta_1^T$, respectively.
@@
 
-\subsection{Results}
-In this section, we present the subset of simulation results that we think is the most telling and interesting; a fuller set of results is available in the online appendix.
+\subsection{Results}\label{sec:simResults}
+In this section, we evaluate estimates of $\eff1$. Results for estimates of $\eff0$ and $\eff1-\eff0$, which are similar, can be found in the online appendix. 
+In replication $b=1,\dots,500$, define the error of a given estimator $\heff1_b$ as $err_b=\heff1_b-\eff1$.
+The bias of $\heff1$, $\EE[\heff1-\eff1]$, is estimated as $\widehat{bias}=\sum_b err_b/500$, and the empirical standard error is defined as $\left\{\sum_b(\heff1_b-\overline{\heff1})^2/499\right\}^{1/2}$, the sample standard deviation of $\heff1$.
+We also report the empirical coverage of confidence intervals, the proportion of replications in which $\eff1$ is within the estimated confidence bounds of $\heff1$.
+Here, we present results on error, bias, empirical standard error, and confidence itnterval coverage; results on root-mean-squared-error %$\left(\widehat{bias}^2+SE_{emp}^2\right)^{1/2}$
+$\left(\sum_b err_b^2/500\right)^{1/2}$
+are available in the online appendix.
 
@@
   \includegraphics[width=4.5in]{../simFigs/biasSEbyB1N.jpg}
-  \caption{Empirical bias and standard error of \textsc{geepers} and Mixture estimates of $\mu_C^1$ as sample size per treatment group $n$ or $\alpha$ varied.
+  \caption{Empirical bias and standard error of \textsc{geepers} and Mixture estimates of $\eff1$ as sample size per treatment group $n$ or $\alpha$ varied.
 There were 500 replications for each $n$ or $\alpha$.
@@
 The first row of Table \ref{tab:coverage} and the first column of Figure \ref{fig:boxplots} show similar results as Figure \ref{fig:alphan}---when all assumptions of both methods are met, \textsc{geepers} is slightly less biased than the parametric mixture model; when $\alpha=0.2$, \textsc{geepers} is much more variable and $\alpha=0.5$ the difference in variance is slight.
-\textsc{psw} has by far the lowest variance and the hightest bias; the latter is due to the violation of principal ignorability.
+\textsc{psw} has by far the lowest variance and the highest bias; the latter is due to the violation of principal ignorability.
 
 The remainder of the results reflect violations modeling assumptions for either \geepers or the parametric mixture model or both: interactions in the data generating model between covariates $X$ and treatment assignment $Z$ and/or principal stratum $S$, either of which would violate assumption \ref{ass:rci} and the linear additive model of the \textsc{pmm}, and/or residuals drawn from a uniform distribution, which violates the normality assumption of \textsc{pmm}.
-For \geepers, when $\alpha=0.2$ the presence of interactions in the data generating model increases sampling variance, but only interactions between $X$ and $Z$ reduce confidence interval coverage.
-When $\alpha=0.5$, the presence of interactions has little to no effect on the performance of the \geepers estimator, and all 95\% confidence intervals approximately acheive their nominal levels.
+For \geepers, when $\alpha=0.2$ the presence of interactions in the data-generating model increases sampling variance, but only interactions between $X$ and $Z$ reduce confidence interval coverage.
+When $\alpha=0.5$, the presence of interactions has little to no effect on the performance of the \geepers estimator, and all 95\% confidence intervals approximately achieve their nominal levels.
 \geepers makes no assumptions about the shape of the residual distribution and behaves similarly under either normal or uniform residuals.
@@
 
-In sum, \psw is the least variable of all three estimator, and does not depend on outcome modeling, but is vulnerable to bias from an unmeasured covariate that predicts both $S$ and $Y$.
+\label{simsum}In sum, \psw is the least variable of all three estimators, and does not depend on outcome modeling, but is vulnerable to bias from an unmeasured covariate that predicts both $S$ and $Y$.
 When all assumptions of both methods are met, \geepers is less efficient, but also less biased, than \pmm, but differences in efficiency decrease with $n$ and $\alpha$.
-Neither \psw nor \pmm are vulnerable to interactions beetween $X$ and $S$.
-\pmm is highly vulnerable to violations of normality and to interactions between $X$ and $Z$, while \psw is robust to non-normality and robust to $X$-$Z$ interactions when $\alpha-0.5$.
+Neither \psw nor \pmm is vulnerable to interactions between $X$ and $S$.
+\pmm is highly vulnerable to violations of normality and to interactions between $X$ and $Z$, while \psw is robust to non-normality and robust to $X$-$Z$ interactions when $\alpha = 0.5$.
+
 %=======
@@
 
-In a similar context, \citet{sales2021student} suggests two alternative appproaches: the average effect of requesting bottom-out hints on posttest scores for students in the ASSISTments condition, and estimating natural direct and indirect effects of assignment to the ASSISTments condition, treating bottom-out hint requests as a mediator.
+In a similar context, \citet{sales2021student} suggests two alternative approaches: the average effect of requesting bottom-out hints on posttest scores for students in the ASSISTments condition, and estimating natural direct and indirect effects of assignment to the ASSISTments condition, treating bottom-out hint requests as a mediator.
 The former approach is an observational study: although students were randomized to the ASSISTments condition, bottom-out hint requests were not randomized.
@@
 
-Another arguable disadvantage of the observational study approach is that it ignores the larger context of the experiment, asking whether bottom-out hints affect posttest scores, rather than asking what role they play in ASSISTments' overall effectiveness.
+Another arguable disadvantage of the observational study approach is that it overlooks the broader context of the experiment, focusing on whether bottom-out hints affect posttest scores rather than examining their role in ASSISTments' overall effectiveness.
 A second approach, mediation analysis, contextualizes the bottom-out effect within the larger experiment.
@@
 A principal stratification approach treats bottom-out hint requests as a type of compliance.
-In other words, if we consider full implementation of the ASSISTments condition to include frequent bottom-out hint requests, students assigned to ASSISTments but requests few bottom-out hints is only partially complying with their treatment assignment.
-Then the question is if students who do not fully comply with their treatment assignment still experiance an effect.
+In other words, if we consider full implementation of the ASSISTments condition to include frequent bottom-out hint requests, students assigned to ASSISTments but who requested few bottom-out hints only partially complied with their treatment assignment.
+The question then is whether students who do not fully comply with their treatment assignment still experience an effect.
 Conversely, might students who only partially comply experience a larger positive effect?
-That would be the case if either bottom-out hints were harmful, or if the treatment was better targeted toward the type of student who tends to need them less.
+That would be the case if either bottom-out hints were harmful or if the treatment was better targeted toward the type of student who tends to need them less.
 
@@
 
+The simulation study suggests that none of the three methods evaluated, \geepers, \nmm, or \psw, dominates the others, and offers some guidance on when each estimator may be preferable.
+\psw is the most precise and requires no assumptions about the distribution of outcomes, but can be quite sensitive to violations of principal ignorability.
+When the determinants of $\st$ are well-understood and measured, \psw seems to be the best option.
+When principal ignorability does not hold, potential outcomes are conditionally normal, and there are no $X:Z$ interactions within principal strata, the \nmm estimator outperforms \geepers, though the difference decreases with $n$ or $\alpha$.
+
 Sampling error prevents firm conclusions regarding principal effects for students who would, if given the opportunity, request frequent bottom-out hints.
@@
 
-\bibliographystyle{plainnat}
+\bibliographystyle{apalike}
 \bibliography{MOM}
@@
 
-\clearpage
-
-\section{Responses to Reviewers}
-We are grateful for the reviewers' and the editor's careful reading
-and comments. We believe that the the substantial changes we made
-chages to the manuscript in response to those comments greatly
-improved the paper. Most significantly, we clarified our presentation
-and added a second application.
-We provide point-by-point responses below.
-
-\subsection{Reviewer 1}
-\begin{enumerate}
-
-  \item \begin{quote}The rationale for using Principal Stratification to estimate treatment
-effects for bottom-outers is insufficient. Additionally, it is
-advisable to compare the Principal Stratification method to other
-causal inference methods, such as instrumental variable and propensity score analysis, to provide a comprehensive understanding of the chosen
-approach.
-\end{quote}
-\item \begin{quote}
-    While the authors effectively describe the novelty of their study, the literature review lacks the historical
-development of Principal Stratification. Including more in-depth reviews and relevant citations would enhance the
-background information and context of the study.
-\end{quote}
-\item
-\begin{quote}
-  The simulation details about how variable Z is simulated and the number of replications for each condition
-are not clearly explained. Providing more information on these aspects would contribute to the transparency and replicability
-of the study.
-\end{quote}
-\item \begin{quote}
-    The rationale for selecting specific values for manipulated factors, such as choosing a sample size range of
-100 to 1000 and using N(0,1/2) instead of N(0,1), is not adequately justified. Including evidence and citations to support
-these choices would strengthen the methodological foundation of the
-study.
-\end{quote}
-\item \begin{quote}
-    A section detailing the evaluation metrics and criteria used to assess simulation results should precede the
-presentation of results. This would provide readers with a clear understanding of the benchmarks against which the findings
-are measured.
-\end{quote}
-\item
- \begin{quote} It is highly recommended that the authors employ a series of ANOVAs to analyze the impact of various
-factors on the accuracy of effect estimation to guide results interpretation. Additionally, a concise paragraph summarizing key
-findings before the application example would enhance the clarity and
-interpretation of the results.
-\end{quote}
-\item
-  \begin{quote}
-    APA Citations and References: The paper does not adhere to the APA 7th style. Authors are advised to consult the APA
-    Formatting and Style Guide %(https://owl.purdue.edu/owl/research_and_citation/apa_style/apa_formatting_and_style_guide/index.html)
-    for proper citation and reference formatting throughout the manuscript. Consistency in style is crucial
-for scholarly publications.
-\end{quote}
-
-\end{enumerate}
-
-\subsection{Reviewer 2}
-\begin{quote}
-Great work. Well written. Please find some of my comments below. I
-hope they are useful.
-\end{quote}
-
-\begin{itemize}
-\item
-  \begin{quote} Please provide more background on M-estimators especially to the extent it evolves to become relevant to the
-    GEEPERs.
-    \end{quote}
-\item
-  \begin{quote}
-I am wondering if doubly robust PSW estimator would reduce the remaining bias difference between GEEPERs and
-PSW, and Mixture and PSW.
-    \end{quote}
-
-\item
-  \begin{quote}
-    GEEPERs seems great, but also, surprised by how efficient Mixture can be. Are there any mechanisms that link M-
-    estimators to Mixture modeling? Just wondering.
-        \end{quote}
-
-\item
-  \begin{quote}
-    Would it be possible to offer some evaluation metric for GEEPERs,
-    which would indicate, under some conditions researchers could
-    possibly opt or seek other estimators
-  \end{quote}
-
-\item
-  \begin{quote}
-    Quickly checked Frangakis and Rubin (2002). Do you think the effect of bottom-outers (or similar experimental
-designs) could be partitioned into associative and dissociative
-effects with GEEPERs?
-    \end{quote}
-
-\item
-  \begin{quote}
-    It would be great if you could share the code for estimation and simulation in the next round of review (if there is). Also
-make sure you have comments in the code so someone reading the article
-can match the narrative with the code.
-    \end{quote}
-
-\end{itemize}
-
-\subsection{Reviewer: 3}
-Comments to the Author
-This paper develops a new estimator for principal stratification based on M-estimation and stacked equations. The proposed
-method is designed to be more robust and simpler to use than standard methods. The method is, however, tailored to a very
-specific set up: one-way noncompliance. In general, principal stratification is a much more general framework that can be
-used for a wide variety of settings where there is conditioning on some post-treatment quantity.
-The method is outlined, and then evaluated through a series of simulations, and then an application is provided. My primary
-criticism is that the new method doesn't appear to outperform extant methods to a compelling degree. I take up this point in
-more detail first for the simulations and second for the empirical
-application.
-
-\subsubsection{Simulations}
-\begin{quote}
-  In my view, the simulations should show the following:
-  \begin{itemize}
-    \item When the assumptions of extant methods are met, GEEPERS should do no worse.
-    \item When the assumptions of extend methods are not met, GEEPERS
-      should do better.
-  \end{itemize}
-Given this criteria, it seems to me that the simulation should focus on the last three factors in the design. That is, I would
-hold the sample size and alpha fixed. Currently too many factors are varied at the same time to give a clear picture. I would
-alter Table 1, to indicate which factors yield assumption violations and then violate these assumptions in isolation instead of
-relaxing them simultaneously. As it currently stands, it is quite hard to understand the relative performance of GEEPERs
-compared to extant methods. Careful relaxation of each assumption allows the reader to see how much bias results from a
-single factor.
-\end{quote}
-
-\subsubsection{Application.}
-I thought there are a number of less than desirable parts to the application. First, the focus on bottoming out is less than
-ideal. This strikes me as a fairly rare phenomenon. I didn't find the concept very intuitive and had to read just the descriptive
-section several times to understand why one might be interested in a principal stratification analysis here. This makes the
-methods seem not particularly general. Why not focus on an application with straightforward oneway noncompliance? This
-is an area where principal stratification is often applied, and the concepts are well understood.
-In addition, it isn't ideal that the ATE in each condition is not estimated very precisely, that is the CI appears to cover zero in
-all cases. Next, all the principal stratification estimates are all pretty close to the same value as the ATE. That serves as
-another reason this application isn't compelling. It is hard to care about an application where principal stratification doesn't
-provide much in the way of insights relative to the ATE. Next, the performance of GEEPERs seems to be quite similar to
-extant methods in each scenario. Overall, to my thinking, there is little to recommend the current application. Again, there is
-little here that demonstrates the value of GEEPERs.
-To my mind, these problems could both be fixed. That is, the simulations could be redesigned to provide more compelling
-evidence. Next, the authors may be able to identify are better application. Therefore, the manuscript should be revised and
-resubmitted.
-
 \end{document}

### 2025-06-17 writeUps/responses.tex

\documentclass[]{article}

\usepackage[letterpaper]{geometry}

\usepackage{xr}
\externaldocument{paper}

\RequirePackage{amsmath,amsfonts,amssymb}
%\RequirePackage[authoryear]{natbib}
\usepackage{longtable}
%\usepackage[notablist,tablesfirst]{endfloat}% To `separate' figures and tables from text if required

\usepackage{epstopdf}% To incorporate .eps illustrations using PDFLaTeX, etc.
\usepackage[caption=false]{subfig}% Support for small, `sub' figures and tables
\usepackage[doublespacing]{setspace}% To produce a `double spaced' document if required
%\setlength\parindent{24pt}% To increase paragraph indentation when line spacing is doubled

\usepackage[longnamesfirst,sort]{natbib}% Citation support using natbib.sty
\bibpunct[, ]{(}{)}{;}{a}{,}{,}% Citation support using natbib.sty
\renewcommand\bibfont{\fontsize{10}{12}\selectfont}% To set the list of references in 10 point font using natbib.sty

% \theoremstyle{plain}% Theorem-like structures provided by asthma.sty
% \newtheorem{theorem}{Theorem}[section]
% \newtheorem{lemma}[theorem]{Lemma}
% \newtheorem{corollary}[theorem]{Corollary}
% \newtheorem{prop}[theorem]{Proposition}

% \theoremstyle{definition}
% \newtheorem{definition}[theorem]{Definition}
% \newtheorem{example}[theorem]{Example}

% \theoremstyle{remark}
% \newtheorem{remark}{Remark}
% \newtheorem{notation}{Notation}

\usepackage{amsthm}
\usepackage[utf8]{inputenc} % set input encoding (not needed with XeLaTeX)
\usepackage{bm}
\usepackage[hidelinks]{hyperref}
\usepackage{graphicx}
\usepackage{mathtools}
%\usepackage{fullpage}
\usepackage{makecell}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{numdef}
%\usepackage{multibib}
\usepackage{rotating}

\usepackage[dvipsnames]{xcolor}

%\startlocaldefs

\newcommand{\rd}[1]{\textcolor{BrickRed}{#1}}

\input{notation.tex}

%\endlocaldefs
%\newcites{supp}{Supplementary References}

%\begin{frontmatter}

\newenvironment{itquote}
  {\begin{quote} \itshape}
  {\end{quote}\ignorespacesafterend}

\title{Responses to Review Comments}
%\runtitle{\textsc{geepers}}
%\maketitle
\date{}

\begin{document}

\maketitle
We are grateful for the reviewers' and the editor's careful reading
and comments. We believe that the the substantial changes we made
chages to the manuscript in response to those comments greatly
improved the paper. Most significantly, we clarified our presentation
and added a second application.
We provide point-by-point responses below.

\section*{Reviewer 1}

The article effectively presents a simulation study comparing three methods and provides a clear application example of applying GEEPERs. However, there are notable areas for improvement, including a lack of literature review on previous simulation studies about Principal Stratification and some unclear presentations of results. Specific comments are outlined below.

\begin{enumerate}

\item    \begin{itquote}  Page 3, Line 10: The rationale for using Principal Stratification to estimate treatment effects for bottom-outers is insufficient. Additionally, it is advisable to compare the Principal Stratification method to other causal inference methods, such as instrumental variable and propensity score analysis, to provide a comprehensive understanding of the chosen approach.
\end{itquote}

\item \begin{itquote}

     Page 5, Line 5: While the authors effectively describe the novelty of their study, the literature review lacks the historical development of Principal Stratification. Including more in-depth reviews and relevant citations would enhance the background information and context of the study.
\end{itquote}

Thanks for this suggestion. We have added material to the introduction, starting from the paragraph beginning ``The conceptual basis'' on page \pageref{background} and through the following page, describing some of the history and background of principal stratification.

\item \begin{itquote}
     Page 15: The simulation details about how variable Z is simulated and the number of replications for each condition are not clearly explained. Providing more information on these aspects would contribute to the transparency and replicability of the study.
\end{itquote}
We agree that the previous draft was unclear in the description of the outcome-generating model (now equation \ref{eq:y-sim}). We have added material to section \ref{sec:dataGeneration} and Table \ref{tab:factor} that includes the necessary details. Should anything still be unclear, replication code is available in the anonymized GitHub repository at \url{https://anonymous.4open.science/r/psGee-B498/}.

\item \begin{itquote}     Page 17: The rationale for selecting specific values for manipulated factors, such as choosing a sample size range of 100 to 1000 and using N(0,1/2) instead of N(0,1), is not adequately justified. Including evidence and citations to support these choices would strengthen the methodological foundation of the study.

\end{itquote}
We added rationale, including citations, to the paragraph beginning ``Potential and observed outcomes...'' on page \pageref{potential} and to section \ref{sec:manipulatedFactors}.

\item \begin{itquote}     Page 18: A section detailing the evaluation metrics and criteria used to assess simulation results should precede the presentation of results. This would provide readers with a clear understanding of the benchmarks against which the findings are measured.

\end{itquote}
We now begin Section \ref{sec:simResults} with definitions of the evaluation metrics, which we hope make the results clearer.

\item \begin{itquote}     Pages 18-20: It is highly recommended that the authors employ a series of ANOVAs to analyze the impact of various factors on the accuracy of effect estimation to guide results interpretation. Additionally, a concise paragraph summarizing key findings before the application example would enhance the clarity and interpretation of the results.
\end{itquote}
Based on your comment, we strongly considered including \textsc{anova} results to gauge the factors' relative importance. However, we decided that such an analysis would not be appropriate in this context. First, the factors were not fully crossed, significantly complicating the analysis. Second, the results would be heavily driven by the factors' variances---specifically, the set of values we chose for $n$, $\alpha$, $\bm\beta$, and $\bm\gamma$; although our choices were well-motivated, they were also, necessarily, somewhat arbitrary. 
Lastly, as Figure \ref{fig:alphan} shows, the effects of $n$ and $\alpha$ on bias and standard error are non-linear, further complicating the \textsc{anova} modeling. For those reasons, we decided our current approach was clearer and simpler.

Your suggestion of a paragraph summarizing the results was quite helpful. We included such a paragraph, on page \pageref{simsum} beginning ``In sum.''

\item \begin{itquote}      APA Citations and References: The paper does not adhere to the APA 7th style. Authors are advised to consult the APA Formatting and Style Guide for proper citation and reference formatting throughout the manuscript. Consistency in style is crucial for scholarly publications.
\end{itquote}
Thank you. We have fixed the formatting.
\end{enumerate}

\section*{Reviewer 2}
Comments to the Author
Great work. Well written. Please find some of my comments below. I hope they are useful.

\begin{enumerate}

\item \begin{itquote}   Please provide more background on M-estimators especially to the extent it evolves to become relevant to the GEEPERs. 
\end{itquote}
Thanks for this suggestion---in response, we added Section \ref{sec:mest}
\item \begin{itquote}      I am wondering if doubly robust PSW estimator would reduce the remaining bias difference between GEEPERs and PSW, and Mixture and PSW.
\end{itquote}
This is an interesting suggestion---you're right that the \psw estimator did not include an outcome model, unlike \geepers and \nmm. However, we doubt that an outcome model would substantially reduce the bias of \psw in the simulation. The advantage of doubly-robust estimation is that an outcome model can correct for model misspecification in the propensity (or principal) score model. However, the principal score model in our simulation study was well-specified---the bias of the \psw estimator was due to an unobserved confounder, $x_3$. Unfortunately, like \psw, doubly-robust estimators rely on ignorability, i.e., that all relevant confounders have been measured and included. 
\item \begin{itquote}      GEEPERs seems great, but also, surprised by how efficient Mixture can be. Are there any mechanisms that link M-estimators to Mixture modeling? Just wondering.
\end{itquote}
We were also surprised by how well the \nmm performed in the simulation, especially given prior literature detailing its vulnerability \citep[][; cited in the paper]{griffin2008application,feller2016principal}. Its similarity to \geepers in some scenarios is less surprising, since they are both essentially mixture model estimators.

\item \begin{itquote}    Would it be possible to offer some evaluation metric for GEEPERs, which would indicate, under some conditions researchers could possibly opt or seek other estimators?
\end{itquote}
This is a great suggestion. 
\item \begin{itquote}     Quickly checked Frangakis and Rubin (2002). Do you think the effect of bottom-outers (or similar experimental designs) could be partitioned into associative and dissociative effects with GEEPERs?
\end{itquote}
\item \begin{itquote}    It would be great if you could share the code for estimation and simulation in the next round of review (if there is). Also make sure you have comments in the code so someone reading the article can match the narrative with the code.
\end{itquote}
\end{enumerate}

\section{Reviewer 3}

Comments to the Author

This paper develops a new estimator for principal stratification based on M-estimation and stacked equations. The proposed method is designed to be more robust and simpler to use than standard methods. The method is, however, tailored to a very specific set up: one-way noncompliance. In general, principal stratification is a much more general framework that can be used for a wide variety of settings where there is conditioning on some post-treatment quantity.

The method is outlined, and then evaluated through a series of simulations, and then an application is provided. My primary criticism is that the new method doesn't appear to outperform extant methods to a compelling degree.  I take up this point in more detail first for the simulations and second for the empirical application.

Simulations.

In my view, the simulations should show the following:

--When the assumptions of extant methods are met, GEEPERS should do no worse.
--When the assumptions of extend methods are not met, GEEPERS should do better.

Given this criteria, it seems to me that the simulation should focus on the last three factors in the design.  That is, I would hold the sample size and alpha fixed.  Currently too many factors are varied at the same time to give a clear picture.  I would alter Table 1, to indicate which factors yield assumption violations and then violate these assumptions in isolation instead of relaxing them simultaneously.  As it currently stands, it is quite hard to understand the relative performance of GEEPERs compared to extant methods. Careful relaxation of each assumption allows the reader to see how much bias results from a single factor.

Application.

I thought there are a number of less than desirable parts to the application. First, the focus on bottoming out is less than ideal. This strikes me as a fairly rare phenomenon. I didn't find the concept very intuitive and had to read just the descriptive section several times to understand why one might be interested in a principal stratification analysis here.  This makes the methods seem not particularly general.  Why not focus on an application with straightforward oneway noncompliance? This is an area where principal stratification is often applied, and the concepts are well understood.

In addition, it isn't ideal that the ATE in each condition is not estimated very precisely, that is the CI appears to cover zero in all cases. Next, all the principal stratification estimates are all pretty close to the same value as the ATE. That serves as another reason this application isn't compelling. It is hard to care about an application where principal stratification doesn't provide much in the way of insights relative to the ATE. Next, the performance of GEEPERs seems to be quite similar to extant methods in each scenario. Overall, to my thinking, there is little to recommend the current application. Again, there is little here that demonstrates the value of GEEPERs.

To my mind, these problems could both be fixed.  That is, the simulations could be redesigned to provide more compelling evidence. Next, the authors may be able to identify are better application.  Therefore, the manuscript should be revised and resubmitted.

\end{document}

### 2025-06-20 writeUps/paper.tex
work on responses

@@
 
-\RequirePackage{amsmath,amsfonts,amssymb}
+\RequirePackage{amsmath,amsfonts,amssymb,amsthm}
 %\RequirePackage[authoryear]{natbib}
@@
 \pmm is highly vulnerable to violations of normality and to interactions between $X$ and $Z$, while \psw is robust to non-normality and robust to $X$-$Z$ interactions when $\alpha = 0.5$.
+These results suggest that researchers who are willing to assume principal ignorability should use \psw, and those who are confident in their parametric model for potential outcomes should use a parametric mixture model.
+Researchers who are hesitant to accept principal ignorability or a parametric outcome model should consider \geepers.
+
+The simulation study suggests that none of the three methods evaluated, \geepers, \pmm, or \psw, dominates the others, and offers some guidance on when each estimator may be preferable.
+\psw is the most precise and requires no assumptions about the distribution of outcomes, but can be quite sensitive to violations of principal ignorability.
+When the determinants of $\st$ are well-understood and measured, \psw seems to be the best option.
+When principal ignorability does not hold, potential outcomes are conditionally normal, and there are no $X:Z$ interactions within principal strata, the \pmm estimator outperforms \geepers, though the difference decreases with $n$ or $\alpha$.
 
@@
 Lastly, can over-fit principal score models cause bias or other estimation problems, and if so, are there ways to protect against overfitting?
-
-The simulation study suggests that none of the three methods evaluated, \geepers, \nmm, or \psw, dominates the others, and offers some guidance on when each estimator may be preferable.
-\psw is the most precise and requires no assumptions about the distribution of outcomes, but can be quite sensitive to violations of principal ignorability.
-When the determinants of $\st$ are well-understood and measured, \psw seems to be the best option.
-When principal ignorability does not hold, potential outcomes are conditionally normal, and there are no $X:Z$ interactions within principal strata, the \nmm estimator outperforms \geepers, though the difference decreases with $n$ or $\alpha$.
 
@@
 \subsection{Full Empirical 95\% Interval Coverage Results}
-The following table gives the empirical coverage of nominal 95\% intervals for \textsc{geepers} and mixture model principal effect estimates under varying data generating models.\\
-
-\input{coverageTabAppendix.tex}
+The following tables give the empirical coverage of nominal 95\% intervals for \textsc{geepers} and mixture model principal effect estimates under varying data generating models.
+First we show results when $n=500$ per condition, and then when $n=1000$ per condition.
+\\
+
+\begin{table}
+  \caption{Empirical coverage of nominal 95\% Confidence intervals for \geepers and \pmm when $n=500$ per condition.}
+  \input{coverageTabAppendix500.tex}
+ \end{table}
+
+ \begin{table}
+  \caption{Empirical coverage of nominal 95\% Confidence intervals for \geepers and \pmm when $n=1000$ per condition.}
+  \input{coverageTabAppendix1000.tex}
+ \end{table}
 
@@
 \subsection{Full RMSE Results}
-The following table gives the root mean squared error (RMSE) for \textsc{geepers}, mixture model, and principal score weighting principal effect estimates under varying data generating models.\\
-
-\input{rmseTabAppendix}
+The following table gives the root mean squared error (RMSE), $\left\{\sum_b (\hat{\tau}-\tau)^2/500\right\}^{1/2}$, for \textsc{geepers}, mixture model, and principal score weighting principal effect estimates under varying data generating models.\\
+
+\begin{table}
+  \caption{Empirical RMSE  for \geepers, \pmm, and \psw when $n=500$ per condition.}
+  \input{rmseTabAppendix500}
+ \end{table}
+
+ \begin{table}
+  \caption{Empirical RMSE  for \geepers, \pmm, and \psw when $n=1000$ per condition.}
+  \input{rmseTabAppendix1000}
+ \end{table}
 

### 2025-06-20 code/simulation/displaySim.r

@@ -533,33 +533,65 @@ rmse%>%filter(n==500,eff==1,errDist!='mix',mu01==0.3,b1>0)%>%
   geom_point()+geom_line()+#geom_hline(yintercept=0.95)+
   facet_grid(interactionS~errDist+interactionZ)
 
-sink('writeUps/rmseTabAppendix.tex')
+sink('writeUps/rmseTabAppendix500.tex')
 cbind(
-
 rmse%>%filter(errDist!='mix',n==500,eff!="Diff",b1==.2)%>%
   transmute(`Residual\nDist.`=c(norm='Normal',unif='Uniform')[errDist],
          `X:Z\nInt.?`=ifelse(intZ,'Yes','No'),
          `X:S\nInt.?`=ifelse(intS,'Yes','No'),
-         estimator=ifelse(estimator=='Mixture','Mix.',estimator),
-
+         estimator=c(Mixture="\\pmm",GEEPERs="\\geepers",PSW="\\psw")[estimator],
          `$\\beta_1$`=mu01,
          `Prin.\nEff`=paste0('$\\tau^',eff,'$'),
          rmse)%>%
 pivot_wider(names_from=estimator,values_from=rmse),
-rmse%>%filter(errDist!='mix',mu01==0.3,eff!="Diff",b1==.5)%>%
+rmse%>%filter(errDist!='mix',n==500,mu01==0.3,eff!="Diff",b1==.5)%>%
   transmute(`Res.\nDist.`=c(norm='Norm.',unif='Unif.')[errDist],
          `X:Z\nInt.?`=ifelse(intZ,'Yes','No'),
          `X:S\nInt.?`=ifelse(intS,'Yes','No'),
-         estimator=ifelse(estimator=='Mixture','Mix.',estimator),
+         estimator=c(Mixture="\\pmm",GEEPERs="\\geepers",PSW="\\psw")[estimator],
          n,
          `$\\beta_1$`=mu01,
          `Prin.\nEff`=paste0('$\\tau^',eff,'$'),
-         rmse)%>%
+         rmse,
+         xx=rep(1:(n()/3),each=3))%>%
+select(xx,estimator,rmse)%>%
 pivot_wider(names_from=estimator,values_from=rmse)%>%
-select(GEEPERs,Mix.,PSW))%>%
-  kbl('latex',booktabs=TRUE,col.names=linebreak(names(.)),escape=FALSE,digits=2)%>%
[…]

### 2025-06-20 code/simulation/readSimFuncs.r

@@ -87,7 +87,7 @@ bayesProc <- function(res1,facs){
 
 
 loadRes <- function(ext1='',ext2='',pswResults){
-    load(paste0('simResults',ext1,'/cases',ext2,'.RData'))
+    load(paste0('simResults/cases',ext1,ext2,'.RData'))
 
     results <- list()
 #    summ <- NULL
@@ -95,7 +95,7 @@ loadRes <- function(ext1='',ext2='',pswResults){
     for(i in 1:nrow(cases)){#length(fn)){
         #if(i %% 10==0)
             cat(round(i/nrow(cases)*100),'%',sep='')#length(fn)*100), '% ')
-      load(paste0('simResults',ext1,'/sim',i,ext2,'.RData'))#fn[i]))
+      load(paste0('simResults/sim',i,ext1,ext2,'.RData'))#fn[i]))
       if(!missing(pswResults))
         res <- map(1:length(res),
                    ~append(res[[.]],list(psw=unlist(pswResults[[i]][.,]))))

### 2025-06-23 writeUps/paper.tex
more responses

@@
 
-\subsubsection{Paremetric Mixture Modeling}
+\subsubsection{Parametric Mixture Modeling}
 %The normal mixture modeling approach to estimating $\muc0$ and $\muc1$ \citep{imbens1997bayesian} assumes that, conditional on $\st$, $Y_C$ is normally distributed.
@@
 Like \geepers, \pmm is insensitive to interactions between $X$ and $S$.
-However, interactions between $X$ and $Z$ in the data generating model skew the \pmm sampling distribution and reduce the coverage of its 95\% confidence intervals, even when $\alpha=0.5$.
+However, interactions between $X$ and $Z$ in the data-generating model skew the \pmm sampling distribution and reduce the coverage of its 95\% confidence intervals, even when $\alpha=0.5$.
 When the residual distribution was uniformly distributed, the \pmm sampling distribution appeared bimodal, and its credible intervals exhibited severe undercoverage.
@@
 
-The simulation study suggests that none of the three methods evaluated, \geepers, \pmm, or \psw, dominates the others, and offers some guidance on when each estimator may be preferable.
+\label{evalMetric}The simulation study suggests that none of the three methods evaluated---\geepers, \pmm, or \psw---dominates the others, and offers some guidance on when each estimator may be preferable.
 \psw is the most precise and requires no assumptions about the distribution of outcomes, but can be quite sensitive to violations of principal ignorability.
@@
 When principal ignorability does not hold, potential outcomes are conditionally normal, and there are no $X:Z$ interactions within principal strata, the \pmm estimator outperforms \geepers, though the difference decreases with $n$ or $\alpha$.
+\geepers may be the best choice when conditional normality may not hold, or if there are $X:Z$ interactions and $\bm{x}$ is sufficiently predictive of $\st$ (AUC $\ge$0.675, corresponding to $\alpha \ge 0.5$, may serve as a threshold).
+These suggestions may depend heavily on the simulation setup here; applying them in general scenarios requires more research.
 

### 2025-06-23 writeUps/responses.tex
more responses

@@
 \end{itquote}
-This is a great suggestion. 
+This is a great suggestion. We added the paragraph beginning ``The simulation study suggests\dots'' on page \pageref{evalMetric} to give some guidance along those lines. 
 \item \begin{itquote}     Quickly checked Frangakis and Rubin (2002).
@@
 \end{itquote}
+This is correct, but only in a narrow technical sense. 
+Under our notation, Frangakis and Rubin (2002) define dissociative effects as contrasts between $\yti$ and $Y_{Ci}$ for the subjects $i$ with $S_{Ci}=\sti$; associative effects are contrasts for subjects with $S_{Ci}\ne\sti$.
+A contrast between associative and dissociative effects ``measures the degree to which a causal effect of treatment on outcome occurs together with a causal effect of treatment on the surrogate."  Since our paper deals only with the one-way noncompliance case, we set $S_{Ci}=0$ for all $i$ (Assumption 1).
+In Section \ref{sec:fh2t}, we define $\st=0$ for non-bottom-outers and $\st=1$ for bottom-outers.
+Therefore, as you wrote, an effect for non-bottom-outers would be dissociative, while an effect for bottom-outers would be associative.
+However, this is, at least in part, an artifact of our notation.
+It's true that anyone assigned to control would, by necessity, request fewer than 11 bottom-out hints (the median in the treatment group)---$S_C=0$---and that treatment assignment affects bottom-out-hint usage less for non-bottom-outers than for bottom-outers.
+However, many non-bottom-outers assigned to treatment will request \emph{some} bottom-out hints, whereas students assigned to control request none---our choice to dichotomize hint requests masks variation that does not exist in the control group.
+Furthermore, our analysis would have been identical had we set $S_{Ci}$ equal to any other constant or set it to NA.
+
 \item \begin{itquote}    It would be great if you could share the code for estimation and simulation in the next round of review (if there is).
@@
 \end{itquote}
+Full replication code is available at \url{FILL IN} 
 \end{enumerate}
@@
 
-Comments to the Author
+\subsection{Simulations}
+\begin{itquote}
+    In my view, the simulations should show the following:
+\begin{itemize}
+   \item When the assumptions of extant methods are met, GEEPERS should do no worse.
+   \item When the assumptions of extent methods are not met, GEEPERS should do better.
+\end{itemize}
+\end{itquote}
+Depending on the meaning of ``do no worse,'' \geepers may or may not meet those criteria. 
+In the scenarios we tested, \geepers is always less biased and its nominal confidence intervals achieve higher empirical coverage than the competitors we tested.
+However, the competing methods often boast smaller standard errors than \geepers. 
+This set of results is typical for semi- or non-parametric methods, which tend to be less efficient than their parametric counterparts when the latter's assumptions are met, but are, say, less biased when those assumptions are violated.
 
-This paper develops a new estimator for principal stratification based on M-estimation and stacked equations.
-The proposed method is designed to be more robust and simpler to use than standard methods.
-The method is, however, tailored to a very specific set up: one-way noncompliance.
-In general, principal stratification is a much more general framework that can be used for a wide variety of settings where there is conditioning on some post-treatment quantity.
+Unfortunately, \geepers may not even meet that criterion relative to \psw.
+Our simulation did not include scenarios in which principal ignorability held, but we expect that in such scenarios---depending on other factors---\psw may outperform \geepers in every respect.
+That said, principal ignorability is untestable and, in our opinion, rather strong. 
+As we stated on page \pageref{evalMetric} (in the paragraph beginning ``The simulation study suggests''), researchers will have to decide in each application which assumptions are the most plausible, and choose estimators based on that decision.
 
-The method is outlined, and then evaluated through a series of simulations, and then an application is provided.
-My primary criticism is that the new method doesn't appear to outperform extant methods to a compelling degree.
-I take up this point in more detail first for the simulations and second for the empirical application.
-
-Simulations.
-
-In my view, the simulations should show the following:
-
---When the assumptions of extant methods are met, GEEPERS should do no worse.
---When the assumptions of extend methods are not met, GEEPERS should do better.
-
-Given this criteria, it seems to me that the simulation should focus on the last three factors in the design.
+\begin{itquote} Given this criterion, it seems to me that the simulation should focus on the last three factors in the design.
 That is, I would hold the sample size and alpha fixed.
-Currently too many factors are varied at the same time to give a clear picture.
-I would alter Table 1, to indicate which factors yield assumption violations and then violate these assumptions in isolation instead of relaxing them simultaneously.
+Currently, too many factors are varied at the same time to give a clear picture.
+I would alter Table 1 to indicate which factors yield assumption violations and then violate these assumptions in isolation instead of relaxing them simultaneously.
 As it currently stands, it is quite hard to understand the relative performance of GEEPERs compared to extant methods.
 Careful relaxation of each assumption allows the reader to see how much bias results from a single factor.
+\end{itquote}
+The results in Figure \ref{fig:boxplots} and in Table \ref{tab:coverage} now hold $n$ fixed at $n=500$ per group and completely cross the other factors, i.e.
+$\alpha=0.2$ or $0.5$, interactions between $\bm{x}$ and $\st$ or $Z$, and the distribution of the residuals.
+Now it is apparent from Figure \ref{fig:boxplots} that, for instance, none of the methods are sensitive to interactions between $\bm{x}$ and $\st$, and that \pmm, but not the others, is sensitive to the distribution of the residuals.
+
+In Table \ref{tab:coverage}, results when one or more assumptions have been violated are highlighted in red.
 
@@
 
-I thought there are a number of less than desirable parts to the application.
+I thought there were a number of less-than-desirable parts to the application.
 First, the focus on bottoming out is less than ideal.
@@
 This makes the methods seem not particularly general.
-Why not focus on an application with straightforward oneway noncompliance?
+Why not focus on an application with straightforward one-way noncompliance?
 This is an area where principal stratification is often applied, and the concepts are well understood.
 
-In addition, it isn't ideal that the ATE in each condition is not estimated very precisely, that is the CI appears to cover zero in all cases.
-Next, all the principal stratification estimates are all pretty close to the same value as the ATE.
+In addition, it isn't ideal that the ATE in each condition is not estimated very precisely; that is, the CI appears to cover zero in all cases.
+Next, all the principal stratification estimates are pretty close to the same value as the ATE.
 That serves as another reason this application isn't compelling.
@@
 That is, the simulations could be redesigned to provide more compelling evidence.
-Next, the authors may be able to identify are better application.
+Next, the authors may be able to identify better applications.
 Therefore, the manuscript should be revised and resubmitted.
 
-\end{document}
+\end{document}}

### 2025-06-23 code/simulation/displaySim.r
fix notation in figures, appendix tables with n=1000

@@ -101,9 +101,9 @@ tab1 <- bind_rows(
   select(measure,everything())
 
 
-  #%>%
-  kbl()%>%
-  collapse_rows(columns=1,latex_hline="major",valign="middle")
+  ## #%>%
+  ## kbl()%>%
+  ## collapse_rows(columns=1,latex_hline="major",valign="middle")
 
 
 bp(pdns,subset=,facet=PE~N,title="Estimation Error by N",ylim=c(-1,1))+
@@ -115,7 +115,7 @@ ggsave('simFigs/byN.jpg',width=6,height=3)
 ### results across B1
 ###############################################################################
 
- pdb1 <- reausultsB1s %>%
+ pdb1 <- resultsB1s %>%
   group_by(b1)%>%
   mutate(AUCf=mean(auc,na.rm=TRUE))%>%
   ungroup()%>%
@@ -222,8 +222,8 @@ pd <- results %>%
     estimator=c(bayes='Mixture',mest='GEEPERs',psw='PSW')[estimator],
     interactionZ=ifelse(intZ,"Z\ninteraction","No Z\ninteraction"),
     interactionS=ifelse(intS,"S\ninteraction","No S\ninteraction"),
-    intAll=ifelse(intZ,ifelse(intS,"X:S & X:Z","X:Z"),ifelse(intS,"X:S","No inter.")),
-    intAll=factor(intAll,levels=c("No inter.","X:Z", "X:S","X:S & X:Z"))
+    intAll=ifelse(intZ,ifelse(intS,"$\\bm{x}:\\st$ & $\\bm{x}:Z$","$\\bm{x}:Z$"),ifelse(intS,"$\\bm{x}:\\st$","No inter.")),
+    intAll=factor(intAll,levels=c("No inter.","$\\bm{x}:Z$", "$\\bm{x}:\\st$","$\\bm{x}:\\st$ & $\\bm{x}:Z$"))
   ) %>%
   filter(estimator=='PSW'|rhat<1.1)
 
@@ -503,8 +503,8 @@ sink('writeUps/rmseTab.tex')
 cbind(
 rmse%>%filter(n==500,eff==1,errDist!='mix',mu01==0.3,b1==0)%>%
   transmute(`Residual\nDist.`=c(norm='Normal',unif='Uniform')[errDist],
-         `X:Z\nInt.?`=ifelse(intZ,'Yes','No'),
[…]

### 2025-06-23 code/simulation/displaySimPublish.r

@@ -1,0 +1,480 @@
+library(tidyverse)
+library(broom)
+library(kableExtra)
+library(gridExtra)
+library(ggpubr)
+
+
+options(
+tikzLatexPackages = c(
+    "\\usepackage{tikz}",
+    "\\usepackage{bm}",
+    "\\usepackage{amsmath}",
+"\\usepackage[active,tightpage]{preview}",
+"\\PreviewEnvironment{pgfpicture}",
+"\\setlength\\PreviewBorder{0pt}"
+),
+tikzXelatexPackages = c(
+"\\usepackage{tikz}\n",
+"\\usepackage[active,tightpage,xetex]{preview}\n",
+"\\usepackage{fontspec,xunicode}\n",
+"\\PreviewEnvironment{pgfpicture}\n",
+"\\setlength\\PreviewBorder{0pt}\n"
+),
+tikzLualatexPackages = c(
+    "\\usepackage{tikz}\n",
+    "\\usepackage{bm}",
+        "\\usepackage{amsmath}",
+"\\usepackage[active,tightpage,psfixbb]{preview}\n",
+"\\usepackage{fontspec,xunicode}\n",
+"\\PreviewEnvironment{pgfpicture}\n",
+"\\setlength\\PreviewBorder{0pt}\n"
+)
+)
+
+source('code/simulation/readSimFuncs.r')
+
+#### after simulation results have been loaded and pre-processed...
+## source('code/simulation/readSim.r')
+load('simResults/fullResults.RData')
+load('simResults/resultsNs.RData')
+load('simResults/resultsB1s.RData')
+
+
+###############################################################################
+### results across ns
+###############################################################################
+ pdns <- resultsNs %>%
+  group_by(b1)%>%
+  mutate(AUCf=mean(auc,na.rm=TRUE))%>%
+  ungroup()%>%
+  mutate(
+    errP = est - pop,
+    B1 = factor(b1,levels=c(0,0.2,0.5),
+                labels=c(bquote(alpha==0),bquote(alpha==0.2),bquote(alpha==0.5))),
+    AUCff=paste0('AUC=',round(AUCf,1)),
+    N=paste0('n=',n),
+    M1=factor(mu01,levels=c("0","0.3"),
[…]

### 2025-06-23 code/simulation/displaySimPublish.r
fix coverage tab

@@ -259,9 +259,9 @@ tikz("simFigs/boxplotsNew.tex",#pdf",
 grid.arrange(norm,unif,nrow=1)
 dev.off()
 
-system("cd simFigs")
+setwd("simFigs")
 system("lualatex boxplotsNew.tex")
-
+setwd("..")
 
 #######################################################
 #### rmse appendix
@@ -374,7 +374,8 @@ cbind(
     kbl('latex',booktabs=TRUE,col.names=linebreak(names(.)),escape=FALSE,digits=2)%>%
     add_header_above(c(" " = 3, #"$\\\\alpha=0$" = 2,
                        "$\\\\alpha=0.2$" = 2, "$\\\\alpha=0.5$" = 2),escape=FALSE)%>%
-  collapse_rows(columns=1,latex_hline="major",valign="middle")
+    collapse_rows(columns=1,latex_hline="major",valign="middle")%>%
+    footnote(general="\\\\footnotesize Based on 500 replications. $n=500$, $\\\\beta_1=0.3$. Simulation standard error $\\\\approx 1$ percentage point. Estimates colored \\\\rd{red} indicate cases where the assumptions of the model are not met.",escape=FALSE,footnote_as_chunk = TRUE)
 sink()
 
 ########################

### 2025-06-23 code/simulation/displaySimPublish.r
fix coverage tab pt2

@@ -375,7 +375,7 @@ cbind(
     add_header_above(c(" " = 3, #"$\\\\alpha=0$" = 2,
                        "$\\\\alpha=0.2$" = 2, "$\\\\alpha=0.5$" = 2),escape=FALSE)%>%
     collapse_rows(columns=1,latex_hline="major",valign="middle")%>%
-    footnote(general="\\\\footnotesize Based on 500 replications. $n=500$, $\\\\beta_1=0.3$. Simulation standard error $\\\\approx 1$ percentage point. Estimates colored \\\\rd{red} indicate cases where the assumptions of the model are not met.",escape=FALSE,footnote_as_chunk = TRUE)
+    footnote(general=c("\\\\footnotesize Based on 500 replications.","$n=500$, $\\\\beta_1=0.3$.", "Simulation standard error $\\\\approx 1$ percentage point.","Estimates colored \\\\rd{red} indicate cases where the assumptions of the model are not met."),escape=FALSE,footnote_as_chunk = TRUE,threeparttable=TRUE)
 sink()
 
 ########################

### 2025-06-24 writeUps/paper.tex
finished first draft of responses

@@
 %\usepackage[notablist,tablesfirst]{endfloat}% To `separate' figures and tables from text if required
-
+\usepackage{threeparttable}
 \usepackage{epstopdf}% To incorporate .eps illustrations using PDFLaTeX, etc.
@@
 
-Using each simulated dataset, we estimated $\eff0$ and $\eff1$ with three methods: a mixture model \eqref{eq:mixtureModelPS} including principal scores, \textsc{geepers}, and the principal score weighting estimator \eqref{eq:psw}.
+Using each simulated dataset, we estimated $\eff0$ and $\eff1$ with three methods: \geepers, a parametric mixture model \eqref{eq:mixtureModelPS} including principal scores---hereafter \pmm---and the principal score weighting estimator \eqref{eq:psw}, hereafter \psw.
 
@@
 
-The mixture estimator then fits two outcome models:
+\pmm, the parametric mixture estimator, then fits two outcome models:
 For the treatment group, the regression model $Y=\beta_0^T+\beta_1^TS+\bm{\gamma}'\bx+\epsilon$ and for the control group, a mixture model of the form \eqref{eq:mixtureModelPS}, but with $\muc0$ and $\muc1$ replaced by regression functions $\beta_0^C+\bm{\gamma}'\bx$ and $\beta_0^C+\beta_1^C+\bm{\gamma}'\bx$, respectively.
@@
 
-For \textsc{geepers} and principal score weighting, the principal score model was fit by itself, to data from the treatment group, using standard maximum likelihood methods.
+For \textsc{geepers} and \psw, the principal score model was fit by itself, to data from the treatment group, using standard maximum likelihood methods.
 For \textsc{geepers}, we used the estimated principal scores in the control group, along with the observed $S$ in the treatment group, to construct $R$, and then fit the model \eqref{eq:reg2} with OLS.
@@
 
-For principal score weighting we plugged estimated principal scores into \eqref{eq:psw} to estimate $\muc0$ and $\muc1$, estimated $\mut0$ and $\mut1$ with the sample means of subjects with $Z=1$ and $S=0$ or $S=1$, respectively, and estimated $\heff0=\hat{\mu}_{Tw}^0-\hat{\mu}_{Cw}^0$ and $\heff1=\hat{\mu}_{Tw}^1-\hat{\mu}_{Cw}^1$.
+For \psw we plugged estimated principal scores into \eqref{eq:psw} to estimate $\muc0$ and $\muc1$, estimated $\mut0$ and $\mut1$ with the sample means of subjects with $Z=1$ and $S=0$ or $S=1$, respectively, and estimated $\heff0=\hat{\mu}_{Tw}^0-\hat{\mu}_{Cw}^0$ and $\heff1=\hat{\mu}_{Tw}^1-\hat{\mu}_{Cw}^1$.
 
@@
 We also report the empirical coverage of confidence intervals, the proportion of replications in which $\eff1$ is within the estimated confidence bounds of $\heff1$.
-Here, we present results on error, bias, empirical standard error, and confidence itnterval coverage; results on root-mean-squared-error %$\left(\widehat{bias}^2+SE_{emp}^2\right)^{1/2}$
+Here, we present results on error, bias, empirical standard error, and confidence interval coverage; results on root-mean-squared-error %$\left(\widehat{bias}^2+SE_{emp}^2\right)^{1/2}$
 $\left(\sum_b err_b^2/500\right)^{1/2}$
@@
 \subsubsection{Varying $n$ and $\alpha$}
-When $\alpha=0$, Assumption \ref{ass:vps} was violated, and \textsc{geepers} estimates were highly erratic; however, as Table \ref{tab:coverage} shows, when other assumptions held, inference remained valid.
+% When $\alpha=0$, Assumption \ref{ass:vps} was violated, and \textsc{geepers} estimates were highly erratic; however, as Table \ref{tab:coverage} shows, when other assumptions held, inference remained valid.
+Our first set of results illustrates the performance of \geepers when all of its assumptions are met.
+Since \geepers is an asymptotic method whose identification depends on $\bm{x}$ as a predictor of $\st$, we examine its bias and standard error as functions of $n$ and $\alpha$.
+We include analogous results for \pmm as a baseline comparison.
 
@@
 
-\subsubsection{Other Factors}
+\subsubsection{Performance when Assumptions do not Hold}
+
+Our next set of results compares the performance of \geepers, \pmm, and \psw when one or more assumptions of each method may be violated.
 
@@
 %<<<<<<< HEAD
-Figure \ref{fig:boxplots} shows violin plots, layered on top of jittered scatter plots, showing estimation error for \textsc{geepers}, the parametric mixture model, and the \textsc{psw} estimator of $\eff1$ for, and Table \ref{tab:coverage} shows empirical coverage of nominal 95\% confidence interval estimates for \textsc{geepers} and a parametric mixture models; when at least one of an estimator's assumptions is violated, the result is colored red.
-
-The first row of Table \ref{tab:coverage} and the first column of Figure \ref{fig:boxplots} show similar results as Figure \ref{fig:alphan}---when all assumptions of both methods are met, \textsc{geepers} is slightly less biased than the parametric mixture model; when $\alpha=0.2$, \textsc{geepers} is much more variable and $\alpha=0.5$ the difference in variance is slight.
+Figure \ref{fig:boxplots} shows violin plots, layered on top of jittered scatter plots, showing estimation error for \geepers, the parametric mixture model \pmm, and the \psw estimator of $\eff1$.
+Table \ref{tab:coverage} shows empirical coverage of nominal 95\% confidence interval estimates for \geepers and \pmm; when at least one of an estimator's assumptions is violated, the result is colored red.
+
+The first column of Figure \ref{fig:boxplots} and the black entries of Table \ref{tab:coverage} show results from scenarios in which all assumptions of \geepers and \pmm are met (\psw relies on principal ignorability, which is violated throughout the simulation study).
+\textsc{geepers} is slightly less biased than the parametric mixture model; when $\alpha=0.2$, \textsc{geepers} is much more variable and $\alpha=0.5$ the difference in variance is slight.
 \textsc{psw} has by far the lowest variance and the highest bias; the latter is due to the violation of principal ignorability. 
-
-The remainder of the results reflect violations modeling assumptions for either \geepers or the parametric mixture model or both: interactions in the data generating model between covariates $X$ and treatment assignment $Z$ and/or principal stratum $S$, either of which would violate assumption \ref{ass:rci} and the linear additive model of the \textsc{pmm}, and/or residuals drawn from a uniform distribution, which violates the normality assumption of \textsc{pmm}.
+When all assumptions are met, confidence intervals from both \geepers and \pmm achieve their nominal levels.
+
+The remainder of the results reflect violations modeling assumptions for either \geepers or the parametric mixture model or both: when $\alpha=0$, violating Assumption \ref{ass:vps}; interactions in the data generating model between covariates $X$ and treatment assignment $Z$ and/or principal stratum $S$, either of which would violate assumption \ref{ass:rci} and the linear additive model of the \textsc{pmm}; and/or residuals drawn from a uniform distribution, which violates the normality assumption of \textsc{pmm}.
+
+We omitted the $\alpha=0$ case from Figures \ref{fig:alphan} and \ref{fig:boxplots} since the variance of \geepers and \pmm estimates was so large as to dominate the plot. 
+However, the first column of results in Table \ref{tab:coverage} shows that, when other assumptions are met, CIs from both methods tend to over-cover when $\alpha=0$. 
+In a randomized experiment, $\alpha$ can be estimated directly, making Assumption \ref{ass:vps} testable.
+
 For \geepers, when $\alpha=0.2$ the presence of interactions in the data-generating model increases sampling variance, but only interactions between $X$ and $Z$ reduce confidence interval coverage.
@@
 \centering
-\begin{tabular}[t]{lllllll}
-\toprule
-\multicolumn{3}{c}{ } & \multicolumn{2}{c}{$\alpha=0.2$} & \multicolumn{2}{c}{$\alpha=0.5$} \\
-\cmidrule(l{3pt}r{3pt}){4-5} \cmidrule(l{3pt}r{3pt}){6-7}
-\makecell[l]{Residual\\Dist.} & \makecell[l]{X:Z\\Int.?} & \makecell[l]{X:S\\Int.?} & \geepers & \pmm & \geepers & \pmm\\
-\midrule
- & No & No & 0.97 & 0.97 & 0.96 & 0.96\\
-
- & Yes & No & \rd{0.88} & \rd{0.74} & \rd{0.94} & \rd{0.89}\\
-
- & No & Yes & \rd{0.98} & \rd{0.97} & \rd{0.97} & \rd{0.96}\\
-
-\multirow{-4}{*}{\raggedright\arraybackslash Normal} & Yes & Yes & \rd{0.9} & \rd{0.77} & \rd{0.94} & \rd{0.88}\\
-\cmidrule{1-7}
- & No & No & 0.97 & \rd{0.4} & 0.95 & \rd{0.61}\\
-
- & Yes & No & \rd{0.87} & \rd{0.2} & \rd{0.93} & \rd{0.48}\\
-
- & No & Yes & \rd{0.98} & \rd{0.53} & \rd{0.96} & \rd{0.57}\\
-
-\multirow{-4}{*}{\raggedright\arraybackslash Uniform} & Yes & Yes & \rd{0.88} & \rd{0.21} & \rd{0.94} & \rd{0.53}\\
-\bottomrule
-\multicolumn{7}{p{3.5in}}{\footnotesize Based on 500 replications.
+% \begin{tabular}[t]{lllllll}
+% \toprule
+% \multicolumn{3}{c}{ } & \multicolumn{2}{c}{$\alpha=0.2$} & \multicolumn{2}{c}{$\alpha=0.5$} \\
+% \cmidrule(l{3pt}r{3pt}){4-5} \cmidrule(l{3pt}r{3pt}){6-7}
+% \makecell[l]{Residual\\Dist.} & \makecell[l]{X:Z\\Int.?} & \makecell[l]{X:S\\Int.?} & \geepers & \pmm & \geepers & \pmm\\
+% \midrule
+%  & No & No & 0.97 & 0.97 & 0.96 & 0.96\\
+
+%  & Yes & No & \rd{0.88} & \rd{0.74} & \rd{0.94} & \rd{0.89}\\
+
+%  & No & Yes & \rd{0.98} & \rd{0.97} & \rd{0.97} & \rd{0.96}\\
+
+% \multirow{-4}{*}{\raggedright\arraybackslash Normal} & Yes & Yes & \rd{0.9} & \rd{0.77} & \rd{0.94} & \rd{0.88}\\
+% \cmidrule{1-7}
+%  & No & No & 0.97 & \rd{0.4} & 0.95 & \rd{0.61}\\
+
+%  & Yes & No & \rd{0.87} & \rd{0.2} & \rd{0.93} & \rd{0.48}\\
+
+%  & No & Yes & \rd{0.98} & \rd{0.53} & \rd{0.96} & \rd{0.57}\\
+
+% \multirow{-4}{*}{\raggedright\arraybackslash Uniform} & Yes & Yes & \rd{0.88} & \rd{0.21} & \rd{0.94} & \rd{0.53}\\
+% \bottomrule
+% \multicolumn{7}{p{3.5in}}{\footnotesize Based on 500 replications.
 $n=500$, $\beta_1=0.3$.
@@
 
-\end{tabular}
+% \end{tabular}
+\input{writeUps/coverageTab}
 \end{table}
@@
 \subsection{Accounting for Imperfect Compliance: Re-Analysis of the Obstetrics and Periodontal Therapy Study}
+\sloppy
 Periodontal disease, a weakening of bone and connective tissue supporting teeth, is highly prevalent among American adults and is associated with adverse pregnancy outcomes, including preterm birth \citep{periodontalEpi}.
@@
 
-\subsubsection{Principal Stratification and Alternatives}
+\subsubsection{Principal Stratification and Alternatives}\label{sec:psInEdTech}
 The goal of principal strafication in this context is to determine if assignment to the ASSISTments condition affected students who tend to request frequent bottom-out hints differently than students who tended to request fewer bottom-out hints.

### 2025-06-24 writeUps/responses.tex
finished first draft of responses

@@
 \usepackage{rotating}
-
+\usepackage{xspace}
 \usepackage[dvipsnames]{xcolor}
@@
 \end{itquote}
+
+We added Section \ref{sec:psInEdTech} to clarify the role principal stratification can play in analyzing implementation data---such as records of which students requested bottom-out hints---from an RCT studying educational effectiveness.
+In that section, we also compared principal stratification to other relevant methods, as you suggested.
 
@@
 \end{itquote}
+The simulation has two parts: in the first part, we estimate the bias and standard error of \geepers and \pmm for a range of values of $\alpha$ and $n$ when the assumptions of both methods are met.
+The goal for the first part is to determine the required sample size and covariate predictability necessary for \geepers to perform well.
+
+The second part presents results for scenarios where assumptions are violated. 
 The results in Figure \ref{fig:boxplots} and in Table \ref{tab:coverage} now hold $n$ fixed at $n=500$ per group and completely cross the other factors, i.e.
@@
 
-In Table \ref{tab:coverage}, results when one or more assumptions have been violated are highlighted in red.
+Following your fine suggestion, we formatted Table \ref{tab:coverage}, highlighting results in red when one or more assumptions have been violated.
+This format makes it clearer that, for instance, \pmm makes distributional assumptions while \geepers does not, and that \geepers is less sensitive to the assumptions that both estimators make, especially when $\alpha=0.5$.
 
-Application.
-
+\subsection{Application}
+\begin{itquote}
 I thought there were a number of less-than-desirable parts to the application.
@@
 Again, there is little here that demonstrates the value of GEEPERs.
+\end{itquote}
 
-To my mind, these problems could both be fixed.
-That is, the simulations could be redesigned to provide more compelling evidence.
-Next, the authors may be able to identify better applications.
-Therefore, the manuscript should be revised and resubmitted.
+These are important points, which we attempted to address in two ways. 
+Most importantly, we included an additional application example, a dental clinical trial with noncompliance, in which \geepers estimated statistically significant principal effects.
+This example is clean and straightforward, showing an instance where \geepers yields interesting results.
+However, our substantive interests and expertise, as well as those of most of the JREE readership, lie in education research rather than dental science, which prevents us from offering a comprehensive discussion of the method's application or interpreting the \emph{a priori} surprising results.
+
+For that reason, we retained the original application. 
+However, we tried to clarify the role PS could play in analyzing student implementation in an RCT studying educational technology. 
+For instance, in the Introduction, we wrote ``Indeed, if the use of bottom-out hints is considered part of the treatment, then those students who rarely used them are partial compliers---could the treatment still benefit them?''
+Most prominently, we added an extensive discussion in Section \ref{sec:psInEdTech}.
 

### 2025-06-24 writeUps/paper.tex
make displaySimPublish.r run

@@
 However, there is no readily identifiable comparison group for bottom-outers, who are only represented in one of the four treatment conditions, and who presumably struggled more frequently, on average, than their non-bottom-outer peers.
-We will formulate a solution using Principal stratification \citep{frangakis}, a framework for defining and estimating effects for post-treatment subgroups, that is, subgroups that may themselves be a function of treatment assignment.
+We will formulate a solution using principal stratification \citep{frangakis}, a framework for defining and estimating effects for post-treatment subgroups, that is, subgroups that may themselves be a function of treatment assignment.
 It relies on the notion of \emph{potential} subgroups called principal strata: which subgroup would a subject be in if they were (perhaps counterfactually) assigned to one treatment condition or another.
@@
 % \end{tabular}
-\input{writeUps/coverageTab}
+\input{coverageTab}
 \end{table}

### 2025-06-24 code/simulation/displaySimPublish.r

@@ -3,6 +3,7 @@ library(broom)
 library(kableExtra)
 library(gridExtra)
 library(ggpubr)
+library(tikzDevice)
 
 
 options(
@@ -127,11 +128,7 @@ ggsave("simFigs/biasSEbyN.jpg",plot=plotByN,width=6,height=3)
   ) %>%
   filter(rhat<1.1)
 
-### figure for appendix
 
-bp(pdb1,facet=PE~B1,title=bquote("Estimation Error by "~alpha),Labeller=label_parsed,labSize=2)+
-  labs(subtitle=bquote("Normal Resid., No Interactions, "~n==500~", "~mu[C]^1-mu[C]^0==0.3))
-ggsave('simFigs/byAlpha.jpg',width=6,height=3)
 
 
 #### figure in the paper
@@ -165,32 +162,8 @@ ggsave("simFigs/biasSEbyB1.jpg",plot=plotByAlpha,width=6,height=3)
 ggarrange(plotByN,plotByAlpha,ncol=1,common.legend = TRUE, legend = "bottom")
 ggsave("simFigs/biasSEbyB1n.jpg",width=5,height=4)
 
-%>%
-ggexport(filename="simFigs/biasSEbyB1n.jpg")
 
-#### table for appendix
-rmseB1 <- pdb1%>%
-  filter(rhat<1.1)%>%
-  group_by(b1,eff,estimator)%>%
-  summarize(rmse=sqrt(mean(errP^2,na.rm=TRUE)))%>%ungroup()
 
- coverageB1 <- pdb1%>%
-  filter(rhat<1.1)%>%
-  group_by(b1,eff,estimator)%>%
-   summarize(coverage=mean(CInormL<=pop & CInormU>=pop,na.rm=TRUE))%>%ungroup()
-
-bind_rows(
-  biasB1%>%filter(eff==1)%>%select(-eff)%>%mutate(b1=paste0(b1,'$'))%>%
-  pivot_wider(names_from=b1,names_prefix="$\\\\alpha=",values_from=bias)%>%mutate(measure="Bias"),
-  seB1%>%filter(eff==1)%>%select(-eff)%>%mutate(b1=paste0(b1,'$'))%>%
-  pivot_wider(names_from=b1,names_prefix="$\\\\alpha=",values_from=se)%>%mutate(measure="SE"),
-  rmseB1%>%filter(eff==1)%>%select(-eff)%>%mutate(b1=paste0(b1,'$'))%>%
[…]

### 2025-06-24 writeUps/paper.tex
edits

@@
 
-This paper expands on those latter results by presenting a new principal effect estimator, that boils down to a simple, three-step procedure: first, estimate principal scores, for instance, with logistic regression; then, impute unknown principal subgroup membership indicators with principal scores; and finally estimate principal effects using ordinary least squares (OLS) regression with interactions.
+This paper expands on those latter results by presenting a new principal effect estimator for the case of one-way noncompliance, in which principal stratum membership is observed in one treatment group but not the other (future work will extend our approach to general scenarios).
+Our estimator boils down to a simple, three-step procedure: first, estimate principal scores, for instance, with logistic regression; then, impute unknown principal subgroup membership indicators with principal scores; and finally estimate principal effects using ordinary least squares (OLS) regression with interactions.
 %, just as in the pre-treatment subgroup case.
@@
 We will refer to the estimators presented here as ``\textsc{geepers}'': general estimating equations for principal effects using regressions.
-\textsc{geepers} estimates rely on the correct specification of regression functions, but unlike common fully parametric methods, they make no assumptions about the shape of the distribution of regression errors.
+\textsc{geepers} estimates rely on the correct specification of regression functions, but unlike common fully-parametric methods, they make no assumptions about the shape of the distribution of regression errors.
 Unlike principal score weighting, \textsc{geepers} does not assume principal ignorability and admits to sandwich-style standard error estimates \citep{stefanskiBoos}, which we provide in the appendix.
@@
 After the initial submission of this manuscript, we became aware of \citet{richardson2023estimating}, which suggests a similar, though not identical, estimator to ours, using G-estimation of a structural nested mean model \citep[e.g.][]{wallace2017r}.
-Our development of \textsc{geepers} builds on \citet{richardson2023estimating} by emphasizing connections to both parametric mixture modeling and principal scores analysis, and comparing \textsc{geepers} to those methods in a simulation study.
+Our development of \textsc{geepers} builds on \citet{richardson2023estimating} by emphasizing connections to both parametric mixture modeling and principal score analysis, and comparing \textsc{geepers} to those methods in a simulation study.
 Moreover, \textsc{geepers}, which, in practice, is a combination of logistic regression and OLS, is more accessible to applied researchers while yielding very similar results.
@@
 
-We will focus on estimation in completely randomized experiments, that is, we will assume
+We will focus on estimation in completely randomized experiments; that is, we will assume
 \begin{ass}[Randomization]\label{ass:rand}
@@
 %The normal mixture modeling approach to estimating $\muc0$ and $\muc1$ \citep{imbens1997bayesian} assumes that, conditional on $\st$, $Y_C$ is normally distributed.
-Say an analyst assumes that, within principal strata, $Y_C$ is drawn from a parametric model with parameter vector $\bm{\delta}^s$, i.e., $f_{Y_C}(y|\st=s)=g^s(y)$.
+Say an analyst assumes that, within principal strata, $Y_C$ is drawn from a parametric model, i.e., $f_{Y_C}(y|\st=s)=g^s(y)$.
 For instance, \citep{imbens1997bayesian} assumes $Y_C$ is normally distributed within strata, so $g^s(y)=\phi\left(\frac{y-\mucs}{\sigma_C^s}\right)$, where $\phi(\cdot)$ is the standard normal density function, and $\sigma_C^0>0$ and $\sigma_C^1>0$ are standard deviations.
@@
 with parameter vector $\bm{\alpha}$.
-An analyst may estimate $\bm{\alpha}$ by fitting model \eqref{eq:prinScoreMod} using observed $\bm{S}$ and $\bxs$ for subjects with $Z=1$, and then compute estimated principal scores $\ppip{\bm{\hat{\alpha}}}$ for subjects with $Z_i=0$.
+An analyst may estimate $\bm{\alpha}$ as $\bm{\hat{\alpha}}$ by fitting model \eqref{eq:prinScoreMod} using observed $\bm{S}$ and $\bxs$ for subjects with $Z=1$, and then compute estimated principal scores $\hppi=\ppip{\bm{\hat{\alpha}}}$ for subjects with $Z_i=0$.
 
@@
 \end{equation}
-where $f_{Y_C}(y|\bm{x})$ and $g^s(y|\bm{x})$, $s=0,1$, are conditional densities which must be specified.
-Most commonly, researchers assume a linear normal model, i.e., $g^s(y|\bm{x})=\phi\left(\frac{y-\mucs-\bm{\gamma}'\bm{x}}{\sigma_C^s}\right)$, where the columns of $\bm{x}$ are centered so that $\mathbb{E}[\bm{x}]=\bm{0}$.
+where $g^s(y|\bm{x})$, $s=0,1$, are conditional densities which must be specified.
+Most commonly, researchers assume a linear normal model, e.g., $g^s(y|\bm{x})=\phi\left(\frac{y-\mucs-\bm{\gamma}'\bm{x}}{\sigma_C^s}\right)$, where the columns of $\bm{x}$ are centered so that $\mathbb{E}[\bm{x}]=\bm{0}$.
 %Ideally, a statistician will estimate parameters $\muc0$, $\muc1$, $\mut0$, $\mut1$, $\sigma^0_C$, $\sigma^1_C$, and $\bm{\alpha}$ simultaneously, using maximum likelihood or Bayesian techniques.
@@
 Alternatively, an analyst may assume ``principal ignorability'' \citep{jo,dingLu}:
-\begin{ass}[Principal Ignorability]\label{ass:PI}
+%\begin{ass}[Principal Ignorability]\label{ass:PI}
 \begin{equation}\label{eq:pi}
@@
 \end{equation}
-\end{ass}
+%\end{ass}
 or a somewhat weaker version, $\EE[Y_C|\st,\bxs]=\EE[Y_C|\bxs]$ \citep{feller2017principal}.
@@
 \begin{equation}\label{eq:psw}
-  \hat{\muc0}_w=\frac{\sum_{i: Z_i=0} Y_i(1-\pp)}{\sum_{i:Z_i=0}1-\pp}\text{ and } \hat{\muc1}_w=\frac{\sum_{i: Z_i=0} Y_i(\pp)}{\sum_{i:Z_i=0}\pp}
+  \hat{\muc0}_w=\frac{\sum_{i: Z_i=0} Y_i(1-\ppi)}{\sum_{i:Z_i=0}1-\ppi}\text{ and } \hat{\muc1}_w=\frac{\sum_{i: Z_i=0} Y_i(\ppi)}{\sum_{i:Z_i=0}\ppi}
 \end{equation}
@@
 %where the superscript $s$ stands for ``stacked.''
-A variance-covariance matrix estimated using the sandwich formula based on $\psi^s$ accounts for uncertainty from both steps of the estimation procedure.
+A variance-covariance matrix estimated using the sandwich formula based on $\psi$ accounts for uncertainty from both steps of the estimation procedure.
 
@@
 The methods all used the same ``observed'' dataset, consisting of outcomes $Y$, treatment assignments $Z$, two covariates $\bxs=\bxy=\bx=[x_1,x_2]$, and $S$ observed only when $Z=1$.
-The third covariate used in the data generating model, $x_3$, was ``unobserved''--since $x_3$ was correlated with both $\st$ and $Y$, Principal Ignorability (Assumption \ref{ass:PI}) was violated.
+The third covariate used in the data generating model, $x_3$, was ``unobserved''--since $x_3$ was correlated with both $\st$ and $Y$, Principal Ignorability \eqref{eq:pi} was violated.
 

### 2025-06-24 writeUps/responses.tex
edits

@@
 Thank you. We have fixed the formatting.
-\end{enumerate}
+\End{enumerate}
 
@@
 \end{itquote}
-This is an interesting suggestion---you're right that the \psw estimator did not include an outcome model, unlike \geepers and \nmm.
+This is an interesting suggestion---you're right that the \psw estimator did not include an outcome model, unlike \geepers and \pmm.
 However, we doubt that an outcome model would substantially reduce the bias of \psw in the simulation.
@@
 \end{itquote}
-We were also surprised by how well the \nmm performed in the simulation, especially given prior literature detailing its vulnerability \citep[][; cited in the paper]{griffin2008application,feller2016principal}.
+We were also surprised by how well the \pmm performed in the simulation, especially given prior literature detailing its vulnerability \citep[][; cited in the paper]{griffin2008application,feller2016principal}.
 Its similarity to \geepers in some scenarios is less surprising, since they are both essentially mixture model estimators.

### 2025-06-25 code/fh2t/estimates.r
take pretest measures out of outcome model for No Pretest analysis

@@ -82,14 +82,11 @@ estimates1rest <-
   lapply(alts,
          function(alt)
            est(getDat(alt),psMod=psModRest,
-               covFormY=update(formula(psModAll)[-2],
+               covFormY=update(formula(psModRest)[-2],
                                .~.
                                -SchIDPre+ClaIDPre
                                -UnexcusedDays6+ns(UnexcusedDays6,3)
-                               -pre_MA_total_score+ns(pre_MA_total_score,3)
                                -pre_MSE_total_score+I(pre_MSE_total_score< -2)+ns(pre_MSE_total_score,3)
-                               -pre.total_math_score+ns(pre.total_math_score,4)
-                               -Scale.Score5+ns(Scale.Score5,3)
                                )
                )
          )

### 2025-06-25 writeUps/paper.tex
Updates from Overleaf

@@
 
-When students get stuck on math problems during homework or classwork, it may be helpful to offer them hints or other ``just-in-time'' support \cite[e.g][]{jit}, possibly helping them overcome specific misconceptions or other sources of confusion and allowing them to move on in the assignment.
+When students get stuck on math problems during homework or classwork, it may be helpful to offer them hints or other ``just-in-time'' support \cite[e.g.,][]{jit}, possibly helping them overcome specific misconceptions or other sources of confusion and allowing them to move on in the assignment.
 The availability of problem-specific support for struggling students is one potential advantage of educational technology over traditional pencil-and-paper practice.
-On the other hand, the ready availability of hints may discourage students from working through difficult problems on their own, thereby inhibiting their learning.
+On the other hand, the ready availability of hints may discourage students from working through difficult problems independently, thereby hindering their learning.
 
@@
 (There are certainly other factors at play as well, so this analysis would not be conclusive.)
-Indeed, if the use of bottom-out hints is considered part of the treatment, then those students who rarely used them are partial compliers---could the treatment still benefit them?
+Indeed, if the use of bottom-out hints is considered part of the treatment, then those students who rarely used them were partial compliers---did the treatment still benefit them?
 
@@
 For instance, \citet{fellerEtAl2016} estimated differential effects of random assignment to the Head Start early childhood care program for children who, if assigned to control, would receive home-based care, and for those who would attend a different early childhood care program.
-Other methodologists have developed estimators for the evaluation of surrogate outcomes \citep{li2010bayesian}, the investigation of causal mechanisms \citep{lidsayPage}, sample attrition \citep{zhangRubin, ding2011}, and other problems.
+Other methodologists have developed estimators for evaluating surrogate outcomes \citep{li2010bayesian}, investigating causal mechanisms \citep{lidsayPage}, addressing sample attrition \citep{zhangRubin, ding2011}, and other related problems.
 Since interest is often in comparing effects between principal strata---a goal that runs counter to the exclusion restriction---the earliest principal stratification estimators built on \citet{imbens1997bayesian}'s parametric mixture models \citep[e.g.][]{Barnard01062003,mealli2004analyzing}, and that approach remains popular \citep{lidsayPage,fellerEtAl2016}.
-However, these models are often hard to specify and hard or impossible to test.
-Even when they are well-specified they can yield estimates that are severely biased and misleading \citep{griffin2008application,feller2016principal}.
+However, these models are often difficult to specify and can be either hard or impossible to test.
+Even when they are well-specified, they can yield estimates that are severely biased and misleading \citep{griffin2008application,feller2016principal}.
 The most common type of parametric model, the normal mixture model, is especially dubious for studying bottom-outers in \citet{impactPaper}'s RCT, whose outcome, the total score on a 10-item posttest, is not normally distributed.
@@
 This paper expands on those latter results by presenting a new principal effect estimator for the case of one-way noncompliance, in which principal stratum membership is observed in one treatment group but not the other (future work will extend our approach to general scenarios).
-Our estimator boils down to a simple, three-step procedure: first, estimate principal scores, for instance, with logistic regression; then, impute unknown principal subgroup membership indicators with principal scores; and finally estimate principal effects using ordinary least squares (OLS) regression with interactions.
+Our estimator boils down to a simple, three-step procedure: first, estimate principal scores, for instance, with logistic regression; then, impute unknown principal subgroup membership indicators with principal scores; and finally, estimate principal effects using ordinary least squares (OLS) regression with interactions.
 %, just as in the pre-treatment subgroup case.
@@
 Lastly, this work includes an extensive simulation study that compares \textsc{geepers} to parametric mixture modeling and principal score weighting under a range of favorable and challenging scenarios.
-The simulation demonstrates the strengths and weaknesses of each method and the tradeoffs between them.
+The simulation demonstrates the strengths and weaknesses of each method, as well as the trade-offs between them.
 
@@
 The average treatment effect in each principal stratum, $\eff1=\EE[\tau|\st=1]$ or $\eff0=\EE[\tau|\st=0]$, is called a principal effect. %
-The challenge in estimating principal effects is because strata membership $\sti$ is only observed when $Z_i=1$, in which case $\sti=S_i$, the observed value of $S$.
+The challenge in estimating principal effects is that strata membership $\sti$ is only observed when $Z_i=1$, in which case $\sti=S_i$, the observed value of $S$.
 
@@
 This approach has several problems.
-First, the assumption of a conditional parametric model %normality assumption 
-is untestable and limiting.
+First, the assumption of a conditional parametric model is untestable and limiting.
 Unfortunately, even when it holds, mixture model estimates can be unreliable and, in some cases, severely biased \citep{griffin2008application,feller2017principal}.
@@
 
-Principal scores can potentially improve inference in a finite mixture model, by re-writing \eqref{eq:mixtureModel} as
+Principal scores can potentially improve inference in a finite mixture model by modifying \eqref{eq:mixtureModel} as
 \begin{equation}\label{eq:mixtureModelPS}
@@
 
-In each run of the simulation, we simulated three independent standard normal covariates, $x_k$ $k=1,\dots,3$.
+In each run of the simulation, we simulated three independent standard normal covariates, $x_k$, $k=1,\dots,3$.
 However, only the first two covariates were ``observed,'' i.e., were included in the analysis model.
@@
 The methods all used the same ``observed'' dataset, consisting of outcomes $Y$, treatment assignments $Z$, two covariates $\bxs=\bxy=\bx=[x_1,x_2]$, and $S$ observed only when $Z=1$.
-The third covariate used in the data generating model, $x_3$, was ``unobserved''--since $x_3$ was correlated with both $\st$ and $Y$, Principal Ignorability \eqref{eq:pi} was violated.
+The third covariate used in the data-generating model, $x_3$, was ``unobserved''--since $x_3$ was correlated with both $\st$ and $Y$, Principal Ignorability \eqref{eq:pi} was violated.
 
@@
 We omitted the $\alpha=0$ case from Figures \ref{fig:alphan} and \ref{fig:boxplots} since the variance of \geepers and \pmm estimates was so large as to dominate the plot. 
-However, the first column of results in Table \ref{tab:coverage} shows that, when other assumptions are met, CIs from both methods tend to over-cover when $\alpha=0$. 
+However, the first column of results in Table \ref{tab:coverage} shows that, when other assumptions are met, confidence intervals from both methods tend to over-cover when $\alpha=0$. 
 In a randomized experiment, $\alpha$ can be estimated directly, making Assumption \ref{ass:vps} testable.
@@
 \geepers may be the best choice when conditional normality may not hold, or if there are $X:Z$ interactions and $\bm{x}$ is sufficiently predictive of $\st$ (AUC $\ge$0.675, corresponding to $\alpha \ge 0.5$, may serve as a threshold).
-These suggestions may depend heavily on the simulation setup here; applying them in general scenarios requires more research.
+These suggestions may depend heavily on the simulation setup; applying them in general scenarios requires further research.
 
@@
 Next, we present a novel analysis of data from the RCT of computer-based math applications reported in \citet{impactPaper}, and look into the role of bottom-out hints, as described in the introduction.
-This analysis aims to shed light on an important question in education research, illustrates model and covariate selection, and includes a discussion on the plausibility of identification assumptions.
+This analysis aims to shed light on an important question in education research, illustrate model and covariate selection, and include a discussion on the plausibility of identification assumptions.
 
@@
 The conclusion of these methods is somewhat surprising: that the effect was larger for subjects who would not complete the treatment than for those who would (the two strata had similar bleeding percentages at baseline, and a logit transformation of the outcome leaves the results qualitatively unchanged).
-One possible explanation is the conjecture that subjects stop showing up for treatment once their periodontal disease has abated.
+One possible explanation is the conjecture that subjects stopped attending treatment once their periodontal disease improved.
 
@@
 \subsubsection{Principal Stratification and Alternatives}\label{sec:psInEdTech}
-The goal of principal strafication in this context is to determine if assignment to the ASSISTments condition affected students who tend to request frequent bottom-out hints differently than students who tended to request fewer bottom-out hints.
+The goal of principal stratification in this context is to determine if assignment to the ASSISTments condition affected students who tended to request frequent bottom-out hints differently than students who tended to request fewer bottom-out hints.
 Is this the right question to ask?
@@
 
-Another arguable disadvantage of the observational study approach is that it overlooks the broader context of the experiment, focusing on whether bottom-out hints affect posttest scores rather than examining their role in ASSISTments' overall effectiveness.
+Another arguable disadvantage of the observational study approach is that it overlooks the broader context of the experiment, focusing solely on whether bottom-out hints affect posttest scores rather than examining their role in the overall effectiveness of ASSISTments.
 A second approach, mediation analysis, contextualizes the bottom-out effect within the larger experiment.
@@
 \emph{A priori}, the student attribute that poses the greatest threat to these modeling choices is prior achievement, that is, scores on the pretest and the fifth-grade state test.
-This is both because they measure a similar construct to the posttest, and because \cite{impactPaper} found a pretest-treatment interaction in a different treatment contrast.
-Following that reasoning, we conducted a third parallel analysis excluding prior achievement covariates.% in $\bxs$ but including all covariates in $\bxy$.
+This is both because they measure a similar construct to the posttest and because \cite{impactPaper} found a pretest-treatment interaction in a different treatment contrast.
+Following that reasoning, we conducted a third parallel analysis excluding prior achievement covariates. % in $\bxs$ but including all covariates in $\bxy$.
 
@@
 
-Outcome models for all analyses adjusted for all of the variables that were in the corresponding principal score models, in addition to measurements of students' prior knowledge; also, the models substituted class fixed effects for school fixed effects.\footnote{Why use a coarser set of fixed effects for the principal score model than the outcome regression?
+Outcome models for all analyses included for all of the variables that were in the corresponding principal score models, in addition to measurements of students' prior knowledge; also, the models substituted class fixed effects for school fixed effects.\footnote{Why use a coarser set of fixed effects for the principal score model than the outcome regression?
 First, the principal score model was fit to a smaller sample (only students in the ``Instant'' condition), so some classrooms only included one student, which can lead to poor fit.
@@
 
-The Bayesian mixture model assumed that outcomes were normally distributed, conditional on covariates and principal stratum, with a residual standard deviation that varied between principal strata.
+The Bayesian mixture model assumed that outcomes were normally distributed, conditional on covariates and principal strata, with a residual standard deviation that varied between principal strata.
 This modeling assumption is necessarily false since post-test scores were integers between 0 and 10.
@@
 The biggest difference between the methods was in the contrast with FH2T, in which the Bayesian mixture model resulted in a larger difference between the two principal effects, as well as the largest standard errors, while the two \textsc{psw} principal effect estimates were quite similar.
-However, these differences are difficult to interpret due to the large uncertainty around all of the point estimates.
+However, these differences are difficult to interpret due to the large uncertainty surrounding all the point estimates.
 
@@
 If, say, $\sti\ge S_{Ci}$, then every subject in the control condition with $S_i=1$ is in the always survive stratum, while those subjects in the treatment condition with $S_i=1$ are a mixture of the always survive and $\sti=1;\;S_{Ci}=0$ stratum.
-This scenario is broadly similar to the strong monotonicity, one-way noncompliance situation that this paper discussed.
+This scenario is broadly similar to the strong monotonicity, one-way noncompliance situation discussed in this paper.
 On the other hand, when no monotonicity assumption holds, both $\st$ and $S_C$ will have to be imputed for every subject; further research is necessary to determine the appropriate way to do so.
@@
 However, the results are consistent with the hypothesis that ASSISTments is less effective for frequent bottom-out requesters when compared to a similar program without hints or immediate feedback.
-We hypothesize that many students who request bottom-out hints frequently are gaming the system, abetted by the availability of bottom-out hints.
+We hypothesize that many students who frequently request bottom-out hints are gaming the system, facilitated by the availability of these hints.
 

### 2025-06-26 code/simulation/displaySimPublish.r
changed simulation to mu01=0

@@ -38,8 +38,8 @@ source('code/simulation/readSimFuncs.r')
 #### after simulation results have been loaded and pre-processed...
 ## source('code/simulation/readSim.r')
 load('simResults/fullResults.RData')
-load('simResults/resultsNs.RData')
-load('simResults/resultsB1s.RData')
+load('simResults/resultsNs_mu01is0.RData')
+load('simResults/resultsB1s_mu01is0.RData')
 
 
 ###############################################################################
@@ -94,7 +94,9 @@ plotByN=bind_rows(
   geom_point()+geom_line()+geom_hline(yintercept=0)+#,linetype="dotted",size=2)+
   scale_y_continuous(name="Bias",sec.axis=sec_axis(trans=~.*3,name='Standard Error'))+
   scale_shape_manual(values=c(0,16))+
-  annotate('text',600,.11,label=list(bquote(atop("Normal Resid., No Interactions, ",alpha==0.5~", "~mu[C]^1-mu[C]^0==0.3))),parse=TRUE)+
+    annotate('text',600,.11,label=list(bquote(paste(#atop(
+                                "Normal Resid., No Interactions, ",alpha==0.5))),#", "~mu[C]^1-mu[C]^0==0.3))),
+             parse=TRUE)+
   scale_x_continuous(name="Sample Size Per Group (n)",breaks=c(seq(100,500,200),1000))+
   #ggtitle("Bias and Standard Error by n")+
   theme(legend.title=element_blank())
@@ -153,8 +155,12 @@ plotByAlpha=bind_rows(
   geom_point()+geom_line()+geom_hline(yintercept=0)+#,linetype="dashed")+
   scale_y_continuous(name="Bias",sec.axis=sec_axis(trans=~.*5,name='Standard Error'))+
   scale_shape_manual(values=c(0,16))+
-  annotate('text',.7,.08,label=list(bquote(atop("Normal Resid., No Interactions, ",n==500~", "~mu[C]^1-mu[C]^0==0.3))),parse=TRUE)+
-  xlab(bquote(alpha))#+ggtitle(bquote("Bias and Standard Error by "~alpha))
+    annotate('text',.7,.08,label=list(bquote(
+                               paste(#atop(
+                                   "Normal Resid., No Interactions, ",n==500))),# "~mu[C]^1-mu[C]^0==0.3))),
+             parse=TRUE)+
[…]

### 2025-06-26 code/simulation/readSim.r

@@ -9,16 +9,16 @@ print(load('simResults/pswResults.RData'))
 results=loadRes(pswResults=pswResults)
 save(results,file='simResults/fullResults.RData')
 
-resultsN100=loadRes(ext2='n100')
-save(resultsN100,file='simResults/resultsN100.RData')
+resultsN100=loadRes(ext2='n100_mu01is0')
+#save(resultsN100,file='simResults/resultsN100.RData')
 
 
-resultsNs=loadRes(ext2='ns')
+resultsNs=loadRes(ext2="ns_mu01is0")
 
-resultsB1s=loadRes(ext2='b1s')
+resultsB1s=loadRes(ext2='b1ss_mu01is0')
 
 ### merge results by n
-load('simResults/casesns.RData')
+load('simResults/casesns_mu01is0.RData')
 casesNs=cases
 load('simResults/casesn100.RData')
 casesN100=cases
@@ -26,18 +26,29 @@ load('simResults/cases.RData')
 
 resultsNs=bind_rows(
   resultsNs,
-  filter(resultsN100,run%in%which(apply(casesN100[,-1],1,function(x) all(x==casesNs[1,-1])))),
-  filter(results,run%in%which(apply(cases[,-1],1,function(x) all(x==casesNs[1,-1])))))
+  filter(resultsN100,mu01%in%resultsNs$mu01[1],
+         errDist%in%resultsNs$errDist,
+         b1%in%resultsNs$b1,
+         intS%in%resultsNs$intS,
+         intZ%in%resultsNs$intZ
+         ),
+  filter(results,mu01==resultsNs$mu01[1],
+         errDist%in%resultsNs$errDist,
+         b1%in%resultsNs$b1,
+         intS%in%resultsNs$intS,
+         intZ%in%resultsNs$intZ
+         ))%>%
+    filter(estimator!="psw")
 
-save(resultsNs,file='simResults/resultsNs.RData')
+save(resultsNs,file='simResults/resultsNs_mu01is0.RData')
 
 ### merge results by B1s
 casesTot=cases
-load('simResults/casesb1s.RData')
+load('simResults/casesb1ss_mu01is0.RData')
 casesB1=cases
 
 resultsB1s=bind_rows(
   resultsB1s,
   filter(results,run%in%(left_join(casesB1[1,],casesTot%>%mutate(run=1:n())%>%select(-b1))%>%pull(run)))
   )
[…]

### 2025-06-26 code/simulation/readSimFuncs.r

@@ -87,7 +87,7 @@ bayesProc <- function(res1,facs){
 
 
 loadRes <- function(ext1='',ext2='',pswResults){
-    load(paste0('simResults/cases',ext1,ext2,'.RData'))
+    load(paste0('simResults',ext1,'/cases',ext2,'.RData'))
 
     results <- list()
 #    summ <- NULL
@@ -95,7 +95,7 @@ loadRes <- function(ext1='',ext2='',pswResults){
     for(i in 1:nrow(cases)){#length(fn)){
         #if(i %% 10==0)
             cat(round(i/nrow(cases)*100),'%',sep='')#length(fn)*100), '% ')
-      load(paste0('simResults/sim',i,ext1,ext2,'.RData'))#fn[i]))
+      load(paste0('simResults',ext1,'/sim',i,ext2,'.RData'))#fn[i]))
       if(!missing(pswResults))
         res <- map(1:length(res),
                    ~append(res[[.]],list(psw=unlist(pswResults[[i]][.,]))))

### 2025-06-26 code/simulation/runSim.r

@@ -1,0 +1,89 @@
+##########################################################################################
+########### run the simulation
+##########################################################################################
+
+
+library(dplyr)
+library(rstan)
+library(parallel)
+
+
+stanPSmod=stan_model("ps.stan",model_name="stanPSmod")
+
+#rstan_options(auto_write = TRUE)
+
+### number of repetitions
+nrep <- 500
+
+### For running with multiple cores:
+
+ncore <-  8#
+
+
+if(.Platform$OS.type=='windows' & ncore>1){
+  print("making cluster")
+  cl <- makeCluster(ncore)
+  clusterEvalQ(cl,library(dplyr))
+  clusterEvalQ(cl,library(rstan))
+  clusterEvalQ(cl,rstan_options(auto_write=TRUE))
+  clusterEvalQ(cl, stan_model('ps.stan'))
+  clusterEvalQ(cl,source('regression.r'))
+  clusterEvalQ(cl,source('simFuncs.r'))
+} else cl <- NULL
+
+
+source('simFuncs.r')
+
+source('regression.r')
+
+
+fullsim(nrep,ns=c(500,1000),mu01=c(0,.3),mu10=0,b1s=c(0,0.2,0.5),
+	ext='',cl=cl,start=1)
+
+
+pswResults=sapply(seq(max(as.numeric(gsub('dat|\\.RData','',list.files('simData/'))),na.rm=TRUE)),
+                  function(i){
+                    print(i)
+                    load(paste0('simData/dat',i,'.RData'))
+                    facs=attributes(datasets[[1]])$facs
+                    out=as.data.frame(t(vapply(datasets,psw,psw(datasets[[1]]))))
+                    attr(out,'facs')=facs
+                    out
+                  },simplify=FALSE)
+save(pswResults,file='simResults/pswResults.RData')
+
+
+################################
+### look at wider ranges of n and alpha
+fullsim(nrep,ns=100,mu01=.3,mu10=0,errDist=c('norm','unif'),b1s=c(0,.2,.5),ext='n100',cl=cl)
+fullsim(nrep,ns=100,mu01=0,mu10=0,errDist=c('norm'),b1s=c(.5),intS=FALSE,intZ=FALSE,ext='n100_mu01is0')#,cl=cl)
[…]

### 2025-06-26 code/simulation/runSimBayes.r

@@ -9,8 +9,9 @@ library(parallel)
 
 #rstan_options(auto_write = TRUE)
 
+nrep <- 500
 
-ncore <-  8#detectCores()/2
+ncore <-  4#detectCores()/2
 #cl <- makeCluster(ncore)
 #     clusterEvalQ(cl,library(tidyverse))
 #     clusterEvalQ(cl,library(rstan))
@@ -31,34 +32,35 @@ source('code/regression.r')
 
 #print(paste("starting at case",max(nums,na.rm=TRUE)+1))
 
-fullsim(500,ns=c(500,1000),mu01=c(0,.3),mu10=0,b1s=c(0,0.2,0.5),
-	ext='',cl=NULL,start=1)
+#fullsim(500,ns=c(500,1000),mu01=c(0,.3),mu10=0,b1s=c(0,0.2,0.5),
+#	ext='',cl=NULL,start=1)
 
 
 #system('drive push --no-clobber -quiet')
 
 
-pswResults=sapply(seq(max(as.numeric(gsub('dat|\\.RData','',list.files('simData/'))),na.rm=TRUE)),
-                  function(i){
-                    print(i)
-                    load(paste0('simData/dat',i,'.RData'))
-                    facs=attributes(datasets[[1]])$facs
-                    out=as.data.frame(t(vapply(datasets,psw,psw(datasets[[1]]))))
-                    attr(out,'facs')=facs
-                    out
-                  },simplify=FALSE)
-save(pswResults,file='simResults/pswResults.RData')
+## pswResults=sapply(seq(max(as.numeric(gsub('dat|\\.RData','',list.files('simData/'))),na.rm=TRUE)),
+##                   function(i){
+##                     print(i)
+##                     load(paste0('simData/dat',i,'.RData'))
+##                     facs=attributes(datasets[[1]])$facs
+##                     out=as.data.frame(t(vapply(datasets,psw,psw(datasets[[1]]))))
+##                     attr(out,'facs')=facs
+##                     out
+##                   },simplify=FALSE)
+## save(pswResults,file='simResults/pswResults.RData')
 
 
[…]

### 2025-06-26 code/simulation/displaySimPublish.r
simulation revisions

@@ -97,7 +97,7 @@ plotByN=bind_rows(
     annotate('text',600,.11,label=list(bquote(paste(#atop(
                                 "Normal Resid., No Interactions, ",alpha==0.5))),#", "~mu[C]^1-mu[C]^0==0.3))),
              parse=TRUE)+
-  scale_x_continuous(name="Sample Size Per Group (n)",breaks=c(seq(100,500,200),1000))+
+  scale_x_continuous(name="Sample Size Per Group (n)",breaks=c(seq(100,700,200),1000))+
   #ggtitle("Bias and Standard Error by n")+
   theme(legend.title=element_blank())
 ggsave("simFigs/biasSEbyN.jpg",plot=plotByN,width=6,height=3)
@@ -150,10 +150,11 @@ full_join(seB1,biasB1)%>%filter(eff==1,estimator!="PSW",b1>0)%>%summarize(across
 plotByAlpha=bind_rows(
   seB1%>%mutate(se=se/5,meas="SE")%>%rename(what=se),
   biasB1%>%mutate(meas="Bias")%>%rename(what=bias))%>%
-  filter(b1>0,eff==1,estimator!="PSW")%>%
+  filter(b1>0.1,eff==1,estimator!="PSW")%>%
   ggplot(aes(b1,what,color=estimator,group=paste0(estimator,meas),linetype=meas,fill=estimator,shape=meas))+
   geom_point()+geom_line()+geom_hline(yintercept=0)+#,linetype="dashed")+
-  scale_y_continuous(name="Bias",sec.axis=sec_axis(trans=~.*5,name='Standard Error'))+
+    scale_y_continuous(name="Bias",sec.axis=sec_axis(trans=~.*5,name='Standard Error'))+
+    scale_x_continuous(breaks=seq(0,1,0.2),minor_breaks=seq(0.1,0.9,0.2),labels=sprintf("%0.1f",seq(0,1,0.2)))+
   scale_shape_manual(values=c(0,16))+
     annotate('text',.7,.08,label=list(bquote(
                                paste(#atop(
@@ -201,7 +202,7 @@ pd <- results %>%
 ### figure for paper
 
 norm<-bp(pd,
-   subset=b1 > 0& errDist=='norm'& eff==1&mu01==0.3&n==500&PE=='Stratum 1',
[…]

### 2025-06-26 code/simulation/readSim.r

@@ -6,8 +6,13 @@ source('code/simulation/readSimFuncs.r')
 #### read, process results from main simulation
 print(load('simResults/pswResults.RData'))
 
-results=loadRes(pswResults=pswResults)
-save(results,file='simResults/fullResults.RData')
+if(file.exists("simResults/fullResults.RData")){
+    load("simResults/fullResults.RData")
+} else{
+    results=loadRes(pswResults=pswResults)
+    save(results,file='simResults/fullResults.RData')
+}
+
 
 resultsN100=loadRes(ext2='n100_mu01is0')
 #save(resultsN100,file='simResults/resultsN100.RData')
@@ -16,6 +21,8 @@ resultsN100=loadRes(ext2='n100_mu01is0')
 resultsNs=loadRes(ext2="ns_mu01is0")
 
 resultsB1s=loadRes(ext2='b1ss_mu01is0')
+
+resB101=loadRes(ext2="b01_mu01is0")
 
 ### merge results by n
 load('simResults/casesns_mu01is0.RData')
@@ -48,7 +55,8 @@ load('simResults/casesb1ss_mu01is0.RData')
 casesB1=cases
 
 resultsB1s=bind_rows(
-  resultsB1s,
+    resultsB1s,
+    resB101,
   filter(results,run%in%(left_join(casesB1[1,],casesTot%>%mutate(run=1:n())%>%select(-b1))%>%pull(run)))
   )
 save(resultsB1s,file='simResults/resultsB1s_mu01is0.RData')

### 2025-06-26 code/simulation/readSimFuncs.r

@@ -87,7 +87,9 @@ bayesProc <- function(res1,facs){
 
 
 loadRes <- function(ext1='',ext2='',pswResults){
-    load(paste0('simResults',ext1,'/cases',ext2,'.RData'))
+    if(file.exists(paste0('simResults',ext1,'/cases',ext2,'.RData'))){
+        load(paste0('simResults',ext1,'/cases',ext2,'.RData'))
+    } else cases <- cbind(1:sum(grepl(paste0('sim[0-9]+',ext2,'\\.RData'),list.files(paste0("simResults",ext1,"/")))))
 
     results <- list()
 #    summ <- NULL

### 2025-06-26 code/simulation/runSim.r

@@ -38,7 +38,7 @@ source('regression.r')
 
 
 fullsim(nrep,ns=c(500,1000),mu01=c(0,.3),mu10=0,b1s=c(0,0.2,0.5),
-	ext='',cl=cl,start=1)
+	ext='',cl=cl,ncores=ncore,start=1)
 
 
 pswResults=sapply(seq(max(as.numeric(gsub('dat|\\.RData','',list.files('simData/'))),na.rm=TRUE)),
@@ -55,16 +55,15 @@ save(pswResults,file='simResults/pswResults.RData')
 
 ################################
 ### look at wider ranges of n and alpha
-fullsim(nrep,ns=100,mu01=.3,mu10=0,errDist=c('norm','unif'),b1s=c(0,.2,.5),ext='n100',cl=cl)
-fullsim(nrep,ns=100,mu01=0,mu10=0,errDist=c('norm'),b1s=c(.5),intS=FALSE,intZ=FALSE,ext='n100_mu01is0')#,cl=cl)
-fullsim(nrep,ns=500,mu01=0,mu10=0,errDist=c('norm'),b1s=c(.1),intS=FALSE,intZ=FALSE,ext='b1s.1_mu01is0')#,cl=cl)
 
-fullsim(nrep,ns=c(200,300,400,600,700,800),errDist='norm',b1s=.5,intS=FALSE,intZ=FALSE,mu01=.3,mu10=0,ext='ns',cl=cl)
+fullsim(nrep,ns=c(100,200,300,400,600,700,800,900),errDist='norm',b1s=.5,intS=FALSE,intZ=FALSE,mu01=.3,mu10=0,ext='ns',cl=cl,
+        ncores=ncore)
 
-fullsim(nrep,ns=c(200,300,400,600,700,800),errDist='norm',b1s=.5,intS=FALSE,intZ=FALSE,mu01=0,mu10=0,ext='ns_mu01is0')#,cl=cl)
+fullsim(nrep,ns=c(200,300,400,600,700,800),errDist='norm',b1s=.5,intS=FALSE,intZ=FALSE,mu01=0,mu10=0,ext='ns_mu01is0',cl=cl,
+            ncores=ncore)
 
[…]

### 2025-06-26 code/simulation/displaySimPublish.r
coverage tab oops

@@ -333,8 +333,6 @@ condRed <- function(coverage,intZ,intS,estimator,errDist,b1=1)
 
 
 sink('writeUps/coverageTab.tex')
-
-
 cbind(
      coverage%>%
       mutate(coverage=condRed(coverage,intZ,intS,estimator,errDist,b1))%>%
@@ -359,7 +357,7 @@ cbind(
     add_header_above(c(" " = 3, "$\\\\alpha=0$" = 2,
                        "$\\\\alpha=0.2$" = 2, "$\\\\alpha=0.5$" = 2),escape=FALSE)%>%
     collapse_rows(columns=1,latex_hline="major",valign="middle")%>%
-    footnote(general=c("\\\\footnotesize Based on 500 replications. $n=500$. Simulation standard error $\\\\approx 1$ percentage point. Estimates colored \\\\rd{red} indicate cases where the assumptions of the model are not met."),escape=FALSE,footnote_as_chunk = TRUE,threeparttable=TRUE)
+    footnote(general=c("\\\\footnotesize Based on 500 replications. $n=500$. Simulation standard error $\\\\approx 1$ percentage point. Estimates colored \\\\rd{red} indicate cases where the assumptions of the model are not met."),escape=FALSE,footnote_as_chunk = TRUE,threeparttable=TRUE)%>%print()
 sink()
 
 ########################

### 2025-06-26 writeUps/paper.tex

@@
 
-\usepackage[dvipsnames]{xcolor}
+\usepackage[dvipsnames,table]{xcolor}
 
 %\startlocaldefs
+\definecolor{lightgray}{gray}{0.9}
 
@@
 The study was designed to answer three sets of overarching questions.
-First, how does the performance of the \textsc{geepers} estimator vary under different conditions, including sample size, the extent to which $\bx$ predicts $\st$, whether and how the principal effects themselves vary, and when important interactions are omitted from the outcome model?
+First, how does the performance of the \textsc{geepers} estimator vary under different conditions, including sample size, the extent to which $\bx$ predicts $\st$, %whether and how the principal effects themselves vary,
+and when important interactions are omitted from the outcome model?
 Second, how does \textsc{geepers} compare with mixture modeling?
@@
 \begin{equation}\label{eq:y-sim}
-Y_i=\beta_1\sti+\beta_3Z_i\sti+(\gamma_1+\gamma_2\sti+\gamma_3Z_i)(x_{1i}+x_{2i})+\frac{1}{\sqrt{6}}x_{3i}+\epsilon_i%\\
+Y_i=0.3 Z_i\sti+(\gamma_1+\gamma_2\sti+\gamma_3Z_i)(x_{1i}+x_{2i})+\frac{1}{\sqrt{6}}x_{3i}+\epsilon_i%\\
 \end{equation}
@@
 % \end{equation}
-with $\beta_3=0.3-\beta_1$, $\EE[\epsilon_i]=0$, and $Var(\epsilon_i)=1/2$.
-The distribution of $\epsilon_i$, $\beta_1$, and values of coefficients $\gamma_1$, $\gamma_2$, and $\gamma_3$ were manipulated factors in the study.
-As described below, in most simulated scenarios, $\gamma_1=1/\sqrt{6}$ and $\gamma_2=\gamma_3=0$.
-In these cases there is no average treatment effect in the $\st=0$ stratum, and, since  $Var(\gamma_1(x_1+x_2+x_3))=1/2$ and $Var(\epsilon_i)=1/2$, covariates explained half of the variance of $Y$ within treatment group and principal stratum, and observed covariates $x_1$ and $x_2$ explained 1/3 of that variance.
-These values are in line with estimates from outcome models in Section \ref{sec:fh2t} and are typical in education contexts \cite{hedgesHedberg}.
+with %$\beta_3=0.3-\beta_1$, 
+$\EE[\epsilon_i]=0$, and $Var(\epsilon_i)=1/2$.
+The distribution of $\epsilon_i$ and values of coefficients $\gamma_1$, $\gamma_2$, and $\gamma_3$ were manipulated factors in the study.
 
 Treatment assignment $Z$ was simulated as independent of all the other variables, with half of the subjects assigned to treatment $Z=1$ and the rest to control $Z=0$.
+
+Since the covariates were centered and independent of $Z$, the true principal effects were 
+$\eff0=\gamma_3\EE[x1+x2|\st=0]$ and $\eff1=0.3+\gamma_3\EE[x1+x2|\st=1]$; in most simulation scenarios $\gamma_3=0$ so $\eff0=0$ and $\eff1=0.3$.
 
@@
   \centering
+  %\rowcolors{1}{}{lightgray}
 \begin{tabular}{r|rl}%{*{3}{c}}
@@
 The distribution of $\epsilon_i$&\multicolumn{2}{l}{ Normal, Uniform}\\
-$\beta_1$ & \multicolumn{2}{l}{0, 0.3}\\
-Interactions in \eqref{eq:y-sim} & None & $\gamma_1=1/\sqrt{6};\gamma_2=\gamma_3=0$\\ 
+%$\beta_1$ & \multicolumn{2}{l}{0, 0.3}\\
+\multirow{4}{*}{Interactions in \eqref{eq:y-sim}} & None & $\gamma_1=1/\sqrt{6};\gamma_2=\gamma_3=0$\\ 
 & $\bm{x}:\st$ & $\gamma_1=3/(4\sqrt{6})$; $\gamma_2=1/(2\sqrt{6})$; $\gamma_3=0$\\
@@
 
-$\bm{\beta_1}$\textbf{:} When $\beta_1=0$, there is only one mixture component in the control
-group, which can be problematic for mixture modeling \citep{griffin2008application,feller2016principal}.
+% $\bm{\beta_1}$\textbf{:} When $\beta_1=0$, there is only one mixture component in the control
+% group, which can be problematic for mixture modeling \citep{griffin2008application,feller2016principal}.
 Conversely, when $\beta_1=0.3$, unless there is an interaction between $Z$ and covariates (i.e., $\gamma_3\ne 0$), $\eff0=\eff1=0$, so there is no average treatment effect for either principal stratum.
 
-\textbf{\emph{Interactions between $\st$ or $Z$ and covariates:}} The last two factors---the presence of interactions between covariates and $\st$ and $Z$, respectively---test the performance of a model that does not include interactions when that modeling assumption is false.
+\textbf{\emph{Interactions between $\st$ or $Z$ and covariates:}} The factor---the presence of interactions between covariates and $\st$ and $Z$, respectively--- tests the performance of a model that does not include interactions when that modeling assumption is false.
+
 When neither interaction was present, $\gamma_2=\gamma_3=0$, then $\gamma_1=1/\sqrt{6}$.
+In these cases, since  $Var(\gamma_1(x_1+x_2+x_3))=1/2$ and $Var(\epsilon_i)=1/2$, covariates explained half of the variance of $Y$ within treatment group and principal stratum, and observed covariates $x_1$ and $x_2$ explained 1/3 of that variance.
+These values align with estimates from outcome models in Section \ref{sec:fh2t} and are typical in education contexts \cite{hedgesHedberg}.
+
 When there was an interaction between covariates and $\sti$, then $\gamma_1=3/(4\sqrt{6})$ and $\gamma_2=1/(2\sqrt{6})$, so that the slopes for $x_1$ and $x_2$ varied by half their magnitude between principal strata, and on average they were equal to $1/\sqrt{6}$, as in the no-interaction case.
@@
 \pmm, the parametric mixture estimator, then fits two outcome models:
-For the treatment group, the regression model $Y=\beta_0^T+\beta_1^TS+\bm{\gamma}'\bx+\epsilon$ and for the control group, a mixture model of the form \eqref{eq:mixtureModelPS}, but with $\muc0$ and $\muc1$ replaced by regression functions $\beta_0^C+\bm{\gamma}'\bx$ and $\beta_0^C+\beta_1^C+\bm{\gamma}'\bx$, respectively.
+For the treatment group, the regression model $Y=\beta_0^T+\beta_1^TS+\bm{\gamma}'\bx+\epsilon$ and for the control group, a mixture model of the form \eqref{eq:mixtureModelPS}, but with linear normal models for $g^s(y|x_1,x_2)$.
 Residual variance was allowed to differ across treatment groups, but not across principal strata.
-These models were fit simultaneously using Bayesian Markov Chain Monte Carlo with Stan \citep{rstan}, and principal effects $\eff0$ and $\eff1$ estimated as $\beta_0^T-\beta_0^C$ and $\beta_0^T+\beta_1^T-\beta_0^C-\beta_1^T$, respectively.
+These models were fit simultaneously using Bayesian Markov Chain Monte Carlo with Stan \citep{rstan}.
 We interpreted the posterior mean as a point estimate and the posterior standard deviation as a standard error.
 
-For \textsc{geepers} and \psw, the principal score model was fit by itself, to data from the treatment group, using standard maximum likelihood methods.
+For \textsc{geepers} and \psw, the principal score model was fitted to data from the treatment group using standard maximum likelihood methods.
 For \textsc{geepers}, we used the estimated principal scores in the control group, along with the observed $S$ in the treatment group, to construct $R$, and then fit the model \eqref{eq:reg2} with OLS.
@@
 
-Figure \ref{fig:alphan} shows the empirical bias and standard error (SE) of the \textsc{geepers} and mixture estimator of $\eff1$ as sample size and $\alpha$ varied.
+Figure \ref{fig:alphan} shows the empirical bias and standard error (SE) of the \textsc{geepers} and \pmm estimators of $\eff1$ as sample size and $\alpha$ varied.
 Across $n$ and $\alpha$, \textsc{geepers} had lower bias and higher SE than the mixture estimator.
@@
 Lastly, Assumption \ref{ass:ci} implies that there is no interaction between $\bxs$ and $\st$ in the outcome model. 
-Further, for the sake of statistical efficiency, we did not consider outcome models with interactions between covariates and $Z$.
 \emph{A priori}, the student attribute that poses the greatest threat to these modeling choices is prior achievement, that is, scores on the pretest and the fifth-grade state test.
@@
 
-Outcome models for all analyses included for all of the variables that were in the corresponding principal score models, in addition to measurements of students' prior knowledge; also, the models substituted class fixed effects for school fixed effects.\footnote{Why use a coarser set of fixed effects for the principal score model than the outcome regression?
+Outcome models for all analyses included all variables from the corresponding principal score models, but substituted class fixed effects for school fixed effects.\footnote{Why use a coarser set of fixed effects for the principal score model than the outcome regression?
 First, the principal score model was fit to a smaller sample (only students in the ``Instant'' condition), so some classrooms only included one student, which can lead to poor fit.

### 2025-07-01 code/OPTS/opts.r
better figures

@@ -1,31 +1,7 @@
-library(rstan)
-library(dplyr)
-library(tibble)
-library(ggplot2)
-library(sandwich)
-library(lmtest)
-library(tableone)
-library(purrr)
-library(kableExtra)
-library(tidyr)
-
-rstan_options(auto_write = FALSE)
 
 
-source('code/regression.r')
-options(mc.cores = 2)
 
-auc <- function(x,y)
-  if(length(unique(y))==2 & length(y)==length(x)){
-    wilcox.test(x~y)$statistic/(sum(y)*(sum(1-y)))
-  }else
-    wilcox.test(x,y)$statistic/(legnth(x)*length(y))
-
-aucMod <- function(mod)
-  auc(mod$linear,1-mod$y)
-
-
-print(load('code/OPTS/OPT_Study_PersonLevel_Data.RData'))
+print(load('data/OPT_Study_PersonLevel_Data.RData'))
 ## downloaded from https://www.causeweb.org/tshs/obstetrics-and-periodontal-therapy/
 
 
@@ -65,10 +41,10 @@ tab1 <- map(list(`Full Data`=opt,`Complete Cases`=cca),
     t())%>%
     do.call("cbind",.)
 
-
+sink("results/optsTable1.tex")
 kbl(tab1[-c(1:2),],format="latex",booktabs=TRUE,col.names=tab1[2,],caption="Descriptive statistics---mean and standard deviation or count and percent---for study variables in the full OPT dataset and in the analysis sample (i.e. complete cases)",label="optTab1" )%>%
     add_header_above(c(" ","Full Data"=2,"Complete Cases"=2))
-
+sink()
 
 
 ## "included 640 participants with nonmissing values for the covariates and
@@ -142,7 +118,7 @@ sdat <- list(
     sdat$ncovY=ncol(Xout)
 
 
-  psStan <- stan(file = 'code/fh2t/psModSimp.stan',data=sdat)
+psStan <- stan(file = 'code/fh2t/psModSimp.stan',data=sdat)
 save(psStan,file='code/OPTS/psStanSimp.RData')
 load('code/OPTS/psStanSimp.RData')
 
@@ -205,9 +181,12 @@ effects%>%pivot_wider(id_cols=eff,names_from=method,values_from=c(estimates,SE))
 effects%>%pivot_wider(id_cols=method,names_from=eff,values_from=c(estimates,SE))%>%
[…]

### 2025-07-01 code/fh2t/plots.r

@@ -51,21 +51,24 @@ bind_rows(
     ymin=estimates-2*SE,
     ymax=estimates+2*SE
   ) %>%
-  ggplot(aes(EFF,estimates,color=PSmodel,ymin=ymin,ymax=ymax))+
-  geom_point()+geom_errorbar(width=0)+
-  geom_point(data=ateDF,aes(EFF,estimates),color='black',inherit.aes=FALSE)+
-  geom_errorbar(data=ateDF,aes(EFF,ymin=ymin,ymax=ymax),width=0,color='black',inherit.aes=FALSE)+
-  geom_hline(yintercept=0)+
-  scale_x_continuous("Principal Stratum",breaks=c(0,.25,.5),minor_breaks=NULL,
-                     labels=c("Non-\nBottom-\nOuter","ATE","Bottom-\nOuter"),limits=c(-.15,.65))+
-  labs(color='PS Model',y='Principal Effect')+
-  facet_wrap(~Alternative,nrow=1)+
-  theme(legend.position="top")
+    ggplot(aes(EFF,estimates,color=PSmodel,ymin=ymin,ymax=ymax))+
+    geom_point()+
+    geom_errorbar(width=0,linewidth=1)+
+    geom_point(data=ateDF,aes(EFF,estimates),color='black',inherit.aes=FALSE)+
+    geom_errorbar(data=ateDF,aes(EFF,ymin=ymin,ymax=ymax),width=0,color='black',inherit.aes=FALSE)+
+    geom_hline(yintercept=0)+
+    scale_x_continuous("Principal Stratum",breaks=c(0,.25,.5),minor_breaks=NULL,
+                       labels=c("Non-\nBottom-\nOuter","ATE","Bottom-\nOuter"),limits=c(-.15,.65))+
+    scale_color_manual(values=c('#1b9e77','#d95f02','#7570b3'))+
+    scale_y_continuous(minor_breaks=c(ateDF$estimates,-0.75,0.75))+
+    labs(color='PS Model',y='Principal Effect')+
+    facet_wrap(~Alternative,nrow=1)+
+    theme(legend.position="top")
 ggsave("figure/prinEffs.pdf",width=5,height=3,units="in")
 
 #################
 ### comparing GEEPERS to other methods
-bind_rows(
+compare <- bind_rows(
   imap_dfr(estimates1all,~effsFromFit(.x)%>%
                         data.frame()%>%
                         rownames_to_column("eff")%>%
[…]

### 2025-07-01 code/regression.r

@@ -311,6 +311,12 @@ psw1 <- function(dat,psMod){
 
 print.psw <- function(x, ...) print(x$coef,...)
 
+bsSchool <- function(dat){
+  datS <- split(dat,dat$SchIDPre)
+  do.call("rbind",
+          lapply(datS,function(x) x[sample(1:nrow(x),nrow(x),replace=TRUE),]))
+}
+
 bsInd=function(data)
   data[sample(1:nrow(data),nrow(data),replace=TRUE),]
 

### 2025-07-01 code/replicate.r

@@ -3,6 +3,8 @@ library(dplyr)
 library(ggplot2)
 library(readr)
 library(tidyr)
+library(sandwich)
+library(lmtest)
 library(forcats)
 library(arm)
 library(randomForest)
@@ -19,14 +21,23 @@ library(tibble)
 library(coefplot)
 library(splines)
 library(tikzDevice)
-library(colorspace)
 select <- dplyr::select
 
-palette <- qualitative_hcl(4)
-names(palette)=c("\\textsc{pmm}","\\textsc{geepers}","\\textsc{psw}","\\textsc{bsiv}")
+
+options(mc.cores = 2)
+
+### NYC subway colors
+### from https://data.ny.gov/widgets/3uhz-sej2
+palette <- c(
+"#0039A6",
+"#00ADD0",
+"#EE352E",
+"#808183"#,
+)#%>%protan()
+#qualitative_hcl(4)
+names(palette)=c("\\textsc{pmm}","\\textsc{psw}","\\textsc{geepers}","\\textsc{bsiv}")
 
 source('code/regression.r')
-source('code/fh2t/psw.r')
 
 auc <- function(x,y)
   if(length(unique(y))==2 & length(y)==length(x)){
@@ -35,7 +46,30 @@ auc <- function(x,y)
     wilcox.test(x,y)$statistic/(legnth(x)*length(y))
 
 aucMod <- function(mod)
-  auc(mod$linear,1-mod$y)
+    auc(mod$linear,1-mod$y)
+
+
+
+############################################################
+############################################################
+###
+### OPTS analysis
+###
+#########################################################
+############################################################
+
+source("code/OPTS/opts.r")
+
+
+
+############################################################
+############################################################
+###
+### FH2T analysis
+###
+#########################################################
+############################################################
+
 
 ############################################################
 ### make dataset

### 2025-07-01 code/simulation/displaySimPublish.r

@@ -64,7 +64,16 @@ load('simResults/resultsB1s_mu01is0.RData')
     interactionS=ifelse(intS,"S interaction","No\nS interaction"),
     N=paste0("n=",n)
   ) %>%
-  filter(rhat<1.1)
+     filter(rhat<1.1)
+
+
+### any statistically significant bias?
+pdns%>%filter(eff=="1")%>%
+    group_by(estimator,n)%>%
+    summarize(bias=mean(errP),pval=t.test(errP)$p.value)%>%
+    mutate(p.adj=p.adjust(pval,method="BH"),
+           p.holm=p.adjust(pval,method="holm"))%>%
+    print(n=Inf)
 
 
 
@@ -130,6 +139,13 @@ ggsave("simFigs/biasSEbyN.jpg",plot=plotByN,width=6,height=3)
   ) %>%
   filter(rhat<1.1)
 
+### any statistically significant bias?
+pdb1%>%filter(eff=="1",rhat<1.1,estimator!="PSW")%>%
+    group_by(estimator,b1)%>%
+    summarize(bias=mean(errP),pval=round(t.test(errP)$p.value,4))%>%
+    mutate(p.adj=p.adjust(pval,method="BH"),
+           p.holm=p.adjust(pval,method="holm"))%>%
+    print(n=Inf)
 
 
 

### 2025-07-01 code/simulation/runSim.r

@@ -65,7 +65,7 @@ fullsim(nrep,ns=c(200,300,400,600,700,800),errDist='norm',b1s=.5,intS=FALSE,intZ
 fullsim(nrep,ns=500,errDist='norm',b1s=c(.1,.3,.4,seq(.6,1,.1)),intS=FALSE,intZ=FALSE,mu01=0,mu10=0,ext='b1ss_mu01is0',cl=cl,
             ncores=ncore)
 
-fullsim(nrep,ns=500,errDist=c("norm","unif"),b1s=c(0.3,0.4),intZ=TRUE,mu10=0,mu01=0.3,ext="alpha34")
+fullsim(nrep,ns=500,errDist=c("norm","unif"),b1s=c(0.3,0.4),intZ=TRUE,mu10=0,mu01=0,ext="alpha34")
 
 
 ##########################################################################################

### 2025-07-01 writeUps/paper.tex

@@
 
-\begin{ass}[Residualized Covariate Independence]\label{ass:rci}
-There exists $\bxy$, a (known) transformation of covariates $\bx$, and an (unknown) vector of coefficients $\bm{\gamma}$ such that
+\begin{assp}{\ref*{ass:ci}$'$}[Residualized Covariate Independence]\label{ass:rci}
+There exists $\bxy$, a (known) multivariate transformation of covariates $\bx$, and an (unknown) vector of coefficients $\bm{\gamma}$ such that
 \begin{equation}\label{eq:ass:rci}
@@
 \end{equation}
-\end{ass}
+\end{assp}
 That is, CI may not hold for potential outcomes themselves, but an analogous assumption holds for \emph{residualized} potential outcomes---i.e., $\yc$ and $\yt$ after subtracting out a linear function of (possibly transformed) covariates.
@@
 
-We conducted a simulation study to compare the performance of \textsc{geepers} with that of a Bayesian mixture model and a principal score weighting estimator, in both favorable and unfavorable circumstances.
+We conducted a simulation study to compare the performance of \textsc{geepers} with that of a parametric mixture model (\pmm) and a principal score weighting (\psw) estimator, in both favorable and unfavorable circumstances.
 The study was designed to answer three sets of overarching questions.
@@
 and when important interactions are omitted from the outcome model?
-Second, how does \textsc{geepers} compare with mixture modeling?
+Second, how does \textsc{geepers} compare with parametric mixture modeling?
 Is it competitive in circumstances favorable to both techniques?
-Does it avoid the pitfalls of mixture modeling when the parametric assumptions of the normal mixture model fail?
-Third, are there ways or circumstances when the principal score weighting estimator outperforms \textsc{geepers} or a mixture estimator even though principal ignorability does not hold?
+Does it avoid the pitfalls of \pmm when the parametric assumptions of the normal mixture model fail?
+Third, are there ways or circumstances when \psw outperforms \textsc{geepers} or \pmm even though principal ignorability does not hold?
 
 \subsection{Simulation Design}
-The simulation study was conducted in \texttt{R} \citep{rcite} and Stan \citep{rstan}, and full replication code is available at %[Redacted]. %
-\url{https://github.com/adamSales/psGee}.
+The simulation study was conducted in \texttt{R} \citep{rcite} and Stan \citep{rstan}, and full replication code is available at [Redacted]. %
+%\url{}.
 We replicated each set of conditions 500 times.
@@
 where $\alpha$ is a manipulated factor.
-When $\alpha$ was higher, $\st$ was more easily predicted by covariates; when $\alpha=0$, Assumption \ref{ass:vps} was violated, i.e., $Var(\pp)=0$.
 Principal stratum $\sti\in \{0,1\}$ was simulated as $S_i\sim Bern(\ppi)$.
@@
 \subsubsection{Manipulated Factors}\label{sec:manipulatedFactors}
-We manipulated five factors in the simulation, listed in Table \ref{tab:factor}.
-The five factors are not completely crossed---in particular, we let $n$ and $\alpha$ each vary across all the levels listed while holding the other four factors fixed.
+We manipulated four factors in the simulation, listed in Table \ref{tab:factor}.
+The factors were completely crossed, except for some levels of $n$ and $\alpha$, which were tested with other factors held fixed.
+
+% \begin{table}
+%     \caption{\label{tab:factor} Manipulated factors in simulation study}
+%   \centering
+%   %\rowcolors{1}{}{lightgray}
+% \begin{tabular}{r|ll}%{*{3}{c}}
+%   \hline
+%   Factor &\multicolumn{2}{l}{Levels}\\
+%   & Part 1 & Part 2\\
+%   \hline
+% Sample size per condition $n$ &$100,200,\dots,1000$ & 500, 1000\\
+% $\alpha$ & $0,0.2,0.3,\dots,1$ & 0, 0.2, 0.5 \\
+% The distribution of $\epsilon_i$& Normal & Normal, Uniform\\
+% %$\beta_1$ & \multicolumn{2}{l}{0, 0.3}\\
+% \multirow{8}{*}{Interactions in \eqref{eq:y-sim}} &  \multirow{8}{1.5in}{None $\gamma_1=1/\sqrt{6};\gamma_2=\gamma_3=0$}& None\\
+% & &$\gamma_1=1/\sqrt{6};\gamma_2=\gamma_3=0$\\ 
+% & & $\bm{x}:\st$\\
+% && $\gamma_1=3/(4\sqrt{6})$; $\gamma_2=1/(2\sqrt{6})$; $\gamma_3=0$\\
+% & & $\bm{x}:Z$\\
+% &&$\gamma_1=1/\sqrt{6};\gamma_2=0$; $\gamma_3=1/(2\sqrt{6})$\\
+% && $\bm{x}:\st$ \& $\bm{x}:Z$\\
+% & & $\gamma_1=3/(4\sqrt{6})$; $\gamma_2=\gamma_3=1/(2\sqrt{6})$\\
+% % between $\st$ and covariates & No, Yes\\
+% % Interactions between $Z$ and covariates& No, Yes\\
+% \hline
+% \end{tabular}
+% \end{table}
 
@@
   %\rowcolors{1}{}{lightgray}
-\begin{tabular}{r|rl}%{*{3}{c}}
+\begin{tabular}{r|ll}%{*{3}{c}}
   \hline
@@
   \hline
-Sample size per condition $n$ &\multicolumn{2}{l}{$100,200,\dots,900,1000$}\\
-$\alpha$ &\multicolumn{2}{l}{$0,0.2,0.3,\dots,0.9,1$}\\
+Sample size per condition $n$ &\multicolumn{2}{l}{$100^*,200^*,300^*,400^*,500,600^*,700^*,800^*,900^*,1000$}\\
+$\alpha$ &\multicolumn{2}{l}{$0,0.2,0.3^\dagger,0.4^\dagger,0.5,0.6^\dagger,0.7^\dagger,0.8^\dagger,0.9^\dagger,1^\dagger$}\\
 The distribution of $\epsilon_i$&\multicolumn{2}{l}{ Normal, Uniform}\\
 %$\beta_1$ & \multicolumn{2}{l}{0, 0.3}\\
-\multirow{4}{*}{Interactions in \eqref{eq:y-sim}} & None & $\gamma_1=1/\sqrt{6};\gamma_2=\gamma_3=0$\\ 
-& $\bm{x}:\st$ & $\gamma_1=3/(4\sqrt{6})$; $\gamma_2=1/(2\sqrt{6})$; $\gamma_3=0$\\
-& $\bm{x}:Z$&$\gamma_1=1/\sqrt{6};\gamma_2=0$; $\gamma_3=1/(2\sqrt{6})$\\
-& $\bm{x}:\st$ \& $\bm{x}:Z$& $\gamma_1=3/(4\sqrt{6})$; $\gamma_2=\gamma_3=1/(2\sqrt{6})$\\
+\multirow{4}{*}{Interactions in \eqref{eq:y-sim}}& None:  $\gamma_1=1/\sqrt{6};\gamma_2=\gamma_3=0$\\ 
+& $\bm{x}:\st$: $\gamma_1=3/(4\sqrt{6})$; $\gamma_2=1/(2\sqrt{6})$; $\gamma_3=0$\\
+& $\bm{x}:Z$: $\gamma_1=1/\sqrt{6};\gamma_2=0$; $\gamma_3=1/(2\sqrt{6})$\\
+& $\bm{x}:\st$ \& $\bm{x}:Z$: $\gamma_1=3/(4\sqrt{6})$; $\gamma_2=\gamma_3=1/(2\sqrt{6})$\\
 % between $\st$ and covariates & No, Yes\\
@@
 \hline
+\multicolumn{3}{p{4.5in}}{\footnotesize Factors fully crossed except where indicated.}\\
+\multicolumn{3}{p{4.5in}}{\footnotesize $*$ Only tested with $\alpha=0.5$, normal $\epsilon$, no interactions in \eqref{eq:y-sim}.}\\
+\multicolumn{3}{p{4.5in}}{\footnotesize $\dagger$ Only tested with $n=500$, normal $\epsilon$, no interactions in \eqref{eq:y-sim}.}
+
 \end{tabular}
@@
 
-$\bm{\alpha}$\textbf{:} As mentioned above, $\alpha$ controls the variance of the principal scores, i.e., the extent to which $\st$ is predictable as a function of covariates.
-When $\alpha=0$, the covariates are unrelated to $\st$ and principal scores do not vary between participants, violating \eqref{ass:vps}.
-We expect that the more predictive the covariates, the better the performance of all three estimators.
+$\bm{\alpha}$\textbf{:} $\alpha$ controlled the extent to which $\st$ is predictable as a function of covariates.
+When $\alpha$ was higher, $\st$ was more easily predicted by covariates; when $\alpha=0$, Assumption \ref{ass:vps} was violated, i.e., $Var(\pp)=0$.
+We expected that the more predictive the covariates, the better the performance of all three estimators.
 To aid in the interpretation of the $\alpha$ parameter, the online appendix plots $\alpha$ against the area under the receiver operating characteristic curve for each fitted model (AUC), a common measure of classification accuracy \citep{bradley1997use}---an AUC of 0.5 implies that the model is no better than random guessing, and AUC of 1 implies perfect prediction.
@@
 However, the shape of $\epsilon$'s distribution varied between the following two possibilities:
-Normal $\epsilon\sim\mathcal{N}(0,1/2)$, as assumed by the mixture model estimator, or Uniform $\epsilon\sim\mathcal{U}(-\sqrt{6}/2,\sqrt{6}/2)$.
+Normal $\epsilon\sim\mathcal{N}(0,1/2)$, as assumed by \pmm, or Uniform $\epsilon\sim\mathcal{U}(-\sqrt{6}/2,\sqrt{6}/2)$.
 %\item[Mixture] $f_\epsilon(x)=\frac{1}{4}\phi\left((x+1/3)\sqrt{6}\right)+\frac{1}{4}\phi\left((x-1)\sqrt{6})\right)$
@@
 
-\textbf{\emph{Interactions between $\st$ or $Z$ and covariates:}} The factor---the presence of interactions between covariates and $\st$ and $Z$, respectively--- tests the performance of a model that does not include interactions when that modeling assumption is false.
+\textbf{\emph{Interactions between $\st$ or $Z$ and covariates:}} This factor was designed to test the vulnerability of the estimators to certain violations or Assumption \ref{ass:rci}.
 
@@
 
-Using each simulated dataset, we estimated $\eff0$ and $\eff1$ with three methods: \geepers, a parametric mixture model \eqref{eq:mixtureModelPS} including principal scores---hereafter \pmm---and the principal score weighting estimator \eqref{eq:psw}, hereafter \psw.
-
-The methods all used the same ``observed'' dataset, consisting of outcomes $Y$, treatment assignments $Z$, two covariates $\bxs=\bxy=\bx=[x_1,x_2]$, and $S$ observed only when $Z=1$.
-The third covariate used in the data-generating model, $x_3$, was ``unobserved''--since $x_3$ was correlated with both $\st$ and $Y$, Principal Ignorability \eqref{eq:pi} was violated.
+Using each simulated dataset, we estimated $\eff0$ and $\eff1$ with three methods: \geepers, \pmm with principal scores \eqref{eq:mixtureModelPS}, and the \psw estimator \eqref{eq:psw}.
+
+The methods all used the same ``observed'' dataset, consisting of outcomes $Y$, treatment assignments $Z$, two covariates $\bx=[x_1,x_2]$, and $S$ observed only when $Z=1$.
+The third covariate used in the data-generating model, $x_3$, was ``unobserved''---since $x_3$ was correlated with both $\st$ and $Y$, Principal Ignorability \eqref{eq:pi} was violated.
 
@@
 \pmm, the parametric mixture estimator, then fits two outcome models:
-For the treatment group, the regression model $Y=\beta_0^T+\beta_1^TS+\bm{\gamma}'\bx+\epsilon$ and for the control group, a mixture model of the form \eqref{eq:mixtureModelPS}, but with linear normal models for $g^s(y|x_1,x_2)$.
+For the treatment group, the regression model $Y=\beta_0^T+\beta_1^TS+\bm{\gamma}'\bx+\epsilon$ and for the control group, a mixture model of the form \eqref{eq:mixtureModelPS} with linear normal models for $g^s(y|x_1,x_2)$.
 Residual variance was allowed to differ across treatment groups, but not across principal strata.
@@
 We also report the empirical coverage of confidence intervals, the proportion of replications in which $\eff1$ is within the estimated confidence bounds of $\heff1$.
-Here, we present results on error, bias, empirical standard error, and confidence interval coverage; results on root-mean-squared-error %$\left(\widehat{bias}^2+SE_{emp}^2\right)^{1/2}$
+Results on root-mean-squared-error %$\left(\widehat{bias}^2+SE_{emp}^2\right)^{1/2}$
 $\left(\sum_b err_b^2/500\right)^{1/2}$
-are available in the online appendix.
-
-\subsubsection{Varying $n$ and $\alpha$}
+are in the online appendix.
+
+\subsubsection{Performance with Varying $n$ and $\alpha$}
 % When $\alpha=0$, Assumption \ref{ass:vps} was violated, and \textsc{geepers} estimates were highly erratic; however, as Table \ref{tab:coverage} shows, when other assumptions held, inference remained valid.
-Our first set of results illustrates the performance of \geepers when all of its assumptions are met.
-Since \geepers is an asymptotic method whose identification depends on $\bm{x}$ as a predictor of $\st$, we examine its bias and standard error as functions of $n$ and $\alpha$.
-We include analogous results for \pmm as a baseline comparison.
+To what extent do the performances of \geepers and \psw depend on $n$ and $\alpha$?
+For this set of results, $\epsilon\sim \mathcal{N}(0,1/2)$ and there were no interactions in \eqref{eq:y-sim}; $\alpha$ was fixed at $0.5$ for varying levels of $n$, and $n$ was fixed at 500 for varying levels of $\alpha$.
 
 Figure \ref{fig:alphan} shows the empirical bias and standard error (SE) of the \textsc{geepers} and \pmm estimators of $\eff1$ as sample size and $\alpha$ varied.
-Across $n$ and $\alpha$, \textsc{geepers} had lower bias and higher SE than the mixture estimator.
-\textsc{geepers} was positively biased for low $n$ with bias converging to close to 0 (up to simulation error) for $n\ge 300$, and was roughly unbiased (up to simulation error) for $\alpha \ge 0.2$, with $n$ fixed at 500.
-The SE of the \textsc{geepers} decreased rapidly for $100\le n \le 500$ and $0.2\le \alpha \le 0.5$ and more gradually for higher values of $n$ and $\alpha$.
-
-For the remainder of the simulation, we focus on the moderate cases of $n=500$ and $\alpha=0.2$ or 0.5.
+Across $n$ and $\alpha$, \textsc{geepers} was approximately unbiased,\footnote{All p-values testing for \geepers bias were $>0.1$, with the exception of when $\alpha=1$, with $\widehat{bias}=0.01$, $p=0.046$ before adjustment for multiple comparisons.} while $\pmm$ estimators exhibited slight positive bias that decreased with $n$.
+
+\geepers had a higher SE than \pmm, particularly when $n=100$ or $\alpha=0.2$.
+The \psw advantage is far less dramatic for higher levels of $n$ and $\alpha$, eventually diminishing to approximately zero.
 
@@
 %<<<<<<< HEAD
-Figure \ref{fig:boxplots} shows violin plots, layered on top of jittered scatter plots, showing estimation error for \geepers, the parametric mixture model \pmm, and the \psw estimator of $\eff1$.
-Table \ref{tab:coverage} shows empirical coverage of nominal 95\% confidence interval estimates for \geepers and \pmm; when at least one of an estimator's assumptions is violated, the result is colored red.
-
-The first column of Figure \ref{fig:boxplots} and the black entries of Table \ref{tab:coverage} show results from scenarios in which all assumptions of \geepers and \pmm are met (\psw relies on principal ignorability, which is violated throughout the simulation study).
-\textsc{geepers} is slightly less biased than the parametric mixture model; when $\alpha=0.2$, \textsc{geepers} is much more variable and $\alpha=0.5$ the difference in variance is slight.
-\textsc{psw} has by far the lowest variance and the highest bias; the latter is due to the violation of principal ignorability. 
-When all assumptions are met, confidence intervals from both \geepers and \pmm achieve their nominal levels.
-
-The remainder of the results reflect violations modeling assumptions for either \geepers or the parametric mixture model or both: when $\alpha=0$, violating Assumption \ref{ass:vps}; interactions in the data generating model between covariates $X$ and treatment assignment $Z$ and/or principal stratum $S$, either of which would violate assumption \ref{ass:rci} and the linear additive model of the \textsc{pmm}; and/or residuals drawn from a uniform distribution, which violates the normality assumption of \textsc{pmm}.
-
-We omitted the $\alpha=0$ case from Figures \ref{fig:alphan} and \ref{fig:boxplots} since the variance of \geepers and \pmm estimates was so large as to dominate the plot. 
-However, the first column of results in Table \ref{tab:coverage} shows that, when other assumptions are met, confidence intervals from both methods tend to over-cover when $\alpha=0$. 
+Figure \ref{fig:boxplots} shows violin plots, layered on top of jittered scatter plots, showing estimation errors for \geepers, \pmm, and \psw.
+Table \ref{tab:coverage} shows empirical coverage rates of nominal 95\% confidence interval estimates for \geepers and \pmm; when at least one of an estimator's assumptions is violated, the result is colored red.
+
+When all of the \geepers and \pmm assumptions were met, that is, when $\epsilon\sim\mathcal{N}(0,1/2)$, $\alpha>0$, and there were no covariate interactions in \eqref{eq:y-sim}, results for $\geepers$ and $\pmm$ were the same as in the previous section; in addition, confidence intervals from both methods achieved their nominal levels.
+\textsc{psw} had by far the lowest variance and the highest bias of the three estimators; the latter is due to the violation of principal ignorability.
+
+%The remainder of the results reflect violations modeling assumptions for either \geepers or the parametric mixture model or both: when $\alpha=0$, violating Assumption \ref{ass:vps}; interactions in the data generating model between covariates $X$ and treatment assignment $Z$ and/or principal stratum $S$, either of which would violate assumption \ref{ass:rci} and the linear additive model of the \textsc{pmm}; and/or residuals drawn from a uniform distribution, which violates the normality assumption of \textsc{pmm}.
+
+%We omitted the $\alpha=0$ case from Figures \ref{fig:alphan} and \ref{fig:boxplots} since the variance of \geepers and \pmm estimates was so large as to dominate the plot. 
+When $\alpha=0$, Assumption \ref{ass:vps} is violated; %, standard errors for \geepers were too high to 
+Nevertheless, as the first column of results in Table \ref{tab:coverage} shows, when other assumptions were met, confidence intervals from both \geepers and \psw tended to over-cover when $\alpha=0$. 
 In a randomized experiment, $\alpha$ can be estimated directly, making Assumption \ref{ass:vps} testable.
 
-For \geepers, when $\alpha=0.2$ the presence of interactions in the data-generating model increases sampling variance, but only interactions between $X$ and $Z$ reduce confidence interval coverage.
-When $\alpha=0.5$, the presence of interactions has little to no effect on the performance of the \geepers estimator, and all 95\% confidence intervals approximately achieve their nominal levels.
-\geepers makes no assumptions about the shape of the residual distribution and behaves similarly under either normal or uniform residuals.
+The impact of covariate interactions in \eqref{eq:y-sim} on the performance of \geepers depended on $\alpha$: when $\alpha=0.2$, the presence of interactions increased sampling variance.
+Still, only interactions between $X$ and $Z$ reduced confidence interval coverage.
+When $\alpha=0.5$, interactions had little to no effect on the \geepers estimator, and all 95\% confidence intervals approximately achieved their nominal levels.
+\geepers performed similarly under either normal or uniform residuals.
 
 Like \geepers, \pmm is insensitive to interactions between $X$ and $S$.
-However, interactions between $X$ and $Z$ in the data-generating model skew the \pmm sampling distribution and reduce the coverage of its 95\% confidence intervals, even when $\alpha=0.5$.
-When the residual distribution was uniformly distributed, the \pmm sampling distribution appeared bimodal, and its credible intervals exhibited severe undercoverage.
+However, interactions between $X$ and $Z$ in the data-generating model skewed its sampling distribution and reduced the coverage of its 95\% confidence intervals for all values of $\alpha$ in the simulation.
+Uniformly-distributed residuals caused bimodality in the \pmm sampling distribution, and caused its credible intervals to exhibit severe undercoverage.
 
@@
 These results suggest that researchers who are willing to assume principal ignorability should use \psw, and those who are confident in their parametric model for potential outcomes should use a parametric mixture model.
-Researchers who are hesitant to accept principal ignorability or a parametric outcome model should consider \geepers.
-
-\label{evalMetric}The simulation study suggests that none of the three methods evaluated---\geepers, \pmm, or \psw---dominates the others, and offers some guidance on when each estimator may be preferable.
+Researchers who are hesitant to accept principal ignorability or a parametric outcome model should consider \geepers, particularly when $n\ge 200$ and covariates can predict $S$ with AUC$>0.6$ (corresponding to $\alpha>0.2$).
+These suggestions may depend heavily on the simulation setup; applying them in general scenarios requires further research.
+
+%\label{evalMetric}The simulation study suggests that none of the three methods evaluated---\geepers, \pmm, or \psw---dominates the others, and offers some guidance on when each estimator may be preferable.
 \psw is the most precise and requires no assumptions about the distribution of outcomes, but can be quite sensitive to violations of principal ignorability.
@@
 \geepers may be the best choice when conditional normality may not hold, or if there are $X:Z$ interactions and $\bm{x}$ is sufficiently predictive of $\st$ (AUC $\ge$0.675, corresponding to $\alpha \ge 0.5$, may serve as a threshold).
-These suggestions may depend heavily on the simulation setup; applying them in general scenarios requires further research.
 
@@
 Periodontal disease, a weakening of bone and connective tissue supporting teeth, is highly prevalent among American adults and is associated with adverse pregnancy outcomes, including preterm birth \citep{periodontalEpi}.
-The Obstetrics and Periodontal Therapy Study \citep[(OPTS)][]{michalowicz2006treatment} investigated the efficacy of nonsurgical periodontal treatment to reduce the preterm birth rate; replication data are available at \url{https://causeweb.org/tshs/obstetrics-and-periodontal-therapy/}.
+The Obstetrics and Periodontal Therapy Study \citep[OPTS; ][]{michalowicz2006treatment} investigated the efficacy of nonsurgical periodontal treatment to reduce the preterm birth rate; replication data are available at \url{https://causeweb.org/tshs/obstetrics-and-periodontal-therapy/}.
 Pregnant patients at four different hospitals who were at least 16 years old, at less than 16 weeks and 6 days of gestation, and diagnosed with periodontal disease were individually randomized to either receive periodontal treatment during the first 21 weeks of pregnancy ($Z=1$, $n_T=413$) or after delivery ($Z=0$, $n_C=410$).
-The primary prespecified outcome of the study was the babies' gestational age at birth; however, the study team also collected data on several secondary birth outcomes, as well as measures of patients' periodontal disease at follow-up.
-
 The main component of the intervention consisted of up to four visits, followed by monthly tooth polishing and other dental care as needed.
-However, not all participants randomized to the treatment condition completed the entire treatment plan.
-However, 96\% of patients in the treatment group received at least some periodontal care during pregnancy---therefore, there is reason to suspect nonzero treatment effects even for those patients who did not (or would not) fully comply.
+Not all participants randomized to the treatment condition completed the entire treatment plan; however, 96\% of them received at least some periodontal care during pregnancy.
 Patients randomized to the control condition did not have access to the intervention.
-\citet{richardson2023estimating} used a method they refer to as ``bespoke instrumental variable'' (BSIV) to estimate the average effects for the principal stratum of subjects who, if randomized to treatment,  would complete the treatment plan and for the principal stratum of subjects who would not.
-Here, we will use \textsc{geepers}, a Bayesian mixture model, and principal score weighting to estimate the same quantities, following the same analysis decisions as \citet{richardson2023estimating}.
-Replication code for our analysis can be found at \url{ https://github.com/adamSales/psGee}.
-
-Following \citep{richardson2023estimating}, we will estimate the principal effects of randomization $Z$ on $Y$, the percentage of probed sites (6 sites per tooth) that were bleeding at visit 5, which was at 29--32 weeks gestation, conditioning on two covariates, Serum measures of endotoxin level and fibrinogen at baseline.
-We omit the 164 cases in which $Y$ was missing as well as 19 additional cases in which one or both of the covariates were missing, leaving a total sample size of 640, with 314 randomized to the treatment condition and 326 randomized to control.
-The compliance variable in the dataset had three categories: ``Yes,'' for subjects in the treatment group who completed the treatment plan (157 subjects), ``No'' for those who did not (4 subjects), and ``Und'' for those who received some therapy, but it is unknown whether they completed the treatment (157 subjects).
-We let $S=1$ when compliance was ``yes''---i.e., when subjects in the treatment group completed the treatment---and $S=0$ otherwise.
+
+Following \citet{richardson2023estimating}, let $\st=1$ for patients who, if randomized to treatment, would complete the full treatment plan, and let $\st=0$ for those who would not.
+Let $Y$, the outcome of interest, be the percentage of probed sites (6 sites per tooth) that were bleeding at visit 5, which was at 29--32 weeks of gestation, with potential values $\yc$ and $\yt$, as above.
+\citet{richardson2023estimating} used a principal stratification estimator they called ``bespoke instrumental variables'' (BSIV) to estimate principal effects $\eff0$ and $\eff1$, the average effects of randomization to the treatment condition for patients who would or would not complete the full treatment plan.
+In this section, we compare their result to estimates using \geepers, \pmm, and \psw.
+
+For simplicity, we followed the same analysis decisions as \citet{richardson2023estimating}. 
+Specifically, we conditioned on two covariates, Serum measures of endotoxin level and fibrinogen at baseline, and
+omitted the 164 cases in which $Y$ was missing as well as 19 additional cases in which one or both of the covariates were missing.
+That left a total sample size of 640, with 314 randomized to the treatment condition and 326 randomized to control.
+%The compliance variable in the dataset had three categories: ``Yes,'' for subjects in the treatment group who completed the treatment plan (157 subjects), ``No'' for those who did not (4 subjects), and ``Und'' for those who received some therapy, but it is unknown whether they completed the treatment (157 subjects).
+%We let $S=1$ when compliance was ``yes''---i.e., when subjects in the treatment group completed the treatment---and $S=0$ otherwise.
 
@@
 
-The ability to bottom out was only one of several differences between the treatment conditions.
-One way to disentangle the role of bottoming out in any treatment effect would be to estimate average treatment effects of being assigned to ASSISTments, relative to any of the other three, separately for subjects who would (if assigned to ASSISTments) bottom out frequently, and for subjects who would not.
+The option to bottom out was only one of several differences between the treatment conditions.
+One way to disentangle the role of bottoming out in any treatment effect would be to estimate the average treatment effects of being assigned to ASSISTments, relative to the other three, separately for subjects who would (if assigned to ASSISTments) bottom out frequently, and for subjects who would not.
 If the availability of bottom-out hints plays a positive role in ASSISTments' efficacy, we might expect larger effects of the condition on students who frequently bottom out than on those who do not; if bottom-out hints are harmful, we might expect the opposite.
@@
 
-Principal score weighting depends on the assumption of principal ignorability, which is equivalent to the no-confounding assumption of the observational study and mediation analysis.
-The mixture model and \textsc{geepers} estimators do not rely on that assumption, instead relying on parametric models and residualized covariate independence, respectively.
-Finally, the instrumental variables estimator of \citet{air} relies on the ``exclusion restriction'' that, for students who request few bottom-out hints, there is no effect of randomization to the ASSISTments condition.
+Principal score weighting depends on the assumption of principal ignorability, which is equivalent to the no-confounding assumption in observational studies and mediation analysis.
+The mixture model and \textsc{geepers} estimators do not rely on this assumption; instead, they rely on parametric models and residualized covariate independence, respectively.
+A final option is the instrumental variables estimator of \citet{air}, which relies on the ``exclusion restriction'' that randomization did not affect students who would request few bottom-out hints.
 Since bottom-out hints are only one of several features of ASSISTments, this assumption is untenable.
@@
 For a description of the dataset and an explanation of how to access it, see \citet{ottmar2023data} and \url{https://osf.io/r3nf2/}.
-Replication code for the analysis is available at %[Redacted].
-\url{https://github.com/adamSales/psGee}.
+Replication code for the analysis is available at [Redacted].
+%\url{https://github.com/adamSales/psGee}.
 
@@
 We estimated principal scores using logistic regression of observed $S$ on school fixed effects and a set of baseline variables $\bxs$.
-Three criteria governed our choice of covariates and regression specifications for estimating principal scores and effects.
-First, Proposition \ref{prop:reg2} requires consistent estimation of principal scores and (via Assumption \ref{ass:rci}) correct specification of the outcome model, that is, the model for $\EE[Y|\bxy,Z,\st]$.
-Hence, it is essential to choose covariates and transformations $\bxs$ and $\bxy$ that lead to a good model fit.
-
-Second, Assumption \ref{ass:vps} requires principal scores to vary with at least some element $\bxs$.
+Three criteria governed our choice of covariates and model specifications for estimating principal scores and effects.
+First, Proposition \ref{prop:reg2} requires consistent estimation of principal scores and (via Assumption \ref{ass:rci}) correct specification of the model for $\EE[Y|\bx,Z,\st]$.
+However, unlike \psw, \geepers does not require that models include all important covariates (for any definition of "important"), so covariates whose inclusion may threaten model specifications or identification assumptions can be dropped from the analysis.
+
+On the other hand, Assumption \ref{ass:vps} requires principal scores to vary with at least some element $\bxs$.
 Moreover, the results of the simulation study suggest that when $\st$ can be predicted more precisely as a function of $\bxs$, \textsc{geepers} principal effect estimators are more precise and more robust to model misspecification.
@@
 
-There is good reason to believe that analogous arguments apply to the selection of outcome covariates $\bxy$.
-
-The first two criteria motivated our choice of two strategies for choosing $\bxs$ and $\bxy$.
+The first two criteria motivated our choice of two strategies for choosing $\bxs$ and model specifications. 
 In our main model, we included every available covariate in the dataset and chose transformations of those covariates based on the results of residual and binned residual plots \citep{arm}.
-
-Including all possible covariates maximizes the model's opportunities to detect covariate relationships that would predict $Y$ or $\st$.
+%
+% Including all possible covariates maximizes the model's opportunities to detect covariate relationships that would predict $Y$ or $\st$.
 However, this strategy can also lead to overfitting, which would decrease out-of-sample predictive accuracy.
-In an alternative analysis, we chose a subset of covariates to include in $\bxs$ and $\bxy$ with a backward stepwise selection algorithm minimizing the AIC \citep{aic} of the principal score model and then modified the resulting models based on residual plots.
-The model selection process is recorded in the GitHub repository cited above.
+In an alternative analysis, we chose a subset of covariates to include in $\bxs$ %and $\bxy$ 
+with a backward stepwise selection algorithm minimizing the AIC \citep{aic} of the principal score model and then modified the resulting models based on residual plots.
 
@@
 \textsc{geepers} is a flexible, easily-implementable method for principal effect estimation for one-way noncompliance, with predictive covariates; extensions to a broader set of circumstances could be a boon to causal modeling.
+
+\section*{Replication Materials}
+Replication materials can be found at
+\url{https://osf.io/fu2sv/?view_only=b2baebc2f3054ef796717863e4a9662f}.
 
@@
   \begin{split}
-    \EE[Y|\bx,Z=0]&=\beta_0+\beta_1\pp+\bm{\gamma_1}'\bm{X^Y}+\bm{\gamma_2}'\bm{X^Y}\pp\\
-                  &=\beta_0+\beta_1R+\bm{\gamma_1}'\bm{X^Y}+\bm{\gamma_2}'\bm{X^Y}R\\
+    \EE[Y|\bx,Z=0]&=\beta_0+\beta_1\pp+\bm{\gamma_1}'\bxy+\bm{\gamma_2}'\bxy\pp\\
+                  &=\beta_0+\beta_1R+\bm{\gamma_1}'\bxy+\bm{\gamma_2}'\bxy R\\
     \EE[Y|\bx,S,Z=1]&=\beta_0+\beta_2+(\beta_1+\beta_3)S+(\bm{\gamma_1}+\bm{\gamma_3})\bxy+(\bm{\gamma_2}+\bm{\gamma_4})\bxy S\\

### 2025-07-02 code/OPTS/opts.r
appendix figure

@@ -90,7 +90,10 @@ psMod <- geepers$psMod
 arm::binnedplot(predict(psMod,type='response'),resid(psMod,type='response'))
 summary(psMod)
 
-aucMod(psMod)
+print(aucMod(psMod))
+
+### standardized coefficients:
+print(glm(S~scale(ETXU_CAT1)+scale(OFIBRIN1),data=mf,family=binomial))
 
 ## bs <- replicate(5000,
 ##                 coef(

### 2025-07-02 code/simulation/displaySimPublish.r

@@ -215,7 +215,10 @@ pd <- results %>%
   ) %>%
   filter(estimator=='PSW'|rhat<1.1)
 
+#######################################################
 ### figure for paper
+#######################################################
+
 
 norm<-bp(pd,
    subset=b1 > 0& errDist=='norm'& eff==1&mu01==0&n==500&PE=='Stratum 1',
@@ -258,6 +261,53 @@ dev.off()
 setwd("simFigs")
 system("lualatex boxplotsNew.tex")
 setwd("..")
+
+
+#######################################################
+### "boxplots" figure with n=1000 for appendix
+#######################################################
+norm1000<-bp(pd,
+   subset=b1 > 0& errDist=='norm'& eff==1&mu01==0&n==1000&PE=='Stratum 1',
+   title="Normal Residuals",ylim=c(-1.5,1.5),#c(-1.5,1.5),
+   facet=B1~intAll,#interactionZ+interactionS,
+   Labeller=labeller(B1="none",#label_parsed,
+                     intAll=label_value),labSize=2.5
+   )+
+    labs(y=bquote("Estimation Error for "~tau^1),#subtitle="No Interactions",
+         x=NULL)+
+  #theme_bw()+
+    theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1),
+          strip.text.y=element_blank(),strip.background.y=element_blank(),
+                                        #plot.subtitle = element_text(size=10),
+        legend.position="none")+
+  scale_fill_manual(values=c('#1b9e77','#d95f02','#7570b3'))+
+  scale_color_manual(values=c('#1b9e77','#d95f02','#7570b3'))
+
+unif1000<-bp(pd,
+   subset=b1 > 0& errDist=='unif'& eff==1&mu01==0&n==1000&PE=='Stratum 1',
+   title="Uniform Residuals",ylim=c(-1.5,1.5),#c(-1.5,1.5),
+   facet=B1~intAll,#interactionZ+interactionS,
+   Labeller=labeller(B1=label_parsed,intAll=label_value),labSize=2.
+   )+
+    labs(y=NULL,#bquote("Estimation Error for "~tau^1),#subtitle="No Interactions",
+         x=NULL)+
+  #theme_bw()+
+  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1),
+                                        #plot.subtitle = element_text(size=10),
[…]

### 2025-07-02 writeUps/appendix.tex
finished?? editing

--- 2025-06-26 writeUps/paper.tex
@@
 \usepackage[letterpaper]{geometry}
-
+\usepackage{xr}
+\externaldocument{writeUps/paper}
 \RequirePackage{amsmath,amsfonts,amssymb,amsthm}
@@
 \usepackage{longtable}
-%\usepackage[notablist,tablesfirst]{endfloat}% To `separate' figures and tables from text if required
 \usepackage{threeparttable}
-\usepackage{epstopdf}% To incorporate .eps illustrations using PDFLaTeX, etc.
-\usepackage[caption=false]{subfig}% Support for small, `sub' figures and tables
-\usepackage[doublespacing]{setspace}% To produce a `double spaced' document if required
-%\setlength\parindent{24pt}% To increase paragraph indentation when line spacing is doubled
+\usepackage[caption=false]{subfig}
+\usepackage[doublespacing]{setspace}
 
-\usepackage[longnamesfirst,sort]{natbib}% Citation support using natbib.sty
-\bibpunct[, ]{(}{)}{;}{a}{,}{,}% Citation support using natbib.sty
+\usepackage[longnamesfirst,sort]{natbib}
+\bibpunct[, ]{(}{)}{;}{a}{,}{,}%
 \renewcommand\bibfont{\fontsize{10}{12}\selectfont}% To set the list of references in 10 point font using natbib.sty
@@
 
-\title{GEEPERs: Principal Stratification using Principal Scores and Stacked Estimating Equations}
+\title{Online Appendices for ``GEEPERs: Principal Stratification using Principal Scores and Stacked Estimating Equations''}
 %\runtitle{\textsc{geepers}}
 %\maketitle
-
-%\begin{aug}
- \author{
- Adam C. Sales\thanks{CONTACT Adam C. Sales. Email: asales@wpi.edu}, Kirk P. Vanacore, and Erin R. Ottmar%\\
- %\affil{Worcester Polytechnic Institute, Worcester, MA, USA}
-}
-
-% \maketitle
-
-\maketitle
-
-\begin{abstract}
-Principal stratification is a framework for making sense of causal effects conditioned on variables that themselves may have been affected by treatment.
-For instance, one component of an educational computer application is the availability of “bottom-out” hints that provide the answer.
-In evaluating a recent experimental evaluation against alternative programs without bottom-out hints, researchers may be interested in estimating separate average treatment effects for students who, if given the opportunity, would request bottom-out hints frequently, and for students who would not.
-Most principal stratification estimators rely on strong structural or modeling assumptions, and many require advanced statistical training to fit and check.
-In this paper, we introduce a new M-estimation principal effect estimator for one-way noncompliance based on a binary indicator.
-Estimates may be computed using conventional regressions (though the standard errors require a specialized sandwich formula) and do not rely on distributional assumptions.
-We present a simulation study that demonstrates the novel method's greater robustness compared to popular alternatives and illustrate the method through two real-data analyses.
-\end{abstract}
-% \begin{keywords}
-% Causal Inference; Principal Stratification; Educational Technology
-% \end{keywords}
-
-\clearpage
-
-\section{Introduction}
-
-When students get stuck on math problems during homework or classwork, it may be helpful to offer them hints or other ``just-in-time'' support \cite[e.g.,][]{jit}, possibly helping them overcome specific misconceptions or other sources of confusion and allowing them to move on in the assignment.
-The availability of problem-specific support for struggling students is one potential advantage of educational technology over traditional pencil-and-paper practice.
-On the other hand, the ready availability of hints may discourage students from working through difficult problems independently, thereby hindering their learning.
-
-The literature on hints in online learning is mixed \citep[see, e.g.,][]{aleven2016help,goldin2012learner,sales2021student}, particularly when it comes to requesting a ``bottom-out'' hint.
-In some online mathematics programs, sequences of hints, which a student can request one at a time, end with a bottom-out hint providing the answer to the problem.
-On the one hand, hints that include the correct answer are essentially worked examples, which can be beneficial to learning \citep[e.g.]{sweller1985use}.
-On the other hand, they allow students to ``game the system,'' by simply requesting enough hints to see the right answer and use the software without ever working to solve any problems \citep[e.g.][]{guo2008trying}.
-
-Data from the randomized controlled trial (RCT) reported in \citet{impactPaper} may provide evidence for one or another side of this debate.
-That RCT contrasted four different applications for computer-based math practice, one of which included bottom-out hints.
-If that condition benefited students who frequently requested bottom-out hints (``bottom-outers'') more than those who didn't, that may suggest that the availability of bottom-out hints played a role in the condition's effectiveness.
-Alternatively, if students who avoided bottom-out hints benefited more, it may be because the availability of bottom-out hints impeded the condition's effectiveness.
-(There are certainly other factors at play as well, so this analysis would not be conclusive.)
-Indeed, if the use of bottom-out hints is considered part of the treatment, then those students who rarely used them were partial compliers---did the treatment still benefit them?
-
-Estimating an effect for bottom-outers, or other partial compliers, is a tricky proposition.
-Causal inference in statistics typically depends on comparisons of outcomes between similar groups of subjects.
-However, there is no readily identifiable comparison group for bottom-outers, who are only represented in one of the four treatment conditions, and who presumably struggled more frequently, on average, than their non-bottom-outer peers.
-We will formulate a solution using principal stratification \citep{frangakis}, a framework for defining and estimating effects for post-treatment subgroups, that is, subgroups that may themselves be a function of treatment assignment.
-It relies on the notion of \emph{potential} subgroups called principal strata: which subgroup would a subject be in if they were (perhaps counterfactually) assigned to one treatment condition or another.
-For example, the appropriate comparison group for bottom-outers is the set of students, assigned to the other treatment arms, who \emph{would have} requested bottom-out hints if only given that option.
-
-Since principal strata membership is based on subjects' responses to both factual and counterfactual treatment assignments, it is unobserved---we don't observe which students in the conditions without bottom-out hints would have bottomed out if given the opportunity.
-Nonetheless, researchers have devised estimators of principal effects---average effects within principal strata---to address a broad range of applied statistical problems.
-
-\label{background}The conceptual basis for principal stratification---though not the terminology---emerged from the study of randomized experiments with noncompliance.
-Specifically, \citet{air} showed that the econometric instrumental variables estimand is the average effect of the treatment on the principal stratum of ``compliers''---those subjects who would receive the treatment if and only if randomized to the treatment condition.
-Moreover, the average effect for compliers is nonparametrically identified if a random treatment assignment only has an effect for compliers---``exclusion restriction''---and if treatment assignment never \emph{prevents} any subjects from receiving treatment (``monotonicity'').
-In a follow-up analysis, \citet{imbens1997bayesian} relaxed the exclusion restriction by way of parametric Bayesian finite mixture modeling, which remains a popular approach.
-
-\citet{frangakis} introduced the terminology of principal stratification and extended the framework well beyond noncompliance, inspiring a broad literature introducing (and debating) novel applications and estimators.
-For instance, \citet{fellerEtAl2016} estimated differential effects of random assignment to the Head Start early childhood care program for children who, if assigned to control, would receive home-based care, and for those who would attend a different early childhood care program.
-Other methodologists have developed estimators for evaluating surrogate outcomes \citep{li2010bayesian}, investigating causal mechanisms \citep{lidsayPage}, addressing sample attrition \citep{zhangRubin, ding2011}, and other related problems.
-Since interest is often in comparing effects between principal strata---a goal that runs counter to the exclusion restriction---the earliest principal stratification estimators built on \citet{imbens1997bayesian}'s parametric mixture models \citep[e.g.][]{Barnard01062003,mealli2004analyzing}, and that approach remains popular \citep{lidsayPage,fellerEtAl2016}.
-However, these models are often difficult to specify and can be either hard or impossible to test.
-Even when they are well-specified, they can yield estimates that are severely biased and misleading \citep{griffin2008application,feller2016principal}.
-The most common type of parametric model, the normal mixture model, is especially dubious for studying bottom-outers in \citet{impactPaper}'s RCT, whose outcome, the total score on a 10-item posttest, is not normally distributed.
-
-These concerns have prompted the development of alternative approaches, including bounding \citep{bounding} and randomization inference \citep{nolen2011randomization}.
-Of particular importance is the ``principal score,'' \citep{jo,dingLu,feller2017principal}, or the probability that a subject belongs to a particular principal stratum, as a function of pre-treatment covariates.
-Under a strong, untestable assumption called ``principal ignorability,''  principal scores can be transformed into weights to compute an unbiased estimate of principal effects.
-Principal ignorability is a direct analogue of the strong ignorability assumption in causal observational studies \citep[e.g.][]{rosenbaum1983central}---that all covariates that predict both principal stratum membership and outcomes have been accounted for---in other words, the precise type of assumption that randomized trials are designed to avoid.
-A separate set of results in \citet{jo2002}, \citet{ding2011}, \citet{jiangDing2021}, and others have shown that, under certain circumstances, researchers can use baseline covariates to identify and estimate principal effects without relying on principal ignorability.
-
-This paper expands on those latter results by presenting a new principal effect estimator for the case of one-way noncompliance, in which principal stratum membership is observed in one treatment group but not the other (future work will extend our approach to general scenarios).
-Our estimator boils down to a simple, three-step procedure: first, estimate principal scores, for instance, with logistic regression; then, impute unknown principal subgroup membership indicators with principal scores; and finally, estimate principal effects using ordinary least squares (OLS) regression with interactions.
-%, just as in the pre-treatment subgroup case.
-We show that this estimator is the solution to a set of estimating equations based on the combined moments of principal scores, indicators for treatment assignment and observed principal stratum membership in the treatment group, outcomes, and (optionally) covariates, and give a set of assumptions under which it is consistent for the true principal effects.
-We will refer to the estimators presented here as ``\textsc{geepers}'': general estimating equations for principal effects using regressions.
-\textsc{geepers} estimates rely on the correct specification of regression functions, but unlike common fully-parametric methods, they make no assumptions about the shape of the distribution of regression errors.
-Unlike principal score weighting, \textsc{geepers} does not assume principal ignorability and admits to sandwich-style standard error estimates \citep{stefanskiBoos}, which we provide in the appendix.
-
-After the initial submission of this manuscript, we became aware of \citet{richardson2023estimating}, which suggests a similar, though not identical, estimator to ours, using G-estimation of a structural nested mean model \citep[e.g.][]{wallace2017r}.
-Our development of \textsc{geepers} builds on \citet{richardson2023estimating} by emphasizing connections to both parametric mixture modeling and principal score analysis, and comparing \textsc{geepers} to those methods in a simulation study.
-Moreover, \textsc{geepers}, which, in practice, is a combination of logistic regression and OLS, is more accessible to applied researchers while yielding very similar results.
-
-This work advances the literature in three ways: first, it makes an explicit connection between principal score methods, M-estimation, and semi-parametric estimation of principal effects using auxiliary data.
-Second, it provides a general estimator that relies on two of the most familiar tools in the applied education researcher's toolbox: linear and logistic regression.
-The \textsc{geepers} method can, therefore, make principal stratification methods accessible for a much broader range of applied education researchers.
-Lastly, this work includes an extensive simulation study that compares \textsc{geepers} to parametric mixture modeling and principal score weighting under a range of favorable and challenging scenarios.
-The simulation demonstrates the strengths and weaknesses of each method, as well as the trade-offs between them.
-
-The following section gives a formal introduction to principal stratification and briefly describes two common estimation techniques, parametric mixture modeling and principal score weighting.
-Next, Section \ref{sec:geepers} develops \textsc{geepers} estimation and gives conditions for its consistency.
-Section \ref{sec:simulation} describes the simulation study.
-% examining the operating characteristics of \textsc{geepers} estimates and comparing them to estimates from parametric mixture modeling and principal score weighting.
-%Under ideal circumstances, estimates from parametric mixture models and weighting estimators are somewhat more efficient than \textsc{geepers}, but when assumptions are violated \textsc{geepers} easily out-performs the other two methods.
-Section \ref{sec:fh2t} illustrates \textsc{geepers} in two applied examples: first, we re-analyze the applied example from \citet{richardson2023estimating}, a randomized trial of a periodontal treatment for pregnant women in which there was significant one-way noncompliance.
-Second, we return to the question of bottom-out hints: does the effect of being randomized to a condition that includes bottom-out hints depend on how often they are requested?
-Section \ref{sec:discussion} concludes.
-
-\section{Background}
-In a randomized experiment with two conditions, let $Z_i=1$ if subject $i$, $i=1,\dots,n$ is assigned to the treatment condition, and let $Z_i=0$ otherwise.
-If interest is in the effect of $Z$ on an outcome $Y$, under the ``stable unit treatment value assumption'' \citep{air} of no interference between units and no hidden versions of the treatment, define potential outcomes \citep{splawa1990application} $\yti$ as the value of $Y$ that $i$ would exhibit if $Z_i=1$, and let $\yci$ be the value if $Z_i=0$.
-Then the observed value of $Y$ satisfies $Y_i=Z_i\yti+(1-Z_i)\yci$, and the treatment effect of $Z$ on $Y$ for subject $i$ may be defined as $\tau_i=\yti-\yci$.
-The average treatment effect (ATE) is $\EE[\tau]$.
-%
-
-Let $S_i\in\{0,1\}$ be a measurement on subject $i$ taken after treatment assignment, so that $S$ may be affected by $Z$.
-Then $S$ itself has potential values $\sti$ and $\sci$, which $i$ would exhibit if $Z_i=1$ or $Z_i=0$, respectively.
-Assume ``strong monotonicity'' \citep[c.f.][]{dingLu} (also referred to as ``one-way noncompliance''):%\citealt{wing2017can}):
-\begin{ass}[Strong Monotonicity]\label{ass:sm}
-\begin{equation*}%\label{eq:strongMonotonicity}
-  \sci=0\text{ for all }i
-\end{equation*}
-\end{ass}
-Like potential outcomes $\yti$ and $\yci$, $\sci$ and $\sti$ are defined for all subjects, though $\sci$ is only observed when $Z_i=0$ and $\sti$ is only observed when $Z_i=1$.
-Under strong monotonicity, since $S$ is binary, every subject belongs to one of two ``principal strata,'' $\sti=1$ or $\sti=0$.
-The average treatment effect in each principal stratum, $\eff1=\EE[\tau|\st=1]$ or $\eff0=\EE[\tau|\st=0]$, is called a principal effect. %
-The challenge in estimating principal effects is that strata membership $\sti$ is only observed when $Z_i=1$, in which case $\sti=S_i$, the observed value of $S$.
-
-\subsection{Estimating Principal Effects}
-In estimating principal effects for a binary $S$, it will be useful to define four conditional means. For $s\in\{0,1\}$, let
-\begin{equation}\label{eq:mus}
-\mucs=\EE[Y_C|\st=s] \mbox{ and }\muts=\EE[Y_T|\st=s]%\mbox{ for }s=0,1
-  % \begin{split}
-  %   \muc0&=\EE[Y_C|\st=0]\\
-  %   \muc1&=\EE[Y_C|\st=1]\\
-  %   \mut0&=\EE[Y_T|\st=0]\\
-  %   \mut1&=\EE[Y_T|\st=1]
-  % \end{split}
-\end{equation}
-Then, since
-\begin{equation*}
-  \effs=\muts-\mucs
-  % \begin{split}
-  %   \eff0&=\mut0-\muc0\\
-  %   \eff1&=\mut1-\muc1
-  % \end{split}
-\end{equation*}
-estimating principal effects requires estimating $\muc0$, $\mut0$, $\muc1$, and $\mut1$.
-
-We will focus on estimation in completely randomized experiments; that is, we will assume
-\begin{ass}[Randomization]\label{ass:rand}
-\begin{equation*}%\label{eq:randomization}
-  \yci,\yti,\sti, \bm{x}_i \independent Z_i
-\end{equation*}
-\end{ass}
-where $\bm{x}_i$ is a vector of pre-treatment covariates, and $\independent$ denotes independence.
-
-Under Assumptions \ref{ass:sm} and \ref{ass:rand}, conditional means $\mut0$ and $\mut1$ are nonparametrically identified.
-In particular, the means of observed outcomes for subjects with $Z=1$ and $S=0$ or $S=1$ are unbiased for $\mut0$ and $\mut1$, respectively, since when $Z_i=1$, $Y_i=\yti$ and $S_i=\sti$.
-In contrast, $\muc0$ and $\muc1$ are not fully identified without further assumptions because $\sti$ and $\yci$ are never observed simultaneously (though they may be nonparametrically bounded; see, e.g., \citealt{bounding}).
-% When $S$ is a measure of compliance, and $\tau_i=0$ whenever $\sti=0$ (the ``exclusion restriction''), $\muc1$---and hence $\eff1$---may be estimated with instrumental variables methods \citep{air}.
-% In this case, $\eff1$ is often called the ``local average treatment effect'' or the average effect of the treatment on the treated.
-We will briefly review two approaches to estimating $\muc0$ and $\muc1$% that do not assume the exclusion restriction
----normal mixture modeling and weighting---with some comments in-between on the role for covariates.
-
-\subsubsection{Parametric Mixture Modeling}
-%The normal mixture modeling approach to estimating $\muc0$ and $\muc1$ \citep{imbens1997bayesian} assumes that, conditional on $\st$, $Y_C$ is normally distributed.
-Say an analyst assumes that, within principal strata, $Y_C$ is drawn from a parametric model, i.e., $f_{Y_C}(y|\st=s)=g^s(y)$.
-For instance, \citep{imbens1997bayesian} assumes $Y_C$ is normally distributed within strata, so $g^s(y)=\phi\left(\frac{y-\mucs}{\sigma_C^s}\right)$, where $\phi(\cdot)$ is the standard normal density function, and $\sigma_C^0>0$ and $\sigma_C^1>0$ are standard deviations.
-Then the probability density of $Y_C$ may be written as
-\begin{equation}\label{eq:mixtureModel}
- % f_{Y_C}(y)=Pr(\st=0)\phi\left(\frac{y-\muc0}{\sigma_C^0}\right)+Pr(\st=1)\phi\left(\frac{y-\muc1}{\sigma_C^1}\right)
- f_{Y_C}(y)=Pr(\st=0)g^0(y)+Pr(\st=1)g^1(y).
-\end{equation}
-The probabilities $Pr(\st=0)$ and $Pr(\st=1)$ may be estimated first using data from the treatment group, since Assumption \ref{ass:rand} implies that $Pr(\st=1|Z=1)=Pr(\st=1)$.
-Given these estimates and \eqref{eq:mixtureModel}, parameters $\muc0$ and $\muc1$ may be estimated using maximum likelihood, method of moments, or Bayesian techniques.
-
-This approach has several problems.
-First, the assumption of a conditional parametric model is untestable and limiting.
-Unfortunately, even when it holds, mixture model estimates can be unreliable and, in some cases, severely biased \citep{griffin2008application,feller2017principal}.
-
-\subsection{Principal Scores}
-The principal score \citep[e.g.][]{jo} is defined as
-\begin{equation}\label{eq:pscore}
-  \ppi=Pr(\sti=1|\bxsi),
-\end{equation}
-%where $\bxsi\subseteq\bm{x}_i$, the set of covariates used to model $S$.
-the probability of a subject with covariates $\bxsi$ being in the $\sti=1$ principal stratum.
-Note that under randomization, $Pr(\sti=1|\bxsi,Z_i)=Pr(\sti=1|\bxsi)$, so a model for principal scores can be estimated using data from the treatment group and extrapolated to the control group.
-Assume a model for principal scores:
-\begin{equation}\label{eq:prinScoreMod}
-  \ppip{\bm{\alpha}}=f(\bxsi;\bm{\alpha})
-\end{equation}
-with parameter vector $\bm{\alpha}$.
-An analyst may estimate $\bm{\alpha}$ as $\bm{\hat{\alpha}}$ by fitting model \eqref{eq:prinScoreMod} using observed $\bm{S}$ and $\bxs$ for subjects with $Z=1$, and then compute estimated principal scores $\hppi=\ppip{\bm{\hat{\alpha}}}$ for subjects with $Z_i=0$.
-
-Going forward, we will occasionally suppress dependence on $\bm{\alpha}$, and write $\ppip{\bm{\alpha}}=\ppi$.
-
-If $\bm{x}$ is informative about $\st$, principal scores will vary between subjects. We will assume this is the case: %the following:
-\begin{ass}[Variable Principal Scores]\label{ass:vps}
- $\pp$ takes at least 3 distinct values
-\end{ass}
-%In some estimators we will discuss below, two distinct values suffice.
-For a broader discussion, see \citet{ding2011,jiangDing2021}.
-
-Principal scores can potentially improve inference in a finite mixture model by modifying \eqref{eq:mixtureModel} as
-\begin{equation}\label{eq:mixtureModelPS}
-  f_{Y_C}(y|\bm{x})=(1-\ppip{\bm{\alpha}})%\phi\left(\frac{y-\muc0}{\sigma_C^0}\right)
-g^0(y|\bm{x})+\ppip{\bm{\alpha}}%\phi\left(\frac{y-\muc1}{\sigma_C^1}\right)
-g^1(y|\bm{x})
-\end{equation}
-where $g^s(y|\bm{x})$, $s=0,1$, are conditional densities which must be specified.
-Most commonly, researchers assume a linear normal model, e.g., $g^s(y|\bm{x})=\phi\left(\frac{y-\mucs-\bm{\gamma}'\bm{x}}{\sigma_C^s}\right)$, where the columns of $\bm{x}$ are centered so that $\mathbb{E}[\bm{x}]=\bm{0}$.
-%Ideally, a statistician will estimate parameters $\muc0$, $\muc1$, $\mut0$, $\mut1$, $\sigma^0_C$, $\sigma^1_C$, and $\bm{\alpha}$ simultaneously, using maximum likelihood or Bayesian techniques.
-
-Alternatively, an analyst may assume ``principal ignorability'' \citep{jo,dingLu}:
-%\begin{ass}[Principal Ignorability]\label{ass:PI}
-\begin{equation}\label{eq:pi}
-  Y_{C}\independent \st |\bxs
-\end{equation}
-%\end{ass}
-or a somewhat weaker version, $\EE[Y_C|\st,\bxs]=\EE[Y_C|\bxs]$ \citep{feller2017principal}.
-That is, the principal stratum is unrelated to control potential outcomes conditional on covariates.
-Principal ignorability \eqref{eq:pi} is reminiscent of ignorability assumptions typical in observational studies. % \citep[e.g.][]{rosenbaum2010design}.
-% Both require a rich set of covariates and background knowledge of the determinants of $S$;
-In particular, an unobserved covariate $u$ that is correlated with both $\st$ and $Y_C$ would invalidate \eqref{eq:pi}.
-
-Under principal ignorability, $\muc0$ and $\muc1$ can be estimated via weighting:
-\begin{equation}\label{eq:psw}
-  \hat{\muc0}_w=\frac{\sum_{i: Z_i=0} Y_i(1-\ppi)}{\sum_{i:Z_i=0}1-\ppi}\text{ and } \hat{\muc1}_w=\frac{\sum_{i: Z_i=0} Y_i(\ppi)}{\sum_{i:Z_i=0}\ppi}
-\end{equation}
-
-Principal score weighting (\textsc{psw}) does not require any distributional assumptions and (after estimating principal scores) is very easy to implement.
-On the other hand, the principal ignorability assumption is strong and restrictive.
-
-\citet{feller2017principal} recommends the case-resampling bootstrap to estimate the sampling variances of $\hat{\mu}_{Cw}^0$, $\hat{\mu}_{Cw}^1$, and their associated principal effect estimators.
-
-\subsection{M-Estimation}\label{sec:mest}
-M-estimation (also generalized method of moments or generalized estimating equations) is a general approach to deriving consistent statistical estimators and their asymptotic distributions.
-In the following section, we will describe \textsc{geepers}, a novel principal effect estimator, and derive its properties using the principles of M-estimation.
-Here, we provide a brief overview of M-estimation, primarily drawn from the excellent treatment in \citet{stefanskiBoos}.
-
-Let $\bm{\mathcal{O}}_i$ be a vector of observed data for subject $i=1,\dots,n$, with $\bm{\mathcal{O}_1},\dots,\bm{\mathcal{O}}_n\overset{\mathrm{iid}}{\sim} P$ for some distribution $P$, %\bm{\mathcal{O}}_i\independent \bm{\mathcal{O}}_{i'}$ if $i\ne i'$,
-let $\bm{\theta}$ be a vector of parameters, 
-and let $\bm{\psi}(\bm{\mathcal{O}},\bm{\theta})$ be a vector field such that $\mathbb{E}_P[\bm{\psi}(\bm{\mathcal{O}},\bm{\theta}_0)]=\bm{0}$ for the true parameter value $\bm{\theta}_0$. 
-Then, an M-estimator $\bm{\hat{\theta}}$ satisfies ``estimating equations'' $\sum_i \psi_j\left(\bm{\mathcal{O}}_i,\bm{\hat{\theta}}\right)=0$. 
-Some common examples are maximum likelihood estimators, where $\bm{\psi}(\bm{\mathcal{O}},\bm{\theta})$ is the score function %, with $\psi_j(\bm{\mathcal{O}},\bm{\theta})=\partial log f(\bm{\mathcal{O}},\bm{\theta})/\partial \theta_j$ and $f(\cdot,\bm{theta})$ is the density of $\bm{\mathcal{O}}$,
-or method of moments estimators, in which, if data $\mathcal{O}$ are unidimensional, $\psi(\mathcal{O},\bm{\theta})=\mathcal{O}-\EE_\theta\left[\mathcal{O}\right]$.
-
-Under suitable regularity conditions, an M-estimator $\bm{\hat{\theta}}$ is asymptotically normal, and its sampling variance-covariance matrix may be estimated using the formula \citep[][ch.
-7]{boosStefanskiBook}:
-\begin{equation}\label{eq:sandwich}
-  \widehat{var}(\bm{\hat{\theta}})=ABA'
-\end{equation}
-%\citep[][ch. 7]{boosStefanskiBook}.
-with matrices
-\begin{equation*}
-    A=\left\{\frac{1}{n}\displaystyle\sum_i\left.
-\frac{\partial \bm{\psi}(\bm{\mathcal{O}_i},\bm{\theta})'}{\partial \bm{\theta}'}\right|_{\bm{\theta}=\bm{\hat{\theta}}}\right\}^{-1}\mbox{ and }B=\frac{1}{n}\sum_i \bm{\psi}(\bm{\mathcal{O}_i},\bm{\hat\theta})\bm{\psi}(\bm{\mathcal{O}_i},\bm{\hat\theta})'.
-\end{equation*}
-Because of its form, \eqref{eq:sandwich} is sometimes called a ``sandwich'' formula, where $A$ is the ``bread'' and $B$ is the ``meat'' \citep[e.g.][]{sandwich}. 
-%To define the matrices $A$ and $B$, let the vector $\bm{\psi}_i=\left\{\psi_j(\bm{\mathcal{O}}_i,\bm{\hat{\theta}})\right\}_{j=1}^J$. Then,
-% \begin{equation*}
-%   A^{-1}=\frac{1}{n}\sum_i\left. \frac{\partial \bm{\psi}_i'}{\partial \bm{\theta}'}\right|_{\bm{\theta}=\bm{\hat{\theta}}}
-% \end{equation*}
-% and
-% \begin{equation*}%\label{eq:Bmat}
-%   B=\frac{1}{n}\sum_i \bm{\psi}_i\bm{\psi}_i'
-% \end{equation*}
-In general, the conditions for asymptotic normality may be difficult to prove; fortunately, they are met for OLS and common generalized linear models such as logistic regression, so long as the dimension of $\bm{\theta}$ does not increase with $n$.
-
-Multiple M-estimators can be combined by ``stacking'' their estimating equations.
-In particular, say an M-estimator with estimating equations $\bm{\psi}^2(\bm{\mathcal{O}},\bm{\hat\theta^2})=\bm{0}$ takes, as input, the results from a previous M-estimator, with estimating equations $\bm{\psi}^1(\bm{\mathcal{O}},\bm{\hat\theta^1})=\bm{0}$.
-The resulting estimator is the solution to the combined estimating equations
-\begin{equation}\label{eq:stacked}
-  \psi(\bm{\mathcal{O}},[\bm{\hat\theta^1},\bm{\hat{\theta}^2}])=[\psi^1(\bm{\mathcal{O}},\bm{\hat\theta^1}),\psi^2(\bm{\mathcal{O}},\bm{\hat\theta^2})]=\bm{0},
-\end{equation}
-%where the superscript $s$ stands for ``stacked.''
-A variance-covariance matrix estimated using the sandwich formula based on $\psi$ accounts for uncertainty from both steps of the estimation procedure.
-
-\section{GEEPERS }\label{sec:geepers}
-
-The approach we introduce here incorporates principal scores into an M-Estimator for the mixture model \eqref{eq:mixtureModel}.
-Like principal score weighting, our M-estimator does not require distributional assumptions but instead requires a conditional independence assumption.
-We will begin by describing a stronger-than-necessary conditional independence assumption to build intuition; in \S~\ref{sec:regression} we will present a considerably weaker alternative.
-
-\subsection{Building Intuition: Covariate Ignorability}\label{sec:ci}
-
-We begin by introducing an assumption that we term ``Covariate Independence,'' or CI. It is a slightly weaker version of \citet{jiangDing2021}'s ``auxiliary independence'' assumption.
-\begin{ass}[Covariate Independence]\label{ass:ci}
-\begin{equation}\label{eq:assumption}
-\EE[\yci|\bxsi,\sti]=\EE[\yci|\sti]=\muc0\text{ or }\muc1
-\end{equation}
-\end{ass}
-i.e., $Y_C$ is mean-independent of $\bxs$ conditional on $\st$.
-Under CI, covariates are not informative of the mean of $Y_C$ within principal strata.
-As stated above, CI will rarely be plausible.
-However, in some circumstances, researchers can identify a subset of observed covariates that satisfy CI, and use those covariates in estimation.
-
-It turns out that under strong monotonicity, randomization, and CI, and given a set of principal scores, the two principal effects $\eff0$ and $\eff1$ can be estimated by a simple OLS regression.
-To see how we'll %first
-derive the estimating equations for $\muc0$, $\muc1$, $\mut0$, and $\mut1$, and show that they are equivalent to estimating equations for a model fit by OLS.
-% assuming that principal scores are known exactly, and then generalize to the case in which they are estimated using data from the treatment group.
-The details of the calculations and the proofs are in the appendix.
-
-The argument stems from a set of expressions for expectations, summarized in the following lemma:
-\begin{lemma}\label{lemma:expectation}
-Let \begin{equation}\label{eq:estEq0}
-\tilde{\Psi}_i=\begin{pmatrix}
-    (1-Z_i)\left[Y_i-\muc0-\ppi(\muc1-\muc0)\right]\\
-    (1-Z_i)\left[\ppi Y_i-\ppi\muc0-\ppi^2(\muc1-\muc0)\right]\\
-    Z_i\left[Y_i-\mut0-\sti (\mut1-\mut0)\right]\\
-    Z_i\left[\sti Y_i -\sti\mut0-\sti^2(\mut1-\mut0)\right]
-  \end{pmatrix}
-\end{equation}
-Then under Assumptions \ref{ass:sm}, \ref{ass:rand}, \ref{ass:vps}, and \ref{ass:ci}, $\EE[\Psi_i]=\bm{0}$.
-\end{lemma}
-
-Lemma \ref{lemma:expectation} implies a set of estimating equations $\sum_{i=1}^n\tilde{\Psi}_i=\bm{0}$.
-
-Now, the estimating equations $\sum_i\tilde{\Psi}_i=\bm{0}$ are algebraically equivalent to the estimating equations for a particular OLS fit, after some transformations, the most important of which follows.
-Let
-\begin{equation}\label{eq:ri}
-\ri=Z_iS_i+(1-Z_i)\ppi=\begin{cases}
-\ppi &Z_i=0\\
-\sti &Z_i=1
-\end{cases}
-\end{equation}
-In general, $(1-Z_i)\ppi=(1-Z_i)\ri$ and $Z_i\sti=Z_i\ri$, allowing $\ri$ to take the place of $\ppi$ and $S_i$ throughout $\tilde{\Psi}_i$.
-Denote estimating equations based on $\ri$ as $\sum_i \psi_i=\bm{0}$.
-
-\begin{prop}\label{prop:reg1}
-  Under Assumptions \ref{ass:sm}, \ref{ass:rand}, \ref{ass:vps}, \ref{ass:ci}, and suitible regularity conditions, if %principal scores are known exactly,
-  $\ri$ is defined as in \eqref{eq:ri} and the model
-\begin{equation}\label{eq:regression0}
-  \EE[Y_i|Z_i,\ri]=\beta_0+\beta_1\ri+\beta_2Z_i+\beta_3Z_i\ri
-\end{equation}
-is fit with OLS, yielding coefficient estimates $\hat{\beta}_k$, $k=0,\dots,3$, then
-\begin{equation}
-  %\begin{split}
-    \heff0_{CI}\equiv \hat{\beta}_2\mbox{ and }
-    \heff1_{CI}\equiv \hat{\beta}_3+\hat{\beta}_2
-%  \end{split}
-\end{equation}
-are M-estimators, with $\heff0_{CI}\rightarrow_p\eff0$ and $\heff1_{CI}\rightarrow_p\eff1$ as $n\rightarrow\infty$
-\end{prop}
-
-Thus, given principal scores, Assumption \ref{ass:ci} enables an analyst to estimate principal effects with a simple regression.
-Proposition \ref{prop:reg1} builds on Theorem 2 of \citet{jiangDing2021} which establishes identification of $\eff0$ and $\eff1$ by representing the principal effect estimators as a familiar regression using estimated principal scores.
-
-% If principal scores were known exactly, the M-estimation structure of $\heff0$ and $\heff1$ implies that their sampling variance could be estimated by the standard sandwich estimator for OLS \citep{stefanskiBoos}.
-% This will rarely be the case.
-%However,
-If the principal scores are modeled as \eqref{eq:prinScoreMod}, and parameters $\bm{\alpha}$ are estimated by M-estimation, with estimating equations $\sum_i Z_i\Omega_i=\bm{0}$, then the estimating equations for the principal score model and principal effect estimation can be stacked as in \eqref{eq:stacked}.
-Sandwich standard error estimates of the form \eqref{eq:sandwich} based on $\left[Z_i\omega_i,\psi_i\right]$ will incorporate uncertainty from both principal score modeling and principal effect estimation conditional on principal scores.
-
-\subsection{Relaxing Covariate Independence with Outcome Modeling}\label{sec:regression}
-In most applications, identifying a set of covariates $\bxs$ satisfying CI will be difficult or impossible.
-Fortunately, CI may be relaxed or obviated by regression (see \citealt[][\S 3.4]{jiangDing2021} for an analogous identification result).
-
-The assumption we'll state here is stronger than necessary, but leads to an attractively simple method; a weaker version of the assumption and corresponding method, along with the proof, can be found in an appendix.
-
-\begin{ass}[Residualized Covariate Independence]\label{ass:rci}
-There exists $\bxy$, a (known) transformation of covariates $\bx$, and an (unknown) vector of coefficients $\bm{\gamma}$ such that
-\begin{equation}\label{eq:ass:rci}
-\begin{split}
-\EE[\yc-\bm{\gamma}'\bxy|\bxs,\st]&=\EE[\yc-\bm{\gamma}'\bxy|\st]\\
-\EE[\yt-\bm{\gamma}'\bxy|\bxs,\st]&=\EE[\yt-\bm{\gamma}'\bxy|\st]
-\end{split}
-\end{equation}
-\end{ass}
-That is, CI may not hold for potential outcomes themselves, but an analogous assumption holds for \emph{residualized} potential outcomes---i.e., $\yc$ and $\yt$ after subtracting out a linear function of (possibly transformed) covariates.
-By forcing $\bm{\gamma}$ to be the same for both treatment groups, and (implicitly) for both principal strata, Assumption \ref{ass:rci} assumes that there are no interactions between columns of $\bxy$ and either condition or principal stratum.
-This is akin to the ``Additivity of Treatment Assignment Effect'' assumption in \citet{jo2002}.
-Proposition \ref{prop:interactions} in the appendix relaxes these requirements, but some preliminary simulations suggest that standard errors from estimates allowing for these interactions will be prohibitively large.
-The simulation results in the following section address cases in which interactions are present in the data-generating model but not in the data analysis.
-
-In any event, Assumption \ref{ass:rci} allows principal effects to be estimated with an OLS model including $\bxy$:
-\begin{prop}{\textsc{geepers}}\label{prop:reg2}
-
-Under Assumptions \ref{ass:sm}, \ref{ass:rand}, \ref{ass:vps}, and \ref{ass:rci}, let principal scores be estimated as in \eqref{eq:prinScoreMod} using data from the treatment group and assume they are linearly independent of $\bxy$.
-Then, say the researcher fits the following model with OLS:
-\begin{equation}\label{eq:reg2}
-Y_i=\beta_0+\beta_1\ri+\beta_2 Z_i+\beta_3Z_i\ri+\bm{\gamma_1}'\bxy_i+\epsilon_i
-\end{equation}
-where $\ri$ is defined as in \eqref{eq:ri}.
-Then let
-\begin{equation}\label{eq:prinEffEst}
-  \begin{split}
-    \heff0_{RCI}&\equiv \hat{\beta}_2\\
-    \heff1_{RCI}&\equiv \hat{\beta}_2+\hat{\beta}_3
-  \end{split}
-   \end{equation}
-  $\heff0_{RCI}$ and $\heff1_{RCI}$ are M-estimators, with $\heff0_{RCI}\rightarrow_p\eff0$ and $\heff1_{RCI}\rightarrow_p\eff1$ as $n\rightarrow\infty$.
-  Under suitable regularity conditions, they are jointly asymptotically normal, with a variance of the form \eqref{eq:sandwich}.
-\end{prop}
-In short, researchers can estimate principal effects under strong monotonicity by simply imputing missing $\st$ values with estimated principal scores, fitting an OLS regression, and estimating standard errors with a sandwich formula.
-For the remainder of the paper, we will refer to principal effect estimators $\heff1_{RCI}$ and $\heff0_{RCI}$---our preferred estimators---as ``\textsc{geepers}.''
-
-\section{A Simulation Study}\label{sec:simulation}
-
-We conducted a simulation study to compare the performance of \textsc{geepers} with that of a Bayesian mixture model and a principal score weighting estimator, in both favorable and unfavorable circumstances.
-The study was designed to answer three sets of overarching questions.
-First, how does the performance of the \textsc{geepers} estimator vary under different conditions, including sample size, the extent to which $\bx$ predicts $\st$, %whether and how the principal effects themselves vary,
-and when important interactions are omitted from the outcome model?
-Second, how does \textsc{geepers} compare with mixture modeling?
-Is it competitive in circumstances favorable to both techniques?
-Does it avoid the pitfalls of mixture modeling when the parametric assumptions of the normal mixture model fail?
-Third, are there ways or circumstances when the principal score weighting estimator outperforms \textsc{geepers} or a mixture estimator even though principal ignorability does not hold?
-
-\subsection{Simulation Design}
-The simulation study was conducted in \texttt{R} \citep{rcite} and Stan \citep{rstan}, and full replication code is available at %[Redacted]. %
-\url{https://github.com/adamSales/psGee}.
-We replicated each set of conditions 500 times.
-
-\subsubsection{Data Generation}\label{sec:dataGeneration}
-
-In each run of the simulation, we simulated three independent standard normal covariates, $x_k$, $k=1,\dots,3$.
-However, only the first two covariates were ``observed,'' i.e., were included in the analysis model.
-Given all three covariates, the principal scores were set as
-\begin{equation*}
-  \ppi=logit^{-1}\left[\alpha(x_{1i}-x_{2i}+x_{3i})\right]
-\end{equation*}
-where $\alpha$ is a manipulated factor.
-When $\alpha$ was higher, $\st$ was more easily predicted by covariates; when $\alpha=0$, Assumption \ref{ass:vps} was violated, i.e., $Var(\pp)=0$.
-Principal stratum $\sti\in \{0,1\}$ was simulated as $S_i\sim Bern(\ppi)$.
-
-\label{potential}Potential and observed outcomes were generated as
-\begin{equation}\label{eq:y-sim}
-Y_i=0.3 Z_i\sti+(\gamma_1+\gamma_2\sti+\gamma_3Z_i)(x_{1i}+x_{2i})+\frac{1}{\sqrt{6}}x_{3i}+\epsilon_i%\\
-\end{equation}
-%   \begin{split}
-%   \yci&=\beta_1\sti+(\gamma_1+\gamma_2\sti)(x_{1i}+x_{2i})+\frac{1}{\sqrt{6}}x_{3i}+\epsilon_i\\
-%   \yti&=0.3\sti+(\gamma_1+\gamma_2\sti+\gamma_3Z_i)(x_{1i}+x_{2i})+\frac{1}{\sqrt{6}}x_{3i}+\epsilon_i\\
-%   Y_i&=Z_i\yti+(1-Z_i)\yci
-%   \end{split}
-% \end{equation}
-with %$\beta_3=0.3-\beta_1$, 
-$\EE[\epsilon_i]=0$, and $Var(\epsilon_i)=1/2$.
-The distribution of $\epsilon_i$ and values of coefficients $\gamma_1$, $\gamma_2$, and $\gamma_3$ were manipulated factors in the study.
-
-Treatment assignment $Z$ was simulated as independent of all the other variables, with half of the subjects assigned to treatment $Z=1$ and the rest to control $Z=0$.
-
-Since the covariates were centered and independent of $Z$, the true principal effects were 
-$\eff0=\gamma_3\EE[x1+x2|\st=0]$ and $\eff1=0.3+\gamma_3\EE[x1+x2|\st=1]$; in most simulation scenarios $\gamma_3=0$ so $\eff0=0$ and $\eff1=0.3$.
-
-\subsubsection{Manipulated Factors}\label{sec:manipulatedFactors}
-We manipulated five factors in the simulation, listed in Table \ref{tab:factor}.
-The five factors are not completely crossed---in particular, we let $n$ and $\alpha$ each vary across all the levels listed while holding the other four factors fixed.
-
-\begin{table}
-    \caption{\label{tab:factor} Manipulated factors in simulation study}
-  \centering
-  %\rowcolors{1}{}{lightgray}
-\begin{tabular}{r|rl}%{*{3}{c}}
-  \hline
-  Factor &\multicolumn{2}{l}{Levels}\\
-  \hline
-Sample size per condition $n$ &\multicolumn{2}{l}{$100,200,\dots,900,1000$}\\
-$\alpha$ &\multicolumn{2}{l}{$0,0.2,0.3,\dots,0.9,1$}\\
-The distribution of $\epsilon_i$&\multicolumn{2}{l}{ Normal, Uniform}\\
-%$\beta_1$ & \multicolumn{2}{l}{0, 0.3}\\
-\multirow{4}{*}{Interactions in \eqref{eq:y-sim}} & None & $\gamma_1=1/\sqrt{6};\gamma_2=\gamma_3=0$\\ 
-& $\bm{x}:\st$ & $\gamma_1=3/(4\sqrt{6})$; $\gamma_2=1/(2\sqrt{6})$; $\gamma_3=0$\\
-& $\bm{x}:Z$&$\gamma_1=1/\sqrt{6};\gamma_2=0$; $\gamma_3=1/(2\sqrt{6})$\\
-& $\bm{x}:\st$ \& $\bm{x}:Z$& $\gamma_1=3/(4\sqrt{6})$; $\gamma_2=\gamma_3=1/(2\sqrt{6})$\\
-% between $\st$ and covariates & No, Yes\\
-% Interactions between $Z$ and covariates& No, Yes\\
-\hline
-\end{tabular}
-\end{table}
-
-\textbf{\emph{Sample Size per condition $\bm{n}$:}} The guarantees of M-estimation are all asymptotic, so we examined the behavior of \textsc{geepers} at a range of sample sizes to determine its finite-sample properties.
-The range of 100--1000 was chosen to represent typical student-level randomized trials in education, for instance, \citet{growthMindsetRuralBurnette} with $n=115$ in the treatment group and $n=107$ in the control group, or \citet{impactPaper}, which initially randomized an average of 1000 students to each condition.
-
-$\bm{\alpha}$\textbf{:} As mentioned above, $\alpha$ controls the variance of the principal scores, i.e., the extent to which $\st$ is predictable as a function of covariates.
-When $\alpha=0$, the covariates are unrelated to $\st$ and principal scores do not vary between participants, violating \eqref{ass:vps}.
-We expect that the more predictive the covariates, the better the performance of all three estimators.
-To aid in the interpretation of the $\alpha$ parameter, the online appendix plots $\alpha$ against the area under the receiver operating characteristic curve for each fitted model (AUC), a common measure of classification accuracy \citep{bradley1997use}---an AUC of 0.5 implies that the model is no better than random guessing, and AUC of 1 implies perfect prediction.
-For $0.5\le \alpha \le 1$, the average AUC varies from roughly 0.53 to 0.77; for $\alpha=0.5$, the average AUC was roughly 0.675.
-
-\textbf{\emph{The distribution of $\bm{\epsilon_i}$:}} $\epsilon$ is the regression error in \eqref{eq:y-sim}.
-Across simulation runs, $E[\epsilon]=0$ and $Var(\epsilon)=1/2$.
-However, the shape of $\epsilon$'s distribution varied between the following two possibilities:
-Normal $\epsilon\sim\mathcal{N}(0,1/2)$, as assumed by the mixture model estimator, or Uniform $\epsilon\sim\mathcal{U}(-\sqrt{6}/2,\sqrt{6}/2)$.
-%\item[Mixture] $f_\epsilon(x)=\frac{1}{4}\phi\left((x+1/3)\sqrt{6}\right)+\frac{1}{4}\phi\left((x-1)\sqrt{6})\right)$
-%The mixture distribution for $\epsilon$ was designed to be particularly difficult for the mixture model estimator.
-
-% $\bm{\beta_1}$\textbf{:} When $\beta_1=0$, there is only one mixture component in the control
-% group, which can be problematic for mixture modeling \citep{griffin2008application,feller2016principal}.
-Conversely, when $\beta_1=0.3$, unless there is an interaction between $Z$ and covariates (i.e., $\gamma_3\ne 0$), $\eff0=\eff1=0$, so there is no average treatment effect for either principal stratum.
-
-\textbf{\emph{Interactions between $\st$ or $Z$ and covariates:}} The factor---the presence of interactions between covariates and $\st$ and $Z$, respectively--- tests the performance of a model that does not include interactions when that modeling assumption is false.
-
-When neither interaction was present, $\gamma_2=\gamma_3=0$, then $\gamma_1=1/\sqrt{6}$.
-In these cases, since  $Var(\gamma_1(x_1+x_2+x_3))=1/2$ and $Var(\epsilon_i)=1/2$, covariates explained half of the variance of $Y$ within treatment group and principal stratum, and observed covariates $x_1$ and $x_2$ explained 1/3 of that variance.
-These values align with estimates from outcome models in Section \ref{sec:fh2t} and are typical in education contexts \cite{hedgesHedberg}.
-
-When there was an interaction between covariates and $\sti$, then $\gamma_1=3/(4\sqrt{6})$ and $\gamma_2=1/(2\sqrt{6})$, so that the slopes for $x_1$ and $x_2$ varied by half their magnitude between principal strata, and on average they were equal to $1/\sqrt{6}$, as in the no-interaction case.
-Interactions between covariates and $Z$ were controlled by $\gamma_3$: when there were interactions, $\gamma_3=1/(2\sqrt{6})$, otherwise $\gamma_3=0$.
-Either type of interaction violates Assumption \ref{ass:rci} since if $\gamma_2\ne 0$ or $\gamma_3 \ne 0$, there is no single vector $\bm{\gamma}$ that can fully capture the dependence of $\yc$ and $\yt$ on $S$---instead, the dependence varies with $\st$ and/or with $Z$.
-
-\subsubsection{Analysis Models}\label{sec:simMods}
-
-Using each simulated dataset, we estimated $\eff0$ and $\eff1$ with three methods: \geepers, a parametric mixture model \eqref{eq:mixtureModelPS} including principal scores---hereafter \pmm---and the principal score weighting estimator \eqref{eq:psw}, hereafter \psw.
-
-The methods all used the same ``observed'' dataset, consisting of outcomes $Y$, treatment assignments $Z$, two covariates $\bxs=\bxy=\bx=[x_1,x_2]$, and $S$ observed only when $Z=1$.
-The third covariate used in the data-generating model, $x_3$, was ``unobserved''--since $x_3$ was correlated with both $\st$ and $Y$, Principal Ignorability \eqref{eq:pi} was violated.
-
-All three methods used the same principal score model, a logistic regression of $S$ on $\bx$, and an intercept, fit using data from the treatment group in which $S=\st$ was observed, i.e., $\ppf{\bm{\alpha}}{\bx}=logit^{-1}(\alpha_0+\bm{\alpha}'\bx)$.
-
-\pmm, the parametric mixture estimator, then fits two outcome models:
-For the treatment group, the regression model $Y=\beta_0^T+\beta_1^TS+\bm{\gamma}'\bx+\epsilon$ and for the control group, a mixture model of the form \eqref{eq:mixtureModelPS}, but with linear normal models for $g^s(y|x_1,x_2)$.
-Residual variance was allowed to differ across treatment groups, but not across principal strata.
-These models were fit simultaneously using Bayesian Markov Chain Monte Carlo with Stan \citep{rstan}.
-We interpreted the posterior mean as a point estimate and the posterior standard deviation as a standard error.
-
-For \textsc{geepers} and \psw, the principal score model was fitted to data from the treatment group using standard maximum likelihood methods.
-For \textsc{geepers}, we used the estimated principal scores in the control group, along with the observed $S$ in the treatment group, to construct $R$, and then fit the model \eqref{eq:reg2} with OLS.
-To estimate standard errors, we computed the sandwich covariance matrix following the procedure detailed in the appendix.
-
-For \psw we plugged estimated principal scores into \eqref{eq:psw} to estimate $\muc0$ and $\muc1$, estimated $\mut0$ and $\mut1$ with the sample means of subjects with $Z=1$ and $S=0$ or $S=1$, respectively, and estimated $\heff0=\hat{\mu}_{Tw}^0-\hat{\mu}_{Cw}^0$ and $\heff1=\hat{\mu}_{Tw}^1-\hat{\mu}_{Cw}^1$.
-
-\subsection{Results}\label{sec:simResults}
-In this section, we evaluate estimates of $\eff1$. Results for estimates of $\eff0$ and $\eff1-\eff0$, which are similar, can be found in the online appendix. 
-In replication $b=1,\dots,500$, define the error of a given estimator $\heff1_b$ as $err_b=\heff1_b-\eff1$.
-The bias of $\heff1$, $\EE[\heff1-\eff1]$, is estimated as $\widehat{bias}=\sum_b err_b/500$, and the empirical standard error is defined as $\left\{\sum_b(\heff1_b-\overline{\heff1})^2/499\right\}^{1/2}$, the sample standard deviation of $\heff1$.
-We also report the empirical coverage of confidence intervals, the proportion of replications in which $\eff1$ is within the estimated confidence bounds of $\heff1$.
-Here, we present results on error, bias, empirical standard error, and confidence interval coverage; results on root-mean-squared-error %$\left(\widehat{bias}^2+SE_{emp}^2\right)^{1/2}$
-$\left(\sum_b err_b^2/500\right)^{1/2}$
-are available in the online appendix.
-
-\subsubsection{Varying $n$ and $\alpha$}
-% When $\alpha=0$, Assumption \ref{ass:vps} was violated, and \textsc{geepers} estimates were highly erratic; however, as Table \ref{tab:coverage} shows, when other assumptions held, inference remained valid.
-Our first set of results illustrates the performance of \geepers when all of its assumptions are met.
-Since \geepers is an asymptotic method whose identification depends on $\bm{x}$ as a predictor of $\st$, we examine its bias and standard error as functions of $n$ and $\alpha$.
-We include analogous results for \pmm as a baseline comparison.
-
-Figure \ref{fig:alphan} shows the empirical bias and standard error (SE) of the \textsc{geepers} and \pmm estimators of $\eff1$ as sample size and $\alpha$ varied.
-Across $n$ and $\alpha$, \textsc{geepers} had lower bias and higher SE than the mixture estimator.
-\textsc{geepers} was positively biased for low $n$ with bias converging to close to 0 (up to simulation error) for $n\ge 300$, and was roughly unbiased (up to simulation error) for $\alpha \ge 0.2$, with $n$ fixed at 500.
-The SE of the \textsc{geepers} decreased rapidly for $100\le n \le 500$ and $0.2\le \alpha \le 0.5$ and more gradually for higher values of $n$ and $\alpha$.
-
-For the remainder of the simulation, we focus on the moderate cases of $n=500$ and $\alpha=0.2$ or 0.5.
-
-\begin{figure}[!ht]
-  \centering
-  \includegraphics[width=4.5in]{../simFigs/biasSEbyB1N.jpg}
-  \caption{Empirical bias and standard error of \textsc{geepers} and Mixture estimates of $\eff1$ as sample size per treatment group $n$ or $\alpha$ varied.
-There were 500 replications for each $n$ or $\alpha$.
-Other factors were held fixed at the noted values.}
-  \label{fig:alphan}
-\end{figure}
-
-% \begin{figure}
-%   \centering
-%   \includegraphics{../simFigs/biasSEbyB1.jpg}
-%   \caption{Empirical bias and standard error of M- and Mixture estimates of $\muc1$ as $\alpha$ varies from 0.1 to 1.
-There were 500 replications for each $n$.
-Other factors were held fixed at the noted values.}
-%   \label{fig:alpha}
-% \end{figure}
-
-\subsubsection{Performance when Assumptions do not Hold}
-
-Our next set of results compares the performance of \geepers, \pmm, and \psw when one or more assumptions of each method may be violated.
-
-\begin{figure}[!ht]
-  \centering
-  \includegraphics[width=5.5in,clip]{../simFigs/boxplotsNew.pdf}
-  \caption{Violin and scatter plots of 500 simulation estimation errors for estimators described in Section~\ref{sec:simMods} under varying conditions.
-For all plots, $n=500$; there were 500 replications for each set of conditions.
-Annotations indicate the number of extreme outliers excluded from the plotting area.}
-  \label{fig:boxplots}
-\end{figure}
-
-%<<<<<<< HEAD
-Figure \ref{fig:boxplots} shows violin plots, layered on top of jittered scatter plots, showing estimation error for \geepers, the parametric mixture model \pmm, and the \psw estimator of $\eff1$.
-Table \ref{tab:coverage} shows empirical coverage of nominal 95\% confidence interval estimates for \geepers and \pmm; when at least one of an estimator's assumptions is violated, the result is colored red.
-
-The first column of Figure \ref{fig:boxplots} and the black entries of Table \ref{tab:coverage} show results from scenarios in which all assumptions of \geepers and \pmm are met (\psw relies on principal ignorability, which is violated throughout the simulation study).
-\textsc{geepers} is slightly less biased than the parametric mixture model; when $\alpha=0.2$, \textsc{geepers} is much more variable and $\alpha=0.5$ the difference in variance is slight.
-\textsc{psw} has by far the lowest variance and the highest bias; the latter is due to the violation of principal ignorability. 
-When all assumptions are met, confidence intervals from both \geepers and \pmm achieve their nominal levels.
-
-The remainder of the results reflect violations modeling assumptions for either \geepers or the parametric mixture model or both: when $\alpha=0$, violating Assumption \ref{ass:vps}; interactions in the data generating model between covariates $X$ and treatment assignment $Z$ and/or principal stratum $S$, either of which would violate assumption \ref{ass:rci} and the linear additive model of the \textsc{pmm}; and/or residuals drawn from a uniform distribution, which violates the normality assumption of \textsc{pmm}.
-
-We omitted the $\alpha=0$ case from Figures \ref{fig:alphan} and \ref{fig:boxplots} since the variance of \geepers and \pmm estimates was so large as to dominate the plot. 
-However, the first column of results in Table \ref{tab:coverage} shows that, when other assumptions are met, confidence intervals from both methods tend to over-cover when $\alpha=0$. 
-In a randomized experiment, $\alpha$ can be estimated directly, making Assumption \ref{ass:vps} testable.
-
-For \geepers, when $\alpha=0.2$ the presence of interactions in the data-generating model increases sampling variance, but only interactions between $X$ and $Z$ reduce confidence interval coverage.
-When $\alpha=0.5$, the presence of interactions has little to no effect on the performance of the \geepers estimator, and all 95\% confidence intervals approximately achieve their nominal levels.
-\geepers makes no assumptions about the shape of the residual distribution and behaves similarly under either normal or uniform residuals.
-
-Like \geepers, \pmm is insensitive to interactions between $X$ and $S$.
-However, interactions between $X$ and $Z$ in the data-generating model skew the \pmm sampling distribution and reduce the coverage of its 95\% confidence intervals, even when $\alpha=0.5$.
-When the residual distribution was uniformly distributed, the \pmm sampling distribution appeared bimodal, and its credible intervals exhibited severe undercoverage.
-
-\label{simsum}In sum, \psw is the least variable of all three estimators, and does not depend on outcome modeling, but is vulnerable to bias from an unmeasured covariate that predicts both $S$ and $Y$.
-When all assumptions of both methods are met, \geepers is less efficient, but also less biased, than \pmm, but differences in efficiency decrease with $n$ and $\alpha$.
-Neither \psw nor \pmm is vulnerable to interactions between $X$ and $S$.
-\pmm is highly vulnerable to violations of normality and to interactions between $X$ and $Z$, while \psw is robust to non-normality and robust to $X$-$Z$ interactions when $\alpha = 0.5$.
-These results suggest that researchers who are willing to assume principal ignorability should use \psw, and those who are confident in their parametric model for potential outcomes should use a parametric mixture model.
-Researchers who are hesitant to accept principal ignorability or a parametric outcome model should consider \geepers.
-
-\label{evalMetric}The simulation study suggests that none of the three methods evaluated---\geepers, \pmm, or \psw---dominates the others, and offers some guidance on when each estimator may be preferable.
-\psw is the most precise and requires no assumptions about the distribution of outcomes, but can be quite sensitive to violations of principal ignorability.
-When the determinants of $\st$ are well-understood and measured, \psw seems to be the best option.
-When principal ignorability does not hold, potential outcomes are conditionally normal, and there are no $X:Z$ interactions within principal strata, the \pmm estimator outperforms \geepers, though the difference decreases with $n$ or $\alpha$.
-\geepers may be the best choice when conditional normality may not hold, or if there are $X:Z$ interactions and $\bm{x}$ is sufficiently predictive of $\st$ (AUC $\ge$0.675, corresponding to $\alpha \ge 0.5$, may serve as a threshold).
-These suggestions may depend heavily on the simulation setup; applying them in general scenarios requires further research.
-
-%=======
-% Figure \ref{fig:boxplots} shows violin plots, layered on top of jittered scatter plots, showing estimation error for \textsc{geepers}, the mixture estimator, and the \textsc{psw} estimator of $\eff1$ under various sets of conditions, and Table \ref{tab:coverage} shows empirical coverage of nominal 95\% confidence interval estimates under a somewhat wider set of circumstances when $\muc1=0.3$.
-More complete results, including root-mean-squared-error estimates, are in the online appendix.
-% The leftmost panel of Figure \ref{fig:boxplots} and the top line of Table \ref{tab:coverage} show results under conditions favorable to both the mixture model and to \textsc{geepers}---normal residuals and no interactions in the data generating model between covariates and either $S$ or $Z$.
-% \textsc{geepers} was roughly unbiased, but less precise than the other two methods---when $\alpha=0.2$ it was much noisier, and when $\alpha=0.5$ it was only slightly less precise.
-% The mixture estimator's bias depended on $\muc1$---when $\muc1=0$, the bias was slight, and when $\muc1=0.3$, the bias was more substantial.
-This result is somewhat surprising, because when $\muc1=0$, there was no separation between the principal strata in the control group, and previous literature \citep{griffin2008application} has identified this scenario as a particular challenge for mixture modeling.
-% The \textsc{psw} estimator was also biased, due, presumably, to the violation of Principal Ignorability, Assumption \ref{ass:PI}, inherent in the data generating model.
-On the other hand, the \textsc{psw} estimator had easily the lowest sampling variance of the three estimators.
-
-% The middle panel of Figure \ref{fig:boxplots} and the fifth line of Table \ref{tab:coverage} show results for when the residuals in \eqref{eq:y-sim} are uniformly distributed.
-% In this case, \textsc{geepers} and \textsc{psw} estimators, which do not assume normality, behaved in a similar fashion as in the normal case, but the mixture estimator showed a bimodal pattern--sometimes over-estimating $\eff1$ and sometimes underestimating, but rarely estimating $\eff1$ accurately.
-% Credible intervals from the mixture model exhibited severe undercoverage, while \textsc{geepers} confidence intervals achieved at least their nominal level.
-
-% %\sloppy
-% The rightmost panel of Figure \ref{fig:boxplots} returns to the case of normal residuals---and only shows results when $\muc1=0.3$ and $\alpha=0.5$---but introduces interactions in the data generating model between covariates and either $\st$, $Z$, or both, violating Assumption \ref{ass:rci}.
-% %Presence of interactions could undermine the \textsc{geepers} by inducing a relationship between principal scores and potential outcomes, even after adjusting the outcomes for covariate effects.
-% Fortunately, these interactions made little difference when $\alpha=0.5$.
-However, Table \ref{tab:coverage} shows that when $\alpha=0$ or 0.2, and interactions between $\bx$ and $Z$ were present (i.e., $\gamma_3\ne 0$), \textsc{geepers} 95\% confidence intervals tended to under-cover.
-Nevertheless, coverage of \textsc{geepers} intervals was nearly always substantially higher than coverage of corresponding 95\% credible intervals from mixture models.%\footnote{}
-% >>>>>>> overleaf-2025-06-12-1948
-
-\begin{table}%[!ht]
-
-\caption{\label{tab:coverage}Empirical coverage of nominal 95\% confidence intervals for estimates of $\tau^1$.}
-\centering
-% \begin{tabular}[t]{lllllll}
-% \toprule
-% \multicolumn{3}{c}{ } & \multicolumn{2}{c}{$\alpha=0.2$} & \multicolumn{2}{c}{$\alpha=0.5$} \\
-% \cmidrule(l{3pt}r{3pt}){4-5} \cmidrule(l{3pt}r{3pt}){6-7}
-% \makecell[l]{Residual\\Dist.} & \makecell[l]{X:Z\\Int.?} & \makecell[l]{X:S\\Int.?} & \geepers & \pmm & \geepers & \pmm\\
-% \midrule
-%  & No & No & 0.97 & 0.97 & 0.96 & 0.96\\
-
-%  & Yes & No & \rd{0.88} & \rd{0.74} & \rd{0.94} & \rd{0.89}\\
-
-%  & No & Yes & \rd{0.98} & \rd{0.97} & \rd{0.97} & \rd{0.96}\\
-
-% \multirow{-4}{*}{\raggedright\arraybackslash Normal} & Yes & Yes & \rd{0.9} & \rd{0.77} & \rd{0.94} & \rd{0.88}\\
-% \cmidrule{1-7}
-%  & No & No & 0.97 & \rd{0.4} & 0.95 & \rd{0.61}\\
-
-%  & Yes & No & \rd{0.87} & \rd{0.2} & \rd{0.93} & \rd{0.48}\\
-
-%  & No & Yes & \rd{0.98} & \rd{0.53} & \rd{0.96} & \rd{0.57}\\
-
-% \multirow{-4}{*}{\raggedright\arraybackslash Uniform} & Yes & Yes & \rd{0.88} & \rd{0.21} & \rd{0.94} & \rd{0.53}\\
-% \bottomrule
-% \multicolumn{7}{p{3.5in}}{\footnotesize Based on 500 replications.
-$n=500$, $\beta_1=0.3$.
-Simulation standard error $\approx 1$ percentage point.
-Estimates colored \rd{red} indicate cases where the assumptions of the model are not met.}
-
-% \end{tabular}
-\input{coverageTab}
-\end{table}
-
-\section{Applications}
-To illustrate \textsc{geepers} and compare it to the other discussed methods, we include two applied examples.
-First, we briefly replicate an analysis from \citet{richardson2023estimating} using the \textsc{geepers} and the other methods discussed above.
-This is a relatively straightforward example of an experiment with imperfect compliance in which \citet{richardson2023estimating} estimated statistically significant effects overall in both principal strata.
-Next, we present a novel analysis of data from the RCT of computer-based math applications reported in \citet{impactPaper}, and look into the role of bottom-out hints, as described in the introduction.
-This analysis aims to shed light on an important question in education research, illustrate model and covariate selection, and include a discussion on the plausibility of identification assumptions.
-
-\subsection{Accounting for Imperfect Compliance: Re-Analysis of the Obstetrics and Periodontal Therapy Study}
-\sloppy
-Periodontal disease, a weakening of bone and connective tissue supporting teeth, is highly prevalent among American adults and is associated with adverse pregnancy outcomes, including preterm birth \citep{periodontalEpi}.
-The Obstetrics and Periodontal Therapy Study \citep[(OPTS)][]{michalowicz2006treatment} investigated the efficacy of nonsurgical periodontal treatment to reduce the preterm birth rate; replication data are available at \url{https://causeweb.org/tshs/obstetrics-and-periodontal-therapy/}.
-Pregnant patients at four different hospitals who were at least 16 years old, at less than 16 weeks and 6 days of gestation, and diagnosed with periodontal disease were individually randomized to either receive periodontal treatment during the first 21 weeks of pregnancy ($Z=1$, $n_T=413$) or after delivery ($Z=0$, $n_C=410$).
-The primary prespecified outcome of the study was the babies' gestational age at birth; however, the study team also collected data on several secondary birth outcomes, as well as measures of patients' periodontal disease at follow-up.
-
-The main component of the intervention consisted of up to four visits, followed by monthly tooth polishing and other dental care as needed.
-However, not all participants randomized to the treatment condition completed the entire treatment plan.
-However, 96\% of patients in the treatment group received at least some periodontal care during pregnancy---therefore, there is reason to suspect nonzero treatment effects even for those patients who did not (or would not) fully comply.
-Patients randomized to the control condition did not have access to the intervention.
-\citet{richardson2023estimating} used a method they refer to as ``bespoke instrumental variable'' (BSIV) to estimate the average effects for the principal stratum of subjects who, if randomized to treatment,  would complete the treatment plan and for the principal stratum of subjects who would not.
-Here, we will use \textsc{geepers}, a Bayesian mixture model, and principal score weighting to estimate the same quantities, following the same analysis decisions as \citet{richardson2023estimating}.
-Replication code for our analysis can be found at \url{ https://github.com/adamSales/psGee}.
-
-Following \citep{richardson2023estimating}, we will estimate the principal effects of randomization $Z$ on $Y$, the percentage of probed sites (6 sites per tooth) that were bleeding at visit 5, which was at 29--32 weeks gestation, conditioning on two covariates, Serum measures of endotoxin level and fibrinogen at baseline.
-We omit the 164 cases in which $Y$ was missing as well as 19 additional cases in which one or both of the covariates were missing, leaving a total sample size of 640, with 314 randomized to the treatment condition and 326 randomized to control.
-The compliance variable in the dataset had three categories: ``Yes,'' for subjects in the treatment group who completed the treatment plan (157 subjects), ``No'' for those who did not (4 subjects), and ``Und'' for those who received some therapy, but it is unknown whether they completed the treatment (157 subjects).
-We let $S=1$ when compliance was ``yes''---i.e., when subjects in the treatment group completed the treatment---and $S=0$ otherwise.
-
-For all three estimation methods, we estimated principal scores with logistic regression of $\st$ on the two covariates.
-For \textsc{geepers} and \textsc{psw}, we used standard maximum likelihood methods and data from the treatment group, achieving an AUC of 0.68.
-For the mixture model, we incorporated the logistic regression specification as a sub-model within a larger Bayesian model fit in Stan.
-
-\begin{figure}
-  \centering
-  \includegraphics{figure/opts.pdf}
-  \caption{Estimated principal effects and 95\% confidence or credible intervals for the Obstetrics and Periodontal Therapy Study, using four estimation methods.
-Results for BSIV were taken from \citet{richardson2023estimating}.}
-  \label{fig:opt}
-\end{figure}
-
-The results are displayed in Figure \ref{fig:opt}% and Table \ref{tab:optEffs}
-: both \textsc{geepers} and the Bayesian mixture model gave similar results to those reported in \citet{richardson2023estimating} from BSIV.
-For the principal stratum of subjects who would not fully complete the treatment, those three methods estimated that assignment to the treatment condition reduced percentage of dental sites with bleeding by 31--34 percentage points; for subjects who would complete the treatment, the three estimated a reduction of 14-16 percentage points.
-The conclusion of these methods is somewhat surprising: that the effect was larger for subjects who would not complete the treatment than for those who would (the two strata had similar bleeding percentages at baseline, and a logit transformation of the outcome leaves the results qualitatively unchanged).
-One possible explanation is the conjecture that subjects stopped attending treatment once their periodontal disease improved.
-
-In contrast, principal score weighting resulted in very similar effects for both strata---a reduction of 23--25 percentage points, which is close to the estimated intent-to-treat effect of 25 percentage points.
-
-\subsection{Estimating Treatment Effects for Implementation Subgroups: Students who Frequently or Rarely Request Bottom-Out Hints}\label{sec:fh2t}
-
-In the educational field experiment reported in \citet{impactPaper}, middle school students completing their math schoolwork on a computer program were randomly assigned to four conditions.
-These included two gamified programs---DragonBox and From Here to There (FH2T)---and two programs in which students used computers to work through a series of algebra problems taken from the district's textbook---``ASSISTments'' and an active control called ``Business as Usual'' (BAU).
-% the active control condition in which students simply used the computer to work on a sequence of math problems.
-In the DragonBox game, students learned to solve algebraic equations by isolating a box containing a dragon; as students progressed in the game, pictures of monsters were replaced by mathematical symbols.
-Another condition, the From Here to There (FH2T) program, used visual cues such as color and spacing to teach algebraic concepts and allowed students to solve equations by manipulating expressions via dragging and dropping.
-%In ASSISTments, students used computers to work through a series of math problems taken from the district's textbook.
-Students in the ASSISTments condition had access to a series of hints for each problem, culminating in a ``bottom out'' hint that contained the answer.
-They received immediate feedback on their answers and had to enter the correct answer to proceed to the next problem.
-In the BAU condition, students had no access to hints and received feedback only after completing the assignment.
-
-In all four conditions, students worked on their assigned math program in class on 12 different occasions spaced throughout the school year: a prior assessment, a mid-test, a post-test, and nine learning sessions.
-In this illustration, we will compare ASSISTments---the only condition featuring bottom-out hints---with each of the three others separately.
-Our goal is to estimate different treatment effects for groups of students who, if assigned to ASSISTments, would request few or many bottom-out hints.
-
-The ability to bottom out was only one of several differences between the treatment conditions.
-One way to disentangle the role of bottoming out in any treatment effect would be to estimate average treatment effects of being assigned to ASSISTments, relative to any of the other three, separately for subjects who would (if assigned to ASSISTments) bottom out frequently, and for subjects who would not.
-If the availability of bottom-out hints plays a positive role in ASSISTments' efficacy, we might expect larger effects of the condition on students who frequently bottom out than on those who do not; if bottom-out hints are harmful, we might expect the opposite.
-If hints (that do not supply the correct answer) and error messages are helpful, but the ability to bottom out is harmful to learning, then we might expect to estimate a positive effect of being assigned to the feedback condition for students who do not bottom out, but a negative effect for students who do.
-If hints are helpful for learning only if they can be read as worked examples, then we might expect no effect of condition on students who generally do not bottom out, but a positive effect for students who do.
-(These hypothetical results may be masked by other treatment effect heterogeneity between students who would or wouldn't bottom out, but the results could still be informative.)
-
-\subsubsection{Principal Stratification and Alternatives}\label{sec:psInEdTech}
-The goal of principal stratification in this context is to determine if assignment to the ASSISTments condition affected students who tended to request frequent bottom-out hints differently than students who tended to request fewer bottom-out hints.
-Is this the right question to ask?
-
-In a similar context, \citet{sales2021student} suggests two alternative approaches: the average effect of requesting bottom-out hints on posttest scores for students in the ASSISTments condition, and estimating natural direct and indirect effects of assignment to the ASSISTments condition, treating bottom-out hint requests as a mediator.
-The former approach is an observational study: although students were randomized to the ASSISTments condition, bottom-out hint requests were not randomized.
-Therefore, estimating the effect of bottom-out hints requires careful control of confounding.
-Unfortunately, confounding is a major concern in this case, since, presumably, students only requested bottom-out hints if they were struggling with the material, and students who struggled with practice problems are likely to score lower on the posttest, on average, than those who needed less help during practice.
-The best available measure of students' baseline ability, their score on the 10-item pretest, is likely insufficient.
-That said, a plausible estimate of the average effect of bottom-out hint requests would be a valuable contribution to the literature, and a suitably innovative observational design \citep[e.g.][]{beck} may bear fruit.
-
-Another arguable disadvantage of the observational study approach is that it overlooks the broader context of the experiment, focusing solely on whether bottom-out hints affect posttest scores rather than examining their role in the overall effectiveness of ASSISTments.
-A second approach, mediation analysis, contextualizes the bottom-out effect within the larger experiment.
-It, too, relies on the same non-confounding assumption as an observational study, and also suffers from an additional challenge: mediation analysis relies on a ``common support'' condition that ``the mediator state must not be a deterministic function of the treatment'' \citep[p.
-225]{celli2022causal}.
-However, bottom-out hints are unavailable in the other three arms of the study, violating this condition (\citealt{sales2021student} offers a solution, which we find unconvincing).
-
-A principal stratification approach treats bottom-out hint requests as a type of compliance.
-In other words, if we consider full implementation of the ASSISTments condition to include frequent bottom-out hint requests, students assigned to ASSISTments but who requested few bottom-out hints only partially complied with their treatment assignment.
-The question then is whether students who do not fully comply with their treatment assignment still experience an effect.
-Conversely, might students who only partially comply experience a larger positive effect?
-That would be the case if either bottom-out hints were harmful or if the treatment was better targeted toward the type of student who tends to need them less.
-
-Principal score weighting depends on the assumption of principal ignorability, which is equivalent to the no-confounding assumption of the observational study and mediation analysis.
-The mixture model and \textsc{geepers} estimators do not rely on that assumption, instead relying on parametric models and residualized covariate independence, respectively.
-Finally, the instrumental variables estimator of \citet{air} relies on the ``exclusion restriction'' that, for students who request few bottom-out hints, there is no effect of randomization to the ASSISTments condition.
-Since bottom-out hints are only one of several features of ASSISTments, this assumption is untenable.
-
-\subsubsection{Data on Bottom-Out Hints and Covariates}
-
-\sloppy
-We gathered dichotomized data on %which students in the treatment condition requested at least one bottom-out hint.
-whether students in ASSISTments requested more than the median number of bottom-out hints (i.e., 11).
-That is, $\st=1$ for students who, if assigned to ASSISTments, would request more than 11 bottom-out hints (``bottom-outers''), and $\st=0$ for students who would request 11 or fewer (``non-bottom-outers'').
-%least one bottom-out hint, and $\st=0$ for students who would not.
-The outcome of interest $Y$ in our example is students' total scores on a ten-item posttest completed within the online tutoring system; $Y$ is an integer ranging from 0 to 10.
-The goal of our analysis will be to determine the principal effects $\eff1=\EE[\yt-\yc|\st=0]$ and $\eff0=\EE[\yt-\yc|\st=0]$, the average effects of assignment to ASSISTments (``$T$'' subscripts) versus the other conditions (``$C$'') for bottom-outers and non-bottom-outers.
-
-We also had access to data on several baseline covariates, including prior achievement, demographics, baseline measurements of student attitudes toward math, and an indicator for whether students began the school year in remote or in-person instruction.
-Students with missing pretest scores were dropped from the analysis; missing data in other covariates were imputed with a Random Forest algorithm \citep{missForest}.
-Summary statistics are displayed in appendix tables \ref{table:tab1fac} and \ref{table:tab1num}.
-
-%The data analysis we present here is intended as a demonstration of \textsc{geepers}, rather than for its substantive conclusions, and our description omits discussion of some important methodological considerations, including attrition and post-selection inference.
-For details on the experimental design, the conditions being compared, sample attrition---which was unusually high due to the COVID-19 pandemic---and the impact analysis, please see \citet{impactPaper}.
-
-For a description of the dataset and an explanation of how to access it, see \citet{ottmar2023data} and \url{https://osf.io/r3nf2/}.
-Replication code for the analysis is available at %[Redacted].
-\url{https://github.com/adamSales/psGee}.
-
-\subsubsection{Data Analysis}
-We estimated principal scores using logistic regression of observed $S$ on school fixed effects and a set of baseline variables $\bxs$.
-Three criteria governed our choice of covariates and regression specifications for estimating principal scores and effects.
-First, Proposition \ref{prop:reg2} requires consistent estimation of principal scores and (via Assumption \ref{ass:rci}) correct specification of the outcome model, that is, the model for $\EE[Y|\bxy,Z,\st]$.
-Hence, it is essential to choose covariates and transformations $\bxs$ and $\bxy$ that lead to a good model fit.
-
-Second, Assumption \ref{ass:vps} requires principal scores to vary with at least some element $\bxs$.
-Moreover, the results of the simulation study suggest that when $\st$ can be predicted more precisely as a function of $\bxs$, \textsc{geepers} principal effect estimators are more precise and more robust to model misspecification.
-Those considerations militate in favor of choosing covariates $\bxs$ to optimize out-of-sample principal score prediction accuracy.
-
-There is good reason to believe that analogous arguments apply to the selection of outcome covariates $\bxy$.
-
-The first two criteria motivated our choice of two strategies for choosing $\bxs$ and $\bxy$.
-In our main model, we included every available covariate in the dataset and chose transformations of those covariates based on the results of residual and binned residual plots \citep{arm}.
-
-Including all possible covariates maximizes the model's opportunities to detect covariate relationships that would predict $Y$ or $\st$.
-However, this strategy can also lead to overfitting, which would decrease out-of-sample predictive accuracy.
-In an alternative analysis, we chose a subset of covariates to include in $\bxs$ and $\bxy$ with a backward stepwise selection algorithm minimizing the AIC \citep{aic} of the principal score model and then modified the resulting models based on residual plots.
-The model selection process is recorded in the GitHub repository cited above.
-
-Lastly, Assumption \ref{ass:ci} implies that there is no interaction between $\bxs$ and $\st$ in the outcome model. 
-\emph{A priori}, the student attribute that poses the greatest threat to these modeling choices is prior achievement, that is, scores on the pretest and the fifth-grade state test.
-This is both because they measure a similar construct to the posttest and because \cite{impactPaper} found a pretest-treatment interaction in a different treatment contrast.
-Following that reasoning, we conducted a third parallel analysis excluding prior achievement covariates. % in $\bxs$ but including all covariates in $\bxy$.
-
-All three principal score models were fit to data from students in the ASSISTments group, for whom bottom-out status ($S$) was observed; hence, the same model fit was used for all three treatment contrasts.
-
-Outcome models for all analyses included all variables from the corresponding principal score models, but substituted class fixed effects for school fixed effects.\footnote{Why use a coarser set of fixed effects for the principal score model than the outcome regression?
-First, the principal score model was fit to a smaller sample (only students in the ``Instant'' condition), so some classrooms only included one student, which can lead to poor fit.
-Second, the asymptotics of logistic regression are more restricted than OLS, so logistic regression tends to perform poorly when the number of parameters increases with the sample size \citep{agresti}.}
-As with principal score models, we checked model fit with residual plots and modified model specifications accordingly.
-
-% \begin{figure}
-%   \centering
-%   \includegraphics{../figure/psModCoef.jpg}
-%   \caption{Principal Score Model Coefficients}
-%   \label{fig:psMod}
-% \end{figure}
-
-\subsubsection{Results}
-
-The three principal score models were fairly successful in distinguishing bottom-outers from non-bottom-outers---the AUC, evaluated using out-of-sample predictions in a 10-fold cross-validation, was roughly 0.795 for the AIC-optimal model, and about 0.745 for the two others.
-That is, in roughly 75--80\% of bottom-outer/non-bottom-outer pairs of subjects, the bottom-outer will have a higher predicted probability from the model.
-Coefficient estimates are displayed in Table \ref{tab:psTab} in the appendix.
-
-Figure \ref{fig:effects} shows principal effect point estimates and 95\% confidence intervals from analyses based on the three principal score models.
-The three covariate selection strategies led to similar point and interval estimates.
-That said, in nearly all cases, the AIC-optimal model led to the smallest standard errors.
-Confidence intervals for all nine principal effect estimates were about twice as wide as for corresponding ATE estimates.
-In no case was there evidence at $\alpha=0.05$---or any other conventional level---for non-zero principal effects or differences between principal effects.
-Table \ref{tab:regTab}, in the appendix, displays the full set of regression coefficients for the outcome models based on the "All Covariates" principal score model, along with nominal standard error estimates.
-
-A comparison of point estimates suggests that assignment to ASSISTments rather than BAU benefits bottom-outers less than non-bottom-outers.
-We speculate that, if true, this pattern may be because many students use bottom-out hints to game the system---entering the provided correct answers rather than attempting to figure out the problems on their own.
-Removing access to hints for bottom-outers forces them to attempt the problems rather than game the system.
-At the same time, hints and instant feedback are helpful for non-bottom-outers, who are less likely to game the system.
-
-On the other hand, point estimates suggest that assignment to ASSISTments, relative to FH2T, hurts non-bottom-outers but not non-bottom-outers.
-Perhaps non-bottom-outers tend to have higher prior knowledge, and, as \cite{impactPaper} shows, FH2T benefits such students more than students with lower prior knowledge.
-
-Perhaps the simplest explanation for observed differences in point estimates is sampling error, which cannot be ruled out.
-
-\begin{figure}
-  \centering
-  \includegraphics{figure/prinEffs.pdf}
-  \caption{Principal effect estimates with 95\% confidence intervals using three different principal score models, alongside ATE estimates and 95\% confidence intervals, for contrasts between the ASSISTments condition and each of the three other conditions in the study.}
-  \label{fig:effects}
-\end{figure}
-
-\subsubsection{Comparing Principal Stratification Methods}
-
-\begin{figure}
-  \centering
-  \includegraphics{figure/compareMethods.pdf}
-  \caption{Principal effect estimates with 95\% confidence or credible intervals using three different principal stratification methods, alongside ATE estimates and 95\% confidence intervals, for contrasts between the ASSISTments condition and each of the three other conditions in the study.
-The \textsc{geepers} estimates used the ``All Covariates'' principal score model.}
-  \label{fig:compare}
-\end{figure}
-
-Figure \ref{fig:compare} shows principal effect estimates from \textsc{geepers} (using the All Covariates principal score model) compared to analogous estimates using Bayesian mixture modeling and \textsc{psw}, alongside estimates of the overall average treatment effects, labeled ``ATE.''
-
-The \textsc{psw} model used the same principal score model as \textsc{geepers}, while the Bayesian model used the same principal score and outcome model specifications.
-
-The Bayesian mixture model assumed that outcomes were normally distributed, conditional on covariates and principal strata, with a residual standard deviation that varied between principal strata.
-This modeling assumption is necessarily false since post-test scores were integers between 0 and 10.
-All model parameters with support in $\mathbb{R}$ were given standard normal priors, while standard deviation parameters were given half-standard normal priors.
-The principal score and outcome models were fit simultaneously using a Markov Chain Monte Carlo algorithm \citep{rstan}.
-
-The \textsc{psw} estimators did not include any covariate outcome modeling. Standard errors for the \textsc{psw} estimates were estimated using the bootstrap, with resampling done within schools.
-
-For all three sets of estimators, approximate 95\% confidence intervals were estimated by adding and subtracting twice the standard error from the point estimate (or twice the posterior standard deviation from the posterior mean, in the Bayesian mixture model).
-
-The other estimation strategies largely agreed with \textsc{geepers} and with each other in the contrast between ASSISTments and BAU, and, to a lesser extent, in the contrast between ASSISTments and DragonBox.
-The biggest difference between the methods was in the contrast with FH2T, in which the Bayesian mixture model resulted in a larger difference between the two principal effects, as well as the largest standard errors, while the two \textsc{psw} principal effect estimates were quite similar.
-However, these differences are difficult to interpret due to the large uncertainty surrounding all the point estimates.
-
-\section{Discussion}\label{sec:discussion}
-\textsc{geepers} is a straightforward approach to principal effect estimation under strong monotonicity, built on widely-used regression models, which is more robust---though sometimes less precise---than alternative approaches under a wide array of scenarios.
-
-There is good reason to hope that extensions to \textsc{geepers}, including cases in which $S$ takes more than two values, may be straightforward.
-For instance, in truncation by death problems \citep[e.g.][]{zhangRubin,ding2011} with weak monotonicity, $S_i=1$ if participant $i$ survives (or, more generally, if the outcome $Y_i$ is measured) and 0 otherwise, interest is typically in the principal effect for the ``always survive'' principal stratum in which $\sti=S_{ci}=1$.
-If, say, $\sti\ge S_{Ci}$, then every subject in the control condition with $S_i=1$ is in the always survive stratum, while those subjects in the treatment condition with $S_i=1$ are a mixture of the always survive and $\sti=1;\;S_{Ci}=0$ stratum.
-This scenario is broadly similar to the strong monotonicity, one-way noncompliance situation discussed in this paper.
-On the other hand, when no monotonicity assumption holds, both $\st$ and $S_C$ will have to be imputed for every subject; further research is necessary to determine the appropriate way to do so.
-
-Another direction of extension involved the principal score model \eqref{eq:pscore}.
-The performance of \textsc{geepers} in the simulation study of Section \ref{sec:simulation} depended heavily on the factor $\alpha$, which controlled the extent to which covariates could predict $S$.
-That suggests that when covariates are high-dimensional, \textsc{geepers}' performance in applications could be optimized with a high-dimensional semi- or non-parametric model.
-If so, several further questions emerge.
-First, can the parameter $\alpha$ be extended to a more general parameter measuring the prediction accuracy of a non-parametric model?
-Second, if the principal score model cannot be formulated as the solution to a set of estimating equations, how should the standard error be computed?
-Lastly, can over-fit principal score models cause bias or other estimation problems, and if so, are there ways to protect against overfitting?
-
-Sampling error prevents firm conclusions regarding principal effects for students who would, if given the opportunity, request frequent bottom-out hints.
-However, the results are consistent with the hypothesis that ASSISTments is less effective for frequent bottom-out requesters when compared to a similar program without hints or immediate feedback.
-We hypothesize that many students who frequently request bottom-out hints are gaming the system, facilitated by the availability of these hints.
-
-\textsc{geepers} is a flexible, easily-implementable method for principal effect estimation for one-way noncompliance, with predictive covariates; extensions to a broader set of circumstances could be a boon to causal modeling.
-
-\bibliographystyle{apalike}
-\bibliography{MOM}
 
@@
   \caption{Empirical coverage of nominal 95\% Confidence intervals for \geepers and \pmm when $n=500$ per condition.}
-  \input{coverageTabAppendix500.tex}
+  \input{writeUps/coverageTabAppendix500.tex}
  \end{table}
@@
   \caption{Empirical coverage of nominal 95\% Confidence intervals for \geepers and \pmm when $n=1000$ per condition.}
-  \input{coverageTabAppendix1000.tex}
+  \input{writeUps/coverageTabAppendix1000.tex}
  \end{table}
@@
   \caption{Empirical RMSE  for \geepers, \pmm, and \psw when $n=500$ per condition.}
-  \input{rmseTabAppendix500}
+  \input{writeUps/rmseTabAppendix500}
  \end{table}
@@
   \caption{Empirical RMSE  for \geepers, \pmm, and \psw when $n=1000$ per condition.}
-  \input{rmseTabAppendix1000}
+  \input{writeUps/rmseTabAppendix1000}
  \end{table}
@@
 
-\section{Additional Results from the Applications}
+\clearpage
+\section{Additional Results from the Bottoming-Out Study}
 \singlespacing

### 2025-07-02 writeUps/paper.tex
finished?? editing

@@
 \usepackage[letterpaper]{geometry}
-
+\usepackage{xr}
+\externaldocument{writeUps/appendix}
 \RequirePackage{amsmath,amsfonts,amssymb,amsthm}
@@
 \usepackage{longtable}
-%\usepackage[notablist,tablesfirst]{endfloat}% To `separate' figures and tables from text if required
 \usepackage{threeparttable}
-\usepackage{epstopdf}% To incorporate .eps illustrations using PDFLaTeX, etc.
-\usepackage[caption=false]{subfig}% Support for small, `sub' figures and tables
-\usepackage[doublespacing]{setspace}% To produce a `double spaced' document if required
-%\setlength\parindent{24pt}% To increase paragraph indentation when line spacing is doubled
-
-\usepackage[longnamesfirst,sort]{natbib}% Citation support using natbib.sty
-\bibpunct[, ]{(}{)}{;}{a}{,}{,}% Citation support using natbib.sty
+\usepackage[caption=false]{subfig}
+\usepackage[doublespacing]{setspace}
+
+\usepackage[longnamesfirst,sort]{natbib}
+\bibpunct[, ]{(}{)}{;}{a}{,}{,}%
 \renewcommand\bibfont{\fontsize{10}{12}\selectfont}% To set the list of references in 10 point font using natbib.sty
@@
 
+\usepackage{endfloat}% To `separate' figures and tables from text if required
+
 %\endlocaldefs
@@
 %\begin{aug}
- \author{
- Adam C. Sales\thanks{CONTACT Adam C. Sales. Email: asales@wpi.edu}, Kirk P. Vanacore, and Erin R. Ottmar%\\
- %\affil{Worcester Polytechnic Institute, Worcester, MA, USA}
-}
+%  \author{
+%  Adam C. Sales\thanks{CONTACT Adam C. Sales. Email: asales@wpi.edu}, Kirk P. Vanacore, and Erin R. Ottmar%\\
+%  %\affil{Worcester Polytechnic Institute, Worcester, MA, USA}
+% }
 
@@
 Sample size per condition $n$ &\multicolumn{2}{l}{$100^*,200^*,300^*,400^*,500,600^*,700^*,800^*,900^*,1000$}\\
-$\alpha$ &\multicolumn{2}{l}{$0,0.2,0.3^\dagger,0.4^\dagger,0.5,0.6^\dagger,0.7^\dagger,0.8^\dagger,0.9^\dagger,1^\dagger$}\\
-The distribution of $\epsilon_i$&\multicolumn{2}{l}{ Normal, Uniform}\\
+$\alpha$ &\multicolumn{2}{l}{$\rd{0},0.2,0.3^\dagger,0.4^\dagger,0.5,0.6^\dagger,0.7^\dagger,0.8^\dagger,0.9^\dagger,1^\dagger$}\\
+The distribution of $\epsilon_i$&\multicolumn{2}{l}{ Normal, 
+\rd{Uniform}}\\
 %$\beta_1$ & \multicolumn{2}{l}{0, 0.3}\\
 \multirow{4}{*}{Interactions in \eqref{eq:y-sim}}& None:  $\gamma_1=1/\sqrt{6};\gamma_2=\gamma_3=0$\\ 
-& $\bm{x}:\st$: $\gamma_1=3/(4\sqrt{6})$; $\gamma_2=1/(2\sqrt{6})$; $\gamma_3=0$\\
-& $\bm{x}:Z$: $\gamma_1=1/\sqrt{6};\gamma_2=0$; $\gamma_3=1/(2\sqrt{6})$\\
-& $\bm{x}:\st$ \& $\bm{x}:Z$: $\gamma_1=3/(4\sqrt{6})$; $\gamma_2=\gamma_3=1/(2\sqrt{6})$\\
+& \rd{$\bm{x}:\st$}: $\gamma_1=3/(4\sqrt{6})$; $\gamma_2=1/(2\sqrt{6})$; $\gamma_3=0$\\
+& \rd{$\bm{x}:Z$}: $\gamma_1=1/\sqrt{6};\gamma_2=0$; $\gamma_3=1/(2\sqrt{6})$\\
+& \rd{$\bm{x}:\st$ \& $\bm{x}:Z$}: $\gamma_1=3/(4\sqrt{6})$; $\gamma_2=\gamma_3=1/(2\sqrt{6})$\\
 % between $\st$ and covariates & No, Yes\\
@@
 \hline
-\multicolumn{3}{p{4.5in}}{\footnotesize Factors fully crossed except where indicated.}\\
+\multicolumn{3}{p{4.5in}}{\footnotesize Factor levels in red violate assumptions of \geepers, \pmm, or both. Factors fully crossed except where indicated.}\\
 \multicolumn{3}{p{4.5in}}{\footnotesize $*$ Only tested with $\alpha=0.5$, normal $\epsilon$, no interactions in \eqref{eq:y-sim}.}\\
@@
 For $0.5\le \alpha \le 1$, the average AUC varies from roughly 0.53 to 0.77; for $\alpha=0.5$, the average AUC was roughly 0.675.
+
+For reference, in the application of Section~\ref{sec:opts}, there were two covariates.
+In the model for principal scores, their standardized coefficients were roughly 0.46 and 0.62, and the AUC was 0.68.
+In the application of Section~\ref{sec:fh2t}, the principal score models (which were far more complex than either the simulation model or the model in Section \ref{sec:opts}) had AUCs between 0.745 and 0.795.
 
@@
 
-% \begin{figure}
-%   \centering
-%   \includegraphics{../simFigs/biasSEbyB1.jpg}
-%   \caption{Empirical bias and standard error of M- and Mixture estimates of $\muc1$ as $\alpha$ varies from 0.1 to 1.
+\begin{figure}
+  \centering
+  \includegraphics{../simFigs/biasSEbyB1.jpg}
+  \caption{Empirical bias and standard error of M- and Mixture estimates of $\muc1$ as $\alpha$ varies from 0.1 to 1.
 There were 500 replications for each $n$.
 Other factors were held fixed at the noted values.}
-%   \label{fig:alpha}
-% \end{figure}
+  \label{fig:alpha}
+\end{figure}
 
@@
 
-\begin{figure}[!ht]
-  \centering
-  \includegraphics[width=5.5in,clip]{../simFigs/boxplotsNew.pdf}
-  \caption{Violin and scatter plots of 500 simulation estimation errors for estimators described in Section~\ref{sec:simMods} under varying conditions.
+% \begin{figure}[!ht]
+%   \centering
+%   \includegraphics[width=5.5in,clip]{../simFigs/boxplotsNew.pdf}
+%   \caption{Violin and scatter plots of 500 simulation estimation errors for estimators described in Section~\ref{sec:simMods} under varying conditions.
 For all plots, $n=500$; there were 500 replications for each set of conditions.
 Annotations indicate the number of extreme outliers excluded from the plotting area.}
-  \label{fig:boxplots}
-\end{figure}
+%   \label{fig:boxplots}
+% \end{figure}
 
@@
 
-\subsection{Accounting for Imperfect Compliance: Re-Analysis of the Obstetrics and Periodontal Therapy Study}
+\subsection{Accounting for Imperfect Compliance: Re-Analysis of the Obstetrics and Periodontal Therapy Study}\label{sec:opts}
 \sloppy
@@
 
-\appendix
-
-\section*{Appendix: Proofs and Calculations}
-\subsection*{Proof for Lemma \ref{lemma:expectation}}
-As a preliminary, note that
-% \begin{align*}
-%   \EE[\st|\hpp]&=\EE\left\{\EE[\st|\pp,\hpp]|\hpp\right\}\\
-%              &=\EE\left\{\EE[\st|\pp]|\hpp\right\}\\
-%              &=\EE[\pp|\hpp]=\hpp
-% \end{align*}
-
-%Then, note
-\begin{align*}
-  \EE[Y_C|\pp]&=\EE\left\{\EE[Y_C|\pp,\st]|\pp\right\}\\
-             &=\EE\left\{\EE[Y_C|\st]|\pp\right\}\tag*{by \eqref{eq:assumption}}\\
-             &=\EE[\muc1\st+\muc0(1-\st)|\pp]\\
-             &=\muc1\pp+\muc0(1-\pp)
-\end{align*}
-
-Then we have
-\begin{equation*}
-  \begin{split}
-    \EE[Y_C]=&\EE\EE[Y_C|\pp]=\muc1\EE\pp+\muc0(1-\EE\pp)\\
-    =&\muc0+\EE\pp(\muc1-\muc0)
-    \end{split}
-\end{equation*}
-
-Next we have
-
-\begin{align*}
-  \EE[Y_C\pp]&=\EE\left\{\EE[Y_C\pp|\pp]\right\}\\
-            &=\EE\left\{\pp\EE[Y_C|\pp]\right\}\\
-            &=\EE\left\{\pp\left[\muc1\pp+\muc0(1-\pp)\right]\right\}\\
-            &=\EE[\pp]\muc0+\EE[\pp^2](\muc1-\muc0)
-\end{align*}
-
-In the treatment group, $\st$ is observed, so
-\begin{align*}
-    \EE[Y_T]=&\mut0+\EE[\st](\mut1-\mut0)\tag*{and}\\
-    \EE[\st Y_T]=&\EE[\st]\mut0+\EE[\st^2](\mut1-\mut0)
-\end{align*}
-
-Due to Assumption \ref{ass:rand} (randomization), $\EE[Y|Z=0]=\EE[Y_C]$, $\EE[Y|Z=1]=\EE[Y_T]$, $\EE[Y\pp|Z=0]=\EE[Y_C\pp]$ and $\EE[YS|Z=1]=\EE[Y_T\st]$, completing the proof.
-
-\subsection*{Proof for Proposition \ref{prop:reg1}}
-
-Replacing $\sti$ and $\pp$ in \eqref{eq:estEq0} with $\ri$, as in \eqref{eq:ri}, and replacing $\tilde{\Psi}_i$ with $\Psi_i=\begin{psmallmatrix} 1 & 0&1&0\\ 0&1&0&1\\0&0&1&0\\0&0&0&1\end{psmallmatrix}\tilde{\Psi}_i$ gives an equivalent set of estimating equations $\sum_{i=1}^{n_C}\Psi_i=\bm{0}$ with $\Psi_i=$
-\begin{equation}\label{eq:estEq1}
-\begin{pmatrix}
-    Y_i-\muc0-\ri(\muc1-\muc0)-Z_i(\mut0-\muc0)-Z_i\ri(\mut1-\mut0-\muc1+\muc0)\\
-    \ri Y_i-\ri\muc0-\ri^2(\muc1-\muc0)-Z_i\ri(\mut0-\muc0)-Z_i\ri^2(\mut0-\mut1-\muc0+\muc1)\\
-    Z_iY_i-Z_i\mut0-Z_i\ri (\mut1-\mut0)\\
-    Z_i\ri Y_i -Z_i\ri\mut0-Z_i\ri^2(\mut1-\mut0)
-
-\end{pmatrix}
-\end{equation}
-These are equivalent to the estimating equations for OLS model \eqref{eq:regression0} with $\beta_0=\muc0$, $\beta_1=\muc1-\muc0$, $\beta_2=\mut0-\muc0$, and $\beta_3=\mut1-\mut0-\muc1-\muc0$.
-Therefore, under standard OLS regularity conditions the estimated parameter vector $\bm{\hat{\beta}}$ is consistent, completing the proof.
-
-\subsection*{A Stronger Version of Proposition \ref{prop:reg2} and a Proof}
-
-\begin{prop}\label{prop:interactions}
-  Say, for $i=1,\dots,n$, principal scores $\ppi$ are generated as \eqref{eq:pscore}, with parameters $\bm{\alpha}$ identified and consistently estimable with M-estimation, and there exist $\beta_0$, $\beta_1$, $\beta_2$, $\beta_3$, $\bm{\gamma_1}$, $\bm{\gamma_2}$, $\bm{\gamma_3}$ and $\bm{\gamma_4}$ such that $\{Y_i,Z_i,\sti,\bxy_i\}_{i=1}^n$ are independent and identically distributed with
-  \begin{equation}\label{eq:interaction}
-    \begin{split}
-    \EE[Y_i|\st,Z,\bx]=&\beta_0+\beta_1\sti+\beta_2 Z_i+\beta_3Z_i\sti\\
-    &+\bm{\gamma_1}'\bxy_i+\bm{\gamma_2}'\bxy_i\sti+
-    \bm{\gamma_3}'\bxy Z_i+\bm{\gamma_4}'\bxy_i Z_i\sti
-    \end{split}
-  \end{equation}
-
-  Then, under Assumptions \ref{ass:sm}, \ref{ass:rand}, and \ref{ass:vps}, if $\ppi$ is linearly independent of $\bxy$, a researcher may follow the following procedure to estimate principal effects:
-  \begin{enumerate}
-  \item Estimate principal scores by fitting model \eqref{eq:pscore} to data from the treatment group
-  \item Replace $\sti$ with $\ri$ (as defined in \ref{eq:ri}) in model \eqref{eq:interaction} and fit with OLS
-  \item Estimate principal effects as:
-   \begin{equation}\label{eq:prinEffEstApp}
-  \begin{split}
-    \heff0_{int}&\equiv \hat{\beta}_2+\bm{\hat{\gamma}_3}'\overline{\bxy}_{Z=1,S=0}\\
-    \heff1_{int}&\equiv \hat{\beta}_2+\hat{\beta}_3+(\bm{\hat{\gamma}_3}+\bm{\hat{\gamma}_4})'\overline{\bxy}_{Z=1,S=1}
-  \end{split}
-   \end{equation}
-   where $\overline{\bxy}_{Z=1,S=0}$ and $\overline{\bxy}_{Z=1,S=0}$ are the vector of covariate sample means for the subsets of subjects with $Z=1$ and $S=0$ or $S=1$, respectively.
-  \end{enumerate}
-  Then $\heff0_{G}$ and $\heff1_{int}$ are M-estimators.
-If the estimating equations for \eqref{eq:prinEffEstApp} are each bounded by an integrable function of $\{\bm{Y},\bxy, \bm{S},\pp,\bm{Z}\}$ that does not depend on $\{\bm{\beta},\bm{\gamma}\}$, then $\heff0_{int}\rightarrow_p\eff0$ and $\heff1_{int}\rightarrow_p\eff1$ as $n\rightarrow\infty$.
-
-  If the parameter estimates of the principal score model are asymptotically normal, second partial derivatives of the estimating equations for \eqref{eq:prinEffEstApp} are bounded by an integrable function of the data for values of $\{\bm{\beta},\bm{\gamma}\}$ in a neighborhood of their probability limits, and the sandwich components of \eqref{eq:sandwich}, $A$ and $B$, exist and are finite, and if $B$ is non-singular, then $\heff0_{int}$ and $\heff1_{int}$ are jointly asymptotically normal, with a variance of the form \eqref{eq:sandwich}.
-\end{prop}
-
-Equation \eqref{eq:interaction} implies Assumption \ref{ass:rci} with $\bm{\gamma_2}=\bm{\gamma_3}=\bm{\gamma_4}=0$.
-
-\begin{proof}
-First of all, by \eqref{eq:interaction},
-\begin{equation*}
-\begin{split}
-  \EE[Y_T-Y_C|\st=0]&\\
-  =&\EE[Y|Z=1,\st=0]-\EE[Y|Z=1,\st=0]\\
-  =&\beta_2+\bm{\gamma_3}'\EE[\bxy|\st=0]
-\end{split}
-\end{equation*}
-and
-\begin{equation*}
-\begin{split}
-  \EE[Y_T-Y_C|\st=1]&\\
-  =&\EE[Y|Z=1,\st=1]-\EE[Y|Z=1,\st=1]\\
-  =&\beta_2+\beta_3+(\bm{\gamma_3}'+\bm{\gamma_4}')\EE[\bxy|\st=1]
-\end{split}
-\end{equation*}
-Furthermore, $\overline{\bxy}_{Z=1,S=0}\rightarrow \EE[\bxy|\st=0]$ and $\overline{\bxy}_{Z=1,S=1}\rightarrow \EE[\bxy|\st=1]$ as $n\rightarrow \infty$.
-
-We will show that the estimated coefficients from model \eqref{eq:interaction}, but with $R$ replacing $\st$, fit with OLS, are consistent for $\bm{\beta}$ and $\bm{\gamma}$ from \eqref{eq:interaction}.
-
-First, note that $\EE[S]=\EE\EE[S|\bx]=\EE[\pp]$ and
-\begin{align*}
-  \EE[\bx S]&=\EE[\bx S|Z=1]=\EE[\bx S|Z=0] \mbox{ (due to randomization)}\\
-  &=\EE[\bx\EE[S|\bx]|Z=0]=\EE[\bx\pp|Z=0]=\EE[\bx\pp]
-\end{align*}
-implying that, according to \eqref{eq:interaction},
-\begin{equation*}
-  \begin{split}
-    \EE[Y|\bx,Z=0]&=\beta_0+\beta_1\pp+\bm{\gamma_1}'\bxy+\bm{\gamma_2}'\bxy\pp\\
-                  &=\beta_0+\beta_1R+\bm{\gamma_1}'\bxy+\bm{\gamma_2}'\bxy R\\
-    \EE[Y|\bx,S,Z=1]&=\beta_0+\beta_2+(\beta_1+\beta_3)S+(\bm{\gamma_1}+\bm{\gamma_3})\bxy+(\bm{\gamma_2}+\bm{\gamma_4})\bxy S\\
-    &=\beta_0+\beta_2+(\beta_1+\beta_3)R+(\bm{\gamma_1}+\bm{\gamma_3})\bxy+(\bm{\gamma_2}+\bm{\gamma_4})\bxy R
-  \end{split}
-\end{equation*}
-Therefore,
-\begin{align*}
-  \EE[Y]=&\EE[Y|Z=0]+\EE[Z]\left\{\EE[Y|Z=1]-\EE[Y|Z=0]\right\}\\
-  =&\beta_0+\beta_1\EE[R]+\beta_2\EE[Z]+\beta_3\EE[ZR]+\bm{\gamma_1}'\EE[\bxy]+\bm{\gamma_2}'\EE[\bxy R]\\
-  &+\bm{\gamma_3}'\EE[\bxy Z]+\bm{\gamma_4}\EE[\bxy RZ]
-\end{align*}
-
-Analogous reasoning leads to expressions for $\EE[RY]$, $\EE[\bxy Y]$, $\EE[ZY]$,  $\EE[ZRY]$, $\EE[\bxy YZ]$, and $\EE[\bxy RZY]$.
-These, in turn, give rise to estimating equations
-\begin{equation*}
-\begin{split}
-  &\psi_i=\\ &\begin{pmatrix}
-  Y_i\\
-  \phantom{}\\
-  Y_i\ri\\
-    \phantom{}\\
-  Y_iZ_i\\
-  \phantom{}\\
-  Y_iZ_i\ri\\
-  \phantom{}\\
-  Y_i\bxy_i\\
-  \phantom{}\\
-  Y_i\bxy_i\ri\\
-  \phantom{}\\
-  Y_i\bxy_iZ_i\\
-  \phantom{}\\
-  Y_i\bxy_iZ_i\ri\\
-  \phantom{}\end{pmatrix} - \begin{pmatrix*}[l]
-    \beta_0+\beta_1R_i+\beta_2Z_i+\beta_3Z_iR_i\\
-    \quad+\left\{\bm{\gamma_1}'+R_i\bm{\gamma_2}'+Z_i\bm{\gamma_3}'+R_iZ_i\bm{\gamma_4}'\right\}\bxy\\
-    \beta_0\ri+\beta_1R_i^2+\beta_2Z_i\ri+\beta_3Z_iR_i^2\\
-    \quad+\ri\left\{\bm{\gamma_1}'+R_i\bm{\gamma_2}'+Z_i\bm{\gamma_3}'+R_iZ_i\bm{\gamma_4}'\right\}\bxy\\
-    (\beta_0+\beta_2)Z_i+(\beta_1+\beta_3)R_iZ_i\\
-    \quad+Z_i\left\{\bm{\gamma_1}'+R_i\bm{\gamma_2}'+Z_i\bm{\gamma_3}'+R_iZ_i\bm{\gamma_4}'\right\}\bxy\\
-    \beta_0Z_i\ri+(\beta_1+\beta_3)Z_iR_i^2+\beta_2Z_i\ri\\
-    \quad+Z_iR_i\left\{\bm{\gamma_1}'+\bm{\gamma_3}'+R_i(\bm{\gamma_2}'+\bm{\gamma_4}')\right\}\bxy \\
-    \beta_0\bxyt+\beta_1R_i\bxyt+\beta_2Z_i\bxyt+\beta_3Z_iR_i\bxyt\\
-    \quad {}+\left\{\bm{\gamma_1}'+R_i\bm{\gamma_2}'+Z_i\bm{\gamma_3}'+R_iZ_i\bm{\gamma_4}'\right\}\bxy\bxyt\\
-    \beta_0\ri\bxyt+\beta_1R_i^2\bxyt+\beta_2Z_i\ri\bxyt+\beta_3Z_iR_i^2\bxyt\\
-    \quad {}+\ri\left\{\bm{\gamma_1}'+R_i\bm{\gamma_2}'+Z_i\bm{\gamma_3}'+R_iZ_i\bm{\gamma_4}'\right\}\bxy\bxyt\\
-        (\beta_0+\beta_2)Z_i\bxyt+(\beta_1+\beta_3)R_iZ_i\bxyt\\
-    \quad {}+Z_i\left\{\bm{\gamma_1}'+R_i\bm{\gamma_2}'+Z_i\bm{\gamma_3}'+R_iZ_i\bm{\gamma_4}'\right\}\bxy\bxyt\\
-\beta_0\ri Z_i\bxyt+\beta_1R_i^2Z_i\bxyt+\beta_2\ri Z_i\bxyt+\beta_3Z_iR_i^2\bxyt\\
-    \quad {}+\ri Z_i\left\{\bm{\gamma_1}'+R_i\bm{\gamma_2}'+Z_i\bm{\gamma_3}'+R_iZ_i\bm{\gamma_4}'\right\}\bxy\bxyt\\
-  \end{pmatrix*}
-  \end{split}
-\end{equation*}
-with $\EE[\psi_i]=0$.
-These are the estimating equations for the regression model \eqref{eq:interaction}, with $R$ replacing $\st$, fit by OLS.
-Consistency and asymptotic normality follow from theorems 7.8.1 and 7.8.2, respectively, of \citet{boosStefanskiBook}
-\end{proof}
-
-\subsection*{Sandwich Matrix Calculations}
-
-Here we will derive the sandwich variance-covariance matrix for the \textsc{geepers} estimate without interactions between $\bx$ and either $Z$ or $\st$---i.e., with $\bm{\gamma_2}=\bm{\gamma_3}=\bm{\gamma_4}=0$ in the notation of \eqref{eq:interaction}---and estimating principal scores using a generalized linear model.
-
-We propose estimating principal effects in two stages.
-First, fit the model
-\begin{equation}\label{eq:psMod}
-  \ppi=Pr(\sti=1|\bxsi)=f(\bm{\alpha}'\bxsit)
-\end{equation}
-for some inverse link function $f(\cdot)$, where $\bxsit=[1,\bxsi]$, using (observed) values from the treatment group, and estimating $\hat{\alpha}$.
-Then let
-\begin{equation}\label{eq:ps}
-  \hat{p}_i=f(\bm{\hat{\alpha}}'\bxsit)
-\end{equation}
-for all subjects in the experiment.
-
-Finally, fit model
-\begin{equation}\label{eq:regression}
-  Y_i=\beta_0+\beta_1r_i+\beta_2Z_i+\beta_3Z_i\ri+\bm{\gamma}'\bxy_i+\epsilon_i
-\end{equation}
-to estimate $\bm{\beta}$ and hence principal effects, where $\bm{x}_i$ is a set of covariates predictive of $Y$ within principal strata.
-Let $\bm{\beta}=[\beta_0,\beta_1,\beta_2,\beta_3,\bm{\gamma}']'$.
-
-\sloppy
-Following \eqref{eq:stacked}, let $\blam(Z_i,S_i,\bxsi,\bxy_i,Y_i;\bm{\alpha},\bm{\beta})=\begin{pmatrix} Z_i\Omega(\bxsi,S_i;\bm{\alpha})\\ \Psi(\bxy_i,\bxsi,Y_i,Z_i,\ri(\bm{\alpha});\bm{\beta})\end{pmatrix}$, the stacked estimating equations of \eqref{eq:psMod} and \eqref{eq:regression}.
-Going forward, for the sake of brevity, we will write $\blam_i(\bm{\alpha},\bm{\beta})=\blam(Z_i,S_i,\bxsi,\bxy_i,Y_i;\bm{\hat{\alpha}},\bm{\hat{\beta}})$, where dependence on the data for $i$ is captured in the subscript $i$, with similar meanings for $\Psi_i(\bm{\alpha},\bm{\beta})$ and $\omega_i(\bm{\alpha})$.
-%Then let $\blamh_i=\blam_i(Z_i,S_i,\bxsi,\bxy_i,Y_i;\bm{\hat{\alpha}},\bm{\hat{\beta}})$ be the estimating equations evaluated at the estimated parameters, and let
-
-%\begin{equation*}
-%\dblam_i=\frac{\partial}{\partial [\alpha,\beta]'} \blam_i\Bigr|_{\substack{\bm{\alpha}=\bm{\hat{\alpha}}\\\bm{\beta}=\bm{\hat{\beta}}}}
-%\end{equation*}
-%the derivative matrix of $\blam$ evaluated at the estimated parameters.
-
-The variance-covariance matrix for $\bm{\hat{\alpha}}$ and $\bm{\hat{\beta}}$ can be estimated as:
-\begin{equation*}
-  \widehat{var}\left([\bm{\hat{\alpha}}',\bm{\hat{\beta}}']'\right)=A^{-1}BA^{-t}
-\end{equation*}
-where
-\begin{equation*}%\label{eq:Amat}
-  A=\sum_i \frac{\partial}{\partial [\bm{\alpha},\bm{\beta}]'} \blam_i\Bigr|_{\substack{\bm{\alpha}=\bm{\hat{\alpha}}\\\bm{\beta}=\bm{\hat{\beta}}}}
-\end{equation*}
-and
-\begin{equation*}%\label{eq:Bmat}
-  B=\sum_i \blam_i(\bm{\hat{\alpha}},\bm{\hat{\beta}})\blam_i(\bm{\hat{\alpha}},\bm{\hat{\beta}})'
-\end{equation*}
-
-Following \citet[][p. 373]{carroll2006measurement}, we %separate the estimating equations $\Psi$ into $\phi$ and $\psi$, estimating equations for models \eqref{
-can decompose the matrices into diagonal elements
-\begin{equation*}
-    \begin{split}
-        A_{1,1}&=\sum_i \partial \Omega_i/\partial \bm{\alpha}|_{\bm{\alpha}=\bm{\hat{\alpha}}}\\
-        A_{2,2}&=\sum_i\partial\Psi_i/\partial \bm{\beta}|_{\bm{\beta}=\bm{\hat{\beta}}}\\
-        B_{1,1}&=\sum_i\Omega_i(\bm{\hat\alpha})\Omega_i(\bm{\hat\alpha})'\\
-        B_{2,2}&=\sum_i \Psi_i(\bm{\hat\alpha},\bm{\hat\beta})\Psi_i(\bm{\hat\alpha},\bm{\hat\beta})'
-    \end{split}
-\end{equation*}
- that pertain to the parameter sets $\bm{\alpha}$ and $\bm{\beta}$ and the estimating equations for models \eqref{eq:psMod} and  \eqref{eq:regression}, respectively, and
- \begin{equation*}
-     \begin{split}
-         A_{21}&=\sum_i\partial\Psi_i/\partial \bm{\alpha}|_{\bm{\alpha}=\bm{\hat{\alpha}}}\\
-         B_{12}=B_{21}'&=\sum_i \Omega_i(\bm{\hat{\alpha}})\Psi_i(\bm{\hat{\alpha}},\bm{\hat{\beta}})'
-     \end{split}
- \end{equation*}
- which capture the dependence of model \eqref{eq:regression} on the parameters $\bm{\alpha}$ from \eqref{eq:psMod} and the covariance between the estimating equations of the two models.
-
- The sub-matrix $A_{12}=\sum_i \partial \Omega_i/\partial \bm{\beta}=0$, since \eqref{eq:psMod} does not depend on $\bm{\beta}$.
-
-The diagonal matrices $A_{1,1}$ and $A_{2,2}$ and $B_{1,1}$ and $B_{2,2}$ are all the typical ``bread'' and ``meat'' matrices from M-estimation of generalized linear models and OLS.
-%In practice, we use the estimates from the \texttt{sandwich} package in \texttt{R}, adjusted in two ways: first, the function \texttt{bread} actually gives $A^{-1}$, not $A$; second, we must pay careful attention to sample sizes, since the sample size for \eqref{eq:psMod} includes only treated observations (with observed $S$) and \eqref{eq:regression} contains all observations.
-Calculation of the matrices $B_{12}$ and $B_{21}$ is straightforward after vectors $\Omega_i(\bm{\hat{\alpha}})$ and $\Psi_i(\bm{\hat{\alpha}},\bm{\hat{\beta}})$ have been calculated.
-Some specialized calculation is necessary for matrix $A_{21}$.
-
-\subsection*{$A_{21}$ Matrix}
-The estimating equations for the regression \eqref{eq:regression} are
-\begin{equation}\label{eq:eeOLS}
-  \psi(Y_i,\bxy_i,\bxsi,\bm{\beta},\alpha)=X_iY_i-X_iX_i'\bm{\beta}
-\end{equation}
-Where $X_i=[1,r_i,Z_i,r_iZ_i,\bxyt_i]'$.
-In other words,
-\begin{align*}
-  \psi(Y_i,&\bxy_i,\bxsi,\bm{\beta},\bm{\alpha})=\\
-  &\left\{
-  \begin{array}{l}
-    Y_i-\left(\beta_0+\beta_1r_i+\beta_2Z_i+\beta_3Z_ir_i+\bm{\gamma}'\bxy_i\right)\\
-    r_iY_i-r_i\left(\beta_0+\beta_2Z_i+\bm{\gamma}'\bxy_i\right)-r_i^2\left(\beta_1+\beta_3Z_i\right)\\
-    Z_iY_i-Z_i\left(\beta_0+\beta_1r_i+\beta_2Z_i+\beta_3Z_ir_i+\bm{\gamma}'\bxy_i\right)\\
-    Z_ir_iY_i-Z_ir_i\left(\beta_0+\beta_2Z_i+\bm{\gamma}'\bxy_i\right)-Z_ir_i^2\left(\beta_1+\beta_3Z_i\right)\\
-    \bxy_iY_i-\bxy_i\left(\beta_0+\beta_1r_i+\beta_2Z_i+\beta_3Z_ir_i+\bm{\gamma}'\bxy_i\right)
-  \end{array}
-  \right\}
-\end{align*}
-(noting that $Z^2=Z$).
-These depend on $\bm{\alpha}$ when $r_i=p_i$, i.e., when $Z_i=0$.% or $S_i$ is missing.
-Then note that, following \eqref{eq:ps}, and letting $\eta_i=\bm{\alpha}'\bxsit$
-\begin{equation}\label{eq:derivP}
-  \frac{\partial p_i}{\partial \bm{\alpha}'}=f'(\eta_i)\frac{\partial \eta_i}{\partial \bm{\alpha}'}=f'(\eta_i)\bxsitp
-\end{equation}
-and that
-\begin{equation}
-  \frac{\partial p_i^2}{\partial \bm{\alpha}'}=2p_i\frac{\partial p_i}{\partial \bm{\alpha}'}=2f(\eta)f'(\eta_i)\bxsitp=2p_if'(\eta_i)\bxsitp
-\end{equation}
-
-Then if $r_i=p_i$,
-% \begin{align*}
-%   \frac{\partial}{\partial \bm{\alpha}'}&\psi(Y_i,\bxy_i,\bm{\beta},\bm{\alpha})=\\
-%   &\left[\begin{array}{c}
-%           -(\beta_1+\beta_3Z_i)\frac{\partial p_i}{\partial \bm{\alpha}'}\\
-%           (Y_i-\beta_0-\beta_2Z_i-\bm{\gamma}'\bm{x}_{i})\partial p/\partial\bm{\alpha}'-(\beta_1+\beta_3Z_i)\partial p^2/\partial \bm{\alpha}'\\
-%           -(\beta_1+\beta_3Z_i)\frac{\partial p_i}{\partial \bm{\alpha}'}Z_i\\
-%           Z_i\left[(Y_i-\beta_0-\beta_2Z_i-\bm{\gamma}'\bm{x}_{i})\partial p/\partial\bm{\alpha}'-(\beta_1+\beta_3Z_i)\partial p^2/\partial \bm{\alpha}'\right]\\
-%           -(\beta_1+\beta_3Z_i)\frac{\partial p_i}{\partial \bm{\alpha}'}\bxy_i
-%         \end{array}\right]\\
-%   =& \left[\begin{array}{c}
-%           -(\beta_1+\beta_3Z_i)f'(\eta_i)\bxsitp\\
-%           \left[Y_i-\beta_0-\beta_2Z_i-\bxy_i'\bm{\gamma}-2p_i(\beta_1+\beta_3Z_i)\right]f'(\eta_i)\bxsitp\\
-%           -Z_i\beta_1f'(\eta_i)\bxsitp\\
-%           Z_i\left[Y_i-\beta_0-\beta_2Z_i-\bxy_i'\bm{\gamma}-2p_i(\beta_1+\beta_3)\right]f'(\eta_i)\bxsitp\\
-%           -(\beta_1+\beta_3Z_i)f'(\eta_i)\bxy_i\bxsitp
-%     \end{array}\right]
-% \end{align*}
-
-% When $S$ is observed for all members of the treatment group, $Z_i=0$ whenever $r_i=p_i$, so the latter expression reduces to
-\begin{align*}
-  \frac{\partial}{\partial \bm{\alpha}'}&\psi(Y_i,\bxy_i,\bm{\beta},\bm{\alpha})=\\
-  & \left[\begin{array}{c}
-          -(\beta_1)f'(\eta_i)\bxsitp\\
-          \left[Y_i-X_i'\bm{\beta}-2p_i(\beta_1)\right]f'(\eta_i)\bxsitp\\
-          0\\
-          0\\
-          -\beta_1f'(\eta_i)\bxy_i\bxsitp
-    \end{array}\right]
-\end{align*}
-
-If $r_i=S_i$, $\frac{\partial}{\partial \bm{\alpha}'}\psi(Y_i,\bxy_i,\bm{\beta},\bm{\alpha})=0$.
-
-\clearpage
-
-%\processdelayedfloats
-
-\setcounter{page}{1}
-
-\begin{center}
-\large
-    \textbf{Online Appendices for ``GEEPERs: Principal Stratification using Principal Scores and Stacked Estimating Equations''}
-\end{center}
-\section{Additional Simulation Results}
-\subsection{Plot of  AUC versus $\alpha$}
-
-\begin{center}
-
-    \includegraphics[width=5.5in]{../simFigs/alphaAUC.pdf}
-
-\end{center}
-%\begin{table}
-%\caption{Coverage of nominal 95\% confidence intervals}
-\clearpage
-
-\subsection{Full Empirical 95\% Interval Coverage Results}
-The following tables give the empirical coverage of nominal 95\% intervals for \textsc{geepers} and mixture model principal effect estimates under varying data generating models.
-First we show results when $n=500$ per condition, and then when $n=1000$ per condition.
-\\
-
-\begin{table}
-  \caption{Empirical coverage of nominal 95\% Confidence intervals for \geepers and \pmm when $n=500$ per condition.}
-  \input{coverageTabAppendix500.tex}
- \end{table}
-
- \begin{table}
-  \caption{Empirical coverage of nominal 95\% Confidence intervals for \geepers and \pmm when $n=1000$ per condition.}
-  \input{coverageTabAppendix1000.tex}
- \end{table}
-
-\clearpage
-\subsection{Full RMSE Results}
-The following table gives the root mean squared error (RMSE), $\left\{\sum_b (\hat{\tau}-\tau)^2/500\right\}^{1/2}$, for \textsc{geepers}, mixture model, and principal score weighting principal effect estimates under varying data generating models.\\
-
-\begin{table}
-  \caption{Empirical RMSE  for \geepers, \pmm, and \psw when $n=500$ per condition.}
-  \input{rmseTabAppendix500}
- \end{table}
-
- \begin{table}
-  \caption{Empirical RMSE  for \geepers, \pmm, and \psw when $n=1000$ per condition.}
-  \input{rmseTabAppendix1000}
- \end{table}
-
-%\end{table}
-\clearpage
-
-\section{Additional Results from the OPT Study}
-\singlespacing
-\subsection{Summary Statistics}
-The following table gives summary statistics for covariates and post-treatment outcomes in two of the conditions from the empirical study.\\
-
-\begin{table}
-
-\caption{\label{tab:optTab1}Descriptive statistics---mean and standard deviation or count and percent---for study variables in the full OPT dataset and in the analysis sample (i.e., complete cases)}
-\centering
-\begin{tabular}[t]{lllll}
-\toprule
-\multicolumn{1}{c}{ } & \multicolumn{2}{c}{Full Data} & \multicolumn{2}{c}{Complete Cases} \\
-\cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5}
-  & Control & Treatment & Control & Treatment\\
-\midrule
-n & 410 & 413 & 326 & 314\\
-Fibrinogen & 21.3 (14.9) & 20.2 (13.3) & 21.5 (15.4) & 20 (13.1)\\
-Endotoxin & 1.8 (1.1) & 1.8 (1.1) & 1.8 (1.1) & 1.8 (1.1)\\
-\% Sites Bleeding & 67.3 (21.1) & 43.9 (20.4) & 67.6 (21.3) & 43.5 (20.3)\\
-Trt. Completed: No & - & 14 (3.4\%) & - & 4 (1.3\%)\\
-\addlinespace
-Trt. Completed: Und & - & 196 (47.5\%) & - & 153 (48.7\%)\\
-Trt. Completed: Yes & - & 185 (44.8\%) & - & 157 (50\%)\\
-\bottomrule
-\end{tabular}
-\end{table}
-
-\section{Additional Results from the Applications}
-\singlespacing
-The following table gives summary statistics for covariates and post-treatment outcomes in two of the conditions from the empirical study.\\
-
-\subsection{Summary Statistics}
-The following table gives summary statistics for covariates and post-treatment outcomes in two of the conditions from the empirical study.\\
-\small
-\input{tab1fac}
-\clearpage
-
-\small
-\input{tab1num}
-
-\subsection{Regression Results}
-Regression estimates from three principal score logit models and three outcome regressions, based on the "All Covariates" principal score model, for \textsc{geepers} estimates.
-Standard errors shown are nominal regression errors, not sandwich corrected.
-Fixed-effect estimates for school (PS-models) or classroom (outcome models) are omitted.
-
-\small
-%\input{writeUps/outcomeRegAppendix}
-\input{psTab}
-
-\input{regTab}
-
-% \bibliographystylesupp{plainnat}
-% \bibliographysupp{MOM}
-
 \end{document}

### 2025-07-02 writeUps/responses.tex
finished?? editing

@@
 \usepackage{xr}
-\externaldocument{paper}
+\externaldocument{writeUps/paper}
 
@@
 \usepackage{longtable}
-%\usepackage[notablist,tablesfirst]{endfloat}% To `separate' figures and tables from text if required
 
-\usepackage{epstopdf}% To incorporate .eps illustrations using PDFLaTeX, etc.
 \usepackage[caption=false]{subfig}% Support for small, `sub' figures and tables
-\usepackage[doublespacing]{setspace}% To produce a `double spaced' document if required
-%\setlength\parindent{24pt}% To increase paragraph indentation when line spacing is doubled
+\usepackage[doublespacing]{setspace}
 
@@
 \renewcommand\bibfont{\fontsize{10}{12}\selectfont}% To set the list of references in 10 point font using natbib.sty
-
-% \theoremstyle{plain}% Theorem-like structures provided by asthma.sty
-% \newtheorem{theorem}{Theorem}[section]
-% \newtheorem{lemma}[theorem]{Lemma}
-% \newtheorem{corollary}[theorem]{Corollary}
-% \newtheorem{prop}[theorem]{Proposition}
-
-% \theoremstyle{definition}
-% \newtheorem{definition}[theorem]{Definition}
-% \newtheorem{example}[theorem]{Example}
-
-% \theoremstyle{remark}
-% \newtheorem{remark}{Remark}
-% \newtheorem{notation}{Notation}
 
@@
 We are grateful for the reviewers' and the editor's careful reading
-and comments. We believe that the the substantial changes we made
-chages to the manuscript in response to those comments greatly
+and comments. We believe that the substantial changes we made
+to the manuscript in response to those comments greatly
 improved the paper. Most significantly, we clarified our presentation
@@
 \section*{Reviewer 1}
-
-The article effectively presents a simulation study comparing three methods and provides a clear application example of applying GEEPERs.
-However, there are notable areas for improvement, including a lack of literature review on previous simulation studies about Principal Stratification and some unclear presentations of results.
-Specific comments are outlined below.
 
@@
 We have added material to section \ref{sec:dataGeneration} and Table \ref{tab:factor} that includes the necessary details.
-Should anything still be unclear, replication code is available in the anonymized GitHub repository at \url{https://anonymous.4open.science/r/psGee-B498/}.
+Should anything still be unclear, replication code is available at \url{https://osf.io/fu2sv/?view_only=b2baebc2f3054ef796717863e4a9662f}.
 
@@
 \end{itquote}
-We added rationale, including citations, to the paragraph beginning ``Potential and observed outcomes...'' on page \pageref{potential} and to section \ref{sec:manipulatedFactors}.
+This was a helpful suggestion. We added rationale, including citations, to section \ref{sec:manipulatedFactors}.
 
@@
 Thank you. We have fixed the formatting.
-\End{enumerate}
+\end{enumerate}
 
@@
 
-\item \begin{itquote}   Please provide more background on M-estimators especially to the extent it evolves to become relevant to the GEEPERs. 
+\item \begin{itquote}   Please provide more background on M-estimators, especially to the extent it evolves to become relevant to the GEEPERs. 
 \end{itquote}
@@
 \end{itquote}
-This is a great suggestion. We added the paragraph beginning ``The simulation study suggests\dots'' on page \pageref{evalMetric} to give some guidance along those lines. 
+This is a great suggestion. We added some suggestions at the end of the paragraph beginning ``In sum," on page \pageref{simsum}.
+
 \item \begin{itquote}     Quickly checked Frangakis and Rubin (2002).
@@
 \end{itquote}
-Full replication code is available at \url{FILL IN} 
+\sloppy
+Full replication code is available at \url{https://osf.io/fu2sv/?view_only=b2baebc2f3054ef796717863e4a9662f}. 
 \end{enumerate}
@@
 However, the competing methods often boast smaller standard errors than \geepers. 
-This set of results is typical for semi- or non-parametric methods, which tend to be less efficient than their parametric counterparts when the latter's assumptions are met, but are, say, less biased when those assumptions are violated.
+This pattern is typical for semi- or non-parametric methods, which tend to be less efficient than their parametric counterparts when the latter's assumptions are met, but are, say, less biased when those assumptions are violated.
 
@@
 That said, principal ignorability is untestable and, in our opinion, rather strong. 
-As we stated on page \pageref{evalMetric} (in the paragraph beginning ``The simulation study suggests''), researchers will have to decide in each application which assumptions are the most plausible, and choose estimators based on that decision.
+As we stated on page \pageref{simsum} (in the paragraph beginning ``In sum''), researchers will have to decide in each application which assumptions are the most plausible, and choose estimators based on that decision.
 
@@
 
-Following your fine suggestion, we formatted Table \ref{tab:coverage}, highlighting results in red when one or more assumptions have been violated.
+Following your fine suggestion, we formatted Tables \ref{tab:factor} and \ref{tab:coverage}, highlighting results in red when one or more assumptions have been violated.
 This format makes it clearer that, for instance, \pmm makes distributional assumptions while \geepers does not, and that \geepers is less sensitive to the assumptions that both estimators make, especially when $\alpha=0.5$.
@@
 
-\end{document}}
+\end{document}

### 2025-07-04 writeUps/paper.tex
resubmitted version JREE

@@
 \usepackage{xr}
-\externaldocument{writeUps/appendix}
+\externaldocument{appendix}
 \RequirePackage{amsmath,amsfonts,amssymb,amsthm}
@@
 \usepackage{xspace}
+
+\usepackage{placeins}
 
@@
 
-\usepackage{endfloat}% To `separate' figures and tables from text if required
+%\usepackage{endfloat}% To `separate' figures and tables from text if required
 
@@
 %\begin{aug}
-%  \author{
-%  Adam C. Sales\thanks{CONTACT Adam C. Sales. Email: asales@wpi.edu}, Kirk P. Vanacore, and Erin R. Ottmar%\\
-%  %\affil{Worcester Polytechnic Institute, Worcester, MA, USA}
-% }
+  \author{
+  Adam C. Sales\thanks{CONTACT Adam C. Sales. Email: asales@wpi.edu}, Kirk P. Vanacore, and Erin R. Ottmar%\\
+   }
 
 %   \maketitle
+\date{}
 
@@
 In evaluating a recent experimental evaluation against alternative programs without bottom-out hints, researchers may be interested in estimating separate average treatment effects for students who, if given the opportunity, would request bottom-out hints frequently, and for students who would not.
-Most principal stratification estimators rely on strong structural or modeling assumptions, and many require advanced statistical training to fit and check.
-In this paper, we introduce a new M-estimation principal effect estimator for one-way noncompliance based on a binary indicator.
+Most principal stratification estimators rely on strong structural or modeling assumptions, and many require advanced statistical training to fit and check, making them inaccessible to many applied researchers.
+In this paper, we introduce a new principal effect estimator for one-way noncompliance based on a binary indicator.
 Estimates may be computed using conventional regressions (though the standard errors require a specialized sandwich formula) and do not rely on distributional assumptions.
@@
 
+Parametric mixture models are often difficult for applied researchers to implement---it is far removed from the regression models that dominate quantiative social science.
+Fitting them may require familiarity with Bayesian Markov Chain Monte Carlo methods or liklihood optimization routines---even when these are automated \citep[e.g.]{PStrata}, they are slow and require an additional round of diagnostics for model convergence.
+Furthermore, the usual model checking techniques of regression analysis, such as examining residuals, are not straightforward to apply in these contexts.
+Combined with their fragility and heavy dependence on model specification, these factors have severely limited their accessibility to applied researchers.
+
 These concerns have prompted the development of alternative approaches, including bounding \citep{bounding} and randomization inference \citep{nolen2011randomization}.
@@
 Principal ignorability is a direct analogue of the strong ignorability assumption in causal observational studies \citep[e.g.][]{rosenbaum1983central}---that all covariates that predict both principal stratum membership and outcomes have been accounted for---in other words, the precise type of assumption that randomized trials are designed to avoid.
-A separate set of results in \citet{jo2002}, \citet{ding2011}, \citet{jiangDing2021}, and others have shown that, under certain circumstances, researchers can use baseline covariates to identify and estimate principal effects without relying on principal ignorability.
+A separate set of results in \citet{jo2002}, \citet{ding2011},
+\citet{jiangDing2021}, \citet{richardson2023estimating}, and others have shown that, under certain circumstances, researchers can use baseline covariates to identify and estimate principal effects without relying on principal ignorability.
 
@@
 We show that this estimator is the solution to a set of estimating equations based on the combined moments of principal scores, indicators for treatment assignment and observed principal stratum membership in the treatment group, outcomes, and (optionally) covariates, and give a set of assumptions under which it is consistent for the true principal effects.
+
 We will refer to the estimators presented here as ``\textsc{geepers}'': general estimating equations for principal effects using regressions.
@@
 Unlike principal score weighting, \textsc{geepers} does not assume principal ignorability and admits to sandwich-style standard error estimates \citep{stefanskiBoos}, which we provide in the appendix.
-
-After the initial submission of this manuscript, we became aware of \citet{richardson2023estimating}, which suggests a similar, though not identical, estimator to ours, using G-estimation of a structural nested mean model \citep[e.g.][]{wallace2017r}.
-Our development of \textsc{geepers} builds on \citet{richardson2023estimating} by emphasizing connections to both parametric mixture modeling and principal score analysis, and comparing \textsc{geepers} to those methods in a simulation study.
-Moreover, \textsc{geepers}, which, in practice, is a combination of logistic regression and OLS, is more accessible to applied researchers while yielding very similar results.
-
-This work advances the literature in three ways: first, it makes an explicit connection between principal score methods, M-estimation, and semi-parametric estimation of principal effects using auxiliary data.
+Unlike any principal stratification estimator we are aware of, \geepers estimates are based on widely-used regression techniques, making them easy to compute, check, and understand for researchers without an extensive background in statistics.
+Existing methods fail to meet three needs: (1) accessibility, (2) robustness, and (3) model flexibility. \textsc{geepers} addresses all three.
+
+%After the initial submission of this manuscript, we became aware of \citet{richardson2023estimating}, which suggests a similar, though not identical, estimator to ours, using G-estimation of a structural nested mean model \citep[e.g.][]{wallace2017r}.
+%Our development of \textsc{geepers} builds on \citet{richardson2023estimating} by emphasizing connections to both parametric mixture modeling and principal score analysis, and comparing \textsc{geepers} to those methods in a simulation study.
+%Moreover, \textsc{geepers}, which, in practice, is a combination of logistic regression and OLS, is more accessible to applied researchers while yielding very similar results.
+
+%This work advances the literature in three ways: first, it makes an explicit connection between principal score methods, M-estimation, and semi-parametric estimation of principal effects using auxiliary data.
 Second, it provides a general estimator that relies on two of the most familiar tools in the applied education researcher's toolbox: linear and logistic regression.
@@
 
-\begin{table}
+\begin{center}
+  [Table \ref{tab:factor} around here]\\
+\end{center}
+
+\textbf{\emph{Sample Size per condition $\bm{n}$:}} The guarantees of M-estimation are all asymptotic, so we examined the behavior of \textsc{geepers} at a range of sample sizes to determine its finite-sample properties.
+The range of 100--1000 was chosen to represent typical student-level randomized trials in education, for instance, \citet{growthMindsetRuralBurnette} with $n=115$ in the treatment group and $n=107$ in the control group, or \citet{impactPaper}, which initially randomized an average of 1000 students to each condition.
+
+$\bm{\alpha}$\textbf{:} $\alpha$ controlled the extent to which $\st$ is predictable as a function of covariates.
+When $\alpha$ was higher, $\st$ was more easily predicted by covariates; when $\alpha=0$, Assumption \ref{ass:vps} was violated, i.e., $Var(\pp)=0$.
+We expected that the more predictive the covariates, the better the performance of all three estimators.
+To aid in the interpretation of the $\alpha$ parameter, the online appendix plots $\alpha$ against the area under the receiver operating characteristic curve for each fitted model (AUC), a common measure of classification accuracy \citep{bradley1997use}---an AUC of 0.5 implies that the model is no better than random guessing, and AUC of 1 implies perfect prediction.
+For $0.5\le \alpha \le 1$, the average AUC varies from roughly 0.53 to 0.77; for $\alpha=0.5$, the average AUC was roughly 0.675.
+
+For reference, in the application of Section~\ref{sec:opts}, there were two covariates.
+In the model for principal scores, their standardized coefficients were roughly 0.46 and 0.62, and the AUC was 0.68.
+In the application of Section~\ref{sec:fh2t}, the principal score models (which were far more complex than either the simulation model or the model in Section \ref{sec:opts}) had AUCs between 0.745 and 0.795.
+
+\textbf{\emph{The distribution of $\bm{\epsilon_i}$:}} $\epsilon$ is the regression error in \eqref{eq:y-sim}.
+Across simulation runs, $E[\epsilon]=0$ and $Var(\epsilon)=1/2$.
+However, the shape of $\epsilon$'s distribution varied between the following two possibilities:
+Normal $\epsilon\sim\mathcal{N}(0,1/2)$, as assumed by \pmm, or Uniform $\epsilon\sim\mathcal{U}(-\sqrt{6}/2,\sqrt{6}/2)$.
+%\item[Mixture] $f_\epsilon(x)=\frac{1}{4}\phi\left((x+1/3)\sqrt{6}\right)+\frac{1}{4}\phi\left((x-1)\sqrt{6})\right)$
+%The mixture distribution for $\epsilon$ was designed to be particularly difficult for the mixture model estimator.
+
+% $\bm{\beta_1}$\textbf{:} When $\beta_1=0$, there is only one mixture component in the control
+% group, which can be problematic for mixture modeling \citep{griffin2008application,feller2016principal}.
+Conversely, when $\beta_1=0.3$, unless there is an interaction between $Z$ and covariates (i.e., $\gamma_3\ne 0$), $\eff0=\eff1=0$, so there is no average treatment effect for either principal stratum.
+
+\textbf{\emph{Interactions between $\st$ or $Z$ and covariates:}} This factor was designed to test the vulnerability of the estimators to certain violations or Assumption \ref{ass:rci}.
+
+When neither interaction was present, $\gamma_2=\gamma_3=0$, then $\gamma_1=1/\sqrt{6}$.
+In these cases, since  $Var(\gamma_1(x_1+x_2+x_3))=1/2$ and $Var(\epsilon_i)=1/2$, covariates explained half of the variance of $Y$ within treatment group and principal stratum, and observed covariates $x_1$ and $x_2$ explained 1/3 of that variance.
+These values align with estimates from outcome models in Section \ref{sec:fh2t} and are typical in education contexts \cite{hedgesHedberg}.
+
+When there was an interaction between covariates and $\sti$, then $\gamma_1=3/(4\sqrt{6})$ and $\gamma_2=1/(2\sqrt{6})$, so that the slopes for $x_1$ and $x_2$ varied by half their magnitude between principal strata, and on average they were equal to $1/\sqrt{6}$, as in the no-interaction case.
+Interactions between covariates and $Z$ were controlled by $\gamma_3$: when there were interactions, $\gamma_3=1/(2\sqrt{6})$, otherwise $\gamma_3=0$.
+Either type of interaction violates Assumption \ref{ass:rci} since if $\gamma_2\ne 0$ or $\gamma_3 \ne 0$, there is no single vector $\bm{\gamma}$ that can fully capture the dependence of $\yc$ and $\yt$ on $S$---instead, the dependence varies with $\st$ and/or with $Z$.
+
+\subsubsection{Analysis Models}\label{sec:simMods}
+
+Using each simulated dataset, we estimated $\eff0$ and $\eff1$ with three methods: \geepers, \pmm with principal scores \eqref{eq:mixtureModelPS}, and the \psw estimator \eqref{eq:psw}.
+
+The methods all used the same ``observed'' dataset, consisting of outcomes $Y$, treatment assignments $Z$, two covariates $\bx=[x_1,x_2]$, and $S$ observed only when $Z=1$.
+The third covariate used in the data-generating model, $x_3$, was ``unobserved''---since $x_3$ was correlated with both $\st$ and $Y$, Principal Ignorability \eqref{eq:pi} was violated.
+
+All three methods used the same principal score model, a logistic regression of $S$ on $\bx$, and an intercept, fit using data from the treatment group in which $S=\st$ was observed, i.e., $\ppf{\bm{\alpha}}{\bx}=logit^{-1}(\alpha_0+\bm{\alpha}'\bx)$.
+
+\pmm, the parametric mixture estimator, then fits two outcome models:
+For the treatment group, the regression model $Y=\beta_0^T+\beta_1^TS+\bm{\gamma}'\bx+\epsilon$ and for the control group, a mixture model of the form \eqref{eq:mixtureModelPS} with linear normal models for $g^s(y|x_1,x_2)$.
+Residual variance was allowed to differ across treatment groups, but not across principal strata.
+These models were fit simultaneously using Bayesian Markov Chain Monte Carlo with Stan \citep{rstan}.
+We interpreted the posterior mean as a point estimate and the posterior standard deviation as a standard error.
+
+For \textsc{geepers} and \psw, the principal score model was fitted to data from the treatment group using standard maximum likelihood methods.
+For \textsc{geepers}, we used the estimated principal scores in the control group, along with the observed $S$ in the treatment group, to construct $R$, and then fit the model \eqref{eq:reg2} with OLS.
+To estimate standard errors, we computed the sandwich covariance matrix following the procedure detailed in the appendix.
+
+For \psw we plugged estimated principal scores into \eqref{eq:psw} to estimate $\muc0$ and $\muc1$, estimated $\mut0$ and $\mut1$ with the sample means of subjects with $Z=1$ and $S=0$ or $S=1$, respectively, and estimated $\heff0=\hat{\mu}_{Tw}^0-\hat{\mu}_{Cw}^0$ and $\heff1=\hat{\mu}_{Tw}^1-\hat{\mu}_{Cw}^1$.
+
+\subsection{Results}\label{sec:simResults}
+In this section, we evaluate estimates of $\eff1$. Results for estimates of $\eff0$, which are similar, can be found in the online appendix. 
+In replication $b=1,\dots,500$, define the error of a given estimator $\heff1_b$ as $err_b=\heff1_b-\eff1$.
+The bias of $\heff1$, $\EE[\heff1-\eff1]$, is estimated as $\widehat{bias}=\sum_b err_b/500$, and the empirical standard error is defined as $\left\{\sum_b(\heff1_b-\overline{\heff1})^2/499\right\}^{1/2}$, the sample standard deviation of $\heff1$.
+We also report the empirical coverage of confidence intervals, the proportion of replications in which $\eff1$ is within the estimated confidence bounds of $\heff1$.
+Results on root-mean-squared-error %$\left(\widehat{bias}^2+SE_{emp}^2\right)^{1/2}$
+$\left(\sum_b err_b^2/500\right)^{1/2}$
+are in the online appendix.
+
+\subsubsection{Performance with Varying $n$ and $\alpha$}
+% When $\alpha=0$, Assumption \ref{ass:vps} was violated, and \textsc{geepers} estimates were highly erratic; however, as Table \ref{tab:coverage} shows, when other assumptions held, inference remained valid.
+To what extent do the performances of \geepers and \psw depend on $n$ and $\alpha$?
+For this set of results, $\epsilon\sim \mathcal{N}(0,1/2)$ and there were no interactions in \eqref{eq:y-sim}; $\alpha$ was fixed at $0.5$ for varying levels of $n$, and $n$ was fixed at 500 for varying levels of $\alpha$.
+
+Figure \ref{fig:alphan} shows the empirical bias and standard error (SE) of the \textsc{geepers} and \pmm estimators of $\eff1$ as sample size and $\alpha$ varied.
+Across $n$ and $\alpha$, \textsc{geepers} was approximately unbiased,\footnote{All p-values testing for \geepers bias were $>0.1$, with the exception of when $\alpha=1$, with $\widehat{bias}=0.01$, $p=0.046$ before adjustment for multiple comparisons.} while $\pmm$ estimators exhibited slight positive bias that decreased with $n$.
+
+\geepers had a higher SE than \pmm, particularly when $n=100$ or $\alpha=0.2$.
+The \psw advantage is far less dramatic for higher levels of $n$ and $\alpha$, eventually diminishing to approximately zero.
+
+\begin{center}
+  [Figure \ref{fig:alphan} around here]\\
+\end{center}
+
+\subsubsection{Performance when Assumptions do not Hold}
+
+Our next set of results compares the performance of \geepers, \pmm,
+and \psw when one or more assumptions of each method may be violated.
+We present results for $n=500$ here; results from when $n=1,000$,
+which are similar, can be found in the online appendix.
+
+\begin{center}
+  [Figure \ref{fig:boxplots} around here]\\
+\end{center}
+
+%<<<<<<< HEAD
+Figure \ref{fig:boxplots} shows violin plots, layered on top of jittered scatter plots, showing estimation errors for \geepers, \pmm, and \psw.
+Table \ref{tab:coverage} shows empirical coverage rates of nominal 95\% confidence interval estimates for \geepers and \pmm; when at least one of an estimator's assumptions is violated, the result is colored red.
+
+When all of the \geepers and \pmm assumptions were met, that is, when $\epsilon\sim\mathcal{N}(0,1/2)$, $\alpha>0$, and there were no covariate interactions in \eqref{eq:y-sim}, results for $\geepers$ and $\pmm$ were the same as in the previous section; in addition, confidence intervals from both methods achieved their nominal levels.
+\textsc{psw} had by far the lowest variance and the highest bias of the three estimators; the latter is due to the violation of principal ignorability.
+
+%The remainder of the results reflect violations modeling assumptions for either \geepers or the parametric mixture model or both: when $\alpha=0$, violating Assumption \ref{ass:vps}; interactions in the data generating model between covariates $X$ and treatment assignment $Z$ and/or principal stratum $S$, either of which would violate assumption \ref{ass:rci} and the linear additive model of the \textsc{pmm}; and/or residuals drawn from a uniform distribution, which violates the normality assumption of \textsc{pmm}.
+
+%We omitted the $\alpha=0$ case from Figures \ref{fig:alphan} and \ref{fig:boxplots} since the variance of \geepers and \pmm estimates was so large as to dominate the plot. 
+When $\alpha=0$, Assumption \ref{ass:vps} is violated; %, standard errors for \geepers were too high to 
+Nevertheless, as the first column of results in Table \ref{tab:coverage} shows, when other assumptions were met, confidence intervals from both \geepers and \psw tended to over-cover when $\alpha=0$. 
+In a randomized experiment, $\alpha$ can be estimated directly, making Assumption \ref{ass:vps} testable.
+
+The impact of covariate interactions in \eqref{eq:y-sim} on the performance of \geepers depended on $\alpha$: when $\alpha=0.2$, the presence of interactions increased sampling variance.
+Still, only interactions between $X$ and $Z$ reduced confidence interval coverage.
+When $\alpha=0.5$, interactions had little to no effect on the \geepers estimator, and all 95\% confidence intervals approximately achieved their nominal levels.
+\geepers performed similarly under either normal or uniform residuals.
+
+Like \geepers, \pmm is insensitive to interactions between $X$ and $S$.
+However, interactions between $X$ and $Z$ in the data-generating model skewed its sampling distribution and reduced the coverage of its 95\% confidence intervals for all values of $\alpha$ in the simulation.
+Uniformly-distributed residuals caused bimodality in the \pmm sampling distribution, and caused its credible intervals to exhibit severe undercoverage.
+
+\label{simsum}In sum, \psw is the least variable of all three estimators, and does not depend on outcome modeling, but is vulnerable to bias from an unmeasured covariate that predicts both $S$ and $Y$.
+When all assumptions of both methods are met, \geepers is less efficient, but also less biased, than \pmm, but differences in efficiency decrease with $n$ and $\alpha$.
+Neither \psw nor \pmm is vulnerable to interactions between $X$ and $S$.
+\pmm is highly vulnerable to violations of normality and to interactions between $X$ and $Z$, while \psw is robust to non-normality and robust to $X$-$Z$ interactions when $\alpha = 0.5$.
+These results suggest that researchers who are willing to assume principal ignorability should use \psw, and those who are confident in their parametric model for potential outcomes should use a parametric mixture model.
+Researchers who are hesitant to accept principal ignorability or a parametric outcome model should consider \geepers, particularly when $n\ge 200$ and covariates can predict $S$ with AUC$>0.6$ (corresponding to $\alpha>0.2$).
+These suggestions may depend heavily on the simulation setup; applying them in general scenarios requires further research.
+
+%\label{evalMetric}The simulation study suggests that none of the three methods evaluated---\geepers, \pmm, or \psw---dominates the others, and offers some guidance on when each estimator may be preferable.
+\psw is the most precise and requires no assumptions about the distribution of outcomes, but can be quite sensitive to violations of principal ignorability.
+When the determinants of $\st$ are well-understood and measured, \psw seems to be the best option.
+When principal ignorability does not hold, potential outcomes are conditionally normal, and there are no $X:Z$ interactions within principal strata, the \pmm estimator outperforms \geepers, though the difference decreases with $n$ or $\alpha$.
+\geepers may be the best choice when conditional normality may not hold, or if there are $X:Z$ interactions and $\bm{x}$ is sufficiently predictive of $\st$ (AUC $\ge$0.675, corresponding to $\alpha \ge 0.5$, may serve as a threshold).
+
+%=======
+% Figure \ref{fig:boxplots} shows violin plots, layered on top of jittered scatter plots, showing estimation error for \textsc{geepers}, the mixture estimator, and the \textsc{psw} estimator of $\eff1$ under various sets of conditions, and Table \ref{tab:coverage} shows empirical coverage of nominal 95\% confidence interval estimates under a somewhat wider set of circumstances when $\muc1=0.3$.
+More complete results, including root-mean-squared-error estimates, are in the online appendix.
+% The leftmost panel of Figure \ref{fig:boxplots} and the top line of Table \ref{tab:coverage} show results under conditions favorable to both the mixture model and to \textsc{geepers}---normal residuals and no interactions in the data generating model between covariates and either $S$ or $Z$.
+% \textsc{geepers} was roughly unbiased, but less precise than the other two methods---when $\alpha=0.2$ it was much noisier, and when $\alpha=0.5$ it was only slightly less precise.
+% The mixture estimator's bias depended on $\muc1$---when $\muc1=0$, the bias was slight, and when $\muc1=0.3$, the bias was more substantial.
+This result is somewhat surprising, because when $\muc1=0$, there was no separation between the principal strata in the control group, and previous literature \citep{griffin2008application} has identified this scenario as a particular challenge for mixture modeling.
+% The \textsc{psw} estimator was also biased, due, presumably, to the violation of Principal Ignorability, Assumption \ref{ass:PI}, inherent in the data generating model.
+On the other hand, the \textsc{psw} estimator had easily the lowest sampling variance of the three estimators.
+
+% The middle panel of Figure \ref{fig:boxplots} and the fifth line of Table \ref{tab:coverage} show results for when the residuals in \eqref{eq:y-sim} are uniformly distributed.
+% In this case, \textsc{geepers} and \textsc{psw} estimators, which do not assume normality, behaved in a similar fashion as in the normal case, but the mixture estimator showed a bimodal pattern--sometimes over-estimating $\eff1$ and sometimes underestimating, but rarely estimating $\eff1$ accurately.
+% Credible intervals from the mixture model exhibited severe undercoverage, while \textsc{geepers} confidence intervals achieved at least their nominal level.
+
+% %\sloppy
+% The rightmost panel of Figure \ref{fig:boxplots} returns to the case of normal residuals---and only shows results when $\muc1=0.3$ and $\alpha=0.5$---but introduces interactions in the data generating model between covariates and either $\st$, $Z$, or both, violating Assumption \ref{ass:rci}.
+% %Presence of interactions could undermine the \textsc{geepers} by inducing a relationship between principal scores and potential outcomes, even after adjusting the outcomes for covariate effects.
+% Fortunately, these interactions made little difference when $\alpha=0.5$.
+However, Table \ref{tab:coverage} shows that when $\alpha=0$ or 0.2, and interactions between $\bx$ and $Z$ were present (i.e., $\gamma_3\ne 0$), \textsc{geepers} 95\% confidence intervals tended to under-cover.
+Nevertheless, coverage of \textsc{geepers} intervals was nearly always substantially higher than coverage of corresponding 95\% credible intervals from mixture models.%\footnote{}
+% >>>>>>> overleaf-2025-06-12-1948
+
+\begin{center}
+  [Table \ref{tab:coverage} around here]\\
+\end{center}
+
+\section{Applications}
+To illustrate \textsc{geepers} and compare it to the other discussed methods, we include two applied examples.
+First, we briefly replicate an analysis from \citet{richardson2023estimating} using the \textsc{geepers} and the other methods discussed above.
+This is a relatively straightforward example of an experiment with imperfect compliance in which \citet{richardson2023estimating} estimated statistically significant effects overall in both principal strata.
+Next, we present a novel analysis of data from the RCT of computer-based math applications reported in \citet{impactPaper}, and look into the role of bottom-out hints, as described in the introduction.
+This analysis aims to shed light on an important question in education research, illustrate model and covariate selection, and include a discussion on the plausibility of identification assumptions.
+
+\subsection{Accounting for Imperfect Compliance: Re-Analysis of the Obstetrics and Periodontal Therapy Study}\label{sec:opts}
+\sloppy
+Periodontal disease, a weakening of bone and connective tissue supporting teeth, is highly prevalent among American adults and is associated with adverse pregnancy outcomes, including preterm birth \citep{periodontalEpi}.
+The Obstetrics and Periodontal Therapy Study \citep[OPTS; ][]{michalowicz2006treatment} investigated the efficacy of nonsurgical periodontal treatment to reduce the preterm birth rate; replication data are available at \url{https://causeweb.org/tshs/obstetrics-and-periodontal-therapy/}.
+Pregnant patients at four different hospitals who were at least 16 years old, at less than 16 weeks and 6 days of gestation, and diagnosed with periodontal disease were individually randomized to either receive periodontal treatment during the first 21 weeks of pregnancy ($Z=1$, $n_T=413$) or after delivery ($Z=0$, $n_C=410$).
+The main component of the intervention consisted of up to four visits, followed by monthly tooth polishing and other dental care as needed.
+Not all participants randomized to the treatment condition completed the entire treatment plan; however, 96\% of them received at least some periodontal care during pregnancy.
+Patients randomized to the control condition did not have access to the intervention.
+
+Following \citet{richardson2023estimating}, let $\st=1$ for patients who, if randomized to treatment, would complete the full treatment plan, and let $\st=0$ for those who would not.
+Let $Y$, the outcome of interest, be the percentage of probed sites (6 sites per tooth) that were bleeding at visit 5, which was at 29--32 weeks of gestation, with potential values $\yc$ and $\yt$, as above.
+\citet{richardson2023estimating} used a principal stratification estimator they called ``bespoke instrumental variables'' (BSIV) to estimate principal effects $\eff0$ and $\eff1$, the average effects of randomization to the treatment condition for patients who would or would not complete the full treatment plan.
+In this section, we compare their result to estimates using \geepers, \pmm, and \psw.
+
+For simplicity, we followed the same analysis decisions as \citet{richardson2023estimating}. 
+Specifically, we conditioned on two covariates, Serum measures of endotoxin level and fibrinogen at baseline, and
+omitted the 164 cases in which $Y$ was missing as well as 19 additional cases in which one or both of the covariates were missing.
+That left a total sample size of 640, with 314 randomized to the treatment condition and 326 randomized to control.
+%The compliance variable in the dataset had three categories: ``Yes,'' for subjects in the treatment group who completed the treatment plan (157 subjects), ``No'' for those who did not (4 subjects), and ``Und'' for those who received some therapy, but it is unknown whether they completed the treatment (157 subjects).
+%We let $S=1$ when compliance was ``yes''---i.e., when subjects in the treatment group completed the treatment---and $S=0$ otherwise.
+
+For all three estimation methods, we estimated principal scores with logistic regression of $\st$ on the two covariates.
+For \textsc{geepers} and \textsc{psw}, we used standard maximum likelihood methods and data from the treatment group, achieving an AUC of 0.68.
+For the mixture model, we incorporated the logistic regression specification as a sub-model within a larger Bayesian model fit in Stan.
+
+\begin{center}
+  [Figure \ref{fig:opt} around here]\\
+\end{center}
+
+The results are displayed in Figure \ref{fig:opt}% and Table \ref{tab:optEffs}
+: both \textsc{geepers} and the Bayesian mixture model gave similar results to those reported in \citet{richardson2023estimating} from BSIV.
+For the principal stratum of subjects who would not fully complete the treatment, those three methods estimated that assignment to the treatment condition reduced percentage of dental sites with bleeding by 31--34 percentage points; for subjects who would complete the treatment, the three estimated a reduction of 14-16 percentage points.
+The conclusion of these methods is somewhat surprising: that the effect was larger for subjects who would not complete the treatment than for those who would (the two strata had similar bleeding percentages at baseline, and a logit transformation of the outcome leaves the results qualitatively unchanged).
+One possible explanation is the conjecture that subjects stopped attending treatment once their periodontal disease improved.
+
+In contrast, principal score weighting resulted in very similar effects for both strata---a reduction of 23--25 percentage points, which is close to the estimated intent-to-treat effect of 25 percentage points.
+
+\subsection{Estimating Treatment Effects for Implementation Subgroups: Students who Frequently or Rarely Request Bottom-Out Hints}\label{sec:fh2t}
+
+In the educational field experiment reported in \citet{impactPaper}, middle school students completing their math schoolwork on a computer program were randomly assigned to four conditions.
+These included two gamified programs---DragonBox and From Here to There (FH2T)---and two programs in which students used computers to work through a series of algebra problems taken from the district's textbook---``ASSISTments'' and an active control called ``Business as Usual'' (BAU).
+% the active control condition in which students simply used the computer to work on a sequence of math problems.
+In the DragonBox game, students learned to solve algebraic equations by isolating a box containing a dragon; as students progressed in the game, pictures of monsters were replaced by mathematical symbols.
+Another condition, the From Here to There (FH2T) program, used visual cues such as color and spacing to teach algebraic concepts and allowed students to solve equations by manipulating expressions via dragging and dropping.
+%In ASSISTments, students used computers to work through a series of math problems taken from the district's textbook.
+Students in the ASSISTments condition had access to a series of hints for each problem, culminating in a ``bottom out'' hint that contained the answer.
+They received immediate feedback on their answers and had to enter the correct answer to proceed to the next problem.
+In the BAU condition, students had no access to hints and received feedback only after completing the assignment.
+
+In all four conditions, students worked on their assigned math program in class on 12 different occasions spaced throughout the school year: a prior assessment, a mid-test, a post-test, and nine learning sessions.
+In this illustration, we will compare ASSISTments---the only condition featuring bottom-out hints---with each of the three others separately.
+Our goal is to estimate different treatment effects for groups of students who, if assigned to ASSISTments, would request few or many bottom-out hints.
+
+The option to bottom out was only one of several differences between the treatment conditions.
+One way to disentangle the role of bottoming out in any treatment effect would be to estimate the average treatment effects of being assigned to ASSISTments, relative to the other three, separately for subjects who would (if assigned to ASSISTments) bottom out frequently, and for subjects who would not.
+If the availability of bottom-out hints plays a positive role in ASSISTments' efficacy, we might expect larger effects of the condition on students who frequently bottom out than on those who do not; if bottom-out hints are harmful, we might expect the opposite.
+If hints (that do not supply the correct answer) and error messages are helpful, but the ability to bottom out is harmful to learning, then we might expect to estimate a positive effect of being assigned to the feedback condition for students who do not bottom out, but a negative effect for students who do.
+If hints are helpful for learning only if they can be read as worked examples, then we might expect no effect of condition on students who generally do not bottom out, but a positive effect for students who do.
+(These hypothetical results may be masked by other treatment effect heterogeneity between students who would or wouldn't bottom out, but the results could still be informative.)
+
+\subsubsection{Principal Stratification and Alternatives}\label{sec:psInEdTech}
+The goal of principal stratification in this context is to determine if assignment to the ASSISTments condition affected students who tended to request frequent bottom-out hints differently than students who tended to request fewer bottom-out hints.
+Is this the right question to ask?
+
+In a similar context, \citet{sales2021student} suggests two alternative approaches: the average effect of requesting bottom-out hints on posttest scores for students in the ASSISTments condition, and estimating natural direct and indirect effects of assignment to the ASSISTments condition, treating bottom-out hint requests as a mediator.
+The former approach is an observational study: although students were randomized to the ASSISTments condition, bottom-out hint requests were not randomized.
+Therefore, estimating the effect of bottom-out hints requires careful control of confounding.
+Unfortunately, confounding is a major concern in this case, since, presumably, students only requested bottom-out hints if they were struggling with the material, and students who struggled with practice problems are likely to score lower on the posttest, on average, than those who needed less help during practice.
+The best available measure of students' baseline ability, their score on the 10-item pretest, is likely insufficient.
+That said, a plausible estimate of the average effect of bottom-out hint requests would be a valuable contribution to the literature, and a suitably innovative observational design \citep[e.g.][]{beck} may bear fruit.
+
+Another arguable disadvantage of the observational study approach is that it overlooks the broader context of the experiment, focusing solely on whether bottom-out hints affect posttest scores rather than examining their role in the overall effectiveness of ASSISTments.
+A second approach, mediation analysis, contextualizes the bottom-out effect within the larger experiment.
+It, too, relies on the same non-confounding assumption as an observational study, and also suffers from an additional challenge: mediation analysis relies on a ``common support'' condition that ``the mediator state must not be a deterministic function of the treatment'' \citep[p.
+225]{celli2022causal}.
+However, bottom-out hints are unavailable in the other three arms of the study, violating this condition (\citealt{sales2021student} offer a solution, which we find unconvincing).
+
+A principal stratification approach treats bottom-out hint requests as a type of compliance.
+In other words, if we consider full implementation of the ASSISTments condition to include frequent bottom-out hint requests, students assigned to ASSISTments but who requested few bottom-out hints only partially complied with their treatment assignment.
+The question then is whether students who do not fully comply with their treatment assignment still experience an effect.
+Conversely, might students who only partially comply experience a larger positive effect?
+That would be the case if either bottom-out hints were harmful or if the treatment was better targeted toward the type of student who tends to need them less.
+
+Principal score weighting depends on the assumption of principal ignorability, which is equivalent to the no-confounding assumption in observational studies and mediation analysis.
+The mixture model and \textsc{geepers} estimators do not rely on this assumption; instead, they rely on parametric models and residualized covariate independence, respectively.
+A final option is the instrumental variables estimator of \citet{air}, which relies on the ``exclusion restriction'' that randomization did not affect students who would request few bottom-out hints.
+Since bottom-out hints are only one of several features of ASSISTments, this assumption is untenable.
+
+\subsubsection{Data on Bottom-Out Hints and Covariates}
+
+\sloppy
+We gathered dichotomized data on %which students in the treatment condition requested at least one bottom-out hint.
+whether students in ASSISTments requested more than the median number of bottom-out hints (i.e., 11).
+That is, $\st=1$ for students who, if assigned to ASSISTments, would request more than 11 bottom-out hints (``bottom-outers''), and $\st=0$ for students who would request 11 or fewer (``non-bottom-outers'').
+%least one bottom-out hint, and $\st=0$ for students who would not.
+The outcome of interest $Y$ in our example is students' total scores on a ten-item posttest completed within the online tutoring system; $Y$ is an integer ranging from 0 to 10.
+The goal of our analysis will be to determine the principal effects $\eff1=\EE[\yt-\yc|\st=0]$ and $\eff0=\EE[\yt-\yc|\st=0]$, the average effects of assignment to ASSISTments (``$T$'' subscripts) versus the other conditions (``$C$'') for bottom-outers and non-bottom-outers.
+
+We also had access to data on several baseline covariates, including prior achievement, demographics, baseline measurements of student attitudes toward math, and an indicator for whether students began the school year in remote or in-person instruction.
+Students with missing pretest scores were dropped from the analysis; missing data in other covariates were imputed with a Random Forest algorithm \citep{missForest}.
+Summary statistics are displayed in appendix tables \ref{table:tab1fac} and \ref{table:tab1num}.
+
+%The data analysis we present here is intended as a demonstration of \textsc{geepers}, rather than for its substantive conclusions, and our description omits discussion of some important methodological considerations, including attrition and post-selection inference.
+For details on the experimental design, the conditions being compared, sample attrition---which was unusually high due to the COVID-19 pandemic---and the impact analysis, please see \citet{impactPaper}.
+
+For a description of the dataset and an explanation of how to access it, see \citet{ottmar2023data} and \url{https://osf.io/r3nf2/}.
+Replication code for the analysis is available at [Redacted].
+%\url{https://github.com/adamSales/psGee}.
+
+\subsubsection{Data Analysis}
+We estimated principal scores using logistic regression of observed $S$ on school fixed effects and a set of baseline variables $\bxs$.
+Three criteria governed our choice of covariates and model specifications for estimating principal scores and effects.
+First, Proposition \ref{prop:reg2} requires consistent estimation of principal scores and (via Assumption \ref{ass:rci}) correct specification of the model for $\EE[Y|\bx,Z,\st]$.
+However, unlike \psw, \geepers does not require that models include all important covariates (for any definition of "important"), so covariates whose inclusion may threaten model specifications or identification assumptions can be dropped from the analysis.
+
+On the other hand, Assumption \ref{ass:vps} requires principal scores to vary with at least some element $\bxs$.
+Moreover, the results of the simulation study suggest that when $\st$ can be predicted more precisely as a function of $\bxs$, \textsc{geepers} principal effect estimators are more precise and more robust to model misspecification.
+Those considerations militate in favor of choosing covariates $\bxs$ to optimize out-of-sample principal score prediction accuracy.
+
+The first two criteria motivated our choice of two strategies for choosing $\bxs$ and model specifications. 
+In our main model, we included every available covariate in the dataset and chose transformations of those covariates based on the results of residual and binned residual plots \citep{arm}.
+%
+% Including all possible covariates maximizes the model's opportunities to detect covariate relationships that would predict $Y$ or $\st$.
+However, this strategy can also lead to overfitting, which would decrease out-of-sample predictive accuracy.
+In an alternative analysis, we chose a subset of covariates to include in $\bxs$ %and $\bxy$ 
+with a backward stepwise selection algorithm minimizing the AIC \citep{aic} of the principal score model and then modified the resulting models based on residual plots.
+
+Lastly, Assumption \ref{ass:ci} implies that there is no interaction between $\bxs$ and $\st$ in the outcome model. 
+\emph{A priori}, the student attribute that poses the greatest threat to these modeling choices is prior achievement, that is, scores on the pretest and the fifth-grade state test.
+This is both because they measure a similar construct to the posttest and because \cite{impactPaper} found a pretest-treatment interaction in a different treatment contrast.
+Following that reasoning, we conducted a third parallel analysis excluding prior achievement covariates. % in $\bxs$ but including all covariates in $\bxy$.
+
+All three principal score models were fit to data from students in the ASSISTments group, for whom bottom-out status ($S$) was observed; hence, the same model fit was used for all three treatment contrasts.
+
+Outcome models for all analyses included all variables from the corresponding principal score models, but substituted class fixed effects for school fixed effects.\footnote{Why use a coarser set of fixed effects for the principal score model than the outcome regression?
+First, the principal score model was fit to a smaller sample (only students in the ``Instant'' condition), so some classrooms only included one student, which can lead to poor fit.
+Second, the asymptotics of logistic regression are more restricted than OLS, so logistic regression tends to perform poorly when the number of parameters increases with the sample size \citep{agresti}.}
+As with principal score models, we checked model fit with residual plots and modified model specifications accordingly.
+
+% \begin{figure}
+%   \centering
+%   \includegraphics{../figure/psModCoef.jpg}
+%   \caption{Principal Score Model Coefficients}
+%   \label{fig:psMod}
+% \end{figure}
+
+\subsubsection{Results}
+
+The three principal score models were fairly successful in distinguishing bottom-outers from non-bottom-outers---the AUC, evaluated using out-of-sample predictions in a 10-fold cross-validation, was roughly 0.795 for the AIC-optimal model, and about 0.745 for the two others.
+That is, in roughly 75--80\% of bottom-outer/non-bottom-outer pairs of subjects, the bottom-outer will have a higher predicted probability from the model.
+Coefficient estimates are displayed in Table \ref{tab:psTab} in the appendix.
+
+Figure \ref{fig:effects} shows principal effect point estimates and 95\% confidence intervals from analyses based on the three principal score models.
+The three covariate selection strategies led to similar point and interval estimates.
+That said, in nearly all cases, the AIC-optimal model led to the smallest standard errors.
+Confidence intervals for all nine principal effect estimates were about twice as wide as for corresponding ATE estimates.
+In no case was there evidence at $\alpha=0.05$---or any other conventional level---for non-zero principal effects or differences between principal effects.
+Table \ref{tab:regTab}, in the appendix, displays the full set of regression coefficients for the outcome models based on the "All Covariates" principal score model, along with nominal standard error estimates.
+
+A comparison of point estimates suggests that assignment to ASSISTments rather than BAU benefits bottom-outers less than non-bottom-outers.
+We speculate that, if true, this pattern may be because many students use bottom-out hints to game the system---entering the provided correct answers rather than attempting to figure out the problems on their own.
+Removing access to hints for bottom-outers forces them to attempt the problems rather than game the system.
+At the same time, hints and instant feedback are helpful for non-bottom-outers, who are less likely to game the system.
+
+On the other hand, point estimates suggest that assignment to ASSISTments, relative to FH2T, hurts non-bottom-outers but not non-bottom-outers.
+Perhaps non-bottom-outers tend to have higher prior knowledge, and, as \cite{impactPaper} shows, FH2T benefits such students more than students with lower prior knowledge.
+
+Perhaps the simplest explanation for observed differences in point estimates is sampling error, which cannot be ruled out.
+
+\begin{center}
+  [Figure \ref{fig:effects} around here]\\
+\end{center}
+
+\subsubsection{Comparing Principal Stratification Methods}
+
+\begin{center}
+  [Figure \ref{fig:compare} around here]\\
+\end{center}
+
+Figure \ref{fig:compare} shows principal effect estimates from \textsc{geepers} (using the All Covariates principal score model) compared to analogous estimates using Bayesian mixture modeling and \textsc{psw}, alongside estimates of the overall average treatment effects, labeled ``ATE.''
+
+The \textsc{psw} model used the same principal score model as \textsc{geepers}, while the Bayesian model used the same principal score and outcome model specifications.
+
+The Bayesian mixture model assumed that outcomes were normally distributed, conditional on covariates and principal strata, with a residual standard deviation that varied between principal strata.
+This modeling assumption is necessarily false since post-test scores were integers between 0 and 10.
+All model parameters with support in $\mathbb{R}$ were given standard normal priors, while standard deviation parameters were given half-standard normal priors.
+The principal score and outcome models were fit simultaneously using a Markov Chain Monte Carlo algorithm \citep{rstan}.
+
+The \textsc{psw} estimators did not include any covariate outcome modeling. Standard errors for the \textsc{psw} estimates were estimated using the bootstrap, with resampling done within schools.
+
+For all three sets of estimators, approximate 95\% confidence intervals were estimated by adding and subtracting twice the standard error from the point estimate (or twice the posterior standard deviation from the posterior mean, in the Bayesian mixture model).
+
+The other estimation strategies largely agreed with \textsc{geepers} and with each other in the contrast between ASSISTments and BAU, and, to a lesser extent, in the contrast between ASSISTments and DragonBox.
+The biggest difference between the methods was in the contrast with FH2T, in which the Bayesian mixture model resulted in a larger difference between the two principal effects, as well as the largest standard errors, while the two \textsc{psw} principal effect estimates were quite similar.
+However, these differences are difficult to interpret due to the large uncertainty surrounding all the point estimates.
+
+\section{Discussion}\label{sec:discussion}
+\textsc{geepers} is a straightforward approach to principal effect estimation under strong monotonicity, which is more robust---though sometimes less precise---than alternative approaches under a wide array of scenarios.
+It is built on widely-used regression models, making it accessible to applied researchers.
+
+There is good reason to hope that extensions to \textsc{geepers}, including cases in which $S$ takes more than two values, may be straightforward.
+For instance, in truncation by death problems \citep[e.g.][]{zhangRubin,ding2011} with weak monotonicity, $S_i=1$ if participant $i$ survives (or, more generally, if the outcome $Y_i$ is measured) and 0 otherwise, interest is typically in the principal effect for the ``always survive'' principal stratum in which $\sti=S_{ci}=1$.
+If, say, $\sti\ge S_{Ci}$, then every subject in the control condition with $S_i=1$ is in the always survive stratum, while those subjects in the treatment condition with $S_i=1$ are a mixture of the always survive and $\sti=1;\;S_{Ci}=0$ stratum.
+This scenario is broadly similar to the strong monotonicity, one-way noncompliance situation discussed in this paper.
+On the other hand, when no monotonicity assumption holds, both $\st$ and $S_C$ will have to be imputed for every subject; further research is necessary to determine the appropriate way to do so.
+
+Another direction of extension involved the principal score model \eqref{eq:pscore}.
+The performance of \textsc{geepers} in the simulation study of Section \ref{sec:simulation} depended heavily on the factor $\alpha$, which controlled the extent to which covariates could predict $S$.
+That suggests that when covariates are high-dimensional, \textsc{geepers}' performance in applications could be optimized with a high-dimensional semi- or non-parametric model.
+If so, several further questions emerge.
+First, can the parameter $\alpha$ be extended to a more general parameter measuring the prediction accuracy of a non-parametric model?
+Second, if the principal score model cannot be formulated as the solution to a set of estimating equations, how should the standard error be computed?
+Lastly, can over-fit principal score models cause bias or other estimation problems, and if so, are there ways to protect against overfitting?
+
+Sampling error prevents firm conclusions regarding principal effects for students who would, if given the opportunity, request frequent bottom-out hints.
+However, the results are consistent with the hypothesis that ASSISTments is less effective for frequent bottom-out requesters when compared to a similar program without hints or immediate feedback.
+We hypothesize that many students who frequently request bottom-out hints are gaming the system, facilitated by the availability of these hints.
+
+\textsc{geepers} is a flexible, easily-implementable method for principal effect estimation for one-way noncompliance, with predictive covariates; extensions to a broader set of circumstances could be a boon to causal modeling.
+
+\section*{Replication Materials}
+Replication materials can be found at
+\url{https://osf.io/fu2sv/?view_only=b2baebc2f3054ef796717863e4a9662f}.
+
+\section*{AI Disclosure}
+We used Grammarly Pro (Version 14.1243.0) and ChatGPT (GPT-4-turbo, current as of July 2025) for editing and language improvement.
+
+\bibliographystyle{apalike}
+\bibliography{MOM}
+
+\clearpage
+
+\FloatBarrier
+\section*{Tables and Figures}
+\FloatBarrier
+
+\begin{table}[t]
     \caption{\label{tab:factor} Manipulated factors in simulation study}
@@
 
-\textbf{\emph{Sample Size per condition $\bm{n}$:}} The guarantees of M-estimation are all asymptotic, so we examined the behavior of \textsc{geepers} at a range of sample sizes to determine its finite-sample properties.
-The range of 100--1000 was chosen to represent typical student-level randomized trials in education, for instance, \citet{growthMindsetRuralBurnette} with $n=115$ in the treatment group and $n=107$ in the control group, or \citet{impactPaper}, which initially randomized an average of 1000 students to each condition.
-
-$\bm{\alpha}$\textbf{:} $\alpha$ controlled the extent to which $\st$ is predictable as a function of covariates.
-When $\alpha$ was higher, $\st$ was more easily predicted by covariates; when $\alpha=0$, Assumption \ref{ass:vps} was violated, i.e., $Var(\pp)=0$.
-We expected that the more predictive the covariates, the better the performance of all three estimators.
-To aid in the interpretation of the $\alpha$ parameter, the online appendix plots $\alpha$ against the area under the receiver operating characteristic curve for each fitted model (AUC), a common measure of classification accuracy \citep{bradley1997use}---an AUC of 0.5 implies that the model is no better than random guessing, and AUC of 1 implies perfect prediction.
-For $0.5\le \alpha \le 1$, the average AUC varies from roughly 0.53 to 0.77; for $\alpha=0.5$, the average AUC was roughly 0.675.
-
-For reference, in the application of Section~\ref{sec:opts}, there were two covariates.
-In the model for principal scores, their standardized coefficients were roughly 0.46 and 0.62, and the AUC was 0.68.
-In the application of Section~\ref{sec:fh2t}, the principal score models (which were far more complex than either the simulation model or the model in Section \ref{sec:opts}) had AUCs between 0.745 and 0.795.
-
-\textbf{\emph{The distribution of $\bm{\epsilon_i}$:}} $\epsilon$ is the regression error in \eqref{eq:y-sim}.
-Across simulation runs, $E[\epsilon]=0$ and $Var(\epsilon)=1/2$.
-However, the shape of $\epsilon$'s distribution varied between the following two possibilities:
-Normal $\epsilon\sim\mathcal{N}(0,1/2)$, as assumed by \pmm, or Uniform $\epsilon\sim\mathcal{U}(-\sqrt{6}/2,\sqrt{6}/2)$.
-%\item[Mixture] $f_\epsilon(x)=\frac{1}{4}\phi\left((x+1/3)\sqrt{6}\right)+\frac{1}{4}\phi\left((x-1)\sqrt{6})\right)$
-%The mixture distribution for $\epsilon$ was designed to be particularly difficult for the mixture model estimator.
-
-% $\bm{\beta_1}$\textbf{:} When $\beta_1=0$, there is only one mixture component in the control
-% group, which can be problematic for mixture modeling \citep{griffin2008application,feller2016principal}.
-Conversely, when $\beta_1=0.3$, unless there is an interaction between $Z$ and covariates (i.e., $\gamma_3\ne 0$), $\eff0=\eff1=0$, so there is no average treatment effect for either principal stratum.
-
-\textbf{\emph{Interactions between $\st$ or $Z$ and covariates:}} This factor was designed to test the vulnerability of the estimators to certain violations or Assumption \ref{ass:rci}.
-
-When neither interaction was present, $\gamma_2=\gamma_3=0$, then $\gamma_1=1/\sqrt{6}$.
-In these cases, since  $Var(\gamma_1(x_1+x_2+x_3))=1/2$ and $Var(\epsilon_i)=1/2$, covariates explained half of the variance of $Y$ within treatment group and principal stratum, and observed covariates $x_1$ and $x_2$ explained 1/3 of that variance.
-These values align with estimates from outcome models in Section \ref{sec:fh2t} and are typical in education contexts \cite{hedgesHedberg}.
-
-When there was an interaction between covariates and $\sti$, then $\gamma_1=3/(4\sqrt{6})$ and $\gamma_2=1/(2\sqrt{6})$, so that the slopes for $x_1$ and $x_2$ varied by half their magnitude between principal strata, and on average they were equal to $1/\sqrt{6}$, as in the no-interaction case.
-Interactions between covariates and $Z$ were controlled by $\gamma_3$: when there were interactions, $\gamma_3=1/(2\sqrt{6})$, otherwise $\gamma_3=0$.
-Either type of interaction violates Assumption \ref{ass:rci} since if $\gamma_2\ne 0$ or $\gamma_3 \ne 0$, there is no single vector $\bm{\gamma}$ that can fully capture the dependence of $\yc$ and $\yt$ on $S$---instead, the dependence varies with $\st$ and/or with $Z$.
-
-\subsubsection{Analysis Models}\label{sec:simMods}
-
-Using each simulated dataset, we estimated $\eff0$ and $\eff1$ with three methods: \geepers, \pmm with principal scores \eqref{eq:mixtureModelPS}, and the \psw estimator \eqref{eq:psw}.
-
-The methods all used the same ``observed'' dataset, consisting of outcomes $Y$, treatment assignments $Z$, two covariates $\bx=[x_1,x_2]$, and $S$ observed only when $Z=1$.
-The third covariate used in the data-generating model, $x_3$, was ``unobserved''---since $x_3$ was correlated with both $\st$ and $Y$, Principal Ignorability \eqref{eq:pi} was violated.
-
-All three methods used the same principal score model, a logistic regression of $S$ on $\bx$, and an intercept, fit using data from the treatment group in which $S=\st$ was observed, i.e., $\ppf{\bm{\alpha}}{\bx}=logit^{-1}(\alpha_0+\bm{\alpha}'\bx)$.
-
-\pmm, the parametric mixture estimator, then fits two outcome models:
-For the treatment group, the regression model $Y=\beta_0^T+\beta_1^TS+\bm{\gamma}'\bx+\epsilon$ and for the control group, a mixture model of the form \eqref{eq:mixtureModelPS} with linear normal models for $g^s(y|x_1,x_2)$.
-Residual variance was allowed to differ across treatment groups, but not across principal strata.
-These models were fit simultaneously using Bayesian Markov Chain Monte Carlo with Stan \citep{rstan}.
-We interpreted the posterior mean as a point estimate and the posterior standard deviation as a standard error.
-
-For \textsc{geepers} and \psw, the principal score model was fitted to data from the treatment group using standard maximum likelihood methods.
-For \textsc{geepers}, we used the estimated principal scores in the control group, along with the observed $S$ in the treatment group, to construct $R$, and then fit the model \eqref{eq:reg2} with OLS.
-To estimate standard errors, we computed the sandwich covariance matrix following the procedure detailed in the appendix.
-
-For \psw we plugged estimated principal scores into \eqref{eq:psw} to estimate $\muc0$ and $\muc1$, estimated $\mut0$ and $\mut1$ with the sample means of subjects with $Z=1$ and $S=0$ or $S=1$, respectively, and estimated $\heff0=\hat{\mu}_{Tw}^0-\hat{\mu}_{Cw}^0$ and $\heff1=\hat{\mu}_{Tw}^1-\hat{\mu}_{Cw}^1$.
-
-\subsection{Results}\label{sec:simResults}
-In this section, we evaluate estimates of $\eff1$. Results for estimates of $\eff0$ and $\eff1-\eff0$, which are similar, can be found in the online appendix. 
-In replication $b=1,\dots,500$, define the error of a given estimator $\heff1_b$ as $err_b=\heff1_b-\eff1$.
-The bias of $\heff1$, $\EE[\heff1-\eff1]$, is estimated as $\widehat{bias}=\sum_b err_b/500$, and the empirical standard error is defined as $\left\{\sum_b(\heff1_b-\overline{\heff1})^2/499\right\}^{1/2}$, the sample standard deviation of $\heff1$.
-We also report the empirical coverage of confidence intervals, the proportion of replications in which $\eff1$ is within the estimated confidence bounds of $\heff1$.
-Results on root-mean-squared-error %$\left(\widehat{bias}^2+SE_{emp}^2\right)^{1/2}$
-$\left(\sum_b err_b^2/500\right)^{1/2}$
-are in the online appendix.
-
-\subsubsection{Performance with Varying $n$ and $\alpha$}
-% When $\alpha=0$, Assumption \ref{ass:vps} was violated, and \textsc{geepers} estimates were highly erratic; however, as Table \ref{tab:coverage} shows, when other assumptions held, inference remained valid.
-To what extent do the performances of \geepers and \psw depend on $n$ and $\alpha$?
-For this set of results, $\epsilon\sim \mathcal{N}(0,1/2)$ and there were no interactions in \eqref{eq:y-sim}; $\alpha$ was fixed at $0.5$ for varying levels of $n$, and $n$ was fixed at 500 for varying levels of $\alpha$.
-
-Figure \ref{fig:alphan} shows the empirical bias and standard error (SE) of the \textsc{geepers} and \pmm estimators of $\eff1$ as sample size and $\alpha$ varied.
-Across $n$ and $\alpha$, \textsc{geepers} was approximately unbiased,\footnote{All p-values testing for \geepers bias were $>0.1$, with the exception of when $\alpha=1$, with $\widehat{bias}=0.01$, $p=0.046$ before adjustment for multiple comparisons.} while $\pmm$ estimators exhibited slight positive bias that decreased with $n$.
-
-\geepers had a higher SE than \pmm, particularly when $n=100$ or $\alpha=0.2$.
-The \psw advantage is far less dramatic for higher levels of $n$ and $\alpha$, eventually diminishing to approximately zero.
-
-\begin{figure}[!ht]
-  \centering
-  \includegraphics[width=4.5in]{../simFigs/biasSEbyB1N.jpg}
-  \caption{Empirical bias and standard error of \textsc{geepers} and Mixture estimates of $\eff1$ as sample size per treatment group $n$ or $\alpha$ varied.
-There were 500 replications for each $n$ or $\alpha$.
-Other factors were held fixed at the noted values.}
-  \label{fig:alphan}
-\end{figure}
-
-\begin{figure}
-  \centering
-  \includegraphics{../simFigs/biasSEbyB1.jpg}
-  \caption{Empirical bias and standard error of M- and Mixture estimates of $\muc1$ as $\alpha$ varies from 0.1 to 1.
-There were 500 replications for each $n$.
-Other factors were held fixed at the noted values.}
-  \label{fig:alpha}
-\end{figure}
-
-\subsubsection{Performance when Assumptions do not Hold}
-
-Our next set of results compares the performance of \geepers, \pmm, and \psw when one or more assumptions of each method may be violated.
-
-% \begin{figure}[!ht]
-%   \centering
-%   \includegraphics[width=5.5in,clip]{../simFigs/boxplotsNew.pdf}
-%   \caption{Violin and scatter plots of 500 simulation estimation errors for estimators described in Section~\ref{sec:simMods} under varying conditions.
-For all plots, $n=500$; there were 500 replications for each set of conditions.
-Annotations indicate the number of extreme outliers excluded from the plotting area.}
-%   \label{fig:boxplots}
-% \end{figure}
-
-%<<<<<<< HEAD
-Figure \ref{fig:boxplots} shows violin plots, layered on top of jittered scatter plots, showing estimation errors for \geepers, \pmm, and \psw.
-Table \ref{tab:coverage} shows empirical coverage rates of nominal 95\% confidence interval estimates for \geepers and \pmm; when at least one of an estimator's assumptions is violated, the result is colored red.
-
-When all of the \geepers and \pmm assumptions were met, that is, when $\epsilon\sim\mathcal{N}(0,1/2)$, $\alpha>0$, and there were no covariate interactions in \eqref{eq:y-sim}, results for $\geepers$ and $\pmm$ were the same as in the previous section; in addition, confidence intervals from both methods achieved their nominal levels.
-\textsc{psw} had by far the lowest variance and the highest bias of the three estimators; the latter is due to the violation of principal ignorability.
-
-%The remainder of the results reflect violations modeling assumptions for either \geepers or the parametric mixture model or both: when $\alpha=0$, violating Assumption \ref{ass:vps}; interactions in the data generating model between covariates $X$ and treatment assignment $Z$ and/or principal stratum $S$, either of which would violate assumption \ref{ass:rci} and the linear additive model of the \textsc{pmm}; and/or residuals drawn from a uniform distribution, which violates the normality assumption of \textsc{pmm}.
-
-%We omitted the $\alpha=0$ case from Figures \ref{fig:alphan} and \ref{fig:boxplots} since the variance of \geepers and \pmm estimates was so large as to dominate the plot. 
-When $\alpha=0$, Assumption \ref{ass:vps} is violated; %, standard errors for \geepers were too high to 
-Nevertheless, as the first column of results in Table \ref{tab:coverage} shows, when other assumptions were met, confidence intervals from both \geepers and \psw tended to over-cover when $\alpha=0$. 
-In a randomized experiment, $\alpha$ can be estimated directly, making Assumption \ref{ass:vps} testable.
-
-The impact of covariate interactions in \eqref{eq:y-sim} on the performance of \geepers depended on $\alpha$: when $\alpha=0.2$, the presence of interactions increased sampling variance.
-Still, only interactions between $X$ and $Z$ reduced confidence interval coverage.
-When $\alpha=0.5$, interactions had little to no effect on the \geepers estimator, and all 95\% confidence intervals approximately achieved their nominal levels.
-\geepers performed similarly under either normal or uniform residuals.
-
-Like \geepers, \pmm is insensitive to interactions between $X$ and $S$.
-However, interactions between $X$ and $Z$ in the data-generating model skewed its sampling distribution and reduced the coverage of its 95\% confidence intervals for all values of $\alpha$ in the simulation.
-Uniformly-distributed residuals caused bimodality in the \pmm sampling distribution, and caused its credible intervals to exhibit severe undercoverage.
-
-\label{simsum}In sum, \psw is the least variable of all three estimators, and does not depend on outcome modeling, but is vulnerable to bias from an unmeasured covariate that predicts both $S$ and $Y$.
-When all assumptions of both methods are met, \geepers is less efficient, but also less biased, than \pmm, but differences in efficiency decrease with $n$ and $\alpha$.
-Neither \psw nor \pmm is vulnerable to interactions between $X$ and $S$.
-\pmm is highly vulnerable to violations of normality and to interactions between $X$ and $Z$, while \psw is robust to non-normality and robust to $X$-$Z$ interactions when $\alpha = 0.5$.
-These results suggest that researchers who are willing to assume principal ignorability should use \psw, and those who are confident in their parametric model for potential outcomes should use a parametric mixture model.
-Researchers who are hesitant to accept principal ignorability or a parametric outcome model should consider \geepers, particularly when $n\ge 200$ and covariates can predict $S$ with AUC$>0.6$ (corresponding to $\alpha>0.2$).
-These suggestions may depend heavily on the simulation setup; applying them in general scenarios requires further research.
-
-%\label{evalMetric}The simulation study suggests that none of the three methods evaluated---\geepers, \pmm, or \psw---dominates the others, and offers some guidance on when each estimator may be preferable.
-\psw is the most precise and requires no assumptions about the distribution of outcomes, but can be quite sensitive to violations of principal ignorability.
-When the determinants of $\st$ are well-understood and measured, \psw seems to be the best option.
-When principal ignorability does not hold, potential outcomes are conditionally normal, and there are no $X:Z$ interactions within principal strata, the \pmm estimator outperforms \geepers, though the difference decreases with $n$ or $\alpha$.
-\geepers may be the best choice when conditional normality may not hold, or if there are $X:Z$ interactions and $\bm{x}$ is sufficiently predictive of $\st$ (AUC $\ge$0.675, corresponding to $\alpha \ge 0.5$, may serve as a threshold).
-
-%=======
-% Figure \ref{fig:boxplots} shows violin plots, layered on top of jittered scatter plots, showing estimation error for \textsc{geepers}, the mixture estimator, and the \textsc{psw} estimator of $\eff1$ under various sets of conditions, and Table \ref{tab:coverage} shows empirical coverage of nominal 95\% confidence interval estimates under a somewhat wider set of circumstances when $\muc1=0.3$.
-More complete results, including root-mean-squared-error estimates, are in the online appendix.
-% The leftmost panel of Figure \ref{fig:boxplots} and the top line of Table \ref{tab:coverage} show results under conditions favorable to both the mixture model and to \textsc{geepers}---normal residuals and no interactions in the data generating model between covariates and either $S$ or $Z$.
-% \textsc{geepers} was roughly unbiased, but less precise than the other two methods---when $\alpha=0.2$ it was much noisier, and when $\alpha=0.5$ it was only slightly less precise.
-% The mixture estimator's bias depended on $\muc1$---when $\muc1=0$, the bias was slight, and when $\muc1=0.3$, the bias was more substantial.
-This result is somewhat surprising, because when $\muc1=0$, there was no separation between the principal strata in the control group, and previous literature \citep{griffin2008application} has identified this scenario as a particular challenge for mixture modeling.
-% The \textsc{psw} estimator was also biased, due, presumably, to the violation of Principal Ignorability, Assumption \ref{ass:PI}, inherent in the data generating model.
-On the other hand, the \textsc{psw} estimator had easily the lowest sampling variance of the three estimators.
-
-% The middle panel of Figure \ref{fig:boxplots} and the fifth line of Table \ref{tab:coverage} show results for when the residuals in \eqref{eq:y-sim} are uniformly distributed.
-% In this case, \textsc{geepers} and \textsc{psw} estimators, which do not assume normality, behaved in a similar fashion as in the normal case, but the mixture estimator showed a bimodal pattern--sometimes over-estimating $\eff1$ and sometimes underestimating, but rarely estimating $\eff1$ accurately.
-% Credible intervals from the mixture model exhibited severe undercoverage, while \textsc{geepers} confidence intervals achieved at least their nominal level.
-
-% %\sloppy
-% The rightmost panel of Figure \ref{fig:boxplots} returns to the case of normal residuals---and only shows results when $\muc1=0.3$ and $\alpha=0.5$---but introduces interactions in the data generating model between covariates and either $\st$, $Z$, or both, violating Assumption \ref{ass:rci}.
-% %Presence of interactions could undermine the \textsc{geepers} by inducing a relationship between principal scores and potential outcomes, even after adjusting the outcomes for covariate effects.
-% Fortunately, these interactions made little difference when $\alpha=0.5$.
-However, Table \ref{tab:coverage} shows that when $\alpha=0$ or 0.2, and interactions between $\bx$ and $Z$ were present (i.e., $\gamma_3\ne 0$), \textsc{geepers} 95\% confidence intervals tended to under-cover.
-Nevertheless, coverage of \textsc{geepers} intervals was nearly always substantially higher than coverage of corresponding 95\% credible intervals from mixture models.%\footnote{}
-% >>>>>>> overleaf-2025-06-12-1948
+\clearpage
 
@@
 
-\section{Applications}
-To illustrate \textsc{geepers} and compare it to the other discussed methods, we include two applied examples.
-First, we briefly replicate an analysis from \citet{richardson2023estimating} using the \textsc{geepers} and the other methods discussed above.
-This is a relatively straightforward example of an experiment with imperfect compliance in which \citet{richardson2023estimating} estimated statistically significant effects overall in both principal strata.
-Next, we present a novel analysis of data from the RCT of computer-based math applications reported in \citet{impactPaper}, and look into the role of bottom-out hints, as described in the introduction.
-This analysis aims to shed light on an important question in education research, illustrate model and covariate selection, and include a discussion on the plausibility of identification assumptions.
-
-\subsection{Accounting for Imperfect Compliance: Re-Analysis of the Obstetrics and Periodontal Therapy Study}\label{sec:opts}
-\sloppy
-Periodontal disease, a weakening of bone and connective tissue supporting teeth, is highly prevalent among American adults and is associated with adverse pregnancy outcomes, including preterm birth \citep{periodontalEpi}.
-The Obstetrics and Periodontal Therapy Study \citep[OPTS; ][]{michalowicz2006treatment} investigated the efficacy of nonsurgical periodontal treatment to reduce the preterm birth rate; replication data are available at \url{https://causeweb.org/tshs/obstetrics-and-periodontal-therapy/}.
-Pregnant patients at four different hospitals who were at least 16 years old, at less than 16 weeks and 6 days of gestation, and diagnosed with periodontal disease were individually randomized to either receive periodontal treatment during the first 21 weeks of pregnancy ($Z=1$, $n_T=413$) or after delivery ($Z=0$, $n_C=410$).
-The main component of the intervention consisted of up to four visits, followed by monthly tooth polishing and other dental care as needed.
-Not all participants randomized to the treatment condition completed the entire treatment plan; however, 96\% of them received at least some periodontal care during pregnancy.
-Patients randomized to the control condition did not have access to the intervention.
-
-Following \citet{richardson2023estimating}, let $\st=1$ for patients who, if randomized to treatment, would complete the full treatment plan, and let $\st=0$ for those who would not.
-Let $Y$, the outcome of interest, be the percentage of probed sites (6 sites per tooth) that were bleeding at visit 5, which was at 29--32 weeks of gestation, with potential values $\yc$ and $\yt$, as above.
-\citet{richardson2023estimating} used a principal stratification estimator they called ``bespoke instrumental variables'' (BSIV) to estimate principal effects $\eff0$ and $\eff1$, the average effects of randomization to the treatment condition for patients who would or would not complete the full treatment plan.
-In this section, we compare their result to estimates using \geepers, \pmm, and \psw.
-
-For simplicity, we followed the same analysis decisions as \citet{richardson2023estimating}. 
-Specifically, we conditioned on two covariates, Serum measures of endotoxin level and fibrinogen at baseline, and
-omitted the 164 cases in which $Y$ was missing as well as 19 additional cases in which one or both of the covariates were missing.
-That left a total sample size of 640, with 314 randomized to the treatment condition and 326 randomized to control.
-%The compliance variable in the dataset had three categories: ``Yes,'' for subjects in the treatment group who completed the treatment plan (157 subjects), ``No'' for those who did not (4 subjects), and ``Und'' for those who received some therapy, but it is unknown whether they completed the treatment (157 subjects).
-%We let $S=1$ when compliance was ``yes''---i.e., when subjects in the treatment group completed the treatment---and $S=0$ otherwise.
-
-For all three estimation methods, we estimated principal scores with logistic regression of $\st$ on the two covariates.
-For \textsc{geepers} and \textsc{psw}, we used standard maximum likelihood methods and data from the treatment group, achieving an AUC of 0.68.
-For the mixture model, we incorporated the logistic regression specification as a sub-model within a larger Bayesian model fit in Stan.
+\clearpage
+
+\begin{figure}[!ht]
+  \centering
+  \includegraphics[width=4.5in]{../simFigs/biasSEbyB1N.jpg}
+  \caption{Empirical bias and standard error of \textsc{geepers} and Mixture estimates of $\eff1$ as sample size per treatment group $n$ or $\alpha$ varied.
+There were 500 replications for each $n$ or $\alpha$.
+Other factors were held fixed at the noted values.}
+  \label{fig:alphan}
+\end{figure}
+
+\clearpage
+
+\begin{figure}[!ht]
+  \centering
+  \includegraphics[width=5.5in,clip]{../simFigs/boxplotsNew.pdf}
+  \caption{Violin and scatter plots of 500 simulation estimation errors for estimators described in Section~\ref{sec:simMods} under varying conditions.
+For all plots, $n=500$; there were 500 replications for each set of conditions.
+Annotations indicate the number of extreme outliers excluded from the plotting area.}
+  \label{fig:boxplots}
+\end{figure}
+\clearpage
 
@@
 
-The results are displayed in Figure \ref{fig:opt}% and Table \ref{tab:optEffs}
-: both \textsc{geepers} and the Bayesian mixture model gave similar results to those reported in \citet{richardson2023estimating} from BSIV.
-For the principal stratum of subjects who would not fully complete the treatment, those three methods estimated that assignment to the treatment condition reduced percentage of dental sites with bleeding by 31--34 percentage points; for subjects who would complete the treatment, the three estimated a reduction of 14-16 percentage points.
-The conclusion of these methods is somewhat surprising: that the effect was larger for subjects who would not complete the treatment than for those who would (the two strata had similar bleeding percentages at baseline, and a logit transformation of the outcome leaves the results qualitatively unchanged).
-One possible explanation is the conjecture that subjects stopped attending treatment once their periodontal disease improved.
-
-In contrast, principal score weighting resulted in very similar effects for both strata---a reduction of 23--25 percentage points, which is close to the estimated intent-to-treat effect of 25 percentage points.
-
-\subsection{Estimating Treatment Effects for Implementation Subgroups: Students who Frequently or Rarely Request Bottom-Out Hints}\label{sec:fh2t}
-
-In the educational field experiment reported in \citet{impactPaper}, middle school students completing their math schoolwork on a computer program were randomly assigned to four conditions.
-These included two gamified programs---DragonBox and From Here to There (FH2T)---and two programs in which students used computers to work through a series of algebra problems taken from the district's textbook---``ASSISTments'' and an active control called ``Business as Usual'' (BAU).
-% the active control condition in which students simply used the computer to work on a sequence of math problems.
-In the DragonBox game, students learned to solve algebraic equations by isolating a box containing a dragon; as students progressed in the game, pictures of monsters were replaced by mathematical symbols.
-Another condition, the From Here to There (FH2T) program, used visual cues such as color and spacing to teach algebraic concepts and allowed students to solve equations by manipulating expressions via dragging and dropping.
-%In ASSISTments, students used computers to work through a series of math problems taken from the district's textbook.
-Students in the ASSISTments condition had access to a series of hints for each problem, culminating in a ``bottom out'' hint that contained the answer.
-They received immediate feedback on their answers and had to enter the correct answer to proceed to the next problem.
-In the BAU condition, students had no access to hints and received feedback only after completing the assignment.
-
-In all four conditions, students worked on their assigned math program in class on 12 different occasions spaced throughout the school year: a prior assessment, a mid-test, a post-test, and nine learning sessions.
-In this illustration, we will compare ASSISTments---the only condition featuring bottom-out hints---with each of the three others separately.
-Our goal is to estimate different treatment effects for groups of students who, if assigned to ASSISTments, would request few or many bottom-out hints.
-
-The option to bottom out was only one of several differences between the treatment conditions.
-One way to disentangle the role of bottoming out in any treatment effect would be to estimate the average treatment effects of being assigned to ASSISTments, relative to the other three, separately for subjects who would (if assigned to ASSISTments) bottom out frequently, and for subjects who would not.
-If the availability of bottom-out hints plays a positive role in ASSISTments' efficacy, we might expect larger effects of the condition on students who frequently bottom out than on those who do not; if bottom-out hints are harmful, we might expect the opposite.
-If hints (that do not supply the correct answer) and error messages are helpful, but the ability to bottom out is harmful to learning, then we might expect to estimate a positive effect of being assigned to the feedback condition for students who do not bottom out, but a negative effect for students who do.
-If hints are helpful for learning only if they can be read as worked examples, then we might expect no effect of condition on students who generally do not bottom out, but a positive effect for students who do.
-(These hypothetical results may be masked by other treatment effect heterogeneity between students who would or wouldn't bottom out, but the results could still be informative.)
-
-\subsubsection{Principal Stratification and Alternatives}\label{sec:psInEdTech}
-The goal of principal stratification in this context is to determine if assignment to the ASSISTments condition affected students who tended to request frequent bottom-out hints differently than students who tended to request fewer bottom-out hints.
-Is this the right question to ask?
-
-In a similar context, \citet{sales2021student} suggests two alternative approaches: the average effect of requesting bottom-out hints on posttest scores for students in the ASSISTments condition, and estimating natural direct and indirect effects of assignment to the ASSISTments condition, treating bottom-out hint requests as a mediator.
-The former approach is an observational study: although students were randomized to the ASSISTments condition, bottom-out hint requests were not randomized.
-Therefore, estimating the effect of bottom-out hints requires careful control of confounding.
-Unfortunately, confounding is a major concern in this case, since, presumably, students only requested bottom-out hints if they were struggling with the material, and students who struggled with practice problems are likely to score lower on the posttest, on average, than those who needed less help during practice.
-The best available measure of students' baseline ability, their score on the 10-item pretest, is likely insufficient.
-That said, a plausible estimate of the average effect of bottom-out hint requests would be a valuable contribution to the literature, and a suitably innovative observational design \citep[e.g.][]{beck} may bear fruit.
-
-Another arguable disadvantage of the observational study approach is that it overlooks the broader context of the experiment, focusing solely on whether bottom-out hints affect posttest scores rather than examining their role in the overall effectiveness of ASSISTments.
-A second approach, mediation analysis, contextualizes the bottom-out effect within the larger experiment.
-It, too, relies on the same non-confounding assumption as an observational study, and also suffers from an additional challenge: mediation analysis relies on a ``common support'' condition that ``the mediator state must not be a deterministic function of the treatment'' \citep[p.
-225]{celli2022causal}.
-However, bottom-out hints are unavailable in the other three arms of the study, violating this condition (\citealt{sales2021student} offers a solution, which we find unconvincing).
-
-A principal stratification approach treats bottom-out hint requests as a type of compliance.
-In other words, if we consider full implementation of the ASSISTments condition to include frequent bottom-out hint requests, students assigned to ASSISTments but who requested few bottom-out hints only partially complied with their treatment assignment.
-The question then is whether students who do not fully comply with their treatment assignment still experience an effect.
-Conversely, might students who only partially comply experience a larger positive effect?
-That would be the case if either bottom-out hints were harmful or if the treatment was better targeted toward the type of student who tends to need them less.
-
-Principal score weighting depends on the assumption of principal ignorability, which is equivalent to the no-confounding assumption in observational studies and mediation analysis.
-The mixture model and \textsc{geepers} estimators do not rely on this assumption; instead, they rely on parametric models and residualized covariate independence, respectively.
-A final option is the instrumental variables estimator of \citet{air}, which relies on the ``exclusion restriction'' that randomization did not affect students who would request few bottom-out hints.
-Since bottom-out hints are only one of several features of ASSISTments, this assumption is untenable.
-
-\subsubsection{Data on Bottom-Out Hints and Covariates}
-
-\sloppy
-We gathered dichotomized data on %which students in the treatment condition requested at least one bottom-out hint.
-whether students in ASSISTments requested more than the median number of bottom-out hints (i.e., 11).
-That is, $\st=1$ for students who, if assigned to ASSISTments, would request more than 11 bottom-out hints (``bottom-outers''), and $\st=0$ for students who would request 11 or fewer (``non-bottom-outers'').
-%least one bottom-out hint, and $\st=0$ for students who would not.
-The outcome of interest $Y$ in our example is students' total scores on a ten-item posttest completed within the online tutoring system; $Y$ is an integer ranging from 0 to 10.
-The goal of our analysis will be to determine the principal effects $\eff1=\EE[\yt-\yc|\st=0]$ and $\eff0=\EE[\yt-\yc|\st=0]$, the average effects of assignment to ASSISTments (``$T$'' subscripts) versus the other conditions (``$C$'') for bottom-outers and non-bottom-outers.
-
-We also had access to data on several baseline covariates, including prior achievement, demographics, baseline measurements of student attitudes toward math, and an indicator for whether students began the school year in remote or in-person instruction.
-Students with missing pretest scores were dropped from the analysis; missing data in other covariates were imputed with a Random Forest algorithm \citep{missForest}.
-Summary statistics are displayed in appendix tables \ref{table:tab1fac} and \ref{table:tab1num}.
-
-%The data analysis we present here is intended as a demonstration of \textsc{geepers}, rather than for its substantive conclusions, and our description omits discussion of some important methodological considerations, including attrition and post-selection inference.
-For details on the experimental design, the conditions being compared, sample attrition---which was unusually high due to the COVID-19 pandemic---and the impact analysis, please see \citet{impactPaper}.
-
-For a description of the dataset and an explanation of how to access it, see \citet{ottmar2023data} and \url{https://osf.io/r3nf2/}.
-Replication code for the analysis is available at [Redacted].
-%\url{https://github.com/adamSales/psGee}.
-
-\subsubsection{Data Analysis}
-We estimated principal scores using logistic regression of observed $S$ on school fixed effects and a set of baseline variables $\bxs$.
-Three criteria governed our choice of covariates and model specifications for estimating principal scores and effects.
-First, Proposition \ref{prop:reg2} requires consistent estimation of principal scores and (via Assumption \ref{ass:rci}) correct specification of the model for $\EE[Y|\bx,Z,\st]$.
-However, unlike \psw, \geepers does not require that models include all important covariates (for any definition of "important"), so covariates whose inclusion may threaten model specifications or identification assumptions can be dropped from the analysis.
-
-On the other hand, Assumption \ref{ass:vps} requires principal scores to vary with at least some element $\bxs$.
-Moreover, the results of the simulation study suggest that when $\st$ can be predicted more precisely as a function of $\bxs$, \textsc{geepers} principal effect estimators are more precise and more robust to model misspecification.
-Those considerations militate in favor of choosing covariates $\bxs$ to optimize out-of-sample principal score prediction accuracy.
-
-The first two criteria motivated our choice of two strategies for choosing $\bxs$ and model specifications. 
-In our main model, we included every available covariate in the dataset and chose transformations of those covariates based on the results of residual and binned residual plots \citep{arm}.
-%
-% Including all possible covariates maximizes the model's opportunities to detect covariate relationships that would predict $Y$ or $\st$.
-However, this strategy can also lead to overfitting, which would decrease out-of-sample predictive accuracy.
-In an alternative analysis, we chose a subset of covariates to include in $\bxs$ %and $\bxy$ 
-with a backward stepwise selection algorithm minimizing the AIC \citep{aic} of the principal score model and then modified the resulting models based on residual plots.
-
-Lastly, Assumption \ref{ass:ci} implies that there is no interaction between $\bxs$ and $\st$ in the outcome model. 
-\emph{A priori}, the student attribute that poses the greatest threat to these modeling choices is prior achievement, that is, scores on the pretest and the fifth-grade state test.
-This is both because they measure a similar construct to the posttest and because \cite{impactPaper} found a pretest-treatment interaction in a different treatment contrast.
-Following that reasoning, we conducted a third parallel analysis excluding prior achievement covariates. % in $\bxs$ but including all covariates in $\bxy$.
-
-All three principal score models were fit to data from students in the ASSISTments group, for whom bottom-out status ($S$) was observed; hence, the same model fit was used for all three treatment contrasts.
-
-Outcome models for all analyses included all variables from the corresponding principal score models, but substituted class fixed effects for school fixed effects.\footnote{Why use a coarser set of fixed effects for the principal score model than the outcome regression?
-First, the principal score model was fit to a smaller sample (only students in the ``Instant'' condition), so some classrooms only included one student, which can lead to poor fit.
-Second, the asymptotics of logistic regression are more restricted than OLS, so logistic regression tends to perform poorly when the number of parameters increases with the sample size \citep{agresti}.}
-As with principal score models, we checked model fit with residual plots and modified model specifications accordingly.
-
-% \begin{figure}
-%   \centering
-%   \includegraphics{../figure/psModCoef.jpg}
-%   \caption{Principal Score Model Coefficients}
-%   \label{fig:psMod}
-% \end{figure}
-
-\subsubsection{Results}
-
-The three principal score models were fairly successful in distinguishing bottom-outers from non-bottom-outers---the AUC, evaluated using out-of-sample predictions in a 10-fold cross-validation, was roughly 0.795 for the AIC-optimal model, and about 0.745 for the two others.
-That is, in roughly 75--80\% of bottom-outer/non-bottom-outer pairs of subjects, the bottom-outer will have a higher predicted probability from the model.
-Coefficient estimates are displayed in Table \ref{tab:psTab} in the appendix.
-
-Figure \ref{fig:effects} shows principal effect point estimates and 95\% confidence intervals from analyses based on the three principal score models.
-The three covariate selection strategies led to similar point and interval estimates.
-That said, in nearly all cases, the AIC-optimal model led to the smallest standard errors.
-Confidence intervals for all nine principal effect estimates were about twice as wide as for corresponding ATE estimates.
-In no case was there evidence at $\alpha=0.05$---or any other conventional level---for non-zero principal effects or differences between principal effects.
-Table \ref{tab:regTab}, in the appendix, displays the full set of regression coefficients for the outcome models based on the "All Covariates" principal score model, along with nominal standard error estimates.
-
-A comparison of point estimates suggests that assignment to ASSISTments rather than BAU benefits bottom-outers less than non-bottom-outers.
-We speculate that, if true, this pattern may be because many students use bottom-out hints to game the system---entering the provided correct answers rather than attempting to figure out the problems on their own.
-Removing access to hints for bottom-outers forces them to attempt the problems rather than game the system.
-At the same time, hints and instant feedback are helpful for non-bottom-outers, who are less likely to game the system.
-
-On the other hand, point estimates suggest that assignment to ASSISTments, relative to FH2T, hurts non-bottom-outers but not non-bottom-outers.
-Perhaps non-bottom-outers tend to have higher prior knowledge, and, as \cite{impactPaper} shows, FH2T benefits such students more than students with lower prior knowledge.
-
-Perhaps the simplest explanation for observed differences in point estimates is sampling error, which cannot be ruled out.
+\clearpage
 
@@
 
-\subsubsection{Comparing Principal Stratification Methods}
+\clearpage
 
@@
 
-Figure \ref{fig:compare} shows principal effect estimates from \textsc{geepers} (using the All Covariates principal score model) compared to analogous estimates using Bayesian mixture modeling and \textsc{psw}, alongside estimates of the overall average treatment effects, labeled ``ATE.''
-
-The \textsc{psw} model used the same principal score model as \textsc{geepers}, while the Bayesian model used the same principal score and outcome model specifications.
-
-The Bayesian mixture model assumed that outcomes were normally distributed, conditional on covariates and principal strata, with a residual standard deviation that varied between principal strata.
-This modeling assumption is necessarily false since post-test scores were integers between 0 and 10.
-All model parameters with support in $\mathbb{R}$ were given standard normal priors, while standard deviation parameters were given half-standard normal priors.
-The principal score and outcome models were fit simultaneously using a Markov Chain Monte Carlo algorithm \citep{rstan}.
-
-The \textsc{psw} estimators did not include any covariate outcome modeling. Standard errors for the \textsc{psw} estimates were estimated using the bootstrap, with resampling done within schools.
-
-For all three sets of estimators, approximate 95\% confidence intervals were estimated by adding and subtracting twice the standard error from the point estimate (or twice the posterior standard deviation from the posterior mean, in the Bayesian mixture model).
-
-The other estimation strategies largely agreed with \textsc{geepers} and with each other in the contrast between ASSISTments and BAU, and, to a lesser extent, in the contrast between ASSISTments and DragonBox.
-The biggest difference between the methods was in the contrast with FH2T, in which the Bayesian mixture model resulted in a larger difference between the two principal effects, as well as the largest standard errors, while the two \textsc{psw} principal effect estimates were quite similar.
-However, these differences are difficult to interpret due to the large uncertainty surrounding all the point estimates.
-
-\section{Discussion}\label{sec:discussion}
-\textsc{geepers} is a straightforward approach to principal effect estimation under strong monotonicity, built on widely-used regression models, which is more robust---though sometimes less precise---than alternative approaches under a wide array of scenarios.
-
-There is good reason to hope that extensions to \textsc{geepers}, including cases in which $S$ takes more than two values, may be straightforward.
-For instance, in truncation by death problems \citep[e.g.][]{zhangRubin,ding2011} with weak monotonicity, $S_i=1$ if participant $i$ survives (or, more generally, if the outcome $Y_i$ is measured) and 0 otherwise, interest is typically in the principal effect for the ``always survive'' principal stratum in which $\sti=S_{ci}=1$.
-If, say, $\sti\ge S_{Ci}$, then every subject in the control condition with $S_i=1$ is in the always survive stratum, while those subjects in the treatment condition with $S_i=1$ are a mixture of the always survive and $\sti=1;\;S_{Ci}=0$ stratum.
-This scenario is broadly similar to the strong monotonicity, one-way noncompliance situation discussed in this paper.
-On the other hand, when no monotonicity assumption holds, both $\st$ and $S_C$ will have to be imputed for every subject; further research is necessary to determine the appropriate way to do so.
-
-Another direction of extension involved the principal score model \eqref{eq:pscore}.
-The performance of \textsc{geepers} in the simulation study of Section \ref{sec:simulation} depended heavily on the factor $\alpha$, which controlled the extent to which covariates could predict $S$.
-That suggests that when covariates are high-dimensional, \textsc{geepers}' performance in applications could be optimized with a high-dimensional semi- or non-parametric model.
-If so, several further questions emerge.
-First, can the parameter $\alpha$ be extended to a more general parameter measuring the prediction accuracy of a non-parametric model?
-Second, if the principal score model cannot be formulated as the solution to a set of estimating equations, how should the standard error be computed?
-Lastly, can over-fit principal score models cause bias or other estimation problems, and if so, are there ways to protect against overfitting?
-
-Sampling error prevents firm conclusions regarding principal effects for students who would, if given the opportunity, request frequent bottom-out hints.
-However, the results are consistent with the hypothesis that ASSISTments is less effective for frequent bottom-out requesters when compared to a similar program without hints or immediate feedback.
-We hypothesize that many students who frequently request bottom-out hints are gaming the system, facilitated by the availability of these hints.
-
-\textsc{geepers} is a flexible, easily-implementable method for principal effect estimation for one-way noncompliance, with predictive covariates; extensions to a broader set of circumstances could be a boon to causal modeling.
-
-\section*{Replication Materials}
-Replication materials can be found at
-\url{https://osf.io/fu2sv/?view_only=b2baebc2f3054ef796717863e4a9662f}.
-
-\bibliographystyle{apalike}
-\bibliography{MOM}
-
 \end{document}

### 2025-07-04 writeUps/appendix.tex
(re)submitted online appendix

@@
 \usepackage[letterpaper]{geometry}
+\usepackage{placeins}
+
 \usepackage{xr}
-\externaldocument{writeUps/paper}
+\externaldocument{paper}
 \RequirePackage{amsmath,amsfonts,amssymb,amsthm}
@@
 
-\title{Online Appendices for ``GEEPERs: Principal Stratification using Principal Scores and Stacked Estimating Equations''}
+\title{Online Appendices for ``GEEPERs: Principal Stratification using
+  Principal Scores and Stacked Estimating Equations''}
+\date{}
 %\runtitle{\textsc{geepers}}
-%\maketitle
+\maketitle
 
@@
 
-\section*{Appendix: Proofs and Calculations}
-\subsection*{Proof for Lemma \ref{lemma:expectation}}
+\section{Appendix: Proofs and Calculations}
+\subsection{Proof for Lemma \ref{lemma:expectation}}
 As a preliminary, note that
@@
 
-\subsection*{Proof for Proposition \ref{prop:reg1}}
+\subsection{Proof for Proposition \ref{prop:reg1}}
 
@@
 
-\subsection*{A Stronger Version of Proposition \ref{prop:reg2} and a Proof}
+\subsection{A Stronger Version of Proposition \ref{prop:reg2} and a Proof}
 
@@
 
-\subsection*{Sandwich Matrix Calculations}
+\subsection{Sandwich Matrix Calculations}
 
@@
 
-\subsection*{$A_{21}$ Matrix}
+\subsection{$A_{21}$ Matrix}
 The estimating equations for the regression \eqref{eq:regression} are
@@
 
-\clearpage
+%\clearpage
 
@@
 
-\setcounter{page}{1}
+%\setcounter{page}{1}
 
-\begin{center}
-\large
-    \textbf{Online Appendices for ``GEEPERs: Principal Stratification using Principal Scores and Stacked Estimating Equations''}
-\end{center}
+% \begin{center}
+% \large
+%     \textbf{Online Appendices for ``GEEPERs: Principal Stratification using Principal Scores and Stacked Estimating Equations''}
+%   \end{center}
+  \FloatBarrier
 \section{Additional Simulation Results}
+\FloatBarrier
 \subsection{Plot of  AUC versus $\alpha$}
-
+\FloatBarrier
 \begin{center}
@@
 %\caption{Coverage of nominal 95\% confidence intervals}
-\clearpage
+%\clearpage
 
+\FloatBarrier
 \subsection{Full Empirical 95\% Interval Coverage Results}
+\FloatBarrier
+
 The following tables give the empirical coverage of nominal 95\% intervals for \textsc{geepers} and mixture model principal effect estimates under varying data generating models.
@@
   \caption{Empirical coverage of nominal 95\% Confidence intervals for \geepers and \pmm when $n=500$ per condition.}
-  \input{writeUps/coverageTabAppendix500.tex}
+  \input{coverageTabAppendix500.tex}
  \end{table}
@@
   \caption{Empirical coverage of nominal 95\% Confidence intervals for \geepers and \pmm when $n=1000$ per condition.}
-  \input{writeUps/coverageTabAppendix1000.tex}
+  \input{coverageTabAppendix1000.tex}
  \end{table}
 
-\clearpage
+%\clearpage
+\FloatBarrier
 \subsection{Full RMSE Results}
+\FloatBarrier
+
 The following table gives the root mean squared error (RMSE), $\left\{\sum_b (\hat{\tau}-\tau)^2/500\right\}^{1/2}$, for \textsc{geepers}, mixture model, and principal score weighting principal effect estimates under varying data generating models.\\
@@
   \caption{Empirical RMSE  for \geepers, \pmm, and \psw when $n=500$ per condition.}
-  \input{writeUps/rmseTabAppendix500}
+  \input{rmseTabAppendix500}
  \end{table}
@@
   \caption{Empirical RMSE  for \geepers, \pmm, and \psw when $n=1000$ per condition.}
-  \input{writeUps/rmseTabAppendix1000}
+  \input{rmseTabAppendix1000}
  \end{table}
@@
 %\end{table}
-\clearpage
+%\clearpage
 
+\FloatBarrier
 \section{Additional Results from the OPT Study}
+\FloatBarrier
 \singlespacing
+\FloatBarrier
 \subsection{Summary Statistics}
+\FloatBarrier
 The following table gives summary statistics for covariates and post-treatment outcomes in two of the conditions from the empirical study.\\
@@
 
-\clearpage
+%\clearpage
+\FloatBarrier
 \section{Additional Results from the Bottoming-Out Study}
@@
 
+\FloatBarrier
 \subsection{Summary Statistics}
+\FloatBarrier
+
 The following table gives summary statistics for covariates and post-treatment outcomes in two of the conditions from the empirical study.\\
@@
 \input{tab1fac}
-\clearpage
+%\clearpage
 
@@
 \input{tab1num}
+\FloatBarrier
+\subsection{Regression Results}
+\FloatBarrier
 
-\subsection{Regression Results}
 Regression estimates from three principal score logit models and three outcome regressions, based on the "All Covariates" principal score model, for \textsc{geepers} estimates.
@@
 \small
-%\input{writeUps/outcomeRegAppendix}
+%\input{outcomeRegAppendix}
 \input{psTab}
@@
 
+\bibliographystyle{apalike}
+\bibliography{MOM}
+
 \end{document}

### 2025-07-04 writeUps/appendix.tex

@@
 \usepackage{xr}
-\externaldocument{paper}
+\externaldocument{writeUps/paper}
+
 \RequirePackage{amsmath,amsfonts,amssymb,amsthm}
@@
 
-\input{notation.tex}
+\input{writeUps/notation.tex}
 
@@
 \subsection{Proof for Lemma \ref{lemma:expectation}}
+
 As a preliminary, note that
@@
 
-%\clearpage
-
-%\processdelayedfloats
-
-%\setcounter{page}{1}
-
-% \begin{center}
-% \large
-%     \textbf{Online Appendices for ``GEEPERs: Principal Stratification using Principal Scores and Stacked Estimating Equations''}
-%   \end{center}
   \FloatBarrier
@@
 \FloatBarrier
+
 \begin{center}
 
-    \includegraphics[width=5.5in]{../simFigs/alphaAUC.pdf}
+    \includegraphics[width=5.5in]{simFigs/alphaAUC.pdf}
 
 \end{center}
-%\begin{table}
-%\caption{Coverage of nominal 95\% confidence intervals}
-%\clearpage
 
@@
 
+=======
 The following table gives the root mean squared error (RMSE), $\left\{\sum_b (\hat{\tau}-\tau)^2/500\right\}^{1/2}$, for \textsc{geepers}, mixture model, and principal score weighting principal effect estimates under varying data generating models.\\
@@
 \FloatBarrier
+
 The following table gives summary statistics for covariates and post-treatment outcomes in two of the conditions from the empirical study.\\
@@
 %\input{outcomeRegAppendix}
+
 \input{psTab}

### 2025-07-04 writeUps/paper.tex
Updates from Overleaf

--- 2025-07-02 writeUps/paper.tex
@@
 
-\begin{figure}[!ht]
-  \centering
-  \includegraphics[width=4.5in]{../simFigs/biasSEbyB1N.jpg}
-  \caption{Empirical bias and standard error of \textsc{geepers} and Mixture estimates of $\eff1$ as sample size per treatment group $n$ or $\alpha$ varied.
+% \begin{figure}[!ht]
+%   \centering
+%   \includegraphics[width=4.5in]{../simFigs/biasSEbyB1N.jpg}
+%   \caption{Empirical bias and standard error of \textsc{geepers} and Mixture estimates of $\eff1$ as sample size per treatment group $n$ or $\alpha$ varied.
 There were 500 replications for each $n$ or $\alpha$.
 Other factors were held fixed at the noted values.}
-  \label{fig:alphan}
-\end{figure}
-
-\begin{figure}
-  \centering
-  \includegraphics{../simFigs/biasSEbyB1.jpg}
-  \caption{Empirical bias and standard error of M- and Mixture estimates of $\muc1$ as $\alpha$ varies from 0.1 to 1.
+%   \label{fig:alphan}
+% \end{figure}
+
+% \begin{figure}
+%   \centering
+%   \includegraphics{../simFigs/biasSEbyB1.jpg}
+%   \caption{Empirical bias and standard error of M- and Mixture estimates of $\muc1$ as $\alpha$ varies from 0.1 to 1.
 There were 500 replications for each $n$.
 Other factors were held fixed at the noted values.}
-  \label{fig:alpha}
-\end{figure}
+%   \label{fig:alpha}
+% \end{figure}
 
@@
 \section*{Replication Materials}
+\sloppy
 Replication materials can be found at
